### 2020-03-20 source/main.tex
first commit

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2019

% ready for submission
\usepackage{neurips_2019}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2019}

% to compile a camera-ready version, add the [final] option, e.g.:
     % \usepackage[final]{neurips_2019}

% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2019}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amsthm}     % for theorems
\usepackage{algorithm,algorithmic}
\usepackage{mathtools}
\usepackage{amsmath}
\allowdisplaybreaks
\usepackage{amssymb}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\usepackage{balance}

\newcommand{\eq}[1]{{Eq~(#1)}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}{Remark}
\newcommand{\lkxcom}[1]{{\color{red}{#1}}}
\input{bmacros}
\newtheorem*{assumption*}{Assumption}

\title{Unified Convergence Analysis of FedAve on Non-IID Data}

\author{}

\begin{document}

\maketitle

% \tableofcontents
% \begin{abstract}
% \end{abstract}

%% Note for ICLR'20 paper
% \input{note_sec_problem}
% \input{note_sec_algorithm}
% \input{note_sec_part}
% \input{note_sec_q}

% converge proof
\input{sec_problem}
\input{sec_overview}
\newpage
% \input{sec_acceleratedFL}
% \input{sec_cvxnonsm}

% accelerated FL survey

% \newpage
\small
\bibliographystyle{plain}
\bibliography{ref.bib}

\end{document}

### 2020-03-20 source/sec_acceleratedFL.tex
first commit

% !TEX ROOT=./main.tex

\section{Accelerated Federated Learning}

\subsection{Literature review}

\cite{felbab2019optimization} applies various accleration method
into federated learning and empirically compare the performance, 
while they did not provide any theoretical analysis of convergence.
\cite{liu2019accelerating} discuss momentum gradient descent
in the context of federated learning. They provide
convergence rate $\cO(\frac{1}{T})$ 
for strongly convex, $L$-smooth under gradient
descent algorithm. 
\cite{huo2020faster} provide convergence rate of momentum accelerated 
stochastic gradient descent for non-convex functions.

\subsection{Framework summarization}

\begin{assumption}
The expected squared norm of stochastic subgradients is uniformly bounded. i.e.,
$\mathbb{E}\|\vg_{t,k}\|^2  \leq G_k^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.  This also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$.
\label{ass:subgrad2}
\end{assumption}

\begin{assumption}[Lipschitz continuous]
	The local function $F_k$ is lipschitz continuous. 
	\label{ass:lipcont}
\end{assumption}

\begin{assumption}[L-smooth]
$F_{1}, \cdots, F_{N}$ are all $L$-smooth: for all  $\mathbf{v}$  and $\mathbf{w}$, $F_{k}(\mathbf{v}) \leq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{L}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$.
\label{ass:lsmooth}
\end{assumption}

\begin{assumption}[Polyak-Lojasiewicz (PL)] The global objective functions $f($ ) is differentiable and satisfy the Polyak-Lojasiewicz condition with constant $\mu,$ i.e., $\frac{1}{2}\|\nabla f(\boldsymbol{w})\|_{2}^{2} \geq \mu\left(f(\boldsymbol{w})-f\left(\boldsymbol{w}^{*}\right)\right)$ holds $\forall \boldsymbol{w} \in \mathbb{R}^{d}$ with
$\boldsymbol{w}^{*}$ being the optimal solution of global objective.
\label{ass:pl}
\end{assumption}

% \begin{assumption}[Strongly-convex]
% $	F_{1}, \cdots, F_{N} \text { are all } \mu \text { -strongly convex: for all v and } \mathbf{w}, F_{k}(\mathbf{v}) \geq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{\mu}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$
% \label{ass:stroncvx}
% \end{assumption}

\begin{assumption}[Bounded Variance]
Let $\xi_{t}^{k}$ be sampled from the k-th device's local data uniformly at random. The variance of stochastic gradients in each device is bounded:
$\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq \sigma_{k}^{2}$, for $k = 1,..., N$.
\label{ass:lgradvar}
\end{assumption}

\begin{assumption}[Expected square norm of gradient]
The expected squared norm of stochastic gradients is uniformly bounded. i.e.,
$\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)\right\|^{2} \leq G^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
\label{ass:squaregrad}
\end{assumption}

\begin{assumption}[loss function f]
	Loss function $F$ has lower bound $F_{\inf}$. This assumption is assumed for all the works.  The existence of $F^*$. We didn't add this 
	in the table.
	\label{ass:flb}
\end{assumption}

\begin{assumption}[Bounded Weighted Gradient Diversity]
	The weighted gradient diversity among local objectives is defined as:
$$
\Lambda(\boldsymbol{w}, \boldsymbol{q}) \triangleq \frac{\sum_{k=1}^{N} p_{k}\left\|\nabla f_{k}(\boldsymbol{w})\right\|_{2}^{2}}{\left\|\sum_{k=1}^{N} p_{k} \nabla f_{k}(\boldsymbol{w})\right\|_{2}^{2}} \leq \lambda
$$
\label{ass:bgd}
\end{assumption}

\begin{table}[h!]
\centering
\small
	\begin{tabular}{|c|c|c|c|c|}\hline
		paper         & Noncvx & Cvx & Strongly Cvx\\ \hline
	Accelerated	SGD   & \cite{huo2020faster} A\ref{ass:lsmooth},\ref{ass:lgradvar} &          &             \\\hline
	Accelerated GD    &        &       &     \cite{liu2019accelerating}A\ref{ass:lipcont},\ref{ass:lsmooth}        \\\hline
	SGD    &   \cite{haddadpour2019convergence} A\ref{ass:lsmooth},\ref{ass:lgradvar},(\ref{ass:pl},\ref{ass:bgd}) ; \cite{huo2020faster}A\ref{ass:lsmooth},\ref{ass:lgradvar}  & A\ref{ass:subgrad2}  &   \cite{haddadpour2019convergence} A\ref{ass:lsmooth},\ref{ass:lgradvar},(\ref{ass:bgd}); \cite{li2019convergence} A\ref{ass:lsmooth},\ref{ass:lgradvar},\ref{ass:squaregrad}     \\\hline
	GD     &   \cite{haddadpour2019convergence}A\ref{ass:lsmooth},(\ref{ass:pl},\ref{ass:bgd})    & \cite{khaled2019first} A\ref{ass:lsmooth} &        \cite{haddadpour2019convergence}A\ref{ass:lsmooth},(\ref{ass:pl},\ref{ass:bgd})    \\\hline
	\end{tabular}
	\caption{Summarize of assumptions on federated learning, heterogeneous data. $(\cdot)$ means that the assumption is not necessary.}
	\label{tb:ass2}
\end{table}

\begin{table}[h!]
\centering
\small
	\begin{tabular}{|c|c|c|c|c|}\hline
		paper         & Noncvx & Cvx and Non-smth & Cvx and L-smth & Strongly Cvx and L-smth\\ \hline
	Accelerated	SGD   & \cite{huo2020faster} A\ref{ass:lsmooth},\ref{ass:lgradvar} &    &      &             \\\hline
	Accelerated GD    &        &    &   &     \cite{liu2019accelerating}A\ref{ass:lipcont},\ref{ass:lsmooth}        \\\hline
	SGD    &   \cite{haddadpour2019convergence} A\ref{ass:lsmooth},\ref{ass:lgradvar},(\ref{ass:pl},\ref{ass:bgd}) ; \cite{huo2020faster}A\ref{ass:lsmooth},\ref{ass:lgradvar}  & A\ref{ass:subgrad2}  &  &   \cite{haddadpour2019convergence} A\ref{ass:lsmooth},\ref{ass:lgradvar},(\ref{ass:bgd}); \cite{li2019convergence} A\ref{ass:lsmooth},\ref{ass:lgradvar},\ref{ass:squaregrad}     \\\hline
	GD     &   \cite{haddadpour2019convergence}A\ref{ass:lsmooth},(\ref{ass:pl},\ref{ass:bgd})     &  & \cite{khaled2019first} A\ref{ass:lsmooth} &        \cite{haddadpour2019convergence}A\ref{ass:lsmooth},(\ref{ass:pl},\ref{ass:bgd})    \\\hline
	\end{tabular}
	\caption{Summarize of assumptions on federated learning, heterogeneous data. $(\cdot)$ means that the assumption is not necessary.}
	\label{tb:ass}
\end{table}
\vspace{-1em}
\begin{table}[h!]
\centering
\small
	\begin{tabular}{|c|c|c|c|c|}\hline
		paper         & Noncvx & Cvx and Non-smth & Cvx and L-smth & Strongly Cvx and L-smth\\ \hline
	Accelerated	SGD   & \cite{huo2020faster} &        &    &             \\\hline
	Accelerated GD    &        &    &    &    \cite{liu2019accelerating}        \\\hline
	SGD    &   \cite{haddadpour2019convergence,huo2020faster}    & Ours &     &    \cite{li2019convergence}       \\\hline
	GD     &   \cite{haddadpour2019convergence}     &  & \cite{khaled2019first}     &         \\\hline
	\end{tabular}
	\caption{Summarize of relatedwork on federated learning, heterogeneous data.}
\end{table}

Note for table~\ref{tb:ass}:
\begin{itemize}
	\item \textbf{GD}: \cite{haddadpour2019convergence} the proof of Noncvx can be used directly for Strongly cvx and L-smooth.
	\item \textbf{SGD}: necessary assumption: Bounded gradient variance assumption. 
	\item \textbf{Non Cvx}: necessary assumption:  L-smooth.
	\item \textbf{Convex}:
\end{itemize}

\textbf{GD}
The proof of convergence rate of local GD in~\cite{haddadpour2019convergence} 
is based A~\ref{ass:lsmooth} and A~\ref{ass:pl}, note that PL is a generalization of $\mu$-strongly convexity. $\mu$-strongly convexity
indicates A~\ref{ass:pl} is satisfied. Therefore, the same proof in~\cite{haddadpour2019convergence} for non-convex function 
can be adapted for strongly convex and $L$-smooth function.

\begin{itemize}
	\item Online Meta learning~\cite{finn2019online} strongly convex case convergence rate~\cite{lin2020collaborative}. 
	\item FL + bandit? 
	\item FL + TD convergence rate.
\end{itemize}

### 2020-03-20 source/sec_convexsmoothsgd.tex
first commit

% !TEX ROOT=./main.tex

\textbf{Full Device Participation}

\begin{itemize}
	\item Stochastic gradient of device $k$ at time step $t$, at point $w_{t,k}$: 
	$$\vg_{t,k} \coloneqq \vg_{t,k}(w_t^k)$$
	$$ \vg_{t, k} = \grad F_{k}\left(w_{t}^{k}, \xi_{t}^{k}\right) $$
	$$\EE \vg_{t, k} = \grad F_{k}\left(w_{t}^{k}\right)$$
\item  One-step stochastic subgradient of all devices.
$$\vg_{t}=\sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) $$
\begin{align}
	\EE \vg_{t}= \EE \sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) \coloneqq \sum_{k=1}^{N} p_{k} \EE \vg_{t, k}
	\label{eq:egtsgd}
\end{align}
\end{itemize}

$$\Delta_{t+1} = \EE \|\overline{\vw}_{t+1} - \vw^*\|^2 = \EE \|\overline{\vv}_{t+1} - \vw^*\|^2,$$
According to the definition of $\overline{\vv}_{t+1}$ in \eq{\ref{eq:vbar}}, we can expand $\Delta_{t+1}$
$$\begin{aligned}\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2 &=\left\|\bar{\vw}_{t}-\eta_{t} \vg_{t}-\vw^{*}\right\|^2 \\ &=\left\|\bar{\vw}_{t}-\vw^{*}\right\|^{2}-2 \eta_{t} \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)+\eta_{t}^{2}\left\|\vg_{t}\right\|^{2} \end{aligned}$$

Take the expectation condition on $\vw_t$ over random samples at all devices:
\begin{align}
\Delta_{t+1} = \EE\left[\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2| \vw_{t}\right]=\|\bar{\vw}_{t}-\vw^{*}\|^{2}-2 \eta_{t} \EE \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)+\eta_{t}^{2} \EE\| \vg_{t} \|^{2}	
\label{eq:expandsgd}
\end{align}

Now we focus on bounding $-2 \eta_{t} \EE \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)$ in \eq{\ref{eq:expandsgd}}:

### 2020-03-20 source/sec_cvxnonsm.tex
first commit

% !TEX ROOT=./main.tex

\textbf{Full Device Participation}

\begin{itemize}
	\item Stochastic Subgradient of device $k$ at time step $t$, at point $w_{t,k}$: 
	$$\vg_{t,k} \coloneqq \vg_{t,k}(w_t^k)$$
	$$ \vg_{t, k} \in \partial F_{k}\left(w_{t}^{k}, \xi_{t}^{k}\right) $$
	$$\EE \vg_{t, k} \in \partial F_{k}\left(w_{t}^{k}\right)$$
\item  One-step stochastic subgradient of all devices.
$$\vg_{t}=\sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) $$
\begin{align}
	\EE \vg_{t}= \EE \sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) \coloneqq \sum_{k=1}^{N} p_{k} \EE \vg_{t, k}
	\label{eq:egt}
\end{align}
\end{itemize}

\begin{assumption}
The expected squared norm of stochastic subgradients is uniformly bounded. i.e.,
$\mathbb{E}\|\vg_{t,k}\|^2  \leq G_k^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.  This also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$.
\label{ass:subgrad2}
\end{assumption}

\begin{theorem}
	Let Assumption~\ref{ass:subgrad2} hold, 
	then the FedAve with full device participation satisfies
	\begin{align}
		 F^*_t - F^* &\leq \frac{RL}{\sqrt{T}}
	\end{align}
	where $$\eta_t = \frac{R}{\sqrt{T}},$$ 
	$$F^*_t \coloneqq \min_{t \in [0, T-1]} F(\overline{\vw}_t),$$
	$$R = \sqrt{ \frac{\Delta_0}{L}},$$ 
	$$L=\sum_{k=1}^N \left( p_k G_k^2 \left(3 + 4(E-1)^2\right)\right).$$ 
	\label{th:cvxnonsmoth}
\end{theorem}
{\color{red}{This bound didn't consider heterogeneous of devices $\Gamma$ in~\cite{li2019convergence}}}

In full device active setting, we have $\overline{\vw}_{t+1}=\overline{\vv}_{t+1}$,
$$\Delta_{t+1} = \EE \|\overline{\vw}_{t+1} - \vw^*\|^2 = \EE \|\overline{\vv}_{t+1} - \vw^*\|^2,$$
According to the definition of $\overline{\vv}_{t+1}$ in \eq{\ref{eq:vbar}}, we can expand $\Delta_{t+1}$
$$\begin{aligned}\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2 &=\left\|\bar{\vw}_{t}-\eta_{t} \vg_{t}-\vw^{*}\right\|^2 \\ &=\left\|\bar{\vw}_{t}-\vw^{*}\right\|^{2}-2 \eta_{t} \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)+\eta_{t}^{2}\left\|\vg_{t}\right\|^{2} \end{aligned}$$

Take the expectation condition on $\vw_t$ over random samples at all devices:
\begin{align}
\Delta_{t+1} = \EE\left[\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2| \vw_{t}\right]=\|\bar{\vw}_{t}-\vw^{*}\|^{2}-2 \eta_{t} \EE \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)+\eta_{t}^{2} \EE\| \vg_{t} \|^{2}	
\label{eq:expand}
\end{align}
Note that since $\EE \vg_t = \EE \sum_{k=1}^{N} p_{k} \vg_{t, k}\left(\vw_{t}^{k}\right) \neq \EE \sum_{k=1}^Np_k \vg_{t,k}(\overline{\vw}_t)$, therefore $\EE \vg_t \notin \partial F(\overline(\vw_t))$, we cannot directly use the convex definition to
upper bound $- \EE \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)$ in \eq{\ref{eq:expand}}, as classic SGD proof did.

Now we focus on bounding $-2 \eta_{t} \EE \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)$ in \eq{\ref{eq:expand}}: 
\begin{align}
	& -2 \eta_{t} \left<\EE \vg_{t}, \bar{\vw}_{t}-\vw^{*} \right>\\
  = & 2 \eta_{t} \sum_{k=1}^N p_k \left(-\left<\EE \vg_{t,k}, \bar{\vw}_{t}-\vw^{*} \right> \right) \\
  \leq& 2 \eta_{t} \sum_{k=1}^N p_k  \left(\frac{1}{2\eta_t} \|\overline{\vw}_t - \vw_{t,k}\|^2 + \frac{\eta_t}{2} \|\EE \vg_{t,k}\|^2 - (F_k(\vw_t^k) - F_k(\vw^*) \right) \\
  =& \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_{t,k}\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2 
  - 2\eta_t\sum_{k=1}^N p_k (F_k(\vw_t^k) - F_k(\vw^*)) \\
  =& \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_{t,k}\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2 
  - 2\eta_t\sum_{k=1}^N p_k (F_k(\vw_t^k) - F^*) \label{eq:exp4}
\end{align}
where in the third line we use the following derivation:
\begin{align}
	& -\left<\EE \vg_{t,k}, \bar{\vw}_{t} - \vw^{*} \right>\\
=& - \left<\EE \vg_{t,k}, \bar{\vw}_{t} - \vw_t^k +  \vw_t^k - \vw^{*} \right>\\
=& - \left<\EE \vg_{t,k}, \bar{\vw}_{t} - \vw_t^k\right> - \left<\EE \vg_{t,k}, \vw_t^k - \vw^{*} \right>\\
\leq & - \left<\EE \vg_{t,k}, \bar{\vw}_{t} - \vw_t^k\right> - (F_k(\vw_t^k) - F_k(\vw^*)) \label{eq:3}\\
\leq & \frac{1}{2\eta_t} \|\vw_t - \vw_{t,k}\|^2 + \frac{\eta_t}{2} \|\EE \vg_{t,k}\|^2
- (F_k(\vw_t^k) - F_k(\vw^*)) \label{eq:4}
\end{align}
where in \eq{\ref{eq:3}} we use the definition of convexity and in \eq{\ref{eq:4}} we use Cauchy-Schwarz inequality and AM-GM inequality $-\langle\va, \vb\rangle=\langle\va,-\vb\rangle \leqslant|\va||\vb| \leqslant \frac{1}{2}\left(|\va|^{2}+|\vb|^{2}\right)$.

Now we plug in \eq{\ref{eq:exp4}} into \eq{\ref{eq:expand}}:
\begin{align}
	\Delta_{t+1} \leq& \|\bar{\vw}_{t}-\vw^{*}\|^{2} + \eta_{t}^{2} \EE\| \vg_{t} \|^{2} \\
	& + \underbrace{\sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_{t,k}\|^2}_{A_1} +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2
  - 2\eta_t\underbrace{\sum_{k=1}^N p_k (F_k(\vw_t^k) - F^*)}_{A_2} \label{eq:main2}
\end{align}

$A_1$ can be bounded by Lemma 3 in~\cite{li2019convergence},

Now we bound $A_2$:
\begin{align}
	& \sum_{k=1}^N p_k (F_k(\vw_t^k) - F^*) \\
 = & \sum_{k=1}^N p_k (F_k(\vw_t^k) - F_k(\overline{w}_t) + F_k(\overline{w}_t) - F^*) \\
 = & \sum_{k=1}^N p_k (F_k(\vw_t^k) - F_k(\overline{w}_t) ) + (F(\overline{w}_t) - F^*)\\ 
 \geq & \sum_{k=1}^{N} P_{k}\left<\EE g_{t k}\left(\overline{\vw}_{t}\right), \vw_{t}^k-\overline{\vw}_{t}\right> + (F(\overline{w}_t) - F^*)\\
 \geq& -\frac{1}{2} \sum_{k=1}^{N} P_{k} (\eta_t \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \frac{1}{\eta_t} \|\vw_{t,k} - \overline{\vw}_t\|^2) + (F(\overline{w}_t) - F^*) \label{eq:a24}
\end{align}
where in fourth line we use Cauchy-Schwarz inequality and AM-GM inequality.

Now we plug in \eq{\ref{eq:a24}} into \eq{\ref{eq:main2}}:
\begin{align}
	\Delta_{t+1} & \leq \Delta_t + 
	\eta_{t}^{2} \EE\| \vg_{t} \|^{2} + \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_{t,k}\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2 \\
	 &- 2 \eta_t \left(-\frac{1}{2} \sum_{k=1}^{N} P_{k} (\eta_t \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \frac{1}{\eta_t} \|\vw_{t,k} - \overline{\vw}_t\|^2) + (F(\overline{w}_t) - F^*) \right)\\
	& \leq \Delta_t + 
	\eta_{t}^{2} \EE\| \vg_{t} \|^{2} + \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_{t,k}\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2 \\
	& +\sum_{k=1}^{N} P_{k} (\eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \|\vw_{t,k} - \overline{\vw}_t\|^2) - 2\eta_t (F(\overline{w}_t) - F^*)\\
	& = \Delta_t + \underbrace{\eta_{t}^{2} \EE\| \vg_{t} \|^{2} +\sum_{k=1}^N p_k  ( 2\|\overline{\vw}_t - \vw_{t,k}\|^2  +  \eta_t^2 \|\EE \vg_{t,k}\|^2 + \eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 )}_{C_1 = \eta_t^2L}  - 2\eta_t (F(\overline{w}_t) - F^*) \label{eq:main3}
\end{align}
Rearrange the last inequality:
\begin{align}
	2\eta_t (F(\overline{w}_t) - F^*) \leq \Delta_t - \Delta_{t+1} + \eta_t^2 L
\end{align}
Sum over two sides from $t=0$ to $t = T- 1$ and define $F^*_t \coloneqq \min_{t \in [0, T-1]} F(\overline{\vw}_t)$
\begin{align}
	\sum_{t=0}^{T-1} 2\eta_t(F(\overline{w}_t) - F^*) & \leq \Delta_0 - \Delta_T + \sum_{t=0}^{T-1} \eta_t^2 L\\
	\sum_{t=0}^{T-1} 2\eta_t(F(\overline{w}_t) - F^*) & \leq \Delta_0 + \sum_{t=0}^{T-1} \eta_t^2 L\\
    F^*_t - F^* &\leq \frac{\Delta_0 + \sum_{t=0}^{T-1} \eta_t^2L}{ \sum_{t=0}^{T-1}  2 \eta_t}\\
    F^*_t - F^* &\leq \frac{RL}{\sqrt{T}}\\
\end{align}
where in last line we let $R = \sqrt{ \frac{\Delta_0}{L}}, \eta_t = \frac{R}{\sqrt{T}}$

Now we discuss how to bound the term $C_1$ in \eq{\ref{eq:main3}}

Assume Assumption~\ref{ass:subgrad2}, $\eta_t$ is non-increasing, and $\eta_t \leq 2\eta_t+E$ for all $t\geq 0$, according to Lemma 3 in~\cite{li2019convergence}, we have

\begin{equation}
\mathbb{E}\left[\sum_{k=1}^{N} p_{k}\left\|\overline{\mathbf{w}}_{t}-\mathbf{w}_{k}^{t}\right\|^{2}\right] \leq \sum_{k=1}^N p_k4 \eta_{t}^{2}(E-1)^{2} G_k^{2}
	\label{eq:g4}
\end{equation}

Assume Assumption~\ref{ass:subgrad2}, we have 
\begin{equation}
	\eta_t^2 \|\EE \vg_{t,k}\|^2 \leq  \eta_t^2 G_k^2
	\label{eq:g3}
\end{equation}
and
\begin{equation}
	\eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t)\|^2 \leq  \eta_t^2 G_k^2
	\label{eq:g2}
\end{equation}

Now the only term left in $C_1$ is $\eta_{t}^{2} \EE\| \vg_{t} \|^{2} $, follow the proof in Lemma 2~\cite{li2019convergence}: 
\begin{align}
	  \eta_{t}^{2} \EE\| \vg_{t} \|^{2} 
	\leq   \eta_{t}^{2} \EE\| \sum_{k=1}^N p_k g_{t,k} \|^{2} 
	\leq   \eta_{t}^{2} \sum_{k=1}^N p_k \EE\| g_{t,k} \|^{2} 
	\leq   \eta_{t}^{2} \sum_{k=1}^N p_k G_k^2  \label{eq:g1}
\end{align}
where in the third inequality we use the convexity of the l2 norm.

Combine \eq{\ref{eq:g4}},\eq{\ref{eq:g3}},\eq{\ref{eq:g2}}, \eq{\ref{eq:g1}},
we have 
\begin{align}
C_1 = \eta_t^2 L  = \eta_t^2 \sum_{k=1}^N p_k G_k^2 \left(3 + 4(E-1)^2\right)	
\end{align}

\textbf{Problems}
\begin{itemize}
	\item The bound in Theorem~\ref{th:cvxnonsmoth} didn't incoporate heterogeneous of data.
	\item \cite{huo2020faster} this paper has no assumption on the heterogeneous, the bounded gradient variance can cover this assumption. 
\end{itemize}

### 2020-03-20 source/sec_problem.tex
first commit

% !TEX ROOT=./main.tex

\section{Problem Setting}

\begin{align}
	\min _{\mathbf{w}}\left\{F(\mathbf{w}) \triangleq \sum_{k=1}^{N} p_{k} F_{k}(\mathbf{w})\right\}
	\label{eq:problem}
\end{align}
where $N$ is the number of devices, and $p_k$ is the weight of the k-th device
such that $p_k \geq 0$ and $\sum_{k=1}^N p_k = 1$. Suppose the k-th device
holds the nk training data: $x_{k,1}, x_{k,2}, \dots, x_{k,n_k}$ . The local
$k \in S_t^K$ objective $F_k(\cdot)$ is defined by

\begin{align}
F_{k}(\mathbf{w}) \triangleq \frac{1}{n_{k}} \sum_{j=1}^{n_{k}} \ell\left(\mathbf{w} ; x_{k, j}\right)	
\label{eq:localloss}
\end{align}

% The summary of all assumptions:

% \begin{assumption}
% $F_{1}, \cdots, F_{N}$ are all $L$-smooth: for all  $\mathbf{v}$  and $\mathbf{w}$, $F_{k}(\mathbf{v}) \leq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{L}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$.
% \end{assumption}

% \begin{assumption}
% $F_1,\dots, F_N$ are all convex: for all $\vv$ and $\vw$, 
% $F_k(\vv) \geq F_k(\vw)+(\vv -\vw)^T \grad F_k(\vw)$. \label{ass:cvx}
% \end{assumption}
% \begin{assumption}
% $	F_{1}, \cdots, F_{N} \text { are all } \mu \text { -strongly convex: for all v and } \mathbf{w}, F_{k}(\mathbf{v}) \geq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{\mu}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$
% \end{assumption}

% \begin{assumption}
% Let $\xi_{t}^{k}$ be sampled from the k-th device's local data uniformly at random. The variance of stochastic gradients in each device is bounded:
% $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq \sigma_{k}^{2}$, for $k = 1,..., N$.
% \end{assumption}

% \begin{assumption}
% The expected squared norm of stochastic gradients is uniformly bounded. i.e.,
% $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)\right\|^{2} \leq G^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
% \end{assumption}

\section{Notations}

\begin{itemize}
	\item $\vw_t^k$: be the model parameter maintained in the k-th device at the t-th step. or the parameter obtained after communication steps. 
	\item $\cI_E$: the set of global synchronization steps, $\cI_E = \{n E|n = 1, 2, 3,\dots \}$. If $t+1 \in \cI_E$, we send $w_{t+1}$ to all devices.
	\item $\vv^k_{t+1}$: the immediate result of one step SGD update from $\vw^k_{t}$.
	\item $\xi_{t}^{k}$: the data sampled from k-th device’s local data 
	uniformly at random.
	\item $\overline{\mathbf{v}}_{t}=\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$.
	\item $\overline{\mathbf{w}}_{t}=\sum_{k=1}^{N} p_{k} \mathbf{w}_{t}^{k}$.
	\item $\overline{\mathbf{v}}_{t+1}$ is one-step SGD from $\overline{\mathbf{w}}_{t}$. $t+1 \in \cI_E$, we can fetch $\overline{\vw}_{t+1}$, we can communicate $\overline{\vw}_{t+1}$ to all devices.
	\item The expected one-step graident, $\overline{\mathbf{g}}_{t}=\sum_{k=1}^{N} p_{k} \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)$,  
	\item One-step stochastic gradient $\mathbf{g}_{t}=\sum_{k=1}^{N} p_{k} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)$. $\EE[\vg_t] = \overline{g}_t$
	\item

	\begin{align}
	\overline{\mathbf{v}}_{t+1}=\overline{\mathbf{w}}_{t}-\eta_{t} \mathbf{g}_{t}	\label{eq:vbar}
	\end{align}

\end{itemize}

\begin{align} 
\mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & (9) \\ \sum_{k=1}^{N} p_{k} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}\end{array}\right.
\end{align}

### 2020-03-27 source/main.tex
accelerated sgd, cvx, non-smoth v1

@@
 \newcommand{\lkxcom}[1]{{\color{red}{#1}}}
+
+\newcommand{\ov}[1]{{\overline{\mathbf{#1}}}}
 \input{bmacros}
@@
 
-%% Note for ICLR'20 paper
-% \input{note_sec_problem}
-% \input{note_sec_algorithm}
-% \input{note_sec_part}
-% \input{note_sec_q}
-
 % converge proof
@@
 \input{sec_overview}
-\newpage
-% \input{sec_acceleratedFL}
-% \input{sec_cvxnonsm}
+
+\input{sec_acceleratedFL}
 
 % accelerated FL survey
+
+% paper
+
+\begin{itemize}
+	\item On Federated Learning of Deep Networks from Non-IID Data: Parameter Divergence and the Effects of Hyperparametric Methods: empirical evaluation 
+	paper, the loss surface could be helpful.
+	\item On the Linear Speedup Analysis of Communication Efficient Momentum SGD for Distributed Non-Convex Optimization: Non-convex, smooth, accelerated SGD, Full device particaption.
+\end{itemize}
 

### 2020-03-30 source/sec_convexsmoothsgd.tex
add SGD convex smooth.

@@
 According to the definition of $\overline{\vv}_{t+1}$ in \eq{\ref{eq:vbar}}, we can expand $\Delta_{t+1}$
-$$\begin{aligned}\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2 &=\left\|\bar{\vw}_{t}-\eta_{t} \vg_{t}-\vw^{*}\right\|^2 \\ &=\left\|\bar{\vw}_{t}-\vw^{*}\right\|^{2}-2 \eta_{t} \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)+\eta_{t}^{2}\left\|\vg_{t}\right\|^{2} \end{aligned}$$
+$$\begin{aligned}\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2 &=\left\|\ov{w}_{t}-\eta_{t} \vg_{t}-\vw^{*}\right\|^2 \\ &=\left\|\ov{w}_{t}-\vw^{*}\right\|^{2}-2 \eta_{t} \vg_{t}^{\top}\left(\ov{w}_{t}-\vw^{*}\right)+\eta_{t}^{2}\left\|\vg_{t}\right\|^{2} \end{aligned}$$
 
@@
 \begin{align}
-\Delta_{t+1} = \EE\left[\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2| \vw_{t}\right]=\|\bar{\vw}_{t}-\vw^{*}\|^{2}-2 \eta_{t} \EE \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)+\eta_{t}^{2} \EE\| \vg_{t} \|^{2}
+\Delta_{t+1} = \EE\left[\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2| \vw_{t}\right]=\|\ov{w}_{t}-\vw^{*}\|^{2}-2 \eta_{t} \EE \vg_{t}^{\top}\left(\ov{w}_{t}-\vw^{*}\right)+\eta_{t}^{2} \EE\| \vg_{t} \|^{2}
 \label{eq:expandsgd}
@@
 
-Now we focus on bounding $-2 \eta_{t} \EE \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)$ in \eq{\ref{eq:expandsgd}}:
+Now we focus on bounding $-2 \eta_{t} \EE \vg_{t}^{\top}\left(\ov{w}_{t}-\vw^{*}\right)$ in \eq{\ref{eq:expandsgd}}:
+
+\begin{align*}
+	& -2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}-\vw^{*}\right> \\
+ =  & -2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}- \vw^{k}_t \right> -2 \eta_{t} \left<\EE \vg_{t}, \vw^{k}_t - \vw^{*}\right>\\
+ \leq & -2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}- \vw^{k}_t \right> + 2 \eta_{t} (F_k(w^*) - F_k(\vw^{k}_t))\\
+ \leq &\sum_{k=1}^N p_k \left[2 \eta_{t} (F_k(\vw^{k}_t) - F_k(\ov{w}_t) + \frac{L}{2} \|\ov{w}_t - \vw^{k}_t\|^2 ) + 2 \eta_{t} (F_k(w^*) - F_k(\vw^{k}_t))\right]\\
+ = & \sum_{k=1}^N p_k \eta_t L \|\ov{w}_t - \vw^{k}_t\|^2 + 2 \eta_{t} \sum_{k=1}^N p_k (F_k(w^*) - F_k(\ov{w}_t))\\
+ = &  \eta_t L \sum_{k=1}^N p_k \|\ov{w}_t - \vw^{k}_t\|^2 + 2 \eta_{t} (F^* - F(\ov{w}_t))
+\end{align*}
+Plug in this upper bound into \eq{\ref{eq:expandsgd}}, we have
+
+\begin{align}
+\EE\left\|\ov{w}_{t+1}-\vw^{*}\right\|^2 &\leq \|\ov{w}_{t}-\vw^{*}\|^{2}+\eta_{t}^{2} \EE\| \vg_{t} \|^{2} + \eta_t L \sum_{k=1}^N p_k \|\ov{w}_t - \vw^{k}_t\|^2 + 2 \eta_{t} (F^* - F(\ov{w}_t))\\
+&\leq \|\ov{w}_{t}-\vw^{*}\|^{2}+\eta_{t}^{2} \EE\| \vg_{t} \|^{2} +  4L\eta_t^3(E-1)^2 G^2 + 2 \eta_{t} (F^* - F(\ov{w}_t))\\
+&\leq \|\ov{w}_{t}-\vw^{*}\|^{2}+\eta_{t}^{2} G^2 +  4L\eta_t^3(E-1)^2 G^2 + 2 \eta_{t} (F^* - F(\ov{w}_t)) \label{eq:cvxsgd1}
+\end{align}
+
+Sum two sides of \eq{\ref{eq:cvxsgd1}}, we assume \lkxcom{$\eta_t < 1$}, set $F^*_t = \min_{t \in [0, T-1]} F(\ov{w}_t)$, 
+\begin{align*}
+	\sum_{t=0}^{T-1} 2\eta_t (F(\ov{w}_t) - F^* ) & \leq \EE \|\ov{w}_{0}-\vw^{*}\|^{2} - \EE\left\|\ov{w}_{T}-\vw^{*}\right\|^2 + \sum_{t=0}^{T-1} (\eta_t^3 4LG^2(E-1)^2 + \eta_t^2 G^2)\\ 
+    & \leq \EE \|\ov{w}_{0}-\vw^{*}\|^{2} - \EE\left\|\ov{w}_{T}-\vw^{*}\right\|^2 + \sum_{t=0}^{T-1} (\eta_t^3 4LG^2(E-1)^2 + \eta_t^2 G^2)\\ 
+    & \leq \EE \|\ov{w}_{0}-\vw^{*}\|^{2} + \sum_{t=0}^{T-1}\eta_t^2G^2 ( 4L(E-1)^2 + 1)\\
+ F_t^* - F^*  & \leq \frac{\EE \|\ov{w}_{0}-\vw^{*}\|^{2}}{2 \sum_{t=0}^{T-1} \eta_t } + ( 4L(E-1)^2 + 1)G^2 \sum_{t=0}^{T-1} \eta_t^2
+\end{align*}
+It will converge under three conditions $ \lim_{T \rightarrow \infty }\sum_{t=0}^{T-1} \eta_t = \infty$
+$ \lim_{T \rightarrow \infty }\sum_{t=0}^{T-1} \eta_t^2 < \infty$. For example, we can set $\eta_t = \frac{1}{t+1}$.

### 2020-03-30 source/sec_cvxnonsm.tex
accelerated sgd, cvx, non-smoth v1
partial sgd v1

@@
 The expected squared norm of stochastic subgradients is uniformly bounded. i.e.,
-$\mathbb{E}\|\vg_{t,k}\|^2  \leq G_k^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.  This also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$.
+$\mathbb{E}\|\vg_{t,k}\|^2  \leq G_k^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
+This also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$.
+Denote $G^2 = \sum_{k=1}^N p_k G_k^2$
 \label{ass:subgrad2}
@@
 \end{align}
-Note that since $\EE \vg_t = \EE \sum_{k=1}^{N} p_{k} \vg_{t, k}\left(\vw_{t}^{k}\right) \neq \EE \sum_{k=1}^Np_k \vg_{t,k}(\overline{\vw}_t)$, therefore $\EE \vg_t \notin \partial F(\overline(\vw_t))$, we cannot directly use the convex definition to
+Note that since $\EE \vg_t = \EE \sum_{k=1}^{N} p_{k} \vg_{t, k}\left(\vw_{t}^{k}\right) \neq \EE \sum_{k=1}^Np_k \vg_{t,k}(\overline{\vw}_t)$, therefore $\EE \vg_t \notin \partial F(\overline{\vw}_t)$, we cannot directly use the convex definition to
 upper bound $- \EE \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)$ in \eq{\ref{eq:expand}}, as classic SGD proof did.
@@
   = & 2 \eta_{t} \sum_{k=1}^N p_k \left(-\left<\EE \vg_{t,k}, \bar{\vw}_{t}-\vw^{*} \right> \right) \\
-  \leq& 2 \eta_{t} \sum_{k=1}^N p_k  \left(\frac{1}{2\eta_t} \|\overline{\vw}_t - \vw_{t,k}\|^2 + \frac{\eta_t}{2} \|\EE \vg_{t,k}\|^2 - (F_k(\vw_t^k) - F_k(\vw^*) \right) \\
-  =& \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_{t,k}\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2 
+  \leq& 2 \eta_{t} \sum_{k=1}^N p_k  \left(\frac{1}{2\eta_t} \|\overline{\vw}_t - \vw_t^k\|^2 + \frac{\eta_t}{2} \|\EE \vg_{t,k}\|^2 - (F_k(\vw_t^k) - F_k(\vw^*) \right) \\
+  =& \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_t^k\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2 
   - 2\eta_t\sum_{k=1}^N p_k (F_k(\vw_t^k) - F_k(\vw^*)) \\
-  =& \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_{t,k}\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2 
+  =& \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_t^k\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2 
   - 2\eta_t\sum_{k=1}^N p_k (F_k(\vw_t^k) - F^*) \label{eq:exp4}
@@
 \leq & - \left<\EE \vg_{t,k}, \bar{\vw}_{t} - \vw_t^k\right> - (F_k(\vw_t^k) - F_k(\vw^*)) \label{eq:3}\\
-\leq & \frac{1}{2\eta_t} \|\vw_t - \vw_{t,k}\|^2 + \frac{\eta_t}{2} \|\EE \vg_{t,k}\|^2
+\leq & \frac{1}{2\eta_t} \|\bar{\vw}_t - \vw_t^k\|^2 + \frac{\eta_t}{2} \|\EE \vg_{t,k}\|^2
 - (F_k(\vw_t^k) - F_k(\vw^*)) \label{eq:4}
@@
 	\Delta_{t+1} \leq& \|\bar{\vw}_{t}-\vw^{*}\|^{2} + \eta_{t}^{2} \EE\| \vg_{t} \|^{2} \\
-	& + \underbrace{\sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_{t,k}\|^2}_{A_1} +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2
+	& + \underbrace{\sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_t^k\|^2}_{A_1} +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2
   - 2\eta_t\underbrace{\sum_{k=1}^N p_k (F_k(\vw_t^k) - F^*)}_{A_2} \label{eq:main2}
@@
  \geq & \sum_{k=1}^{N} P_{k}\left<\EE g_{t k}\left(\overline{\vw}_{t}\right), \vw_{t}^k-\overline{\vw}_{t}\right> + (F(\overline{w}_t) - F^*)\\
- \geq& -\frac{1}{2} \sum_{k=1}^{N} P_{k} (\eta_t \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \frac{1}{\eta_t} \|\vw_{t,k} - \overline{\vw}_t\|^2) + (F(\overline{w}_t) - F^*) \label{eq:a24}
+ \geq& -\frac{1}{2} \sum_{k=1}^{N} P_{k} (\eta_t \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \frac{1}{\eta_t} \|\vw_t^k - \overline{\vw}_t\|^2) + (F(\overline{w}_t) - F^*) \label{eq:a24}
 \end{align}
@@
 	\Delta_{t+1} & \leq \Delta_t + 
-	\eta_{t}^{2} \EE\| \vg_{t} \|^{2} + \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_{t,k}\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2 \\
-	 &- 2 \eta_t \left(-\frac{1}{2} \sum_{k=1}^{N} P_{k} (\eta_t \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \frac{1}{\eta_t} \|\vw_{t,k} - \overline{\vw}_t\|^2) + (F(\overline{w}_t) - F^*) \right)\\
+	\eta_{t}^{2} \EE\| \vg_{t} \|^{2} + \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_t^k\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2 \\
+	 &- 2 \eta_t \left(-\frac{1}{2} \sum_{k=1}^{N} P_{k} (\eta_t \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \frac{1}{\eta_t} \|\vw_t^k - \overline{\vw}_t\|^2) + (F(\overline{w}_t) - F^*) \right)\\
 	& \leq \Delta_t + 
-	\eta_{t}^{2} \EE\| \vg_{t} \|^{2} + \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_{t,k}\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2 \\
-	& +\sum_{k=1}^{N} P_{k} (\eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \|\vw_{t,k} - \overline{\vw}_t\|^2) - 2\eta_t (F(\overline{w}_t) - F^*)\\
-	& = \Delta_t + \underbrace{\eta_{t}^{2} \EE\| \vg_{t} \|^{2} +\sum_{k=1}^N p_k  ( 2\|\overline{\vw}_t - \vw_{t,k}\|^2  +  \eta_t^2 \|\EE \vg_{t,k}\|^2 + \eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 )}_{C_1 = \eta_t^2L}  - 2\eta_t (F(\overline{w}_t) - F^*) \label{eq:main3}
+	\eta_{t}^{2} \EE\| \vg_{t} \|^{2} + \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_t^k\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2 \\
+	& +\sum_{k=1}^{N} P_{k} (\eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \|\vw_t^k - \overline{\vw}_t\|^2) - 2\eta_t (F(\overline{w}_t) - F^*)\\
+	& = \Delta_t + \underbrace{\eta_{t}^{2} \EE\| \vg_{t} \|^{2} +\sum_{k=1}^N p_k  ( 2\|\overline{\vw}_t - \vw_t^k\|^2  +  \eta_t^2 \|\EE \vg_{t,k}\|^2 + \eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 )}_{C_1 = \eta_t^2L}  - 2\eta_t (F(\overline{w}_t) - F^*) \label{eq:main3}
 \end{align}

### 2020-04-01 source/sec_convexsmoothsgd.tex
partial sgd v1
add partial activation to cvx smth.

--- 2020-03-20 source/sec_convexsmoothsgd.tex
@@
 
-\textbf{Full Device Participation}
+Stochastic gradient of device $k$ at time step $t$, at point $\vw_t^k$: 
+	$\vg_{t,k} \coloneqq \vg_{t,k}(w_t^k)$,
+	$ \vg_{t, k} = \grad F_{k}\left(w_{t}^{k}, \xi_{t}^{k}\right) $,
+	$\EE \vg_{t, k} = \grad F_{k}\left(w_{t}^{k}\right)$.
+One-step stochastic gradient of all devices:
+$\vg_{t}=\sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) $, $\EE \vg_{t}= \EE \sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) \coloneqq \sum_{k=1}^{N} p_{k} \EE \vg_{t, k}$
 
-\begin{itemize}
-	\item Stochastic gradient of device $k$ at time step $t$, at point $w_{t,k}$: 
-	$$\vg_{t,k} \coloneqq \vg_{t,k}(w_t^k)$$
-	$$ \vg_{t, k} = \grad F_{k}\left(w_{t}^{k}, \xi_{t}^{k}\right) $$
-	$$\EE \vg_{t, k} = \grad F_{k}\left(w_{t}^{k}\right)$$
-\item  One-step stochastic subgradient of all devices.
-$$\vg_{t}=\sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) $$
+\begin{theorem}
+Let Assumption~\ref{ass:subgrad2} and Assumption~\ref{ass:lsmooth} hold, suppose we have a bound 
+on our starting distance, i.e., $\|\vw_{0} - \vw^*\|^2 \leq \Delta_0$, set learning rate $\eta_t = \sqrt{ \frac{\Delta_0}{ T [G^2( 4L(E-1)^2 + 1) + C] }}$, we have,
+$$\EE[ F_t^*] - F^*  \leq \frac{R [G^2( 4L(E-1)^2 + 1) + C] }{\sqrt{T}}$$
+where we denote $R = \sqrt{ \frac{\Delta_0}{ G^2( 4L(E-1)^2 + 1) + C }}$.
+\end{theorem}
+
+\begin{proof}
+
 \begin{align}
-	\EE \vg_{t}= \EE \sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) \coloneqq \sum_{k=1}^{N} p_{k} \EE \vg_{t, k}
-	\label{eq:egtsgd}
+\EE_{\cS_{t+1}, \xi_{t}} \|\overline{\vw}_{t+1} - \vw^*\|^2 &= \EE_{\cS_{t+1}, \xi_{t}} \|\overline{\vw}_{t+1} - \overline{\vv}_{t+1} + \overline{\vv}_{t+1} - \vw^*\|^2\\
+&= \EE_{\cS_{t+1}, \xi_{t}} \left[\|\overline{\vw}_{t+1} - \overline{\vv}_{t+1}\|^2 + \|\overline{\vv}_{t+1} - \vw^*\|^2\right] \\
+& \hspace{3em}+ 2\EE_{\xi_{t}} \left<\EE_{\cS_{t+1}} \overline{\vw}_{t+1} - \overline{\vv}_{t+1},   \overline{\vv}_{t+1} - \vw^*\right> \\
+& = \EE_{\cS_{t+1}, \xi_{t}} \|\overline{\vw}_{t+1} - \overline{\vv}_{t+1}\|^2 + \EE_{\xi_t} \|\overline{\vv}_{t+1} - \vw^*\|^2 \\
+& \leq  \eta_t^2 C + \EE_{\xi_t} \|\overline{\vv}_{t+1} - \vw^*\|^2 \label{eq:sgdcvxsmth1}
 \end{align}
-\end{itemize}
+where in the second line we use $\EE_{\cS_{t+1}} \overline{\vw}_{t+1}  = \overline{\vv}_{t+1}$. 
+The first term can be bounded by consider different sampling scheme. 
+According to the definition of $\overline{\vv}_{t+1}$ in \eq{\ref{eq:vbar}}, we can expand the second term in \eq{\ref{eq:sgdcvxsmth1}} as follows:
+$$\begin{aligned}\left\|\ov{v}_{t+1}-\vw^{*}\right\|^2 &=\left\|\ov{w}_{t}-\eta_{t} \vg_{t}-\vw^{*}\right\|^2 \\ &=\left\|\ov{w}_{t}-\vw^{*}\right\|^{2}-2 \eta_{t} \vg_{t}^{\top}\left(\ov{w}_{t}-\vw^{*}\right)+\eta_{t}^{2}\left\|\vg_{t}\right\|^{2} \end{aligned}$$
 
-$$\Delta_{t+1} = \EE \|\overline{\vw}_{t+1} - \vw^*\|^2 = \EE \|\overline{\vv}_{t+1} - \vw^*\|^2,$$
-According to the definition of $\overline{\vv}_{t+1}$ in \eq{\ref{eq:vbar}}, we can expand $\Delta_{t+1}$
-$$\begin{aligned}\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2 &=\left\|\bar{\vw}_{t}-\eta_{t} \vg_{t}-\vw^{*}\right\|^2 \\ &=\left\|\bar{\vw}_{t}-\vw^{*}\right\|^{2}-2 \eta_{t} \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)+\eta_{t}^{2}\left\|\vg_{t}\right\|^{2} \end{aligned}$$
-
-Take the expectation condition on $\vw_t$ over random samples at all devices:
+Take the expectation condition on $\vw_t$ over $\xi_t$, i.e., random samples at all devices:
 \begin{align}
-\Delta_{t+1} = \EE\left[\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2| \vw_{t}\right]=\|\bar{\vw}_{t}-\vw^{*}\|^{2}-2 \eta_{t} \EE \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)+\eta_{t}^{2} \EE\| \vg_{t} \|^{2}
+\EE\left[\left\|\ov{v}_{t+1}-\vw^{*}\right\|^2| \vw_{t}\right]=\|\ov{w}_{t}-\vw^{*}\|^{2}-2 \eta_{t} \EE \vg_{t}^{\top}\left(\ov{w}_{t}-\vw^{*}\right)+\eta_{t}^{2} \EE\| \vg_{t} \|^{2}	
 \label{eq:expandsgd}
@@
 
-Now we focus on bounding $-2 \eta_{t} \EE \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)$ in \eq{\ref{eq:expandsgd}}:
+Now we focus on bounding $-2 \eta_{t} \EE \vg_{t}^{\top}\left(\ov{w}_{t}-\vw^{*}\right)$ in \eq{\ref{eq:expandsgd}}:
+
+\begin{align*}
+	& -2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}-\vw^{*}\right> \\
+ =  & -2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}- \vw^{k}_t \right> -2 \eta_{t} \left<\EE \vg_{t}, \vw^{k}_t - \vw^{*}\right>\\
+ \leq & -2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}- \vw^{k}_t \right> + 2 \eta_{t} (F_k(w^*) - F_k(\vw^{k}_t))\\
+ \leq &\sum_{k=1}^N p_k \left[2 \eta_{t} (F_k(\vw^{k}_t) - F_k(\ov{w}_t) + \frac{L}{2} \|\ov{w}_t - \vw^{k}_t\|^2 ) + 2 \eta_{t} (F_k(w^*) - F_k(\vw^{k}_t))\right]\\
+ = & \sum_{k=1}^N p_k \eta_t L \|\ov{w}_t - \vw^{k}_t\|^2 + 2 \eta_{t} \sum_{k=1}^N p_k (F_k(w^*) - F_k(\ov{w}_t))\\
+ = &  \eta_t L \sum_{k=1}^N p_k \|\ov{w}_t - \vw^{k}_t\|^2 + 2 \eta_{t} (F^* - F(\ov{w}_t))
+\end{align*}
+Plug in this upper bound into \eq{\ref{eq:expandsgd}}, \eq{\ref{eq:sgdcvxsmth1}} and take totoal expectation over all samples at all iterations, we have
+\begin{align}
+\EE\left\|\ov{w}_{t+1}-\vw^{*}\right\|^2 &\leq \EE\|\ov{w}_{t}-\vw^{*}\|^{2}+\eta_{t}^{2} \EE\| \vg_{t} \|^{2} + \eta_t L \sum_{k=1}^N p_k \EE \|\ov{w}_t - \vw^{k}_t\|^2 + 2 \eta_{t} (F^* - \EE F(\ov{w}_t)) + \eta_t^2 C\\
+&\leq\EE \|\ov{w}_{t}-\vw^{*}\|^{2}+\eta_{t}^{2} \EE\| \vg_{t} \|^{2} +  4L\eta_t^3(E-1)^2 G^2 + 2 \eta_{t} (F^* - \EE F(\ov{w}_t)) + \eta_t^2 C\\
+&\leq \EE\|\ov{w}_{t}-\vw^{*}\|^{2}+\eta_{t}^{2} G^2 +  4L\eta_t^3(E-1)^2 G^2 + 2 \eta_{t} (F^* - \EE F(\ov{w}_t))+\eta_t^2 C \label{eq:cvxsgd1}
+\end{align}
+
+Sum two sides of \eq{\ref{eq:cvxsgd1}}, we require $\eta_t < 1$ and set $F^*_t = \min_{t \in [0, T-1]} F(\ov{w}_t)$, 
+\begin{align*}
+	\sum_{t=0}^{T-1} 2\eta_t (\EE[F(\ov{w}_t)] - F^* ) & \leq \EE \|\ov{w}_{0}-\vw^{*}\|^{2} - \EE\left\|\ov{w}_{T}-\vw^{*}\right\|^2 + \sum_{t=0}^{T-1} (\eta_t^3 4LG^2(E-1)^2 + \eta_t^2 G^2+ \eta_t^2 C)\\
+    & \leq \EE \|\ov{w}_{0}-\vw^{*}\|^{2} - \EE\left\|\ov{w}_{T}-\vw^{*}\right\|^2 + \sum_{t=0}^{T-1} (\eta_t^3 4LG^2(E-1)^2 + \eta_t^2 G^2+ \eta_t^2 C)\\ 
+    & \leq \EE \|\ov{w}_{0}-\vw^{*}\|^{2} + \sum_{t=0}^{T-1}\eta_t^2\left[ G^2 ( 4L(E-1)^2 + 1) + C\right]\\
+    \sum_{t=0}^{T-1} 2\eta_t (\EE[F^*_t] - F^* ) & \leq \EE \|\ov{w}_{0}-\vw^{*}\|^{2} + \sum_{t=0}^{T-1}\eta_t^2\left[ G^2 ( 4L(E-1)^2 + 1) + C\right]\\
+\EE[ F_t^*] - F^*  & \leq \frac{\|\ov{w}_{0}-\vw^{*}\|^{2}}{2 \sum_{t=0}^{T-1} \eta_t } + \frac{\left[G^2( 4L(E-1)^2 + 1) + C\right] \sum_{t=0}^{T-1} \eta_t^2}{2 \sum_{t=0}^{T-1} \eta_t }
+\end{align*}
+It will converge under the following conditions: $ \lim_{T \rightarrow \infty }\sum_{t=0}^{T-1} \eta_t = \infty$, 
+$ \lim_{T \rightarrow \infty }\sum_{t=0}^{T-1} \eta_t^2 < \infty$. 
+\begin{align*}
+	\EE[ F_t^*] - F^*  & \leq \frac{\|\ov{w}_{0}-\vw^{*}\|^{2}}{2 \sum_{t=0}^{T-1} \eta_t } + \frac{\left[G^2( 4L(E-1)^2 + 1) + C\right] \sum_{t=0}^{T-1} \eta_t^2}{2 \sum_{t=0}^{T-1} \eta_t } \\
+	\EE[ F_t^*] - F^*  & \leq \frac{\Delta_0 + \left[G^2( 4L(E-1)^2 + 1) + C\right] \sum_{t=0}^{T-1} \eta_t^2 }{2 \sum_{t=0}^{T-1} \eta_t }\\
+	& \leq \frac{R [G^2( 4L(E-1)^2 + 1) + C] }{\sqrt{T}}
+\end{align*}
+where in the last line we set the learning rate satisfying $\eta_t = \sqrt{ \frac{\Delta_0}{ T [G^2( 4L(E-1)^2 + 1) + C] }}$, 
+and denote $R = \sqrt{ \frac{\Delta_0}{ G^2( 4L(E-1)^2 + 1) + C }}$ .
+\end{proof}

### 2020-04-01 source/sec_problem.tex
add partial activation to cvx smth.

@@
 
-\begin{itemize}
-	\item $\vw_t^k$: be the model parameter maintained in the k-th device at the t-th step. or the parameter obtained after communication steps. 
-	\item $\cI_E$: the set of global synchronization steps, $\cI_E = \{n E|n = 1, 2, 3,\dots \}$. If $t+1 \in \cI_E$, we send $w_{t+1}$ to all devices.
-	\item $\vv^k_{t+1}$: the immediate result of one step SGD update from $\vw^k_{t}$.
-	\item $\xi_{t}^{k}$: the data sampled from k-th device’s local data 
-	uniformly at random.
-	\item $\overline{\mathbf{v}}_{t}=\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$.
-	\item $\overline{\mathbf{w}}_{t}=\sum_{k=1}^{N} p_{k} \mathbf{w}_{t}^{k}$.
-	\item $\overline{\mathbf{v}}_{t+1}$ is one-step SGD from $\overline{\mathbf{w}}_{t}$. $t+1 \in \cI_E$, we can fetch $\overline{\vw}_{t+1}$, we can communicate $\overline{\vw}_{t+1}$ to all devices.
-	\item The expected one-step graident, $\overline{\mathbf{g}}_{t}=\sum_{k=1}^{N} p_{k} \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)$,  
-	\item One-step stochastic gradient $\mathbf{g}_{t}=\sum_{k=1}^{N} p_{k} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)$. $\EE[\vg_t] = \overline{g}_t$
-	\item
-
+Let $\vw_t^k$ be the model parameter maintained in the k-th device at the t-th step.
+$\cI_E$ is a set of global synchronization steps, i.e., $\cI_E = \{n E|n = 1, 2, 3,\dots \}$.
+If $t+1 \in \cI_E$, which means we communicate the server with (all) clients at time step $t+1$.
+$\vv^k_{t+1}$ is the immediate result of one step SGD update from $\vw^k_{t}$.
+$\xi_{t}^{k}$ denotes the data sampled from k-th device’s local data uniformly at random.
+Follow the common practice, we define two virtual sequences $\overline{\mathbf{v}}_{t}$ and $\overline{\mathbf{w}}_{t}$. For full device participation and $t \notin \cI_E$,
+$\ov{v}_t = \ov{w}_t =\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$.
+In partial participation, $t \in \cI_E$, $\ov{w}_t \neq \ov{v}_t$ since $\ov{v}_t=\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$ while $\sum_{k\in \cS_t}\mathbf{w}_{t}^{k}$.
+However, we can
+set unbiased sampling strategy such that $ \EE_{\cS_t} \ov{w}_t = \ov{v}_t$.
+$\overline{\mathbf{v}}_{t+1}$ is one-step SGD from $\overline{\mathbf{w}}_{t}$. 
 \begin{align}
@@
 \end{align}
+% $t+1 \in \cI_E$, we can fetch $\overline{\vw}_{t+1}$, we can communicate $\overline{\vw}_{t+1}$ to all devices.
+where $\vg_{t} = \sum_{k=1}^{N} p_{k} \vg_{t,k} $ is one-step stochastic gradient, averaged over all devices. 
+\begin{align}
+\vg_{t,k} \left\{\begin{array}{ll} 
+ = \nabla F_{k}\left(\mathbf{w}_{t}^{k},\xi_{t}^{k} \right)  &  \text{smooth}\\
+ \in \partial F_{k}\left(w_{t}^{k}, \xi_{t}^{k}\right)  & \text{non-smooth}
+ \end{array}\right.
+\end{align}
+Similarly, we denote the expected one-step gradient $\ov{g}_{t}= \EE_{\xi_t}[\vg_t] = \sum_{k=1}^{N} p_{k} \EE_{\xi_{t}^{k}} \vg_{t,k}$, where
+\begin{align}
+\EE_{\xi_{t}^{k}} \vg_{t,k}  \left\{\begin{array}{ll} 
+ = \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)  &  \text{smooth}\\
+ \in \partial F_{k}\left(w_{t}^{k}\right)  & \text{non-smooth}
+ \end{array}\right.
+\end{align}
+and we use $\xi_t = \{\xi_t^k\}_{k=1}^N$ denotes samples at all devices at time step $t$.
 
-\end{itemize}
+The updates of FedAve with partial device activation is given by: 
+\begin{align} 
+\mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\
+\sum_{k \in \cS_{t+1}} p_{k} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}\end{array}\right.
+\end{align}
 
+In~\cite{li2019convergence}, two types of unbiased sampling strategies are considered in Lemma 5. 
+The sampling scheme I establishes $\cS_{t+1}$ by i.i.d. sampling the devices with replacement,
+in this case the upper bound of expected square norm of $\ov{w}_{t+1} - \ov{v}_{t+1}$ is given by:
 \begin{align}
-\mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & (9) \\ \sum_{k=1}^{N} p_{k} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}\end{array}\right.
+\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4}{K} \eta_t^2 E^2G^2
 \end{align}
+The sampling scheme II establishes $\cS_{t+1}$ by uniformly sampling all devices without
+replacement, in which we have the 
+\begin{align}
+\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4(N - K)}{K(N-1)} \eta_t^2 E^2G^2
+\end{align}
+We denote this upper bound as follows for concise presentation. 
+\begin{align}
+	\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2 \leq  \eta_t^2 C
+	\label{eq:partialsample}
+\end{align}
+
+\lkxcom{quantify biased sampling scheme ?}

### 2020-04-02 source/sec_convexsmoothsgd.tex
update convex partial

@@
 $$\EE[ F_t^*] - F^*  \leq \frac{R [G^2( 4L(E-1)^2 + 1) + C] }{\sqrt{T}}$$
-where we denote $R = \sqrt{ \frac{\Delta_0}{ G^2( 4L(E-1)^2 + 1) + C }}$.
+where we denote $F^*_t = \min_{t \in [0, T-1]} F(\ov{w}_t)$ and $R = \sqrt{ \frac{\Delta_0}{ G^2( 4L(E-1)^2 + 1) + C }}$.
+\label{th:sgdcvxsmth}
 \end{theorem}

### 2020-04-02 source/sec_cvxnonsm.tex
add partial activation to cvx smth.
update convex partial

@@
 
-\textbf{Full Device Participation}
+% \textbf{Full Device Participation}
 
-\begin{itemize}
-	\item Stochastic Subgradient of device $k$ at time step $t$, at point $w_{t,k}$: 
-	$$\vg_{t,k} \coloneqq \vg_{t,k}(w_t^k)$$
-	$$ \vg_{t, k} \in \partial F_{k}\left(w_{t}^{k}, \xi_{t}^{k}\right) $$
-	$$\EE \vg_{t, k} \in \partial F_{k}\left(w_{t}^{k}\right)$$
-\item  One-step stochastic subgradient of all devices.
-$$\vg_{t}=\sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) $$
-\begin{align}
-	\EE \vg_{t}= \EE \sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) \coloneqq \sum_{k=1}^{N} p_{k} \EE \vg_{t, k}
-	\label{eq:egt}
-\end{align}
-\end{itemize}
+% \begin{itemize}
+% 	\item Stochastic Subgradient of device $k$ at time step $t$, at point $w_{t,k}$: 
+% 	$$\vg_{t,k} \coloneqq \vg_{t,k}(w_t^k)$$
+% 	$$ \vg_{t, k} \in \partial F_{k}\left(w_{t}^{k}, \xi_{t}^{k}\right) $$
+% 	$$\EE \vg_{t, k} \in \partial F_{k}\left(w_{t}^{k}\right)$$
+% \item  One-step stochastic subgradient of all devices.
+% $$\vg_{t}=\sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) $$
+% \begin{align}
+% 	\EE \vg_{t}= \EE \sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) \coloneqq \sum_{k=1}^{N} p_{k} \EE \vg_{t, k}
+% 	\label{eq:egt}
+% \end{align}
+% \end{itemize}
 
@@
 	\end{align}
-	where $$\eta_t = \frac{R}{\sqrt{T}},$$ 
-	$$F^*_t \coloneqq \min_{t \in [0, T-1]} F(\overline{\vw}_t),$$
-	$$R = \sqrt{ \frac{\Delta_0}{L}},$$ 
-	$$L=\sum_{k=1}^N \left( p_k G_k^2 \left(3 + 4(E-1)^2\right)\right).$$ 
+	where $\eta_t = \frac{R}{\sqrt{T}},$
+	$F^*_t \coloneqq \min_{t \in [0, T-1]} F(\overline{\vw}_t),$
+	$R = \sqrt{ \frac{\Delta_0}{L}},$
+	$L=\sum_{k=1}^N \left( G^2 \left(3 + 4(E-1)^2\right)\right).$
 	\label{th:cvxnonsmoth}
 \end{theorem}
-{\color{red}{This bound didn't consider heterogeneous of devices $\Gamma$ in~\cite{li2019convergence}}}
-
-In full device active setting, we have $\overline{\vw}_{t+1}=\overline{\vv}_{t+1}$,
-$$\Delta_{t+1} = \EE \|\overline{\vw}_{t+1} - \vw^*\|^2 = \EE \|\overline{\vv}_{t+1} - \vw^*\|^2,$$
-According to the definition of $\overline{\vv}_{t+1}$ in \eq{\ref{eq:vbar}}, we can expand $\Delta_{t+1}$
+\begin{proof}
+% In full device active setting, we have $\overline{\vw}_{t+1}=\overline{\vv}_{t+1}$,
+Follow the same reasoning in the proof of Theorem~\ref{th:sgdcvxsmth} \eq{\ref{eq:sgdcvxsmth1}}
+$$\EE_{\cS_{t+1}, \xi_{t}} \|\ov{w}_{t+1} - \vw^*\|^2 \leq  \eta_t^2 C + \EE_{\xi_t} \|\overline{\vv}_{t+1} - \vw^*\|^2 $$
+where $C$ is defined in \eq{\ref{eq:partialsample}}.
+For convenience, we denote $\Delta_{t+1} = \EE \|\overline{\vw}_{t+1} - \vw^*\|^2$.
+We now focus on bounding $\left\|\ov{v}_{t+1}-\vw^{*}\right\|^2$.
+According to the definition of $\overline{\vv}_{t+1}$ in \eq{\ref{eq:vbar}}, we have,
 $$\begin{aligned}\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2 &=\left\|\bar{\vw}_{t}-\eta_{t} \vg_{t}-\vw^{*}\right\|^2 \\ &=\left\|\bar{\vw}_{t}-\vw^{*}\right\|^{2}-2 \eta_{t} \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)+\eta_{t}^{2}\left\|\vg_{t}\right\|^{2} \end{aligned}$$
@@
 \begin{align}
-\Delta_{t+1} = \EE\left[\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2| \vw_{t}\right]=\|\bar{\vw}_{t}-\vw^{*}\|^{2}-2 \eta_{t} \EE \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)+\eta_{t}^{2} \EE\| \vg_{t} \|^{2}
+\EE\left[\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2| \vw_{t}\right]=\|\bar{\vw}_{t}-\vw^{*}\|^{2}-2 \eta_{t} \EE \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)+\eta_{t}^{2} \EE\| \vg_{t} \|^{2}	
 \label{eq:expand}
@@
 \begin{align}
-	\Delta_{t+1} \leq& \|\bar{\vw}_{t}-\vw^{*}\|^{2} + \eta_{t}^{2} \EE\| \vg_{t} \|^{2} \\
+	\Delta_{t+1} \leq& \|\bar{\vw}_{t}-\vw^{*}\|^{2} + \eta_{t}^{2} \EE\| \vg_{t} \|^{2}  + \eta_t^2 C \\
 	& + \underbrace{\sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_t^k\|^2}_{A_1} +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2
@@
 	\Delta_{t+1} & \leq \Delta_t + 
-	\eta_{t}^{2} \EE\| \vg_{t} \|^{2} + \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_t^k\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2 \\
-	 &- 2 \eta_t \left(-\frac{1}{2} \sum_{k=1}^{N} P_{k} (\eta_t \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \frac{1}{\eta_t} \|\vw_t^k - \overline{\vw}_t\|^2) + (F(\overline{w}_t) - F^*) \right)\\
+	\eta_{t}^{2} \EE\| \vg_{t} \|^{2} + \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_t^k\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2 + \eta_t^2 C \\
+	 &- 2 \eta_t \left(-\frac{1}{2} \sum_{k=1}^{N} P_{k} (\eta_t \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \frac{1}{\eta_t} \|\vw_t^k - \overline{\vw}_t\|^2) + (F(\overline{w}_t) - F^*) \right)  + \eta_t^2 C \\
 	& \leq \Delta_t + 
-	\eta_{t}^{2} \EE\| \vg_{t} \|^{2} + \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_t^k\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2 \\
-	& +\sum_{k=1}^{N} P_{k} (\eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \|\vw_t^k - \overline{\vw}_t\|^2) - 2\eta_t (F(\overline{w}_t) - F^*)\\
-	& = \Delta_t + \underbrace{\eta_{t}^{2} \EE\| \vg_{t} \|^{2} +\sum_{k=1}^N p_k  ( 2\|\overline{\vw}_t - \vw_t^k\|^2  +  \eta_t^2 \|\EE \vg_{t,k}\|^2 + \eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 )}_{C_1 = \eta_t^2L}  - 2\eta_t (F(\overline{w}_t) - F^*) \label{eq:main3}
+	\eta_{t}^{2} \EE\| \vg_{t} \|^{2} + \sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_t^k\|^2 +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2  + \eta_t^2 C  \\
+	& +\sum_{k=1}^{N} P_{k} (\eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \|\vw_t^k - \overline{\vw}_t\|^2) - 2\eta_t (F(\overline{w}_t) - F^*) \\
+	& = \Delta_t + \underbrace{\eta_{t}^{2} \EE\| \vg_{t} \|^{2} +\sum_{k=1}^N p_k  ( 2\|\overline{\vw}_t - \vw_t^k\|^2  +  \eta_t^2 \|\EE \vg_{t,k}\|^2 + \eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 ) + \eta_t^2 C }_{C_1 = \eta_t^2L}  - 2\eta_t (F(\overline{w}_t) - F^*) \label{eq:main3}
 \end{align}
@@
     F^*_t - F^* &\leq \frac{\Delta_0 + \sum_{t=0}^{T-1} \eta_t^2L}{ \sum_{t=0}^{T-1}  2 \eta_t}\\
-    F^*_t - F^* &\leq \frac{RL}{\sqrt{T}}\\
+    F^*_t - F^* &\leq \frac{RL}{\sqrt{T}}
 \end{align}
@@
 \begin{align}
-C_1 = \eta_t^2 L  = \eta_t^2 \sum_{k=1}^N p_k G_k^2 \left(3 + 4(E-1)^2\right)	
+C_1 = \eta_t^2 L  = \eta_t^2 [\sum_{k=1}^N p_k G_k^2 \left(3 + 4(E-1)^2\right) + C]  = \eta_t^2 [C+ G^2 \left(3 + 4(E-1)^2\right)]
 \end{align}
 
-\textbf{Problems}
-\begin{itemize}
-	\item The bound in Theorem~\ref{th:cvxnonsmoth} didn't incoporate heterogeneous of data.
-	\item \cite{huo2020faster} this paper has no assumption on the heterogeneous, the bounded gradient variance can cover this assumption. 
-\end{itemize}
+\end{proof}

### 2020-04-03 source/sec_convexsmoothsgd.tex
partial cvx smth nasgd

@@
 According to the definition of $\overline{\vv}_{t+1}$ in \eq{\ref{eq:vbar}}, we can expand the second term in \eq{\ref{eq:sgdcvxsmth1}} as follows:
-$$\begin{aligned}\left\|\ov{v}_{t+1}-\vw^{*}\right\|^2 &=\left\|\ov{w}_{t}-\eta_{t} \vg_{t}-\vw^{*}\right\|^2 \\ &=\left\|\ov{w}_{t}-\vw^{*}\right\|^{2}-2 \eta_{t} \vg_{t}^{\top}\left(\ov{w}_{t}-\vw^{*}\right)+\eta_{t}^{2}\left\|\vg_{t}\right\|^{2} \end{aligned}$$
+$$\begin{aligned}\left\|\ov{v}_{t+1}-\vw^{*}\right\|^2 &=\left\|\ov{w}_{t}-\eta_{t} \vg_{t}-\vw^{*}\right\|^2 \\ &=\left\|\ov{w}_{t}-\vw^{*}\right\|^{2}-2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}-\vw^{*} \right>+\eta_{t}^{2}\left\|\vg_{t}\right\|^{2} \end{aligned}$$
 
@@
 \begin{align}
-\EE\left[\left\|\ov{v}_{t+1}-\vw^{*}\right\|^2| \vw_{t}\right]=\|\ov{w}_{t}-\vw^{*}\|^{2}-2 \eta_{t} \EE \vg_{t}^{\top}\left(\ov{w}_{t}-\vw^{*}\right)+\eta_{t}^{2} \EE\| \vg_{t} \|^{2}	
+\EE\left[\left\|\ov{v}_{t+1}-\vw^{*}\right\|^2| \vw_{t}\right]=\|\ov{w}_{t}-\vw^{*}\|^{2}-2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}-\vw^{*} \right> +\eta_{t}^{2} \EE\| \vg_{t} \|^{2}	
 \label{eq:expandsgd}
@@
 
-Now we focus on bounding $-2 \eta_{t} \EE \vg_{t}^{\top}\left(\ov{w}_{t}-\vw^{*}\right)$ in \eq{\ref{eq:expandsgd}}:
+Now we focus on bounding $-2 \eta_{t} \left< \EE \vg_{t}, \ov{w}_{t}-\vw^{*}\right>$ in \eq{\ref{eq:expandsgd}}:
 

### 2020-04-03 source/sec_theorems.tex
partial cvx smth nasgd

% !TEX ROOT=./main.tex

\section{Convergence Results}

\begin{table}[h!]
\centering
\small
	\begin{tabular}{|c|c|c|c|c|}\hline
		paper         &  Cvx Non-smth & Cvx L-smth & Strongly Cvx Non-smth& Strongly Cvx L-smth \\ \hline
	Accelerated	SGD   &   $\cO(\frac{1}{\sqrt{T}})$     & $\cO(\frac{1}{\sqrt{T}})$   &    $\cO(\frac{1}{T})$   &  $\cO(\frac{1}{T})$    \\\hline
	SGD    &    $\cO(\frac{1}{\sqrt{T}})$  &   $\cO(\frac{1}{\sqrt{T}})$  &  $\cO(\frac{1}{T})$  & $\cO(\frac{1}{T})$\cite{li2019convergence,haddadpour2019convergence}      \\\hline
	\end{tabular}
	\caption{Summarize of related work on federated learning, heterogeneous data.}
\end{table}

\lkxcom{Problems}
\begin{itemize}
	\item The bound in SGD didn't interpret heterogeneous of data.
	\item \cite{huo2020faster} this paper has no assumption on the heterogeneous, the bounded gradient variance can cover this assumption. 
	\item quantify biased sampling scheme?
\end{itemize}

### 2020-04-03 source/sec_problem.tex
update convex partial
reorg

@@
 \mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\
-\sum_{k \in \cS_{t+1}} p_{k} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}\end{array}\right.
+\sum_{k \in \cS_{t+1}} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}\end{array}\right.
 \end{align}
@@
 
-\lkxcom{quantify biased sampling scheme ?}
+\lkxcom{Questions and next steps}
+\begin{itemize}
+	\item Consider heterogeneity of data: with strongly convex and smooth, we can use optimality gap. with non-smooth, we can use weighted gradient diversity. 
+	\item Nonconvex imply strongly convex: KDD~\cite{zhuo2019federated} has constant learning rate, in our case, we have a diminishing learning rate. 
+	\item Quantify biased sampling scheme?
+	\item Constant learning rate in other non-strongly convex case. 
+\end{itemize}

### 2020-04-03 source/sec_theorems.tex
reorg

@@
 \section{Convergence Results}
-
+This section summarizes the current results and assumptions.
 \begin{table}[h!]
@@
 
-\lkxcom{Problems}
-\begin{itemize}
-	\item The bound in SGD didn't interpret heterogeneous of data.
-	\item \cite{huo2020faster} this paper has no assumption on the heterogeneous, the bounded gradient variance can cover this assumption. 
-	\item quantify biased sampling scheme?
-\end{itemize}
+\begin{assumption}
+The expected squared norm of stochastic subgradients is uniformly bounded. i.e.,
+$\mathbb{E}\|\vg_{t,k}\|^2  \leq G_k^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.  This also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$.
+\label{ass:subgrad2}
+\end{assumption}
+
+\begin{assumption}[Expected square norm of gradient]
+The expected squared norm of stochastic gradients is uniformly bounded. i.e.,
+$\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)\right\|^{2} \leq G^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
+\label{ass:squaregrad}
+\end{assumption}
+
+\begin{assumption}[L-smooth]
+$F_{1}, \cdots, F_{N}$ are all $L$-smooth: for all  $\mathbf{v}$  and $\mathbf{w}$, $F_{k}(\mathbf{v}) \leq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{L}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$.
+\label{ass:lsmooth}
+\end{assumption}
+\begin{assumption}[Strongly-convex]
+$	F_{1}, \cdots, F_{N} \text { are all } \mu \text { -strongly convex: for all v and } \mathbf{w}, F_{k}(\mathbf{v}) \geq F_{k}(\mathbf{w})+(\mathbf{v}-\mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{\mu}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$
+\label{ass:stroncvx}
+\end{assumption}
+
+\subsection{Stochastic gradient descent}
+\subsubsection{Convex}
+\begin{theorem}
+Let Assumption~\ref{ass:subgrad2} and Assumption~\ref{ass:lsmooth} hold, suppose we have a bound 
+on our starting distance, i.e., $\|\vw_{0} - \vw^*\|^2 \leq \Delta_0$, set learning rate $\eta_t = \sqrt{ \frac{\Delta_0}{ T [G^2( 4L(E-1)^2 + 1) + C] }}$, we have,
+$$\EE[ F_t^*] - F^*  \leq \frac{R [G^2( 4L(E-1)^2 + 1) + C] }{\sqrt{T}}$$
+where we denote $F^*_t = \min_{t \in [0, T-1]} F(\ov{w}_t)$ and $R = \sqrt{ \frac{\Delta_0}{ G^2( 4L(E-1)^2 + 1) + C }}$.
+\label{th:sgdcvxsmth}
+\end{theorem}
+
+\subsubsection{Strongly Convex}
+\cite{li2019convergence}
+
+\subsection{Stochastic Subgradient methods}
+
+\subsubsection{Convex}
+\begin{theorem}
+	Let Assumption~\ref{ass:subgrad2} hold, 
+	then the FedAve with full device participation satisfies
+	\begin{align}
+		 \EE[F^*_t] - F^* &\leq \frac{RL}{\sqrt{T}}
+	\end{align}
+	where $\eta_t = \frac{R}{\sqrt{T}},$
+	$F^*_t \coloneqq \min_{t \in [0, T-1]} F(\overline{\vw}_t),$
+	$R = \sqrt{ \frac{\Delta_0}{L}},$
+	$L=\sum_{k=1}^N \left( G^2 \left(3 + 4(E-1)^2\right)\right).$
+	\label{th:cvxnonsmoth}
+\end{theorem}
+
+\subsubsection{Strongly Convex}
+\begin{theorem}
+	Let assumption~\ref{ass:stroncvx} and assumption~\ref{ass:subgrad2} hold. Then,
+	\begin{align}
+		\EE[F(\hat{\vw}_T)] - F^* \leq \frac{2(B + C)}{\mu(T+1)}.
+	\end{align}
+	where $B =  G^2 (3  + 8 (E-1)^2)$, for sampling scheme I $C =\frac{4(N - K)}{K(N-1)} \eta_t^2 E^2G^2 $ or
+sampling scheme II $C = \frac{4}{K} \eta_t^2 E^2G^2$, for full device participation $C= 0$.
+\end{theorem}
+
+\section{Accelerated methods}
+\subsection{stochastic gradient descent}
+\subsubsection{Convex}
+\begin{theorem}
+	Let Assumption~\ref{ass:lsmooth} and Assumption~\ref{ass:squaregrad} hold,  choose the learning rate $\eta = \frac{\alpha}{1 - \beta} = \sqrt{\frac{\Delta_0}{(D+C)T}}$, $\beta \in (0, 1)$ and $\beta \leq \min\{1, \frac{1}{1 + \sqrt{\frac{\Delta_0}{(D+C)T}}}\}$, then the FedNestrovAve with partial device participation satisfies
+	\begin{align}
+		 \EE F(\hat{\vw}_T) - F^* &\leq \sqrt{\frac{\Delta_0(D + C)}{T}} + \frac{\Delta_0^{3/2}}{2\sqrt{D+C}}\frac{1}{T^{3/2}} 
+	\end{align}
+	where $\hat{\vw}_T = \frac{1}{T}\sum_{t=0}^{T-1} \ov{w}_t$, $C$ is defined in \eq{\ref{eq:partialsample}},
+$D$ is given by $D =  G^2[6 + (4(E-1)^2+1)L]$.
+\end{theorem}
+
+\subsubsection{Strongly Convex}
+\begin{theorem}
+	(Partial device participation) Suppose $F_{k}$ is $L$-smooth and
+	$\mu$-strongly convex for all $k$. Let $\kappa=\frac{L}{\mu}$,
+	$\gamma=\max\{8\kappa-1,E\}$ where $E$ is the communication interval,
+	and learning rates 
+	\begin{align*}
+	\alpha_{t} & =\frac{c}{\mu(\gamma+t)}\\
+	\beta_{t} & \leq\alpha_{t}
+	\end{align*}
+	so that $\alpha_{t}\leq2\alpha_{t+E}$, and where $0\leq c\leq1$
+	is small enough such that the following hold: 
+	\begin{align*}
+	\alpha_{t}^{2}+\beta_{t-1}^{2} & \leq\frac{1}{2}\\
+	\alpha_{t} & \leq\frac{1}{4L}\\
+	4\alpha_{t-1}^{2} & \leq\alpha_{t}
+	\end{align*}
+	for all $t\geq0$. Suppose also that $G$ is a constant satisfying
+	$\mathbb{E}\|w_{0}-\alpha_{0}g_{0,k}\|^{2}=\mathbb{E}\|w_{0}-\alpha_{0}\nabla F_{k}(w_{0},\xi_{0}^{k})\|^{2}\leq G^{2}$
+	for all $k$, and $\mathbb{E}\|g_{t,k}\|^{2}\leq G^{2}$ for all $t,k$.
+
+	Then with the partial device participation scheme described above,
+	\begin{align*}
+	\mathbb{E}F(w_{T})-F^{\ast} & \leq\frac{\kappa}{\gamma+T}(\frac{B'+C}{\mu}+4L(\|w_{0}-w^{\ast}\|^{2})\\
+	B' & =8L\Gamma+32(E-1)^{2}G^{2}+3G^{2}+2K^{2}\\
+	C & =\frac{16}{S}E^{2}G^{2}
+	\end{align*}
+	and $K$ is such that 
+	\begin{align*}
+	\alpha_{0}B+2K\cdot G & \leq\mu K^{2}\\
+	B & =8L\Gamma+32(E-1)^{2}G^{2}+3G^{2}
+	\end{align*}
+	and
+	\begin{align*}
+	K & \geq\max\{\|w_{0}-w^{\ast}\|^{2},\frac{G}{2\alpha_{0}}\}
+	\end{align*}
+\end{theorem}
+
+\subsection{Stochastic Subgradient methods}
+
+\subsubsection{Convex}
+\begin{theorem}
+	Let Assumption~\ref{ass:subgrad2} hold, choose the learning rate $\eta = \frac{\alpha}{1 - \beta} = \sqrt{\frac{\Delta_0}{(D+C)T}}$, $\beta \in (0, 1)$ and $\beta \leq \min\{1, \frac{1}{1 + \sqrt{\frac{\Delta_0}{(D+C)T}}}\}$, then the FedNestrovAve with partial device participation satisfies
+	\begin{align}
+		 \EE F(\hat{\vw}_T) - F^* &\leq \sqrt{\frac{\Delta_0(D + C)}{T}} + \frac{\Delta_0^{3/2}}{2\sqrt{D+C}}\frac{1}{T^{3/2}} 
+	\end{align}
+	where $\hat{\vw}_T = \frac{1}{T}\sum_{t=0}^{T-1} \ov{w}_t$, $C$ is defined in \eq{\ref{eq:partialsample}},
+$D$ is given by $D = (8 + 8(E-1)^2)G^2$.
+	\label{th:nasgcvxnonsmoth}
+\end{theorem}
+
+\subsubsection{Strongly Convex}
+% \input{sec_nasgd_scvx_nonsmooth}
+\begin{thm}
+	(Partial device participation) Let the parameters satisfy the assumptions
+	in the ICLR paper, $\kappa=\frac{L}{\mu}$, $\gamma=\max\{8\kappa,E\}$
+	and learning rate $\alpha_{t}=\frac{c}{\mu(\gamma+t)}$, $\beta_{t}\leq\alpha_{t}$
+	where $c$ is small enough such that $\alpha_{t}^{2}+\beta_{t-1}^{2}\leq\frac{1}{2}$
+	and $2\alpha_{t-1}^{2}\leq\alpha_{t}$ for all $t$. Then with the
+	partial device participation scheme described above,
+	\begin{align*}
+	\mathbb{E}F(w_{T})-F^{\ast} & \leq\frac{2\kappa}{\gamma+T}(\frac{B'+C}{\mu}+2L(\|w_{0}-w^{\ast}\|^{2})\\
+	B' & =\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+9L\Gamma+32(E-1)^{2}G^{2}+2+2G^{2}+2GK\\
+	C & =\frac{16}{S}E^{2}G^{2}
+	\end{align*}
+	and $K$ is such that 
+	\begin{align*}
+	\alpha_{0}B+2\sqrt{K}\cdot G & \leq\mu K\\
+	B & =\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+9L\Gamma+32(E-1)^{2}G^{2}
+	\end{align*}
+	and
+	\begin{align*}
+	\|w_{0}-w^{\ast}\|^{2} & \leq K
+	\end{align*}
+\end{thm}

### 2020-04-03 source/sec_theorems.tex
reorg
updated theorem summary

@@
 \begin{theorem}
-	Let assumption~\ref{ass:stroncvx} and assumption~\ref{ass:subgrad2} hold. Then,
+	Let assumption~\ref{ass:stroncvx} and assumption~\ref{ass:subgrad2} hold, choose learning rate $\eta_t = \frac{2}{\mu(t+1)} $.Then,
 	\begin{align}
@@
 % \input{sec_nasgd_scvx_nonsmooth}
-\begin{thm}
-	(Partial device participation) Let the parameters satisfy the assumptions
-	in the ICLR paper, $\kappa=\frac{L}{\mu}$, $\gamma=\max\{8\kappa,E\}$
-	and learning rate $\alpha_{t}=\frac{c}{\mu(\gamma+t)}$, $\beta_{t}\leq\alpha_{t}$
-	where $c$ is small enough such that $\alpha_{t}^{2}+\beta_{t-1}^{2}\leq\frac{1}{2}$
-	and $2\alpha_{t-1}^{2}\leq\alpha_{t}$ for all $t$. Then with the
-	partial device participation scheme described above,
+\begin{theorem}
+	(Partial device participation) Suppose $F_{k}$ is $\mu$-strongly
+	convex for all $k$. Let $E$ be the communication interval, and learning
+	rates 
 	\begin{align*}
-	\mathbb{E}F(w_{T})-F^{\ast} & \leq\frac{2\kappa}{\gamma+T}(\frac{B'+C}{\mu}+2L(\|w_{0}-w^{\ast}\|^{2})\\
-	B' & =\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+9L\Gamma+32(E-1)^{2}G^{2}+2+2G^{2}+2GK\\
+	\alpha_{t} & =\frac{c}{\mu(E+t)}\\
+	\beta_{t} & \leq\alpha_{t}
+	\end{align*}
+	so that $\alpha_{t}\leq2\alpha_{t+E}$, and where $c$ is small enough
+	such that the following hold: 
+	\begin{align*}
+	\alpha_{t}^{2}+\beta_{t-1}^{2} & \leq\frac{1}{2}\\
+	4\alpha_{t-1}^{2} & \leq\alpha_{t}
+	\end{align*}
+	for all $t\geq0$. Suppose also that $G$ is a constant satisfying
+	$\mathbb{E}\|w_{0}-\alpha_{0}g_{0,k}\|^{2}=\mathbb{E}\|w_{0}-\alpha_{0}\nabla F_{k}(w_{0},\xi_{0}^{k})\|^{2}\leq G^{2}$
+	for all $k$, and $\mathbb{E}\|\nabla F_{k}(w,\xi_{t}^{k})\|^{2}\leq G^{2}$
+	for $w=\overline{w}_{t}$ or $w=w_{t}^{k}$ and all $t,k$.
+
+	Then with partial device participation, \textbf{
+		\begin{align*}
+		F(\sum_{t=1}^{T}\frac{2t}{T(T+1)}\overline{w}_{t})-F^{\ast} & \leq\frac{2(B'+C)}{\mu(T+1)}
+		\end{align*}
+	} where
+	\begin{align*}
+	B' & =6G^{2}+32(E-1)^{2}G^{2}+2K^{2}\\
 	C & =\frac{16}{S}E^{2}G^{2}
@@
 	\begin{align*}
-	\alpha_{0}B+2\sqrt{K}\cdot G & \leq\mu K\\
-	B & =\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+9L\Gamma+32(E-1)^{2}G^{2}
+	\alpha_{0}B+2K\cdot G & \leq\mu K^{2}\\
+	B & =6G^{2}+32(E-1)^{2}G^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\|w_{0}-w^{\ast}\|^{2} & \leq K
+	K & \geq\max\{\|w_{0}-w^{\ast}\|^{2},\frac{G}{2\alpha_{0}}\}
 	\end{align*}
-\end{thm}
+\end{theorem}

### 2020-04-03 source/sec_theorems.tex
Update sec_theorems.tex

--- 2020-04-03 source/sec_theorems.tex
@@
 
-\begin{assumption}
-The expected squared norm of stochastic subgradients is uniformly bounded. i.e.,
-$\mathbb{E}\|\vg_{t,k}\|^2  \leq G_k^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.  This also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$.
-\label{ass:subgrad2}
-\end{assumption}
+% \begin{assumption}
+% The expected squared norm of stochastic subgradients is uniformly bounded. i.e.,
+% $\mathbb{E}\|\vg_{t,k}\|^2  \leq G_k^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
+This also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$.
+Denote $G^2 = \sum_{k=1}^N p_k G_k^2$
+% \label{ass:subgrad2}
+% \end{assumption}
 
@@
 $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)\right\|^{2} \leq G^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
-\label{ass:squaregrad}
+\label{ass:subgrad2}
 \end{assumption}
@@
 Let Assumption~\ref{ass:subgrad2} and Assumption~\ref{ass:lsmooth} hold, suppose we have a bound 
-on our starting distance, i.e., $\|\vw_{0} - \vw^*\|^2 \leq \Delta_0$, set learning rate $\eta_t = \sqrt{ \frac{\Delta_0}{ T [G^2( 4L(E-1)^2 + 1) + C] }}$, we have,
+on our starting distance, i.e., $\|\vw_{0} - \vw^*\|^2 \leq \Delta_0$, set learning rate $\eta_t = \left(\frac{\Delta_0}{ T [G^2( 4L(E-1)^2 + 1) + C] }\right)^{1/2}$, we have,
 $$\EE[ F_t^*] - F^*  \leq \frac{R [G^2( 4L(E-1)^2 + 1) + C] }{\sqrt{T}}$$
-where we denote $F^*_t = \min_{t \in [0, T-1]} F(\ov{w}_t)$ and $R = \sqrt{ \frac{\Delta_0}{ G^2( 4L(E-1)^2 + 1) + C }}$.
+where we denote $F^*_t = \min_{t \in [0, T-1]} F(\ov{w}_t)$ and $R = \left(\frac{\Delta_0}{ G^2( 4L(E-1)^2 + 1) + C }\right)$.
 \label{th:sgdcvxsmth}
@@
 \begin{theorem}
-	Let Assumption~\ref{ass:lsmooth} and Assumption~\ref{ass:squaregrad} hold,  choose the learning rate $\eta = \frac{\alpha}{1 - \beta} = \sqrt{\frac{\Delta_0}{(D+C)T}}$, $\beta \in (0, 1)$ and $\beta \leq \min\{1, \frac{1}{1 + \sqrt{\frac{\Delta_0}{(D+C)T}}}\}$, then the FedNestrovAve with partial device participation satisfies
+	Let Assumption~\ref{ass:lsmooth} and Assumption~\ref{ass:subgrad2} hold,  choose the learning rate $\eta = \frac{\alpha}{1 - \beta} = \sqrt{\frac{\Delta_0}{(D+C)T}}$, $\beta \in (0, 1)$ and $\beta \leq \min\{1, \frac{1}{1 + \sqrt{\frac{\Delta_0}{(D+C)T}}}\}$, then the FedNestrovAve with partial device participation satisfies
 	\begin{align}

### 2020-04-04 source/sec_convexsmoothsgd.tex
refine proof v1.

@@
 Let Assumption~\ref{ass:subgrad2} and Assumption~\ref{ass:lsmooth} hold, suppose we have a bound 
-on our starting distance, i.e., $\|\vw_{0} - \vw^*\|^2 \leq \Delta_0$, set learning rate $\eta_t = \sqrt{ \frac{\Delta_0}{ T [G^2( 4L(E-1)^2 + 1) + C] }}$, we have,
-$$\EE[ F_t^*] - F^*  \leq \frac{R [G^2( 4L(E-1)^2 + 1) + C] }{\sqrt{T}}$$
-where we denote $F^*_t = \min_{t \in [0, T-1]} F(\ov{w}_t)$ and $R = \sqrt{ \frac{\Delta_0}{ G^2( 4L(E-1)^2 + 1) + C }}$.
-\label{th:sgdcvxsmth}
+on our starting distance, i.e., $\|\vw_{0} - \vw^*\|^2 \leq \Delta_0$, set learning rate $\eta_t =  \left(\frac{\Delta_0}{ T [G^2( 4L(E-1)^2 + 1) + C]}\right)^{1/2}$, we have,
+$$\EE[ F_t^*] - F^*  \leq \left(\frac{ \Delta_0 [G^2( 4L(E-1)^2 + 1) + C] }{T}\right)^{1/2}$$
+where we denote $F^*_t = \min_{t \in [0, T-1]} F(\ov{w}_t)$.
+% \label{th:sgdcvxsmth}
 \end{theorem}
@@
 	\EE[ F_t^*] - F^*  & \leq \frac{\Delta_0 + \left[G^2( 4L(E-1)^2 + 1) + C\right] \sum_{t=0}^{T-1} \eta_t^2 }{2 \sum_{t=0}^{T-1} \eta_t }\\
-	& \leq \frac{R [G^2( 4L(E-1)^2 + 1) + C] }{\sqrt{T}}
+	& \leq \left(\frac{ \Delta_0 [G^2( 4L(E-1)^2 + 1) + C] }{T}\right)^{1/2}
 \end{align*}
-where in the last line we set the learning rate satisfying $\eta_t = \sqrt{ \frac{\Delta_0}{ T [G^2( 4L(E-1)^2 + 1) + C] }}$, 
-and denote $R = \sqrt{ \frac{\Delta_0}{ G^2( 4L(E-1)^2 + 1) + C }}$ .
+where in the last line we set the learning rate satisfying $\eta_t =  \left(\frac{\Delta_0}{ T [G^2( 4L(E-1)^2 + 1) + C]}\right)^{1/2}$.
 \end{proof}

### 2020-04-04 source/sec_cvxnonsm.tex
reorg
refine proof v1.

@@
 
-\begin{assumption}
-The expected squared norm of stochastic subgradients is uniformly bounded. i.e.,
-$\mathbb{E}\|\vg_{t,k}\|^2  \leq G_k^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
-This also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$.
-Denote $G^2 = \sum_{k=1}^N p_k G_k^2$
-\label{ass:subgrad2}
-\end{assumption}
-
 \begin{theorem}
 	Let Assumption~\ref{ass:subgrad2} hold, 
-	then the FedAve with full device participation satisfies
+	then the FedAve with partial device participation satisfies
 	\begin{align}
-		 F^*_t - F^* &\leq \frac{RL}{\sqrt{T}}
+		 \EE[F^*_t] - F^* &\leq  \left(\frac{\Delta_0[C+ G^2 \left(3 + 4(E-1)^2\right)]}{T}\right)^{1/2}
 	\end{align}
-	where $\eta_t = \frac{R}{\sqrt{T}},$
-	$F^*_t \coloneqq \min_{t \in [0, T-1]} F(\overline{\vw}_t),$
-	$R = \sqrt{ \frac{\Delta_0}{L}},$
-	$L=\sum_{k=1}^N \left( G^2 \left(3 + 4(E-1)^2\right)\right).$
-	\label{th:cvxnonsmoth}
+	where $\eta_t = \eta_t  = \sqrt{ \frac{\Delta_0}{T[C+ G^2 \left(3 + 4(E-1)^2\right)]}}$,
+	$F^*_t \coloneqq \min_{t \in [0, T-1]} F(\overline{\vw}_t)$.
+	% $R = \sqrt{ \frac{\Delta_0}{L}},$
+	% $L=\sum_{k=1}^N \left( G^2 \left(3 + 4(E-1)^2\right)\right).$
+	% \label{th:cvxnonsmoth}
 \end{theorem}
@@
 	& +\sum_{k=1}^{N} P_{k} (\eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \|\vw_t^k - \overline{\vw}_t\|^2) - 2\eta_t (F(\overline{w}_t) - F^*) \\
-	& = \Delta_t + \underbrace{\eta_{t}^{2} \EE\| \vg_{t} \|^{2} +\sum_{k=1}^N p_k  ( 2\|\overline{\vw}_t - \vw_t^k\|^2  +  \eta_t^2 \|\EE \vg_{t,k}\|^2 + \eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 ) + \eta_t^2 C }_{C_1 = \eta_t^2L}  - 2\eta_t (F(\overline{w}_t) - F^*) \label{eq:main3}
+	& = \Delta_t + \underbrace{\eta_{t}^{2} \EE\| \vg_{t} \|^{2} +\sum_{k=1}^N p_k  ( 2\|\overline{\vw}_t - \vw_t^k\|^2  +  \eta_t^2 \|\EE \vg_{t,k}\|^2 + \eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 ) + \eta_t^2 C }_{\eta_t^2C_1}  - 2\eta_t (F(\overline{w}_t) - F^*) \label{eq:main3}
 \end{align}
@@
 \begin{align}
-	2\eta_t (F(\overline{w}_t) - F^*) \leq \Delta_t - \Delta_{t+1} + \eta_t^2 L
+	2\eta_t (\EE[F(\overline{w}_t)] - F^*) \leq \Delta_t - \Delta_{t+1} + \eta_t^2 C_1 
 \end{align}
@@
 \begin{align}
-	\sum_{t=0}^{T-1} 2\eta_t(F(\overline{w}_t) - F^*) & \leq \Delta_0 - \Delta_T + \sum_{t=0}^{T-1} \eta_t^2 L\\
-	\sum_{t=0}^{T-1} 2\eta_t(F(\overline{w}_t) - F^*) & \leq \Delta_0 + \sum_{t=0}^{T-1} \eta_t^2 L\\
-    F^*_t - F^* &\leq \frac{\Delta_0 + \sum_{t=0}^{T-1} \eta_t^2L}{ \sum_{t=0}^{T-1}  2 \eta_t}\\
-    F^*_t - F^* &\leq \frac{RL}{\sqrt{T}}
+	\sum_{t=0}^{T-1} 2\eta_t(\EE[F(\overline{w}_t)] - F^*) & \leq \Delta_0 - \Delta_T + \sum_{t=0}^{T-1} \eta_t^2 C_1 \\
+	\sum_{t=0}^{T-1} 2\eta_t(\EE[F(\overline{w}_t)]- F^*) & \leq \Delta_0 + \sum_{t=0}^{T-1} \eta_t^2 C_1 \\
+    \EE[F^*_t] - F^* &\leq \frac{\Delta_0 + \sum_{t=0}^{T-1} \eta_t^2C_1 }{ \sum_{t=0}^{T-1}  2 \eta_t}\\
+    \EE[F^*_t] - F^* &\leq  \left(\frac{\Delta_0[C+ G^2 \left(3 + 4(E-1)^2\right)]}{T}\right)^{1/2}
 \end{align}
-where in last line we let $R = \sqrt{ \frac{\Delta_0}{L}}, \eta_t = \frac{R}{\sqrt{T}}$
+where in last line we let $\eta_t  = \sqrt{ \frac{\Delta_0}{T[C+ G^2 \left(3 + 4(E-1)^2\right)]}}$.
 
@@
 \begin{align}
-C_1 = \eta_t^2 L  = \eta_t^2 [\sum_{k=1}^N p_k G_k^2 \left(3 + 4(E-1)^2\right) + C]  = \eta_t^2 [C+ G^2 \left(3 + 4(E-1)^2\right)]
+\eta_t^2 C_1  = \eta_t^2 [\sum_{k=1}^N p_k G_k^2 \left(3 + 4(E-1)^2\right) + C]  = \eta_t^2 [C+ G^2 \left(3 + 4(E-1)^2\right)]
 \end{align}

### 2020-04-04 source/sec_theorems.tex
refine proof v1.

@@
 Let Assumption~\ref{ass:subgrad2} and Assumption~\ref{ass:lsmooth} hold, suppose we have a bound 
-on our starting distance, i.e., $\|\vw_{0} - \vw^*\|^2 \leq \Delta_0$, set learning rate $\eta_t = \left(\frac{\Delta_0}{ T [G^2( 4L(E-1)^2 + 1) + C] }\right)^{1/2}$, we have,
-$$\EE[ F_t^*] - F^*  \leq \frac{R [G^2( 4L(E-1)^2 + 1) + C] }{\sqrt{T}}$$
-where we denote $F^*_t = \min_{t \in [0, T-1]} F(\ov{w}_t)$ and $R = \left(\frac{\Delta_0}{ G^2( 4L(E-1)^2 + 1) + C }\right)$.
+on our starting distance, i.e., $\|\vw_{0} - \vw^*\|^2 \leq \Delta_0$, set learning rate $\eta_t =  \left(\frac{\Delta_0}{ T [G^2( 4L(E-1)^2 + 1) + C]}\right)^{1/2}$, we have,
+$$\EE[ F_t^*] - F^*  \leq \left(\frac{ \Delta_0 [G^2( 4L(E-1)^2 + 1) + C] }{T}\right)^{1/2}$$
+where we denote $F^*_t = \min_{t \in [0, T-1]} F(\ov{w}_t)$.
 \label{th:sgdcvxsmth}
 \end{theorem}
+The proof is deferred to Section~\ref{sec:convexsmoothsgd} in the Appendix.
 
@@
 \subsubsection{Convex}
+
 \begin{theorem}
-	Let Assumption~\ref{ass:subgrad2} hold, 
-	then the FedAve with full device participation satisfies
+	Let Assumption~\ref{ass:subgrad2} hold, choose learning rate $\eta_t  = (\frac{\Delta_0}{T[C+ G^2 \left(3 + 4(E-1)^2\right)]})^{1/2}$,
+	then the FedAve with partial device participation satisfies
 	\begin{align}
-		 \EE[F^*_t] - F^* &\leq \frac{RL}{\sqrt{T}}
+		 \EE[F^*_t] - F^* &\leq  \left(\frac{\Delta_0[C+ G^2 \left(3 + 4(E-1)^2\right)]}{T}\right)^{1/2}
 	\end{align}
-	where $\eta_t = \frac{R}{\sqrt{T}},$
-	$F^*_t \coloneqq \min_{t \in [0, T-1]} F(\overline{\vw}_t),$
-	$R = \sqrt{ \frac{\Delta_0}{L}},$
-	$L=\sum_{k=1}^N \left( G^2 \left(3 + 4(E-1)^2\right)\right).$
-	\label{th:cvxnonsmoth}
+	where $F^*_t \coloneqq \min_{t \in [0, T-1]} F(\overline{\vw}_t)$.
+	% $R = \sqrt{ \frac{\Delta_0}{L}},$
+	% $L=\sum_{k=1}^N \left( G^2 \left(3 + 4(E-1)^2\right)\right).$
+	% \label{th:cvxnonsmoth}
 \end{theorem}
+The proof is deferred to Section~\ref{sec:sgdcvxnonsmth} in the Appendix.
 

### 2020-04-18 source/sec_cvxnonsm.tex
convex smooth nonsmoth reorg

@@
 % In full device active setting, we have $\overline{\vw}_{t+1}=\overline{\vv}_{t+1}$,
-Follow the same reasoning in the proof of Theorem~\ref{th:sgdcvxsmth} \eq{\ref{eq:sgdcvxsmth1}}
-$$\EE_{\cS_{t+1}, \xi_{t}} \|\ov{w}_{t+1} - \vw^*\|^2 \leq  \eta_t^2 C + \EE_{\xi_t} \|\overline{\vv}_{t+1} - \vw^*\|^2 $$
-where $C$ is defined in \eq{\ref{eq:partialsample}}.
-For convenience, we denote $\Delta_{t+1} = \EE \|\overline{\vw}_{t+1} - \vw^*\|^2$.
-We now focus on bounding $\left\|\ov{v}_{t+1}-\vw^{*}\right\|^2$.
-According to the definition of $\overline{\vv}_{t+1}$ in \eq{\ref{eq:vbar}}, we have,
-$$\begin{aligned}\left\|\bar{\vv}_{t+1}-\vw^{*}\right\|^2 &=\left\|\bar{\vw}_{t}-\eta_{t} \vg_{t}-\vw^{*}\right\|^2 \\ &=\left\|\bar{\vw}_{t}-\vw^{*}\right\|^{2}-2 \eta_{t} \vg_{t}^{\top}\left(\bar{\vw}_{t}-\vw^{*}\right)+\eta_{t}^{2}\left\|\vg_{t}\right\|^{2} \end{aligned}$$
-
-Take the expectation condition on $\vw_t$ over random samples at all devices:
+Follow the same reasoning in the proof of Theorem~\ref{th:sgdcvxsmth} \eq{\ref{eq:expandsgd}}, we have
 \begin{align}
@@
 \end{align}
+Apply Lemma~\ref{lma:gradient} and Lemma~\ref{lma:l3iclr} in the last line, we can bound $C_1$ as follows:
+\begin{align}
+\eta_t^2 C_1  = \eta_t^2 [\sum_{k=1}^N p_k G_k^2 \left(3 + 4(E-1)^2\right) + C]  = \eta_t^2 [C+ G^2 \left(3 + 4(E-1)^2\right)]
+\end{align}
 Rearrange the last inequality:
@@
 
-Now we discuss how to bound the term $C_1$ in \eq{\ref{eq:main3}}
+% Now we discuss how to bound the term $C_1$ in \eq{\ref{eq:main3}}
+% Assume Assumption~\ref{ass:subgrad2}, $\eta_t$ is non-increasing, and $\eta_t \leq 2\eta_{t+E}$ for all $t\geq 0$, according to Lemma 3 in~\cite{li2019convergence}, we have
 
-Assume Assumption~\ref{ass:subgrad2}, $\eta_t$ is non-increasing, and $\eta_t \leq 2\eta_t+E$ for all $t\geq 0$, according to Lemma 3 in~\cite{li2019convergence}, we have
+% \begin{equation}
+% \mathbb{E}\left[\sum_{k=1}^{N} p_{k}\left\|\overline{\mathbf{w}}_{t}-\mathbf{w}_{k}^{t}\right\|^{2}\right] \leq \sum_{k=1}^N p_k4 \eta_{t}^{2}(E-1)^{2} G_k^{2}
+% 	\label{eq:g4}
+% \end{equation}
 
-\begin{equation}
-\mathbb{E}\left[\sum_{k=1}^{N} p_{k}\left\|\overline{\mathbf{w}}_{t}-\mathbf{w}_{k}^{t}\right\|^{2}\right] \leq \sum_{k=1}^N p_k4 \eta_{t}^{2}(E-1)^{2} G_k^{2}
-	\label{eq:g4}
-\end{equation}
+% Assume Assumption~\ref{ass:subgrad2}, we have 
+% \begin{equation}
+% 	\eta_t^2 \|\EE \vg_{t,k}\|^2 \leq  \eta_t^2 G_k^2
+% 	\label{eq:g3}
+% \end{equation}
+% and
+% \begin{equation}
+% 	\eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t)\|^2 \leq  \eta_t^2 G_k^2
+% 	\label{eq:g2}
+% \end{equation}
 
-Assume Assumption~\ref{ass:subgrad2}, we have 
-\begin{equation}
-	\eta_t^2 \|\EE \vg_{t,k}\|^2 \leq  \eta_t^2 G_k^2
-	\label{eq:g3}
-\end{equation}
-and
-\begin{equation}
-	\eta_t^2 \|\EE \vg_{t,k}(\overline{\vw}_t)\|^2 \leq  \eta_t^2 G_k^2
-	\label{eq:g2}
-\end{equation}
+% Now the only term left in $C_1$ is $\eta_{t}^{2} \EE\| \vg_{t} \|^{2} $, follow the proof in Lemma 2~\cite{li2019convergence}: 
+% \begin{align}
+% 	  \eta_{t}^{2} \EE\| \vg_{t} \|^{2} 
+% 	\leq   \eta_{t}^{2} \EE\| \sum_{k=1}^N p_k g_{t,k} \|^{2} 
+% 	\leq   \eta_{t}^{2} \sum_{k=1}^N p_k \EE\| g_{t,k} \|^{2} 
+% 	\leq   \eta_{t}^{2} \sum_{k=1}^N p_k G_k^2  \label{eq:g1}
+% \end{align}
+% where in the third inequality we use the convexity of the l2 norm.
 
-Now the only term left in $C_1$ is $\eta_{t}^{2} \EE\| \vg_{t} \|^{2} $, follow the proof in Lemma 2~\cite{li2019convergence}: 
-\begin{align}
-	  \eta_{t}^{2} \EE\| \vg_{t} \|^{2} 
-	\leq   \eta_{t}^{2} \EE\| \sum_{k=1}^N p_k g_{t,k} \|^{2} 
-	\leq   \eta_{t}^{2} \sum_{k=1}^N p_k \EE\| g_{t,k} \|^{2} 
-	\leq   \eta_{t}^{2} \sum_{k=1}^N p_k G_k^2  \label{eq:g1}
-\end{align}
-where in the third inequality we use the convexity of the l2 norm.
-
-Combine \eq{\ref{eq:g4}},\eq{\ref{eq:g3}},\eq{\ref{eq:g2}}, \eq{\ref{eq:g1}},
-we have 
-\begin{align}
-\eta_t^2 C_1  = \eta_t^2 [\sum_{k=1}^N p_k G_k^2 \left(3 + 4(E-1)^2\right) + C]  = \eta_t^2 [C+ G^2 \left(3 + 4(E-1)^2\right)]
-\end{align}
+% Combine \eq{\ref{eq:g4}},\eq{\ref{eq:g3}},\eq{\ref{eq:g2}}, \eq{\ref{eq:g1}},
+% we have 
+% \begin{align}
+% \eta_t^2 C_1  = \eta_t^2 [\sum_{k=1}^N p_k G_k^2 \left(3 + 4(E-1)^2\right) + C]  = \eta_t^2 [C+ G^2 \left(3 + 4(E-1)^2\right)]
+% \end{align}
 

### 2020-04-18 source/sec_theorems.tex
overview re arrange.

--- 2020-04-03 source/sec_theorems.tex
@@
 
+\begin{assumption}[Bounded local variance]
+Let $\xi_{t}^{k}$ be sampled from the $k$ -th device's local data uniformly at random.
+The variance of stochastic gradients in each device is bounded: $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq \sigma_{k}^{2}$,
+for $k=1, \cdots, N$.	
+\end{assumption}
+
 % \begin{assumption}
@@
 $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)\right\|^{2} \leq G^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
+This also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$.
 \label{ass:subgrad2}
@@
 
-\subsection{Stochastic gradient descent}
-\subsubsection{Convex}
+\begin{definition}[Squared Weighted Gradient Diversity]
+	We indicate the following quantity as weighted gradient diversity among the local 
+	objectives:
+	\begin{align*}
+		\frac{\sum_{k=1}^K p_k  \mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}, \xi^{k}\right)\right\|^{2} }{\mathbb{E}\left\| \sum_{k=1}^K p_k   \nabla F_{k}\left(\mathbf{w}, \xi^{k}\right)\right\|^{2} } \leq \lambda
+	\end{align*}
+\end{definition}
+This definition is similar to the weighted gradient diversity defined in~\cite{haddadpour2019convergence}.
+
+\begin{definition}[Gradient Dissimilarity]
+	We indicate the following quantity as weighted gradient diversity among the local 
+	objectives:
+	\begin{align*}
+		\frac{\sum_{k=1}^K p_k \left\|\nabla F_{k}\left(\mathbf{w}\right)\right\|^{2} }{\left\|   \nabla F\left(\mathbf{w}\right)\right\|^{2} } \leq \lambda 
+	\end{align*}
+\end{definition}
+
+\section{The Convergence of FedAve for General Convex Functions}
+This section summarizes the convergence of FedAve.
+The updates of FedAve with partial device activation is given by: 
+\begin{align} 
+\mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\
+\sum_{k \in \cS_{t+1}} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}\end{array}\right.
+\end{align}
+
+\subsection{The Convergence of FedAve for Convex Functions}
+In this subsection, we discuss the convergence of FedAve for convex functions,
+we consider both smooth (Theorem~\ref{th:sgdcvxsmth}) and non-smooth (Theorem~\ref{th:sgdcvxnonsmoth})
+functions.
+
 \begin{theorem}
 Let Assumption~\ref{ass:subgrad2} and Assumption~\ref{ass:lsmooth} hold, suppose we have a bound 
-on our starting distance, i.e., $\|\vw_{0} - \vw^*\|^2 \leq \Delta_0$, set learning rate $\eta_t = \left(\frac{\Delta_0}{ T [G^2( 4L(E-1)^2 + 1) + C] }\right)^{1/2}$, we have,
-$$\EE[ F_t^*] - F^*  \leq \frac{R [G^2( 4L(E-1)^2 + 1) + C] }{\sqrt{T}}$$
-where we denote $F^*_t = \min_{t \in [0, T-1]} F(\ov{w}_t)$ and $R = \left(\frac{\Delta_0}{ G^2( 4L(E-1)^2 + 1) + C }\right)$.
+on our starting distance, i.e., $\|\vw_{0} - \vw^*\|^2 \leq \Delta_0$, set learning rate $\eta_t =  \left(\frac{\Delta_0}{ T A}\right)^{1/2}$, we have,
+$$\EE[ F_T^*] - F^*  \leq \left(\frac{ \Delta_0 A }{T}\right)^{1/2},$$
+where we denote $F^*_T = \min_{t \in [0, T-1]} F(\ov{w}_t)$, $A = G^2( 4L(E-1)^2 + 1)) + C$.
 \label{th:sgdcvxsmth}
@@
 
-\subsubsection{Strongly Convex}
+\begin{theorem}
+If we remove the smoothness assumption in Theorem~\ref{th:sgdcvxsmth}, i.e., 
+	Let Assumption~\ref{ass:subgrad2} hold, replacing the constant $A$ with $A=G^2 \left(4(E-1)^2 + 3\right) + C$, then the same convergence rate in Theorem~\ref{th:sgdcvxsmth}.
+	\label{th:sgdcvxnonsmoth}
+\end{theorem}
+The proof is deferred to Section~\ref{sec:convexsmoothsgd} in the Appendix.
+
+\subsection{The Convergence of FedAve for Strongly Convex Functions}
+
+\subsection{Smooth}
 \cite{li2019convergence}
 
-\subsection{Stochastic Subgradient methods}
-
-\subsubsection{Convex}
-\begin{theorem}
-	Let Assumption~\ref{ass:subgrad2} hold, 
-	then the FedAve with full device participation satisfies
-	\begin{align}
-		 \EE[F^*_t] - F^* &\leq \frac{RL}{\sqrt{T}}
-	\end{align}
-	where $\eta_t = \frac{R}{\sqrt{T}},$
-	$F^*_t \coloneqq \min_{t \in [0, T-1]} F(\overline{\vw}_t),$
-	$R = \sqrt{ \frac{\Delta_0}{L}},$
-	$L=\sum_{k=1}^N \left( G^2 \left(3 + 4(E-1)^2\right)\right).$
-	\label{th:cvxnonsmoth}
-\end{theorem}
-
-\subsubsection{Strongly Convex}
+\subsection{Non-smooth}
 \begin{theorem}

### 2020-04-18 source/sec_theorems.tex
linear speedup for scvx smooth
updated result table 1

@@
 	Accelerated	SGD   &   $\cO(\frac{1}{\sqrt{T}})$     & $\cO(\frac{1}{\sqrt{T}})$   &    $\cO(\frac{1}{T})$   &  $\cO(\frac{1}{T})$    \\\hline
-	SGD    &    $\cO(\frac{1}{\sqrt{T}})$  &   $\cO(\frac{1}{\sqrt{T}})$  &  $\cO(\frac{1}{T})$  & $\cO(\frac{1}{T})$\cite{li2019convergence,haddadpour2019convergence}      \\\hline
+	SGD    &    $\cO(\frac{1}{\sqrt{NT}})$  &   $\cO(\frac{1}{\sqrt{T}})$  &  $\cO(\frac{1}{T})$  & $\cO(\frac{1}{T})$\cite{li2019convergence,haddadpour2019convergence}, $\cO(\frac{1}{NT})$ ours    \\\hline
 	\end{tabular}
@@
 \subsection{Smooth}
-\cite{li2019convergence}
+We improve on the result of \cite{li2019convergence} and show the linear speedup of FedAvg with the number of devices.
+
+\begin{theorem}
+	(Full device participation) Suppose $F_{k}$ is $L$-smooth and $\mu$-strongly
+	convex for all $k$. Let $\kappa=\frac{L}{\mu}$, $\gamma=\max\{8\kappa-1,E\}$
+	where $E$ is the communication interval, and learning rates 
+	\begin{align*}
+	\alpha_{t} & =\frac{c}{\mu(\gamma+t)}
+	\end{align*}
+	so that $\alpha_{t}\leq2\alpha_{t+E}$, and where $0\leq c\leq1$
+	is small enough such that the following hold: 
+	\begin{align*}
+	\alpha_{t} & \leq\frac{1}{4L}\\
+	\alpha_{t} & \leq\frac{1}{N}
+	\end{align*}
+	for all $t\geq0$.
+
+	Then with full device participation,
+	\begin{align*}
+	\mathbb{E}F(w_{T})-F^{\ast} & \leq C\frac{\kappa}{N(\gamma+t)}\\
+	C & =4E^{2}LG^{2}+\nu_{max}^{2}\sigma^{2}+LG^{2}
+	\end{align*}
+
+\end{theorem}
 

### 2020-04-18 source/sec_theorems.tex
gradi dis

--- 2020-04-18 source/sec_theorems.tex
@@
 
-\begin{definition}[Gradient Dissimilarity]
-	We indicate the following quantity as weighted gradient diversity among the local 
+\begin{definition}[Gradient Dissimilarity~\cite{li2018federated}]
+	We indicate the following quantity as gradient diversity among the local 
 	objectives:
@@
 	\end{align*}
+	For some $\epsilon>0$, for all the points $\vw \in \mathcal{S}_{\epsilon}^{c}=\left\{\vw |\|\nabla F(\vw)\|^{2}>\epsilon\right\}$
 \end{definition}

### 2020-04-18 source/sec_sgd_scvx_smooth.tex

We show that FedAvg with SGD has $O(1/NT)$ rate under $\mu$-strong
convexity and $L$-smoothness. The result improves on the ICLR paper's
$O(1/T)$ with a linear speedup in the number of devices $N$. {[}TODO:
Add more comment here.{]} The FedAv algorithm with SGD follows the
updates
\begin{align*}
y_{t+1}^{k} & =w_{t}^{k}-\alpha_{t}g_{t,k}\\
w_{t+1}^{k} & =\begin{cases}
y_{t+1}^{k} & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}y_{t+1}^{k} & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}
\end{align*}
 where 
\begin{align*}
g_{t,k} & :=\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})
\end{align*}
 is the stochastic gradient.

Define the virtual sequences $\overline{y}_{t}=\sum_{k=1}^{N}p_{k}y_{t}^{k}$,
$\overline{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$, and $\overline{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$.
We have $\mathbb{E}g_{t}=\overline{g}_{t}$ and $\overline{y}_{t+1}=\overline{w}_{t}-\alpha_{t}g_{t}$,
and $\overline{w}_{t+1}=\overline{y}_{t+1}$ for all $t$.

%
\begin{assumption}
The stochastic gradients have bounded variance 
\begin{align*}
\mathbb{E}\|\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})-\nabla F_{k}(w_{t}^{k})\|^{2} & \leq\sigma_{k}^{2}
\end{align*}
for all $k$. Moreover, the expected squared norm of stochastic gradients
is uniformly bounded, 
\begin{align*}
\mathbb{E}\|\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})\|^{2} & \leq G^{2}
\end{align*}
\end{assumption}
\begin{theorem}
(Full device participation) Suppose $F_{k}$ is $L$-smooth and $\mu$-strongly
convex for all $k$. Let $\kappa=\frac{L}{\mu}$, $\gamma=\max\{8\kappa-1,E\}$
where $E$ is the communication interval, and learning rates 
\begin{align*}
\alpha_{t} & =\frac{c}{\mu(\gamma+t)}
\end{align*}
so that $\alpha_{t}\leq2\alpha_{t+E}$, and where $0\leq c\leq1$
is small enough such that the following hold: 
\begin{align*}
\alpha_{t} & \leq\frac{1}{4L}\\
\alpha_{t} & \leq\frac{1}{N}
\end{align*}
 for all $t\geq0$.

Then with full device participation,
\begin{align*}
\mathbb{E}F(w_{T})-F^{\ast} & \leq C\frac{\kappa}{N(\gamma+t)}\\
C & =4E^{2}LG^{2}+\nu_{max}^{2}\sigma^{2}+LG^{2}
\end{align*}

\end{theorem}
\begin{proof}
The idea of the proof is to observe that the $L$-smoothness of $F$
provides the upper bound
\begin{align*}
\mathbb{E}(F(\overline{w}_{t}))-F^{\ast} & =\mathbb{E}(F(\overline{w}_{t})-F(w^{\ast}))\\
 & \leq\frac{L}{2}\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}
\end{align*}
 and to show that $\mathbb{E}\|\overline{w}_{T}-w^{\ast}\|^{2}=O(\frac{1}{NT})$.

Our main step is to prove the bound 
\begin{align*}
\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}^{2}B
\end{align*}
 where $B$ is define in the statement of the theorem.

We have 
\begin{align*}
\|\overline{w}_{t+1}-w^{\ast}\|^{2} & =\|(\overline{w}_{t}-\alpha_{t}g_{t})-w^{\ast}\|^{2}\\
 & =\|(\overline{w}_{t}-\alpha_{t}\overline{g}_{t}-w^{\ast})-\alpha_{t}(g_{t}-\overline{g}_{t})\|^{2}\\
 & =A_{1}+A_{2}+A_{3}
\end{align*}
where 
\begin{align*}
A_{1} & =\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2}\\
A_{2} & =2\alpha_{t}\langle\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t},\overline{g}_{t}-g_{t}\rangle\\
A_{3} & =\alpha_{t}^{2}\|g_{t}-\overline{g}_{t}\|^{2}
\end{align*}
 $\mathbb{E}A_{2}=0$ by definition of $g_{t}$ and $\overline{g}_{t}$,
while for $A_{3}$ we have
\begin{align*}
\alpha_{t}^{2}\mathbb{E}\|g_{t}-\overline{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|g_{t}-\mathbb{E}g_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
\end{align*}
 again by Jensen's inequality and using the independence of $g_{t,k},g_{t,k'}$.

Next we bound $A_{1}$: 
\begin{align*}
\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & =\|\overline{w}_{t}-w^{\ast}\|^{2}+2\langle\overline{w}_{t}-w^{\ast},-\alpha_{t}\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2}
\end{align*}
and we will show that the third term $\|\alpha_{t}\overline{g}_{t}\|^{2}$
can be canceled by an upper bound of the second term. %

Now 
\begin{align*}
-2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle & =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
 & =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
 & \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(w^{\ast})-F_{k}(w_{t}^{k}))\\
 & \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w_{t}^{k})-F_{k}(\overline{w}_{t})+\frac{L}{2}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+F_{k}(w^{\ast})-F_{k}(w_{t}^{k})\right]\\
 & =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]
\end{align*}
For the second term, which is negative, we can ignore it, but this
yields a suboptimal bound that fails to provide the desired linear
speedup. This is the flaw in the analysis of the ICLR paper. Instead,
we upper bound it using the following derivation: 
\begin{align*}
2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right] & \leq2\alpha_{t}\left[F(\overline{w}_{t+1})-F(\overline{w}_{t})\right]\\
 & \leq2\alpha_{t}\mathbb{E}\langle\nabla F(\overline{w}_{t}),\overline{w}_{t+1}-\overline{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\overline{w}_{t+1}-\overline{w}_{t}\|^{2}\\
 & =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\overline{w}_{t}),g_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
 & =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\overline{w}_{t}),\overline{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
 & =-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\|\nabla F(\overline{w}_{t})-\overline{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
 & =-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\|\nabla F(\overline{w}_{t})-\sum_{k}p_{k}\nabla F(w_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
 & \leq-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\overline{w}_{t})-\nabla F(w_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
 & \leq-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
 & \leq-\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}
\end{align*}
where we have used the smoothness of $F$ twice.

Note that the term $-\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}$ exactly
cancels the $\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}$ in the bound
for $A_{1}$, so that plugging in the bound for $-2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle$,
we have so far proved 
\begin{align*}
\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
 & +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}
\end{align*}
 The term $\mathbb{E}\|g_{t}\|^{2}\leq G^{2}$ by assumption.

Now we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}$.
Since communication is done every $E$ steps, for any $t\geq0$, we
can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $w_{t_{0}}^{k}=\overline{w}_{t_{0}}$for
all $k$. Moreover, using $\eta_{t}$ is non-increasing and $\alpha_{t_{0}}\leq\alpha{}_{t}$
for any $t-t_{0}\leq E-1$, we have 
\begin{align*}
\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2} & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}-(\overline{w}_{t}-\overline{w}_{t_{0}})\|^{2}\\
 & \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}\|^{2}\\
 & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-w_{t_{0}}^{k}\|^{2}\\
 & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\
 & \leq\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}\|g_{i,k}\|^{2}\\
 & \leq\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{t_{0}}^{2}G^{2}\\
 & \leq4E^{2}\alpha_{t}^{2}G^{2}
\end{align*}

Using the bound on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}$,
we can conclude that, with $\nu_{max}:=N\cdot\max_{k}p_{k}$ and $\nu_{min}:=N\cdot\min_{k}p_{k}$,

\begin{align*}
\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
 & =\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\sum_{k=1}^{N}(p_{k}N)^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
 & \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\nu_{max}^{2}\sum_{k=1}^{N}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
 & \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\alpha_{t}^{3}LG^{2}
\end{align*}
 where $\sigma^{2}=\max_{k}\sigma_{k}^{2}$. The leading term in the
bound is $\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}$, which
one summed using the recursive relations gives $O(\alpha_{t})$, but
the constant now is scaled down by $1/N$. To achieve this linear
speedup, we need $\sum_{t=0}^{T}(1-\mu\alpha_{t})\|w_{0}-w^{\ast}\|^{2}$
to be small enough.

To complete the proof of convergence, for $\alpha_{t}=\frac{\delta}{t+\gamma}$
for $\delta>\frac{1}{\mu}$ and $\gamma>0$ such that $\alpha_{1}\leq\min\{\frac{1}{\mu},\frac{1}{4L}\}$and
$\alpha_{t}\leq2\alpha_{t+E}$, we can show that 
\begin{align*}
\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2} & \leq\frac{v}{N(\gamma+t)}
\end{align*}
 where $v=\max\{\frac{\delta^{2}B'}{\delta\mu-c+1},(\gamma+1)\|w_{0}-w^{\ast}\|^{2}\}$
by induction. The definition of $v$ ensures that 
\begin{align*}
\|w_{0}-w^{\ast}\|^{2} & \leq\frac{v}{\gamma}
\end{align*}
 and suppose $\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}\leq\frac{v}{N(\gamma+t)}$
holds for some $t\geq0$. Then 
\begin{align*}
\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2}\leq & (1-\alpha_{t}\mu)\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}^{2}B'\\
 & \leq(1-\frac{\delta\mu}{t+\gamma})\frac{v}{t+\gamma}+\frac{\delta^{2}B'}{(t+\gamma)^{2}}\\
 & =\frac{t+\gamma-1}{(t+\gamma)^{2}}v+\left[\frac{\delta^{2}B'}{(t+\gamma)^{2}}-\frac{\delta\mu-1}{(t+\gamma)^{2}}v\right]\\
 & \leq\frac{v}{t+\gamma+1}
\end{align*}
 by definition of $v$.

Finally, the $L$-smoothness of $F$ implies 
\begin{align*}
\mathbb{E}(F(\overline{w}_{t}))-F^{\ast} & =\mathbb{E}(F(\overline{w}_{t})-F(w^{\ast}))\\
 & \leq\frac{L}{2}\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}\leq\frac{L}{2}\frac{v}{\gamma+t}
\end{align*}
 Now if we choose $\delta=\frac{c}{\mu}$ where $c\leq1$ is small
enough such that the required conditions 
\begin{align*}
\alpha_{t}^{2}+\beta_{t-1}^{2} & \leq\frac{1}{2}\\
\alpha_{t} & \leq\frac{1}{4L}\\
4\alpha_{t-1}^{2} & \leq\alpha_{t}
\end{align*}
hold for all $t\geq0$, $\gamma=\max\{8\kappa-1,E\}$, then $\alpha_{t}=\frac{c}{\mu}\frac{1}{\gamma+t}$
and recalling $v=\max\{\frac{\delta^{2}B'}{\delta\mu-c+1},(\gamma+1)\|w_{0}-w^{\ast}\|^{2}\}$,
we finally have 
\begin{align*}
\mathbb{E}(F(\overline{w}_{t}))-F^{\ast} & \leq\frac{L}{2}\frac{v}{\gamma+t}\leq\frac{L/2}{\gamma+t}(\frac{\delta^{2}B'}{\delta\mu-c+1}+(\gamma+1)\|w_{0}-w^{\ast}\|^{2})\\
 & \leq\frac{\kappa}{\gamma+t}(\frac{B'}{\mu}+4L\|w_{0}-w^{\ast}\|^{2})
\end{align*}
\end{proof}

### 2020-04-18 source/sec_theorems.tex
updated table 1

--- 2020-04-18 source/sec_theorems.tex
@@
 	Accelerated	SGD   &   $\cO(\frac{1}{\sqrt{T}})$     & $\cO(\frac{1}{\sqrt{T}})$   &    $\cO(\frac{1}{T})$   &  $\cO(\frac{1}{T})$    \\\hline
-	SGD    &    $\cO(\frac{1}{\sqrt{NT}})$  &   $\cO(\frac{1}{\sqrt{T}})$  &  $\cO(\frac{1}{T})$  & $\cO(\frac{1}{T})$\cite{li2019convergence,haddadpour2019convergence}, $\cO(\frac{1}{NT})$ ours    \\\hline
+	SGD    &    $\cO(\frac{1}{\sqrt{T}})$  &   $\cO(\frac{1}{\sqrt{NT}})$  &  $\cO(\frac{1}{T})$  & $\cO(\frac{1}{T})$\cite{li2019convergence,haddadpour2019convergence}, $\cO(\frac{1}{NT})$ ours    \\\hline
 	\end{tabular}

### 2020-04-19 source/sec_sgd_scvx_smooth.tex

@@
 
-%
+We first summarize some properties of $L$-smooth and $\mu$-strongly
+convex functions.
+
 \begin{assumption}
@@
 \begin{theorem}
-(Full device participation) Suppose $F_{k}$ is $L$-smooth and $\mu$-strongly
-convex for all $k$. Let $\kappa=\frac{L}{\mu}$, $\gamma=\max\{8\kappa-1,E\}$
-where $E$ is the communication interval, and learning rates 
+	Suppose $F_{k}$ is $L$-smooth and $\mu$-strongly convex for all
+	$k$. Let $\kappa=\frac{L}{\mu}$, $\gamma=\max\{8\kappa-1,E\}$ where
+	$E$ is the communication interval, and learning rates 
 	\begin{align*}
@@
 
-Then with full device participation,
+	Then with full device participation, as long as $E\leq O(\sqrt{\frac{N}{T}})$,
 	\begin{align*}
 	\mathbb{E}F(w_{T})-F^{\ast} & \leq C\frac{\kappa}{N(\gamma+t)}\\
-C & =4E^{2}LG^{2}+\nu_{max}^{2}\sigma^{2}+LG^{2}
+	C & =5LG^{2}+\nu_{max}^{2}\sigma^{2}
 	\end{align*}
-
+	and with partial participation, with $E=O(1)$ and $K$ sampled devices
+	at each communication round, 
+	\begin{align*}
+	\mathbb{E}F(w_{T})-F^{\ast} & \leq(4E^{2}G^{2}+\nu_{max}^{2}\sigma^{2})\frac{\kappa}{K(\gamma+t)}+\frac{4E^{2}LG^{2}}{t^{2}}
+	\end{align*}
 \end{theorem}
@@
 	speedup, we need $\sum_{t=0}^{T}(1-\mu\alpha_{t})\|w_{0}-w^{\ast}\|^{2}$
-to be small enough.
+	to be small enough. \textbf{Note also that the above bound implies
+		that $E$ can be chosen to be as large as $O(\sqrt{\frac{T}{N}})$
+		without degrading the performance, since $E^{2}\alpha_{t}^{3}=O(\alpha_{t}^{2}/N)$
+		which is the order of the leading term. As we will see, in the partial
+		participation case, this is no longer true, and $E$ cannot be chosen
+		$O(T^{\beta})$ for any $\beta>0$. }
 
@@
 	\end{align*}
+
+	With partial participation, the update at each communication round
+	is now given by averages over a subset of sampled devices. The key
+	is to bound 
+	\begin{align*}
+	\mathbb{E}\|\overline{w}_{t}-\overline{y}_{t}\|^{2} & =\frac{1}{K}\sum_{k}p^{k}\|w_{t}^{k}-\sum_{k}p_{k}w_{t}^{k}\|^{2}
+	\end{align*}
+	This can be bounded by the same argument as before and yield 
+	\begin{align*}
+	\mathbb{E}\|\overline{w}_{t}-\overline{y}_{t}\|^{2} & \leq\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
+	\end{align*}
+	and so this is also a leading term, and together with the term $\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}$
+	determines the convergence rate, which is now $O(\frac{(\nu_{max}^{2}\sigma^{2}+4E^{2}G^{2})}{KT})$.
+	This implies that $E$ canot be chosen $O(T^{\beta})$ for any $\beta>0$
+	without degrading the performance. This should be checked in experiments,
+	whether with partial participation if the communication round is set
+	to scale with $T$, the convergence deteriorates. This is in constrast
+	with the full participation case, where $E=O(\sqrt{\frac{T}{N}})$
+	is allowed. \textbf{If we can confirm that full participation allows
+		linear speedup with $\nu=N\cdot\max_{k}p_{k}\approx1$ and $E=O(\sqrt{\frac{T}{N}})$,
+		whereas partial participation only allows $E=O(1)$, then this would
+		an interesting phenomenon that is not reported by previous studies!}
 \end{proof}

### 2020-04-20 source/sec_convexsmoothsgd.tex
change cvx smooth assumption to variance

@@
 One-step stochastic gradient of all devices:
-$\vg_{t}=\sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) $, $\EE \vg_{t}= \EE \sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) \coloneqq \sum_{k=1}^{N} p_{k} \EE \vg_{t, k}$
+$\vg_{t}=\sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) $, $\EE \vg_{t}= \EE \sum_{k=1}^{N} p_{k} \vg_{t, k}\left(w_{t}^{k}\right) \coloneqq \sum_{k=1}^{N} p_{k} \EE \vg_{t, k} = \ov{g}_t$.
+
+\begin{lemma}
+Under Assumption~\ref{ass:subgrad2}, we have the following bound
+\begin{align}
+	 \|\EE \vg_{t,k}\|^2 & \leq G_k^2	\label{eq:g3} \\
+	\|\EE \vg_{t,k}(\overline{\vw}_t)\|^2 & \leq  G_k^2 \label{eq:g2} \\
+   \EE\| \vg_{t} \|^{2} &  \leq  \sum_{k=1}^N p_k G_k^2  \label{eq:g1}
+\end{align}
+\label{lma:gradient}
+\end{lemma}
+The first two inequalities directly follows from Assumption~\ref{ass:subgrad2} and the inequality
+can be obtained by applying the convexity of the l2 norm and Jensen's inequality.
+
+\begin{lemma}
+	With unbiased device sampling scheme, we have the following bound: 
+	$$ \EE_{\cS_{t+1}, \xi_{t}} \|\overline{\vw}_{t+1} - \vw^*\|^2 \leq \eta_t^2 C + \EE_{\xi_t} \|\overline{\vv}_{t+1} - \vw^*\|^2 $$
+\label{lma:wdistance}
+\end{lemma}
+\begin{proof}
+\begin{align*}
+& \EE_{\cS_{t+1}, \xi_{t}} \|\overline{\vw}_{t+1} - \vw^*\|^2 \\
+=& \EE_{\cS_{t+1}, \xi_{t}} \|\overline{\vw}_{t+1} - \overline{\vv}_{t+1} + \overline{\vv}_{t+1} - \vw^*\|^2\\
+=& \EE_{\cS_{t+1}, \xi_{t}} \left[\|\overline{\vw}_{t+1} - \overline{\vv}_{t+1}\|^2 + \|\overline{\vv}_{t+1} - \vw^*\|^2\right] + 2\EE_{\xi_{t}} \left<\EE_{\cS_{t+1}} \overline{\vw}_{t+1} - \overline{\vv}_{t+1},   \overline{\vv}_{t+1} - \vw^*\right> \\
+=& \EE_{\cS_{t+1}, \xi_{t}} \|\overline{\vw}_{t+1} - \overline{\vv}_{t+1}\|^2 + \EE_{\xi_t} \|\overline{\vv}_{t+1} - \vw^*\|^2 \\
+\leq&  \eta_t^2 C + \EE_{\xi_t} \|\overline{\vv}_{t+1} - \vw^*\|^2 %\label{eq:sgdcvxsmth1}
+\end{align*}
+where in the second line we use $\EE_{\cS_{t+1}} \overline{\vw}_{t+1}  = \overline{\vv}_{t+1}$. 
+The first term can be bounded by consider different sampling scheme as in \eq{\ref{eq:partialsample}}.
+\end{proof}
+
+\begin{lemma}[Lemma 3~\cite{li2019convergence}]
+Let $\eta_t$ and $E$ satisfy $\eta_0 \leq 2 \eta_t$ for all $t- t_0 \leq E - 1$, we have
+$$\mathbb{E}\left[\sum_{k=1}^{N} p_{k}\left\|\overline{\mathbf{w}}_{t}-\mathbf{w}_{k}^{t}\right\|^{2}\right] \leq \sum_{k=1}^N p_k4 \eta_{t}^{2}(E-1)^{2} G_k^{2}$$
+\label{lma:l3iclr}
+\end{lemma}
 
@@
 \end{theorem}
-
 \begin{proof}
-
+With Lemma~\ref{lma:wdistance}, 
 \begin{align}
-\EE_{\cS_{t+1}, \xi_{t}} \|\overline{\vw}_{t+1} - \vw^*\|^2 &= \EE_{\cS_{t+1}, \xi_{t}} \|\overline{\vw}_{t+1} - \overline{\vv}_{t+1} + \overline{\vv}_{t+1} - \vw^*\|^2\\
-&= \EE_{\cS_{t+1}, \xi_{t}} \left[\|\overline{\vw}_{t+1} - \overline{\vv}_{t+1}\|^2 + \|\overline{\vv}_{t+1} - \vw^*\|^2\right] \\
-& \hspace{3em}+ 2\EE_{\xi_{t}} \left<\EE_{\cS_{t+1}} \overline{\vw}_{t+1} - \overline{\vv}_{t+1},   \overline{\vv}_{t+1} - \vw^*\right> \\
-& = \EE_{\cS_{t+1}, \xi_{t}} \|\overline{\vw}_{t+1} - \overline{\vv}_{t+1}\|^2 + \EE_{\xi_t} \|\overline{\vv}_{t+1} - \vw^*\|^2 \\
-& \leq  \eta_t^2 C + \EE_{\xi_t} \|\overline{\vv}_{t+1} - \vw^*\|^2 \label{eq:sgdcvxsmth1}
+	\EE_{\cS_{t+1}, \xi_{t}} \|\overline{\vw}_{t+1} - \vw^*\|^2 \leq \eta_t^2 C + \EE_{\xi_t} \|\overline{\vv}_{t+1} - \vw^*\|^2\label{eq:sgdcvxsmth1}
 \end{align}
-where in the second line we use $\EE_{\cS_{t+1}} \overline{\vw}_{t+1}  = \overline{\vv}_{t+1}$. 
-The first term can be bounded by consider different sampling scheme. 
-According to the definition of $\overline{\vv}_{t+1}$ in \eq{\ref{eq:vbar}}, we can expand the second term in \eq{\ref{eq:sgdcvxsmth1}} as follows:
-$$\begin{aligned}\left\|\ov{v}_{t+1}-\vw^{*}\right\|^2 &=\left\|\ov{w}_{t}-\eta_{t} \vg_{t}-\vw^{*}\right\|^2 \\ &=\left\|\ov{w}_{t}-\vw^{*}\right\|^{2}-2 \eta_{t} \left<\EE \vg_{t}, \ov{w}_{t}-\vw^{*} \right>+\eta_{t}^{2}\left\|\vg_{t}\right\|^{2} \end{aligned}$$
-
-Take the expectation condition on $\vw_t$ over $\xi_t$, i.e., random samples at all devices:
+and according to the definition of $\overline{\vv}_{t+1}$ in \eq{\ref{eq:vbar}}, we can expand the second term in \eq{\ref{eq:sgdcvxsmth1}} as follows:
+\begin{align*}
+\left\|\ov{v}_{t+1}-\vw^{*}\right\|^2 
+ &=\left\|\ov{w}_{t}-\eta_{t} \vg_{t}-\vw^{*}\right\|^2 \\
+ &=\left\|\ov{w}_{t}-\vw^{*}\right\|^{2}-2 \eta_{t} \left< \vg_{t}, \ov{w}_{t}-\vw^{*} \right>+\eta_{t}^{2}\left\|\vg_{t}\right\|^{2} 
+\end{align*}
+We take the expectation condition on $\vw_t$ over $\xi_t$, i.e., random samples at all devices:
 \begin{align}
@@
 \end{align}
+where in the second line and third line we use Lemma~\ref{lma:gradient}.
 

### 2020-04-25 source/sec_theorems.tex
added result summary

@@
 \end{table}
+
+Linear speedup of SGD and Nesterov SGD for strongly convex and smooth
+objectives: full participation $O(\frac{\nu_{max}^{2}\sigma^{2}}{NT}+\frac{E^{2}LG^{2}}{T^{2}})$
+with $\nu_{max}=N\cdot\max_{k}p_{k}$; partial participation $O(\frac{\nu_{max}^{2}\sigma^{2}+4E^{2}G^{2}L}{KT})$.
+In particular, this implies full participation allows $E=O(\sqrt{\frac{T}{N}})$
+to achieve $O(1/NT)$, whereas partial participation only allows $E=O(1)$
+to achieve $O(1/KT)$.
+
+Other accelerations likely not possible.
 

### 2020-04-25 source/sec_theorems.tex
updated result summary
Update sec_theorems.tex

@@
 		paper         &  Cvx Non-smth & Cvx L-smth & Strongly Cvx Non-smth& Strongly Cvx L-smth \\ \hline
-	Accelerated	SGD   &   $\cO(\frac{1}{\sqrt{T}})$     & $\cO(\frac{1}{\sqrt{T}})$   &    $\cO(\frac{1}{T})$   &  $\cO(\frac{1}{T})$    \\\hline
-	SGD    &    $\cO(\frac{1}{\sqrt{T}})$  &   $\cO(\frac{1}{\sqrt{NT}})$  &  $\cO(\frac{1}{T})$  & $\cO(\frac{1}{T})$\cite{li2019convergence,haddadpour2019convergence}, $\cO(\frac{1}{NT})$ ours    \\\hline
+	Accelerated	SGD   &   $\cO(\frac{1}{\sqrt{T}})$     & $\cO(\frac{1}{\sqrt{T}})$   &    $\cO(\frac{1}{T})$   &  $\cO(\frac{1}{KT})$    \\\hline
+	SGD    &    $\cO(\frac{1}{\sqrt{T}})$  &   $\cO(\frac{1}{\sqrt{T}})$  &  $\cO(\frac{1}{T})$  & $\cO(\frac{1}{T})$\cite{li2019convergence,haddadpour2019convergence}, $\cO(\frac{1}{KT})$ ours    \\\hline
 	\end{tabular}
@@
 \end{table}
+
+Acceleration of MaSS for linear regression, partial participation $O((1+\frac{4}{K}\cdot(\frac{t}{E}-1))\exp(-\frac{t}{\sqrt{\kappa_{m}\tilde{\kappa}_{m}}}))$
+over $O((1+\frac{4}{K}\cdot(\frac{t}{E}-1))\exp(-\frac{t}{\kappa_{m}}))$.
 

### 2020-04-25 source/sec_exp.tex
add experiments section

% !TEX ROOT=./main.tex

\section{Numerical Experiments}

\begin{figure}
\begin{tabular}{cc}
	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_loss.pdf} & 
	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_accuracy.pdf} \\
	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_lossuser30TuneE.pdf} & 
	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_lossuser40TuneE.pdf} \\
\end{tabular}
\caption{Each user has 250 samples, non-iid setting, full device participation. Constant learning rate 0.01.}
\end{figure}

### 2020-05-07 source/sec_exp.tex
learning speedup v1

@@
 
+% \begin{figure}
+% \begin{tabular}{cc}
+% 	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_loss.pdf} & 
+% 	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_accuracy.pdf} \\
+% 	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_lossuser30TuneE.pdf} & 
+% 	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_lossuser40TuneE.pdf} \\
+% \end{tabular}
+% \caption{Each user has 250 samples, non-iid setting, full device participation. Constant learning rate 0.01.}
+% \end{figure}
+
+We examine the practical speedup on a linear regression problem, 
+$ F(\vw) = \sum_{k=1}^N p_kF_k(\vw)$, the objective function on each local device $k$ is given by, $F_k(\vw) = \frac{1}{N_k} \sum_{i=1}^{N_k} (\vw^T\vx_i^k + b  - y_i^k)^2$, where we generated i.i.d.
+data $\vx_i^k$ for all devices.
+At each communication round, all devices participated
+in synchronization.
+
 \begin{figure}
-\begin{tabular}{cc}
-	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_loss.pdf} & 
-	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_accuracy.pdf} \\
-	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_lossuser30TuneE.pdf} & 
-	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_lossuser40TuneE.pdf} \\
-\end{tabular}
-\caption{Each user has 250 samples, non-iid setting, full device participation. Constant learning rate 0.01.}
+\centering
+	\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_1k_6k-epsilon01-logTrue.pdf}
+	\caption{The linear speed up w.r.t the number of nodes.
+The synthetic dataset has $6000$ samples, evenly distributed on $10, 100, 1000$ devices.
+The figure shows the number of iterations needed to converge to $\epsilon=0.01$.
+The learning rate is decayed as the $\eta_t = \frac{1}{c + t \times a}$, where we extensively search the best learning rate $c \in \{1, 10\}$ and $a \in \{1\mathrm{e}{-2}, 1\mathrm{e}{-3}, 1\mathrm{e}{-4}, 1\mathrm{e}{-5}, 1\mathrm{e}{-6}\}$ for each configuration.}
 \end{figure}

### 2020-05-08 source/sec_exp.tex
add epsilon=0.001

@@
 \centering
-	\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_1k_6k-epsilon01-logTrue.pdf}
+\begin{tabular}{cc}
+	\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_1k_6k-epsilon01-logTrue.pdf} & 
+\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_1k_6k-epsilon001-logTrue-epoch-1.pdf} \\
+(a) $\epsilon=0.01$  & (b) $\epsilon=0.001$
+\end{tabular}
 	\caption{The linear speed up w.r.t the number of nodes.
 The synthetic dataset has $6000$ samples, evenly distributed on $10, 100, 1000$ devices.
-The figure shows the number of iterations needed to converge to $\epsilon=0.01$.
+The figure shows the number of iterations needed to converge to $\epsilon$.
 The learning rate is decayed as the $\eta_t = \frac{1}{c + t \times a}$, where we extensively search the best learning rate $c \in \{1, 10\}$ and $a \in \{1\mathrm{e}{-2}, 1\mathrm{e}{-3}, 1\mathrm{e}{-4}, 1\mathrm{e}{-5}, 1\mathrm{e}{-6}\}$ for each configuration.}

### 2020-05-10 source/sec_exp.tex
noise

@@
 
+\subsection{SGD}
 We examine the practical speedup on a linear regression problem, 
@@
 The figure shows the number of iterations needed to converge to $\epsilon$.
-The learning rate is decayed as the $\eta_t = \frac{1}{c + t \times a}$, where we extensively search the best learning rate $c \in \{1, 10\}$ and $a \in \{1\mathrm{e}{-2}, 1\mathrm{e}{-3}, 1\mathrm{e}{-4}, 1\mathrm{e}{-5}, 1\mathrm{e}{-6}\}$ for each configuration.}
+The learning rate is decayed as the $\eta_t = \frac{1}{c + t \times a}$, where we extensively search the best learning rate $c \in \{1, 10\}$ and $a \in \{1\mathrm{e}{-2}, 1\mathrm{e}{-3}, 1\mathrm{e}{-4}, 1\mathrm{e}{-5}, 1\mathrm{e}{-6}\}$ for each configuration.
+Batch size is 6 in this case.}
 \end{figure}
+
+Linear regression with noise data.
+
+\begin{figure}
+\centering
+% \begin{tabular}{cc}
+\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_noise1e-02-epsilon1e-02-logTrue-epoch-1-b4}  
+% \includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_1k_6k-epsilon001-logTrue-epoch-1.pdf} \\
+% (a) $\epsilon=0.01$  & (b) $\epsilon=0.001$
+% \end{tabular}
+	\caption{The linear speed up w.r.t the number of nodes.
+The synthetic dataset has $6000$ samples, evenly distributed on $10, 100, 1000$ devices.
+The figure shows the number of iterations needed to converge to $\epsilon$.
+The learning rate is decayed as the $\eta_t = \frac{1}{c + t \times a}$, where we extensively search the best learning rate $c \in \{1, 10\}$ and $a \in \{1\mathrm{e}{-2}, 1\mathrm{e}{-3}, 1\mathrm{e}{-4}, 1\mathrm{e}{-5}, 1\mathrm{e}{-6}\}$ for each configuration.
+Batch size is 4 in this case.
+	We add Gaussian noise with zero mean and 0.01 standard deviation.}
+\end{figure}
+
+% \subsection{Nesterov}

### 2020-05-12 source/sec_exp.tex
convex smooth exp

@@
 	\caption{The linear speed up w.r.t the number of nodes.
-The synthetic dataset has $6000$ samples, evenly distributed on $10, 100, 1000$ devices.
+The synthetic dataset has $6000$ samples, evenly distributed on $10, 50, 100, 500, 1000$ devices.
 The figure shows the number of iterations needed to converge to $\epsilon$.
@@
 
-Linear regression with noise data.
-
 \begin{figure}
@@
 % \begin{tabular}{cc}
-\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_noise1e-02-epsilon1e-02-logTrue-epoch-1-b4}  
-% \includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_1k_6k-epsilon001-logTrue-epoch-1.pdf} \\
-% (a) $\epsilon=0.01$  & (b) $\epsilon=0.001$
+% \includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_noise1e-02-epsilon1e-02-logTrue-epoch-1-b4} \\ 
+\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_noise1e-02-epsilon001-logTrue-epoch-1-b4-adapt0.pdf} \\ 
 % \end{tabular}
+(a) $\epsilon=0.01$ \\
 	\caption{The linear speed up w.r.t the number of nodes.
-The synthetic dataset has $6000$ samples, evenly distributed on $10, 100, 1000$ devices.
+The synthetic dataset has $6000$ samples, evenly distributed on $10, 50, 100, 500, 1000$ devices.
 The figure shows the number of iterations needed to converge to $\epsilon$.
@@
 
+What is the relationship between $N/E$ and iteration $T$ in strongly convex function? $linear regression$, no noise. 
+\begin{itemize}
+	\item $N/E = 10$
+	\item $N=10, E=1$, $N=20, E=2$, $N=30, E=3$, $N=40, E=4$, $N=50, E=5$, $N=100, E=10$.
+\end{itemize}
+
+Does convex smooth function have linear speedup?
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{fig/synthetic_logistic_regression_small-epsilon09-logTrue-epoch-1-b4-regularization0.pdf} \\ 
+(a) $\epsilon=0.9$ \\
+\caption{Convex smooth objectives (logistic regression with zero l2-regularization) The number of iterations w.r.t the number of nodes.
+The synthetic dataset has $6000$ samples, evenly distributed on $10, 50, 100, 500, 1000$ devices.
+The figure shows the number of iterations needed to converge to $\epsilon-$accuracy.
+The learning rate is decayed as the $\eta_t = \frac{1}{c + t \times a}$, where we extensively search the best learning rate $c \in \{1, 10\}$ and $a \in \{1\mathrm{e}{-2}, 1\mathrm{e}{-3}, 1\mathrm{e}{-4}, 1\mathrm{e}{-5}, 1\mathrm{e}{-6}\}$ for each configuration.
+Batch size is 4 in this case.}
+\end{figure}
+
+% /media/Research/linkaixi/env/py36/bin/python /home/linkaixi/Dropbox/recsys/convexFed/main.py --dataset=synthetic_logistic_regression_small --epsilon=0.01 --num_rounds=500000 --num_epochs=1 --number_user=10 --learning_rate=0.001 --is_decay=True --lrconst=1 --machine=gpu --batch_size=4 --num_iteration=50000 --model=logisticregression --optimizer=fedave --dimension=60 --adapt=0
 % \subsection{Nesterov}

### 2020-05-16 source/sec_exp.tex
add more data point

@@
 \centering
+\begin{tabular}{cc}
+\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_1k_6k-epsilon001-logTrue-epoch-1-2-3-4-b4-adapt0.pdf} & 
+\includegraphics[width=0.5\textwidth]{fig/iteration_EandN-synthetic_linear_regression_1k_6k-epsilon01-epoch-1-2-3-4-5-b4-adapt0.pdf} 
+ \\
+(a) $\epsilon=0.01$  & (b) $N/E = 10$ 
+\end{tabular}
+	\caption{The linear speed up w.r.t the number of nodes.
+The synthetic dataset has $6000$ samples, evenly distributed on $10, 20, 30, 40, 50, 60, 80, 100, 200, 300, 400, 500, 600, 750, 1000$ devices.
+The figure shows the number of iterations needed to converge to $\epsilon$.
+The learning rate is decayed as the $\eta_t = \frac{1}{c + t \times a}$, where we extensively search the best learning rate $c \in \{1, 10\}$ and $a \in \{1\mathrm{e}{-2}, 1\mathrm{e}{-3}, 1\mathrm{e}{-4}, 1\mathrm{e}{-5}, 1\mathrm{e}{-6}\}$ for each configuration.
+Batch size is 4 in this case.}
+\end{figure}
+
+\begin{figure}
+\centering
 % \begin{tabular}{cc}
 % \includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_noise1e-02-epsilon1e-02-logTrue-epoch-1-b4} \\ 
-\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_noise1e-02-epsilon001-logTrue-epoch-1-b4-adapt0.pdf} \\ 
+\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_noise1e-02-epsilon01008-logTrue-epoch-1-b4-adapt0.pdf} \\ 
 % \end{tabular}

### 2020-05-22 source/sec_exp.tex
fix figure n/e
update exp
add strongly convex results

--- 2020-05-10 source/sec_exp.tex
@@
 
-% \begin{figure}
-% \begin{tabular}{cc}
-% 	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_loss.pdf} & 
-% 	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_accuracy.pdf} \\
-% 	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_lossuser30TuneE.pdf} & 
-% 	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_lossuser40TuneE.pdf} \\
-% \end{tabular}
-% \caption{Each user has 250 samples, non-iid setting, full device participation. Constant learning rate 0.01.}
-% \end{figure}
+\subsection{Strongly Convex}
 
-\subsection{SGD}
-We examine the practical speedup on a linear regression problem, 
-$ F(\vw) = \sum_{k=1}^N p_kF_k(\vw)$, the objective function on each local device $k$ is given by, $F_k(\vw) = \frac{1}{N_k} \sum_{i=1}^{N_k} (\vw^T\vx_i^k + b  - y_i^k)^2$, where we generated i.i.d.
-data $\vx_i^k$ for all devices.
-At each communication round, all devices participated
-in synchronization.
-
-\begin{figure}
+\begin{figure}[h!]
 \centering
 \begin{tabular}{cc}
-	\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_1k_6k-epsilon01-logTrue.pdf} & 
-\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_1k_6k-epsilon001-logTrue-epoch-1.pdf} \\
-(a) $\epsilon=0.01$  & (b) $\epsilon=0.001$
+	\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupNodesT-min-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} & 
+\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupEpochsT-min-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} 
 \end{tabular}
-	\caption{The linear speed up w.r.t the number of nodes.
-The synthetic dataset has $6000$ samples, evenly distributed on $10, 100, 1000$ devices.
-The figure shows the number of iterations needed to converge to $\epsilon$.
-The learning rate is decayed as the $\eta_t = \frac{1}{c + t \times a}$, where we extensively search the best learning rate $c \in \{1, 10\}$ and $a \in \{1\mathrm{e}{-2}, 1\mathrm{e}{-3}, 1\mathrm{e}{-4}, 1\mathrm{e}{-5}, 1\mathrm{e}{-6}\}$ for each configuration.
-Batch size is 6 in this case.}
+	\caption{Logistic regression with regularization $1e-5$.
+The linear speed up w.r.t the number of nodes and number of epochs.
+The synthetic dataset has $N=49749$ samples, evenly distributed on $1, 2, 4, 8, 16, 32, 64, 128, 256, 512, 1024$ devices.
+The figure shows the number of iterations/rounds needed to converge to $\epsilon-$accuracy.
+The learning rate is decayed as the $\eta_t = \min(32, \frac{Nc}{1 + t})$, where we extensively search the best learning rate $c \in \{2^{-1}c, 2^{-2}c, c, 2c, 2^{2}c\}$ and for each configuration.
+Batch size is 4 in this case.}
 \end{figure}
 
-Linear regression with noise data.
+\subsection{Convex smooth objective}
 
-\begin{figure}
-\centering
-% \begin{tabular}{cc}
-\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_noise1e-02-epsilon1e-02-logTrue-epoch-1-b4}  
-% \includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_1k_6k-epsilon001-logTrue-epoch-1.pdf} \\
-% (a) $\epsilon=0.01$  & (b) $\epsilon=0.001$
-% \end{tabular}
-	\caption{The linear speed up w.r.t the number of nodes.
-The synthetic dataset has $6000$ samples, evenly distributed on $10, 100, 1000$ devices.
-The figure shows the number of iterations needed to converge to $\epsilon$.
-The learning rate is decayed as the $\eta_t = \frac{1}{c + t \times a}$, where we extensively search the best learning rate $c \in \{1, 10\}$ and $a \in \{1\mathrm{e}{-2}, 1\mathrm{e}{-3}, 1\mathrm{e}{-4}, 1\mathrm{e}{-5}, 1\mathrm{e}{-6}\}$ for each configuration.
-Batch size is 4 in this case.
-	We add Gaussian noise with zero mean and 0.01 standard deviation.}
-\end{figure}
-
-% \subsection{Nesterov}
+\subsection{Linear regression}

### 2020-05-22 source/sec_exp_note.tex
add strongly convex results

% !TEX ROOT=./main.tex

\section{Numerical Experiments}

% \begin{figure}
% \begin{tabular}{cc}
% 	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_loss.pdf} & 
% 	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_accuracy.pdf} \\
% 	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_lossuser30TuneE.pdf} & 
% 	\includegraphics[width=0.5\textwidth]{fig/synthetic_balance_0_0_lossuser40TuneE.pdf} \\
% \end{tabular}
% \caption{Each user has 250 samples, non-iid setting, full device participation. Constant learning rate 0.01.}
% \end{figure}

\subsection{SGD}
We examine the practical speedup on a linear regression problem, 
$ F(\vw) = \sum_{k=1}^N p_kF_k(\vw)$, the objective function on each local device $k$ is given by, $F_k(\vw) = \frac{1}{N_k} \sum_{i=1}^{N_k} (\vw^T\vx_i^k + b  - y_i^k)^2$, where we generated i.i.d. data $\vx_i^k$ for all devices. At each communication round, all devices participated
in synchronization.

\begin{figure}
\centering
\begin{tabular}{cc}
	\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_1k_6k-epsilon01-logTrue.pdf} & 
\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_1k_6k-epsilon001-logTrue-epoch-1.pdf} \\
(a) $\epsilon=0.01$  & (b) $\epsilon=0.001$
\end{tabular}
	\caption{The linear speed up w.r.t the number of nodes. The synthetic dataset has $6000$ samples, evenly distributed on $10, 50, 100, 500, 1000$ devices. The figure shows the number of iterations needed to converge to $\epsilon$. The learning rate is decayed as the $\eta_t = \frac{1}{c + t \times a}$, where we extensively search the best learning rate $c \in \{1, 10\}$ and $a \in \{1\mathrm{e}{-2}, 1\mathrm{e}{-3}, 1\mathrm{e}{-4}, 1\mathrm{e}{-5}, 1\mathrm{e}{-6}\}$ for each configuration. Batch size is 6 in this case.}
\end{figure}

\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_1k_6k-epsilon001-logTrue-epoch-1-2-3-4-b4-adapt0.pdf} & 
\includegraphics[width=0.5\textwidth]{fig/iteration_EandN-synthetic_linear_regression_1k_6k-epsilon01.pdf} 
 \\
(a) $\epsilon=0.01$  & (b) $N/E = 10$ 
\end{tabular}
	\caption{The linear speed up w.r.t the number of nodes. The synthetic dataset has $6000$ samples, evenly distributed on $10, 20, 30, 40, 50, 60, 80, 100, 200, 300, 400, 500, 600, 750, 1000$ devices. The figure shows the number of iterations needed to converge to $\epsilon$. The learning rate is decayed as the $\eta_t = \frac{1}{c + t \times a}$, where we extensively search the best learning rate $c \in \{1, 10\}$ and $a \in \{1\mathrm{e}{-2}, 1\mathrm{e}{-3}, 1\mathrm{e}{-4}, 1\mathrm{e}{-5}, 1\mathrm{e}{-6}\}$ for each configuration. Batch size is 4 in this case.}
\end{figure}

\begin{figure}
\centering
% \begin{tabular}{cc}
% \includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_noise1e-02-epsilon1e-02-logTrue-epoch-1-b4} \\ 
\includegraphics[width=0.5\textwidth]{fig/synthetic_linear_regression_noise1e-02-epsilon01008-logTrue-epoch-1-b4-adapt0.pdf} \\ 
% \end{tabular}
(a) $\epsilon=0.01$ \\
	\caption{The linear speed up w.r.t the number of nodes. The synthetic dataset has $6000$ samples, evenly distributed on $10, 20, 30, 40, 50, 60, 80, 100, 200, 300, 400, 500, 600, 750, 1000$ devices. The figure shows the number of iterations needed to converge to $\epsilon$. The learning rate is decayed as the $\eta_t = \frac{1}{c + t \times a}$, where we extensively search the best learning rate $c \in \{1, 10\}$ and $a \in \{1\mathrm{e}{-2}, 1\mathrm{e}{-3}, 1\mathrm{e}{-4}, 1\mathrm{e}{-5}, 1\mathrm{e}{-6}\}$ for each configuration. Batch size is 4 in this case. 
	We add Gaussian noise with zero mean and 0.01 standard deviation.}
\end{figure}

What is the relationship between $N/E$ and iteration $T$ in strongly convex function? $linear regression$, no noise. 
\begin{itemize}
	\item $N/E = 10$
	\item $N=10, E=1$, $N=20, E=2$, $N=30, E=3$, $N=40, E=4$, $N=50, E=5$, $N=100, E=10$.
\end{itemize}

Does convex smooth function have linear speedup?

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{fig/synthetic_logistic_regression_small-epsilon09-logTrue-epoch-1-b4-regularization0.pdf} \\ 
(a) $\epsilon=0.9$ \\
\caption{Convex smooth objectives (logistic regression with zero l2-regularization) The number of iterations w.r.t the number of nodes. The synthetic dataset has $6000$ samples, evenly distributed on $10, 50, 100, 500, 1000$ devices. The figure shows the number of iterations needed to converge to $\epsilon-$accuracy. The learning rate is decayed as the $\eta_t = \frac{1}{c + t \times a}$, where we extensively search the best learning rate $c \in \{1, 10\}$ and $a \in \{1\mathrm{e}{-2}, 1\mathrm{e}{-3}, 1\mathrm{e}{-4}, 1\mathrm{e}{-5}, 1\mathrm{e}{-6}\}$ for each configuration. Batch size is 4 in this case.}
\end{figure}

\begin{figure}
\centering
\begin{tabular}{cc}
	\includegraphics[width=0.5\textwidth]{fig/speedupEpochsT-synthetic_logistic_regression_iclr300-epsilon07-b4-adapt0.pdf} & 
\includegraphics[width=0.5\textwidth]{fig/speedupEpochsRounds-synthetic_logistic_regression_iclr300-epsilon07-b4-adapt0.pdf} \\
	\includegraphics[width=0.5\textwidth]{fig/speedupNodesT-synthetic_logistic_regression_iclr300-epsilon07-b4-adapt0.pdf} & 
\includegraphics[width=0.5\textwidth]{fig/speedupNodesRounds-synthetic_logistic_regression_iclr300-epsilon07-b4-adapt0.pdf} \\
\end{tabular}
	\caption{Logistic regression with regularization $0.001$. The linear speed up w.r.t the number of nodes and number of epochs. The synthetic dataset has $42000$ samples, evenly distributed on $10, 20, 30, 40, 50, 100, 1000$ devices. The figure shows the number of iterations/rounds needed to converge to $\epsilon-$accuracy. The learning rate is decayed as the $\eta_t = \frac{1}{1 + t \times a}$, where we extensively search the best learning rate $c \in \{1, 10\}$ and $a \in \{1, 0.1, 0.01\}$ for each configuration. Batch size is 4 in this case.}
\end{figure}

\begin{figure}
\centering
\begin{tabular}{cc}
	\includegraphics[width=0.5\textwidth]{fig/speedupEpochsT-synthetic_logistic_regression_iclr300-epsilon07-b4-reg0-adapt0.pdf} & 
\includegraphics[width=0.5\textwidth]{fig/speedupEpochsRounds-synthetic_logistic_regression_iclr300-epsilon07-b4-reg0-adapt0.pdf} \\
	\includegraphics[width=0.5\textwidth]{fig/speedupNodesT-synthetic_logistic_regression_iclr300-epsilon07-b4-reg0-adapt0.pdf} & 
\includegraphics[width=0.5\textwidth]{fig/speedupNodesRounds-synthetic_logistic_regression_iclr300-epsilon07-b4-reg0-adapt0.pdf} \\
\end{tabular}
	\caption{Logistic regression with regularization $0$. The linear speed up w.r.t the number of nodes and number of epochs. The synthetic dataset has $42000$ samples, evenly distributed on $10, 20, 30, 40, 50, 100, 1000$ devices. The figure shows the number of iterations/rounds needed to converge to $\epsilon-$accuracy. The learning rate is decayed as the $\eta_t = \frac{1}{1 + t \times a}$, where we extensively search the best learning rate $c \in \{1, 10\}$ and $a \in \{1, 0.1, 0.01\}$ for each configuration. Batch size is 4 in this case.}
\end{figure}

% /media/Research/linkaixi/env/py36/bin/python /home/linkaixi/Dropbox/recsys/convexFed/main.py --dataset=synthetic_logistic_regression_small --epsilon=0.01 --num_rounds=500000 --num_epochs=1 --number_user=10 --learning_rate=0.001 --is_decay=True --lrconst=1 --machine=gpu --batch_size=4 --num_iteration=50000 --model=logisticregression --optimizer=fedave --dimension=60 --adapt=0
% \subsection{Nesterov}

### 2020-05-23 source/save.tex
first draft

--- 2020-04-18 source/sec_theorems.tex
@@
-% !TEX ROOT=./main.tex
-
-\section{Convergence Results}
-This section summarizes the current results and assumptions.
-\begin{table}[h!]
-\centering
-\small
-	\begin{tabular}{|c|c|c|c|c|}\hline
-		paper         &  Cvx Non-smth & Cvx L-smth & Strongly Cvx Non-smth& Strongly Cvx L-smth \\ \hline
-	Accelerated	SGD   &   $\cO(\frac{1}{\sqrt{T}})$     & $\cO(\frac{1}{\sqrt{T}})$   &    $\cO(\frac{1}{T})$   &  $\cO(\frac{1}{T})$    \\\hline
-	SGD    &    $\cO(\frac{1}{\sqrt{T}})$  &   $\cO(\frac{1}{\sqrt{NT}})$  &  $\cO(\frac{1}{T})$  & $\cO(\frac{1}{T})$\cite{li2019convergence,haddadpour2019convergence}, $\cO(\frac{1}{NT})$ ours    \\\hline
-	\end{tabular}
-	\caption{Summarize of related work on federated learning, heterogeneous data.}
-\end{table}
-
-\begin{assumption}[Bounded local variance]
-Let $\xi_{t}^{k}$ be sampled from the $k$ -th device's local data uniformly at random.
-The variance of stochastic gradients in each device is bounded: $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq \sigma_{k}^{2}$,
-for $k=1, \cdots, N$.	
-\end{assumption}
-
-% \begin{assumption}
-% The expected squared norm of stochastic subgradients is uniformly bounded. i.e.,
-% $\mathbb{E}\|\vg_{t,k}\|^2  \leq G_k^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
-This also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$.
-Denote $G^2 = \sum_{k=1}^N p_k G_k^2$
-% \label{ass:subgrad2}
-% \end{assumption}
-
-\begin{assumption}[Expected square norm of gradient]
-The expected squared norm of stochastic gradients is uniformly bounded. i.e.,
-$\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)\right\|^{2} \leq G^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
-This also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$.
-\label{ass:subgrad2}
-\end{assumption}
-
-\begin{assumption}[L-smooth]
-$F_{1}, \cdots, F_{N}$ are all $L$-smooth: for all  $\mathbf{v}$  and $\mathbf{w}$, $F_{k}(\mathbf{v}) \leq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{L}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$.
-\label{ass:lsmooth}
-\end{assumption}
-\begin{assumption}[Strongly-convex]
-$	F_{1}, \cdots, F_{N} \text { are all } \mu \text { -strongly convex: for all v and } \mathbf{w}, F_{k}(\mathbf{v}) \geq F_{k}(\mathbf{w})+(\mathbf{v}-\mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{\mu}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$
-\label{ass:stroncvx}
-\end{assumption}
-
-\begin{definition}[Squared Weighted Gradient Diversity]
-	We indicate the following quantity as weighted gradient diversity among the local 
-	objectives:
-	\begin{align*}
-		\frac{\sum_{k=1}^K p_k  \mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}, \xi^{k}\right)\right\|^{2} }{\mathbb{E}\left\| \sum_{k=1}^K p_k   \nabla F_{k}\left(\mathbf{w}, \xi^{k}\right)\right\|^{2} } \leq \lambda
-	\end{align*}
-\end{definition}
-This definition is similar to the weighted gradient diversity defined in~\cite{haddadpour2019convergence}.
-
-\begin{definition}[Gradient Dissimilarity~\cite{li2018federated}]
-	We indicate the following quantity as gradient diversity among the local 
-	objectives:
-	\begin{align*}
-		\frac{\sum_{k=1}^K p_k \left\|\nabla F_{k}\left(\mathbf{w}\right)\right\|^{2} }{\left\|   \nabla F\left(\mathbf{w}\right)\right\|^{2} } \leq \lambda 
-	\end{align*}
-	For some $\epsilon>0$, for all the points $\vw \in \mathcal{S}_{\epsilon}^{c}=\left\{\vw |\|\nabla F(\vw)\|^{2}>\epsilon\right\}$
-\end{definition}
-
 \section{The Convergence of FedAve for General Convex Functions}
-This section summarizes the convergence of FedAve.
-The updates of FedAve with partial device activation is given by: 
-\begin{align} 
-\mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\
-\sum_{k \in \cS_{t+1}} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}\end{array}\right.
-\end{align}
 

### 2020-05-23 source/sec_exp.tex
first draft

@@
 	\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupNodesT-min-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} & 
-\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupEpochsT-min-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} 
+\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupEpochsT-min-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} \\
+	\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupNodesRounds-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} & 
+\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupEpochsRounds-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} \\
 \end{tabular}

### 2020-05-23 source/sec_linear_regression.tex
first draft

% !TEX ROOT=./main.tex

\section{Linear Regression}

In contrast to the gradient descent setting, in the stochastic gradient
setting, Nesterov update is known to fail to accelerate over stochastic
gradient. Thus in general we cannot hope to obtain acceleration results
for the FedAvg algorithm with Nesterov updates. However, in special
problem settings, there have been a number of works showing that accelerated
stochastic gradient descent methods achieves better rates than stochastic
gradient descent. One such case is the so-called interpolation setting,
where the objective has an optimal value of 0, e.g. due to overparameterization.
The works of Belkin et.al have shown that under interpolation, SGD
achieves exponential convergence under constant step-size, thanks
to the automatic variance reduction property that is not available
in generic optimization problems. In this section, we extend their
result to the federated learning problem and prove the exponential
convergence of SGD for a class of convex objectives in the interpolation
setting.

In the non-federated learning setting, {[}TODO: CITE{]} have provided
negative results on the acceleration of Nesterov and Heavy Ball updates
over plain SGD for certain problems. In {[}Liu\&Belkin{]}, the authors
show that in the interpolation setting, stochastic Nesterov update
is not able to achieve accleration over vanilla SGD. They introduce
the MaSS algorithm, which is a variant of the Nesterov acceleration,
where in the update a non-negative multiple of the gradient is added
to the Nesterov parameter update to correct for the ``over-descent''
of the Nesterov update. They show that the MaSS algorithm is able
to achieve acceleration over the exponential convergence of SGD both
theoretically and empirically.

The next natural question is then whether similar conclusions can
be drawn in the federated learning setting. We demonstrate that this
is possible by extending the results of Belkin et al. to show that
for the federated learning problem in the interpolation setting, Nesterov
SGD cannot in general achieve acceleration over plain SGD. We then
adapt the MaSS algorithm to the federated learning setting, and prove
that for certain class of convex objectives with 0 global objective
value, the FedAvg algorithm with MaSS updates achieves exponential
convergence and acceleration of convergence rate over SGD.

We first show that FedAvg with MaSS has exponential convergence for
linear regression when the global objective has 0 as its global minimum.

\subsection{Notation and Definitions}

Following Liu\&Belkin and Jain et al., we define some condition number
related quantities. Let $H^{k}=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}x_{k,j}x_{k,j}$
be the Hessian matrix of $F_{k}$. Let $L^{k}$ and $\mu^{k}$ be
any upper bound and lower bound of the non-zero eigenvalues of the
Hessians $H^{k}$. For a mini-batch $\{\tilde{x}_{j}\}_{j=1}^{m}$
of $m$ samples from device $k$, let $\tilde{H}_{m}^{k}=\frac{1}{m}\sum_{j=1}^{m}\tilde{x}_{j}\tilde{x}_{j}^{T}$
be the unbiased mini-batch estimate of $H^{k}$. Let $L_{1}^{k}$
be the smallest positive number such that 
\begin{align*}
\mathbb{E}\left[\|\tilde{x}\|^{2}\tilde{x}\tilde{x}^{T}\right] & \preceq L_{1}^{k}H^{k}
\end{align*}
and define 
\begin{align*}
L_{m}^{k} & =L_{1}^{k}/m+(m-1)L^{k}/m
\end{align*}

Define the $m$-stochastic condition number as $\kappa_{m}^{k}:=L_{m}^{k}/\mu$.
Define $L_{m}=\max_{k}L_{m}^{k}$ and $\kappa_{m}=\max_{k}\kappa_{m}^{k}$.
Define the statistical condition number $\tilde{\kappa}^{k}$ as the
smallest positive real number such that 
\begin{align*}
\mathbb{E}\left[\langle\tilde{x}(H^{k})^{-1},\tilde{x}\rangle\tilde{x}\tilde{x}^{T}\right] & \preceq\tilde{\kappa}^{k}H^{k}
\end{align*}

In order to use hyperparameters that are universal across devices,
we further define $L=\max_{k}L^{k}$, $\mu=\min_{k}\mu^{k}$, $L_{m}=\max_{k}L_{m}^{k}$
and $\kappa_{m}=\max_{k}\kappa_{m}^{k}$.

\subsection{FedAvg with MaSS Updates}

The FedAvg algorithm with MaSS follows the updates
\begin{align*}
w_{t+1}^{k} & =\begin{cases}
u_{t}^{k}-\eta_{1}^{k}g_{t,k} & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}\left[u_{t}^{k}-\eta_{1}^{k}g_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}\\
u_{t+1}^{k} & =w_{t+1}^{k}+\gamma^{k}(w_{t+1}^{k}-w_{t}^{k})+\eta_{2}^{k}g_{t,k}
\end{align*}
where we note that the natural parameter is $w_{t}$, while $u_{t}$
is an auxiliary parameter, which we initialize to be $u_{0}^{k}$,
and 
\begin{align*}
g_{t,k} & :=\nabla F_{k}(u_{t}^{k},\xi_{t}^{k})
\end{align*}
is the stochastic gradient and 
\begin{align*}
g_{t} & =\sum_{k=1}^{N}p_{k}g_{t,k}
\end{align*}
is the averaged stochastic gradient. When $\eta_{2}^{k}\equiv0$,
this reduces to the FedAvg algorithm with Nesterov updates.

We note that the update can equivalently be written as 
\begin{align*}
v_{t+1}^{k} & =(1-\alpha^{k})v_{t}^{k}+\alpha^{k}u_{t}^{k}-\delta^{k}g_{t,k}\\
w_{t+1}^{k} & =\begin{cases}
u_{t}^{k}-\eta^{k}g_{t,k} & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}\left[u_{t}^{k}-\eta^{k}g_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}\\
u_{t+1}^{k} & =\frac{\alpha^{k}}{1+\alpha^{k}}v_{t+1}^{k}+\frac{1}{1+\alpha^{k}}w_{t+1}^{k}
\end{align*}
where there is a bijection between the parameters 
\begin{align*}
\frac{1-\alpha^{k}}{1+\alpha^{k}} & =\gamma^{k}\\
\eta^{k} & =\eta_{1}^{k}\\
\frac{\eta^{k}-\alpha^{k}\delta^{k}}{1+\alpha^{k}} & =\eta_{2}^{k}
\end{align*}
and we further introduce an auxiliary parameter $v_{t}^{k}$, which
is initialized at $v_{0}^{k}$. We also note that when $\delta^{k}=\frac{\eta^{k}}{\alpha^{k}}$,
the update reduces to the Nesterov accelerated SGD. This version of
the FedAvg with MaSS algorithm is used for analyzing the exponential
convergence.

\subsection{Exponential Convergence of FedAvg with MaSS}

We now present the exponential convergence result in linear regression
using FedAvg with MaSS updates. On each device, local data is stored
and mini-batch gradient descent with batch size $m$ is performed.
We assume that the batch size is the same across devices. 
\begin{theorem}
	(FedAvg with MaSS, Linear Regression) Let $\tilde{\kappa}_{m}:=\tilde{\kappa}/m+(m-1)/m$,
	and let the hyperparameters satisfy 
	\begin{align*}
	\eta^{k}(m)=\frac{1}{L_{m}}, & \alpha^{k}(m)=\frac{1}{\sqrt{\kappa_{m}\tilde{\kappa}_{m}}},\delta^{k}(m)=\frac{\eta^{k}}{\alpha^{k}\tilde{\kappa}_{m}}
	\end{align*}
	for all $k$. Let $\mathcal{I}_{E}=\{\ell E\mid\ell\in\mathbb{N}\}$
	be the set of communication rounds, then the FedAvg algorithm with
	MaSS and full participation solves the problem 
	\begin{align*}
	F(w) & =\sum_{k=1}^{N}p_{k}F_{k}(w)\\
	F_{k}(w) & =\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(w^{T}x_{k,j}-z_{k,j})^{2}
	\end{align*}
	with exponential convergence
	\begin{align*}
	\mathbb{E}F(\overline{w}_{t})\leq\frac{L}{2}\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2} & \leq C\cdot\exp(-\frac{t}{\sqrt{(\kappa_{N}+(E-1)^{2})(\tilde{\kappa}_{N}+(E-1)^{2})}})
	\end{align*}
	where $w^{\ast}$ is such that $F(w^{\ast})=0$ and $C=\frac{L}{2}\cdot\frac{\alpha}{\delta}\sum_{k=1}^{N}p_{k}(\|v_{0}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta}{\alpha}\|w_{0}^{k}-w^{\ast}\|^{2})$.

In particular, this implies that for $E=1$, there is linear speedup, i.e. 
	\begin{align*}
	\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq C\cdot\exp(-\frac{t}{\sqrt{(\kappa_{N})(\tilde{\kappa}_{N})}})\leq C\cdot\exp(-\frac{Nt}{\sqrt{(\kappa_{1})(\tilde{\kappa})}})
	\end{align*}

	With partial participation of $K$ devices each communication round,
	the convergence rate is also 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{t})\leq\frac{L}{2}\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2} & \leq C\cdot(1+\frac{4}{K}\cdot(\frac{t}{E}-1))\cdot(1-\frac{1}{\sqrt{\kappa_{K}\tilde{\kappa}_{K}}})^{t}=O(\exp(-\frac{t}{\sqrt{(\kappa_{N}+(E-1)^{2})(\tilde{\kappa}_{N}+(E-1)^{2})}}))
	\end{align*}

\end{theorem}

### 2020-05-23 source/sec_problem.tex
reorg
first draft

--- 2020-04-01 source/sec_problem.tex
@@
 
-\section{Problem Setting}
+\section{Problem Setup}
+
+\subsection{The Federated Learning Problem}
+
+The federated learning problem, given local objectives $F_k$ and device weights $p_k$, is to minimize
 
@@
 \end{align}
-where $N$ is the number of devices, and $p_k$ is the weight of the k-th device
-such that $p_k \geq 0$ and $\sum_{k=1}^N p_k = 1$. Suppose the k-th device
-holds the nk training data: $x_{k,1}, x_{k,2}, \dots, x_{k,n_k}$ . The local
-$k \in S_t^K$ objective $F_k(\cdot)$ is defined by
+
+where $N$ is the number of devices, and device weights $p_k$ satisfy $p_k \geq 0$ and $\sum_{k=1}^N p_k = 1$. Suppose the $k$-th device
+holds the $n_k$ data points: $\xi_{k,1}, \xi_{k,2}, \dots, \xi_{k,n_k}$ . We assume that the local objective $F_k(\cdot)$ is defined by
 
 \begin{align}
-F_{k}(\mathbf{w}) \triangleq \frac{1}{n_{k}} \sum_{j=1}^{n_{k}} \ell\left(\mathbf{w} ; x_{k, j}\right)	
+F_{k}(\mathbf{w}) \triangleq \frac{1}{n_{k}} \sum_{j=1}^{n_{k}} \ell\left(\mathbf{w} ; \xi_{k, j}\right)	
 \label{eq:localloss}
@@
 
-% The summary of all assumptions:
+where $\ell$ is some loss function.
+Thus, in the federated learning setting, each device stores its local data, which gives rise to its own local objective $F_k$.
+A key difference with many problems in the distributed optimization setting is that data on local devices is heterogeneous, i.e.
+$F_k(w)$ may have different minimizers $w^{\ast}_k$.
+Moreover, each local device only has access to stochastic gradients based on data stored on the local device.
+In this paper, we investigate local objectives that are convex or strongly convex and smooth, although convergence results for non-smooth objectives are also given in the appendix.
 
-% \begin{assumption}
-% $F_{1}, \cdots, F_{N}$ are all $L$-smooth: for all  $\mathbf{v}$  and $\mathbf{w}$, $F_{k}(\mathbf{v}) \leq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{L}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$.
-% \end{assumption}
+\subsection{The FedAvg Algorithm}
 
-% \begin{assumption}
-% $F_1,\dots, F_N$ are all convex: for all $\vv$ and $\vw$, 
-% $F_k(\vv) \geq F_k(\vw)+(\vv -\vw)^T \grad F_k(\vw)$. \label{ass:cvx}
-% \end{assumption}
-% \begin{assumption}
-% $	F_{1}, \cdots, F_{N} \text { are all } \mu \text { -strongly convex: for all v and } \mathbf{w}, F_{k}(\mathbf{v}) \geq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{\mu}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$
-% \end{assumption}
+Despite the heterogeneity in local objectives, the federated learning
+problem can still be solved efficiently in a distributed manner with
+the simple Federated Average (FedAvg) algorithm, which consists of
+the iterations
 
-% \begin{assumption}
-% Let $\xi_{t}^{k}$ be sampled from the k-th device's local data uniformly at random. The variance of stochastic gradients in each device is bounded:
-% $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq \sigma_{k}^{2}$, for $k = 1,..., N$.
-% \end{assumption}
+\begin{align}
+\vv_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}g_{t,k}\\
+\vw_{t+1}^{k} & =\begin{cases}
+\vv_{t+1}^{k} & \text{if }t+1\notin\mathcal{I}_{E}\\
+\sum_{k=1}^{N}p_{k}\vv_{t+1}^{k} & \text{if }t+1\in\mathcal{I}_{E}
+\end{cases}
+\end{align}
 
-% \begin{assumption}
-% The expected squared norm of stochastic gradients is uniformly bounded. i.e.,
-% $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)\right\|^{2} \leq G^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
-% \end{assumption}
+where $\vw_t^k$ is the local model parameter maintained in the $k$-th device during the $t$-th iteration, 
+\begin{align}
+g_{t,k} & :=\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})
+\end{align} 
+is the stochastic gradient based on $\xi_{t}^{k}$, the data sampled from k-th device’s local data uniformly at random. $\mathcal{I}_{E}=\{E,2E,\dots\}$
+is the set of global synchronization steps. Thus when $t+1 \notin \mathcal{I}_{E}$, each device updates its local parameter based on local
+stochastic gradient descent, and every $E$ iterations, parameters are aggregated
+across devices based on weights $p_{k}$ and then broadcast back to
+each local device.
 
-\section{Notations}
+Communication and broadcasting of local parameters
+is costly, so the efficiency of the FedAvg algorithm requires that
+$E$ is relatively large, while at the same time ensuring convergence
+to a global minimizer of the global objective $\sum_{k}p_{k}F_{k}(w)$.
 
-Let $\vw_t^k$ be the model parameter maintained in the k-th device at the t-th step.
-$\cI_E$ is a set of global synchronization steps, i.e., $\cI_E = \{n E|n = 1, 2, 3,\dots \}$.
-If $t+1 \in \cI_E$, which means we communicate the server with (all) clients at time step $t+1$.
-$\vv^k_{t+1}$ is the immediate result of one step SGD update from $\vw^k_{t}$.
-$\xi_{t}^{k}$ denotes the data sampled from k-th device’s local data uniformly at random.
-Follow the common practice, we define two virtual sequences $\overline{\mathbf{v}}_{t}$ and $\overline{\mathbf{w}}_{t}$. For full device participation and $t \notin \cI_E$,
+In practical applications, it may not be possible to collect local
+parameters from \emph{all }devices at a specified communication time.
+Thus the above \emph{full participation }FedAvg algorithm may not
+be feasible. Instead, a more practical algorithm is to allow only
+a random subset $\mathcal{S}_{t+1}$ of all devices to participate in parameter aggregation, while
+broadcasting the aggregated parameter to all devices. This gives rise
+to the \emph{partial participation }FedAvg algorithm.
+
+The updates of FedAvg with partial device activation is given by: 
+\begin{align} 
+\mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\
+\sum_{k \in \cS_{t+1}} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}\end{array}\right.
+\end{align}
+
+where $\mathcal{S}_{t+1}$ is obtained using some sampling scheme.
+We consider two types of common unbiased sampling strategies.
+The first one samples each device with probability $p_k$, with replacement, while the second one samples each device uniformly without replacement.
+Both of these sampling schemes guarantee that updates in FedAvg with partial participation are unbiased stochastic versions of updates in FedAvg with full participation, during the communication round.
+
+\subsection{Notation and Assumptions}
+
+Following common practice, we define two virtual sequences $\overline{\mathbf{v}}_{t}$ and $\overline{\mathbf{w}}_{t}$. For full device participation and $t \notin \cI_E$,
 $\ov{v}_t = \ov{w}_t =\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$.
-In partial participation, $t \in \cI_E$, $\ov{w}_t \neq \ov{v}_t$ since $\ov{v}_t=\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$ while $\sum_{k\in \cS_t}\mathbf{w}_{t}^{k}$.
+For partial participation, $t \in \cI_E$, $\ov{w}_t \neq \ov{v}_t$ since $\ov{v}_t=\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$ while $\ov{w}_t=\sum_{k\in \cS_t}\mathbf{w}_{t}^{k}$.
 However, we can
@@
 \end{align}
-and we use $\xi_t = \{\xi_t^k\}_{k=1}^N$ denotes samples at all devices at time step $t$.
+and $\xi_t = \{\xi_t^k\}_{k=1}^N$ denotes samples at all devices at time step $t$.
 
-The updates of FedAve with partial device activation is given by: 
-\begin{align} 
-\mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\
-\sum_{k \in \cS_{t+1}} p_{k} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}\end{array}\right.
-\end{align}
+We will require some standard assumptions on the local objectives and their gradients.
 
-In~\cite{li2019convergence}, two types of unbiased sampling strategies are considered in Lemma 5. 
-The sampling scheme I establishes $\cS_{t+1}$ by i.i.d. sampling the devices with replacement,
-in this case the upper bound of expected square norm of $\ov{w}_{t+1} - \ov{v}_{t+1}$ is given by:
-\begin{align}
-\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4}{K} \eta_t^2 E^2G^2
-\end{align}
-The sampling scheme II establishes $\cS_{t+1}$ by uniformly sampling all devices without
-replacement, in which we have the 
-\begin{align}
-\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4(N - K)}{K(N-1)} \eta_t^2 E^2G^2
-\end{align}
-We denote this upper bound as follows for concise presentation. 
-\begin{align}
-	\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2 \leq  \eta_t^2 C
-	\label{eq:partialsample}
-\end{align}
+\begin{assumption}[Bounded local variance]
+	Let $\xi_{t}^{k}$ be sampled from the $k$-th device's local data uniformly at random.
+The variance of stochastic gradients in each device is bounded: $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq \sigma_{k}^{2}$,
+	for $k=1, \cdots, N$ and any $\mathbf{w}_{t}^{k}$.
+\end{assumption}
 
-\lkxcom{quantify biased sampling scheme ?}
+\begin{assumption}[Expected square norm of gradient]
+	The expected squared norm of stochastic gradients is uniformly bounded. i.e.,
+	$\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)\right\|^{2} \leq G^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
+	\label{ass:subgrad2}
+\end{assumption}
+
+\begin{assumption}[L-smooth]
+	$F_{1}, \cdots, F_{N}$ are all $L$-smooth: for all  $\mathbf{v}$  and $\mathbf{w}$, $F_{k}(\mathbf{v}) \leq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{L}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$.
+	\label{ass:lsmooth}
+\end{assumption}
+
+\begin{assumption}[Strongly-convex]
+	$	F_{1}, \cdots, F_{N} \text { are all } \mu \text { -strongly convex: for all v and } \mathbf{w}, F_{k}(\mathbf{v}) \geq F_{k}(\mathbf{w})+(\mathbf{v}-\mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{\mu}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$
+	\label{ass:stroncvx}
+\end{assumption}
+
+Even though Assumption \ref{ass:subgrad2} also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$, which further implies $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq G^2$, the two assumptions are stated separately because it makes explicit the dependence of the optimality gap on different terms.

### 2020-05-23 source/sec_strongly_convex_smooth.tex
first draft

% !TEX ROOT=./main.tex

\section{Strongly Convex Smooth Objectives}

In this section, we show that FedAvg with SGD and Nesterov Accelerated SGD has $O(1/NT)$ rate under $\mu$-strong
convexity and $L$-smoothness. The result improves on previous results of
$O(1/T)$ with a linear speedup in the number of devices $N$.

\subsection{SGD}

\begin{theorem}
	Suppose $F_{k}$ is $L$-smooth and $\mu$-strongly convex for all
	$k$. Let $\kappa=\frac{L}{\mu}$, $\gamma=\max\{8\kappa-1,E\}$ where
	$E$ is the communication interval, and learning rates 
	\begin{align*}
	\alpha_{t} & =\frac{c}{\mu(\gamma+t)}
	\end{align*}
	so that $\alpha_{t}\leq2\alpha_{t+E}$, and where $0\leq c\leq1$
	is small enough such that
	$\alpha_{t} \leq\frac{1}{4L}$ and 
	$\alpha_{t} \leq\frac{1}{N}$ for all $t\geq0$.

	Then with full device participation, if $\nu_{max}=\max_{k}\frac{1}{N}p_{k}$,
	we have 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq C\frac{\kappa}{\gamma+T}\cdot(8E^{2}LG^{2}/T+\frac{1}{N}\nu_{max}^{2}\sigma^{2})
	\end{align*}
	so that as long as $E=O(\sqrt{\frac{N}{T}})$, 
	\begin{align*}
	\mathbb{E}F(w_{T})-F^{\ast} & \leq C(8LG^{2}+\nu_{max}^{2}\sigma^{2})\frac{\kappa}{N(\gamma+T)}
	\end{align*}
	and with partial participation with $K$ sampled devices at each
	communication round, 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq C\frac{\kappa}{\gamma+T}\cdot(8E^{2}LG^{2}/T+\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{4}{K}E^{2}G^{2})
	\end{align*}
	so that if $E=O(1)$, 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast} & \leq C(4E^{2}G^{2}+\nu_{max}^{2}\sigma^{2})\frac{\kappa}{K(\gamma+T)}+\frac{8E^{2}LG^{2}}{T^{2}}
	\end{align*}
\end{theorem}
\begin{proof}
	See the appendix.
\end{proof}

The convergence result implies that with fixed $E$, as long as $N$ satisfies $E=O(\sqrt{T/N})$, there is linear speedup. When $N$ exceeds an upper bound, however, linear speedup may fail to happen, and $T/N$ may remain constant, i.e. the number of iterations required for convergence may increase with $N$.

\subsection{Accelerated SGD}
In this section, we show that FedAvg with Nesterov accelerated SGD updates enjoys similar convergence rates with linear speedup.

The FedAv algorithm with Nesterov Accelerated
SGD follows the updates
\begin{align*}
y_{t+1}^{k} & =w_{t}^{k}-\alpha_{t}g_{t,k}\\
w_{t+1}^{k} & =\begin{cases}
y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}\left[y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}
\end{align*}
where 
\begin{align*}
g_{t,k} & :=\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})
\end{align*}
is the stochastic gradient.

\begin{theorem}
	Suppose $F_{k}$ is $L$-smooth and $\mu$-strongly convex for all
	$k$. Let $\kappa=\frac{L}{\mu}$, $\gamma=\max\{8\kappa-1,E\}$ where
	$E$ is the communication interval, and learning rates 
	\begin{align*}
	\alpha_{t} & =\frac{c}{\mu(\gamma+t)}\\
	\beta_{t} & \leq\alpha_{t}
	\end{align*}
	so that $\alpha_{t}\leq2\alpha_{t+E}$, and where $0\leq c\leq1$
	is small enough such that the following hold: 
	\begin{align*}
	\alpha_{t}^{2}+\beta_{t-1}^{2} & \leq\frac{1}{2}\\
	\alpha_{t} & \leq\frac{1}{4L}\\
	4\alpha_{t-1}^{2} & \leq\alpha_{t}
	\end{align*}
	for all $t\geq0$. Suppose also that $G$ is a constant satisfying
	$\mathbb{E}\|w_{0}-\alpha_{0}g_{0,k}\|^{2}=\mathbb{E}\|w_{0}-\alpha_{0}\nabla F_{k}(w_{0},\xi_{0}^{k})\|^{2}\leq G^{2}$
	for all $k$, and $\mathbb{E}\|g_{t,k}\|^{2}\leq G^{2}$ for all $t,k$.

	Then with full device participation, if $\nu_{max}=\max_{k}\frac{1}{N}p_{k}$,
	we have 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq C\frac{\kappa}{\gamma+T}\cdot(32E^{2}LG^{2}/T+\frac{1}{N}\nu_{max}^{2}\sigma^{2})
	\end{align*}
	so that as long as $E=O(\sqrt{\frac{N}{T}})$, 
	\begin{align*}
	\mathbb{E}F(w_{T})-F^{\ast} & \leq C(32LG^{2}+\nu_{max}^{2}\sigma^{2})\frac{\kappa}{N(\gamma+T)}
	\end{align*}
	and with partial participation with $K$ sampled devices at each
	communication round, 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq C\frac{\kappa}{\gamma+T}\cdot(32E^{2}LG^{2}/T+\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{16}{K}E^{2}G^{2})
	\end{align*}
	so that if $E=O(1)$, 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast} & \leq C(16E^{2}G^{2}+\nu_{max}^{2}\sigma^{2})\frac{\kappa}{K(\gamma+T)}+\frac{32E^{2}LG^{2}}{T^{2}}
	\end{align*}
\end{theorem}

### 2020-05-24 source/main.tex
partial cvx smth nasgd
reorg
Update main.tex
refine proof v1.
overview re arrange.
add experiments section
add strongly convex results
first draft
add abstract v1

--- 2020-03-20 source/main.tex
@@
 \usepackage{balance}
-
+\usepackage{comment}
 \newcommand{\eq}[1]{{Eq~(#1)}}
@@
 \newtheorem{theorem}{Theorem}
+\newtheorem{thm}{Theorem}
 \newtheorem{lemma}[theorem]{Lemma}
@@
 % \tableofcontents
-% \begin{abstract}
-% \end{abstract}
-
-%% Note for ICLR'20 paper
-% \input{note_sec_problem}
-% \input{note_sec_algorithm}
-% \input{note_sec_part}
-% \input{note_sec_q}
+\begin{abstract}
+Federated learning is a distributed learning paradigm that enables learning a
+model jointly over a large number of workers without sharing the private data
+on local workers. The characteristics of non-identically distributed data
+across the network and low device participation bring significant challenges
+in analyzing the convergence of federated learning algorithms. Prior works
+mainly analyze its convergence rate in limited scenarios.
+In this work, we provide a comprehensive analysis of the convergence rate of SGD and Nesterov SGD for general convex problems.
+Theoretically, we firstly show 
+that the convergence rates enjoy a linear speedup for
+all convex smooth problems for both algorithms.
+Furthermore, we propose a new accelerated SGD algorithm improving the convergence rate over SGD in the overparameterized setting. 
+Empirical evaluation of the algorithms in various settings match our theoretical
+results.
+\end{abstract}
 
 % converge proof
+% Introduction.
+
+% accelerated 
+% convex strongly 
+% smooth non-smooth. 
 \input{sec_problem}
+\input{sec_theorems}
+\newpage
+\input{sec_exp}
+\input{sec_exp_note}
+
+\newpage
+\appendix
 \input{sec_overview}
-\newpage
 % \input{sec_acceleratedFL}
-% \input{sec_cvxnonsm}
 
 % accelerated FL survey
+
+% paper
+
+% \begin{itemize}
+% 	\item On Federated Learning of Deep Networks from Non-IID Data: Parameter Divergence and the Effects of Hyperparametric Methods: empirical evaluation 
+% 	paper, the loss surface could be helpful.
+% 	\item On the Linear Speedup Analysis of Communication Efficient Momentum SGD for Distributed Non-Convex Optimization: Non-convex, smooth, accelerated SGD, Full device particaption.
+% \end{itemize}
 

### 2020-05-25 source/main.tex
add stronlgy convex and convex smooth results

@@
 in analyzing the convergence of federated learning algorithms. Prior works
-mainly analyze its convergence rate in limited scenarios.
-In this work, we provide a comprehensive analysis of the convergence rate of SGD and Nesterov SGD for general convex problems.
-Theoretically, we firstly show 
-that the convergence rates enjoy a linear speedup for
-all convex smooth problems for both algorithms.
-Furthermore, we propose a new accelerated SGD algorithm improving the convergence rate over SGD in the overparameterized setting. 
-Empirical evaluation of the algorithms in various settings match our theoretical
-results.
+have either analyzed the convergence with i.i.d. data, or with full device
+participation, and often with suboptimal rates. In this work, we provide a
+comprehensive analysis of the convergence rate of Federated Averaging (FedAvg)
+with SGD and Nesterov SGD updates for general convex problems with arbitrarily
+heterogeneous data. Firstly we show that the convergence rates enjoy a linear
+speedup in the number of participating workers for all convex smooth problems
+for both algorithms. Furthermore, we propose a new accelerated federated
+learning algorithm improving the convergence rate over FedAvg with SGD in the
+overparameterized setting. Empirical evaluation of the algorithms in various
+settings match our theoretical results.
 \end{abstract}
@@
 \input{sec_exp}
-\input{sec_exp_note}
+% \input{sec_exp_note}
 

### 2020-05-25 source/sec_exp.tex
add stronlgy convex and convex smooth results

--- 2020-05-22 source/sec_exp.tex
@@
 \begin{tabular}{cc}
-	\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupNodesT-min-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} & 
-\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupEpochsT-min-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} 
+	\includegraphics[width=0.45\textwidth]{fig/paper-stronglycvxsmthspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} & 
+\includegraphics[width=0.45\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsT-min-w8a-epsilon0131-reg1e-05.pdf} \\
 \end{tabular}
@@
 The linear speed up w.r.t the number of nodes and number of epochs.
-The synthetic dataset has $N=49749$ samples, evenly distributed on $1, 2, 4, 8, 16, 32, 64, 128, 256, 512, 1024$ devices.
-The figure shows the number of iterations/rounds needed to converge to $\epsilon-$accuracy.
+The synthetic dataset has $N=49749$ samples, evenly distributed on $1, 2, 4, 8, 16, 32$ devices.
+The figure shows the number of iterations/rounds needed to converge to $\epsilon-$accuracy, where $\epsilon=0.005$.
 The learning rate is decayed as the $\eta_t = \min(32, \frac{Nc}{1 + t})$, where we extensively search the best learning rate $c \in \{2^{-1}c, 2^{-2}c, c, 2c, 2^{2}c\}$ and for each configuration.
@@
 
+\begin{figure}[h!]
+\centering
+\begin{tabular}{cc}
+	\includegraphics[width=0.45\textwidth]{fig/paper-cvxsmoothspeedupNodesT-min-w8a-epsilon013-b4-reg0.pdf} & 
+\includegraphics[width=0.45\textwidth]{fig/paper-cvxsmoothspeedupEpochsT-min-w8a-epsilon013-b4-reg0.pdf} \\
+\end{tabular}
+	\caption{Logistic regression without regularization.
+The linear speed up w.r.t the number of nodes and number of epochs.
+The w8a dataset has $N=49749$ samples, evenly distributed on $1, 2, 4, 8, 16, 32, 64, 128, 256, 512, 1024$ devices.
+The figure shows the number of iterations needed to converge to $\epsilon-$accuracy, where $\epsilon=0.02$.
+The learning rate is decayed as the $\eta_t = \min(\eta_0, \frac{Nc}{1 + t})$, where we extensively search the best learning rate $c \in \{2^{-1}c, 2^{-2}c, c, 2c, 2^{2}c\}$, $\eta_0 \in \{1, 32\}$ and for each configuration.
+Batch size is 4 in this case.}
+\end{figure}
+
 \subsection{Linear regression}

### 2020-05-25 source/sec_exp.tex
add linear regression

@@
 \section{Numerical Experiments}
+
+We examine the practical speedup on a linear regression problem, 
+$ F(\vw) = \sum_{k=1}^N p_kF_k(\vw)$, the objective function on each local device $k$ is given by, $F_k(\vw) = \frac{1}{N_k} \sum_{i=1}^{N_k} (\vw^T\vx_i^k + b  - y_i^k)^2$, where we generated i.i.d.
+data $\vx_i^k$ for all devices.
+At each communication round, all devices participated
+in synchronization.
+
+\begin{figure}
+\centering
+	\begin{tabular}{ccc}
+	\includegraphics[width=0.33\textwidth]{fig/paper-linregressionspeedupEpochsT-min-linearregressionw8a-epsilon002-reg0.pdf} & \includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsT-min-w8a-epsilon0131-reg1e-05.pdf} &
+	\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupEpochsT-min-w8a-epsilon0134-reg0.pdf}
+\\
+\includegraphics[width=0.33\textwidth]{fig/paper-linregressionspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}
+&
+\includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} &
+\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupNodesT-min-w8a-epsilon0134-reg0.pdf} \\
+(a) Linear regression  & (b) Strongly convex objective & (c) Convex smooth objective
+	\end{tabular}
+
+\end{figure}
 
@@
 \begin{tabular}{cc}
-	\includegraphics[width=0.45\textwidth]{fig/paper-cvxsmoothspeedupNodesT-min-w8a-epsilon013-b4-reg0.pdf} & 
-\includegraphics[width=0.45\textwidth]{fig/paper-cvxsmoothspeedupEpochsT-min-w8a-epsilon013-b4-reg0.pdf} \\
+	\includegraphics[width=0.45\textwidth]{fig/paper-cvxsmoothspeedupNodesT-min-w8a-epsilon0134-reg0.pdf} & 
+\includegraphics[width=0.45\textwidth]{fig/paper-cvxsmoothspeedupEpochsT-min-w8a-epsilon0134-reg0.pdf} \\
 \end{tabular}

### 2020-05-26 source/sec_exp_note.tex
add linear regression
exp section version 1

@@
 \section{Numerical Experiments}
+\subsection{Strongly Convex}
+
+\begin{figure}[h!]
+\centering
+\begin{tabular}{cc}
+	\includegraphics[width=0.45\textwidth]{fig/paper-stronglycvxsmthspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} & 
+\includegraphics[width=0.45\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsT-min-w8a-epsilon0131-reg1e-05.pdf} \\
+\end{tabular}
+	\caption{Logistic regression with regularization $1e-5$.
+The linear speed up w.r.t the number of nodes and number of epochs.
+The synthetic dataset has $N=49749$ samples, evenly distributed on $1, 2, 4, 8, 16, 32$ devices.
+The figure shows the number of iterations/rounds needed to converge to $\epsilon-$accuracy, where $\epsilon=0.005$.
+The learning rate is decayed as the $\eta_t = \min(32, \frac{Nc}{1 + t})$, where we extensively search the best learning rate $c \in \{2^{-1}c, 2^{-2}c, c, 2c, 2^{2}c\}$ and for each configuration.
+Batch size is 4 in this case.}
+\end{figure}
+
+\subsection{Convex smooth objective}
+
+\begin{figure}[h!]
+\centering
+\begin{tabular}{cc}
+	\includegraphics[width=0.45\textwidth]{fig/paper-cvxsmoothspeedupNodesT-min-w8a-epsilon0134-reg0.pdf} & 
+\includegraphics[width=0.45\textwidth]{fig/paper-cvxsmoothspeedupEpochsT-min-w8a-epsilon0134-reg0.pdf} \\
+\end{tabular}
+	\caption{Logistic regression without regularization.
+The linear speed up w.r.t the number of nodes and number of epochs.
+The w8a dataset has $N=49749$ samples, evenly distributed on $1, 2, 4, 8, 16, 32, 64, 128, 256, 512, 1024$ devices.
+The figure shows the number of iterations needed to converge to $\epsilon-$accuracy, where $\epsilon=0.02$.
+The learning rate is decayed as the $\eta_t = \min(\eta_0, \frac{Nc}{1 + t})$, where we extensively search the best learning rate $c \in \{2^{-1}c, 2^{-2}c, c, 2c, 2^{2}c\}$, $\eta_0 \in \{1, 32\}$ and for each configuration.
+Batch size is 4 in this case.}
+\end{figure}
+
+\subsection{Linear regression}
 
@@
 \end{tabular}
-	\caption{The linear speed up w.r.t the number of nodes.
+	\caption{linear regression.
+The linear speed up w.r.t the number of nodes.
 The synthetic dataset has $6000$ samples, evenly distributed on $10, 20, 30, 40, 50, 60, 80, 100, 200, 300, 400, 500, 600, 750, 1000$ devices.

### 2020-05-26 source/sec_exp.tex
exp version 1.

--- 2020-05-23 source/sec_exp.tex
@@
 
-\subsection{Strongly Convex}
+In this section, we empirically examine the theoretical convergence rate of FedAve on various scenarios, including strongly convex function, convex smooth function, and overparameterized setting, as presented in previous sections.
 
-\begin{figure}[h!]
+\textbf{Setup.} Follow the experimental setting in~\cite{stich2018local}, We
+conduct experiments on both synthetic datasets and real-world dataset
+w8a~\cite{platt1998fast} $(d=300, n=49749)$. We consider the distributed
+objectives $F(\vw) = \sum_{k=1}^N p_kF_k(\vw)$, and the objective function on
+each local device $k$ includes three cases: 1) \textbf{Strongly convex
+objective}: the regularized binary logistic regression problem, $F_k(\vw) =
+\frac{1}{N_k} \sum_{i=1}^{N_k} \log( 1+ \exp(-y_i^k \vw^T\vx_i^k) + \lambda
+\|\vw\|^2$. The regularization parameter is set to $\lambda = 1/n \approx
+1e-5$. 2) \textbf{Convex smooth objective}: the binary logistic regression
+problem without regularization. $F_k(\vw) = \frac{1}{N_k} \sum_{i=1}^{N_k}
+\log( 1+ \exp(-y_i^k \vw^T\vx_i^k)$. 3) \textbf{Overparameterized setting}:
+the linear regression problem without adding noise to the label, $F_k(\vw) =
+\frac{1}{N_k} \sum_{i=1}^{N_k} (\vw^T\vx_i^k + b  - y_i^k)^2$.  In the first
+two cases, we use the w8a dataset to conduct empirical evaluation. 
+In the linear regression problem, we use features from w8a dataset and 
+generate a ground truth $w$ to label each instance, which satisfies the
+overparameterized setting.
+
+% where we generated i.i.d. data $\vx_i^k$ for all devices. At each communication round, all devices participated
+% in synchronization. 
+\textbf{Linear speedup.} To verify the linear speedup results in theorems {\color{red}add theorem id}, we evaluate the number of iterations needed to reach
+$\epsilon$-accuracy in three objectives. We initialize all runs with $w_0 = \textbf{0}_d$ and measure. For each configuration, we run three random seeds and report
+the best seed results for each configuration $(E, K)$. In each configure, we
+extensive search the stepsize from $\min(\eta_0, \frac{nc}{1 + t})$, where
+$\eta_0 \in \{0.1, 0.12, 1, 32 \}$ according to different problems and $c$ can
+take the values $c = 2^u \forall i \in \ZZ$. For the detailed experimental setup,
+please refer to the appendix. As the results shown in Figure~\ref{fig:speedup},
+the number of iterations to reach the target accuracy $\epsilon$ decreasing 
+as the number of workers increase.
+
+\textbf{The impact of $E$.} We further examine how does the number of epochs 
+affect the convergence. As shown in Figure~\ref{fig:e}, the number of iterations
+increases as the number of epochs increase, which slow the convergence in
+terms of gradient computation. However, it can save communication cost as
+the number of rounds decreased when the number of epochs increases.
+
+\begin{figure}
 \centering
-\begin{tabular}{cc}
-	\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupNodesT-min-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} & 
-\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupEpochsT-min-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} \\
-	\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupNodesRounds-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} & 
-\includegraphics[width=0.5\textwidth]{fig/E1-4-16speedupEpochsRounds-w8a-epsilon013-b4-reg1e-05-adapt0.pdf} \\
+	\begin{tabular}{ccc}
+	\hspace{-2em}\includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} &
+\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupNodesT-min-w8a-epsilon0134-reg0.pdf} &
+\includegraphics[width=0.33\textwidth]{fig/paper-linregressionspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
+(a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
 	\end{tabular}
-	\caption{Logistic regression with regularization $1e-5$.
-The linear speed up w.r.t the number of nodes and number of epochs.
-The synthetic dataset has $N=49749$ samples, evenly distributed on $1, 2, 4, 8, 16, 32, 64, 128, 256, 512, 1024$ devices.
-The figure shows the number of iterations/rounds needed to converge to $\epsilon-$accuracy.
-The learning rate is decayed as the $\eta_t = \min(32, \frac{Nc}{1 + t})$, where we extensively search the best learning rate $c \in \{2^{-1}c, 2^{-2}c, c, 2c, 2^{2}c\}$ and for each configuration.
-Batch size is 4 in this case.}
+\caption{The linear speedup convergence of FedAve w.r.t the number of nodes. }
+\label{fig:speedup}
 \end{figure}
 
-\subsection{Convex smooth objective}
-
-\subsection{Linear regression}
+\begin{figure}
+\centering
+	\begin{tabular}{ccc}
+	\hspace{-2em} \includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsT-min-w8a-epsilon0131-reg1e-05.pdf} &
+	\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupEpochsT-min-w8a-epsilon0134-reg0.pdf} & 
+	\includegraphics[width=0.33\textwidth]{fig/paper-linregressionspeedupEpochsT-min-linearregressionw8a-epsilon002-reg0.pdf} \\
+	\hspace{-2em} \includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsRounds-min-w8a-epsilon0131-reg1e-05.pdf} &
+	\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupEpochsRounds-min-w8a-epsilon0134-reg0.pdf} & 
+	\includegraphics[width=0.33\textwidth]{fig/paper-linregression-newspeedupEpochsRounds-min-linearregressionw8a-epsilon002-reg0.pdf} \\
+(a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
+	\end{tabular}
+\caption{The convergence of FedAve w.r.t the number of epochs. }
+\label{fig:e}
+\end{figure}

### 2020-05-26 source/sec_exp.tex
update figures

@@
 % in synchronization. 
-\textbf{Linear speedup.} To verify the linear speedup results in theorems {\color{red}add theorem id}, we evaluate the number of iterations needed to reach
+\textbf{Linear speedup.} To verify the linear speedup results in Theorem~\ref{th:strongcvxsmth}~\ref{th:cvxsmoth}~\ref{th:fedmasslr}, we evaluate the number of iterations needed to reach
 $\epsilon$-accuracy in three objectives. We initialize all runs with $w_0 = \textbf{0}_d$ and measure. For each configuration, we run three random seeds and report

### 2020-05-26 source/main.tex
add linear regression
abs
change to FedAvg

@@
 
-\title{Unified Convergence Analysis of FedAve on Non-IID Data}
+\title{Blessing of FedAvg's Linear Speedup in Federated Learning}
 
@@
 on local workers. The characteristics of non-identically distributed data
-across the network and low device participation bring significant challenges
+across the network and low device participation brings significant challenges
 in analyzing the convergence of federated learning algorithms. Prior works
@@
 % smooth non-smooth. 
+\input{sec_intro}
 \input{sec_problem}
-\input{sec_theorems}
-\newpage
+\input{sec_strongly_convex_smooth}
+\input{sec_convex_smooth}
+\input{sec_linear_regression}
 \input{sec_exp}
-% \input{sec_exp_note}
 
@@
 \appendix
-\input{sec_overview}
+\input{sec_appendix}
 % \input{sec_acceleratedFL}

### 2020-05-26 source/sec_cvxnonsm.tex
change cvx smooth assumption to variance
change to FedAvg

@@
 	Let Assumption~\ref{ass:subgrad2} hold, 
-	then the FedAve with partial device participation satisfies
+	then the FedAvg with partial device participation satisfies
 	\begin{align}
@@
 	& + \underbrace{\sum_{k=1}^N p_k \|\overline{\vw}_t - \vw_t^k\|^2}_{A_1} +  \sum_{k=1}^N p_k \eta_t^2 \|\EE \vg_{t,k}\|^2
-  - 2\eta_t\underbrace{\sum_{k=1}^N p_k (F_k(\vw_t^k) - F^*)}_{A_2} \label{eq:main2}
+  - 2\eta_t\underbrace{\sum_{k=1}^N p_k (F_k(\vw_t^k) - F^*)}_{A_2} \label{eq:cvxmain2}
 \end{align}
@@
  \geq & \sum_{k=1}^{N} P_{k}\left<\EE g_{t k}\left(\overline{\vw}_{t}\right), \vw_{t}^k-\overline{\vw}_{t}\right> + (F(\overline{w}_t) - F^*)\\
- \geq& -\frac{1}{2} \sum_{k=1}^{N} P_{k} (\eta_t \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \frac{1}{\eta_t} \|\vw_t^k - \overline{\vw}_t\|^2) + (F(\overline{w}_t) - F^*) \label{eq:a24}
+ \geq& -\frac{1}{2} \sum_{k=1}^{N} P_{k} (\eta_t \|\EE \vg_{t,k}(\overline{\vw}_t) \|^2 + \frac{1}{\eta_t} \|\vw_t^k - \overline{\vw}_t\|^2) + (F(\overline{w}_t) - F^*) \label{eq:cvxa24}
 \end{align}
@@
 
-Now we plug in \eq{\ref{eq:a24}} into \eq{\ref{eq:main2}}:
+Now we plug in \eq{\ref{eq:cvxa24}} into \eq{\ref{eq:cvxmain2}}:
 \begin{align}

### 2020-05-26 source/sec_exp.tex
change to FedAvg

@@
 
-In this section, we empirically examine the theoretical convergence rate of FedAve on various scenarios, including strongly convex function, convex smooth function, and overparameterized setting, as presented in previous sections.
+In this section, we empirically examine the theoretical convergence rate of FedAvg on various scenarios, including strongly convex function, convex smooth function, and overparameterized setting, as presented in previous sections.
 
@@
 	\end{tabular}
-\caption{The linear speedup convergence of FedAve w.r.t the number of nodes. }
+\caption{The linear speedup convergence of FedAvg w.r.t the number of nodes. }
 \label{fig:speedup}
@@
 	\end{tabular}
-\caption{The convergence of FedAve w.r.t the number of epochs. }
+\caption{The convergence of FedAvg w.r.t the number of epochs. }
 \label{fig:e}

### 2020-05-26 source/sec_theorems.tex
MaSS_linear_speedup
change to FedAvg

@@
 \end{table}
+
+Linear speedup of SGD and MaSS for linear regression, full participation: 
+\begin{align*}
+\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq C\cdot\exp(-\frac{t}{\sqrt{(\kappa_{N}+(E-1)^{2})(\tilde{\kappa}_{N}+(E-1)^{2})}})
+\end{align*}
+
+ which implies that for $E=1$, there is linear speedup, i.e. 
+\begin{align*}
+\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq C\cdot\exp(-\frac{t}{\sqrt{(\kappa_{N})(\tilde{\kappa}_{N})}})\leq C\cdot\exp(-\frac{Nt}{\sqrt{(\kappa_{1})(\tilde{\kappa})}})
+\end{align*}
+when $N$ satifies $N=O(\min(L_{1}/L,\tilde{\kappa}))$, and for $(E-1)^{2}\leq\min(\kappa_{N},\tilde{\kappa}_{N})$,
+\begin{align*}
+\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq C\cdot\exp(-\frac{Nt}{4\sqrt{\kappa_{1}\tilde{\kappa}}})
+\end{align*}
 
@@
 
-\section{The Convergence of FedAve for General Convex Functions}
-This section summarizes the convergence of FedAve.
-The updates of FedAve with partial device activation is given by: 
+\section{The Convergence of FedAvg for General Convex Functions}
+This section summarizes the convergence of FedAvg.
+The updates of FedAvg with partial device activation is given by: 
 \begin{align} 
@@
 
-\subsection{The Convergence of FedAve for Convex Functions}
-In this subsection, we discuss the convergence of FedAve for convex functions,
+\subsection{The Convergence of FedAvg for Convex Functions}
+In this subsection, we discuss the convergence of FedAvg for convex functions,
 we consider both smooth (Theorem~\ref{th:sgdcvxsmth}) and non-smooth (Theorem~\ref{th:sgdcvxnonsmoth})
@@
 
-\subsection{The Convergence of FedAve for Strongly Convex Functions}
+\subsection{The Convergence of FedAvg for Strongly Convex Functions}
 

### 2020-05-27 source/sec_intro.tex
intro v1.

% !TEX ROOT=./main.tex

\section{Introduction}
% Motivate FL from the security and privacy perspective. FedAvg
Federated learning (FL) is a distributed learning paradigm that can collaboratively solve a machine learning problem by leveraging the data
from a large amount of workers without compromising the users' (aka workers/local devices) privacy~\cite{kairouz2019advances}. 
Among all the leading algorithms in FL, Federated Averaging~\cite{mcmahan2016communication} (FedAvg)
is perhaps the most widely adopted optimization algorithm, which runs local
Stochastic Gradient Descent (SGD) updates on a subset of devices
and synchronize the local models once in a while. 
The empirical success of FedAvg and its variants have
attracted lots of efforts~\cite{li2018federated,stich2018local,khaled2019first,yu2019parallel,haddadpour2019convergence,li2019convergence,huo2020faster} on analyzing its convergence properties. 
There are few key challenges that differentiates the theoretical analysis
of FL from the tradition distribution optimization: 
1) The data is non-identically distributed across the workers, which means the
data in each local device cannot be regarded as samples drawn from
a same distribution. 
2) The workers are not active at every communication
round. In FL, the central server has no control over the local devices. 
It is more practical to assume only a subset of workers is active during
each communication round.

% The current gaps in FL.
However, most of prior works~\cite{li2018federated,stich2018local,khaled2019first,yu2019parallel,haddadpour2019convergence} either assume
the data is identically distributed or the all devices are active, which
violates the practical characteristic in FL. The most recent work~\cite{li2019convergence} firstly presents $O(1/T)$ convergence guarantee without 
making those unrealistic assumptions while their analysis focused 
on strongly convex case only and the relation between the number of active
workers and the convergence rate is not clearly discussed.

% Our contributions
Our work aims to systematically fill the gap in the convergence analysis
of FedAvg and its accelerated variants. 
Our theoretical results not only cover all general convex objectives, including strongly convex cases, convex smooth cases and convex non-smooth cases, but also provide tighter convergence guarantee all convex smooth objectives, which shows the convergence rate enjoys a linear speedup 
Furthermore, we studied a popular over-parameterized setting~\cite{liu2018accelerating} where the optimal solution can obtain zero training loss. In this scenario, we provide a novel accelerated FL algorithm improving
the convergence rate of FedAvg. Last but not least, we conduct extensive
evaluation on both synthetic and real-world dataset, which demonstrates that our theoretical indications is well-aligned with the empirical observations.

### 2020-05-27 source/sec_exp.tex
parital device participation

@@
 
+\textbf{Partial participations.} We verify the linear speedup in the
+partial participation settings, where we set only $50\%$ of devices
+are active. As the results shown in Figure~\ref{fig:partial}, the 
+FedAve enjoys linear speedup in various settings even with partial
+device participation.
+
 \begin{figure}
@@
 	\end{tabular}
-\caption{The linear speedup convergence of FedAvg w.r.t the number of nodes. }
+\caption{The linear speedup convergence of FedAvg w.r.t the number of workers. }
 \label{fig:speedup}
@@
 \end{figure}
+
+\begin{figure}
+\centering
+	\begin{tabular}{ccc}
+	\hspace{-2em}\includegraphics[width=0.33\textwidth]{fig/paper-partialstronglycvxsmthspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} &
+\includegraphics[width=0.33\textwidth]{fig/paper-partialcvxsmoothspeedupNodesT-min-w8a-epsilon0134-reg0.pdf} &
+\includegraphics[width=0.33\textwidth]{fig/paper-partiallinregressionspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
+(a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
+	\end{tabular}
+\caption{The linear speedup convergence of FedAvg w.r.t the number of active workers. }
+\label{fig:partial}
+\end{figure}

### 2020-05-27 source/sec_impact.tex
social impact

% !TEX ROOT=./main.tex

% Broader impact: Authors are asked to include a section in their submissions discussing the broader impact of their work, including possible societal consequences — -both positive and negative.
\section{Broader impact}
With the widespread use of mobile devices, the powerful sensors
collected enormous amount of data empowered with machine learning
solutions can transform people's life in many aspects. The proper
data integration can produce positive social
impact in terms of better personalization~\cite{fallah2020personalized}, more efficient information solicitation~\cite{chen2018federated}, more robust and stable service~\cite{47586}.

In the meanwhile, the increasing concerns of security and
privacy leaking from both large-companies and the entire society
present significant challenges of proper utilization of local
data. This trade-off between efficiency and privacy or security
lies in the central of many industrial applications, including 
but not limited to recommender systems~\cite{chen2018federated}, virtual assistants~\cite{lamautonomy}, and mobile keyboard prediction~\cite{47586}

In this work, we establish solid theoretical foundation for federated
learning that achieves efficiency and privacy at the same time. 
On one hand, the data is isolated in each local devices, thus prevent
the potential leaking of private information. On the otherhand, the
convergence rates improved substantially if the number of
active local devices increases. In conclusion, our work 
shows a faithful pathway to simultaneously secure people's privacy and improve
users' experience.

### 2020-05-27 source/main.tex
update nips template
social impact
final abstract submission

@@
 %     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2019
+% before loading neurips_2020
 
 % ready for submission
-\usepackage{neurips_2019}
+% \usepackage{neurips_2020}
 
@@
 % [preprint] option:
-%     \usepackage[preprint]{neurips_2019}
+%     \usepackage[preprint]{neurips_2020}
 
 % to compile a camera-ready version, add the [final] option, e.g.:
-     % \usepackage[final]{neurips_2019}
+%     \usepackage[final]{neurips_2020}
 
 % to avoid loading the natbib package, add option nonatbib:
-%     \usepackage[nonatbib]{neurips_2019}
+     \usepackage[nonatbib]{neurips_2020}
 
@@
 
-% \tableofcontents
 \begin{abstract}
-Federated learning is a distributed learning paradigm that enables learning a
-model jointly over a large number of workers without sharing the private data
-on local workers. The characteristics of non-identically distributed data
+Federated learning is a distributed learning paradigm that enables learning
+a model jointly over a large number of workers without sharing the private
+data on local workers. The characteristics of non-identically distributed data
 across the network and low device participation brings significant challenges
@@
 speedup in the number of participating workers for all convex smooth problems
-for both algorithms. Furthermore, we propose a new accelerated federated
-learning algorithm improving the convergence rate over FedAvg with SGD in the
-overparameterized setting. Empirical evaluation of the algorithms in various
-settings match our theoretical results.
+for both algorithms.
+Furthermore, in the overparamterized setting, we prove that the convergence under FedAvg is in fact exponential, also with a speedup factor linear in the number of workers.
+We also propose a new accelerated federated learning algorithm that further improves this convergence rate with provable guarantees in the linear regression setting.
+Empirical evaluation of the algorithms in various settings match our theoretical results.
 \end{abstract}
 
-% converge proof
-% Introduction.
-
-% accelerated 
-% convex strongly 
-% smooth non-smooth. 
 \input{sec_intro}
@@
 
+\input{sec_impact}
 \newpage
-\appendix
-\input{sec_appendix}
-% \input{sec_acceleratedFL}
-
-% accelerated FL survey
-
-% paper
-
-% \begin{itemize}
-% 	\item On Federated Learning of Deep Networks from Non-IID Data: Parameter Divergence and the Effects of Hyperparametric Methods: empirical evaluation 
-% 	paper, the loss surface could be helpful.
-% 	\item On the Linear Speedup Analysis of Communication Efficient Momentum SGD for Distributed Non-Convex Optimization: Non-convex, smooth, accelerated SGD, Full device particaption.
-% \end{itemize}
-
-% \newpage
-\small
 \bibliographystyle{plain}
@@
 
+\appendix
+\input{sec_appendix}
 \end{document}

### 2020-05-27 source/sec_linear_regression.tex
update figures
interpolation results and proofs updated

@@
 
-\section{Linear Regression}
+\section{Exponential Convergence of FedAvg in the Overparameterized Setting}
 
@@
 
-We first show that FedAvg with MaSS has exponential convergence for
+We first show that FedAvg with SGD has exponential convergence for
 linear regression when the global objective has 0 as its global minimum.
+
+\subsection{Exponential Convergence of FedAvg with SGD}
+
+In Belkin et al., the authors show that in the interpolation setting,
+SGD achieves exponential convergence with linear speedup batch size
+that is below some critical threshold. In this section, we extend
+their result to the federated learning setting and show that similar
+exponential convergence with linear speedup in the number of workers
+hold using FedAvg with SGD updates holds.
+
+We consider the standard setting of empirical risk minimization where
+$\ell(w;\xi_{k,j})$ is non-negative, $l$-smooth, and convex, and
+as before, each $F_{k}(w)=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\ell(w;\xi_{k,j})$
+is $L$-smooth and $\mu$-strongly convex. Note that $l\geq L$. This
+setting includes many important problems in practice, such as linear
+regression with a full rank sample covariance matrix. In the interpolation
+or overparameterized setting, we assume that there exists $w^{\ast}\in\arg\min_{w}\sum_{k=1}^{N}p_{k}F_{k}(w)$
+such that $\ell(w^{\ast};\xi_{k,j})=0$ for all $\xi_{k,j}$. We first
+present a general result on the exponential convergence with linear
+speedup in the number of workers, before giving a sharper analysis
+for linear regression problems. 
+\begin{theorem}
+	For the overparameterized setting with general strongly convex and
+	smooth objectives, FedAvg with local SGD updates and communication
+	every $E$ iterations with constant step size $\overline{\alpha}=\frac{1}{2E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})}$
+	gives the exponential convergence guarantee 
+	\begin{align*}
+	\mathbb{E}F(\overline{w}_{t}) & \leq\frac{L}{2}(1-\mu\overline{\alpha})^{t}\|w_{0}-w^{\ast}\|^{2}=O(\exp(-\frac{\mu}{2E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})}t)\cdot\|w_{0}-w^{\ast}\|^{2})
+	\end{align*}
+\end{theorem}
+%
+We see that when $l>L$, the speedup factor is on the order of $\frac{N}{E(l\nu_{\max}+L(N-\nu_{\min}))}/\frac{1}{l\nu_{\max}+L(1-\nu_{\min})}\approx\frac{Nl}{E(l+L(N-1))}=O(N/E)$
+for $N\leq\frac{l}{L}+1$, i.e. FedAvg with $N$ workers and communication
+every $E$ iterations provide an exponential convergence speedup factor
+of $O(N/E)$, for $N\leq\frac{l}{L}+1$.
+
+Next we show that specialized to the overparameterized linear regression
+setting, the exponential convergence bound can be further sharpened,
+with a similar linear speedup in the number of workers. In this setting,
+we have 
+\begin{align*}
+F(w) & =\sum_{k=1}^{N}p_{k}F_{k}(w)\\
+F_{k}(w) & =\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(w^{T}x_{k,j}-z_{k,j})^{2}
+\end{align*}
+and there exists $w^{\ast}$ such that $F_{k}(w^{\ast})\equiv0$
+for all $k$. 
+\begin{theorem}
+	(FedAvg with SGD, Linear Regression) For the overparamterized linear
+	regression problem, FedAvg with local SGD updates and communication
+	every $E$ iterations with constant step size $\overline{\alpha}=\frac{1}{2E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})}$
+	gives the exponential convergence guarantee 
+	\begin{align*}
+	\mathbb{E}F(\overline{w}_{t}) & \leq L(1-c\frac{N}{E\kappa})^{t}\|w_{0}-w^{\ast}\|^{2}
+	\end{align*}
+	where $\kappa=L/\mu$ and $c$ is some universal constant. 
+\end{theorem}
+
+Next we show FedAvg with MaSS has exponential convergence but with an improved rate.
 
@@
 \begin{theorem}
-	(FedAvg with MaSS, Linear Regression) Let $\tilde{\kappa}_{m}:=\tilde{\kappa}/m+(m-1)/m$,
-	and let the hyperparameters satisfy 
+	(FedAvg with MaSS, Linear Regression) For the overparamterized linear
+	regression problem, FedAvg with local MaSS updates and communication
+	every $E$ iterations with constant step size $\overline{\alpha}=\frac{1}{2E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})}$
+	gives the exponential convergence guarantee 
 	\begin{align*}
-	\eta^{k}(m)=\frac{1}{L_{m}}, & \alpha^{k}(m)=\frac{1}{\sqrt{\kappa_{m}\tilde{\kappa}_{m}}},\delta^{k}(m)=\frac{\eta^{k}}{\alpha^{k}\tilde{\kappa}_{m}}
+	\mathbb{E}F(\overline{w}_{t}) & \leq L(1-c\frac{N}{E\sqrt{\kappa\tilde{\kappa}}})^{t}\|w_{0}-w^{\ast}\|^{2}
 	\end{align*}
-	for all $k$. Let $\mathcal{I}_{E}=\{\ell E\mid\ell\in\mathbb{N}\}$
-	be the set of communication rounds, then the FedAvg algorithm with
-	MaSS and full participation solves the problem 
-	\begin{align*}
-	F(w) & =\sum_{k=1}^{N}p_{k}F_{k}(w)\\
-	F_{k}(w) & =\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(w^{T}x_{k,j}-z_{k,j})^{2}
-	\end{align*}
-	with exponential convergence
-	\begin{align*}
-	\mathbb{E}F(\overline{w}_{t})\leq\frac{L}{2}\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2} & \leq C\cdot\exp(-\frac{t}{\sqrt{(\kappa_{N}+(E-1)^{2})(\tilde{\kappa}_{N}+(E-1)^{2})}})
-	\end{align*}
-	where $w^{\ast}$ is such that $F(w^{\ast})=0$ and $C=\frac{L}{2}\cdot\frac{\alpha}{\delta}\sum_{k=1}^{N}p_{k}(\|v_{0}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta}{\alpha}\|w_{0}^{k}-w^{\ast}\|^{2})$.
+	where $\kappa=L/\mu$, $c$ is some universal constant, and $\tilde{\kappa}$
+	is the smallest positive number such that $\mathbb{E}\|\tilde{x}\|_{H^{-1}}^{2}\tilde{x}\tilde{x}^{T}\preceq\tilde{\kappa}H$. 
+	\label{th:fedmasslr}
+\end{theorem}
 
-In particular, this implies that for $E=1$, there is linear speedup, i.e. 
-	\begin{align*}
-	\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq C\cdot\exp(-\frac{t}{\sqrt{(\kappa_{N})(\tilde{\kappa}_{N})}})\leq C\cdot\exp(-\frac{Nt}{\sqrt{(\kappa_{1})(\tilde{\kappa})}})
-	\end{align*}
-
-	With partial participation of $K$ devices each communication round,
-	the convergence rate is also 
-	\begin{align*}
-	\mathbb{E}F(\overline{w}_{t})\leq\frac{L}{2}\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2} & \leq C\cdot(1+\frac{4}{K}\cdot(\frac{t}{E}-1))\cdot(1-\frac{1}{\sqrt{\kappa_{K}\tilde{\kappa}_{K}}})^{t}=O(\exp(-\frac{t}{\sqrt{(\kappa_{N}+(E-1)^{2})(\tilde{\kappa}_{N}+(E-1)^{2})}}))
-	\end{align*}
-
-\end{theorem}
+We immediately see that since $\tilde{\kappa}\leq\kappa$, FedAvg with MaSS provides an improvement over FedAvg with SGD in the the interpolation setting, while enjoying the same speedup factor in the number of workers.

### 2020-05-28 source/sec_intro.tex
Intro setup v0

@@
 of FedAvg and its accelerated variants. 
-Our theoretical results not only cover all general convex objectives, including strongly convex cases, convex smooth cases and convex non-smooth cases, but also provide tighter convergence guarantee all convex smooth objectives, which shows the convergence rate enjoys a linear speedup
+Our theoretical results not only cover all general convex objectives, including strongly convex cases, convex smooth cases and convex non-smooth cases, but also provide tighter convergence guarantee for all convex smooth objectives, which shows the convergence rate enjoys a linear speedup w.r.t.
+the number of workers.
 Furthermore, we studied a popular over-parameterized setting~\cite{liu2018accelerating} where the optimal solution can obtain zero training loss.

### 2020-05-28 source/sec_problem.tex
Intro setup v0

@@
 
-\section{Problem Setup}
+\section{Setup}
 
-\subsection{The Federated Learning Problem}
+% \subsection{The Federated Learning Problem}
 
-The federated learning problem, given local objectives $F_k$ and device weights $p_k$, is to minimize
-
+We study the federated learning problem
 \begin{align}
-	\min _{\mathbf{w}}\left\{F(\mathbf{w}) \triangleq \sum_{k=1}^{N} p_{k} F_{k}(\mathbf{w})\right\}
+	\min _{\mathbf{w}}\left\{F(\mathbf{w}) \triangleq \sum_{k=1}^{N} p_{k} F_{k}(\mathbf{w})\right\},
 	\label{eq:problem}
 \end{align}
+where $N$ is the number of local devices (users/nodes/workers) and $p_k$ is the $k$-th device's weight satisfying $p_k \geq 0$ and $\sum_{k=1}^N p_k = 1$. 
+In $k$-th local device, there are a finite number of $n_k$ data points:
+$\vx_k^1, \vx_k^2, \dots, \vx_k^{n_k}$.  
+The local objective $F_k(\cdot)$ is defined as:
+% \begin{align}
+$F_{k}(\mathbf{w}) \triangleq \frac{1}{n_{k}} \sum_{j=1}^{n_{k}} \ell\left(\mathbf{w} ; \vx_{k, j}\right)$,
+% \label{eq:localloss}
+% \end{align}
+where $\ell$ denotes a user-specified loss function.
+In federated learning, each device stores its local data, which gives rise to its own local objective $F_k$.
+Note that we do not make any assumption on the
+data distributions of each local device. The local minimum
+$\vw^{*}_k= \argmin_{\vw \in R^d} F_k(\vw)$ can be drifted far from
+the global minimum of \eq{\ref{eq:problem}}, which quantifies the 
+heterogeneity of the problem.
+In this paper, we systematically investigate the convergence results, 
+when the local objectives are convex smooth, strongly convex and convex but
+non-smooth.
 
-where $N$ is the number of devices, and device weights $p_k$ satisfy $p_k \geq 0$ and $\sum_{k=1}^N p_k = 1$. Suppose the $k$-th device
-holds the $n_k$ data points: $\xi_{k,1}, \xi_{k,2}, \dots, \xi_{k,n_k}$ . We assume that the local objective $F_k(\cdot)$ is defined by
-
-\begin{align}
-F_{k}(\mathbf{w}) \triangleq \frac{1}{n_{k}} \sum_{j=1}^{n_{k}} \ell\left(\mathbf{w} ; \xi_{k, j}\right)	
-\label{eq:localloss}
-\end{align}
-
-where $\ell$ is some loss function.
-Thus, in the federated learning setting, each device stores its local data, which gives rise to its own local objective $F_k$.
-A key difference with many problems in the distributed optimization setting is that data on local devices is heterogeneous, i.e.
+% A key difference with many problems in the distributed optimization setting is that data on local devices is heterogeneous, i.e.
 $F_k(w)$ may have different minimizers $w^{\ast}_k$.
@@
 
-\subsection{The FedAvg Algorithm}
-
-Despite the heterogeneity in local objectives, the federated learning
-problem can still be solved efficiently in a distributed manner with
-the simple Federated Average (FedAvg) algorithm, which consists of
-the iterations
-
+\subsection{The Federated Averaging (FedAvg) Algorithm}
+Although the local objectives are heterogeneous, federated learning
+problem can be solved efficiently in a distributed manner with
+the Federated Average (FedAvg) algorithm, which consists of
+the following procedures at $t$-th round: 
 \begin{align}

### 2020-05-28 source/sec_exp.tex
nesterov sgd linear regression results

@@
 \textbf{Partial participations.} We verify the linear speedup in the
-partial participation settings, where we set only $50\%$ of devices
+partial participation settings, where we set $50\%$ of devices
 are active. As the results shown in Figure~\ref{fig:partial}, the 
@@
 \end{figure}
+
+\begin{figure}
+\centering
+\begin{tabular}{ccc}
+ &  & 
+\includegraphics[width=0.33\textwidth]{fig/paper-lrnesterovspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
+(a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
+	\end{tabular}
+\caption{The linear speedup convergence of Accelerated FedAvg w.r.t the number of workers. }
+\label{fig:nesterov}
+\end{figure}

### 2020-05-28 source/sec_appendix_notations.tex
problem set up v1.

--- 2020-04-03 source/sec_problem.tex
@@
 
-\section{Problem Setting}
-
-\begin{align}
-	\min _{\mathbf{w}}\left\{F(\mathbf{w}) \triangleq \sum_{k=1}^{N} p_{k} F_{k}(\mathbf{w})\right\}
-	\label{eq:problem}
-\end{align}
-where $N$ is the number of devices, and $p_k$ is the weight of the k-th device
-such that $p_k \geq 0$ and $\sum_{k=1}^N p_k = 1$. Suppose the k-th device
-holds the nk training data: $x_{k,1}, x_{k,2}, \dots, x_{k,n_k}$ . The local
-$k \in S_t^K$ objective $F_k(\cdot)$ is defined by
-
-\begin{align}
-F_{k}(\mathbf{w}) \triangleq \frac{1}{n_{k}} \sum_{j=1}^{n_{k}} \ell\left(\mathbf{w} ; x_{k, j}\right)	
-\label{eq:localloss}
-\end{align}
-
-% The summary of all assumptions:
-
-% \begin{assumption}
-% $F_{1}, \cdots, F_{N}$ are all $L$-smooth: for all  $\mathbf{v}$  and $\mathbf{w}$, $F_{k}(\mathbf{v}) \leq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{L}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$.
-% \end{assumption}
-
-% \begin{assumption}
-% $F_1,\dots, F_N$ are all convex: for all $\vv$ and $\vw$, 
-% $F_k(\vv) \geq F_k(\vw)+(\vv -\vw)^T \grad F_k(\vw)$. \label{ass:cvx}
-% \end{assumption}
-% \begin{assumption}
-% $	F_{1}, \cdots, F_{N} \text { are all } \mu \text { -strongly convex: for all v and } \mathbf{w}, F_{k}(\mathbf{v}) \geq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{\mu}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$
-% \end{assumption}
-
-% \begin{assumption}
-% Let $\xi_{t}^{k}$ be sampled from the k-th device's local data uniformly at random. The variance of stochastic gradients in each device is bounded:
-% $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq \sigma_{k}^{2}$, for $k = 1,..., N$.
-% \end{assumption}
-
-% \begin{assumption}
-% The expected squared norm of stochastic gradients is uniformly bounded. i.e.,
-% $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)\right\|^{2} \leq G^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
-% \end{assumption}
-
-\section{Notations}
-
-Let $\vw_t^k$ be the model parameter maintained in the k-th device at the t-th step.
-$\cI_E$ is a set of global synchronization steps, i.e., $\cI_E = \{n E|n = 1, 2, 3,\dots \}$.
-If $t+1 \in \cI_E$, which means we communicate the server with (all) clients at time step $t+1$.
-$\vv^k_{t+1}$ is the immediate result of one step SGD update from $\vw^k_{t}$.
-$\xi_{t}^{k}$ denotes the data sampled from k-th device’s local data uniformly at random.
-Follow the common practice, we define two virtual sequences $\overline{\mathbf{v}}_{t}$ and $\overline{\mathbf{w}}_{t}$. For full device participation and $t \notin \cI_E$,
+\section{Additional Notations}
+In this section, we introduce additional notations that are used throughout
+the proof.
+Following common practice~\cite{stich2018local,li2019convergence}, we define two virtual sequences $\overline{\mathbf{v}}_{t}$ and $\overline{\mathbf{w}}_{t}$.
+For full device participation and $t \notin \cI_E$,
 $\ov{v}_t = \ov{w}_t =\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$.
-In partial participation, $t \in \cI_E$, $\ov{w}_t \neq \ov{v}_t$ since $\ov{v}_t=\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$ while $\sum_{k\in \cS_t}\mathbf{w}_{t}^{k}$.
+For partial participation, $t \in \cI_E$, $\ov{w}_t \neq \ov{v}_t$ since $\ov{v}_t=\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$ while $\ov{w}_t=\sum_{k\in \cS_t}\mathbf{w}_{t}^{k}$.
 However, we can
@@
 \end{align}
-and we use $\xi_t = \{\xi_t^k\}_{k=1}^N$ denotes samples at all devices at time step $t$.
-
-The updates of FedAve with partial device activation is given by: 
-\begin{align} 
-\mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\
-\sum_{k \in \cS_{t+1}} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}\end{array}\right.
-\end{align}
-
-In~\cite{li2019convergence}, two types of unbiased sampling strategies are considered in Lemma 5. 
-The sampling scheme I establishes $\cS_{t+1}$ by i.i.d. sampling the devices with replacement,
-in this case the upper bound of expected square norm of $\ov{w}_{t+1} - \ov{v}_{t+1}$ is given by:
-\begin{align}
-\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4}{K} \eta_t^2 E^2G^2
-\end{align}
-The sampling scheme II establishes $\cS_{t+1}$ by uniformly sampling all devices without
-replacement, in which we have the 
-\begin{align}
-\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4(N - K)}{K(N-1)} \eta_t^2 E^2G^2
-\end{align}
-We denote this upper bound as follows for concise presentation. 
-\begin{align}
-	\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2 \leq  \eta_t^2 C
-	\label{eq:partialsample}
-\end{align}
-
-\lkxcom{Questions and next steps}
-\begin{itemize}
-	\item Consider heterogeneity of data: with strongly convex and smooth, we can use optimality gap. with non-smooth, we can use weighted gradient diversity. 
-	\item Nonconvex imply strongly convex: KDD~\cite{zhuo2019federated} has constant learning rate, in our case, we have a diminishing learning rate. 
-	\item Quantify biased sampling scheme?
-	\item Constant learning rate in other non-strongly convex case. 
-\end{itemize}
+and $\xi_t = \{\xi_t^k\}_{k=1}^N$ denotes random samples at all devices at time step $t$.

### 2020-05-28 source/sec_problem.tex
problem set up v1.

@@
 % \begin{align}
-$F_{k}(\mathbf{w}) \triangleq \frac{1}{n_{k}} \sum_{j=1}^{n_{k}} \ell\left(\mathbf{w} ; \vx_{k, j}\right)$,
+$F_{k}(\mathbf{w}) \triangleq \frac{1}{n_{k}} \sum_{j=1}^{n_{k}} \ell\left(\mathbf{w} ; \vx_k^j\right)$,
 % \label{eq:localloss}
@@
 \subsection{The Federated Averaging (FedAvg) Algorithm}
-Although the local objectives are heterogeneous, federated learning
-problem can be solved efficiently in a distributed manner with
-the Federated Average (FedAvg) algorithm, which consists of
-the following procedures at $t$-th round: 
+We first introduce the standard Federated Average (FedAvg) algorithm.
+The FedAvg updates the model in each device by local Stochastic Gradient Descent (SGD) and send the latest model to the central server every $E$
+steps. The central server conducts a weighted average over the models
+received from active devices and broadcasts the latest averaged model to all devices.
+Formally, the updates of FedAvg at round $t$ is described as follows:
 \begin{align}
 \vv_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}g_{t,k}\\
-\vw_{t+1}^{k} & =\begin{cases}
-\vv_{t+1}^{k} & \text{if }t+1\notin\mathcal{I}_{E}\\
-\sum_{k=1}^{N}p_{k}\vv_{t+1}^{k} & \text{if }t+1\in\mathcal{I}_{E}
-\end{cases}
+\mathbf{w}_{t+1}^{k} &=\left\{
+\begin{array}{ll}
+\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\ 
+\sum_{k \in \cS_{t+1}} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}
+\end{array}\right.
 \end{align}
+where $\vw_t^k$ is the local model parameter maintained in the $k$-th device at the $t$-th iteration, $\vg_{t,k}:=\nabla F_{k}(\vw_{t}^{k},\xi_{t}^{k})$ is the stochastic gradient based on $\xi_{t}^{k}$, the data sampled from $k$-th device’s local data uniformly at random.
+$\mathcal{I}_{E}=\{E,2E,\dots\}$ is the set of global synchronization steps.
+We use $\mathcal{S}_{t+1}$ to represent the set of active devices at $t+1$.
 
-where $\vw_t^k$ is the local model parameter maintained in the $k$-th device during the $t$-th iteration, 
-\begin{align}
-g_{t,k} & :=\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})
-\end{align} 
-is the stochastic gradient based on $\xi_{t}^{k}$, the data sampled from k-th device’s local data uniformly at random. $\mathcal{I}_{E}=\{E,2E,\dots\}$
-is the set of global synchronization steps. Thus when $t+1 \notin \mathcal{I}_{E}$, each device updates its local parameter based on local
-stochastic gradient descent, and every $E$ iterations, parameters are aggregated
-across devices based on weights $p_{k}$ and then broadcast back to
-each local device.
+Since federated learning usually involves an enormous amount of 
+local devices, it is more realistic to consider only a subset of 
+local devices is active at each communication round. In this work,
+we consider both the case of \textbf{full participation} where the model is
+averaging over all devices at the communication round, i.e.,
+$\mathbf{w}_{t+1}^{k} = \sum_{k=1}^N p_k \mathbf{v}_{t+1}^{k}$, and
+the case of \textbf{partial participation} where $|\cS_{t+1}| < N$. 
+In the partial participation scenario, $\cS_{t+1}$ is obtained by two types of
+sampling schemes to simulate the practical scenarios~\cite{li2019convergence}. The first one samples
+each device with probability $p_k$, with replacement. The second one samples
+each device uniformly without replacement. Both of these sampling schemes
+guarantee that gradient updates in FedAvg are unbiased stochastic versions of
+updates in FedAvg with full participation, during the communication round.
 
-Communication and broadcasting of local parameters
-is costly, so the efficiency of the FedAvg algorithm requires that
-$E$ is relatively large, while at the same time ensuring convergence
-to a global minimizer of the global objective $\sum_{k}p_{k}F_{k}(w)$.
+% Thus when $t+1 \notin \mathcal{I}_{E}$, each device updates its local parameter based on local
+% stochastic gradient descent, and every $E$ iterations, parameters are aggregated
+% across devices based on weights $p_{k}$ and then broadcast back to
+% each local device.
 
-In practical applications, it may not be possible to collect local
-parameters from \emph{all }devices at a specified communication time.
-Thus the above \emph{full participation }FedAvg algorithm may not
-be feasible. Instead, a more practical algorithm is to allow only
-a random subset $\mathcal{S}_{t+1}$ of all devices to participate in parameter aggregation, while
-broadcasting the aggregated parameter to all devices. This gives rise
-to the \emph{partial participation }FedAvg algorithm.
+% Communication and broadcasting of local parameters
+% is costly, so the efficiency of the FedAvg algorithm requires that
+% $E$ is relatively large, while at the same time ensuring convergence
+% to a global minimizer of the global objective $\sum_{k}p_{k}F_{k}(w)$.
 
-The updates of FedAvg with partial device activation is given by: 
-\begin{align} 
-\mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\
-\sum_{k \in \cS_{t+1}} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}\end{array}\right.
-\end{align}
+% In practical applications, it may not be possible to collect local
+% parameters from \emph{all }devices at a specified communication time.
+% Thus the above \emph{full participation }FedAvg algorithm may not
+% be feasible. Instead, a more practical algorithm is to allow only
+% a random subset $\mathcal{S}_{t+1}$ of all devices to participate in parameter aggregation, while
+% broadcasting the aggregated parameter to all devices. This gives rise
+% to the \emph{partial participation }FedAvg algorithm.
 
-where $\mathcal{S}_{t+1}$ is obtained using some sampling scheme.
+% The updates of FedAvg with partial device activation is given by: 
+% \begin{align} 
+% \mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\
+% \sum_{k \in \cS_{t+1}} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}\end{array}\right.
+% \end{align}
+
+% where $\mathcal{S}_{t+1}$ is obtained using some sampling scheme.
 We consider two types of common unbiased sampling strategies.
@@
 
-\subsection{Notation and Assumptions}
+\subsection{Assumptions}
 
-Following common practice, we define two virtual sequences $\overline{\mathbf{v}}_{t}$ and $\overline{\mathbf{w}}_{t}$. For full device participation and $t \notin \cI_E$,
-$\ov{v}_t = \ov{w}_t =\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$.
-For partial participation, $t \in \cI_E$, $\ov{w}_t \neq \ov{v}_t$ since $\ov{v}_t=\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$ while $\ov{w}_t=\sum_{k\in \cS_t}\mathbf{w}_{t}^{k}$.
-However, we can
-set unbiased sampling strategy such that $ \EE_{\cS_t} \ov{w}_t = \ov{v}_t$.
-$\overline{\mathbf{v}}_{t+1}$ is one-step SGD from $\overline{\mathbf{w}}_{t}$. 
-\begin{align}
-\overline{\mathbf{v}}_{t+1}=\overline{\mathbf{w}}_{t}-\eta_{t} \mathbf{g}_{t}	\label{eq:vbar}
-\end{align}
-% $t+1 \in \cI_E$, we can fetch $\overline{\vw}_{t+1}$, we can communicate $\overline{\vw}_{t+1}$ to all devices.
-where $\vg_{t} = \sum_{k=1}^{N} p_{k} \vg_{t,k} $ is one-step stochastic gradient, averaged over all devices. 
-\begin{align}
-\vg_{t,k} \left\{\begin{array}{ll} 
- = \nabla F_{k}\left(\mathbf{w}_{t}^{k},\xi_{t}^{k} \right)  &  \text{smooth}\\
- \in \partial F_{k}\left(w_{t}^{k}, \xi_{t}^{k}\right)  & \text{non-smooth}
- \end{array}\right.
-\end{align}
-Similarly, we denote the expected one-step gradient $\ov{g}_{t}= \EE_{\xi_t}[\vg_t] = \sum_{k=1}^{N} p_{k} \EE_{\xi_{t}^{k}} \vg_{t,k}$, where
-\begin{align}
-\EE_{\xi_{t}^{k}} \vg_{t,k}  \left\{\begin{array}{ll} 
- = \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)  &  \text{smooth}\\
- \in \partial F_{k}\left(w_{t}^{k}\right)  & \text{non-smooth}
- \end{array}\right.
-\end{align}
-and $\xi_t = \{\xi_t^k\}_{k=1}^N$ denotes samples at all devices at time step $t$.
-
-We will require some standard assumptions on the local objectives and their gradients.
+We make the following standard assumptions on the objective function $F_1,\dots, F_N$.
+Assumption~\ref{ass:boundedvariance} and Assumption~\ref{ass:subgrad2} have been made by many prior works in federated learning~\cite{yu2019parallel,li2019convergence,stich2018local}.
+Assumption~\ref{ass:lsmooth} and Assumption~\ref{ass:stroncvx} are common assumptions satisfied by a range of popular objective functions, such as $l-2$
+norm regularized logisitc regressions, logistic regressions and cross-entropy loss functions.
 
@@
 	for $k=1, \cdots, N$ and any $\mathbf{w}_{t}^{k}$.
+	\label{ass:boundedvariance}
 \end{assumption}
-
 \begin{assumption}[Expected square norm of gradient]
@@
 \end{assumption}
-
 \begin{assumption}[L-smooth]
@@
 \end{assumption}
-
 \begin{assumption}[Strongly-convex]
@@
 \end{assumption}
-
-Even though Assumption \ref{ass:subgrad2} also implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$, which further implies $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq G^2$, the two assumptions are stated separately because it makes explicit the dependence of the optimality gap on different terms.
+Assumption~\ref{ass:lsmooth} is made to achieve linear speedup 
+of convergence rate with respect to the number of workers, though it
+is not necessary for FedAvg and its variants to convergence.
+Assumption \ref{ass:subgrad2} implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$, which further implies $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq G^2$, the two assumptions are stated separately because it makes explicit the dependence of the optimality gap on different terms.

### 2020-05-28 source/sec_overparameterized.tex
changed file name

% !TEX ROOT=./main.tex

\section{Exponential Convergence of FedAvg in the Overparameterized Setting}

So far we have shown $O(1/\sqrt{NT})$ and $O(1/NT)$ convergence
rates for FedAvg with local SGD and Nesterov udpates in the convex
and strongly convex settings, respectively. In this section, we turn
to the special setting of overparameterized problems, where a non-negative
loss function admits a zero-loss global minimizer. Such problems are
very common in machine learning systems where the number of parameters
far exceeds the number of data points, and perfect fitting is possible.
There have been a line of recent works {[}CITE{]} showing that SGD
and accelerated methods achieve exponential convergence in this case,
thanks to the property of ``automatic variance reduction'' property
{[}CITE{]}. The natural question is whether such a result also holds
in the federated learning setting. We first show that this is indeed
the case and establish the exponential (linear) convergence of FedAvg
with local SGD updates with constant step size for general strongly
convex and smooth overparameterized problems. In addition, we show
that the convergence rate speeds up linearly in the number of workers
$N$ when $N$ is below some problem-dependent threhold, while decreases
with $E$ through $1/E$. To our knowledge, this is the first result
on the exponential convergence of FedAvg algorithms in the interpolation
setting with linear speedup in the number of workers and explicit
dependence on the communication interval $E$. We next sharpen this
exponential convergence rate in the special case of linear regression,
and show that it is $O(\exp(-\frac{NT}{E\kappa_{1}}))$, for $\kappa_{1}$
an appropriately defined condition number of the system. Lastly, we
turn to the question of whether FedAvg with momentum-based local updates
can outperform FedAvg with SGD updates. In contrast to the gradient
descent setting, in the single-agent stochastic gradient setting,
Nesterov and Heavy Ball updates are known to fail to accelerate over
SGD, both in the overparameterized setting and standard convex setting{[}CITE{]}.
Thus in general we cannot hope to obtain acceleration results for
the FedAvg algorithm with Nesterov and Heavy Ball updates. On the
other hand, in the overparameterized setting, {[}CITE Belkin{]} introduced
the MaSS algorithm, which is a modification of the Nesterov update,
where in the update a non-negative multiple of the gradient is added
to the Nesterov parameter update to correct for the ``over-descent''
of the Nesterov update. For quadratic objectives, the authors show
that the MaSS algorithm is able to achieve acceleration over the exponential
convergence of SGD both theoretically and empirically. In the last
part of this section, we introduce a new FedAvg algorithm with momentum
updates by adapting the MaSS algorithm to the federated learning setting.
For this new FedAvg algorithm with local MaSS updates, we show that
it achieves exponential convergence for overparameterized quadratic
problems with rate $O(\exp(-\frac{NT}{E\sqrt{\kappa_{1}\tilde{\kappa}}}))$,
where $\tilde{\kappa}$ is the ``statistical condition number''
introduced in {[}CITE{]} that satisfies $\tilde{\kappa}\leq\kappa_{1}$.
Thus the FedAvg algorithm with local MaSS updates preserves the linear
speedup property in the number of workers and the $1/E$ dependence
on $E$, while achieving a speedup of factor $\sqrt{\frac{\kappa_{1}}{\tilde{\kappa}}}$
over FedAvg with local SGD updates.

\subsection{Exponential Convergence of FedAvg with local SGD: Overparameterized
	Strongly Convex Smooth Objective}

In {[}CITE Belkin{]}, the authors show that in the overparameterized
setting, SGD achieves exponential convergence with linear speedup
in batch size that is below some critical threshold. In this section,
we extend their result to the federated learning setting and show
that a similar exponential convergence with linear speedup in the
number of workers using FedAvg with local SGD updates holds.

Recall the federated learning problem 
\begin{align*}
\min_{w}\sum_{k=1}^{N}p_{k}F_{k}(w)\\
F_{k}(w)=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\ell(w;\xi_{k,j})
\end{align*}
In this section, we consider the standard ERM setting where each
$\ell(w;\xi_{k,j})$ is non-negative, $l$-smooth, and convex, and
as before, we assume that each $F_{k}(w)=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\ell(w;\xi_{k,j})$
is $L$-smooth and $\mu$-strongly convex. Note that $l\geq L$. This
setting includes many important problems in practice, such as linear
regression with a full rank sample covariance matrix. In the overparameterized
setting, we assume that there exists $w^{\ast}\in\arg\min_{w}\sum_{k=1}^{N}p_{k}F_{k}(w)$
such that $\ell(w^{\ast};\xi_{k,j})=0$ for all $\xi_{k,j}$. Given
these assumptions, we first show that FedAvg with local SGD updates
and communication every $E$ iterations achieves exponential convergence
with linear speedup in the number of workers.
\begin{theorem}
	For the overparameterized setting with general strongly convex and
	smooth objectives, FedAvg with local SGD updates and communication
	every $E$ iterations with constant step size $\overline{\alpha}=\frac{1}{4E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})}$
	gives the exponential convergence guarantee 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{t}) & \leq\frac{L}{2}(1-\mu\overline{\alpha})^{t}\|w_{0}-w^{\ast}\|^{2}=O(L\exp(-\frac{\mu}{4E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})}t)\cdot\|w_{0}-w^{\ast}\|^{2})
	\end{align*}
\end{theorem}
%
\begin{remark}
	We see that when $l>L$, the speedup factor is on the order of $\frac{N}{E(l\nu_{\max}+L(N-\nu_{\min}))}/\frac{1}{l\nu_{\max}+L(1-\nu_{\min})}\approx\frac{Nl}{E(l+L(N-1))}=O(N/E)$
	for $N\leq\frac{l}{L}+1$, i.e. FedAvg with $N$ workers and communication
	every $E$ iterations provides an exponential convergence speedup
	factor of $O(N/E)$, for $N\leq\frac{l}{L}+1$. When $N$ is above
	this threhold, however, the speedup is almost constant in the number
	of woerkers. Our bound also illustrates the tradeoff between increasing
	the number of workers and the communication latency. When the cost
	of communication is high, i.e. it is costly to aggregate local parameters
	and broadcast the average to all active devices, increasing the number
	of workers by some factor achieves the same order of speedup as decreasing
	the communication lag by the same factor. 
\end{remark}
In the next part, we specialize to the problem of linear regression,
and demonstrate how the above bound can be sharpened to explicitly
depend on the condition number of the system. This result also paves
the way for the analysis of a new FedAvg with local momentum-based
updates in the following part.

\subsection{Overparameterized Quadratic Problems}

We now consider the federated learning problem with overparameterized
quadratic objectives. In other words, the local device objectives
are given by the sum of squares. 
\begin{align*}
F_{k}(w) & =\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(w^{T}x_{k,j}-z_{k,j})^{2}
\end{align*}
and there exists $w^{\ast}$ such that $F(w^{\ast})=\sum_{k}p_{k}F_{k}(w^{\ast})\equiv0$.
In this section, we gather some necessary definitions and observations
useful in the analysis of exponential convergence of FedAvg algorithms.

Define the local Hessian matrix as 
\begin{align*}
H^{k} & :=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}x_{k,j}(x_{k,j})^{T}
\end{align*}
and the averaged Hessian matrix as 
\begin{align*}
H & :=\sum_{k=1}^{N}p_{k}H^{k}
\end{align*}

In general $H$ has zero eigevalues. However, because the null space
of $H$ and range of $H$ are orthogonal, in our subsequence analysis
it suffices to project $\overline{w}_{t}-w^{\ast}$ onto the range
of $H$, thus we may restrict to the non-zero eigenvalue of $H$.

We can use $w^{\ast T}x_{k,j}-z_{k,j}\equiv0$ to rewrite the local
objectives 
\begin{align*}
F_{k}(w) & =\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(w^{T}x_{k,j}-z_{k,j}-(w^{\ast T}x_{k,j}-z_{k,j}))^{2}\\
& =\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}((w-w^{\ast})^{T}x_{k,j})^{2}=\frac{1}{2}\langle w-w^{\ast},H^{k}(w-w^{\ast})\rangle=\frac{1}{2}\|w-w^{\ast}\|_{H^{k}}^{2}
\end{align*}
so that 
\begin{align*}
F(w) & =\sum_{k=1}^{N}p_{k}F_{k}(w)=\sum_{k=1}^{N}p_{k}\frac{1}{2}(w-w^{\ast})^{T}H^{k}(w-w^{\ast})=\frac{1}{2}(w-w^{\ast})^{T}H(w-w^{\ast})=\frac{1}{2}\|w-w^{\ast}\|_{H}^{2}
\end{align*}

Let $x_{t}^{k}$ be the stochastic sample on the $k$th device during
iteration $t$, and define $\tilde{H}_{t}^{k}:=x_{t}^{k}(x_{t}^{k})^{T}$
as the stochastic Hessian matrix corresponding to the sample $x_{t}^{k}$.
Note that $\mathbb{E}\tilde{H}_{t}^{k}=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}x_{k,j}(x_{k,j})^{T}=H^{k}$
so that $\tilde{H}_{t}^{k}$ is an unbiased estimate of $H^{k}$.
Moreover, we have
\begin{align*}
g_{t,k} & =\nabla F_{k}(w_{t}^{k},x_{t}^{k})=\tilde{H}_{t}^{k}(w_{t}^{k}-w^{\ast})\\
g_{t} & =\sum_{k=1}^{N}p_{k}\nabla F_{k}(w_{t}^{k},x_{t}^{k})=\sum_{k=1}^{N}p_{k}\tilde{H}_{t}^{k}(w_{t}^{k}-w^{\ast})
\end{align*}

Define $l$ to be the smallest positive number such that $\mathbb{E}\|x_{t}^{k}\|^{2}$$x_{t}^{k}$($x_{t}^{k})^{T}\preceq lH^{k}$
for all $k,t$. Note that $l\leq\max_{k=1,\dots,N}\max_{1\leq j\leq n_{k}}\|x_{k,j}\|^{2}$,
i.e. $l$ is upper bounded by the maximum norm squared among all feature
vectors on all devices.

Let $L^{k}$ and $\mu^{k}$ to be the largest and smallest non-zero
eigenvalues of $H^{k}$, and $L:=\max_{k}L^{k}$, $\mu:=\min_{k}\mu^{k}$.
Define $\kappa_{1}:=l/\mu$ and $\kappa:=L/\mu$.

Following {[}CITE: Liu\&Belkin and Jain et al{]}, we define the statistical
condition number $\tilde{\kappa}^{k}$ as the smallest positive real
number such that 
\begin{align*}
\mathbb{E}\left[\langle x_{t}^{k}(H^{k})^{-1},x_{t}^{k}\rangle x_{t}^{k}(x_{t}^{k})^{T}\right] & \preceq\tilde{\kappa}^{k}H^{k}
\end{align*}
and define $\tilde{\kappa}:=\max_{k}\tilde{\kappa}^{k}$. The condition
numbers $\kappa_{1}$ and $\tilde{\kappa}$ are important in the characterization
of convergence rates for FedAvg algorithms.

Note that $\kappa_{1}>\kappa$ and $\kappa_{1}>\tilde{\kappa}$.

\subsection{Exponential Convergence of FedAvg with local SGD: Over-parameterized
	Quadratic Objective}

Next we show that specialized to overparameterized quadratic objectives,
the exponential convergence bound can be further sharpened, with a
similar linear speedup in the number of workers. Recall that in this
setting, we have 
\begin{align*}
F(w) & =\sum_{k=1}^{N}p_{k}F_{k}(w)\\
F_{k}(w) & =\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(w^{T}x_{k,j}-z_{k,j})^{2}
\end{align*}
and there exists $w^{\ast}$ such that $F_{k}(w^{\ast})\equiv0$
for all $k$. 
\begin{theorem}
	For the overparamterized linear regression problem, FedAvg with local
	SGD updates and communication every $E$ iterations with constant
	step size $\alpha_{t}=\frac{1}{4E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})}$
	gives the exponential convergence guarantee 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{t}) & \leq L(1-\frac{N}{4E(\nu_{\max}\kappa_{1}+(N-\nu_{\min}))})^{t}\|w_{0}-w^{\ast}\|^{2}
	\end{align*}
	where $\kappa_{1}=l/\mu$. 
\end{theorem}
%
We see that when $N=O(\kappa_{1})$, the convergence rate is $O((1-\frac{N}{E\kappa_{1}})^{t})=O(\exp(-\frac{N}{E\kappa_{1}}t))$,
which exhibits linear speedup in the number of workers, as well as
a $1/\kappa_{1}$ dependence on the condition number $\kappa_{1}$.
In the last part of this section, we will demonstrate that a new FedAvg
algorithm based on a local momentum update {[}CITE Belkin{]} improves
upon this dependence to $1/\sqrt{\kappa_{1}\tilde{\kappa}}$.

 \subsection{Exponential Convergence of FedAvg with local MaSS: Over-parameterized
 	Quadratic Objective}

 Following {[}CITE{]}, we propose the FedAvg algorithm with local MaSS
 updates: 
 \begin{align*}
 w_{t+1}^{k} & =\begin{cases}
 u_{t}^{k}-\eta_{1}^{k}g_{t,k} & \text{if }t+1\notin\mathcal{I}_{E}\\
 \sum_{k=1}^{N}p_{k}\left[u_{t}^{k}-\eta_{1}^{k}g_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E}
 \end{cases}\\
 u_{t+1}^{k} & =w_{t+1}^{k}+\gamma^{k}(w_{t+1}^{k}-w_{t}^{k})+\eta_{2}^{k}g_{t,k}
 \end{align*}
 where we note that the natural parameter is $w_{t}$, while $u_{t}$
 is an auxiliary parameter, which we initialize to be $u_{0}^{k}$,
 $g_{t,k}=\nabla F_{k}(u_{t}^{k},x_{t}^{k})$ is the stochastic gradient
 and $g_{t}=\sum_{k=1}^{N}p_{k}g_{t,k}$ is the averaged stochastic
 gradient. When $\eta_{2}^{k}\equiv0$, this reduces to the FedAvg
 algorithm with Nesterov updates.

 Also note that the update can equivalently be written as 
 \begin{align*}
 v_{t+1}^{k} & =(1-\alpha^{k})v_{t}^{k}+\alpha^{k}u_{t}^{k}-\delta^{k}g_{t,k}\\
 w_{t+1}^{k} & =\begin{cases}
 u_{t}^{k}-\eta^{k}g_{t,k} & \text{if }t+1\notin\mathcal{I}_{E}\\
 \sum_{k=1}^{N}p_{k}\left[u_{t}^{k}-\eta^{k}g_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E}
 \end{cases}\\
 u_{t+1}^{k} & =\frac{\alpha^{k}}{1+\alpha^{k}}v_{t+1}^{k}+\frac{1}{1+\alpha^{k}}w_{t+1}^{k}
 \end{align*}
 where there is a bijection between the parameters 
 \begin{align*}
 \frac{1-\alpha^{k}}{1+\alpha^{k}} & =\gamma^{k}\\
 \eta^{k} & =\eta_{1}^{k}\\
 \frac{\eta^{k}-\alpha^{k}\delta^{k}}{1+\alpha^{k}} & =\eta_{2}^{k}
 \end{align*}
 and we further introduce an auxiliary parameter $v_{t}^{k}$, which
 is initialized at $v_{0}^{k}$. We also note that when $\delta^{k}=\frac{\eta^{k}}{\alpha^{k}}$,
 the update reduces to the Nesterov accelerated SGD. This version of
 the FedAvg algorithm with local MaSS updates is used for analyzing
 the exponential convergence.

 As before, define the virtual sequences $\overline{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$,
 $\overline{v}_{t}=\sum_{k=1}^{N}p_{k}v_{t}^{k}$, $\overline{u}_{t}=\sum_{k=1}^{N}p_{k}u_{t}^{k}$,
 and $\overline{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$. We have
 $\mathbb{E}g_{t}=\overline{g}_{t}$ and $\overline{w}_{t+1}=\overline{u}_{t}-\eta_{t}g_{t}$,
 $\overline{v}_{t+1}=(1-\alpha^{k})\overline{v}_{t}+\alpha^{k}\overline{w}_{t}-\delta^{k}g_{t}$,
 and $\overline{u}_{t+1}=\frac{\alpha^{k}}{1+\alpha^{k}}\overline{v}_{t+1}+\frac{1}{1+\alpha^{k}}\overline{w}_{t+1}$.

 We now present the exponential convergence result in overparameterized
 quadratic problems using FedAvg with local MaSS updates. 
 \begin{theorem}
 	(FedAvg with MaSS, Linear Regression) For the overparamterized quadratic
 	problem, FedAvg with local MaSS updates and communication every $E$
 	iterations with constant step sizes 
 	\begin{align*}
 	\eta_{t}=\frac{1}{4E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})}, & \alpha_{t}=\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}},\delta_{t}=\frac{\eta_{t}}{\alpha_{t}\tilde{\kappa}}
 	\end{align*}
 	gives the exponential convergence guarantee 
 	\begin{align*}
 	\mathbb{E}F(\overline{w}_{t}) & \leq C(1-\frac{1}{4E}\frac{N}{l\nu_{\max}\sqrt{\kappa_{1}\tilde{\kappa}}+(N-\nu_{\min})})^{t}\|w_{0}-w^{\ast}\|^{2}
 	\end{align*}
 \end{theorem}
 Compared to the convergence rate of FedAvg with local SGD updates,
 we see that when $N=O(\sqrt{\kappa_{1}\tilde{\kappa}})$, the convergence
 rate is $O((1-\frac{N}{E\sqrt{\kappa_{1}\tilde{\kappa}}})^{t})=O(\exp(-\frac{N}{E\sqrt{\kappa_{1}\tilde{\kappa}}}t))$
 as opposd to $O(\exp(-\frac{N}{E\kappa_{1}}t))$. Since $\kappa_{1}\geq\tilde{\kappa}$,
 this implies a speedup factor of $\sqrt{\frac{\kappa_{1}}{\tilde{\kappa}}}$
 for the FedAvg with local MaSS updates. On the other hand, the same
 linear speedup in the number of workers holds for $N$ in a smaller
 range of values.

### 2020-05-29 source/sec_strongly_convex_smooth.tex
update figures
updated strongly convex smooth section

@@
 
-In this section, we show that FedAvg with SGD and Nesterov Accelerated SGD has $O(1/NT)$ rate under $\mu$-strong
-convexity and $L$-smoothness. The result improves on previous results of
-$O(1/T)$ with a linear speedup in the number of devices $N$.
-
-\subsection{SGD}
-
+In this section, we show that FedAvg with local SGD has $O(1/NT)$ convergence
+rate for $\mu$-strongly convex and $L$-smooth objectives. The result
+improves on the $O(1/T)$ result of {[}ICLR{]} with a linear speedup
+in the number of devices $N$. Moreover, it makes a distinction on
+the choice of $E$ with partial and full participation. With full
+participation, $E$ can be chosen as large as $O(\sqrt{\frac{T}{N}})$
+without degrading the linear speedup in the number of workers. On
+the other hand, with partial participation, $E$ must be $O(1)$. 
 \begin{theorem}
@@
 	$k$. Let $\kappa=\frac{L}{\mu}$, $\gamma=\max\{8\kappa-1,E\}$ where
-	$E$ is the communication interval, and learning rates 
+	$E$ is the communication delay, and learning rates 
 	\begin{align*}
@@
 	so that $\alpha_{t}\leq2\alpha_{t+E}$, and where $0\leq c\leq1$
-	is small enough such that
-	$\alpha_{t} \leq\frac{1}{4L}$ and 
-	$\alpha_{t} \leq\frac{1}{N}$ for all $t\geq0$.
+	is such that $\alpha_{t}\leq\frac{1}{8L}$ for all $t\geq0$.
 
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq C\frac{\kappa}{\gamma+T}\cdot(8E^{2}LG^{2}/T+\frac{1}{N}\nu_{max}^{2}\sigma^{2})
-	\end{align*}
-	so that as long as $E=O(\sqrt{\frac{N}{T}})$, 
-	\begin{align*}
-	\mathbb{E}F(w_{T})-F^{\ast} & \leq C(8LG^{2}+\nu_{max}^{2}\sigma^{2})\frac{\kappa}{N(\gamma+T)}
+	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq C\cdot(\frac{\kappa\nu_{max}^{2}\sigma^{2}}{NT}+\frac{5\kappa E^{2}LG^{2}}{T^{2}})
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq C\frac{\kappa}{\gamma+T}\cdot(8E^{2}LG^{2}/T+\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{4}{K}E^{2}G^{2})
-	\end{align*}
-	so that if $E=O(1)$, 
-	\begin{align*}
-	\mathbb{E}F(\overline{w}_{T})-F^{\ast} & \leq C(4E^{2}G^{2}+\nu_{max}^{2}\sigma^{2})\frac{\kappa}{K(\gamma+T)}+\frac{8E^{2}LG^{2}}{T^{2}}
+	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq C(\frac{\kappa\nu_{max}^{2}\sigma^{2}}{NT}+\frac{4\kappa E^{2}G^{2}}{KT}+\frac{5\kappa E^{2}LG^{2}}{T^{2}})
 	\end{align*}
 \end{theorem}
-\begin{proof}
-	See the appendix.
-\end{proof}
+%
+\begin{remark}
+	We remark on the linear speedup and the choice of $E$. In full participation,
+	the leading term is $\frac{1}{NT}\nu_{max}^{2}\sigma^{2}$ whereas
+	$\frac{5\kappa E^{2}LG^{2}}{T^{2}}$ is a lower order term if $E$
+	is chosen constant. This implies that $E$ can be chosen to be as
+	large as $E=O(\sqrt{\frac{T}{N}})$ which results in $\frac{5\kappa E^{2}LG^{2}}{T^{2}}=O(\frac{1}{NT})$,
+	which will be on the same order as the leading term. This is in contrast
+	with the convergence rate in {[}ICLR{]} where $E$ cannot scale with
+	$T$ and must be $O(1)$, even with full participation. Thus our bound
+	yields a more efficient communication complexity than previous analyses.
+	We note however that this is only possible when there is full participation,
+	and when $\nu_{max}$ is close to 1. When there is partial participation,
+	however, $E$ has to stay $O(1)$, as the term leading term in the
+	bound for partial participation contains $\frac{4\kappa E^{2}G^{2}}{KT}$.
+	Thus our result hightlights the difference between the communication
+	efficiency of full and partial participation when an $O(1/NT)$ bound
+	on the optimality gap is desired.
 
-The convergence result implies that with fixed $E$, as long as $N$ satisfies $E=O(\sqrt{T/N})$, there is linear speedup.
-When $N$ exceeds an upper bound, however, linear speedup may fail to happen, and $T/N$ may remain constant, i.e.
-the number of iterations required for convergence may increase with $N$.
+	\textbf{}%
+	\begin{comment}
+	This implies that $E$ canot be chosen $O(T^{\beta})$ for any $\beta>0$
+	without degrading the performance. This should be checked in experiments,
+	whether with partial participation if the communication round is set
+	to scale with $T$, the convergence deteriorates. This is in constrast
+	with the full participation case, where $E=O(\sqrt{\frac{T}{N}})$
+	is allowed. \textbf{If we can confirm that full participation allows
+	linear speedup with $\nu=N\cdot\max_{k}p_{k}\approx1$ and $E=O(\sqrt{\frac{T}{N}})$,
+	whereas partial participation only allows $E=O(1)$, then this would
+	an interesting phenomenon that is not reported by previous studies!}
+	\end{comment}
+\end{remark}
 

### 2020-05-29 source/sec_exp.tex
nesterov sgd
add exp appendix

@@
 \section{Numerical Experiments}
+\label{sec:exp}
 
@@
 
-\textbf{Setup.} Follow the experimental setting in~\cite{stich2018local}, We
+\textbf{Setup.} Follow the experimental setting in~\cite{stich2018local}, we
 conduct experiments on both synthetic datasets and real-world dataset
@@
 \begin{tabular}{ccc}
- &  & 
+\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} & 
+\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0134-reg0.pdf}
+& 
 \includegraphics[width=0.33\textwidth]{fig/paper-lrnesterovspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\

### 2020-05-29 source/sec_exp_details.tex
add exp appendix

% !TEX ROOT=./main.tex

\section{Details on Experiments}
% \subsection{Experiments}
\label{sec:expsupp}

We described the precise procedure to reproduce the results in this paper.
As we mentioned in Section~\ref{sec:exp}, we empirically verified the
linear speed up on various convex settings for both FedAvg and its
accelerated variants. For all the results, we set random seeds as $0, 1, 2$
and report the best convergence rate across the three folds. For each
run, we initialize $\vw_0 = \mathbf{0}$ and measure the number of iteration
to reach the target accuracy $\epsilon$. We use the small-scale dataset
w8a~\cite{platt1998fast}, which consists of $n = 49749 $ samples with
feature dimension $d = 300$. The label is either positive one or negative one.
The dataset has sparse binary features in $\{0, 1\}$. Each sample
has 11.15 non-zero feature values out of $300$ features on average.
We set the batch size equal to four across all experiments.
In the next following subsections,
we introduce parametering searching in each objective separately.

\subsection{Strongly Convex Objectives}
We first consider the strongly convex objective function, where we use
a regularized binary logistic regression with regularization $\lambda=1/n\approx 1e-5$. We evenly distributed on $1, 2, 4, 8, 16, 32$ devices and  report the number of iterations/rounds needed to converge to $\epsilon-$accuracy, where $\epsilon=0.005$. The optimal objective function value $f^*$
is set as $f^* = 0.126433176216545$. This is determined numerically and we follow the setting in~\cite{stich2018local}. The learning rate is decayed as the $\eta_t = \min(\eta_0, \frac{nc}{1 + t})$, where we extensively search the best learning rate $c \in \{2^{-1}c_0, 2^{-2}c_0, c_0, 2c_0, 2^{2}c_0\}$. In this case, we search the initial learning rate $\eta_0\in \{1, 32\}$ and $c_0 = 1/8$.

\subsection{Convex Smooth Objectives}
We also use binary logistic regression without regularization.
The setting is almost same as its regularized counter part. We also evenly distributed all the samples on $1, 2, 4, 8, 16, 32$ devices. The figure shows the number of iterations needed to converge to $\epsilon-$accuracy, where $\epsilon=0.02$. The optiaml objective function value is set as $f^*=0.11379089057514849$, determined numerically. 
The learning rate is decayed as the $\eta_t = \min(\eta_0, \frac{nc}{1 + t})$, where we extensively search the best learning rate $c \in \{2^{-1}c_0, 2^{-2}c_0, c_0, 2c_0, 2^{2}c_0\}$. In this case, we search the initial learning rate $\eta_0\in \{1, 32\}$ and $c_0 = 1/8$.

\subsection{Linear regression}
For linear regression, we use the same feature vectors from w8a dataset 
and generate ground truth $[\vw^*, b^*]$ from a multivariate normal distribution
with zero mean and standard deviation one. Then we generate label 
based on $y_i = \vx_i^t\vw^* + b^*$. This procedure will ensure we satisfy
the over-parameterized setting as required in our theorems. 
We also evenly distributed all the samples on $1, 2, 4, 8, 16, 32$ devices. The figure shows the number of iterations needed to converge to $\epsilon-$accuracy, where $\epsilon=0.02$. The optiaml objective function value is $f^*=0$. 
The learning rate is decayed as the $\eta_t = \min(\eta_0, \frac{nc}{1 + t})$, where we extensively search the best learning rate $c \in \{2^{-1}c_0, 2^{-2}c_0, c_0, 2c_0, 2^{2}c_0\}$. In this case, we search the initial learning rate $\eta_0\in \{0.1, 0.12\}$ and $c_0 = 1/256$.

### 2020-05-29 source/sec_exp_details.tex
experiments appendix details

@@
 In this case, we search the initial learning rate $\eta_0\in \{0.1, 0.12\}$ and $c_0 = 1/256$.
+
+\subsection{Partial Participation}
+To examine the linear speedup of FedAvg in partial participation experiments,
+we evenly distributed data on $4, 8, 16, 32, 64, 128$ devices and 
+uniformly sample $50\%$ devices without replacement. All other hyperparameters
+are the same as previous sections.
+
+\subsection{Nesterov accelerated FedAvg}
+The experiments of Nesterov accelerated FedAvg (the update formula is given as follows) uses the same setting as
+previous three sections for vanilia FedAvg.
+\begin{align*}
+y_{t+1}^{k} & =w_{t}^{k}-\alpha_{t}g_{t,k}\\
+w_{t+1}^{k} & =\begin{cases}
+y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E}\\
+\sum_{k\in\mathcal{S}_{t+1}}\left(y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k})\right) & \text{if }t+1\in\mathcal{I}_{E}
+\end{cases}
+\end{align*}
+We set $\beta_t = 0.1$ and search $\alpha_t$ in the same way as $\eta_t$
+in FedAvg.

### 2020-05-29 source/sec_exp.tex
experiments appendix details
add conclusion

@@
 generate a ground truth $w$ to label each instance, which satisfies the
-overparameterized setting.
+overparameterized setting. For the detailed experimental setup and
+hyperparameter setting, please refer to the section~\ref{sec:expsupp} in the
+appendix.
 
-% where we generated i.i.d. data $\vx_i^k$ for all devices. At each communication round, all devices participated
-% in synchronization. 
-\textbf{Linear speedup.} To verify the linear speedup results in Theorem~\ref{th:strongcvxsmth}~\ref{th:cvxsmoth}~\ref{th:fedmasslr}, we evaluate the number of iterations needed to reach
-$\epsilon$-accuracy in three objectives. We initialize all runs with $w_0 = \textbf{0}_d$ and measure. For each configuration, we run three random seeds and report
+% ~\ref{th:strongcvxsmth}~\ref{th:cvxsmoth}~\ref{th:fedmasslr}
+\textbf{Linear speedup.} To verify the linear speedup results in Theorems
+\lkxcom{cite sgd theorems}, we evaluate the number of iterations needed to reach
+$\epsilon$-accuracy in three objectives. We initialize all runs with $\vw_0 = \textbf{0}_d$ and measure. For each configuration, we run three random seeds and report
 the best seed results for each configuration $(E, K)$. In each configure, we
@@
 $\eta_0 \in \{0.1, 0.12, 1, 32 \}$ according to different problems and $c$ can
-take the values $c = 2^u \forall i \in \ZZ$. For the detailed experimental setup,
-please refer to the appendix. As the results shown in Figure~\ref{fig:speedup},
+take the values $c = 2^i \ \forall i \in \ZZ$. As the results shown in Figure~\ref{fig:speedup},
 the number of iterations to reach the target accuracy $\epsilon$ decreasing 
@@
 device participation.
+
+\textbf{Nesterov accelerated FedAvg.} In Figure~\ref{fig:nesterov}, 
+report the convergence of Nesterov accelerated FedAvg. The empirical 
+observation is aligned with our theorems \lkxcom{cite accelerated theorems} that the acclerated version of FedAvg can also achieve the lienar speedup
+w.r.t the number of nodes.
 
@@
 \begin{tabular}{ccc}
-\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} & 
+\hspace{-2em}\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} & 
 \includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0134-reg0.pdf}
@@
 	\end{tabular}
-\caption{The linear speedup convergence of Accelerated FedAvg w.r.t the number of workers. }
+\caption{The linear speedup convergence of Nesterov accelerated FedAvg w.r.t the number of workers. }
 \label{fig:nesterov}

### 2020-05-29 source/sec_conclude.tex

% !TEX ROOT=./main.tex

\section{Conclusions}
This paper provides a comprehensive analysis of the convergence rate of FedAvg
and its accelerated variants. We show that both accelerated FedAvg and FedAvg
can achieve speedup when the number of nodes increases, i.e., $O(1/\sqrt{NT})$
convergence for convex smooth problems and $O(1/NT)$ convergence for strongly 
convex problems. Furthermore, we show FedAvg can achieve exponential 
convergence for overparameterized strongly convex smooth problems, and we propose Mass accelerated Fedavg, which show improved convergence rate over
FedAvg on the linear regression problem. Last but not least, we empirically
verify the linear speedup of FedAvg and Nesterov accelerated FedAvg for strongly convex, convex smooth, and linear regression problems. The empirical results are well-aligned with our theories.

### 2020-05-29 source/sec_SGD.tex
Updated FedAvg with SGD results and discussions

% !TEX ROOT=./main.tex

\section{Linear Speedup Analysis of FedAvg }

In this section, we provide convergence analysis of FedAvg with local
SGD updates. We show that for strongly convex and smooth objectives,
the convergence of the optimality gap of averaged parameters across
devices is $O(1/NT)$ where $N$ is the number of active devices during
each communication round, while for convex and smooth objectives,
the rate is $O(1/\sqrt{NT})$.

\subsection{Strongly Convex and Smooth Objectives}

We first show that FedAvg with local SGD updates has $O(1/NT)$ convergence
rate for $\mu$-strongly convex and $L$-smooth objectives. The result
improves on the $O(1/T)$ result of {[}ICLR{]} with a linear speedup
in the number of devices $N$. Moreover, it implies a distinction
in the choice of $E$ that guarantees this linear speedup for FedAvg
with full and partial device participation. With full participation,
$E$ can be chosen as large as $O(\sqrt{\frac{T}{N}})$ without degrading
the linear speedup in the number of workers. On the other hand, with
partial participation, $E$ must be $O(1)$. 
\begin{theorem}
	Suppose $F_{k}$ is $L$-smooth and $\mu$-strongly convex for all
	$k$, and let $\nu_{\max}=\max_{k}\frac{1}{N}p_{k}$. Let $\kappa=\frac{L}{\mu}$,
	$\gamma=\max\{32\kappa,E\}$ where $E$ is the communication delay,
	and diminishing learning rates $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$.
	Let $\overline{w}_{T}=\sum_{k=1}^{N}p_{k}w_{T}^{k}$ be the average
	of local parameters at an arbitrary time $T$.

	With full device participation, 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
	\end{align*}
	and with partial device participation with $K$ sampled devices at
	each communication round, 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
	\end{align*}
\end{theorem}
%
\begin{remark}
	\textbf{Linear speedup. }We first compare our bound with that in {[}ICLR{]},
	which is $O(\frac{1}{NT}\nu_{\max}^{2}\sigma^{2}\kappa/\mu+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{E^{2}G^{2}}{T}\kappa/\mu)$.
	Because the third term $\frac{E^{2}G^{2}}{T}\kappa/\mu$ is also $O(1/T)$
	without a dependence on $N$, for any choice of $E$ their bound cannot
	achieve linear speedup. The improvement of our bound comes from the
	term $\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}$, which now is $O(1/T^{2})$.
	As a result, all leading terms scale with $1/N$ in the full device
	particiaption setting, and with $1/K$ in the partial participation
	setting. This implies that in both settings, there is a linear speedup
	in the number of active workers during the communication round.
\end{remark}
%
\begin{remark}
	\textbf{Choice of $E$.} Our bound implies a distinction in the choice
	of $E$ between the full and partial participation settings. In full
	participation, as long as $E=O(\sqrt{\frac{T}{N}})$, the term $\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}=O(\frac{1}{NT})$,
	which is on the same order as the leading term $\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}$.
	Thus to achieve a linear speedup in the number of workers $N$, it
	suffices to communicate every $O(\sqrt{\frac{T}{N}})$ iterations.
	In contrast, the bound in {[}ICLR{]} does not allow $E$ to scale
	with $\sqrt{T}$, even for full participation. Thus our bound yields
	a more efficient communication complexity, while at the same time
	providing a linear speedup in the convergence. On the other hand,
	with partial participation, the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$
	is also a leading term, and so $E$ must be $O(1)$. In this setting,
	our bound still yields a linear speedup in $K$, which is not the
	case for previous analyses. 
\end{remark}
%
\textbf{}%
\begin{comment}
This implies that $E$ canot be chosen $O(T^{\beta})$ for any $\beta>0$
without degrading the performance. This should be checked in experiments,
whether with partial participation if the communication round is set
to scale with $T$, the convergence deteriorates. This is in constrast
with the full participation case, where $E=O(\sqrt{\frac{T}{N}})$
is allowed. \textbf{If we can confirm that full participation allows
linear speedup with $\nu=N\cdot\max_{k}p_{k}\approx1$ and $E=O(\sqrt{\frac{T}{N}})$,
whereas partial participation only allows $E=O(1)$, then this would
an interesting phenomenon that is not reported by previous studies!}

The convergence result implies that with fixed $E$, as long as \$N\$
satisfies \$E=O(\textbackslash sqrt\{T/N\})\$, there is linear speedup.
When \$N\$ exceeds an upper bound, however, linear speedup may fail
to happen, and \$T/N\$ may remain constant, i.e. the number of iterations
required for convergence may increase with \$N\$. 
\end{comment}

\subsection{Convex Smooth Objectives}

In this section, we provide linear speedup analyses of FedAvg with
convex and smooth objectives and show that the optimality gap is $O(1/\sqrt{NT})$
where $N$ is the number of participating devices. This result complements
the strongly convex case in the previous part, as well as the non-convex
smooth setting in {[}CITE Yu and others{]}, where a similar $O(1/\sqrt{NT})$
rate is given in terms of averaged gradient norm. 
\begin{theorem}
	Suppose $F_{k}$ is $L$-smooth and convex for all $k$, and let $\nu_{\max}=\max_{k}\frac{1}{N}p_{k}$.
	Set learning rate $\alpha_{t}=O(\sqrt{\frac{N}{T}})$. With full device participation, 
	\begin{align*}
	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
	\end{align*}
	and with partial device participation with $K$ sampled devices at
	each communication round, 
	\begin{align*}
	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}G^{2}}{KT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
	\end{align*}
\end{theorem}

\begin{remark}
	\textbf{Choice of $E$ and linear speedup. }Similar in the strongly
	convex case, we see that in the full participation setting, as long
	as $E=O(\sqrt{\frac{T^{1/3}}{N}})$, the convergence benefits from
	a linear speedup in the number of active workers. In the partial participation
	setting, $E$ must be $O(1)$ in order to allow for linear speedup.
\end{remark}
%
\begin{remark}
	\textbf{Learning rate. }The learning rate now depends on the final
	horizon $T$ of the convergence statement, whereas before the learning
	rate was set to $O(1/t)$ for the $t$-th iteration. We also note
	that the requirement on $\alpha_{t}=O(\sqrt{\frac{N}{T}})$ is also
	present in the work of {[}Yu{]} where they derive similar $O(1/\sqrt{NT})$
	linear speedup convergence results for nonconvex smooth federated
	learning problems.
\end{remark}
For non-smooth objectives, we also have convergence results with rates
$O(1/T)$ and $O(1/\sqrt{T})$ for strongly convex and convex cases,
which can be found in the appendix. Without smoothness, however, the
linear speedup is no longer guaranteed. Together these results complete
the picture of convergence analyses for FedAvg with local SGD updates
for general convex heterogeneous objectives. We next investigate the
convergence of FedAvg with Nesterov updates, and provide similar linear
speedup analyses for convex smooth objectives.

### 2020-05-29 source/sec_SGD.tex
minor changes

@@
 \end{remark}
-%
-\textbf{}%
-\begin{comment}
-This implies that $E$ canot be chosen $O(T^{\beta})$ for any $\beta>0$
-without degrading the performance. This should be checked in experiments,
-whether with partial participation if the communication round is set
-to scale with $T$, the convergence deteriorates. This is in constrast
-with the full participation case, where $E=O(\sqrt{\frac{T}{N}})$
-is allowed. \textbf{If we can confirm that full participation allows
-linear speedup with $\nu=N\cdot\max_{k}p_{k}\approx1$ and $E=O(\sqrt{\frac{T}{N}})$,
-whereas partial participation only allows $E=O(1)$, then this would
-an interesting phenomenon that is not reported by previous studies!}
-
-The convergence result implies that with fixed $E$, as long as \$N\$
-satisfies \$E=O(\textbackslash sqrt\{T/N\})\$, there is linear speedup.
-When \$N\$ exceeds an upper bound, however, linear speedup may fail
-to happen, and \$T/N\$ may remain constant, i.e. the number of iterations
-required for convergence may increase with \$N\$. 
-\end{comment}
 

### 2020-05-30 source/sec_conclude.tex
updated conclusion

@@
 This paper provides a comprehensive analysis of the convergence rate of FedAvg
-and its accelerated variants. We show that both accelerated FedAvg and FedAvg
-can achieve speedup when the number of nodes increases, i.e., $O(1/\sqrt{NT})$
+and its accelerated variants in a general federated learning problem with heterogeneous local data and partial participation. We show that both accelerated FedAvg with Nesterov updates and FedAvg
+can achieve linear speedup when the number of nodes increases, i.e., $O(1/\sqrt{NT})$
 convergence for convex smooth problems and $O(1/NT)$ convergence for strongly 
-convex problems. Furthermore, we show FedAvg can achieve exponential 
-convergence for overparameterized strongly convex smooth problems, and we propose Mass accelerated Fedavg, which show improved convergence rate over
-FedAvg on the linear regression problem. Last but not least, we empirically
+convex smooth problems. Furthermore, we show that FedAvg can in fact achieve exponential 
+convergence for overparameterized strongly convex smooth problems, and we propose the MaSS accelerated Fedavg algorithm, which has provable speedup in convergence rate over
+FedAvg on quadratic problems. Last but not least, we empirically
 verify the linear speedup of FedAvg and Nesterov accelerated FedAvg for strongly convex, convex smooth, and linear regression problems. The empirical results are well-aligned with our theories.

### 2020-05-30 source/sec_relatedwork.tex
related work v0

% !TEX ROOT=./main.tex

\section{Related Works}

% related work on convergence rate
Federated learning (FL) was originally proposed
in~\cite{mcmahan2016communication} for learning a single, global statistal
model from the isolated data stored in a massive number of devices.  The
empirical success of FL~\cite{chen2018federated,47586} has attracted much
attentions on the rigorous theoretical understanding of the leading algorithm 
in the field: Federated Averaging (FedAvg) and its accelerated variants~\cite{liu2019accelerating,haddadpour2019convergence,khaled2019first,li2019convergence,huo2020faster,yu2019linear,yu2019parallel,koloskova2020unified}.

% Non-convex linear speedup
In particular, we focus on the convergence analysis of FedAvg, which
has been discussed in various settings. 
In \cite{yu2019parallel,wang2019adaptive}, the authors provides convergence
rate of FedAvg for non-convex problem, given the data are i.i.d. distributed.

### 2020-05-30 source/sec_overparameterized.tex

@@
 far exceeds the number of data points, and perfect fitting is possible.
+
 There have been a line of recent works {[}CITE{]} showing that SGD
 and accelerated methods achieve exponential convergence in this case,
-thanks to the property of ``automatic variance reduction'' property
-{[}CITE{]}. The natural question is whether such a result also holds
-in the federated learning setting. We first show that this is indeed
-the case and establish the exponential (linear) convergence of FedAvg
-with local SGD updates with constant step size for general strongly
-convex and smooth overparameterized problems. In addition, we show
-that the convergence rate speeds up linearly in the number of workers
-$N$ when $N$ is below some problem-dependent threhold, while decreases
-with $E$ through $1/E$. To our knowledge, this is the first result
-on the exponential convergence of FedAvg algorithms in the interpolation
-setting with linear speedup in the number of workers and explicit
-dependence on the communication interval $E$. We next sharpen this
-exponential convergence rate in the special case of linear regression,
-and show that it is $O(\exp(-\frac{NT}{E\kappa_{1}}))$, for $\kappa_{1}$
-an appropriately defined condition number of the system. Lastly, we
-turn to the question of whether FedAvg with momentum-based local updates
-can outperform FedAvg with SGD updates. In contrast to the gradient
-descent setting, in the single-agent stochastic gradient setting,
-Nesterov and Heavy Ball updates are known to fail to accelerate over
-SGD, both in the overparameterized setting and standard convex setting{[}CITE{]}.
-Thus in general we cannot hope to obtain acceleration results for
-the FedAvg algorithm with Nesterov and Heavy Ball updates. On the
-other hand, in the overparameterized setting, {[}CITE Belkin{]} introduced
-the MaSS algorithm, which is a modification of the Nesterov update,
-where in the update a non-negative multiple of the gradient is added
-to the Nesterov parameter update to correct for the ``over-descent''
+thanks to the property of ``automatic variance reduction'' {[}CITE{]}.
+The natural question is whether such a result also holds in the federated
+learning setting. We first show that this is indeed the case and establish
+the exponential (linear) convergence of FedAvg with local SGD updates
+with constant step size for general strongly convex and smooth overparameterized
+problems. In addition, we show that the convergence rate speeds up
+linearly in the number of workers $N$ when $N$ is below some problem-dependent
+threhold, while decreases with $E$ through $1/E$.
+
+To our knowledge, this is the first result on the exponential convergence
+of FedAvg algorithms in the interpolation setting with linear speedup
+in the number of workers and explicit dependence on the communication
+interval $E$. We next sharpen this exponential convergence rate in
+the special case of linear regression, and show that it is $O(\exp(-\frac{NT}{E\kappa_{1}}))$,
+for $\kappa_{1}$ an appropriately defined condition number of the
+system.
+
+Lastly, we turn to the question of whether FedAvg with momentum-based
+local updates can outperform FedAvg with SGD updates. In contrast
+to the gradient descent setting, in the single-agent stochastic gradient
+setting, Nesterov and Heavy Ball updates are known to fail to accelerate
+over SGD, both in the overparameterized setting and standard convex
+setting{[}CITE{]}. Thus in general we cannot hope to obtain acceleration
+results for the FedAvg algorithm with Nesterov and Heavy Ball updates.
+On the other hand, in the overparameterized setting, {[}CITE Belkin{]}
+introduced the MaSS algorithm, which is a modification of the Nesterov
+update, where in the update a non-negative multiple of the gradient
+is added to the Nesterov parameter update to correct for the ``over-descent''
 of the Nesterov update. For quadratic objectives, the authors show
@@
 convergence of SGD both theoretically and empirically. In the last
-part of this section, we introduce a new FedAvg algorithm with momentum
-updates by adapting the MaSS algorithm to the federated learning setting.
-For this new FedAvg algorithm with local MaSS updates, we show that
-it achieves exponential convergence for overparameterized quadratic
-problems with rate $O(\exp(-\frac{NT}{E\sqrt{\kappa_{1}\tilde{\kappa}}}))$,
+part of this section, we introduce a new accelerated FedAvg algorithm
+by adapting the MaSS algorithm to the federated learning setting.
+For this new accelerated FedAvg algorithm with local MaSS updates,
+we show that it achieves exponential convergence for overparameterized
+quadratic problems with rate $O(\exp(-\frac{NT}{E\sqrt{\kappa_{1}\tilde{\kappa}}}))$,
 where $\tilde{\kappa}$ is the ``statistical condition number''
@@
 
-\subsection{Exponential Convergence of FedAvg with local SGD: Overparameterized
-	Strongly Convex Smooth Objective}
+\subsection{Exponential Convergence and Linear Speedup of FedAvg with local SGD }
 
@@
 
-Recall the federated learning problem 
-\begin{align*}
-\min_{w}\sum_{k=1}^{N}p_{k}F_{k}(w)\\
-F_{k}(w)=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\ell(w;\xi_{k,j})
-\end{align*}
-In this section, we consider the standard ERM setting where each
-$\ell(w;\xi_{k,j})$ is non-negative, $l$-smooth, and convex, and
-as before, we assume that each $F_{k}(w)=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\ell(w;\xi_{k,j})$
+Recall the federated learning problem $\min_{w}\sum_{k=1}^{N}p_{k}F_{k}(w)$
+with $F_{k}(w)=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\ell(w;\xi_{k,j})$.
+In this section, we consider the standard ERM setting where each $\ell(w;\xi_{k,j})$
+is non-negative, $l$-smooth, and convex, and as before, we assume
+that each $F_{k}(w)=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\ell(w;\xi_{k,j})$
 is $L$-smooth and $\mu$-strongly convex. Note that $l\geq L$. This
@@
 \begin{theorem}
-	For the overparameterized setting with general strongly convex and
-	smooth objectives, FedAvg with local SGD updates and communication
+	For the overparameterized setting with general $\mu$-strongly convex
+	and $L$-smooth objectives, FedAvg with local SGD updates and communication
 	every $E$ iterations with constant step size $\overline{\alpha}=\frac{1}{4E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})}$
@@
 \begin{remark}
-	We see that when $l>L$, the speedup factor is on the order of $\frac{N}{E(l\nu_{\max}+L(N-\nu_{\min}))}/\frac{1}{l\nu_{\max}+L(1-\nu_{\min})}\approx\frac{Nl}{E(l+L(N-1))}=O(N/E)$
+	When $l>L$, the speedup factor is on the order of $\frac{N}{E(l\nu_{\max}+L(N-\nu_{\min}))}/\frac{1}{l\nu_{\max}+L(1-\nu_{\min})}\approx\frac{Nl}{E(l+L(N-1))}=O(N/E)$
 	for $N\leq\frac{l}{L}+1$, i.e. FedAvg with $N$ workers and communication
@@
 and there exists $w^{\ast}$ such that $F(w^{\ast})=\sum_{k}p_{k}F_{k}(w^{\ast})\equiv0$.
-In this section, we gather some necessary definitions and observations
+For such quadratic problems, we will present a sharper bound for FedAvg
+with explicit dependence on the condition number, and demonstrate
+how the convergence rate can be further improved using a momentum-based
+accelerated FedAvg algorithm based on the MaSS algorithm of {[}CITE
+Belkin{]}.
+
+In this section, we first gather some necessary definitions and observations
 useful in the analysis of exponential convergence of FedAvg algorithms.
-
-Define the local Hessian matrix as 
-\begin{align*}
-H^{k} & :=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}x_{k,j}(x_{k,j})^{T}
-\end{align*}
-and the averaged Hessian matrix as 
-\begin{align*}
-H & :=\sum_{k=1}^{N}p_{k}H^{k}
-\end{align*}
-
+Define the local Hessian matrix as $H^{k}:=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}x_{k,j}(x_{k,j})^{T}$
+and the averaged Hessian matrix as $H:=\sum_{k=1}^{N}p_{k}H^{k}$.
 In general $H$ has zero eigevalues. However, because the null space
@@
 \begin{align*}
-F_{k}(w) & =\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(w^{T}x_{k,j}-z_{k,j}-(w^{\ast T}x_{k,j}-z_{k,j}))^{2}\\
-& =\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}((w-w^{\ast})^{T}x_{k,j})^{2}=\frac{1}{2}\langle w-w^{\ast},H^{k}(w-w^{\ast})\rangle=\frac{1}{2}\|w-w^{\ast}\|_{H^{k}}^{2}
+F_{k}(w) & =\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(w^{T}x_{k,j}-z_{k,j}-(w^{\ast T}x_{k,j}-z_{k,j}))^{2}=\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}((w-w^{\ast})^{T}x_{k,j})^{2}\\
+& =\frac{1}{2}\langle w-w^{\ast},H^{k}(w-w^{\ast})\rangle=\frac{1}{2}\|w-w^{\ast}\|_{H^{k}}^{2}
 \end{align*}
-so that 
-\begin{align*}
-F(w) & =\sum_{k=1}^{N}p_{k}F_{k}(w)=\sum_{k=1}^{N}p_{k}\frac{1}{2}(w-w^{\ast})^{T}H^{k}(w-w^{\ast})=\frac{1}{2}(w-w^{\ast})^{T}H(w-w^{\ast})=\frac{1}{2}\|w-w^{\ast}\|_{H}^{2}
-\end{align*}
+so that $F(w)=\frac{1}{2}\|w-w^{\ast}\|_{H}^{2}$.
 
@@
 numbers $\kappa_{1}$ and $\tilde{\kappa}$ are important in the characterization
-of convergence rates for FedAvg algorithms.
+of convergence rates for FedAvg algorithms. Note that $\kappa_{1}>\kappa$ and $\kappa_{1}>\tilde{\kappa}$.
 
-Note that $\kappa_{1}>\kappa$ and $\kappa_{1}>\tilde{\kappa}$.
+\subsection{Acceleration of FedAvg with Momentum-based Updates}
 
-\subsection{Exponential Convergence of FedAvg with local SGD: Over-parameterized
-	Quadratic Objective}
-
-Next we show that specialized to overparameterized quadratic objectives,
-the exponential convergence bound can be further sharpened, with a
-similar linear speedup in the number of workers. Recall that in this
-setting, we have 
-\begin{align*}
-F(w) & =\sum_{k=1}^{N}p_{k}F_{k}(w)\\
-F_{k}(w) & =\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(w^{T}x_{k,j}-z_{k,j})^{2}
-\end{align*}
-and there exists $w^{\ast}$ such that $F_{k}(w^{\ast})\equiv0$
-for all $k$. 
+In the single-agent stochastic optimization setting, Nesterov updates
+are known to fail to accelerate over SGD updates, so that in the federated
+learning setting, we cannot hope for acceleration results of FedAvg
+with Nesterov updates in general. The natural question is then whether
+there are accelerated versions of the FedAvg algorithm that have provable
+improvements over the FedAvg algorithm based on local SGD updates.
+We answer this question in the affirmative in the overparameterized
+quadratic problem setting, extending recent results {[}CITE: Sidford,
+Belkin{]} in the single agent setting. In particular, we show that
+specialized to overparameterized quadratic objectives, the exponential
+convergence bound can be further sharpened, with a similar linear
+speedup in the number of workers. This rate will then be further improved
+using a new accelerated FedAvg algorithm with momentum-based updates. 
 \begin{theorem}
-	For the overparamterized linear regression problem, FedAvg with local
-	SGD updates and communication every $E$ iterations with constant
-	step size $\alpha_{t}=\frac{1}{4E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})}$
+	For the overparamterized quadratic problem, FedAvg with local SGD
+	updates and communication every $E$ iterations with constant step
+	size $\alpha_{t}=\frac{1}{4E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})}$
 	gives the exponential convergence guarantee 
@@
 	\end{align*}
-	where $\kappa_{1}=l/\mu$. 
 \end{theorem}
 %
+
 We see that when $N=O(\kappa_{1})$, the convergence rate is $O((1-\frac{N}{E\kappa_{1}})^{t})=O(\exp(-\frac{N}{E\kappa_{1}}t))$,
@@
 
- \subsection{Exponential Convergence of FedAvg with local MaSS: Over-parameterized
- 	Quadratic Objective}
-
- Following {[}CITE{]}, we propose the FedAvg algorithm with local MaSS
- updates: 
+Following {[}CITE{]}, we propose the accelerated FedAvg algorithm
+with local MaSS updates: 
 \begin{align*}
@@
 and $g_{t}=\sum_{k=1}^{N}p_{k}g_{t,k}$ is the averaged stochastic
- gradient. When $\eta_{2}^{k}\equiv0$, this reduces to the FedAvg
- algorithm with Nesterov updates.
+gradient. When $\eta_{2}^{k}\equiv0$, this reduces to the accelerated
+FedAvg algorithm with Nesterov updates.
 
@@
 \end{align*}
- where there is a bijection between the parameters 
- \begin{align*}
- \frac{1-\alpha^{k}}{1+\alpha^{k}} & =\gamma^{k}\\
- \eta^{k} & =\eta_{1}^{k}\\
- \frac{\eta^{k}-\alpha^{k}\delta^{k}}{1+\alpha^{k}} & =\eta_{2}^{k}
- \end{align*}
+where there is a bijection between the parameters $\frac{1-\alpha^{k}}{1+\alpha^{k}}=\gamma^{k},\eta^{k}=\eta_{1}^{k},\frac{\eta^{k}-\alpha^{k}\delta^{k}}{1+\alpha^{k}}=\eta_{2}^{k}$,
 and we further introduce an auxiliary parameter $v_{t}^{k}$, which
@@
 \begin{theorem}
- 	(FedAvg with MaSS, Linear Regression) For the overparamterized quadratic
- 	problem, FedAvg with local MaSS updates and communication every $E$
- 	iterations with constant step sizes 
+For the overparamterized quadratic problem, FedAvg with local MaSS
+updates and communication every $E$ iterations with constant step
+sizes 
 \begin{align*}
@@
 rate is $O((1-\frac{N}{E\sqrt{\kappa_{1}\tilde{\kappa}}})^{t})=O(\exp(-\frac{N}{E\sqrt{\kappa_{1}\tilde{\kappa}}}t))$
- as opposd to $O(\exp(-\frac{N}{E\kappa_{1}}t))$. Since $\kappa_{1}\geq\tilde{\kappa}$,
+as opposed to $O(\exp(-\frac{N}{E\kappa_{1}}t))$. Since $\kappa_{1}\geq\tilde{\kappa}$,
 this implies a speedup factor of $\sqrt{\frac{\kappa_{1}}{\tilde{\kappa}}}$

### 2020-05-30 source/sec_relatedwork.tex
Update sec_relatedwork.tex
related work v1

@@
 Federated learning (FL) was originally proposed
-in~\cite{mcmahan2016communication} for learning a single, global statistal
+in~\cite{mcmahan2016communication} for learning a single, global statistical
 model from the isolated data stored in a massive number of devices.  The
-empirical success of FL~\cite{chen2018federated,47586} has attracted much
-attentions on the rigorous theoretical understanding of the leading algorithm 
-in the field: Federated Averaging (FedAvg) and its accelerated variants~\cite{liu2019accelerating,haddadpour2019convergence,khaled2019first,li2019convergence,huo2020faster,yu2019linear,yu2019parallel,koloskova2020unified}.
+empirical success of FL~\cite{chen2018federated,47586} has attracted much attention to the rigorous theoretical understanding of the leading algorithm in the field: Federated Averaging (FedAvg) and its accelerated variants~\cite{liu2019accelerating,haddadpour2019convergence,khaled2019first,li2019convergence,huo2020faster,yu2019linear,yu2019parallel}.
 
@@
 has been discussed in various settings. 
-In \cite{yu2019parallel,wang2019adaptive}, the authors provides convergence
-rate of FedAvg for non-convex problem, given the data are i.i.d. distributed.
+In~\cite{yu2019parallel,wang2019adaptive}, the authors provides convergence
+rate of FedAvg for non-convex problem, given all the devices are active at
+each communication round. Recently, \citep{huo2020faster} provide 
+convergence guarantee for Nesterov accelerated FedAvg for non-convex 
+setting in partial participation setting. Although the empirical
+results is improved over FedAvg, the accelerated version admitting a similar 
+convergence rate as FedAvg. In this work, we provide an novel accelerated
+FedAvg with improved convergence rate under a popular overparameterized 
+setting. In addition, \cite{yu2019linear} provides $O(1/\sqrt{NT})$ 
+convergence (i.e., linear speedup) of FedAvg for non-convex problem in the full participation setting. 
+For convex smooth problems, \cite{khaled2019first} prove the 
+convergence rate of local GD (a full batch version of FedAvg with full participation) on heterogeneous data. 
+Concurrently, \cite{li2019convergence} firstly provide exponential convergence of FedAvg for strongly convex problem,
+under the more practical setting of partial participation and Non-IID data,
+while their results did not achieve linear speedup. 
+\cite{stich2018local} prove linear speedup convergence of FedAvg for strongly convex problem, while they assume full participation and i.i.d. data.
+Comparing to prior work, our results provide the linear speedup on strongly convex problem in the practical partial participation setting and heterogeneous data.

### 2020-05-30 source/sec_relatedwork.tex
related work add ref

@@
 Comparing to prior work, our results provide the linear speedup on strongly convex problem in the practical partial participation setting and heterogeneous data.
+In \cite{liu2019accelerating}, the authors provide convergence rate of accelerated FedAvg for stronlgy convex problem, while it is considered in full participation setting.
+To the best of our knowledge,
+we provide the first linear speedup convergence results for 
+accelerated FedAvg in practical setting.

### 2020-05-30 source/save.tex

@@
-\section{The Convergence of FedAve for General Convex Functions}
+\section{The Convergence of FedAvg for General Convex Functions}
 
-\subsection{The Convergence of FedAve for Convex Functions}
-In this subsection, we discuss the convergence of FedAve for convex functions,
+\subsection{The Convergence of FedAvg for Convex Functions}
+In this subsection, we discuss the convergence of FedAvg for convex functions,
 we consider both smooth (Theorem~\ref{th:sgdcvxsmth}) and non-smooth (Theorem~\ref{th:sgdcvxnonsmoth})

### 2020-05-30 source/sec_Nesterov.tex

% !TEX ROOT=./main.tex

\section{Linear Speedup Analysis of Nesterov Accelerated FedAvg}

A natural extension of the FedAvg algorithm is to use momentum-based
local updates instead of local SGD updates. To our knowledge, the
only formal analyses of accelerated FedAvg with momentum-based stochastic
updates {[}CITE kdd papaer{]} focus on the non-convex smooth case.
In this section, we complete the picture by providing convergence
results of Nesterov-accelerated FedAvg for convex objectives. We show
that for strongly convex and smooth objectives, the convergence of
the optimality gap of averaged parameters across devices is $O(1/NT)$
where $N$ is the number of active devices during each communication
round, while for convex and smooth objectives, the rate is $O(1/\sqrt{NT})$.
Thus Nesterov accelerated FedAvg algorithms enjoy the same convergence
rates as the original FedAvg algorithm. As we know from stochastic
optimization that Nesterov updates may fail to accelerate over SGD,
in general we cannot hope to obtain acceleration results in the federated
learning setting. However, in the following section we specialize
to the overparameterized setting where we demonstrate that a particular
accelerated FedAvg algorithm with momentum-based updates is able to
improve on the original FedAvg algorithm.

\subsection{Strongly Convex and Smooth Objectives}

We first show that the Nesterov accelerated FedAvg has $O(1/NT)$
convergence rate for $\mu$-strongly convex and $L$-smooth objectives.
The Nesterov Accelerated FedAvg algorithm follows the updates
\begin{align*}
y_{t+1}^{k} & =w_{t}^{k}-\alpha_{t}g_{t,k}\\
w_{t+1}^{k} & =\begin{cases}
y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}\left[y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}
\end{align*}
where $g_{t,k}:=\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})$ is the stochastic
gradient. 
\begin{theorem}
	Assume that $F_{k}$ is $L$-smooth and $\mu$-strongly convex for
	all $k$, and let $\nu_{\max}=\max_{k}\frac{1}{N}p_{k}$. Let $\kappa=\frac{L}{\mu}$,
	$\gamma=\max\{32\kappa,E\}$ where $E$ is the communication delay,
	and diminishing learning rates $\alpha_{t}=\beta_{t}=\frac{1}{4\mu(\gamma+t)}$.
	Let $\overline{w}_{T}=\sum_{k=1}^{N}p_{k}w_{T}^{k}$ be the average
	of local parameters at an arbitrary time $T$.

	With full device participation, 
	\begin{align*}
	\mathbb{E}F(\overline{y}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
	\end{align*}
	and with partial device participation with $K$ sampled devices at
	each communication round, 
	\begin{align*}
	\mathbb{E}F(\overline{y}_{T})-F^{\ast}\leq O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
	\end{align*}
\end{theorem}
%
\begin{remark}
To our knowledge, this is the first convergence result for Nesterov accelerated FedAvg in the convex and smooth setting. The same discussion about linear speedup and choice of $E$ from the
	previous section applies to the Nesterov accelerated FedAvg algorithm. 
\end{remark}

\subsection{Convex Smooth Objectives}

In this section, we provide linear speedup analyses of Nesterov Accelerated
FedAvg with convex and smooth objectives and show that the optimality
gap is $O(1/\sqrt{NT})$ where $N$ is the number of participating
devices. This result complements the strongly convex case in the previous
part, as well as the non-convex smooth setting in {[}CITE kdd and
others{]}, where a similar $O(1/\sqrt{NT})$ rate is given in terms
of averaged gradient norm. 
\begin{theorem}
	Suppose $F_{k}$ is $L$-smooth and convex for all $k$, and let $\nu_{\max}=\max_{k}\frac{1}{N}p_{k}$.
	Set learning rate $\alpha_{t}=O(\sqrt{\frac{N}{T}})$.

	With full device participation, 
	\begin{align*}
	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
	\end{align*}
	and with partial device participation with $K$ sampled devices at
	each communication round, 
	\begin{align*}
	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}G^{2}}{KT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
	\end{align*}
\end{theorem}
%
\begin{remark}
	It is possible to extend the analysis to accelerated FedAvg algorithms
	with other momentum-based updates. However, in line with the stochastic
	optimization setting, none of these methods will be able to achieve
	a better rate than the original FedAvg algorithm in general. For this
	reason, we will instead turn to the overparameterized setting {[}CITE
	Belkin{]} where we show that it is possible to improve the convergence
	rate of FedAvg with momentum based local updates. 
\end{remark}

### 2020-05-30 source/sec_conclude.tex

@@
 convergence for convex smooth problems and $O(1/NT)$ convergence for strongly 
-convex smooth problems. Furthermore, we show that FedAvg can in fact achieve exponential 
+convex smooth problems. [TODO:Add discussion about communication efficiency.]Furthermore, we show that FedAvg can in fact achieve exponential 
 convergence for overparameterized strongly convex smooth problems, and we propose the MaSS accelerated Fedavg algorithm, which has provable speedup in convergence rate over

### 2020-05-30 source/sec_impact.tex

@@
 
-In the meanwhile, the increasing concerns of security and
+In the meantime, the increasing concerns of security and
 privacy leaking from both large-companies and the entire society
@@
 data. This trade-off between efficiency and privacy or security
-lies in the central of many industrial applications, including 
+lies in the center of many industrial applications, including 
 but not limited to recommender systems~\cite{chen2018federated}, virtual assistants~\cite{lamautonomy}, and mobile keyboard prediction~\cite{47586}
 
-In this work, we establish solid theoretical foundation for federated
+In this work, we establish solid theoretical foundations for federated
 learning that achieves efficiency and privacy at the same time. 

### 2020-05-30 source/sec_overparameterized.tex

@@
 
-There have been a line of recent works {[}CITE{]} showing that SGD
+There have been a line of recent works {[}CITE{\cite{ma2017power}]} showing that SGD
 and accelerated methods achieve exponential convergence in this case,

### 2020-05-30 source/sec_relatedwork.tex

--- 2020-05-30 source/sec_relatedwork.tex
@@
 has been discussed in various settings. 
-In~\cite{yu2019parallel,wang2019adaptive}, the authors provides convergence
-rate of FedAvg for non-convex problem, given all the devices are active at
-each communication round. Recently, \citep{huo2020faster} provide 
+In~\cite{yu2019parallel,wang2019adaptive}, the authors provide the convergence rate of FedAvg for the non-convex problem, given all the devices are active at each communication round.
+Recently, \citep{huo2020faster} provide
 convergence guarantee for Nesterov accelerated FedAvg for non-convex 
-setting in partial participation setting. Although the empirical
-results is improved over FedAvg, the accelerated version admitting a similar 
-convergence rate as FedAvg. In this work, we provide an novel accelerated
-FedAvg with improved convergence rate under a popular overparameterized 
-setting. In addition, \cite{yu2019linear} provides $O(1/\sqrt{NT})$ 
-convergence (i.e., linear speedup) of FedAvg for non-convex problem in the full participation setting. 
+setting in a partial participation setting. Although the empirical
+results are improved over FedAvg, the accelerated version admitting a similar convergence rate as FedAvg. In this work, we provide a novel accelerated
+FedAvg with improved convergence rates under a popular overparameterized setting. In addition, \cite{yu2019linear} provides $O(1/\sqrt{NT})$ 
+convergence (i.e., linear speedup) of FedAvg for the non-convex problem in the full participation setting. 
 For convex smooth problems, \cite{khaled2019first} prove the 
 convergence rate of local GD (a full batch version of FedAvg with full participation) on heterogeneous data. 
-Concurrently, \cite{li2019convergence} firstly provide exponential convergence of FedAvg for strongly convex problem,
-under the more practical setting of partial participation and Non-IID data,
-while their results did not achieve linear speedup. 
-\cite{stich2018local} prove linear speedup convergence of FedAvg for strongly convex problem, while they assume full participation and i.i.d. data.
-Comparing to prior work, our results provide the linear speedup on strongly convex problem in the practical partial participation setting and heterogeneous data.
+Concurrently, \cite{li2019convergence} firstly provides exponential convergence of FedAvg for the strongly convex problem,
+under the more practical setting of partial participation and Non-IID data while their results did not achieve linear speedup. 
+\cite{stich2018local} prove linear speedup convergence of FedAvg for the strongly convex problem, while they assume full participation and identically distributed data.
+Comparing to previous works, our results provide the linear speedup on strongly convex problems in the practical partial participation setting and heterogeneous data.
+In \cite{liu2019accelerating}, the authors provide a convergence rate of accelerated FedAvg for strongly convex problems, while it is considered in full participation setting.
+To the best of our knowledge,
+we provide the first linear speedup convergence results for accelerated FedAvg in the practical setting.

### 2020-05-30 source/sec_relatedwork.tex
Update sec_relatedwork.tex

@@
 In particular, we focus on the convergence analysis of FedAvg, which
-has been discussed in various settings. 
+has been studied in various scenarios. 
 In~\cite{yu2019parallel,wang2019adaptive}, the authors provide the convergence rate of FedAvg for the non-convex problem, given all the devices are active at each communication round.
+\cite{haddadpour2019convergence}
+improves the convergence analysis for non-convex problem under the setting of partial participation. 
 Recently, \citep{huo2020faster} provide

### 2020-05-30 source/sec_relatedwork.tex
test overleaf change

@@
 convergence rate of local GD (a full batch version of FedAvg with full participation) on heterogeneous data. 
-Concurrently, \cite{li2019convergence} firstly provides exponential convergence of FedAvg for the strongly convex problem,
+Concurrently, \cite{li2019convergence} provides exponential convergence of FedAvg for the strongly convex problem,
 under the more practical setting of partial participation and Non-IID data while their results did not achieve linear speedup. 

### 2020-05-30 source/app_sgd_cvx_smooth.tex
sgd proofs complete

\begin{theorem}
	Suppose $F_{k}$ is $L$-smooth and convex for all $k$, and let $\nu_{\max}=\max_{k}\frac{1}{N}p_{k}$.
	Set learning rate $\alpha_{t}=O(\sqrt{\frac{N}{T}})$.

	With full device participation, 
	\begin{align*}
	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
	\end{align*}
	and with partial device participation with $K$ sampled devices at
	each communication round, 
	\begin{align*}
	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}G^{2}}{KT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
	\end{align*}
\end{theorem}

\begin{proof}
	We again start by bounding the term
	\begin{align*}
	\|\overline{w}_{t+1}-w^{\ast}\|^{2} & =\|(\overline{w}_{t}-\alpha_{t}g_{t})-w^{\ast}\|^{2}\\
	& =\|(\overline{w}_{t}-\alpha_{t}\overline{g}_{t}-w^{\ast})-\alpha_{t}(g_{t}-\overline{g}_{t})\|^{2}\\
	& =A_{1}+A_{2}+A_{3}
	\end{align*}
	where 
	\begin{align*}
	A_{1} & =\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2}\\
	A_{2} & =2\alpha_{t}\langle\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t},\overline{g}_{t}-g_{t}\rangle\\
	A_{3} & =\alpha_{t}^{2}\|g_{t}-\overline{g}_{t}\|^{2}
	\end{align*}
	$\mathbb{E}A_{2}=0$ by definition of $g_{t}$ and $\overline{g}_{t}$,
	while for $A_{3}$ we have
	\begin{align*}
	\alpha_{t}^{2}\mathbb{E}\|g_{t}-\overline{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|g_{t}-\mathbb{E}g_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
	\end{align*}
	again by Jensen's inequality and using the independence of $g_{t,k},g_{t,k'}$.

	Next we bound $A_{1}$: 
	\begin{align*}
	\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & =\|\overline{w}_{t}-w^{\ast}\|^{2}+2\langle\overline{w}_{t}-w^{\ast},-\alpha_{t}\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2}
	\end{align*}
	Using the convexity and $L$-smoothness of $F_{k}$, 
	\begin{align*}
	-2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle & =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(w^{\ast})-F_{k}(w_{t}^{k}))\\
	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w_{t}^{k})-F_{k}(\overline{w}_{t})+\frac{L}{2}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+F_{k}(w^{\ast})-F_{k}(w_{t}^{k})\right]\\
	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]
	\end{align*}
	which results in 
	\begin{align*}
	\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]+\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
	\end{align*}

	The difference of this bound with that in the strongly convex case
	is that we no longer have a contraction factor in front of $\|\overline{w}_{t}-w^{\ast}\|^{2}$.
	In the strongly convex case, we were able to cancel $\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}$
	with $2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]$
	and obtain only lower order terms. In the convex case, we use a different
	strategy and preserve $\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]$
	in order to obtain a telescoping sum.

	We have
	\begin{align*}
	\|\overline{g}_{t}\|^{2} & =\|\sum_{k}p_{k}\nabla F_{k}(w_{t}^{k})\|^{2}\\
	& =\|\sum_{k}p_{k}\nabla F_{k}(w_{t}^{k})-\sum_{k}p_{k}\nabla F_{k}(\overline{w}_{t})+\sum_{k}p_{k}\nabla F_{k}(\overline{w}_{t})\|^{2}\\
	& \leq2\|\sum_{k}p_{k}\nabla F_{k}(w_{t}^{k})-\sum_{k}p_{k}\nabla F_{k}(\overline{w}_{t})\|^{2}+2\|\sum_{k}p_{k}\nabla F_{k}(\overline{w}_{t})\|^{2}\\
	& \leq2L^{2}\sum_{k}p_{k}\|w_{t}^{k}-\overline{w}_{t}\|^{2}+2\|\sum_{k}p_{k}\nabla F_{k}(\overline{w}_{t})\|^{2}\\
	& =2L^{2}\sum_{k}p_{k}\|w_{t}^{k}-\overline{w}_{t}\|^{2}+2\|\nabla F(\overline{w}_{t})\|^{2}
	\end{align*}
	using $\nabla F(w^{\ast})=0$. Now using the $L$ smoothness of $F$,
	we have $\|\nabla F(\overline{w}_{t})\|^{2}\leq2L(F(\overline{w}_{t})-F(w^{\ast}))$,
	so that 
	\begin{align*}
	\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]\\
	& +2\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|w_{t}^{k}-\overline{w}_{t}\|^{2}+4\alpha_{t}^{2}L(F(\overline{w}_{t})-F(w^{\ast}))+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
	& =\|\overline{w}_{t}-w^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
	& +\alpha_{t}(1-4\alpha_{t}L)(F(w^{\ast})-F(\overline{w}_{t}))
	\end{align*}
	Since $F(w^{\ast})\leq F(\overline{w}_{t})$, as long as $4\alpha_{t}L\leq1$,
	we can ignore the last term, and rearrange the inequality to obtain
	\begin{align*}
	\|\overline{w}_{t+1}-w^{\ast}\|^{2}+\alpha_{t}(F(\overline{w}_{t})-F(w^{\ast})) & \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
	& \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+\frac{3}{2}\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
	\end{align*}

	The same argument as before yields $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}\leq4E^{2}\alpha_{t}^{2}G^{2}$
	which gives 
	\begin{align*}
	\|\overline{w}_{t+1}-w^{\ast}\|^{2}+\alpha_{t}(F(\overline{w}_{t})-F(w^{\ast})) & \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+6\alpha_{t}^{3}E^{2}LG^{2}\\
	& \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
	\end{align*}
	Summing the inequalities from $t=0$ to $t=T$, we obtain 
	\begin{align*}
	\sum_{t=0}^{T}\alpha_{t}(F(\overline{w}_{t})-F(w^{\ast})) & \leq\|w_{0}-w^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}
	\end{align*}
	so that
	\begin{align*}
	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|w_{0}-w^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right)
	\end{align*}

	By setting the constant learning rate $\alpha_{t}\equiv\sqrt{\frac{\|w_{0}-w^{\ast}\|^{2}}{\frac{1}{N}\nu_{\max}^{2}\sigma^{2}T+6E^{2}LG^{2}T^{2/3}}}=O(\sqrt{\frac{N}{T}})$,
	we have

	\begin{align*}
	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & \leq\frac{1}{T\sqrt{\frac{\|w_{0}-w^{\ast}\|^{2}}{\frac{1}{N}\nu_{\max}^{2}\sigma^{2}T+6E^{2}LG^{2}T^{2/3}}}}\cdot\left(\|w_{0}-w^{\ast}\|^{2}+T\frac{\|w_{0}-w^{\ast}\|^{2}}{\frac{1}{N}\nu_{\max}^{2}\sigma^{2}T}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6E^{2}LG^{2}T\frac{\|w_{0}-w^{\ast}\|^{2}}{6E^{2}LG^{2}T}\right)\\
	& =\frac{3\|w_{0}-w^{\ast}\|^{2}}{T\sqrt{\frac{\|w_{0}-w^{\ast}\|^{2}}{\frac{1}{N}\nu_{\max}^{2}\sigma^{2}T+6E^{2}LG^{2}T^{2/3}}}}=3\|w_{0}-w^{\ast}\|\sqrt{(\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{6E^{2}LG^{2}}{T^{4/3}})}
	\end{align*}

	Similarly, for partial participation, we just need to replace $\frac{\nu_{\max}^{2}\sigma^{2}}{NT}$
	with $\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+C\frac{E^{2}G^{2}}{KT}$,
	so that we have 
	\begin{align*}
	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}G^{2}}{KT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
	\end{align*}
\end{proof}

### 2020-05-30 source/app_sgd_scvx_smooth.tex
sgd proofs complete

We first summarize some properties of $L$-smooth and $\mu$-strongly
convex functions. 
\begin{lemma}
	Let $F$ be a convex $L$-smooth function. Then we have the following
	inequalities:

	1. Quadratic upper bound: $0\leq F(w)-F(w')-\langle\nabla F(w'),w-w'\rangle\leq\frac{L}{2}\|w-w'\|^{2}$.

	2. Coercivity: $\frac{1}{L}\|\nabla F(w)-\nabla F(w')\|^{2}\leq\langle\nabla F(w)-\nabla F(w'),w-w'\rangle$.

	3. Lower bound: $F(w)\geq F(w')+\langle\nabla F(w'),w-w'\rangle+\frac{1}{2L}\|\nabla F(w)-\nabla F(w')\|^{2}$.
	In particular, $\|\nabla F(w)\|^{2}\leq2L(F(w)-F(w^{\ast}))$.

	4. Optimality gap: $F(w)-F(w^{\ast})\leq$$\langle\nabla F(w),w-w^{\ast}\rangle$.
\end{lemma}
%
\begin{lemma}
	Let $F$ be a $\mu$-strongly convex function. Then 
	\begin{align*}
	F(w) & \leq F(w')+\langle\nabla F(w'),w-w'\rangle+\frac{1}{2\mu}\|\nabla F(w)-\nabla F(w')\|^{2}\\
	F(w)-F(w^{\ast}) & \leq\frac{1}{2\mu}\|\nabla F(w)\|^{2}
	\end{align*}
\end{lemma}
%
\begin{assumption}
	The stochastic gradients have bounded variance 
	\begin{align*}
	\mathbb{E}\|\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})-\nabla F_{k}(w_{t}^{k})\|^{2} & \leq\sigma_{k}^{2}
	\end{align*}
	for all $k$. Moreover, the expected squared norm of stochastic gradients
	is uniformly bounded, 
	\begin{align*}
	\mathbb{E}\|\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})\|^{2} & \leq G^{2}
	\end{align*}
\end{assumption}

\begin{theorem}
	Assume that $F_{k}$ is $L$-smooth and $\mu$-strongly convex for
	all $k$, and let $\nu_{\max}=\max_{k}\frac{1}{N}p_{k}$. Let $\kappa=\frac{L}{\mu}$,
	$\gamma=\max\{32\kappa,E\}$ where $E$ is the communication delay,
	and diminishing learning rates $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$.
	Let $\overline{w}_{T}=\sum_{k=1}^{N}p_{k}w_{T}^{k}$ be the average
	of local parameters at an arbitrary time $T$.

	With full device participation, 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
	\end{align*}
	and with partial device participation with $K$ sampled devices at
	each communication round, 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
	\end{align*}
\end{theorem}

\begin{proof}
		The idea of the proof is to observe that the $L$-smoothness of $F$
		provides the upper bound
		\begin{align*}
		\mathbb{E}(F(\overline{w}_{t}))-F^{\ast} & =\mathbb{E}(F(\overline{w}_{t})-F(w^{\ast}))\\
		& \leq\frac{L}{2}\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}
		\end{align*}
		and bound $\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}$.

		Our main step is to prove the bound 
		\begin{align*}
		\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}L\alpha_{t}^{3}G^{2}
		\end{align*}

		We have 
		\begin{align*}
		\|\overline{w}_{t+1}-w^{\ast}\|^{2} & =\|(\overline{w}_{t}-\alpha_{t}g_{t})-w^{\ast}\|^{2}\\
		& =\|(\overline{w}_{t}-\alpha_{t}\overline{g}_{t}-w^{\ast})-\alpha_{t}(g_{t}-\overline{g}_{t})\|^{2}\\
		& =A_{1}+A_{2}+A_{3}
		\end{align*}
		where 
		\begin{align*}
		A_{1} & =\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2}\\
		A_{2} & =2\alpha_{t}\langle\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t},\overline{g}_{t}-g_{t}\rangle\\
		A_{3} & =\alpha_{t}^{2}\|g_{t}-\overline{g}_{t}\|^{2}
		\end{align*}
		$\mathbb{E}A_{2}=0$ by definition of $g_{t}$ and $\overline{g}_{t}$,
		while for $A_{3}$ we have
		\begin{align*}
		\alpha_{t}^{2}\mathbb{E}\|g_{t}-\overline{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|g_{t}-\mathbb{E}g_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
		\end{align*}
		again by Jensen's inequality and using the independence of $g_{t,k},g_{t,k'}$.

		Next we bound $A_{1}$: 
		\begin{align*}
		\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & =\|\overline{w}_{t}-w^{\ast}\|^{2}+2\langle\overline{w}_{t}-w^{\ast},-\alpha_{t}\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2}
		\end{align*}
		and we will show that the third term $\|\alpha_{t}\overline{g}_{t}\|^{2}$
		can be canceled by an upper bound of the second term. %
		\begin{comment}
		The last term is straightforward to bound by the convexity of $\|\cdot\|^{2}$
		and $L$-smoothness of $F_{k}$,
		\begin{align*}
		\alpha_{t}^{2}\|\overline{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(w_{t}^{k})\|^{2}\leq2L\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}(F_{k}(w_{t}^{k})-F_{k}^{\ast})
		\end{align*}
		or 
		\begin{align*}
		\alpha_{t}^{2}\|\overline{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(w_{t}^{k})\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\mathbb{E}\|\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})\|^{2}\leq\alpha_{t}^{2}G^{2}
		\end{align*}
		\end{comment}

		Now 
		\begin{align*}
		-2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle & =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
		& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
		& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(w^{\ast})-F_{k}(w_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-w^{\ast}\|^{2}\\
		& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w_{t}^{k})-F_{k}(\overline{w}_{t})+\frac{L}{2}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+F_{k}(w^{\ast})-F_{k}(w_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}w_{t}^{k}-w^{\ast}\|^{2}\\
		& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]-\alpha_{t}\mu\|\overline{w}_{t}-w^{\ast}\|^{2}
		\end{align*}
		For the second term, which is negative, we can ignore it, but this
		yields a suboptimal bound that fails to provide the desired linear
		speedup. Instead, we upper bound it using the following derivation:
		\begin{align*}
		2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right] & \leq2\alpha_{t}\left[F(\overline{w}_{t+1})-F(\overline{w}_{t})\right]\\
		& \leq2\alpha_{t}\mathbb{E}\langle\nabla F(\overline{w}_{t}),\overline{w}_{t+1}-\overline{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\overline{w}_{t+1}-\overline{w}_{t}\|^{2}\\
		& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\overline{w}_{t}),g_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
		& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\overline{w}_{t}),\overline{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
		& =-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\|\nabla F(\overline{w}_{t})-\overline{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
		& =-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\|\nabla F(\overline{w}_{t})-\sum_{k}p_{k}\nabla F(w_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
		& \leq-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\overline{w}_{t})-\nabla F(w_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
		& \leq-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
		& \leq-\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\overline{w}_{t})\|^{2}
		\end{align*}
		where we have used the smoothness of $F$ twice.

		Note that the term $-\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}$ exactly
		cancels the $\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}$ in the bound
		for $A_{1}$, so that plugging in the bound for $-2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle$,
		we have so far proved 
		\begin{align*}
		\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
		& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\overline{w}_{t})\|^{2}
		\end{align*}
		The term $\mathbb{E}\|g_{t}\|^{2}\leq G^{2}$ by assumption.

		Now we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}$.
		Since communication is done every $E$ steps, for any $t\geq0$, we
		can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $w_{t_{0}}^{k}=\overline{w}_{t_{0}}$for
		all $k$. Moreover, using $\alpha_{t}$ is non-increasing and $\alpha_{t_{0}}\leq2\alpha{}_{t}$
		for any $t-t_{0}\leq E-1$, we have 
		\begin{align*}
		\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2} & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}-(\overline{w}_{t}-\overline{w}_{t_{0}})\|^{2}\\
		& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}\|^{2}\\
		& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-w_{t_{0}}^{k}\|^{2}\\
		& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\
		& \leq\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}\|g_{i,k}\|^{2}\\
		& \leq\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{t_{0}}^{2}G^{2}\\
		& \leq4E^{2}\alpha_{t}^{2}G^{2}
		\end{align*}

		Using the bound on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}$,
		we can conclude that, with $\nu_{max}:=N\cdot\max_{k}p_{k}$ and $\nu_{min}:=N\cdot\min_{k}p_{k}$,

		\begin{align*}
		\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
		& =\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\sum_{k=1}^{N}(p_{k}N)^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
		& \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\nu_{max}^{2}\sum_{k=1}^{N}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
		& \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
		\end{align*}
		where $\sigma^{2}=\max_{k}\sigma_{k}^{2}$, and if $\alpha_{t}$
		satisfies $L\alpha_{t}\leq\frac{1}{8}$. We show next that $\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}}{t^{2}})$.

		Let $C\equiv5E^{2}LG^{2}$ and $D\equiv\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
		Suppose that we have shown $\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}\leq b\cdot(\alpha_{t}D+\alpha_{t}^{2}C)$
		for some constant $b$ and $\alpha_{t}$. Then 
		\begin{align*}
		\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq b(1-\mu\alpha_{t})(\alpha_{t}D+\alpha_{t}^{2}C)+\alpha_{t}^{2}D+\alpha_{t}^{3}C\\
		& =(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}D+(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}C
		\end{align*}
		and so it remains to choose $\alpha_{t}$ and $b$ such that $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}\leq b\alpha_{t+1}$.
		Recall that we require $\alpha_{t_{0}}\leq\alpha_{t}$ for any $t-t_{0}\leq E-1$,
		and $L\alpha_{t}\leq\frac{1}{8}$. If we let $\alpha_{t}=\frac{4}{\mu(t+\gamma)}$
		where $\gamma=\max\{E,32\kappa\}$, then we may check that $\alpha_{t}$
		satisfies both requirements.

		Setting $b=\frac{4}{\mu}$, we have 
		\begin{align*}
		(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t} & =\left(b(1-\frac{4}{t+\gamma})+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}\\
		& =\left(b\frac{t+\gamma-4}{t+\gamma}+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}\\
		& =b(\frac{t+\gamma-2}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
		& \leq b(\frac{t+\gamma-1}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
		& \leq b\frac{4}{\mu(t+\gamma+1)}=b\alpha_{t+1}
		\end{align*}
		and 
		\begin{align*}
		(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2} & =\left(b(1-\frac{4}{t+\gamma})+\frac{4}{\mu(t+\gamma)}\right)\frac{16}{\mu^{2}(t+\gamma)^{2}}\\
		& =\left(b\frac{t+\gamma-4}{t+\gamma}+\frac{4}{\mu(t+\gamma)}\right)\frac{16}{\mu^{2}(t+\gamma)^{2}}\\
		& =b(\frac{t+\gamma-2}{t+\gamma})\frac{16}{\mu^{2}(t+\gamma)^{2}}\\
		& \leq b\frac{16}{\mu^{2}(t+\gamma+1)^{2}}=b\alpha_{t+1}^{2}
		\end{align*}
		and where we have used the facts that 
		\begin{align*}
		\frac{t+\gamma-1}{(t+\gamma)^{2}} & \leq\frac{1}{(t+\gamma+1)}\\
		\frac{t+\gamma-2}{(t+\gamma)^{3}} & \leq\frac{1}{(t+\gamma+1)^{2}}
		\end{align*}
		for $\gamma\geq1$.

		Thus we have shown 
		\begin{align*}
		\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq b\cdot(\alpha_{t+1}D+\alpha_{t+1}^{2}C)
		\end{align*}
		for our choice of $\alpha_{t}$ and $b$. Now to ensure 
		\begin{align*}
		\|w_{0}-w^{\ast}\|^{2} & \leq b\cdot(\alpha_{0}D+\alpha_{0}^{2}C)\\
		& =b\cdot(\frac{4}{\mu\gamma}D+\frac{16}{\mu^{2}\gamma^{2}}C)
		\end{align*}
		we can simply scale $b$ by $c\|w_{0}-w^{\ast}\|^{2}$ for a constant
		$c$ large enough and the induction step still holds.

		It follows that 
		\begin{align*}
		\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2} & \leq c\|w_{0}-w^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{t}+C\alpha_{t}^{2})
		\end{align*}
		for all $t\geq0$.

		Finally, the $L$-smoothness of $F$ implies 
		\begin{align*}
		\mathbb{E}(F(\overline{w}_{T}))-F^{\ast} & =\mathbb{E}(F(\overline{w}_{T})-F(w^{\ast}))\\
		& \leq\frac{L}{2}\mathbb{E}\|\overline{w}_{T}-w^{\ast}\|^{2}\leq\frac{L}{2}c\|w_{0}-w^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
		& =2c\|w_{0}-w^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
		& \leq2c\|w_{0}-w^{\ast}\|^{2}\kappa\left[\frac{4}{\mu T}\cdot\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}LG^{2}\cdot(\frac{4}{\mu T})^{2}\right]\\
		& =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa}{\mu^{2}}E^{2}LG^{2}\cdot\frac{1}{T^{2}})
		\end{align*}

		With partial participation, the update at each communication round
		is now given by averages over a subset of sampled devices. The key
		is to bound 
		\begin{align*}
		\mathbb{E}\|\overline{w}_{t}-\overline{y}_{t}\|^{2} & =\frac{1}{K}\sum_{k}p^{k}\|w_{t}^{k}-\sum_{k}p_{k}w_{t}^{k}\|^{2}
		\end{align*}
		This can be bounded by the same argument as before and yield 
		\begin{align*}
		\mathbb{E}\|\overline{w}_{t}-\overline{y}_{t}\|^{2} & \leq\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
		\end{align*}
		which results in the bound 
		\begin{align*}
		\mathbb{E}F(\overline{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
		\end{align*}
	\end{proof}

### 2020-05-30 source/sec_exp.tex
update exp
reorg exp

@@
 
-In this section, we empirically examine the theoretical convergence rate of FedAvg on various scenarios, including strongly convex function, convex smooth function, and overparameterized setting, as presented in previous sections.
+In this section, we empirically examine the linear speedup convergence of FedAvg and Nesterov accelerated FedAvg on various scenarios, including strongly convex function, convex smooth function, and overparameterized setting, as proved in previous theorems.
 
@@
 objectives $F(\vw) = \sum_{k=1}^N p_kF_k(\vw)$, and the objective function on
-each local device $k$ includes three cases: 1) \textbf{Strongly convex
+$k$-th local device includes three cases: 1) \textbf{Strongly convex
 objective}: the regularized binary logistic regression problem, $F_k(\vw) =
-\frac{1}{N_k} \sum_{i=1}^{N_k} \log( 1+ \exp(-y_i^k \vw^T\vx_i^k) + \lambda
+\frac{1}{N_k} \sum_{i=1}^{N_k} \log( 1+ \exp(-y_i^k \vw^T\vx_i^k) + \frac{\lambda}{2}
 \|\vw\|^2$. The regularization parameter is set to $\lambda = 1/n \approx
@@
 $\eta_0 \in \{0.1, 0.12, 1, 32 \}$ according to different problems and $c$ can
-take the values $c = 2^i \ \forall i \in \ZZ$. As the results shown in Figure~\ref{fig:speedup},
+take the values $c = 2^i \ \forall i \in \ZZ$. As the results shown in Figure~\ref{fig:speedup} (1st row),
 the number of iterations to reach the target accuracy $\epsilon$ decreasing 
@@
 
-\textbf{The impact of $E$.} We further examine how does the number of epochs 
-affect the convergence. As shown in Figure~\ref{fig:e}, the number of iterations
-increases as the number of epochs increase, which slow the convergence in
-terms of gradient computation. However, it can save communication cost as
-the number of rounds decreased when the number of epochs increases.
-
 \textbf{Partial participations.} We verify the linear speedup in the
 partial participation settings, where we set $50\%$ of devices
-are active. As the results shown in Figure~\ref{fig:partial}, the 
+are active. As the results are shown in Figure~\ref{fig:speedup} (2nd row), the 
 FedAve enjoys linear speedup in various settings even with partial
@@
 
-\textbf{Nesterov accelerated FedAvg.} In Figure~\ref{fig:nesterov}, 
-report the convergence of Nesterov accelerated FedAvg. The empirical 
-observation is aligned with our theorems \lkxcom{cite accelerated theorems} that the acclerated version of FedAvg can also achieve the lienar speedup
-w.r.t the number of nodes.
+\textbf{Nesterov accelerated FedAvg.} In the third row of Figure~\ref{fig:speedup}, 
+we report the last iteration to converge to $\epsilon$-accuracy of Nesterov accelerated FedAvg.
+The empirical observations align with our theorems \lkxcom{cite accelerated theorems} that the accelerated version of FedAvg can also achieve the linear speedup
+w.r.t the number of workers.
 
-\begin{figure}
+\textbf{The impact of $E$.} We further examine how does the number of local steps ($E$) 
+affect convergence. As shown in Figure~\ref{fig:e}, the number of iterations increases as $E$ increase, which slow down the convergence in
+terms of gradient computation. However, it can save communication costs as
+the number of rounds decreased when the $E$ increases. This showcase that
+we need a proper choice of $E$ to trade-off the communication cost and
+convergence speed.
+
+\begin{figure}[t!]
 \centering
@@
 \includegraphics[width=0.33\textwidth]{fig/paper-linregressionspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
+\hspace{-2em}\includegraphics[width=0.33\textwidth]{fig/paper-partialstronglycvxsmthspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} &
+\includegraphics[width=0.33\textwidth]{fig/paper-partialcvxsmoothspeedupNodesT-min-w8a-epsilon0134-reg0.pdf} &
+\includegraphics[width=0.33\textwidth]{fig/paper-partiallinregressionspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
+\hspace{-2em}\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} & 
+\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0134-reg0.pdf}
+& 
+\includegraphics[width=0.33\textwidth]{fig/paper-lrnesterovspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
 (a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
 	\end{tabular}
-\caption{The linear speedup convergence of FedAvg w.r.t the number of workers. }
+\caption{\textbf{Linear speedup} The first row reports the linear speedup convergence of FedAvg w.r.t the number of workers.
+The second row reports the linear speedup convergence of FedAvg w.r.t the number of active workers in the partial parcipation setting.
+The third row reports the linear speedup convergence of Nesterov accelerated FedAvg in full participation setting.
+}
 \label{fig:speedup}
 \end{figure}
+
+% \begin{figure}
+% \centering
+% 	\begin{tabular}{ccc}
+% 	\hspace{-2em}\includegraphics[width=0.33\textwidth]{fig/paper-partialstronglycvxsmthspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} &
+% \includegraphics[width=0.33\textwidth]{fig/paper-partialcvxsmoothspeedupNodesT-min-w8a-epsilon0134-reg0.pdf} &
+% \includegraphics[width=0.33\textwidth]{fig/paper-partiallinregressionspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
+% (a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
+% 	\end{tabular}
+% \caption{The linear speedup convergence of FedAvg w.r.t the number of active workers. }
+% \vspace{-2em}
+% \label{fig:partial}
+% \end{figure}
+
+% \begin{figure}
+% \centering
+% \begin{tabular}{ccc}
+% \hspace{-2em}\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} & 
+% \includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0134-reg0.pdf}
+% & 
+% \includegraphics[width=0.33\textwidth]{fig/paper-lrnesterovspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
+% (a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
+% 	\end{tabular}
+% \caption{The linear speedup convergence of Nesterov accelerated FedAvg w.r.t the number of workers. }
+% \label{fig:nesterov}
+% \end{figure}
 
@@
 	\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupEpochsT-min-w8a-epsilon0134-reg0.pdf} & 
-	\includegraphics[width=0.33\textwidth]{fig/paper-linregressionspeedupEpochsT-min-linearregressionw8a-epsilon002-reg0.pdf} \\
+	\includegraphics[width=0.33\textwidth]{fig/paper-linregression-newspeedupEpochsT-min-linearregressionw8a-epsilon002-reg0.pdf} \\
 	\hspace{-2em} \includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsRounds-min-w8a-epsilon0131-reg1e-05.pdf} &
@@
 	\end{tabular}
-\caption{The convergence of FedAvg w.r.t the number of epochs. }
+\caption{The convergence of FedAvg w.r.t the number of local steps $E$. }
 \label{fig:e}
 \end{figure}
-
-\begin{figure}
-\centering
-	\begin{tabular}{ccc}
-	\hspace{-2em}\includegraphics[width=0.33\textwidth]{fig/paper-partialstronglycvxsmthspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} &
-\includegraphics[width=0.33\textwidth]{fig/paper-partialcvxsmoothspeedupNodesT-min-w8a-epsilon0134-reg0.pdf} &
-\includegraphics[width=0.33\textwidth]{fig/paper-partiallinregressionspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
-(a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
-	\end{tabular}
-\caption{The linear speedup convergence of FedAvg w.r.t the number of active workers. }
-\label{fig:partial}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{tabular}{ccc}
-\hspace{-2em}\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0131-reg1e-05.pdf} & 
-\includegraphics[width=0.33\textwidth]{fig/paper-nesterovspeedupNodesT-min-w8a-epsilon0134-reg0.pdf}
-& 
-\includegraphics[width=0.33\textwidth]{fig/paper-lrnesterovspeedupNodesT-min-linearregressionw8a-epsilon002-reg0.pdf}\\
-(a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
-	\end{tabular}
-\caption{The linear speedup convergence of Nesterov accelerated FedAvg w.r.t the number of workers. }
-\label{fig:nesterov}
-\end{figure}

### 2020-05-31 source/sec_intro.tex
Update sec_intro.tex

--- 2020-05-27 source/sec_intro.tex
@@
 \section{Introduction}
-% Motivate FL from the security and privacy perspective. FedAvg
-Federated learning (FL) is a distributed learning paradigm that can collaboratively solve a machine learning problem by leveraging the data
-from a large amount of workers without compromising the users' (aka workers/local devices) privacy~\cite{kairouz2019advances}. 
-Among all the leading algorithms in FL, Federated Averaging~\cite{mcmahan2016communication} (FedAvg)
-is perhaps the most widely adopted optimization algorithm, which runs local
-Stochastic Gradient Descent (SGD) updates on a subset of devices
-and synchronize the local models once in a while. 
-The empirical success of FedAvg and its variants have
-attracted lots of efforts~\cite{li2018federated,stich2018local,khaled2019first,yu2019parallel,haddadpour2019convergence,li2019convergence,huo2020faster} on analyzing its convergence properties. 
-There are few key challenges that differentiates the theoretical analysis
-of FL from the tradition distribution optimization: 
-1) The data is non-identically distributed across the workers, which means the
-data in each local device cannot be regarded as samples drawn from
-a same distribution. 
-2) The workers are not active at every communication
-round. In FL, the central server has no control over the local devices. 
-It is more practical to assume only a subset of workers is active during
-each communication round.
+% Motivate FL
+The widespread usage of smart home devices and mobile phones constitutes
+large-scale distributed networks that generate datasets of unprecedented size.
+The emergence of those massive amounts of data has empowered machine learning
+solutions to develop many successful applications such as personalized
+recommendations~\cite{chen2018federated}, keyboard predictions~\cite{47586},
+etc. Due to the increasing awareness of the privacy and rapid development of
+mobile devices' computational power, it becomes more and more attractive to
+push the computation to the edge. Recent years have witnessed the rise of
+federated learning~\cite{mcmahan2016communication} to full fill this demand.
+
+% Main challenge in FL.
+As a distributed learning paradigm, Federated Learning (FL) collaboratively
+learns a single statistical model leveraging the data from all devices while
+keeping the training data isolated.  Thus, FL can integrate cross-device
+information for acquiring a better model  without compromising the users'
+privacy~\cite{kairouz2019advances}. Comparing to the traditional distribution optimization,  FL confronts unique
+challenges~\cite{li2019convergence,smith2017federated}, including statistical heterogeneity: the data distribution in each local device can be significantly different, and system heterogeneity: only a subset of devices are accessible to the central server for model training.
+
+% Gaps in the theoretical understanding. 
+Towards resolving the aforementioned challenges, Federated
+Averaging~\cite{mcmahan2016communication} (FedAvg) was proposed as an
+effective heuristic and its great empirical success~\cite{47586} has attracted
+lots of efforts on building up the theoretical understanding of FedAvg
+and its accelerated variants. 
+Roughly, the current convergence guarantee of FedAvg obtained by prior arts either established on simplified setting~\cite{stich2018local,khaled2019first,wang2018cooperative,yu2019parallel} or proved a suboptimal
+convergence rate~\cite{li2019convergence}. Furthermore, the convergence of
+FedAvg's accelerated variants~\cite{yu2019linear,huo2020faster,liu2019accelerating} under the realistic setting is still waiting to be solved.
+
+\textbf{Notations}
+Throughout the paper, we adopt the following notation. Let $N$ be the total
+number of local devices, and $K \leq N$ is the maximal number of devices
+that are accessible to the central server.
+Let $T$ be the total number of  Stochastic Gradient Descent (SGD) performed on each local device, $E$ be the local steps between two communication steps.
+Thus, $\frac{T}{E}$ is
+the number of communications.
+
+% Our contributions
+\subsection{Contributions}
+This work focus on providing comprehensive convergence analysis
+of FedAvg and accelerated variants under the realistic setting, showing
+the less number of communication round is needed to converge to the
+target accuracy and preserving the linear speedup at the same time.
+More specifically, for strongly convex and smooth problems, we provide the {\small{$\cO(\frac{1}{NT})$}} convergence  for FedAvg and the local steps can be as large as {\small{$\cO(\sqrt{\frac{T}{N}})$}}.
+For convex and smooth objectives, we prove the {\small{$\cO(\sqrt{\frac{1}{NT}})$}} and the local steps can be as large as {\small{$\cO(\sqrt{\frac{T}{N}})$}}.
+The same conclusions can be equally applied for Nesterov accelerated FedAvg.
+To the best of our knowledge, this is the first linear speedup results for FedAvg with Nesterov accelerated SGD.
+Furthermore,
+we also provide the first improved convergence rate for Nesterov accelerated FedAvg comparing to FedAvg under a popular overparamterized setting. 
+Last but not least, we systematically examine the convergence rate of
+FedAvg and it accelerated variants on both a real-world dataset and a synthetic dataset. The empirical observations are well-aligned with theoretical indications.
+
+% Our theoretical results not only cover all general convex objectives, including strongly convex cases, convex smooth cases and convex non-smooth cases, but also provide tighter convergence guarantee for all convex smooth objectives, which shows the convergence rate enjoys a linear speedup w.r.t.
+% the number of workers.
+% Furthermore, we studied a popular over-parameterized setting~\cite{liu2018accelerating} where the optimal solution can obtain zero training loss.
+In this scenario, we provide a novel accelerated FL algorithm improving
+% the convergence rate of FedAvg. Last but not least, we conduct extensive
+% evaluation on both synthetic and real-world dataset, which demonstrates that our theoretical indications is well-aligned with the empirical observations.
+
+% \textbf{Organization} 
+% The reminader of this paper is organized as follows. In Section
+
+% Intro and related work.
+% sec 4. linear speedup of FedAvg,  sec 5. linear speedup of Accelerated FedAvg.
+% sec 6 linear regression in 
+% 1. this is the first result on the exponential convergence of FedAvg algorithms in the
+% interpolation setting with linear speedup in the number of workers and explicit dependence on the
+% communication interval E.  
+% 2. Improved convergence rate over FedAvg for For the overparamterized quadratic problem.
+
+% without considering statistical heterogeneity
+% or system heterogeneity~\cite{stich2018local,khaled2019first,wang2018cooperative,yu2019parallel,yu2019linear} or suboptimal 
+% convergence rate~\cite{li2019convergence}.
+
+% is perhaps the most widely adopted optimization algorithm, which runs local
+% Stochastic Gradient Descent (SGD) updates on a subset of devices
+% and synchronize the local models once in a while. 
+% The empirical success of FedAvg and its variants have
+% attracted lots of efforts~\cite{li2018federated,stich2018local,khaled2019first,yu2019parallel,haddadpour2019convergence,li2019convergence,huo2020faster} on analyzing its convergence properties. 
+% There are few key challenges that differentiates the theoretical analysis
+% of FL from the tradition distribution optimization: 
+% 1) The data is non-identically distributed across the workers, which means the
+% data in each local device cannot be regarded as samples drawn from
+% a same distribution. 
+% 2) The workers are not active at every communication
+% round. In FL, the central server has no control over the local devices. 
+% It is more practical to assume only a subset of workers is active during
+% each communication round.
 
 % The current gaps in FL.
-However, most of prior works~\cite{li2018federated,stich2018local,khaled2019first,yu2019parallel,haddadpour2019convergence} either assume
-the data is identically distributed or the all devices are active, which
-violates the practical characteristic in FL. The most recent work~\cite{li2019convergence} firstly presents $O(1/T)$ convergence guarantee without 
-making those unrealistic assumptions while their analysis focused 
-on strongly convex case only and the relation between the number of active
-workers and the convergence rate is not clearly discussed.
-
-% Our contributions
-Our work aims to systematically fill the gap in the convergence analysis
-of FedAvg and its accelerated variants. 
-Our theoretical results not only cover all general convex objectives, including strongly convex cases, convex smooth cases and convex non-smooth cases, but also provide tighter convergence guarantee all convex smooth objectives, which shows the convergence rate enjoys a linear speedup
-Furthermore, we studied a popular over-parameterized setting~\cite{liu2018accelerating} where the optimal solution can obtain zero training loss.
-In this scenario, we provide a novel accelerated FL algorithm improving
-the convergence rate of FedAvg. Last but not least, we conduct extensive
-evaluation on both synthetic and real-world dataset, which demonstrates that our theoretical indications is well-aligned with the empirical observations.
+% However, most of prior works~\cite{li2018federated,stich2018local,khaled2019first,yu2019parallel,haddadpour2019convergence} either assume
+% the data is identically distributed or the all devices are active, which
+% violates the practical characteristic in FL. The most recent work~\cite{li2019convergence} firstly presents $O(1/T)$ convergence guarantee without 
+% making those unrealistic assumptions while their analysis focused 
+% on strongly convex case only and the relation between the number of active
+% workers and the convergence rate is not clearly discussed.

### 2020-05-31 source/app_nasgd_scvx_smooth.tex
updated nesterov section and proofs

We show that FedAvg with Accelerated SGD has $O(1/T)$ rate under
$\mu$-strong convexity and $L$-smoothness. The proof follows the
framework of the ICLR paper. The FedAv algorithm with Nesterov Accelerated
SGD (NASGD) follows the updates
\begin{align*}
y_{t+1}^{k} & =w_{t}^{k}-\alpha_{t}g_{t,k}\\
w_{t+1}^{k} & =\begin{cases}
y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}\left[y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}
\end{align*}
where 
\begin{align*}
g_{t,k} & :=\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})
\end{align*}
is the stochastic gradient.

Define the virtual sequences $\overline{y}_{t}=\sum_{k=1}^{N}p_{k}y_{t}^{k}$,
$\overline{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$, and $\overline{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$.
We have $\mathbb{E}g_{t}=\overline{g}_{t}$ and $\overline{y}_{t+1}=\overline{w}_{t}-\alpha_{t}g_{t}$,
and $\overline{w}_{t+1}=\overline{y}_{t+1}+\beta_{t}(\overline{y}_{t+1}-\overline{y}_{t})$.

\begin{theorem}
	Assume that $F_{k}$ is $L$-smooth and $\mu$-strongly convex for
	all $k$, and let $\nu_{\max}=\max_{k}\frac{1}{N}p_{k}$. Let $\kappa=\frac{L}{\mu}$,
	$\gamma=\max\{32\kappa,E\}$ where $E$ is the communication delay,
	and diminishing learning rates $\alpha_{t}=\frac{9}{\mu}\frac{1}{t+\gamma}$
	and $\beta_{t-1}=\frac{9}{14\max\{\mu,1\}}\frac{1}{t+\gamma}$. Let
	$\overline{y}_{T}=\sum_{k=1}^{N}p_{k}y_{T}^{k}$ be the average of
	local parameters at an arbitrary time $T$.

	With full device participation, 
	\begin{align*}
	\mathbb{E}F(\overline{y}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
	\end{align*}
	and with partial device participation with $K$ sampled devices at
	each communication round, 
	\begin{align*}
	\mathbb{E}F(\overline{y}_{T})-F^{\ast}\leq O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
	\end{align*}
\end{theorem}
%
%
\textbf{}%
\begin{comment}
This implies that $E$ canot be chosen $O(T^{\beta})$ for any $\beta>0$
without degrading the performance. This should be checked in experiments,
whether with partial participation if the communication round is set
to scale with $T$, the convergence deteriorates. This is in constrast
with the full participation case, where $E=O(\sqrt{\frac{T}{N}})$
is allowed. \textbf{If we can confirm that full participation allows
linear speedup with $\nu=N\cdot\max_{k}p_{k}\approx1$ and $E=O(\sqrt{\frac{T}{N}})$,
whereas partial participation only allows $E=O(1)$, then this would
an interesting phenomenon that is not reported by previous studies!}

The convergence result implies that with fixed $E$, as long as \$N\$
satisfies \$E=O(\textbackslash sqrt\{T/N\})\$, there is linear speedup.
When \$N\$ exceeds an upper bound, however, linear speedup may fail
to happen, and \$T/N\$ may remain constant, i.e. the number of iterations
required for convergence may increase with \$N\$. 
\end{comment}

\begin{proof}
	The proof again uses the $L$-smoothness of $F$ to bound 
	\begin{align*}
	\mathbb{E}(F(\overline{y}_{t}))-F^{\ast} & =\mathbb{E}(F(\overline{y}_{t})-F(w^{\ast}))\\
	& \leq\frac{L}{2}\mathbb{E}\|\overline{y}_{t}-w^{\ast}\|^{2}
	\end{align*}

	Our main step is to prove the bound 
	\begin{align*}
	\mathbb{E}\|\overline{y}_{t+1}-w^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\overline{y}_{t}-w^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}L\alpha_{t}^{3}G^{2}
	\end{align*}

	We have 
	\begin{align*}
	\|\overline{y}_{t+1}-w^{\ast}\|^{2} & =\|(\overline{w}_{t}-\alpha_{t}g_{t})-w^{\ast}\|^{2}\\
	& =\|(\overline{w}_{t}-\alpha_{t}\overline{g}_{t}-w^{\ast})-\alpha_{t}(g_{t}-\overline{g}_{t})\|^{2}\\
	& =A_{1}+A_{2}+A_{3}
	\end{align*}
	where 
	\begin{align*}
	A_{1} & =\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2}\\
	A_{2} & =2\alpha_{t}\langle\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t},\overline{g}_{t}-g_{t}\rangle\\
	A_{3} & =\alpha_{t}^{2}\|g_{t}-\overline{g}_{t}\|^{2}
	\end{align*}
	$\mathbb{E}A_{2}=0$ by definition of $g_{t}$ and $\overline{g}_{t}$,
	while for $A_{3}$ we have
	\begin{align*}
	\alpha_{t}^{2}\mathbb{E}\|g_{t}-\overline{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|g_{t}-\mathbb{E}g_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
	\end{align*}
	again by Jensen's inequality and using the independence of $g_{t,k},g_{t,k'}$.

	Next we bound $A_{1}$: 
	\begin{align*}
	\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & =\|\overline{w}_{t}-w^{\ast}\|^{2}+2\langle\overline{w}_{t}-w^{\ast},-\alpha_{t}\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2}
	\end{align*}
	Same as the SGD case, 
	\begin{align*}
	-2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2} & \leq\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}\mu\|\overline{w}_{t}-w^{\ast}\|^{2}
	\end{align*}
	so that 
	\begin{align*}
	\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & \leq(1-\alpha_{t}\mu)\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}
	\end{align*}

	Different from the SGD case, we have 
	\begin{align*}
	\|\overline{w}_{t}-w^{\ast}\|^{2} & =\|\overline{y}_{t}+\beta_{t-1}(\overline{y}_{t}-\overline{y}_{t-1})-w^{\ast}\|^{2}\\
	& =\|(1+\beta_{t-1})(\overline{y}_{t}-w^{\ast})-\beta_{t-1}(\overline{y}_{t-1}-w^{\ast})\|^{2}\\
	& =(1+\beta_{t-1})^{2}\|\overline{y}_{t}-w^{\ast}\|^{2}-2\beta_{t-1}(1+\beta_{t-1})\langle\overline{y}_{t}-w^{\ast},\overline{y}_{t-1}-w^{\ast}\rangle+\beta_{t-1}^{2}\|(\overline{y}_{t-1}-w^{\ast})\|^{2}\\
	& \leq(1+\beta_{t-1})^{2}\|\overline{y}_{t}-w^{\ast}\|^{2}+2\beta_{t-1}(1+\beta_{t-1})\|\overline{y}_{t}-w^{\ast}\|\cdot\|\overline{y}_{t-1}-w^{\ast}\|+\beta_{t-1}^{2}\|(\overline{y}_{t-1}-w^{\ast})\|^{2}
	\end{align*}
	which gives 
	\begin{align*}
	\|\overline{y}_{t+1}-w^{\ast}\|^{2} & \leq(1-\alpha_{t}\mu)(1+\beta_{t-1})^{2}\|\overline{y}_{t}-w^{\ast}\|^{2}+2(1-\alpha_{t}\mu)\beta_{t-1}(1+\beta_{t-1})\|\overline{y}_{t}-w^{\ast}\|\cdot\|\overline{y}_{t-1}-w^{\ast}\|+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
	& +\beta_{t-1}^{2}(1-\alpha_{t}\mu)\|(\overline{y}_{t-1}-w^{\ast})\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}LG^{2}
	\end{align*}
	and we will using this recursive relation to obtain the desired bound.

	First we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}$.
	Since communication is done every $E$ steps, for any $t\geq0$, we
	can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $w_{t_{0}}^{k}=\overline{w}_{t_{0}}$for
	all $k$. Moreover, using $\alpha_{t}$ is non-increasing and $\alpha_{t_{0}}\leq2\alpha{}_{t}$
	for any $t-t_{0}\leq E-1$, we have 
	\begin{align*}
	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2} & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}-(\overline{w}_{t}-\overline{w}_{t_{0}})\|^{2}\\
	& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}\|^{2}\\
	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-w_{t_{0}}^{k}\|^{2}\\
	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\sum_{i=t_{0}}^{t-1}\beta_{i}(y_{i+1}^{k}-y_{i}^{k})-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\
	& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}\|g_{i,k}\|^{2}+2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\beta_{i}^{2}\|(y_{i+1}^{k}-y_{i}^{k})\|^{2}\\
	& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}(\|g_{i,k}\|^{2}+\|(y_{i+1}^{k}-y_{i}^{k})\|^{2})\\
	& \leq4\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}G^{2}\\
	& \leq4(E-1)^{2}\alpha_{t_{0}}^{2}G^{2}\leq16(E-1)^{2}\alpha_{t}^{2}G^{2}
	\end{align*}
	where we have used $\mathbb{E}\|\overline{y}_{t+1}-\overline{y}_{t}\|^{2}\leq G^{2}$.
	To see this identity for appropriate $\alpha_{t},\beta_{t}$, note
	the recursion 
	\begin{align*}
	y_{t+1}^{k}-y_{t}^{k} & =w_{t}^{k}-w_{t-1}^{k}-(\alpha_{t}g_{t,k}-\alpha_{t-1}g_{t-1,k})\\
	w_{t+1}^{k}-w_{t}^{k} & =-\alpha_{t}g_{t,k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k})
	\end{align*}
	so that 
	\begin{align*}
	y_{t+1}^{k}-y_{t}^{k} & =-\alpha_{t-1}g_{t-1,k}+\beta_{t-1}(y_{t}^{k}-y_{t-1}^{k})-(\alpha_{t}g_{t,k}-\alpha_{t-1}g_{t-1,k})\\
	& =\beta_{t-1}(y_{t}^{k}-y_{t-1}^{k})-\alpha_{t}g_{t,k}
	\end{align*}
	Since the identity $y_{t+1}^{k}-y_{t}^{k}=\beta_{t-1}(y_{t}^{k}-y_{t-1}^{k})-\alpha_{t}g_{t,k}$
	implies 
	\begin{align*}
	\mathbb{E}\|y_{t+1}^{k}-y_{t}^{k}\|^{2} & \leq2\beta_{t-1}^{2}\mathbb{E}\|y_{t}^{k}-y_{t-1}^{k}\|^{2}+2\alpha_{t}^{2}G^{2}
	\end{align*}
	as long as $\alpha_{t},\beta_{t}$ satisfy $2\beta_{t-1}^{2}+2\alpha_{t}^{2}\leq1$,
	we can guarantee that $\mathbb{E}\|y_{t}^{k}-y_{t-1}^{k}\|^{2}\leq G^{2}$
	for all $k$. This together with Jensen's inequality gives $\mathbb{E}\|\overline{y}_{t}-\overline{y}_{t-1}\|^{2}\leq G^{2}$
	for all $t$.

	Using the bound on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}$,
	we can conclude that, with $\nu_{max}:=N\cdot\max_{k}p_{k}$ and $\nu_{min}:=N\cdot\min_{k}p_{k}$,

	\begin{align*}
	\mathbb{E}\|\overline{y}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\overline{y}_{t}-w^{\ast}\|^{2}+16E^{2}L\alpha_{t}^{3}G^{2}+16E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{3}LG^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{y}_{t-1}-w^{\ast})\|^{2}\\
	& +\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{y}_{t}-w^{\ast}\|\cdot\|\overline{y}_{t-1}-w^{\ast}\|\\
	& \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\overline{y}_{t}-w^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{y}_{t-1}-w^{\ast})\|^{2}\\
	& +\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{y}_{t}-w^{\ast}\|\cdot\|\overline{y}_{t-1}-w^{\ast}\|
	\end{align*}
	where $\sigma^{2}=\max_{k}\sigma_{k}^{2}$, and $\alpha_{t}$ satisfies
	$L\alpha_{t}\leq\frac{1}{8}$. We show next that $\mathbb{E}\|\overline{y}_{t}-w^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{1}{t^{2}})$
	by induction.

	Assume that we have shown 
	\begin{align*}
	\mathbb{E}\|\overline{y}_{t}-w^{\ast}\|^{2} & \leq b(C\alpha_{t}^{2}+D\alpha_{t})
	\end{align*}
	where $C=20E^{2}LG^{2}$, $D=\frac{1}{N}\nu_{max}^{2}\sigma^{2}$,
	and $b$ is to be chosen later. For step sizes we choose $\alpha_{t}=\frac{9}{\mu}\frac{1}{t+\gamma}$
	and $\beta_{t-1}=\frac{9}{14\mu}\frac{1}{t+\gamma}$ if $\mu>1$ and
	$\frac{9}{14}\frac{1}{t+\gamma}$ if $\mu<1$, so that $\beta_{t-1}\leq\alpha_{t}$
	and $(1-\mu\alpha_{t})(1+14\beta_{t-1})^{2}\leq(1-\mu\alpha_{t})(1+\mu\alpha_{t})$.
	Moreover, $\|\overline{y}_{t-1}-w^{\ast}\|\leq2\|\overline{y}_{t}-w^{\ast}\|$.
	Recall that we require $\alpha_{t_{0}}\leq\alpha_{t}$ for any $t-t_{0}\leq E-1$,
	and $L\alpha_{t}\leq\frac{1}{8}$.

	Then the bound for $\mathbb{E}\|\overline{y}_{t+1}-w^{\ast}\|^{2}$
	can be further simplified with 
	\begin{align*}
	2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{y}_{t}-w^{\ast}\|\cdot\|\overline{y}_{t-1}-w^{\ast}\| & \leq4\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{y}_{t}-w^{\ast}\|^{2}
	\end{align*}
	and 
	\begin{align*}
	(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{y}_{t-1}-w^{\ast})\|^{2} & \leq4(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{y}_{t}-w^{\ast})\|^{2}
	\end{align*}
	Thus 
	\begin{align*}
	\mathbb{E}\|\overline{y}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})((1+\beta_{t-1})^{2}+4\beta_{t-1}(1+\beta_{t-1})+4\beta_{t-1}^{2})\|(\overline{y}_{t}-w^{\ast})\|^{2}\\
	& +20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\\
	& \leq\mathbb{E}(1-\mu\alpha_{t})(1+14\beta_{t-1})\|(\overline{y}_{t}-w^{\ast})\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\\
	& \leq b(1-\mu\alpha_{t})^{2}(C\alpha_{t}^{2}+D\alpha_{t})+C\alpha_{t}^{3}+D\alpha_{t}^{2}\\
	& =(b(1-\mu\alpha_{t})^{2}+\alpha_{t})\alpha_{t}^{2}C+(b(1-\mu\alpha_{t})^{2}+\alpha_{t})\alpha_{t}D
	\end{align*}
	and so it remains to choose $b$ such that 
	\begin{align*}
	(b(1-\mu\alpha_{t})^{2}+\alpha_{t})\alpha_{t} & \leq b\alpha_{t+1}\\
	(b(1-\mu\alpha_{t})^{2}+\alpha_{t})\alpha_{t}^{2} & \leq b\alpha_{t+1}^{2}
	\end{align*}

	We have 
	\begin{align*}
	(b(1-\mu\alpha_{t})^{2}+\alpha_{t})\alpha_{t} & =(b(1-\frac{4}{t+\gamma})^{2}+\frac{4}{\mu(t+\gamma)})\frac{4}{\mu(t+\gamma)}\\
	& =(b\frac{(t+\gamma-4)^{2}}{(t+\gamma)^{2}}+\frac{4}{\mu(t+\gamma)})\frac{4}{\mu(t+\gamma)}\\
	& =b(\frac{(t+\gamma-4)^{2}+t+\gamma}{(t+\gamma)^{2}})\frac{4}{\mu(t+\gamma)}\\
	& =b\frac{4}{\mu(t+\gamma+1)}
	\end{align*}
	where we have used 
	\begin{align*}
	\frac{(t+\gamma-4)^{2}+t+\gamma}{(t+\gamma)^{2}}\frac{1}{t+\gamma} & \leq\frac{1}{t+\gamma+1}
	\end{align*}
	Similarly $(b(1-\mu\alpha_{t})^{2}+\alpha_{t})\alpha_{t}^{2}\leq b\alpha_{t+1}^{2}$.
	It follows that $\mathbb{E}\|\overline{y}_{t}-w^{\ast}\|^{2}\leq b(C\alpha_{t}^{2}+D\alpha_{t})$
	for all $t$. Thus 
	\begin{align*}
	\mathbb{E}(F(\overline{y}_{T}))-F^{\ast} & =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa}{\mu^{2}}E^{2}LG^{2}\cdot\frac{1}{T^{2}})
	\end{align*}

	With partial participation, the update at each communication round
	is now given by averages over a subset of sampled devices. The key
	is to bound 
	\begin{align*}
	\mathbb{E}\|\overline{w}_{t}-\overline{y}_{t}\|^{2} & =\frac{1}{K}\sum_{k}p^{k}\|w_{t}^{k}-\sum_{k}p_{k}w_{t}^{k}\|^{2}
	\end{align*}
	This can be bounded by the same argument as before and yield 
	\begin{align*}
	\mathbb{E}\|\overline{w}_{t}-\overline{y}_{t}\|^{2} & \leq\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
	\end{align*}
	which results in the bound 
	\begin{align*}
	\mathbb{E}F(\overline{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
	\end{align*}
\end{proof}

### 2020-05-31 source/sec_Nesterov.tex
updated nesterov section and proofs

@@
 	$\gamma=\max\{32\kappa,E\}$ where $E$ is the communication delay,
-	and diminishing learning rates $\alpha_{t}=\beta_{t}=\frac{1}{4\mu(\gamma+t)}$.
-	Let $\overline{w}_{T}=\sum_{k=1}^{N}p_{k}w_{T}^{k}$ be the average
-	of local parameters at an arbitrary time $T$.
+	and diminishing learning rates $\alpha_{t}=\frac{9}{\mu}\frac{1}{t+\gamma}$
+	and $\beta_{t-1}=\frac{9}{14\max\{\mu,1\}}\frac{1}{t+\gamma}$. Let
+	$\overline{y}_{T}=\sum_{k=1}^{N}p_{k}y_{T}^{k}$ be the average of
+	local parameters at an arbitrary time $T$.
 
@@
 	Suppose $F_{k}$ is $L$-smooth and convex for all $k$, and let $\nu_{\max}=\max_{k}\frac{1}{N}p_{k}$.
-	Set learning rate $\alpha_{t}=O(\sqrt{\frac{N}{T}})$.
+	Set learning rate $\alpha_{t},\beta_{t}=O(\sqrt{\frac{N}{T}})$.
 
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
+	\min_{t\leq T}F(\overline{y}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
 	\end{align*}
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}G^{2}}{KT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
+	\min_{t\leq T}F(\overline{y}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}G^{2}}{KT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
 	\end{align*}

### 2020-05-31 source/sec_overparameterized.tex
streamline section 6

@@
 We can use $w^{\ast T}x_{k,j}-z_{k,j}\equiv0$ to rewrite the local
-objectives 
+objectives $F_{k}(w)=\frac{1}{2}\|w-w^{\ast}\|_{H^{k}}^{2}$ so that
+$F(w)=\frac{1}{2}\|w-w^{\ast}\|_{H}^{2}$. %
+\begin{comment}
 \begin{align*}
@@
 \end{align*}
-so that $F(w)=\frac{1}{2}\|w-w^{\ast}\|_{H}^{2}$.
+\end{comment}
 
@@
 FedAvg algorithm with Nesterov updates.
-
+\begin{comment}
 Also note that the update can equivalently be written as 
@@
 and $\overline{u}_{t+1}=\frac{\alpha^{k}}{1+\alpha^{k}}\overline{v}_{t+1}+\frac{1}{1+\alpha^{k}}\overline{w}_{t+1}$.
+\end{comment}
 

### 2020-05-31 source/sec_intro.tex
Update sec_intro.tex

@@
 \textbf{Notations}
-Throughout the paper, we adopt the following notation. Let $N$ be the total
+Throughout the paper, we adopt the following notations. Let $N$ be the total
 number of local devices, and $K \leq N$ is the maximal number of devices
@@
 More specifically, for strongly convex and smooth problems, we provide the {\small{$\cO(\frac{1}{NT})$}} convergence  for FedAvg and the local steps can be as large as {\small{$\cO(\sqrt{\frac{T}{N}})$}}.
-For convex and smooth objectives, we prove the {\small{$\cO(\sqrt{\frac{1}{NT}})$}} and the local steps can be as large as {\small{$\cO(\sqrt{\frac{T}{N}})$}}.
+For convex and smooth objectives, we prove the {\small{$\cO(\frac{1}{\sqrt{NT}})$}} and the local steps can be as large as {\small{$\cO(\sqrt{\frac{T}{N}})$}}.
 The same conclusions can be equally applied for Nesterov accelerated FedAvg.

### 2020-05-31 source/sec_Nesterov.tex
Nesterov section revision round 1 complete

--- 2020-05-30 source/sec_Nesterov.tex
@@
 local updates instead of local SGD updates. To our knowledge, the
-only formal analyses of accelerated FedAvg with momentum-based stochastic
-updates {[}CITE kdd papaer{]} focus on the non-convex smooth case.
-In this section, we complete the picture by providing convergence
-results of Nesterov-accelerated FedAvg for convex objectives. We show
-that for strongly convex and smooth objectives, the convergence of
-the optimality gap of averaged parameters across devices is $O(1/NT)$
-where $N$ is the number of active devices during each communication
-round, while for convex and smooth objectives, the rate is $O(1/\sqrt{NT})$.
-Thus Nesterov accelerated FedAvg algorithms enjoy the same convergence
-rates as the original FedAvg algorithm. As we know from stochastic
-optimization that Nesterov updates may fail to accelerate over SGD,
-in general we cannot hope to obtain acceleration results in the federated
-learning setting. However, in the following section we specialize
-to the overparameterized setting where we demonstrate that a particular
-accelerated FedAvg algorithm with momentum-based updates is able to
-improve on the original FedAvg algorithm.
+only convergence analyses of FedAvg with momentum-based stochastic
+updates focus on the non-convex smooth case \cite{yu2019linear,huo2020faster,li2018federated}.
+In this section, we complete the picture with $\mathcal{O}(1/NT)$
+and $\mathcal{O}(1/\sqrt{NT})$ convergence results for Nesterov-accelerated
+FedAvg for convex objectives. As we know from stochastic optimization,
+Nesterov updates may fail to accelerate over SGD, so in general we
+cannot hope to obtain acceleration results over FedAvg. However, in
+the following section we specialize to overparameterized problems
+where we demonstrate that a particular FedAvg variant with momentum
+updates is able to accelerate over the original FedAvg algorithm.
 
@@
 
-We first show that the Nesterov accelerated FedAvg has $O(1/NT)$
+We first show that the Nesterov accelerated FedAvg has $\mathcal{O}(1/NT)$
 convergence rate for $\mu$-strongly convex and $L$-smooth objectives.
@@
 \begin{align*}
-y_{t+1}^{k} & =w_{t}^{k}-\alpha_{t}g_{t,k}\\
-w_{t+1}^{k} & =\begin{cases}
-y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E}\\
-\sum_{k=1}^{N}p_{k}\left[y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E}
+\mathbf{v}_{t+1}^{k} & =\mathbf{w}_{t}^{k}-\alpha_{t}g_{t,k}\\
+\mathbf{w}_{t+1}^{k} & =\begin{cases}
+\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E}\\
+\sum_{k=1}^{N}p_{k}\left[\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E}
 \end{cases}
 \end{align*}
-where $g_{t,k}:=\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})$ is the stochastic
-gradient. 
+where $g_{t,k}:=\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})$ is
+the stochastic gradient sampled on the $k$-th device at time $t$.  
 \begin{theorem}
-	Assume that $F_{k}$ is $L$-smooth and $\mu$-strongly convex for
-	all $k$, and let $\nu_{\max}=\max_{k}\frac{1}{N}p_{k}$. Let $\kappa=\frac{L}{\mu}$,
-	$\gamma=\max\{32\kappa,E\}$ where $E$ is the communication delay,
-	and diminishing learning rates $\alpha_{t}=\beta_{t}=\frac{1}{4\mu(\gamma+t)}$.
-	Let $\overline{w}_{T}=\sum_{k=1}^{N}p_{k}w_{T}^{k}$ be the average
-	of local parameters at an arbitrary time $T$.
-
-	With full device participation, 
+	Let $\overline{\mathbf{v}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{v}_{T}^{k}$
+	be the average of local parameters of Nesterov accelerated FedAvg
+	with learning rates $\alpha_{t}=\frac{6}{\mu}\frac{1}{t+\gamma}$
+	and $\beta_{t-1}=\frac{3}{14(t+\gamma)(1-\frac{6}{t+\gamma})\max\{\mu,1\}}$,
+	$\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$. Then with
+	full device participation, 
 	\begin{align*}
-	\mathbb{E}F(\overline{y}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{y}_{T})-F^{\ast}\leq O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}\leq O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
 	\end{align*}
@@
 \begin{remark}
-To our knowledge, this is the first convergence result for Nesterov accelerated FedAvg in the convex and smooth setting. The same discussion about linear speedup and choice of $E$ from the
-	previous section applies to the Nesterov accelerated FedAvg algorithm. 
+To our knowledge, this is the first convergence result for Nesterov
+accelerated FedAvg in the strongly convex and smooth setting. The
+same discussion about linear speedup of FedAvg applies to the Nesterov
+accelerated variant. In particular, to achieve $\mathcal{O}(1/NT)$
+linear speedup, $T$ iterations of the algorithm requires only $\mathcal{O}(\sqrt{NT})$
+communication rounds with full participation. 
 \end{remark}
@@
 
-In this section, we provide linear speedup analyses of Nesterov Accelerated
-FedAvg with convex and smooth objectives and show that the optimality
-gap is $O(1/\sqrt{NT})$ where $N$ is the number of participating
+We now provide linear speedup analysis of Nesterov Accelerated FedAvg
+with convex and smooth objectives and show that the optimality gap
+is $\mathcal{O}(1/\sqrt{NT})$ where $N$ is the number of participating
 devices. This result complements the strongly convex case in the previous
-part, as well as the non-convex smooth setting in {[}CITE kdd and
-others{]}, where a similar $O(1/\sqrt{NT})$ rate is given in terms
+part, as well as the non-convex smooth setting in \cite{huo2020faster,yu2019linear,li2018federated},
+where a similar $\mathcal{O}(1/\sqrt{NT})$ rate is given in terms
 of averaged gradient norm. 
 \begin{theorem}
-	Suppose $F_{k}$ is $L$-smooth and convex for all $k$, and let $\nu_{\max}=\max_{k}\frac{1}{N}p_{k}$.
-	Set learning rate $\alpha_{t}=O(\sqrt{\frac{N}{T}})$.
-
-	With full device participation, 
+	Set learning rates $\alpha_{t}=\beta_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$
+	for Accelerated FedAvg with convex and $L$-smooth objectives. With
+	full device participation, 
 	\begin{align*}
-	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F^{\ast} & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
 	\end{align*}
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}G^{2}}{KT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F^{\ast} & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}G^{2}}{KT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
 	\end{align*}
@@
 \begin{remark}
-	It is possible to extend the analysis to accelerated FedAvg algorithms
-	with other momentum-based updates. However, in line with the stochastic
-	optimization setting, none of these methods will be able to achieve
-	a better rate than the original FedAvg algorithm in general. For this
-	reason, we will instead turn to the overparameterized setting {[}CITE
-	Belkin{]} where we show that it is possible to improve the convergence
-	rate of FedAvg with momentum based local updates. 
+	It is possible to extend the analyses in this section to accelerated
+	FedAvg algorithms with other momentum-based updates. However, in the
+	stochastic optimization setting, none of these methods will be able
+	to achieve a better rate than the original FedAvg with SGD in general.
+	For this reason, we will instead turn to the overparameterized setting
+	\cite{ma2017power,liu2018accelerating,canziani2016analysis} in the
+	next section where we show that FedAvg enjoys exponential convergence
+	and it is possible to improve its convergence rate with momentum based
+	local updates. 
 \end{remark}

### 2020-05-31 source/sec_relatedwork.tex
related work

--- 2020-05-30 source/sec_relatedwork.tex
@@
 
+% Version 1.
 % related work on convergence rate
-Federated learning (FL) was originally proposed
-in~\cite{mcmahan2016communication} for learning a single, global statistical
-model from the isolated data stored in a massive number of devices.  The
-empirical success of FL~\cite{chen2018federated,47586} has attracted much attention to the rigorous theoretical understanding of the leading algorithm in the field: Federated Averaging (FedAvg) and its accelerated variants~\cite{liu2019accelerating,haddadpour2019convergence,khaled2019first,li2019convergence,huo2020faster,yu2019linear,yu2019parallel}.
+% Federated learning (FL) was originally proposed
+% in~\cite{mcmahan2016communication} for learning a single, global statistical
+% model from the isolated data stored in a massive number of devices.  The
+% empirical success of FL~\cite{chen2018federated,47586} has attracted much attention to the rigorous theoretical understanding of the leading algorithm in the field: Federated Averaging (FedAvg) and its accelerated variants~\cite{liu2019accelerating,haddadpour2019convergence,khaled2019first,li2019convergence,huo2020faster,yu2019linear,yu2019parallel}.
 
-% Non-convex linear speedup
-In particular, we focus on the convergence analysis of FedAvg, which
-has been discussed in various settings. 
-In~\cite{yu2019parallel,wang2019adaptive}, the authors provide the convergence rate of FedAvg for the non-convex problem, given all the devices are active at each communication round.
-Recently, \citep{huo2020faster} provide
-convergence guarantee for Nesterov accelerated FedAvg for non-convex 
-setting in a partial participation setting. Although the empirical
-results are improved over FedAvg, the accelerated version admitting a similar convergence rate as FedAvg. In this work, we provide a novel accelerated
-FedAvg with improved convergence rates under a popular overparameterized setting. In addition, \cite{yu2019linear} provides $O(1/\sqrt{NT})$ 
-convergence (i.e., linear speedup) of FedAvg for the non-convex problem in the full participation setting. 
-For convex smooth problems, \cite{khaled2019first} prove the 
-convergence rate of local GD (a full batch version of FedAvg with full participation) on heterogeneous data. 
-Concurrently, \cite{li2019convergence} firstly provides exponential convergence of FedAvg for the strongly convex problem,
-under the more practical setting of partial participation and Non-IID data while their results did not achieve linear speedup. 
-\cite{stich2018local} prove linear speedup convergence of FedAvg for the strongly convex problem, while they assume full participation and identically distributed data.
+% % Non-convex linear speedup
+% In particular, we focus on the convergence analysis of FedAvg, which
+% has been studied in various scenarios. 
+% In~\cite{yu2019parallel,wang2019adaptive}, the authors provide the convergence rate of FedAvg for the non-convex problem, given all the devices are active at each communication round.
+\cite{haddadpour2019convergence}
+% improves the convergence analysis for non-convex problem under the setting of partial participation. 
+% In addition, \cite{yu2019linear} provides $O(1/\sqrt{NT})$ 
+% convergence (i.e., linear speedup) of FedAvg for the non-convex problem in the full participation setting. 
+% For convex smooth problems, \cite{khaled2019first} prove the 
+% convergence rate of local GD (a full batch version of FedAvg with full participation) on heterogeneous data. 
+% Concurrently, \cite{li2019convergence} provides exponential convergence of FedAvg for the strongly convex problem,
+% under the more practical setting of partial participation and Non-IID data while their results did not achieve linear speedup. 
+% \cite{stich2018local} prove linear speedup convergence of FedAvg for the strongly convex problem, while they assume full participation and identically distributed data.
 Comparing to previous works, our results provide the linear speedup on strongly convex problems in the practical partial participation setting and heterogeneous data.
@@
 To the best of our knowledge,
-we provide the first linear speedup convergence results for accelerated FedAvg in the practical setting.
+% we provide the first linear speedup convergence results for accelerated FedAvg in the practical setting.
+
+\textbf{The Convergence of FedAvg and Linear Speedup}
+There has been a vast amount of convergence analysis for Federated averaging
+(FedAvg) in different settings. However, most of the prior works provide the
+convergence guarantee under the simplified setting, where either statistical
+heterogeneity~\cite{stich2018local} or system
+heterogeneity~\cite{yu2019parallel,wang2019adaptive,khaled2019first} are not
+considered. This simplified setting is also referred as local SGD. 
+In~\cite{li2019convergence},
+the authors firstly provide convergence of FedAvg for strongly convex problem considering both statistical heterogeneity and system heterogeneity, while their results
+didn't achieve linear speedup. In~\cite{haddadpour2019convergence,liang2019variance,huo2020faster}, the convergence of FedAvg for non-convex problems has been studied. 
+However, \cite{huo2020faster} did not discuss the linear speedup and the theorems in \cite{haddadpour2019convergence,liang2019variance} failed to cover the accelerated version of
+FedAvg.
+
+\textbf{Federated Learning Beyond Local SGD}
+% The accelerated SGD methods has achieved optimal convergence rate in strongly convex
+% optimization~\cite{nesterov1983method}.
+In \cite{liu2019accelerating}, the authors provide a convergence rate of accelerated FedAvg for strongly convex problems, while it is considered in full participation setting.
+\cite{yu2019linear} and \citep{huo2020faster} provide
+convergence guarantee for Nesterov accelerated FedAvg for non-convex. 
+However, \cite{yu2019linear} did not consider the system heterogeneity and \cite{huo2020faster} haven't discuss the linear speedup.
+Furthermore, the accelerated FedAvg admitting a similar convergence rate as FedAvg in all previous works.
+In this work, we provide a novel accelerated FedAvg with improved convergence rates under a popular overparameterized setting.

### 2020-05-31 source/sec_conclude.tex
Update sec_conclude.tex

--- 2020-05-30 source/sec_conclude.tex
@@
 This paper provides a comprehensive analysis of the convergence rate of FedAvg
-and its accelerated variants in a general federated learning problem with heterogeneous local data and partial participation. We show that both accelerated FedAvg with Nesterov updates and FedAvg
-can achieve linear speedup when the number of nodes increases, i.e., $O(1/\sqrt{NT})$
-convergence for convex smooth problems and $O(1/NT)$ convergence for strongly 
-convex smooth problems. Furthermore, we show that FedAvg can in fact achieve exponential 
-convergence for overparameterized strongly convex smooth problems, and we propose the MaSS accelerated Fedavg algorithm, which has provable speedup in convergence rate over
-FedAvg on quadratic problems. Last but not least, we empirically
+and its accelerated variants in a general federated learning problem with heterogeneous local data and partial participation. We show that both Nesterov accelerated FedAvg and FedAvg
+can achieve {\small{$\cO(\frac{1}{\sqrt{NT}})$}} linear speedup convergence for convex smooth problems and {\small{$\cO(\frac{1}{NT})$}} convergence for strongly 
+convex smooth problems. In addition, we show that the local steps for stronlgy convex and convex smoth problems can be as large as {\small{$\cO(\sqrt{\frac{T}{N}})$}} and 
+{\small{$\cO(\sqrt{\frac{T}{N}})$}}, respectively, which can save communication cost substantially comparing to prior results. 
+Furthermore, this work also makes algorithmic contributions.
+We not only prove that FedAvg can achieve exponential convergence for overparameterized strongly convex smooth problems, but also propose the MaSS accelerated Fedavg algorithm, which has provable speedup in convergence rate over FedAvg on quadratic problems.
+Last but not least, we empirically
 verify the linear speedup of FedAvg and Nesterov accelerated FedAvg for strongly convex, convex smooth, and linear regression problems. The empirical results are well-aligned with our theories.

### 2020-05-31 source/sec_conclude.tex
Update sec_conclude.tex

@@
 can achieve {\small{$\cO(\frac{1}{\sqrt{NT}})$}} linear speedup convergence for convex smooth problems and {\small{$\cO(\frac{1}{NT})$}} convergence for strongly 
-convex smooth problems. In addition, we show that the local steps for stronlgy convex and convex smoth problems can be as large as {\small{$\cO(\sqrt{\frac{T}{N}})$}} and 
+convex smooth problems. In addition, we show that the local steps for stronlgy convex and convex smooth problems can be as large as {\small{$\cO(\sqrt{\frac{T}{N}})$}} and 
 {\small{$\cO(\sqrt{\frac{T}{N}})$}}, respectively, which can save communication cost substantially comparing to prior results. 

### 2020-05-31 source/sec_conclude.tex
Update sec_conclude.tex

--- 2020-05-31 source/sec_conclude.tex
@@
 can achieve {\small{$\cO(\frac{1}{\sqrt{NT}})$}} linear speedup convergence for convex smooth problems and {\small{$\cO(\frac{1}{NT})$}} convergence for strongly 
-convex smooth problems. In addition, we show that the local steps for stronlgy convex and convex smoth problems can be as large as {\small{$\cO(\sqrt{\frac{T}{N}})$}} and 
-{\small{$\cO(\sqrt{\frac{T}{N}})$}}, respectively, which can save communication cost substantially comparing to prior results. 
+convex smooth problems.
+In addition, we show that the local steps for stronlgy convex and convex smooth problems can be as large as {\small{$\cO(\sqrt{\frac{T}{N}})$}}, which can save communication cost substantially comparing to prior results.
 Furthermore, this work also makes algorithmic contributions.

### 2020-05-31 source/sec_intro.tex
intro notations

@@
 Let $T$ be the total number of  Stochastic Gradient Descent (SGD) performed on each local device, $E$ be the local steps between two communication steps.
-Thus, $\frac{T}{E}$ is
+Thus, $T/E$ is
 the number of communications.
@@
 target accuracy and preserving the linear speedup at the same time.
-More specifically, for strongly convex and smooth problems, we provide the {\small{$\cO(\frac{1}{NT})$}} convergence  for FedAvg and the local steps can be as large as {\small{$\cO(\sqrt{\frac{T}{N}})$}}.
-For convex and smooth objectives, we prove the {\small{$\cO(\frac{1}{\sqrt{NT}})$}} and the local steps can be as large as {\small{$\cO(\sqrt{\frac{T}{N}})$}}.
+More specifically, for strongly convex and smooth problems, we provide the {\small{$\cO(1/NT)$}} convergence  for FedAvg and the local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}.
+For convex and smooth objectives, we prove the {\small{$\cO(1/\sqrt{NT})$}} and the local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}.
 The same conclusions can be equally applied for Nesterov accelerated FedAvg.

### 2020-05-31 source/sec_Nesterov.tex
updated Nesterov section

@@
 \begin{theorem}
-	Let $\overline{\mathbf{v}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{v}_{T}^{k}$
+	\label{thm:nesterov_scvx}Let $\overline{\mathbf{v}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{v}_{T}^{k}$
 	be the average of local parameters of Nesterov accelerated FedAvg
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}\leq O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}\leq\mathcal{O}(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
 	\end{align*}
@@
 \begin{theorem}
-	Set learning rates $\alpha_{t}=\beta_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$
+	\label{thm:Nesterov_cvx}Set learning rates $\alpha_{t}=\beta_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$
 	for Accelerated FedAvg with convex and $L$-smooth objectives. With
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F^{\ast} & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F^{\ast} & =\mathcal{O}(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
 	\end{align*}
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F^{\ast} & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}G^{2}}{KT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F^{\ast} & =\mathcal{O}(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}G^{2}}{KT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
 	\end{align*}

### 2020-05-31 source/sec_relatedwork.tex
updated reference

@@
 convergence guarantee under the simplified setting, where either statistical
-heterogeneity~\cite{stich2018local} or system
-heterogeneity~\cite{yu2019parallel,wang2019adaptive,khaled2019first} are not
+heterogeneity~\cite{stich2018local,zhou2017convergence,wang2018cooperative,woodworth2018graph} or system
+heterogeneity~\cite{yu2019parallel,wang2019adaptive,khaled2019first,jiang2018linear} are not
 considered. This simplified setting is also referred as local SGD. 
@@
 the authors firstly provide convergence of FedAvg for strongly convex problem considering both statistical heterogeneity and system heterogeneity, while their results
-didn't achieve linear speedup. In~\cite{haddadpour2019convergence,liang2019variance,huo2020faster}, the convergence of FedAvg for non-convex problems has been studied. 
+didn't achieve linear speedup. In~\cite{haddadpour2019convergence,liang2019variance,huo2020faster,jiang2018linear}, the convergence of FedAvg for non-convex problems has been studied. 
 However, \cite{huo2020faster} did not discuss the linear speedup and the theorems in \cite{haddadpour2019convergence,liang2019variance} failed to cover the accelerated version of

### 2020-06-01 source/sec_problem.tex
experiments appendix details
Nesterov section revision round 1 complete
update setup and additional notations

@@
 \begin{align}
-\vv_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}g_{t,k}\\
+\vv_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}\vg_{t,k}\\
 \mathbf{w}_{t+1}^{k} &=\left\{
@@
 local devices, it is more realistic to consider only a subset of 
-local devices is active at each communication round. In this work,
+local devices is active at each communication round (system heterogeneity). In this work,
 we consider both the case of \textbf{full participation} where the model is
@@
 the case of \textbf{partial participation} where $|\cS_{t+1}| < N$. 
-In the partial participation scenario, $\cS_{t+1}$ is obtained by two types of
-sampling schemes to simulate the practical scenarios~\cite{li2019convergence}. The first one samples
-each device with probability $p_k$, with replacement. The second one samples
-each device uniformly without replacement. Both of these sampling schemes
+In the partial participation scenario, $\cS_{t+1}$ is obtained by two types
+sampling schemes to simulate practical scenarios~\cite{li2019convergence}. 
+The sampling scheme I establishes $\cS_{t+1}$ by i.i.d. sampling the devices according to
+the weight $p_k$ with replacement. 
+The sampling scheme II establishes $\cS_{t+1}$ by uniformly sampling all devices without
+replacement. Both of the sampling schemes
 guarantee that gradient updates in FedAvg are unbiased stochastic versions of
-updates in FedAvg with full participation, during the communication round.
+updates in FedAvg with full participation. For 
+more details on the notations and setup, please also refer to the Section~\ref{sec:app:notations} in the appendix.
 
@@
 The variance of stochastic gradients in each device is bounded: $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq \sigma_{k}^{2}$,
-	for $k=1, \cdots, N$ and any $\mathbf{w}_{t}^{k}$.
+	for $k=1, \cdots, N$ and any $\mathbf{w}_{t}^{k}$. Let $\sigma^2=\max_k\sigma_{k}^{2}$.
 	\label{ass:boundedvariance}

### 2020-06-01 source/main.tex
Intro setup v0
add exp appendix
add conclusion
changed file names
Updated FedAvg with SGD results and discussions
related work v0
Update on Overleaf.
test overleaf change

@@
 
-\title{Blessing of FedAvg's Linear Speedup in Federated Learning}
+\title{The Blessing of FedAvg's Linear Speedup in Federated Learning}
 
@@
 \begin{abstract}
-Federated learning is a distributed learning paradigm that enables learning
-a model jointly over a large number of workers without sharing the private
-data on local workers. The characteristics of non-identically distributed data
-across the network and low device participation brings significant challenges
-in analyzing the convergence of federated learning algorithms. Prior works
-have either analyzed the convergence with i.i.d. data, or with full device
-participation, and often with suboptimal rates. In this work, we provide a
-comprehensive analysis of the convergence rate of Federated Averaging (FedAvg)
-with SGD and Nesterov SGD updates for general convex problems with arbitrarily
-heterogeneous data. Firstly we show that the convergence rates enjoy a linear
-speedup in the number of participating workers for all convex smooth problems
-for both algorithms.
-Furthermore, in the overparamterized setting, we prove that the convergence under FedAvg is in fact exponential, also with a speedup factor linear in the number of workers.
-We also propose a new accelerated federated learning algorithm that further improves this convergence rate with provable guarantees in the linear regression setting.
-Empirical evaluation of the algorithms in various settings match our theoretical results.
+Federated learning learns
+a model jointly from a set of distributed workers without sharing their private
+data. The characteristics of non-\emph{i.i.d.} data
+across the network and low device participation bring significant challenges
+in understanding the convergence of federated learning algorithms. Prior work
+analyzed the convergence with restrained settings,
+%either \emph{i.i.d.} data, or with full device participation, 
+and often with suboptimal rates. 
+In this work, we provide a
+comprehensive convergence analysis of the widely used framework Federated Averaging (FedAvg),
+using SGD or Nesterov SGD updates, for general convex problems with arbitrarily
+heterogeneous data. 
+We first show that the convergence rates of both algorithms enjoy a linear speedup w.r.t. the number of participating workers for all convex smooth problems. 
+Furthermore, in the overparamterized setting, we show that the convergence of FedAvg is in fact exponential, also with a speedup factor linear in the number of workers. 
+We propose a new accelerated federated learning algorithm that further improves this convergence rate with provable guarantees in the linear regression setting. 
+Empirical studies of the algorithms in various settings have supported our theoretical results.
 \end{abstract}
@@
 \input{sec_intro}
+\input{sec_relatedwork}
 \input{sec_problem}
-\input{sec_strongly_convex_smooth}
-\input{sec_convex_smooth}
-\input{sec_linear_regression}
+\input{sec_SGD}
+\input{sec_Nesterov}
+\input{sec_overparameterized}
 \input{sec_exp}
+
+\input{sec_conclude}
 
@@
 \end{document}
+
+\iffalse
+
+\endif

### 2020-06-01 source/sec_SGD.tex
Update on Overleaf.
updated SGD section 4

@@
 
-\section{Linear Speedup Analysis of FedAvg }
+\section{Linear Speedup Analysis of FedAvg}
 
-In this section, we provide convergence analysis of FedAvg with local
-SGD updates. We show that for strongly convex and smooth objectives,
+In this section, we provide convergence analyses of FedAvg with local
+SGD updates in the general setting with heterogeneous data and partial
+participation. We show that for strongly convex and smooth objectives,
 the convergence of the optimality gap of averaged parameters across
-devices is $O(1/NT)$ where $N$ is the number of active devices during
-each communication round, while for convex and smooth objectives,
-the rate is $O(1/\sqrt{NT})$.
+devices is $\mathcal{O}(1/NT)$ where $N$ is the number of active
+devices during each communication round, while for convex and smooth
+objectives, the rate is $\mathcal{O}(1/\sqrt{NT})$.
 
@@
 
-We first show that FedAvg with local SGD updates has $O(1/NT)$ convergence
-rate for $\mu$-strongly convex and $L$-smooth objectives. The result
-improves on the $O(1/T)$ result of {[}ICLR{]} with a linear speedup
-in the number of devices $N$. Moreover, it implies a distinction
-in the choice of $E$ that guarantees this linear speedup for FedAvg
-with full and partial device participation. With full participation,
-$E$ can be chosen as large as $O(\sqrt{\frac{T}{N}})$ without degrading
-the linear speedup in the number of workers. On the other hand, with
-partial participation, $E$ must be $O(1)$. 
+We first show that FedAvg with local SGD updates has $\mathcal{O}(1/NT)$
+convergence rate for $\mu$-strongly convex and $L$-smooth objectives.
+The result improves on the $\mathcal{O}(1/T)$ result of \cite{li2019convergence}
+with a linear speedup in the number of devices $N$. Moreover, it
+implies a distinction in communication efficiency that guarantees
+this linear speedup for FedAvg with full and partial device participations.
+With full participation, $E$ can be chosen as large as $\mathcal{O}(\sqrt{T/N})$
+without degrading the linear speedup in the number of workers. On
+the other hand, with partial participation, $E$ must be $\mathcal{O}(1)$. 
 \begin{theorem}
-	Suppose $F_{k}$ is $L$-smooth and $\mu$-strongly convex for all
-	$k$, and let $\nu_{\max}=\max_{k}\frac{1}{N}p_{k}$. Let $\kappa=\frac{L}{\mu}$,
-	$\gamma=\max\{32\kappa,E\}$ where $E$ is the communication delay,
-	and diminishing learning rates $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$.
-	Let $\overline{w}_{T}=\sum_{k=1}^{N}p_{k}w_{T}^{k}$ be the average
-	of local parameters at an arbitrary time $T$.
-
-	With full device participation, 
+	\label{thm:SGD_scvx}Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$
+	be the average of local parameters of FedAvg with learning rates $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$,
+	where $\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$. Then
+	with full device participation, 
 	\begin{align*}
-	\mathbb{E}F(\overline{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}\leq\mathcal{O}(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
 	\end{align*}
+	where $\nu_{\max}=\max_{k}Np_{k}$.
 \end{theorem}
@@
 \begin{remark}
-	\textbf{Linear speedup. }We first compare our bound with that in {[}ICLR{]},
-	which is $O(\frac{1}{NT}\nu_{\max}^{2}\sigma^{2}\kappa/\mu+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{E^{2}G^{2}}{T}\kappa/\mu)$.
-	Because the third term $\frac{E^{2}G^{2}}{T}\kappa/\mu$ is also $O(1/T)$
+	\textbf{Linear speedup. }We first compare our bound with that in \cite{li2019convergence},
+	which is $\mathcal{O}(\frac{1}{NT}\nu_{\max}^{2}\sigma^{2}\kappa/\mu+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{E^{2}G^{2}}{T}\kappa/\mu)$.
+	Because the third term $\frac{E^{2}G^{2}}{T}\kappa/\mu$ is also $\mathcal{O}(1/T)$
 	without a dependence on $N$, for any choice of $E$ their bound cannot
 	achieve linear speedup. The improvement of our bound comes from the
-	term $\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}$, which now is $O(1/T^{2})$.
+	term $\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}$, which now is $\mathcal{O}(1/T^{2})$.
 	As a result, all leading terms scale with $1/N$ in the full device
@@
 \begin{remark}
-	\textbf{Choice of $E$.} Our bound implies a distinction in the choice
-	of $E$ between the full and partial participation settings. In full
-	participation, as long as $E=O(\sqrt{\frac{T}{N}})$, the term $\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}=O(\frac{1}{NT})$,
-	which is on the same order as the leading term $\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}$.
-	Thus to achieve a linear speedup in the number of workers $N$, it
-	suffices to communicate every $O(\sqrt{\frac{T}{N}})$ iterations.
-	In contrast, the bound in {[}ICLR{]} does not allow $E$ to scale
-	with $\sqrt{T}$, even for full participation. Thus our bound yields
-	a more efficient communication complexity, while at the same time
-	providing a linear speedup in the convergence. On the other hand,
-	with partial participation, the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$
-	is also a leading term, and so $E$ must be $O(1)$. In this setting,
-	our bound still yields a linear speedup in $K$, which is not the
-	case for previous analyses. 
+	\textbf{Communication Complexity.} Our bound implies a distinction
+	in the choice of $E$ between the full and partial participation settings.
+	With full participation there is linear speedup $\mathcal{O}(1/N)$
+	as long as $E=\mathcal{O}(\sqrt{T/N})$. This corresponds to a communication
+	complexity of $\mathcal{O}(\sqrt{NT})$. In contrast, the bound in
+	\cite{li2019convergence} does not allow $E$ to scale with $\sqrt{T}$,
+	even for full participation. On the other hand, with partial participation,
+	the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$ is also a leading term,
+	and so $E$ must be $\mathcal{O}(1)$. In this setting, our bound
+	still yields a linear speedup in $K$, which is not the case for previous
+	analyses. 
+\end{remark}
+%
+\begin{remark}
+	\textbf{Comparison with related works.} There have been a line of
+	works that demonstrate that linear speedup is possible in distributed
+	optimization with infrequent communication. In particular, \cite{haddadpour2019convergence}
+	is the only other work that demonstrates linear speedup for strongly
+	convex and smooth objectives in the heterogeneous data with partial
+	participation setting, with $\mathcal{O}(N^{1/3}T^{2/3})$ communication
+	complexity. However, their results are derived based on a gradient
+	diversity assumption, and the learning rates are carefully chosen
+	based on an upper bound of the gradient diversity. Thus our results
+	are not directly comparable as we use the gradient bound assumption
+	instead, and our learning rate does not depend on the bound $G$.
 \end{remark}
@@
 \subsection{Convex Smooth Objectives}
-
-In this section, we provide linear speedup analyses of FedAvg with
-convex and smooth objectives and show that the optimality gap is $O(1/\sqrt{NT})$
+Next we provide linear speedup analyses of FedAvg with convex and
+smooth objectives and show that the optimality gap is $\mathcal{O}(1/\sqrt{NT})$
 where $N$ is the number of participating devices. This result complements
-the strongly convex case in the previous part, as well as the non-convex
-smooth setting in {[}CITE Yu and others{]}, where a similar $O(1/\sqrt{NT})$
-rate is given in terms of averaged gradient norm. 
+the strongly convex case in the previous part, as well as the nonconvex
+smooth setting in \cite{jiang2018linear,yu2019parallel,haddadpour2019convergence},
+where $\mathcal{O}(1/\sqrt{NT})$ results are given in terms of averaged
+gradient norm. 
 \begin{theorem}
-	Suppose $F_{k}$ is $L$-smooth and convex for all $k$, and let $\nu_{\max}=\max_{k}\frac{1}{N}p_{k}$.
-	Set learning rate $\alpha_{t}=O(\sqrt{\frac{N}{T}})$. With full device participation, 
+	\label{thm:SGD_cvx}With full device participation and constant learning
+	rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$, 
 	\begin{align*}
-	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T})
 	\end{align*}
 	and with partial device participation with $K$ sampled devices at
-	each communication round, 
+	each communication round and learning rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{K}{T}})$,
 	\begin{align*}
-	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}G^{2}}{KT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T})
 	\end{align*}
 \end{theorem}
-
+%
 \begin{remark}
-	\textbf{Choice of $E$ and linear speedup. }Similar in the strongly
-	convex case, we see that in the full participation setting, as long
-	as $E=O(\sqrt{\frac{T^{1/3}}{N}})$, the convergence benefits from
-	a linear speedup in the number of active workers. In the partial participation
-	setting, $E$ must be $O(1)$ in order to allow for linear speedup.
+	\textbf{Choice of $E$ and linear speedup. }With full participation,
+	as long as $E=\mathcal{O}(\sqrt{T^{1/2}/N^{3/2}})$, the convergence
+	rate is $\mathcal{O}(1/\sqrt{NT})$ with $\mathcal{O}(N^{3/4}T^{3/4})$
+	communication rounds. In the partial participation setting, $E$ must
+	be $O(1)$ in order to achieve linear speedup of $\mathcal{O}(1/\sqrt{KT})$.
+	Our result again demonstrates the difference in communication efficiencies
+	between full and partial participation, and is to our knowledge the
+	first result on linear speedup in the general federated learning setting
+	with both heterogeneous data and partial participation. 
 \end{remark}
@@
 	horizon $T$ of the convergence statement, whereas before the learning
-	rate was set to $O(1/t)$ for the $t$-th iteration. We also note
-	that the requirement on $\alpha_{t}=O(\sqrt{\frac{N}{T}})$ is also
-	present in the work of {[}Yu{]} where they derive similar $O(1/\sqrt{NT})$
-	linear speedup convergence results for nonconvex smooth federated
-	learning problems.
+	rate was set to $\mathcal{O}(1/t)$ for the $t$-th iteration. We
+	also note that the requirement on $\alpha_{t}=\mathcal{O}(\sqrt{N/T})$
+	is present in the works \cite{haddadpour2019convergence,yu2019parallel}
+	on nonconvex problems with $O(1/\sqrt{NT})$ linear speedup convergence
+	results. 
 \end{remark}
-For non-smooth objectives, we also have convergence results with rates
-$O(1/T)$ and $O(1/\sqrt{T})$ for strongly convex and convex cases,
-which can be found in the appendix. Without smoothness, however, the
-linear speedup is no longer guaranteed. Together these results complete
-the picture of convergence analyses for FedAvg with local SGD updates
-for general convex heterogeneous objectives. We next investigate the
-convergence of FedAvg with Nesterov updates, and provide similar linear
-speedup analyses for convex smooth objectives.

### 2020-06-01 source/sec_intro.tex

@@
 The widespread usage of smart home devices and mobile phones constitutes
-large-scale distributed networks that generate datasets of unprecedented size.
-The emergence of those massive amounts of data has empowered machine learning
-solutions to develop many successful applications such as personalized
+large-scale distributed networks that generate amount of personal data.
+Models learned from such data lead to 
+% The emergence of those massive amounts of data has empowered machine learning
+% solutions to develop 
+many successful applications such as personalized
 recommendations~\cite{chen2018federated}, keyboard predictions~\cite{47586},
-etc. Due to the increasing awareness of the privacy and rapid development of
-mobile devices' computational power, it becomes more and more attractive to
-push the computation to the edge. Recent years have witnessed the rise of
-federated learning~\cite{mcmahan2016communication} to full fill this demand.
+and etc. Due to the rising awareness of the privacy and rapid growth of
+computational power in mobile devices, it becomes increasingly attractive to
+push the computation to the edge. Recent years witnessed the rise of
+federated learning~\cite{mcmahan2016communication} to fulfill this demand.
 
@@
 learns a single statistical model leveraging the data from all devices while
-keeping the training data isolated.  Thus, FL can integrate cross-device
-information for acquiring a better model  without compromising the users'
+keeping the training data isolated.  Thus, FL integrates cross-device
+information for acquiring improved models without transferring training data from individual devices, thus improves user
 privacy~\cite{kairouz2019advances}. Comparing to the traditional distribution optimization,  FL confronts unique
-challenges~\cite{li2019convergence,smith2017federated}, including statistical heterogeneity: the data distribution in each local device can be significantly different, and system heterogeneity: only a subset of devices are accessible to the central server for model training.
+challenges~\cite{li2019convergence,smith2017federated}, including statistical heterogeneity: the data distributions in different devices can be significantly different, and system heterogeneity: only a subset of devices may be accessible to the central server for model training at each time.
 
 % Gaps in the theoretical understanding. 
-Towards resolving the aforementioned challenges, Federated
+To address the aforementioned challenges, Federated
 Averaging~\cite{mcmahan2016communication} (FedAvg) was proposed as an
-effective heuristic and its great empirical success~\cite{47586} has attracted
-lots of efforts on building up the theoretical understanding of FedAvg
-and its accelerated variants. 
-Roughly, the current convergence guarantee of FedAvg obtained by prior arts either established on simplified setting~\cite{stich2018local,khaled2019first,wang2018cooperative,yu2019parallel} or proved a suboptimal
+effective heuristic and its great empirical success~\cite{47586} attracted
+lots of efforts on building up its theoretical understanding 
+and accelerated variants. 
+Existing convergence guarantees of FedAvg were largely established on simplified setting~\cite{stich2018local,khaled2019first,wang2018cooperative,yu2019parallel} or provides a suboptimal 
 convergence rate~\cite{li2019convergence}. Furthermore, the convergence of
-FedAvg's accelerated variants~\cite{yu2019linear,huo2020faster,liu2019accelerating} under the realistic setting is still waiting to be solved.
-
-\textbf{Notations}
-Throughout the paper, we adopt the following notations. Let $N$ be the total
-number of local devices, and $K \leq N$ is the maximal number of devices
-that are accessible to the central server.
-Let $T$ be the total number of  Stochastic Gradient Descent (SGD) performed on each local device, $E$ be the local steps between two communication steps.
-Thus, $T/E$ is
-the number of communications.
+FedAvg's accelerated variants~\cite{yu2019linear,huo2020faster,liu2019accelerating} under the realistic setting is yet to be analyzed.
 
 % Our contributions
-\subsection{Contributions}
+% \subsection{Contributions}
+\textbf{Contributions.}
 This work focus on providing comprehensive convergence analysis
-of FedAvg and accelerated variants under the realistic setting, showing
-the less number of communication round is needed to converge to the
-target accuracy and preserving the linear speedup at the same time.
-More specifically, for strongly convex and smooth problems, we provide the {\small{$\cO(1/NT)$}} convergence  for FedAvg and the local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}.
-For convex and smooth objectives, we prove the {\small{$\cO(1/\sqrt{NT})$}} and the local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}.
-The same conclusions can be equally applied for Nesterov accelerated FedAvg.
-To the best of our knowledge, this is the first linear speedup results for FedAvg with Nesterov accelerated SGD.
-Furthermore,
-we also provide the first improved convergence rate for Nesterov accelerated FedAvg comparing to FedAvg under a popular overparamterized setting. 
+of FedAvg and its accelerated variants under a realistic setting.
+Our results show
+that the less number of communication rounds is needed to converge to a
+target accuracy and preserve the linear speedup w.r.t. participating devices at the same time.
+More specifically, for strongly convex and smooth problems, we show 
+that FedAvg admits a convergence rate of 
+{\small{$\cO(1/NT)$}} convergence, and local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}.
+For convex and smooth objectives, we show the rate of {\small{$\cO(1/\sqrt{NT})$}} and local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}.
+Moreover, we show that FedAvg with Nesterov acceleration also enjoys the same rates and speedup. 
+%The same conclusions are equally applied for Nesterov accelerated FedAvg. 
+To the best of our knowledge, this is the first linear speedup results for FedAvg with Nesterov accelerated SGD. Furthermore, 
+we provide the first improved convergence rate for Nesterov accelerated FedAvg comparing to FedAvg under the popular overparamterized setting. 
 Last but not least, we systematically examine the convergence rate of
@@
 % evaluation on both synthetic and real-world dataset, which demonstrates that our theoretical indications is well-aligned with the empirical observations.
+
+\textbf{Notations.}
+Throughout the paper, we adopt the following notations. Let $N$ be the total
+number of local devices, and $K \leq N$ is the maximal number of devices
+that are accessible to the central server.
+Let $T$ be the total number of  Stochastic Gradient Descent (SGD) iterations performed by each local device, $E$ be the local steps between two consecutive server communications.
+Thus, $T/E$ is
+the number of communications.
 

### 2020-06-01 source/main.tex
Update on Overleaf.

@@
 In this work, we provide a
-comprehensive convergence analysis of the widely used framework Federated Averaging (FedAvg),
-using SGD or Nesterov SGD updates, for general convex problems with arbitrarily
+comprehensive convergence analysis on Federated Averaging (FedAvg),
+a widely used framework in which SGD or Nesterov SGD updates are used for general convex problems with arbitrarily
 heterogeneous data. 

### 2020-06-01 source/sec_exp.tex
exp and sec 6 add threm id

@@
 % ~\ref{th:strongcvxsmth}~\ref{th:cvxsmoth}~\ref{th:fedmasslr}
-\textbf{Linear speedup.} To verify the linear speedup results in Theorems
-\lkxcom{cite sgd theorems}, we evaluate the number of iterations needed to reach
+\textbf{Linear speedup.} To verify the linear speedup results of FedAvg in Theorem~\ref{thm:SGD_scvx}~\ref{thm:SGD_cvx}, we evaluate the number of iterations needed to reach
 $\epsilon$-accuracy in three objectives. We initialize all runs with $\vw_0 = \textbf{0}_d$ and measure. For each configuration, we run three random seeds and report
@@
 we report the last iteration to converge to $\epsilon$-accuracy of Nesterov accelerated FedAvg.
-The empirical observations align with our theorems \lkxcom{cite accelerated theorems} that the accelerated version of FedAvg can also achieve the linear speedup
+The empirical observations align with Theorem~\ref{thm:nesterov_scvx}~\ref{thm:Nesterov_cvx} that the accelerated version of FedAvg can also achieve the linear speedup
 w.r.t the number of workers.

### 2020-06-01 source/sec_impact.tex
reorg exp
Update on Overleaf.

@@
 data integration can produce positive social
-impact in terms of better personalization~\cite{fallah2020personalized}, more efficient information solicitation~\cite{chen2018federated}, more robust and stable service~\cite{47586}.
+impact in terms of better personalization~\cite{fallah2020personalized}, more efficient information solicitation~\cite{chen2018federated}, and enhanced service~\cite{47586}.
 
 In the meantime, the increasing concerns of security and
-privacy leaking from both large-companies and the entire society
+privacy leaking from both large companies and the entire society
 present significant challenges of proper utilization of local
-data. This trade-off between efficiency and privacy or security
+data. This trade-off between efficiency and privacy/security
 lies in the center of many industrial applications, including 
-but not limited to recommender systems~\cite{chen2018federated}, virtual assistants~\cite{lamautonomy}, and mobile keyboard prediction~\cite{47586}
+but not limited to recommendation systems~\cite{chen2018federated}, virtual assistants~\cite{lamautonomy}, and mobile keyboard prediction~\cite{47586}.
 
@@
 On one hand, the data is isolated in each local devices, thus prevent
-the potential leaking of private information. On the otherhand, the
-convergence rates improved substantially if the number of
+the potential leakage of private information. On the other hand, the
+convergence rates can be improved substantially if the number of
 active local devices increases. In conclusion, our work 

### 2020-06-01 source/sec_intro.tex
Update on Overleaf.

@@
 The widespread usage of smart home devices and mobile phones constitutes
-large-scale distributed networks that generate amount of personal data.
-Models learned from such data lead to 
+large-scale distributed networks that generate significant amount of personal data.
+Models learned from such data have led to 
 % The emergence of those massive amounts of data has empowered machine learning
@@
 many successful applications such as personalized
-recommendations~\cite{chen2018federated}, keyboard predictions~\cite{47586},
-and etc. Due to the rising awareness of the privacy and rapid growth of
+recommendations~\cite{chen2018federated}, virtual assistants~\cite{lamautonomy}, and keyboard predictions~\cite{47586}, etc. Due to the rising awareness of  privacy and the rapid growth of
 computational power in mobile devices, it becomes increasingly attractive to
-push the computation to the edge. Recent years witnessed the rise of
+push the computation to the edge. Recent years witnessed the emergence of
 federated learning~\cite{mcmahan2016communication} to fulfill this demand.
@@
 As a distributed learning paradigm, Federated Learning (FL) collaboratively
-learns a single statistical model leveraging the data from all devices while
-keeping the training data isolated.  Thus, FL integrates cross-device
-information for acquiring improved models without transferring training data from individual devices, thus improves user
-privacy~\cite{kairouz2019advances}. Comparing to the traditional distribution optimization,  FL confronts unique
-challenges~\cite{li2019convergence,smith2017federated}, including statistical heterogeneity: the data distributions in different devices can be significantly different, and system heterogeneity: only a subset of devices may be accessible to the central server for model training at each time.
+learns a single statistical model leveraging the data from a massive amount of devices while
+keeping the training data isolated.  In particular, FL integrates cross-device
+information to achieve improved model quality without transferring training data from individual devices, thus improves user
+privacy~\cite{kairouz2019advances}. While promising, however, comparing to the traditional distribution optimization,  FL confronts unique
+challenges~\cite{li2019convergence,smith2017federated}, including statistical heterogeneity, that is, the data distributions in different devices can be significantly different, and system heterogeneity, that is, only a subset of devices may be accessible to the central server for model training at each time.
 
@@
 To address the aforementioned challenges, Federated
-Averaging~\cite{mcmahan2016communication} (FedAvg) was proposed as an
-effective heuristic and its great empirical success~\cite{47586} attracted
+Averaging (FedAvg)~\cite{mcmahan2016communication} was proposed as an
+effective heuristic which has enjoyed great empirical success~\cite{47586}, attracting
 lots of efforts on building up its theoretical understanding 
 and accelerated variants. 
-Existing convergence guarantees of FedAvg were largely established on simplified setting~\cite{stich2018local,khaled2019first,wang2018cooperative,yu2019parallel} or provides a suboptimal 
+Existing convergence guarantees of FedAvg were either largely established on simplified settings~\cite{stich2018local,khaled2019first,wang2018cooperative,yu2019parallel} or provides a suboptimal 
 convergence rate~\cite{li2019convergence}. Furthermore, the convergence of
@@
 \textbf{Contributions.}
-This work focus on providing comprehensive convergence analysis
+This work focuses on providing comprehensive convergence analysis
 of FedAvg and its accelerated variants under a realistic setting.
@@
 that the less number of communication rounds is needed to converge to a
-target accuracy and preserve the linear speedup w.r.t. participating devices at the same time.
+target accuracy while preserving the linear speedup w.r.t. the number of participating devices.
 More specifically, for strongly convex and smooth problems, we show 
 that FedAvg admits a convergence rate of 
-{\small{$\cO(1/NT)$}} convergence, and local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}.
+{\small{$\cO(1/NT)$}}, and local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}.
 For convex and smooth objectives, we show the rate of {\small{$\cO(1/\sqrt{NT})$}} and local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}.
@@
 Last but not least, we systematically examine the convergence rate of
-FedAvg and it accelerated variants on both a real-world dataset and a synthetic dataset. The empirical observations are well-aligned with theoretical indications.
+FedAvg and its accelerated variants on both a synthetic dataset and  a real-world dataset. The empirical observations are well-aligned with the obtained theoretical findings.
 
@@
 that are accessible to the central server.
-Let $T$ be the total number of  Stochastic Gradient Descent (SGD) iterations performed by each local device, $E$ be the local steps between two consecutive server communications.
+Let $T$ be the total number of SGD iterations performed by each local device, $E$ be the local steps between two consecutive server communications.
 Thus, $T/E$ is

### 2020-06-01 source/sec_relatedwork.tex
Update on Overleaf.

@@
 
-\section{Related Works}
+\section{Related Work}
 
@@
 
-\textbf{The Convergence of FedAvg and Linear Speedup}
-There has been a vast amount of convergence analysis for Federated averaging
-(FedAvg) in different settings. However, most of the prior works provide the
-convergence guarantee under the simplified setting, where either statistical
+\textbf{The Convergence of FedAvg and Linear Speedup.}
+There has been a vast amount of convergence analysis for FedAvg in different settings. However, most of the prior works provide the
+convergence guarantee under a simplified setting, where either statistical
 heterogeneity~\cite{stich2018local,zhou2017convergence,wang2018cooperative,woodworth2018graph} or system
@@
 In~\cite{li2019convergence},
-the authors firstly provide convergence of FedAvg for strongly convex problem considering both statistical heterogeneity and system heterogeneity, while their results
+the authors for the first time provide convergence of FedAvg for strongly convex problem considering both statistical heterogeneity and system heterogeneity, yet it
 didn't achieve linear speedup. In~\cite{haddadpour2019convergence,liang2019variance,huo2020faster,jiang2018linear}, the convergence of FedAvg for non-convex problems has been studied. 
@@
 
-\textbf{Federated Learning Beyond Local SGD}
+\textbf{Federated Learning Beyond Local SGD.}
 % The accelerated SGD methods has achieved optimal convergence rate in strongly convex

### 2020-06-01 source/main.tex

@@
 heterogeneous data. 
-We first show that the convergence rates of both algorithms enjoy a linear speedup w.r.t. the number of participating workers for all convex smooth problems. 
-Furthermore, in the overparamterized setting, we show that the convergence of FedAvg is in fact exponential, also with a speedup factor linear in the number of workers. 
-We propose a new accelerated federated learning algorithm that further improves this convergence rate with provable guarantees in the linear regression setting. 
+We first show that the convergence rates using both updates enjoy a linear speedup w.r.t. the number of participating workers for all convex smooth problems. 
+In addition, in the overparamterized setting, we show that the convergence of FedAvg is in fact exponential, also with a speedup factor linear in the number of workers. 
+Furthermore, we propose a new accelerated federated learning algorithm that further improves this convergence rate with provable guarantees in the linear regression setting. 
 Empirical studies of the algorithms in various settings have supported our theoretical results.

### 2020-06-01 source/sec_conclude.tex

@@
 convex smooth problems.
-In addition, we show that the local steps for stronlgy convex and convex smooth problems can be as large as {\small{$\cO(\sqrt{\frac{T}{N}})$}}, which can save communication cost substantially comparing to prior results.
+In addition, we show that the local steps for stronlgy convex and convex smooth problems can be as large as {\small{$\cO(\sqrt{\frac{T}{N}})$}}, which can substantially save communication cost comparing to prior results.
 Furthermore, this work also makes algorithmic contributions.

### 2020-06-01 source/sec_problem.tex
Update on Overleaf.
edited assumptions and updated sections 4 and 5

@@
 
-We study the federated learning problem
+In this paper, we study the following federated learning problem:
 \begin{align}
-	\min _{\mathbf{w}}\left\{F(\mathbf{w}) \triangleq \sum_{k=1}^{N} p_{k} F_{k}(\mathbf{w})\right\},
+	\min _{\mathbf{w}}\left\{F(\mathbf{w}) \triangleq \sum\nolimits_{k=1}^{N} p_{k} F_{k}(\mathbf{w})\right\},
 	\label{eq:problem}
@@
 We make the following standard assumptions on the objective function $F_1,\dots, F_N$.
-Assumption~\ref{ass:boundedvariance} and Assumption~\ref{ass:subgrad2} have been made by many prior works in federated learning~\cite{yu2019parallel,li2019convergence,stich2018local}.
-Assumption~\ref{ass:lsmooth} and Assumption~\ref{ass:stroncvx} are common assumptions satisfied by a range of popular objective functions, such as $l-2$
-norm regularized logisitc regressions, logistic regressions and cross-entropy loss functions.
+Assumption~\ref{ass:lsmooth} and Assumption~\ref{ass:stroncvx} are common assumptions satisfied by a range of popular objective functions, such as $\ell^{2}$ regularized logistic regression and cross-entropy loss functions.
 
-\begin{assumption}[Bounded local variance]
-	Let $\xi_{t}^{k}$ be sampled from the $k$-th device's local data uniformly at random.
-The variance of stochastic gradients in each device is bounded: $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq \sigma_{k}^{2}$,
-	for $k=1, \cdots, N$ and any $\mathbf{w}_{t}^{k}$. Let $\sigma^2=\max_k\sigma_{k}^{2}$.
-	\label{ass:boundedvariance}
-\end{assumption}
-\begin{assumption}[Expected square norm of gradient]
-	The expected squared norm of stochastic gradients is uniformly bounded. i.e.,
-	$\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)\right\|^{2} \leq G^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
-	\label{ass:subgrad2}
-\end{assumption}
+ %Assumption~\ref{ass:lsmooth} is made to achieve linear speedup of convergence rate with respect to the number of workers, though it is not necessary for FedAvg and its variants to convergence.
+
 \begin{assumption}[L-smooth]
@@
 \end{assumption}
-Assumption~\ref{ass:lsmooth} is made to achieve linear speedup 
-of convergence rate with respect to the number of workers, though it
-is not necessary for FedAvg and its variants to convergence.
-Assumption \ref{ass:subgrad2} implies $\left\| \mathbb{E}\vg_{t,k}\right\|^2  \leq \mathbb{E}\|\vg_{t,k}\|^2 \leq G_k^2$, which further implies $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq G^2$, the two assumptions are stated separately because it makes explicit the dependence of the optimality gap on different terms.
+\begin{assumption}[Bounded local variance]
+	Let $\xi_{t}^{k}$ be sampled from the $k$-th device's local data uniformly at random.
+The variance of stochastic gradients in each device is bounded: $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq \sigma_{k}^{2}$,
+	for $k=1, \cdots, N$ and any $\mathbf{w}_{t}^{k}$. Let $\sigma^2=\sum_{k=1}^{N}p_k\sigma_{k}^{2}$.
+	\label{ass:boundedvariance}
+\end{assumption}
+\begin{assumption}[Bounded local gradient]
+	The expected squared norm of stochastic gradients is uniformly bounded. i.e.,
+	$\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)\right\|^{2} \leq G^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
+	\label{ass:subgrad2}
+\end{assumption}
+
+\textbf{Justfication for Assumptions~\ref{ass:boundedvariance} and~\ref{ass:subgrad2}.} These standard assumptions have been made in many previous
+works in federated learning ~\cite{yu2019parallel,li2019convergence,stich2018local}. Here we provide one more justification
+for their generality: even though they require bounds for all $\mathbf{w}$,
+they are not much stronger than requiring bounds at the global optimal
+parameter $\mathbf{w}^{\ast}$. As model average parameters become
+closer to $\mathbf{w}^{\ast}$, the $L$-smoothness property implies
+that $\mathbb{E}\|\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})\|^{2}$
+and $\mathbb{E}\|\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})-\nabla F_{k}(\mathbf{w}_{t}^{k})\|^{2}$
+will become close to $\mathbb{E}\|\nabla F_{k}(\mathbf{w}^{\ast},\xi_{t}^{k})\|^{2}$
+and $\mathbb{E}\|\nabla F_{k}(\mathbf{w}^{\ast},\xi_{t}^{k})-\nabla F_{k}(\mathbf{w}^{\ast})\|^{2}$.
+Therefore, there is no substantial difference between these assumptions
+and assuming the bounds at $\mathbf{w}^{\ast}$ only. On the other
+hand, some works such as \cite{haddadpour2019convergence} assume
+bounds on the ``gradient diversity'', defined as $\frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$.
+Such assumptions are more restrictive than our assumptions, because
+they imply $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$ for all local
+objectives. For such ``overparameterized'' settings, we also prove
+that the convergence is in fact exponential.

### 2020-06-01 source/sec_SGD.tex

@@
 \section{Linear Speedup Analysis of FedAvg}
+\label{sec:sgd}
 
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right)
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}\leq\mathcal{O}(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}\leq\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right)
 	\end{align*}
@@
 %
-\begin{remark}
 \textbf{Linear speedup. }We first compare our bound with that in \cite{li2019convergence},
@@
 setting. This implies that in both settings, there is a linear speedup
-	in the number of active workers during the communication round. 
-\end{remark}
-%
-\begin{remark}
+in the number of active workers during the communication round. To
+our knowledge, this is the first result that explicitly demonstrates
+linear speedup in the number of active workers in the setting with
+both non i.i.d. data and partial participation. \\
 \textbf{Communication Complexity.} Our bound implies a distinction
@@
 still yields a linear speedup in $K$, which is not the case for previous
-	analyses. 
-\end{remark}
-%
-\begin{remark}
-	\textbf{Comparison with related works.} There have been a line of
-	works that demonstrate that linear speedup is possible in distributed
-	optimization with infrequent communication. In particular, \cite{haddadpour2019convergence}
-	is the only other work that demonstrates linear speedup for strongly
+analyses. \\
+\textbf{Comparison with related works.} \cite{haddadpour2019convergence}
+also proves a linear speedup $\mathcal{O}(1/NT)$ result for strongly
 convex and smooth objectives in the heterogeneous data with partial
 participation setting, with $\mathcal{O}(N^{1/3}T^{2/3})$ communication
-	complexity. However, their results are derived based on a gradient
-	diversity assumption, and the learning rates are carefully chosen
-	based on an upper bound of the gradient diversity. Thus our results
-	are not directly comparable as we use the gradient bound assumption
-	instead, and our learning rate does not depend on the bound $G$.
-\end{remark}
+complexity. However, their results are derived based on the more restrictive
+bounded gradient diversity assumption. As discussed before, if $\frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$
+is bounded, then $L$-smoothness implies $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$
+for all $k$. In this overparameterized setting, we show in Section
+\label{sec:overparameterized} that the convergence is in fact exponential,
+which improves on the $\mathcal{O}(1/NT)$ rate in \cite{haddadpour2019convergence}
+with better communication complexity.
 
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T})
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T}\right)
 	\end{align*}
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T})
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right)
 	\end{align*}
@@
 %
-\begin{remark}
 \textbf{Choice of $E$ and linear speedup. }With full participation,
@@
 be $O(1)$ in order to achieve linear speedup of $\mathcal{O}(1/\sqrt{KT})$.
-	Our result again demonstrates the difference in communication efficiencies
+Our result again demonstrates the difference in communication complexities
 between full and partial participation, and is to our knowledge the
 first result on linear speedup in the general federated learning setting
-	with both heterogeneous data and partial participation. 
-\end{remark}
-%
-\begin{remark}
+with both heterogeneous data and partial participation. \\
 \textbf{Learning rate. }The learning rate now depends on the final
-	horizon $T$ of the convergence statement, whereas before the learning
-	rate was set to $\mathcal{O}(1/t)$ for the $t$-th iteration. We
-	also note that the requirement on $\alpha_{t}=\mathcal{O}(\sqrt{N/T})$
-	is present in the works \cite{haddadpour2019convergence,yu2019parallel}
-	on nonconvex problems with $O(1/\sqrt{NT})$ linear speedup convergence
-	results. 
-\end{remark}
+horizon $T$ of the convergence statement, whereas in the strongly
+convex case the learning rate decays as $\mathcal{O}(1/t)$. Such
+a requirement $\alpha_{t}=\mathcal{O}(\sqrt{N/T})$ is also present
+in the works \cite{haddadpour2019convergence,yu2019parallel} on nonconvex
+problems with $\mathcal{O}(1/\sqrt{NT})$ linear speedup convergence results.

### 2020-06-01 source/sec_intro.tex

@@
 \section{Introduction}
-% Motivate FL
-The widespread usage of smart home devices and mobile phones constitutes
-large-scale distributed networks that generate significant amount of personal data.
-Models learned from such data have led to 
-% The emergence of those massive amounts of data has empowered machine learning
-% solutions to develop 
-many successful applications such as personalized
-recommendations~\cite{chen2018federated}, virtual assistants~\cite{lamautonomy}, and keyboard predictions~\cite{47586}, etc. Due to the rising awareness of  privacy and the rapid growth of
-computational power in mobile devices, it becomes increasingly attractive to
-push the computation to the edge. Recent years witnessed the emergence of
-federated learning~\cite{mcmahan2016communication} to fulfill this demand.
+Federated learning (FL) is a machine learning setting where many clients (e.g.
+mobile devices or whole organizations) collaboratively train a model under the orchestration of a central server (e.g.
+service provider), while keeping the training data decentralized~\cite{kairouz2019advances}.
+In recent years, FL has swiftly emerged to be an important learning paradigm~\cite{mcmahan2016communication}--one that enjoys widespread success in applications such as personalized recommendations~\cite{chen2018federated}, virtual assistants~\cite{lamautonomy}, keyboard predictions~\cite{47586} to name a few--for at least two reasons: First, the rapid proliferation of smart devices that are equipped with both computing power and data-capturing capabilities provided the infrastructure core for FL.
+Second, the rising awareness of privacy and the exponential growth of computational power (blessed by Moore's law) in mobile devices have made it increasingly attractive to push the computation to the edge.
 
-% Main challenge in FL.
-As a distributed learning paradigm, Federated Learning (FL) collaboratively
-learns a single statistical model leveraging the data from a massive amount of devices while
-keeping the training data isolated.  In particular, FL integrates cross-device
-information to achieve improved model quality without transferring training data from individual devices, thus improves user
-privacy~\cite{kairouz2019advances}. While promising, however, comparing to the traditional distribution optimization,  FL confronts unique
-challenges~\cite{li2019convergence,smith2017federated}, including statistical heterogeneity, that is, the data distributions in different devices can be significantly different, and system heterogeneity, that is, only a subset of devices may be accessible to the central server for model training at each time.
+Despite its promise and broad applicability in our current era, the potential value FL delivers is coupled with the unique challenges it brings forth. In particular,
+when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated (and hence private)~\cite{kairouz2019advances}, it faces two challenges that are absent in centralized optimization and
+traditional distributed optimization~\cite{li2019convergence,smith2017federated}:
+\begin{enumerate}
+\item\textbf{Statistical heterogeneity:} data distributions in different devices are different.
+\item\textbf{System heterogeneity:} only a subset of devices may access the central server at each time. 
+\end{enumerate}
 
 % Gaps in the theoretical understanding. 
-To address the aforementioned challenges, Federated
-Averaging (FedAvg)~\cite{mcmahan2016communication} was proposed as an
-effective heuristic which has enjoyed great empirical success~\cite{47586}, attracting
-lots of efforts on building up its theoretical understanding 
-and accelerated variants. 
+To address these challenges, Federated Averaging (FedAvg)~\cite{mcmahan2016communication} was proposed as a particularly effective heuristic, which has enjoyed great empirical success~\cite{47586}. 
+Its success has since motivated a growing line of research efforts into understanding its theoretical convergence guarantees in various settings.
+
  Existing convergence guarantees of FedAvg were either largely established on simplified settings~\cite{stich2018local,khaled2019first,wang2018cooperative,yu2019parallel} or provides a suboptimal 

### 2020-06-01 source/sec_intro.tex
Update on Overleaf.

@@
 Despite its promise and broad applicability in our current era, the potential value FL delivers is coupled with the unique challenges it brings forth. In particular,
-when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated (and hence private)~\cite{kairouz2019advances}, it faces two challenges that are absent in centralized optimization and
-traditional distributed optimization~\cite{li2019convergence,smith2017federated}:
+when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated (and hence private)~\cite{kairouz2019advances}, it faces two challenges that are absent in centralized optimization and traditional distributed optimization~\cite{li2019convergence,smith2017federated}:
 \begin{enumerate}
@@
 FedAvg's accelerated variants~\cite{yu2019linear,huo2020faster,liu2019accelerating} under the realistic setting is yet to be analyzed.
+
+\subsection{Related Work}
+\textbf{The Convergence of FedAvg and Linear Speedup.}
+There has been a vast amount of convergence analysis for FedAvg in different settings. However, most of the prior works provide the
+convergence guarantee under a simplified setting, where either statistical
+heterogeneity~\cite{stich2018local,zhou2017convergence,wang2018cooperative,woodworth2018graph} or system
+heterogeneity~\cite{yu2019parallel,wang2019adaptive,khaled2019first,jiang2018linear} are not
+considered. This simplified setting is also referred as local SGD. 
+In~\cite{li2019convergence},
+the authors for the first time provide convergence of FedAvg for strongly convex problem considering both statistical heterogeneity and system heterogeneity, yet it
+didn't achieve linear speedup. In~\cite{haddadpour2019convergence,liang2019variance,huo2020faster,jiang2018linear}, the convergence of FedAvg for non-convex problems has been studied. 
+However, \cite{huo2020faster} did not discuss the linear speedup and the theorems in \cite{haddadpour2019convergence,liang2019variance} failed to cover the accelerated version of
+FedAvg.
+
+\textbf{Federated Learning Beyond Local SGD.}
+% The accelerated SGD methods has achieved optimal convergence rate in strongly convex
+% optimization~\cite{nesterov1983method}.
+In \cite{liu2019accelerating}, the authors provide a convergence rate of accelerated FedAvg for strongly convex problems, while it is considered in full participation setting.
+\cite{yu2019linear} and \citep{huo2020faster} provide
+convergence guarantee for Nesterov accelerated FedAvg for non-convex. 
+However, \cite{yu2019linear} did not consider the system heterogeneity and \cite{huo2020faster} haven't discuss the linear speedup.
+Furthermore, the accelerated FedAvg admitting a similar convergence rate as FedAvg in all previous works.
+In this work, we provide a novel accelerated FedAvg with improved convergence rates under a popular overparameterized setting.
 

### 2020-06-01 source/app_notations.tex
add convergence table 1.

--- 2020-05-28 source/sec_appendix_notations.tex
@@
 \section{Additional Notations}
+\label{sec:app:notations}
 In this section, we introduce additional notations that are used throughout
@@
 and $\xi_t = \{\xi_t^k\}_{k=1}^N$ denotes random samples at all devices at time step $t$. 
+Since in this work, we also consider the case of partial participation. The sampling
+strategy to approximate the system heterogeneity can also affect the convergence. Here we
+follow the prior arts~\cite{haddadpour2019convergence} considering two types of sampling
+schemes. 
+The sampling scheme I establishes $\cS_{t+1}$ by i.i.d. sampling the devices with replacement,
+in this case the upper bound of expected square norm of $\ov{w}_{t+1} - \ov{v}_{t+1}$ is given by:
+\begin{align}
+\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4}{K} \eta_t^2 E^2G^2
+\end{align}
+The sampling scheme II establishes $\cS_{t+1}$ by uniformly sampling all devices without
+replacement, in which we have the 
+\begin{align}
+\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4(N - K)}{K(N-1)} \eta_t^2 E^2G^2
+\end{align}
+We denote this upper bound as follows for concise presentation. 
+\begin{align}
+	\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2 \leq  \eta_t^2 C
+	\label{eq:partialsample}
+\end{align}

### 2020-06-01 source/app_sumarization.tex
add convergence table 1.

% !TEX ROOT=./main.tex

\section{Comparison of convergence rates}

\begin{table}
\centering
\begin{tabular}{|c|c|c|}\hline 
	Algorithm \textbackslash Setting            & Strongly Convex        & Convex \\ \hline \hline
	FedAvg                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
	Nesterov accelerated FedAvg    & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    & $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
\end{tabular}
\caption{The summary of the convergence results in this work.}
\end{table}

\begin{table}
\centering
{\small
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline Reference                 & Convergence rate    & E                           & Non-IID & Participation & Extra Assumptions  		  & Setting  \\ \hline\hline 
\cite{li2019convergence}         & $\cO(\frac{E^2}{T})$& $\cO(1)$                    & \cmark  & Partial       & Bounded gradient   		  & Strongly convex  \\ \hline
\cite{haddadpour2019convergence} & $\cO(\frac{1}{KT})$ & $\cO(K^{1/3}T^{2/3})^*$     & \cmark  & Partial       & Bounded gradient diversity   & Strongly convex  \\ \hline
This work                        & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{**}$ & \cmark  & Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
\cite{zanette2019tighter}        &                     &                             & \cmark  & Partial       & Bounded gradient             & Convex  \\\hline\hline

% Khaled et al. (2020) & & & \\
% Li et al. (2019) & & & \\
% this work & $\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}}{NT\mu}+\frac{\kappa E^{2}G^{2}}{KT\mu}+\frac{\kappa^{2}E^{2}G^{2}}{T^{2}\mu}\right)$ &  &  \\
\end{tabular}
}
\caption{$^*$ This $E$ is obtained under i.i.d. setting. 
         $^{**}$: This $E$ is obtained under full participation setting.}
\end{table}

### 2020-06-01 source/sec_SGD.tex
minor edits

@@
 bounded gradient diversity assumption. As discussed before, if $\frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$
-is bounded, then $L$-smoothness implies $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$
-for all $k$. In this overparameterized setting, we show in Section
-\label{sec:overparameterized} that the convergence is in fact exponential,
-which improves on the $\mathcal{O}(1/NT)$ rate in \cite{haddadpour2019convergence}
-with better communication complexity.
+is bounded, then $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$ for all
+$k$. In this overparameterized setting, we show in Section~\ref{sec:overparameterized}
+that the convergence is in fact exponential, thus improving on the
+$\mathcal{O}(1/NT)$ rate in \cite{haddadpour2019convergence} with
+better communication complexity.
 

### 2020-06-01 source/sec_problem.tex
minor edits

@@
 and assuming the bounds at $\mathbf{w}^{\ast}$ only. On the other
-hand, some works such as \cite{haddadpour2019convergence} assume
-bounds on the ``gradient diversity'', defined as $\frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$.
+hand, some works such as \cite{haddadpour2019convergence,li2018federated}
+assume bounds on the ``gradient diversity'' or ``dissimilarity'',
+defined as $\frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$.
 Such assumptions are more restrictive than our assumptions, because
 they imply $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$ for all local
-objectives. For such ``overparameterized'' settings, we also prove
-that the convergence is in fact exponential.
+objectives. For such ``overparameterized'' settings, we will prove
+that the convergence is in fact geometric.

### 2020-06-01 source/app_sumarization.tex
Update app_sumarization.tex

@@
 \begin{tabular}{|c|c|c|}\hline 
-	Algorithm \textbackslash Setting            & Strongly Convex        & Convex \\ \hline \hline
+	Algorithm \textbackslash\  Objective function        & Strongly Convex        & Convex \\ \hline \hline
 	FedAvg                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
@@
 \caption{The summary of the convergence results in this work.}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{|c|c|c|}\hline 
+	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex \\ \hline \hline
+	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
+	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$      \\ \hline
+\end{tabular}
+\caption{The summary of the convergence results in this work. The}
 \end{table}

### 2020-06-01 source/app_sumarization.tex
Update app_sumarization.tex

@@
 \end{tabular}
-\caption{The summary of the convergence results in this work.}
+\caption{The summary of the convergence results under full participation setting in this work.}
 \end{table}
@@
 \end{tabular}
-\caption{The summary of the convergence results in this work. The}
+\caption{The summary of the convergence results of FedAvg and Nesterov accelerated FedAvg in this work.}
 \end{table}

### 2020-06-02 source/sec_exp.tex
exp cut

@@
 \textbf{The impact of $E$.} We further examine how does the number of local steps ($E$) 
-affect convergence. As shown in Figure~\ref{fig:e}, the number of iterations increases as $E$ increase, which slow down the convergence in
+affect convergence. As shown in Figure~\ref{fig:e} in appendix section~\ref{sec:expsupp}, the number of iterations increases as $E$ increase, which slow down the convergence in
 terms of gradient computation. However, it can save communication costs as
@@
 
-\begin{figure}
-\centering
-	\begin{tabular}{ccc}
-	\hspace{-2em} \includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsT-min-w8a-epsilon0131-reg1e-05.pdf} &
-	\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupEpochsT-min-w8a-epsilon0134-reg0.pdf} & 
-	\includegraphics[width=0.33\textwidth]{fig/paper-linregression-newspeedupEpochsT-min-linearregressionw8a-epsilon002-reg0.pdf} \\
-	\hspace{-2em} \includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsRounds-min-w8a-epsilon0131-reg1e-05.pdf} &
-	\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupEpochsRounds-min-w8a-epsilon0134-reg0.pdf} & 
-	\includegraphics[width=0.33\textwidth]{fig/paper-linregression-newspeedupEpochsRounds-min-linearregressionw8a-epsilon002-reg0.pdf} \\
-(a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
-	\end{tabular}
-\caption{The convergence of FedAvg w.r.t the number of local steps $E$. }
-\label{fig:e}
-\end{figure}
+% \begin{figure}
+% \centering
+% 	\begin{tabular}{ccc}
+% 	\hspace{-2em} \includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsT-min-w8a-epsilon0131-reg1e-05.pdf} &
+% 	\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupEpochsT-min-w8a-epsilon0134-reg0.pdf} & 
+% 	\includegraphics[width=0.33\textwidth]{fig/paper-linregression-newspeedupEpochsT-min-linearregressionw8a-epsilon002-reg0.pdf} \\
+% 	\hspace{-2em} \includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsRounds-min-w8a-epsilon0131-reg1e-05.pdf} &
+% 	\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupEpochsRounds-min-w8a-epsilon0134-reg0.pdf} & 
+% 	\includegraphics[width=0.33\textwidth]{fig/paper-linregression-newspeedupEpochsRounds-min-linearregressionw8a-epsilon002-reg0.pdf} \\
+% (a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
+% 	\end{tabular}
+% \caption{The convergence of FedAvg w.r.t the number of local steps $E$. }
+% \label{fig:e}
+% \end{figure}

### 2020-06-02 source/app_sumarization.tex
Update app_sumarization.tex

--- 2020-06-01 source/app_sumarization.tex
@@
 
-\section{Comparison of convergence rates}
+% \section{Comparison of convergence rates}
 
-\begin{table}
+\begin{table}[h!]
 \centering
@@
 \end{tabular}
-\caption{The summary of the convergence results in this work.}
+\caption{A high level summary of the convergence results for full participation in this paper. This table only hightlights the
+dependence on $T$ (number of iterations), $E$ (the largest number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices.}
+\label{tb:convergenceratev1}
 \end{table}
 
-\begin{table}
+\begin{table}[h!]
 \centering
@@
 \end{tabular}
-\caption{The summary of the convergence results in this work. The}
+\caption{A high level summary of the convergence results in this paper. This table only hightlights the
+dependence on $T$ (number of iterations), $E$ (the largest number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices.
+These convergence rates applied to both FedAvg and Nesterov accelerated FedAvg.}
+\label{tb:convergenceratev2}
 \end{table}
 
-\begin{table}
+\begin{table}[h!]
 \centering
+\hspace{-2em}
 {\small
 \begin{tabular}{|c|c|c|c|c|c|c|}
-\hline Reference                 & Convergence rate    & E                           & Non-IID & Participation & Extra Assumptions  		  & Setting  \\ \hline\hline 
-\cite{li2019convergence}         & $\cO(\frac{E^2}{T})$& $\cO(1)$                    & \cmark  & Partial       & Bounded gradient   		  & Strongly convex  \\ \hline
-\cite{haddadpour2019convergence} & $\cO(\frac{1}{KT})$ & $\cO(K^{1/3}T^{2/3})^*$     & \cmark  & Partial       & Bounded gradient diversity   & Strongly convex  \\ \hline
-This work                        & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{**}$ & \cmark  & Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
-\cite{zanette2019tighter}        &                     &                             & \cmark  & Partial       & Bounded gradient             & Convex  \\\hline\hline
-
-% Khaled et al. (2020) & & & \\
-% Li et al. (2019) & & & \\
-% this work & $\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}}{NT\mu}+\frac{\kappa E^{2}G^{2}}{KT\mu}+\frac{\kappa^{2}E^{2}G^{2}}{T^{2}\mu}\right)$ &  &  \\
+\hline Reference                 & Convergence rate    & E                           			& Participation & Extra Assumptions  		  & Setting  \\ \hline\hline 
+\cite{li2019convergence}         & $\cO(\frac{E^2}{T})$& $\cO(1)$                    			& Partial       & Bounded gradient   		  & Strongly convex  \\ \hline
+\cite{haddadpour2019convergence} & $\cO(\frac{1}{KT})$ & $\cO(K^{1/3}T^{2/3})^{\dagger}$     			& Partial       & Bounded gradient diversity   & Strongly convex$^{\mathsection}$  \\ \hline
+This work                        & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{\dagger\dagger}$ 			& Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
+\cite{zanette2019tighter}  & $\cO(\frac{1}{\sqrt{NT}})$ & $\cO(N^{-3/2}T^{1/2})$     			& Full        & Bounded gradient             & Convex  \\\hline
+This work      & $\cO\left(\frac{1}{\sqrt{NT}}\right)$ &  $\cO(N^{-1/2}T^{1/2})^{\dagger\dagger}$			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
+This work & $\cO\left(\exp(-\frac{NT}{E\kappa_1})\right)$ & $ \cO(T^{\beta})$                   &  Partial     & Bounded gradient    & Overparameterized (LR)$^{\ddagger}$ \\ \hline
+This work & $\cO\left(\exp(-\frac{NT}{E\sqrt{\kappa_1\tilde{\kappa}}})\right)$ & $ \cO(T^{\beta})$  &  Partial     & Bounded gradient    & Overparameterized LR \\ \hline
 \end{tabular}
 }
-\caption{$^*$ This $E$ is obtained under i.i.d. setting. 
-         $^{**}$: This $E$ is obtained under full participation setting.}
+\caption{A high level summary of the convergence results in this paper and their comparison to prior state-of-the-art FL algorithms, considering heterogeneous data. This table only hightlights the
+dependence on $T$ (number of iterations), $E$ (the largest number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices. 
+$\kappa$ is the condition number of the system and $\beta \in (0,1)$. 
+We note that all our results for strongly convex and convex smooth problems are unified results for both FedAvg and accelerated FedAvg. The converegence results for prior arts only apply for FedAvg.}
+{\raggedright 
+         $^{\dagger}$ This $E$ is obtained under i.i.d. setting. \\
+         $^{\dagger\dagger}$ This $E$ is obtained under full participation setting. \\ 
+         $^{\mathsection}$ In~\cite{haddadpour2019convergence}, the convergence rate is for non-convex smooth problem with PL condition.
+Thus, it is also applied for strongly convex problem.
+For unified comparison, we denote it as strongly convex setting here.
+\\
+         $^{\ddagger}$ The convergence of FedAvg for general overparameterized strongly convex and overparameterized linear regression share the same result.  \par}
+\label{tb:convergenceratev3}
 \end{table}

### 2020-06-02 source/app_sumarization.tex

@@
 \cite{li2019convergence}         & $\cO(\frac{E^2}{T})$& $\cO(1)$                    			& Partial       & Bounded gradient   		  & Strongly convex  \\ \hline
-\cite{haddadpour2019convergence} & $\cO(\frac{1}{KT})$ & $\cO(K^{1/3}T^{2/3})^{\dagger}$     			& Partial       & Bounded gradient diversity   & Strongly convex$^{\mathsection}$  \\ \hline
-This work                        & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{\dagger\dagger}$ 			& Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
+\cite{haddadpour2019convergence} & $\cO(\frac{1}{KT})$ & $\cO(K^{1/3}T^{2/3})^1$     			& Partial       & Bounded gradient diversity   & Strongly convex$^{3}$  \\ \hline
+This work                        & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{2}$ 			& Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
 \cite{zanette2019tighter}  & $\cO(\frac{1}{\sqrt{NT}})$ & $\cO(N^{-3/2}T^{1/2})$     			& Full        & Bounded gradient             & Convex  \\\hline
-This work      & $\cO\left(\frac{1}{\sqrt{NT}}\right)$ &  $\cO(N^{-1/2}T^{1/2})^{\dagger\dagger}$			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
-This work & $\cO\left(\exp(-\frac{NT}{E\kappa_1})\right)$ & $ \cO(T^{\beta})$                   &  Partial     & Bounded gradient    & Overparameterized (LR)$^{\ddagger}$ \\ \hline
+This work      & $\cO\left(\frac{1}{\sqrt{NT}}\right)$ &  $\cO(N^{-1/2}T^{1/2})^{2}$			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
+This work & $\cO\left(\exp(-\frac{NT}{E\kappa_1})\right)$ & $ \cO(T^{\beta})$                   &  Partial     & Bounded gradient    & Overparameterized (LR)$^{4}$ \\ \hline
 This work & $\cO\left(\exp(-\frac{NT}{E\sqrt{\kappa_1\tilde{\kappa}}})\right)$ & $ \cO(T^{\beta})$  &  Partial     & Bounded gradient    & Overparameterized LR \\ \hline
@@
 {\raggedright 
-         $^{\dagger}$ This $E$ is obtained under i.i.d. setting. \\
-         $^{\dagger\dagger}$ This $E$ is obtained under full participation setting. \\ 
-         $^{\mathsection}$ In~\cite{haddadpour2019convergence}, the convergence rate is for non-convex smooth problem with PL condition.
+         $^1$ This $E$ is obtained under i.i.d. setting. \\
+         $^{2}$ This $E$ is obtained under full participation setting. \\ 
+         $^{3}$ In~\cite{haddadpour2019convergence}, the convergence rate is for non-convex smooth problem with PL condition.
 Thus, it is also applied for strongly convex problem.
@@
 \\
-         $^{\ddagger}$ The convergence of FedAvg for general overparameterized strongly convex and overparameterized linear regression share the same result.  \par}
+         $^{4}$ The convergence of FedAvg for general overparameterized strongly convex and overparameterized linear regression share the same result.  \par}
 \label{tb:convergenceratev3}

### 2020-06-02 source/sec_exp_details.tex

@@
 
-\section{Details on Experiments}
+\section{Details on Experiments and additional results}
 % \subsection{Experiments}
@@
 in FedAvg.
+
+\subsection{The impact of $E$.}
+In this subsection, we further examine how does the number of local steps ($E$) 
+affect convergence. As shown in Figure~\ref{fig:e}, the number of iterations increases as $E$ increase, which slow down the convergence in
+terms of gradient computation. However, it can save communication costs as
+the number of rounds decreased when the $E$ increases. This showcase that
+we need a proper choice of $E$ to trade-off the communication cost and
+convergence speed. 
+\begin{figure}
+\centering
+	\begin{tabular}{ccc}
+	\hspace{-2em} \includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsT-min-w8a-epsilon0131-reg1e-05.pdf} &
+	\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupEpochsT-min-w8a-epsilon0134-reg0.pdf} & 
+	\includegraphics[width=0.33\textwidth]{fig/paper-linregression-newspeedupEpochsT-min-linearregressionw8a-epsilon002-reg0.pdf} \\
+	\hspace{-2em} \includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsRounds-min-w8a-epsilon0131-reg1e-05.pdf} &
+	\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupEpochsRounds-min-w8a-epsilon0134-reg0.pdf} & 
+	\includegraphics[width=0.33\textwidth]{fig/paper-linregression-newspeedupEpochsRounds-min-linearregressionw8a-epsilon002-reg0.pdf} \\
+(a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
+	\end{tabular}
+\caption{The convergence of FedAvg w.r.t the number of local steps $E$. }
+\label{fig:e}
+\end{figure}

### 2020-06-02 source/sec_overparameterized.tex
Update on Overleaf.

--- 2020-05-30 source/sec_overparameterized.tex
@@
 \section{Exponential Convergence of FedAvg in the Overparameterized Setting}
+\label{sec:overparameterized}
 
-So far we have shown $O(1/\sqrt{NT})$ and $O(1/NT)$ convergence
-rates for FedAvg with local SGD and Nesterov udpates in the convex
-and strongly convex settings, respectively. In this section, we turn
-to the special setting of overparameterized problems, where a non-negative
-loss function admits a zero-loss global minimizer. Such problems are
-very common in machine learning systems where the number of parameters
-far exceeds the number of data points, and perfect fitting is possible.
-
-There have been a line of recent works {[}CITE{]} showing that SGD
-and accelerated methods achieve exponential convergence in this case,
-thanks to the property of ``automatic variance reduction'' {[}CITE{]}.
+In this section, we turn to the special setting of overparameterized
+problems, where a non-negative loss function admits a zero-loss global
+minimizer. Such problems are very common in machine learning systems
+where the number of parameters far exceeds the number of data points,
+and perfect fitting is possible. There have been a line of recent
+works showing that SGD and accelerated methods achieve exponential
+convergence in similar settings thanks to the property of ``automatic
+variance reduction'' \cite{ma2017power,moulines2011non,needell2014stochastic,schmidt2013fast,strohmer2009randomized}.
 The natural question is whether such a result also holds in the federated
-learning setting. We first show that this is indeed the case and establish
-the exponential (linear) convergence of FedAvg with local SGD updates
+learning setting. We show that this is indeed the case and establish
+the geometric (linear) convergence of FedAvg with local SGD updates
 with constant step size for general strongly convex and smooth overparameterized
@@
 linearly in the number of workers $N$ when $N$ is below some problem-dependent
-threhold, while decreases with $E$ through $1/E$.
+threshold, while it decreases with $E$ through $1/E$.
 
 To our knowledge, this is the first result on the exponential convergence
-of FedAvg algorithms in the interpolation setting with linear speedup
-in the number of workers and explicit dependence on the communication
-interval $E$. We next sharpen this exponential convergence rate in
-the special case of linear regression, and show that it is $O(\exp(-\frac{NT}{E\kappa_{1}}))$,
-for $\kappa_{1}$ an appropriately defined condition number of the
-system.
+of FedAvg algorithms in the overparameterized setting with linear
+speedup in the number of workers and explicit dependence on the communication
+interval $E$. We next sharpen this result in the special case of
+quadratic problems, and show that it is $\mathcal{O}(\exp(-\frac{NT}{E\kappa_{1}}))$,
+for $\kappa_{1}$ a condition number of the system.
 
-Lastly, we turn to the question of whether FedAvg with momentum-based
+Lastly, we deal with the question of whether FedAvg with momentum-based
 local updates can outperform FedAvg with SGD updates. In contrast
-to the gradient descent setting, in the single-agent stochastic gradient
-setting, Nesterov and Heavy Ball updates are known to fail to accelerate
-over SGD, both in the overparameterized setting and standard convex
-setting{[}CITE{]}. Thus in general we cannot hope to obtain acceleration
-results for the FedAvg algorithm with Nesterov and Heavy Ball updates.
-On the other hand, in the overparameterized setting, {[}CITE Belkin{]}
-introduced the MaSS algorithm, which is a modification of the Nesterov
-update, where in the update a non-negative multiple of the gradient
-is added to the Nesterov parameter update to correct for the ``over-descent''
-of the Nesterov update. For quadratic objectives, the authors show
-that the MaSS algorithm is able to achieve acceleration over the exponential
-convergence of SGD both theoretically and empirically. In the last
-part of this section, we introduce a new accelerated FedAvg algorithm
-by adapting the MaSS algorithm to the federated learning setting.
-For this new accelerated FedAvg algorithm with local MaSS updates,
-we show that it achieves exponential convergence for overparameterized
-quadratic problems with rate $O(\exp(-\frac{NT}{E\sqrt{\kappa_{1}\tilde{\kappa}}}))$,
-where $\tilde{\kappa}$ is the ``statistical condition number''
-introduced in {[}CITE{]} that satisfies $\tilde{\kappa}\leq\kappa_{1}$.
-Thus the FedAvg algorithm with local MaSS updates preserves the linear
-speedup property in the number of workers and the $1/E$ dependence
-on $E$, while achieving a speedup of factor $\sqrt{\frac{\kappa_{1}}{\tilde{\kappa}}}$
+to the gradient descent setting, Nesterov and Heavy Ball updates are
+known to fail to accelerate over SGD, both in the overparameterized
+setting and standard convex setting \cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2016influence}.
+Thus in general one cannot hope to obtain acceleration results for
+the FedAvg algorithm with Nesterov and Heavy Ball updates. On the
+hopeful side, \cite{jain2017accelerating,liu2018accelerating} introduced
+similar algorithms that make modifications of the Nesterov updates
+to correct for its ``over-descent''. For quadratic objectives in
+the overparamterized setting, \cite{liu2018accelerating} show their
+algorithm achieves acceleration over the geometric convergence of
+SGD that recovers the well-known $\mathcal{O}(\exp(-t/\sqrt{\kappa})$
+acceleration of Nesterov over GD. In the last part, we introduce a
+new accelerated FedAvg algorithm by adapting the MaSS algorithm of
+\cite{liu2018accelerating} to the federated learning setting. We
+show that it achieves geometric convergence for overparameterized
+quadratic problems with rate $\mathcal{O}(\exp(-\frac{NT}{E\sqrt{\kappa_{1}\tilde{\kappa}}}))$,
+where $\tilde{\kappa}$ is a ``statistical condition number''\cite{liu2018accelerating,jain2017accelerating}
+that satisfies $\tilde{\kappa}\leq\kappa_{1}$. Thus our new FedAvg
+algorithm achieves a speedup of factor $\sqrt{\kappa_{1}/\tilde{\kappa}}$
 over FedAvg with local SGD updates.
 
-\subsection{Exponential Convergence and Linear Speedup of FedAvg with local SGD }
-
-In {[}CITE Belkin{]}, the authors show that in the overparameterized
-setting, SGD achieves exponential convergence with linear speedup
-in batch size that is below some critical threshold. In this section,
-we extend their result to the federated learning setting and show
-that a similar exponential convergence with linear speedup in the
-number of workers using FedAvg with local SGD updates holds.
-
-Recall the federated learning problem $\min_{w}\sum_{k=1}^{N}p_{k}F_{k}(w)$
-with $F_{k}(w)=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\ell(w;\xi_{k,j})$.
-In this section, we consider the standard ERM setting where each $\ell(w;\xi_{k,j})$
+\subsection{Geometric Convergence of FedAvg in the Overparameterized Setting}
+Recall the federated learning problem $\min_{w}\sum_{k=1}^{N}p_{k}F_{k}(\mathbf{w})$
+with $F_{k}(\mathbf{w})=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\ell(\mathbf{w};\mathbf{x}_{k}^{j})$.
+In this section, we consider the standard ERM setting where each $\ell$
 is non-negative, $l$-smooth, and convex, and as before, we assume
-that each $F_{k}(w)=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\ell(w;\xi_{k,j})$
+that each $F_{k}(\mathbf{w})=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\ell(\mathbf{w};\mathbf{x}_{k}^{j})$
 is $L$-smooth and $\mu$-strongly convex. Note that $l\geq L$. This
-setting includes many important problems in practice, such as linear
-regression with a full rank sample covariance matrix. In the overparameterized
-setting, we assume that there exists $w^{\ast}\in\arg\min_{w}\sum_{k=1}^{N}p_{k}F_{k}(w)$
-such that $\ell(w^{\ast};\xi_{k,j})=0$ for all $\xi_{k,j}$. Given
-these assumptions, we first show that FedAvg with local SGD updates
-and communication every $E$ iterations achieves exponential convergence
+setup includes many important problems in practice. In the overparameterized
+setting, we assume that there exists $\mathbf{w}^{\ast}\in\arg\min_{w}\sum_{k=1}^{N}p_{k}F_{k}(\mathbf{w})$
+such that $\ell(\mathbf{w}^{\ast};\mathbf{x}_{k}^{j})=0$ for all
+$\mathbf{x}_{k}^{j}$. We first show that FedAvg with local SGD updates
+and communication every $E$ iterations achieves geometric convergence
 with linear speedup in the number of workers. 
 \begin{theorem}
-	For the overparameterized setting with general $\mu$-strongly convex
-	and $L$-smooth objectives, FedAvg with local SGD updates and communication
-	every $E$ iterations with constant step size $\overline{\alpha}=\frac{1}{4E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})}$
-	gives the exponential convergence guarantee 
+	\label{thm:overparameterized_general}In the overparameterized setting,
+	FedAvg with local SGD updates and communication every $E$ iterations
+	with constant step size $\overline{\alpha}=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})})$
+	has geometric convergence
 	\begin{align*}
-	\mathbb{E}F(\overline{w}_{t}) & \leq\frac{L}{2}(1-\mu\overline{\alpha})^{t}\|w_{0}-w^{\ast}\|^{2}=O(L\exp(-\frac{\mu}{4E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})}t)\cdot\|w_{0}-w^{\ast}\|^{2})
+	\mathbb{E}F(\overline{\mathbf{w}}_{T}) & \leq\frac{L}{2}(1-\overline{\alpha})^{T}\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}=\mathcal{O}\left(L\exp(-\frac{\mu}{E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})}t)\cdot\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right)
 	\end{align*}
@@
 %
-\begin{remark}
-	When $l>L$, the speedup factor is on the order of $\frac{N}{E(l\nu_{\max}+L(N-\nu_{\min}))}/\frac{1}{l\nu_{\max}+L(1-\nu_{\min})}\approx\frac{Nl}{E(l+L(N-1))}=O(N/E)$
+\textbf{Linear speedup and Communication Complexity} When $l>L$,
+the speedup factor is on the order of $\frac{N}{E(l\nu_{\max}+L(N-\nu_{\min}))}/\frac{1}{l\nu_{\max}+L(1-\nu_{\min})}\approx\frac{Nl}{E(l+L(N-1))}=\mathcal{O}(N/E)$
 for $N\leq\frac{l}{L}+1$, i.e. FedAvg with $N$ workers and communication
-	every $E$ iterations provides an exponential convergence speedup
-	factor of $O(N/E)$, for $N\leq\frac{l}{L}+1$. When $N$ is above
-	this threhold, however, the speedup is almost constant in the number
-	of woerkers. Our bound also illustrates the tradeoff between increasing
-	the number of workers and the communication latency. When the cost
-	of communication is high, i.e. it is costly to aggregate local parameters
-	and broadcast the average to all active devices, increasing the number
-	of workers by some factor achieves the same order of speedup as decreasing
-	the communication lag by the same factor. 
-\end{remark}
-In the next part, we specialize to the problem of linear regression,
-and demonstrate how the above bound can be sharpened to explicitly
-depend on the condition number of the system. This result also paves
-the way for the analysis of a new FedAvg with local momentum-based
-updates in the following part.
+every $E$ iterations provides a geometric convergence speedup factor
+of $\mathcal{O}(N/E)$, for $N\leq\frac{l}{L}+1$. When $N$ is above
+this threshold, however, the speedup is almost constant in the number
+of workers. This matches the findings in \cite{ma2017power}. Our
+result also illustrates that $E$ can be taken $\mathcal{O}(t^{\beta})$
+for any $\beta<1$, achieving better communication efficiency than
+the standard setting.
 
@@
 
-We now consider the federated learning problem with overparameterized
-quadratic objectives. In other words, the local device objectives
-are given by the sum of squares. 
-\begin{align*}
-F_{k}(w) & =\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(w^{T}x_{k,j}-z_{k,j})^{2}
-\end{align*}
-and there exists $w^{\ast}$ such that $F(w^{\ast})=\sum_{k}p_{k}F_{k}(w^{\ast})\equiv0$.
-For such quadratic problems, we will present a sharper bound for FedAvg
-with explicit dependence on the condition number, and demonstrate
-how the convergence rate can be further improved using a momentum-based
-accelerated FedAvg algorithm based on the MaSS algorithm of {[}CITE
-Belkin{]}.
+In the remaining part of the section, we specialize to quadratic problems
+and demonstrate how the general bound in \ref{thm:overparameterized_general}
+can be sharpened to explicitly depend on the condition number of the
+system, and how this rate can be further improved with a new FedAvg
+algorithm with momentum updates.
 
-In this section, we first gather some necessary definitions and observations
-useful in the analysis of exponential convergence of FedAvg algorithms.
-Define the local Hessian matrix as $H^{k}:=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}x_{k,j}(x_{k,j})^{T}$
+The local device objectives are now given by the sum of squares $F_{k}(\mathbf{w})=\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(\mathbf{w}^{T}\mathbf{x}_{k}^{j}-z_{k}^{j})^{2}$,
+and there exists $\mathbf{w}^{\ast}$ such that $F(\mathbf{w}^{\ast})=\sum_{k}p_{k}F_{k}(\mathbf{w}^{\ast})\equiv0$.
+Define the local Hessian matrix as $H^{k}:=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\mathbf{x}_{k}^{j}(\mathbf{x}_{k}^{j})^{T}$
 and the averaged Hessian matrix as $H:=\sum_{k=1}^{N}p_{k}H^{k}$.
+\begin{comment}
 In general $H$ has zero eigevalues. However, because the null space
 of $H$ and range of $H$ are orthogonal, in our subsequence analysis
-it suffices to project $\overline{w}_{t}-w^{\ast}$ onto the range
-of $H$, thus we may restrict to the non-zero eigenvalue of $H$.
-
-We can use $w^{\ast T}x_{k,j}-z_{k,j}\equiv0$ to rewrite the local
-objectives 
+it suffices to project $\overline{\mathbf{w}}_{t}-\mathbf{w}^{\ast}$
+onto the range of $H$, thus we may restrict to the non-zero eigenvalue
+of $H$. We can use $\mathbf{w}^{\ast T}\mathbf{x}_{k}^{j}-z_{k}^{j}\equiv0$
+to rewrite the local objectives as $F_{k}(\mathbf{w})=\frac{1}{2}\langle\mathbf{w}-\mathbf{w}^{\ast},H^{k}(\mathbf{w}-\mathbf{w}^{\ast})\rangle\equiv\frac{1}{2}\|\mathbf{w}-\mathbf{w}^{\ast}\|_{H^{k}}^{2}$
+so that $F(\mathbf{w})=\frac{1}{2}\|\mathbf{w}-\mathbf{w}^{\ast}\|_{H}^{2}$.
+\end{comment}
+{} %
+\begin{comment}
 \begin{align*}
@@
 \end{align*}
-so that $F(w)=\frac{1}{2}\|w-w^{\ast}\|_{H}^{2}$.
-
-Let $x_{t}^{k}$ be the stochastic sample on the $k$th device during
-iteration $t$, and define $\tilde{H}_{t}^{k}:=x_{t}^{k}(x_{t}^{k})^{T}$
-as the stochastic Hessian matrix corresponding to the sample $x_{t}^{k}$.
-Note that $\mathbb{E}\tilde{H}_{t}^{k}=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}x_{k,j}(x_{k,j})^{T}=H^{k}$
-so that $\tilde{H}_{t}^{k}$ is an unbiased estimate of $H^{k}$.
-Moreover, we have
+\end{comment}
+Let $\xi_{t}^{k}$ be the stochastic sample on the $k$th device at
+time $t$, and define $\tilde{H}_{t}^{k}:=\xi_{t}^{k}(\xi_{t}^{k})^{T}$
+as the stochastic Hessian matrix. %
+\begin{comment}
+Note that $\mathbb{E}\tilde{H}_{t}^{k}=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\mathbf{x}_{k}^{j}(\mathbf{x}_{k}^{j})^{T}=H^{k}$
+and $\mathbf{g}_{t,k}=\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})=\tilde{H}_{t}^{k}(\mathbf{w}_{t}^{k}-\mathbf{w}^{\ast})$
+while $\mathbf{g}_{t}=\sum_{k=1}^{N}p_{k}\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})=\sum_{k=1}^{N}p_{k}\tilde{H}_{t}^{k}(\mathbf{w}_{t}^{k}-\mathbf{w}^{\ast})$.
+\end{comment}
+{} Define $l$ to be the smallest positive number such that $\mathbb{E}\|\xi_{t}^{k}\|^{2}$$\mathbf{\xi}_{t}^{k}$($\mathbf{\xi}_{t}^{k})^{T}\preceq lH^{k}$
+for all $k$. Note that $l\leq\max_{k,j}\|\mathbf{x}_{k}^{j}\|^{2}$.
+Let $L$ and $\mu$ be lower and upper bounds of non-zero eigenvalues
+of $H^{k}$. Define $\kappa_{1}:=l/\mu$ and $\kappa:=L/\mu$. Following
+\cite{liu2018accelerating,jain2017accelerating}, we define the statistical
+condition number $\tilde{\kappa}$ as the smallest positive real number
+such that 
 \begin{align*}
-g_{t,k} & =\nabla F_{k}(w_{t}^{k},x_{t}^{k})=\tilde{H}_{t}^{k}(w_{t}^{k}-w^{\ast})\\
-g_{t} & =\sum_{k=1}^{N}p_{k}\nabla F_{k}(w_{t}^{k},x_{t}^{k})=\sum_{k=1}^{N}p_{k}\tilde{H}_{t}^{k}(w_{t}^{k}-w^{\ast})
+\mathbb{E}\left[\langle\xi_{t}^{k}(H^{k})^{-1},\xi_{t}^{k}\rangle\xi_{t}^{k}(\xi_{t}^{k})^{T}\right] & \preceq\tilde{\kappa}H^{k}
 \end{align*}
-
-Define $l$ to be the smallest positive number such that $\mathbb{E}\|x_{t}^{k}\|^{2}$$x_{t}^{k}$($x_{t}^{k})^{T}\preceq lH^{k}$
-for all $k,t$. Note that $l\leq\max_{k=1,\dots,N}\max_{1\leq j\leq n_{k}}\|x_{k,j}\|^{2}$,
-i.e. $l$ is upper bounded by the maximum norm squared among all feature
-vectors on all devices.
-
-Let $L^{k}$ and $\mu^{k}$ to be the largest and smallest non-zero
-eigenvalues of $H^{k}$, and $L:=\max_{k}L^{k}$, $\mu:=\min_{k}\mu^{k}$.
-Define $\kappa_{1}:=l/\mu$ and $\kappa:=L/\mu$.
-
-Following {[}CITE: Liu\&Belkin and Jain et al{]}, we define the statistical
-condition number $\tilde{\kappa}^{k}$ as the smallest positive real
-number such that 
-\begin{align*}
-\mathbb{E}\left[\langle x_{t}^{k}(H^{k})^{-1},x_{t}^{k}\rangle x_{t}^{k}(x_{t}^{k})^{T}\right] & \preceq\tilde{\kappa}^{k}H^{k}
-\end{align*}
-and define $\tilde{\kappa}:=\max_{k}\tilde{\kappa}^{k}$. The condition
-numbers $\kappa_{1}$ and $\tilde{\kappa}$ are important in the characterization
-of convergence rates for FedAvg algorithms. Note that $\kappa_{1}>\kappa$ and $\kappa_{1}>\tilde{\kappa}$.
+for all $k$. The condition numbers $\kappa_{1}$ and $\tilde{\kappa}$
+are important in the characterization of convergence rates for FedAvg
+algorithms. Note that $\kappa_{1}>\kappa$ and $\kappa_{1}>\tilde{\kappa}$.
 
@@
 
-In the single-agent stochastic optimization setting, Nesterov updates
-are known to fail to accelerate over SGD updates, so that in the federated
-learning setting, we cannot hope for acceleration results of FedAvg
-with Nesterov updates in general. The natural question is then whether
-there are accelerated versions of the FedAvg algorithm that have provable
-improvements over the FedAvg algorithm based on local SGD updates.
-We answer this question in the affirmative in the overparameterized
-quadratic problem setting, extending recent results {[}CITE: Sidford,
-Belkin{]} in the single agent setting. In particular, we show that
-specialized to overparameterized quadratic objectives, the exponential
-convergence bound can be further sharpened, with a similar linear
-speedup in the number of workers. This rate will then be further improved
-using a new accelerated FedAvg algorithm with momentum-based updates. 
+We show that the bound in \ref{thm:overparameterized_general} can
+be improved for quadratic problems with explicit dependence on condition
+number, which is then used to illustrate the acceleration of our proposed
+FedAvg with MaSS updates over SGD updates.
 \begin{theorem}
-	For the overparamterized quadratic problem, FedAvg with local SGD
-	updates and communication every $E$ iterations with constant step
-	size $\alpha_{t}=\frac{1}{4E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})}$
-	gives the exponential convergence guarantee 
+	\label{thm:overparameterized_quadratic}For the overparamterized quadratic
+	problem, FedAvg with local SGD updates and communication every $E$
+	iterations with constant step size $\overline{\alpha}=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})})$
+	has geometric convergence 
 	\begin{align*}
-	\mathbb{E}F(\overline{w}_{t}) & \leq L(1-\frac{N}{4E(\nu_{\max}\kappa_{1}+(N-\nu_{\min}))})^{t}\|w_{0}-w^{\ast}\|^{2}
+	\mathbb{E}F(\overline{\mathbf{w}}_{t}) & \leq\mathcal{O}\left(L(1-\frac{N}{E(\nu_{\max}\kappa_{1}+(N-\nu_{\min}))})^{t}\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right)
 	\end{align*}
 \end{theorem}
-%
 
-We see that when $N=O(\kappa_{1})$, the convergence rate is $O((1-\frac{N}{E\kappa_{1}})^{t})=O(\exp(-\frac{N}{E\kappa_{1}}t))$,
+When $N=\mathcal{O}(\kappa_{1})$, the convergence rate is $\mathcal{O}((1-\frac{N}{E\kappa_{1}})^{t})=\mathcal{O}(\exp(-\frac{N}{E\kappa_{1}}t))$,
 which exhibits linear speedup in the number of workers, as well as
 a $1/\kappa_{1}$ dependence on the condition number $\kappa_{1}$.
-In the last part of this section, we will demonstrate that a new FedAvg
-algorithm based on a local momentum update {[}CITE Belkin{]} improves
-upon this dependence to $1/\sqrt{\kappa_{1}\tilde{\kappa}}$.
-
-Following {[}CITE{]}, we propose the accelerated FedAvg algorithm
-with local MaSS updates: 
+Following \cite{liu2018accelerating}, we propose the \textbf{MaSS
+	accelerated FedAvg algorithm} (FedMaSS):
 \begin{align*}
-w_{t+1}^{k} & =\begin{cases}
-u_{t}^{k}-\eta_{1}^{k}g_{t,k} & \text{if }t+1\notin\mathcal{I}_{E}\\
-\sum_{k=1}^{N}p_{k}\left[u_{t}^{k}-\eta_{1}^{k}g_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E}
+\mathbf{w}_{t+1}^{k} & =\begin{cases}
+\mathbf{u}_{t}^{k}-\eta_{1}^{k}\mathbf{g}_{t,k} & \text{if }t+1\notin\mathcal{I}_{E}\\
+\sum_{k=1}^{N}p_{k}\left[\mathbf{u}_{t}^{k}-\eta_{1}^{k}\mathbf{g}_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E}
 \end{cases}\\
-u_{t+1}^{k} & =w_{t+1}^{k}+\gamma^{k}(w_{t+1}^{k}-w_{t}^{k})+\eta_{2}^{k}g_{t,k}
+\mathbf{u}_{t+1}^{k} & =\mathbf{w}_{t+1}^{k}+\gamma^{k}(\mathbf{w}_{t+1}^{k}-\mathbf{w}_{t}^{k})+\eta_{2}^{k}\mathbf{g}_{t,k}
 \end{align*}
-where we note that the natural parameter is $w_{t}$, while $u_{t}$
-is an auxiliary parameter, which we initialize to be $u_{0}^{k}$,
-$g_{t,k}=\nabla F_{k}(u_{t}^{k},x_{t}^{k})$ is the stochastic gradient
-and $g_{t}=\sum_{k=1}^{N}p_{k}g_{t,k}$ is the averaged stochastic
-gradient. When $\eta_{2}^{k}\equiv0$, this reduces to the accelerated
-FedAvg algorithm with Nesterov updates.
-
-Also note that the update can equivalently be written as 
+When $\eta_{2}^{k}\equiv0$, this algorithm reduces to the Nesterov
+accelerated FedAvg algorithm. In the next theorem, we demonstrate
+that FedMaSS improves the convergence to $\mathcal{O}(\exp(-\frac{N}{E\sqrt{\kappa_{1}\tilde{\kappa}}}t))$.
+To our knowledge, this represents the first acceleration result of
+FedAvg with momentum updates over SGD updates.
+\begin{theorem}
+	\label{thm:overparameterized_MaSS}For the overparamterized quadratic
+	problem, FedMaSS with communication every $E$ iterations and constant
+	step sizes $\overline{\eta_{1}}=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})}),\overline{\eta_{2}}=\frac{\overline{\eta_{1}}(1-\frac{1}{\tilde{\kappa}})}{1+\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}}},\overline{\gamma}=\frac{1-\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}}}{1+\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}}}$
+	has geometric convergence
 	\begin{align*}
-v_{t+1}^{k} & =(1-\alpha^{k})v_{t}^{k}+\alpha^{k}u_{t}^{k}-\delta^{k}g_{t,k}\\
-w_{t+1}^{k} & =\begin{cases}
-u_{t}^{k}-\eta^{k}g_{t,k} & \text{if }t+1\notin\mathcal{I}_{E}\\
-\sum_{k=1}^{N}p_{k}\left[u_{t}^{k}-\eta^{k}g_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E}
-\end{cases}\\
-u_{t+1}^{k} & =\frac{\alpha^{k}}{1+\alpha^{k}}v_{t+1}^{k}+\frac{1}{1+\alpha^{k}}w_{t+1}^{k}
-\end{align*}
-where there is a bijection between the parameters $\frac{1-\alpha^{k}}{1+\alpha^{k}}=\gamma^{k},\eta^{k}=\eta_{1}^{k},\frac{\eta^{k}-\alpha^{k}\delta^{k}}{1+\alpha^{k}}=\eta_{2}^{k}$,
-and we further introduce an auxiliary parameter $v_{t}^{k}$, which
-is initialized at $v_{0}^{k}$. We also note that when $\delta^{k}=\frac{\eta^{k}}{\alpha^{k}}$,
-the update reduces to the Nesterov accelerated SGD. This version of
-the FedAvg algorithm with local MaSS updates is used for analyzing
-the exponential convergence.
-
-As before, define the virtual sequences $\overline{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$,
-$\overline{v}_{t}=\sum_{k=1}^{N}p_{k}v_{t}^{k}$, $\overline{u}_{t}=\sum_{k=1}^{N}p_{k}u_{t}^{k}$,
-and $\overline{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$. We have
-$\mathbb{E}g_{t}=\overline{g}_{t}$ and $\overline{w}_{t+1}=\overline{u}_{t}-\eta_{t}g_{t}$,
-$\overline{v}_{t+1}=(1-\alpha^{k})\overline{v}_{t}+\alpha^{k}\overline{w}_{t}-\delta^{k}g_{t}$,
-and $\overline{u}_{t+1}=\frac{\alpha^{k}}{1+\alpha^{k}}\overline{v}_{t+1}+\frac{1}{1+\alpha^{k}}\overline{w}_{t+1}$.
-
-We now present the exponential convergence result in overparameterized
-quadratic problems using FedAvg with local MaSS updates. 
-\begin{theorem}
-For the overparamterized quadratic problem, FedAvg with local MaSS
-updates and communication every $E$ iterations with constant step
-sizes 
-\begin{align*}
-\eta_{t}=\frac{1}{4E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})}, & \alpha_{t}=\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}},\delta_{t}=\frac{\eta_{t}}{\alpha_{t}\tilde{\kappa}}
-\end{align*}
-gives the exponential convergence guarantee 
-\begin{align*}
-\mathbb{E}F(\overline{w}_{t}) & \leq C(1-\frac{1}{4E}\frac{N}{l\nu_{\max}\sqrt{\kappa_{1}\tilde{\kappa}}+(N-\nu_{\min})})^{t}\|w_{0}-w^{\ast}\|^{2}
+	\mathbb{E}F(\overline{w}_{t}) & \leq\mathcal{O}\left((1-\frac{1}{E}\frac{N}{l\nu_{\max}\sqrt{\kappa_{1}\tilde{\kappa}}+(N-\nu_{\min})})^{t}\|w_{0}-w^{\ast}\|^{2}\right)
 	\end{align*}
 \end{theorem}
-Compared to the convergence rate of FedAvg with local SGD updates,
-we see that when $N=O(\sqrt{\kappa_{1}\tilde{\kappa}})$, the convergence
-rate is $O((1-\frac{N}{E\sqrt{\kappa_{1}\tilde{\kappa}}})^{t})=O(\exp(-\frac{N}{E\sqrt{\kappa_{1}\tilde{\kappa}}}t))$
-as opposed to $O(\exp(-\frac{N}{E\kappa_{1}}t))$. Since $\kappa_{1}\geq\tilde{\kappa}$,
-this implies a speedup factor of $\sqrt{\frac{\kappa_{1}}{\tilde{\kappa}}}$
-for the FedAvg with local MaSS updates. On the other hand, the same
-linear speedup in the number of workers holds for $N$ in a smaller
-range of values.
+\textbf{Speedup of FedMaSS over FedAvg} Compared to the convergence
+rate of FedAvg with local SGD updates, when $N=\mathcal{O}(\sqrt{\kappa_{1}\tilde{\kappa}})$,
+the convergence rate is $\mathcal{O}((1-\frac{N}{E\sqrt{\kappa_{1}\tilde{\kappa}}})^{t})=\mathcal{O}(\exp(-\frac{N}{E\sqrt{\kappa_{1}\tilde{\kappa}}}t))$
+as opposed to $\mathcal{O}(\exp(-\frac{N}{E\kappa_{1}}t))$. Since
+$\kappa_{1}\geq\tilde{\kappa}$, this implies a speedup factor of
+$\sqrt{\frac{\kappa_{1}}{\tilde{\kappa}}}$ for FedMaSS. On the other
+hand, the same linear speedup in the number of workers holds for $N$
+in a smaller range of values.

### 2020-06-02 source/app_sumarization.tex
Update app_sumarization.tex

--- 2020-06-02 source/app_sumarization.tex
@@
 \cite{zanette2019tighter}  & $\cO(\frac{1}{\sqrt{NT}})$ & $\cO(N^{-3/2}T^{1/2})$     			& Full        & Bounded gradient             & Convex  \\\hline
-This work      & $\cO\left(\frac{1}{\sqrt{NT}}\right)$ &  $\cO(N^{-1/2}T^{1/2})^{\dagger\dagger}$			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
+This work      & $\cO\left(\frac{1}{\sqrt{NT}}\right)$ &  $\cO(N^{-3/4}T^{1/4})^{\dagger\dagger}$			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
 This work & $\cO\left(\exp(-\frac{NT}{E\kappa_1})\right)$ & $ \cO(T^{\beta})$                   &  Partial     & Bounded gradient    & Overparameterized (LR)$^{\ddagger}$ \\ \hline

### 2020-06-02 source/sec_Nesterov.tex
minor edits
updated nesterov statement
updated sgd and nesterov

@@
 \section{Linear Speedup Analysis of Nesterov Accelerated FedAvg}
+\label{sec:Nesterov}
 
@@
 only convergence analyses of FedAvg with momentum-based stochastic
-updates focus on the non-convex smooth case \cite{yu2019linear,huo2020faster,li2018federated}.
+updates focus on the non-convex smooth case \cite{huo2020faster,yu2019linear,li2018federated}.
 In this section, we complete the picture with $\mathcal{O}(1/NT)$
 and $\mathcal{O}(1/\sqrt{NT})$ convergence results for Nesterov-accelerated
-FedAvg for convex objectives. As we know from stochastic optimization,
-Nesterov updates may fail to accelerate over SGD, so in general we
-cannot hope to obtain acceleration results over FedAvg. However, in
-the following section we specialize to overparameterized problems
-where we demonstrate that a particular FedAvg variant with momentum
-updates is able to accelerate over the original FedAvg algorithm.
+FedAvg for convex objectives that match the original FedAvg results.
+As we know from stochastic optimization, Nesterov and other momumtum
+updates may fail to accelerate over SGD \cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2016influence}.
+Therefore in Section~\ref{sec:overparameterized} we will specialize to overparameterized
+problems where we demonstrate that a particular FedAvg variant with
+momentum updates is able to accelerate over the original FedAvg algorithm.
 
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right)
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}\leq\mathcal{O}(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}\leq\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right)
 	\end{align*}
@@
 %
-\begin{remark}
 To our knowledge, this is the first convergence result for Nesterov
@@
 communication rounds with full participation.
-\end{remark}
 
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F^{\ast} & =\mathcal{O}(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
+	\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T})
 	\end{align*}
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F^{\ast} & =\mathcal{O}(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}G^{2}}{KT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
+	\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T})
 	\end{align*}
@@
 %
-\begin{remark}
 	It is possible to extend the analyses in this section to accelerated
@@
 	and it is possible to improve its convergence rate with momentum based
-	local updates. 
-\end{remark}

### 2020-06-02 source/sec_SGD.tex
updated sgd and nesterov

@@
 \textbf{Choice of $E$ and linear speedup. }With full participation,
-as long as $E=\mathcal{O}(\sqrt{T^{1/2}/N^{3/2}})$, the convergence
+as long as $E=\mathcal{O}(T^{1/4}/N^{3/4})$, the convergence
 rate is $\mathcal{O}(1/\sqrt{NT})$ with $\mathcal{O}(N^{3/4}T^{3/4})$

### 2020-06-02 source/sec_problem.tex
updated assmption section

@@
 \textbf{Justfication for Assumptions~\ref{ass:boundedvariance} and~\ref{ass:subgrad2}.} These standard assumptions have been made in many previous
-works in federated learning ~\cite{yu2019parallel,li2019convergence,stich2018local}. Here we provide one more justification
+works in federated learning\cite{yu2019parallel,li2019convergence,stich2018local}. Here we provide one more justification
 for their generality: even though they require bounds for all $\mathbf{w}$,
@@
 Such assumptions are more restrictive than our assumptions, because
-they imply $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$ for all local
+they effectively require $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$ for all local
 objectives. For such ``overparameterized'' settings, we will prove

### 2020-06-02 source/sec_SGD.tex
sec 3 minor
sec 3.1:
1. remove the claim of first obtained linear speedup results in L152
2. rephrase comparison with related work.
3. typo in communication complexity of [19].

sec 3.2 remove N
change nonconvex to non-convex to be consistant accross the paper.

@@
 the convergence of the optimality gap of averaged parameters across
-devices is $\mathcal{O}(1/NT)$ where $N$ is the number of active
-devices during each communication round, while for convex and smooth
+devices is $\mathcal{O}(1/NT)$, while for convex and smooth
 objectives, the rate is $\mathcal{O}(1/\sqrt{NT})$. 
+The proof of FedAvg is deferred in Appendix Section~\ref{sec:app:fedavg}.
 
@@
 setting. This implies that in both settings, there is a linear speedup
-in the number of active workers during the communication round. To
-our knowledge, this is the first result that explicitly demonstrates
-linear speedup in the number of active workers in the setting with
-both non i.i.d. data and partial participation. \\
+in the number of active workers during the communication round. \\
+% To our knowledge, this is the first result that explicitly demonstrates
+% linear speedup in the number of active workers in the setting with
+% both non i.i.d. data and partial participation. \\
 \textbf{Communication Complexity.} Our bound implies a distinction
@@
 analyses. \\
-\textbf{Comparison with related works.} \cite{haddadpour2019convergence}
-also proves a linear speedup $\mathcal{O}(1/NT)$ result for strongly
-convex and smooth objectives in the heterogeneous data with partial
-participation setting, with $\mathcal{O}(N^{1/3}T^{2/3})$ communication
-complexity. However, their results are derived based on the more restrictive
-bounded gradient diversity assumption. As discussed before, if $\frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$
-is bounded, then $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$ for all
-$k$. In this overparameterized setting, we show in Section~\ref{sec:overparameterized}
-that the convergence is in fact exponential, thus improving on the
-$\mathcal{O}(1/NT)$ rate in \cite{haddadpour2019convergence} with
+\textbf{Comparison with related works.} To better understand the significance of obtained bound, we compare 
+our rates to the best-known results in the same setting. 
+In~\cite{haddadpour2019convergence}, the authors prove a linear speedup $\mathcal{O}(1/NT)$ result for strongly
+convex and smooth objectives, with $\mathcal{O}(N^{1/3}T^{1/3})$ communication complexity when considering i.i.d. data. 
+However, their results build upon the bounded gradient diversity assumption, which leads to a more restricted overparameterized setting (see discussions in Section~\ref{sec:assumptions}).
+In this overparameterized setting, we prove a geometric convergence rate (see Section~\ref{sec:overparameterized}),
+thus improving on the rate in \cite{haddadpour2019convergence} with
 better communication complexity.
@@
 Next we provide linear speedup analyses of FedAvg with convex and
-smooth objectives and show that the optimality gap is $\mathcal{O}(1/\sqrt{NT})$
-where $N$ is the number of participating devices. This result complements
-the strongly convex case in the previous part, as well as the nonconvex
+smooth objectives and show that the optimality gap is $\mathcal{O}(1/\sqrt{NT})$. 
+This result complements the strongly convex case in the previous part, as well as the non-convex
 smooth setting in \cite{jiang2018linear,yu2019parallel,haddadpour2019convergence},
@@
 a requirement $\alpha_{t}=\mathcal{O}(\sqrt{N/T})$ is also present
-in the works \cite{haddadpour2019convergence,yu2019parallel} on nonconvex
+in the works \cite{haddadpour2019convergence,yu2019parallel} on non-convex
 problems with $\mathcal{O}(1/\sqrt{NT})$ linear speedup convergence results.

### 2020-06-02 source/sec_Nesterov.tex
minor update
Sec 4. typos.

@@
 FedAvg for convex objectives that match the original FedAvg results.
-As we know from stochastic optimization, Nesterov and other momumtum
+As we know from stochastic optimization, Nesterov and other momentum
 updates may fail to accelerate over SGD \cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2016influence}.
@@
 momentum updates is able to accelerate over the original FedAvg algorithm. 
+The convergence proof of Nesterov-accelerated FedAvg is deferred in Appendix Section~\ref{sec:app:Nesterovfedavg}.
 
@@
 \begin{align*}
-\mathbf{v}_{t+1}^{k} & =\mathbf{w}_{t}^{k}-\alpha_{t}g_{t,k}\\
+\mathbf{v}_{t+1}^{k} & =\mathbf{w}_{t}^{k}-\alpha_{t}\mathbf{g}_{t,k}\\
 \mathbf{w}_{t+1}^{k} & =\begin{cases}
@@
     stochastic optimization setting, none of these methods will be able
-	to achieve a better rate than the original FedAvg with SGD in general.
+    to achieve a better rate than the original FedAvg with SGD updates in general.
     For this reason, we will instead turn to the overparameterized setting
     \cite{ma2017power,liu2018accelerating,canziani2016analysis} in the
-	next section where we show that FedAvg enjoys exponential convergence
-	and it is possible to improve its convergence rate with momentum based
+    next section where we show that FedAvg enjoys geometric convergence
+    and it is possible to improve its convergence rate with momentum-based updates.

### 2020-06-03 source/sec_overparameterized.tex
Sec 5
1.  Shrink the summary in first paragraph.
2.  Move discussion of related work on accelerated FedAvg from intro to 5.3.
3. remove some computation in L244.

@@
 
-\section{Exponential Convergence of FedAvg in the Overparameterized Setting}
+\section{Geometric Convergence of FedAvg in the Overparameterized Setting}
 \label{sec:overparameterized}
 
-In this section, we turn to the special setting of overparameterized
-problems, where a non-negative loss function admits a zero-loss global
-minimizer. Such problems are very common in machine learning systems
-where the number of parameters far exceeds the number of data points,
-and perfect fitting is possible. There have been a line of recent
-works showing that SGD and accelerated methods achieve exponential
-convergence in similar settings thanks to the property of ``automatic
-variance reduction'' \cite{ma2017power,moulines2011non,needell2014stochastic,schmidt2013fast,strohmer2009randomized}.
-The natural question is whether such a result also holds in the federated
-learning setting. We show that this is indeed the case and establish
-the geometric (linear) convergence of FedAvg with local SGD updates
-with constant step size for general strongly convex and smooth overparameterized
-problems. In addition, we show that the convergence rate speeds up
-linearly in the number of workers $N$ when $N$ is below some problem-dependent
-threshold, while it decreases with $E$ through $1/E$.
+% In this section, we turn to the special setting of overparameterized
+% problems, where a non-negative loss function admits a zero-loss global
+% minimizer. Such problems are prevalent especially at the age of deep learning,
+% where the number of parameters far exceeds the number of data points,
+% and perfect fitting is possible. There have been a line of recent
+% works showing that SGD and accelerated methods achieve geometric
+% convergence in similar settings thanks to the property of \textit{automatic
+% variance reduction}~\cite{ma2017power,moulines2011non,needell2014stochastic,schmidt2013fast,strohmer2009randomized}.
+% The natural question is whether such a result also holds in the federated
+% learning setting. 
+% We show that this is indeed the case and establish
+% the geometric (linear) convergence of FedAvg with local SGD updates
+% with constant step size for general strongly convex and smooth overparameterized
+% problems. In addition, we show that the convergence rate speeds up
+% linearly in the number of workers $N$ when $N$ is below some problem-dependent
+% threshold, while it decreases with $E$ through $1/E$.
 
-To our knowledge, this is the first result on the exponential convergence
-of FedAvg algorithms in the overparameterized setting with linear
-speedup in the number of workers and explicit dependence on the communication
-interval $E$. We next sharpen this result in the special case of
-quadratic problems, and show that it is $\mathcal{O}(\exp(-\frac{NT}{E\kappa_{1}}))$,
-for $\kappa_{1}$ a condition number of the system.
-
-Lastly, we deal with the question of whether FedAvg with momentum-based
+Overparameterization is a prevalent machine learning setting where the
+statistical model has much more parameters than the number of training samples
+and the existence of parameter choices with zero training loss is
+assured~\cite{allen2018convergence,zhang2016understanding}.  Due to the
+property of \textit{automatic variance reduction} in overparameterization,  a
+line of recent works proved that SGD and accelerated methods achieve geometric
+convergence~\cite{ma2017power,moulines2011non,needell2014stochastic,schmidt2013fast,strohmer2009randomized}.
+A natural question is whether such a result still holds in the federated
+learning setting. 
+In this section, we provide the first geometric convergence rate
+of FedAvg for the overparameterized strongly convex and smooth problems,
+and we show that it preserves the linear speedup at the same time.  
+We then sharpen this result in the special case of
+quadratic problems.
+Inspiring by the recent advances in SGD~\cite{liu2019accelerating,jain2017accelerating}, 
+we further propose a novel momentum-based FedAvg algorithm, which enjoys an
+improved convergence rate over FedAvg. 
+\begin{comment}
+we deal with the question of whether FedAvg with momentum-based
 local updates can outperform FedAvg with SGD updates. In contrast
@@
 over FedAvg with local SGD updates. 
+\end{comment}
 
@@
 with $F_{k}(\mathbf{w})=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\ell(\mathbf{w};\mathbf{x}_{k}^{j})$.
-In this section, we consider the standard ERM setting where each $\ell$
+In this section, we consider the standard Empirical Risk Minimization (ERM) setting where $\ell$
 is non-negative, $l$-smooth, and convex, and as before, we assume
-that each $F_{k}(\mathbf{w})=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\ell(\mathbf{w};\mathbf{x}_{k}^{j})$
-is $L$-smooth and $\mu$-strongly convex. Note that $l\geq L$. This
+that each $F_{k}(\mathbf{w})$ is $L$-smooth and $\mu$-strongly convex. Note that $l\geq L$. This
 setup includes many important problems in practice. In the overparameterized
-setting, we assume that there exists $\mathbf{w}^{\ast}\in\arg\min_{w}\sum_{k=1}^{N}p_{k}F_{k}(\mathbf{w})$
+setting, there exists $\mathbf{w}^{\ast}\in\arg\min_{w}\sum_{k=1}^{N}p_{k}F_{k}(\mathbf{w})$
 such that $\ell(\mathbf{w}^{\ast};\mathbf{x}_{k}^{j})=0$ for all
@@
 	with constant step size $\overline{\alpha}=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})})$
-	has geometric convergence
+	has geometric convergence,
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{w}}_{T}) & \leq\frac{L}{2}(1-\overline{\alpha})^{T}\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}=\mathcal{O}\left(L\exp(-\frac{\mu}{E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})}t)\cdot\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right)
+	\mathbb{E}F(\overline{\mathbf{w}}_{T}) & \leq\frac{L}{2}(1-\overline{\alpha})^{T}\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}=\mathcal{O}\left(L\exp(-\frac{\mu}{E}\frac{NT}{l\nu_{\max}+L(N-\nu_{\min})})\cdot\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right).
 	\end{align*}
@@
 \textbf{Linear speedup and Communication Complexity} When $l>L$,
-the speedup factor is on the order of $\frac{N}{E(l\nu_{\max}+L(N-\nu_{\min}))}/\frac{1}{l\nu_{\max}+L(1-\nu_{\min})}\approx\frac{Nl}{E(l+L(N-1))}=\mathcal{O}(N/E)$
+the speedup factor is on the order of $\mathcal{O}(N/E)$ %\frac{N}{E(l\nu_{\max}+L(N-\nu_{\min}))}/\frac{1}{l\nu_{\max}+L(1-\nu_{\min})}\approx\frac{Nl}{E(l+L(N-1))}=
 for $N\leq\frac{l}{L}+1$, i.e. FedAvg with $N$ workers and communication
@@
 of workers. This matches the findings in \cite{ma2017power}. Our
-result also illustrates that $E$ can be taken $\mathcal{O}(t^{\beta})$
+result also illustrates that $E$ can be taken $\mathcal{O}(T^{\beta})$
 for any $\beta<1$, achieving better communication efficiency than
@@
 \subsection{Overparameterized Quadratic Problems}
-
-In the remaining part of the section, we specialize to quadratic problems
-and demonstrate how the general bound in \ref{thm:overparameterized_general}
+We now turn to quadratic problems
+and demonstrate how the general bound in Theorem~\ref{thm:overparameterized_general}
 can be sharpened to explicitly depend on the condition number of the
@@
 
-The local device objectives are now given by the sum of squares $F_{k}(\mathbf{w})=\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(\mathbf{w}^{T}\mathbf{x}_{k}^{j}-z_{k}^{j})^{2}$,
-and there exists $\mathbf{w}^{\ast}$ such that $F(\mathbf{w}^{\ast})=\sum_{k}p_{k}F_{k}(\mathbf{w}^{\ast})\equiv0$.
-Define the local Hessian matrix as $H^{k}:=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\mathbf{x}_{k}^{j}(\mathbf{x}_{k}^{j})^{T}$
-and the averaged Hessian matrix as $H:=\sum_{k=1}^{N}p_{k}H^{k}$.
+The local device objectives are now given by the sum of squares {\small$F_{k}(\mathbf{w})=\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(\mathbf{w}^{T}\mathbf{x}_{k}^{j}-z_{k}^{j})^{2}$},
+and there exists $\mathbf{w}^{\ast}$ such that $F(\mathbf{w}^{\ast})\equiv0$.
+Define the local Hessian matrix as $H^{k}:=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\mathbf{x}_{k}^{j}(\mathbf{x}_{k}^{j})^{T}$,
+the averaged Hessian matrix as $H:=\sum_{k=1}^{N}p_{k}H^{k}$, and the stochastic Hessian matrix as $\tilde{H}_{t}^{k}:=\xi_{t}^{k}(\xi_{t}^{k})^{T}$. 
 \begin{comment}
@@
 \end{comment}
-{} %
 \begin{comment}
@@
 \end{comment}
+\begin{comment}
 Let $\xi_{t}^{k}$ be the stochastic sample on the $k$th device at
@@
 as the stochastic Hessian matrix. %	
+\end{comment}
 \begin{comment}
@@
 \end{comment}
-{} Define $l$ to be the smallest positive number such that $\mathbb{E}\|\xi_{t}^{k}\|^{2}$$\mathbf{\xi}_{t}^{k}$($\mathbf{\xi}_{t}^{k})^{T}\preceq lH^{k}$
+Define $l$ to be the smallest positive number such that $\mathbb{E}\|\xi_{t}^{k}\|^{2}$$\mathbf{\xi}_{t}^{k}$($\mathbf{\xi}_{t}^{k})^{T}\preceq lH^{k}$
 for all $k$. Note that $l\leq\max_{k,j}\|\mathbf{x}_{k}^{j}\|^{2}$.
@@
 condition number $\tilde{\kappa}$ as the smallest positive real number
-such that 
-\begin{align*}
-\mathbb{E}\left[\langle\xi_{t}^{k}(H^{k})^{-1},\xi_{t}^{k}\rangle\xi_{t}^{k}(\xi_{t}^{k})^{T}\right] & \preceq\tilde{\kappa}H^{k}
-\end{align*}
-for all $k$. The condition numbers $\kappa_{1}$ and $\tilde{\kappa}$
+such that $\mathbb{E}\left[\langle\xi_{t}^{k}(H^{k})^{-1},\xi_{t}^{k}\rangle\xi_{t}^{k}(\xi_{t}^{k})^{T}\right] \  \preceq\tilde{\kappa}H^{k}$, for all $k$. 
+The condition numbers $\kappa_{1}$ and $\tilde{\kappa}$
 are important in the characterization of convergence rates for FedAvg
@@
 	iterations with constant step size $\overline{\alpha}=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})})$
-	has geometric convergence 
+	has geometric convergence:
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{w}}_{t}) & \leq\mathcal{O}\left(L(1-\frac{N}{E(\nu_{\max}\kappa_{1}+(N-\nu_{\min}))})^{t}\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right)
+	\mathbb{E}F(\overline{\mathbf{w}}_{T}) & \leq\mathcal{O}\left(L(1-\frac{N}{E(\nu_{\max}\kappa_{1}+(N-\nu_{\min}))})^{T}\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right).
 	\end{align*}
@@
 
-When $N=\mathcal{O}(\kappa_{1})$, the convergence rate is $\mathcal{O}((1-\frac{N}{E\kappa_{1}})^{t})=\mathcal{O}(\exp(-\frac{N}{E\kappa_{1}}t))$,
+When $N=\mathcal{O}(\kappa_{1})$, the convergence rate is $\mathcal{O}((1-\frac{N}{E\kappa_{1}})^{T})=\mathcal{O}(\exp(-\frac{NT}{E\kappa_{1}}))$,
 which exhibits linear speedup in the number of workers, as well as
 a $1/\kappa_{1}$ dependence on the condition number $\kappa_{1}$.
-Following \cite{liu2018accelerating}, we propose the \textbf{MaSS
+Inspired by \cite{liu2018accelerating}, we propose the \textbf{MaSS
 	accelerated FedAvg algorithm} (FedMaSS):
@@
 \mathbf{w}_{t+1}^{k} & =\begin{cases}
-\mathbf{u}_{t}^{k}-\eta_{1}^{k}\mathbf{g}_{t,k} & \text{if }t+1\notin\mathcal{I}_{E}\\
-\sum_{k=1}^{N}p_{k}\left[\mathbf{u}_{t}^{k}-\eta_{1}^{k}\mathbf{g}_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E}
+\mathbf{u}_{t}^{k}-\eta_{1}^{k}\mathbf{g}_{t,k} & \text{if }t+1\notin\mathcal{I}_{E},\\
+\sum_{k=1}^{N}p_{k}\left[\mathbf{u}_{t}^{k}-\eta_{1}^{k}\mathbf{g}_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E},
 \end{cases}\\
-\mathbf{u}_{t+1}^{k} & =\mathbf{w}_{t+1}^{k}+\gamma^{k}(\mathbf{w}_{t+1}^{k}-\mathbf{w}_{t}^{k})+\eta_{2}^{k}\mathbf{g}_{t,k}
+\mathbf{u}_{t+1}^{k} & =\mathbf{w}_{t+1}^{k}+\gamma^{k}(\mathbf{w}_{t+1}^{k}-\mathbf{w}_{t}^{k})+\eta_{2}^{k}\mathbf{g}_{t,k}. 
 \end{align*}
@@
 accelerated FedAvg algorithm. In the next theorem, we demonstrate
-that FedMaSS improves the convergence to $\mathcal{O}(\exp(-\frac{N}{E\sqrt{\kappa_{1}\tilde{\kappa}}}t))$.
-To our knowledge, this represents the first acceleration result of
+that FedMaSS improves the convergence to $\mathcal{O}(\exp(-\frac{NT}{E\sqrt{\kappa_{1}\tilde{\kappa}}}))$.
+To our knowledge, this is the first accelerated result of
 FedAvg with momentum updates over SGD updates.
@@
 	problem, FedMaSS with communication every $E$ iterations and constant
-	step sizes $\overline{\eta_{1}}=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})}),\overline{\eta_{2}}=\frac{\overline{\eta_{1}}(1-\frac{1}{\tilde{\kappa}})}{1+\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}}},\overline{\gamma}=\frac{1-\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}}}{1+\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}}}$
-	has geometric convergence
+	step sizes $\overline{\eta}_1=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})}),\overline{\eta}_{2}=\frac{\overline{\eta}_{1}(1-\frac{1}{\tilde{\kappa}})}{1+\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}}},\overline{\gamma}=\frac{1-\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}}}{1+\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}}}$
+	has geometric convergence:
 	\begin{align*}
-	\mathbb{E}F(\overline{w}_{t}) & \leq\mathcal{O}\left((1-\frac{1}{E}\frac{N}{l\nu_{\max}\sqrt{\kappa_{1}\tilde{\kappa}}+(N-\nu_{\min})})^{t}\|w_{0}-w^{\ast}\|^{2}\right)
+	\mathbb{E}F(\overline{\vw}_{T}) & \leq\mathcal{O}\left((1-\frac{1}{E}\frac{N}{l\nu_{\max}\sqrt{\kappa_{1}\tilde{\kappa}}+(N-\nu_{\min})})^{T}\|w_{0}-w^{\ast}\|^{2}\right).
 	\end{align*}
 \end{theorem}
-\textbf{Speedup of FedMaSS over FedAvg} Compared to the convergence
-rate of FedAvg with local SGD updates, when $N=\mathcal{O}(\sqrt{\kappa_{1}\tilde{\kappa}})$,
-the convergence rate is $\mathcal{O}((1-\frac{N}{E\sqrt{\kappa_{1}\tilde{\kappa}}})^{t})=\mathcal{O}(\exp(-\frac{N}{E\sqrt{\kappa_{1}\tilde{\kappa}}}t))$
-as opposed to $\mathcal{O}(\exp(-\frac{N}{E\kappa_{1}}t))$. Since
+\textbf{Speedup of FedMaSS over FedAvg} 
+To better understand the significance of the above result, we briefly discuss the related works in SGD.
+Nesterov and Heavy Ball updates are known to fail to accelerate over SGD in both the overparameterized
+and convex settings~\cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2016influence}.
+Thus in general one cannot hope to obtain acceleration results for
+the FedAvg algorithm with Nesterov and Heavy Ball updates. 
+Luckily, recent works in SGD~\cite{jain2017accelerating,liu2018accelerating} introduced
+an additional compensation term to the Nesterov updates to address the non-acceleration issue. Surprisingly, 
+we show the same approach can effectively improve the rate of FedAvg.
+Comparing the convergence rate of FedMass (Theorem~\ref{thm:overparameterized_MaSS}) and FedAvg (Theorem~\ref{thm:overparameterized_quadratic}), when $N=\mathcal{O}(\sqrt{\kappa_{1}\tilde{\kappa}})$,
+the convergence rate is $\mathcal{O}((1-\frac{N}{E\sqrt{\kappa_{1}\tilde{\kappa}}})^{T})=\mathcal{O}(\exp(-\frac{NT}{E\sqrt{\kappa_{1}\tilde{\kappa}}}))$
+as opposed to $\mathcal{O}(\exp(-\frac{NT}{E\kappa_{1}}))$. Since
 $\kappa_{1}\geq\tilde{\kappa}$, this implies a speedup factor of

### 2020-06-03 source/sec_overparameterized.tex
sec3,4,5 typo

@@
 \label{sec:overparameterized}
-
 % In this section, we turn to the special setting of overparameterized
@@
 % threshold, while it decreases with $E$ through $1/E$. 
-
 Overparameterization is a prevalent machine learning setting where the
@@
 we further propose a novel momentum-based FedAvg algorithm, which enjoys an
-improved convergence rate over FedAvg. 
+improved convergence rate over FedAvg. The proof is deferred to Appendix Section~\ref{sec:interpolation}.
 \begin{comment}

### 2020-06-03 source/sec_Nesterov.tex
Sec 4, 8 minor
sec3,4,5 typo
minor updates

@@
 momentum updates is able to accelerate over the original FedAvg algorithm. 
-The convergence proof of Nesterov-accelerated FedAvg is deferred in Appendix Section~\ref{sec:app:Nesterovfedavg}.
+The convergence proof of Nesterov-accelerated FedAvg is deferred to Appendix Section~\ref{sec:app:Nesterovfedavg}.
 
@@
 \begin{align*}
-\mathbf{v}_{t+1}^{k} & =\mathbf{w}_{t}^{k}-\alpha_{t}\mathbf{g}_{t,k}\\
+\mathbf{v}_{t+1}^{k} & =\mathbf{w}_{t}^{k}-\alpha_{t}\mathbf{g}_{t,k},\\
 \mathbf{w}_{t+1}^{k} & =\begin{cases}
-\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E}\\
-\sum_{k=1}^{N}p_{k}\left[\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E}
+\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E},\\
+\sum_{k=1}^{N}p_{k}\left[\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E}.
 \end{cases}
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right)
+	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right),
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}\leq\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right)
+	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}\leq\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).
 	\end{align*}
@@
 
-We now provide linear speedup analysis of Nesterov Accelerated FedAvg
-with convex and smooth objectives and show that the optimality gap
-is $\mathcal{O}(1/\sqrt{NT})$ where $N$ is the number of participating
-devices. This result complements the strongly convex case in the previous
+We now show that the optimality gap of Nesterov Accelerated FedAvg has $\mathcal{O}(1/\sqrt{NT})$ rate. This result complements the strongly convex case in the previous
 part, as well as the non-convex smooth setting in \cite{huo2020faster,yu2019linear,li2018federated},
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T})
+	\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T}\right),
 	\end{align*}
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T})
+	\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right).
 	\end{align*}
@@
 %
-	It is possible to extend the analyses in this section to accelerated
+	It is possible to extend the results in this section to accelerated
     FedAvg algorithms with other momentum-based updates. However, in the

### 2020-06-03 source/sec_overparameterized.tex
minor updates

@@
 quadratic problems.
-Inspiring by the recent advances in SGD~\cite{liu2019accelerating,jain2017accelerating}, 
+Inspired by recent advances in accelerating SGD~\cite{liu2019accelerating,jain2017accelerating}, 
 we further propose a novel momentum-based FedAvg algorithm, which enjoys an
@@
 the speedup factor is on the order of $\mathcal{O}(N/E)$ %\frac{N}{E(l\nu_{\max}+L(N-\nu_{\min}))}/\frac{1}{l\nu_{\max}+L(1-\nu_{\min})}\approx\frac{Nl}{E(l+L(N-1))}=
-for $N\leq\frac{l}{L}+1$, i.e. FedAvg with $N$ workers and communication
+for $N\leq\mathcal{O}(\frac{l}{L})$, i.e. FedAvg with $N$ workers and communication
 every $E$ iterations provides a geometric convergence speedup factor
-of $\mathcal{O}(N/E)$, for $N\leq\frac{l}{L}+1$. When $N$ is above
+of $\mathcal{O}(N/E)$, for $N\leq\mathcal{O}(\frac{l}{L})$. When $N$ is above
 this threshold, however, the speedup is almost constant in the number
@@
 for any $\beta<1$, achieving better communication efficiency than
-the standard setting.
+the standard FL setting.
 
@@
 
-We show that the bound in \ref{thm:overparameterized_general} can
-be improved for quadratic problems with explicit dependence on condition
-number, which is then used to illustrate the acceleration of our proposed
-FedAvg with MaSS updates over SGD updates.
+We show that the bound in Theorem \ref{thm:overparameterized_general} can be improved to $\mathcal{O}(\exp(-\frac{N}{E\kappa_{1}}t))$ for quadratic problems, then propose a variant of FedAvg that has provable acceleration over FedAvg with SGD updates.
+
 \begin{theorem}
@@
 \textbf{Speedup of FedMaSS over FedAvg} 
-To better understand the significance of the above result, we briefly discuss the related works in SGD.
+To better understand the significance of the above result, we briefly discuss related works on accelerating SGD.
 Nesterov and Heavy Ball updates are known to fail to accelerate over SGD in both the overparameterized

### 2020-06-03 source/sec_conclude.tex
exp and conclusions

@@
 \section{Conclusions}
-This paper provides a comprehensive analysis of the convergence rate of FedAvg
+\begin{comment}
+This paper provides a comprehensive and unified analysis of the convergence rate of FedAvg
 and its accelerated variants in a general federated learning problem with heterogeneous local data and partial participation. We show that both Nesterov accelerated FedAvg and FedAvg
@@
 verify the linear speedup of FedAvg and Nesterov accelerated FedAvg for strongly convex, convex smooth, and linear regression problems. The empirical results are well-aligned with our theories.		
+\end{comment}
+This paper provides a comprehensive and unified analysis of the convergence rate of FedAvg
+and its accelerated variants in a general federated learning problem with heterogeneous local data and partial participation. 
+We show that both Nesterov accelerated FedAvg and FedAvg can achieve linear speedup convergence for convex smooth problems and strongly 
+convex smooth problems.  
+Furthermore, this work also makes algorithmic contributions. We not only prove that FedAvg can achieve geometric convergence for overparameterized strongly convex smooth problems 
+but also propose the MaSS accelerated Fedavg algorithm, which has provable speedup in convergence rate over FedAvg on quadratic problems. Last but not least, extensive empirical 
+evaluations demonstrate the linear speedup of FedAvg and Nesterov accelerated FedAvg under various settings.

### 2020-06-03 source/sec_exp.tex
exp and conclusions

@@
 \|\vw\|^2$. The regularization parameter is set to $\lambda = 1/n \approx
-1e-5$. 2) \textbf{Convex smooth objective}: the binary logistic regression
-problem without regularization. $F_k(\vw) = \frac{1}{N_k} \sum_{i=1}^{N_k}
-\log( 1+ \exp(-y_i^k \vw^T\vx_i^k)$. 3) \textbf{Overparameterized setting}:
+2e-5$. 2) \textbf{Convex smooth objective}: the binary logistic regression
+problem without regularization. 3) \textbf{Overparameterized setting}:
 the linear regression problem without adding noise to the label, $F_k(\vw) =
-\frac{1}{N_k} \sum_{i=1}^{N_k} (\vw^T\vx_i^k + b  - y_i^k)^2$.  In the first
-two cases, we use the w8a dataset to conduct empirical evaluation. 
-In the linear regression problem, we use features from w8a dataset and 
-generate a ground truth $w$ to label each instance, which satisfies the
-overparameterized setting. For the detailed experimental setup and
-hyperparameter setting, please refer to the section~\ref{sec:expsupp} in the
-appendix.
+\frac{1}{N_k} \sum_{i=1}^{N_k} (\vw^T\vx_i^k + b  - y_i^k)^2$.  
+% In the first
+% two cases, we use the w8a dataset to conduct empirical evaluation. 
+% In the overparameterized setting, we use features from w8a dataset and 
+% generate a ground truth $w$ to label each instance, which satisfies the
+% overparameterized setting.
 
 % ~\ref{th:strongcvxsmth}~\ref{th:cvxsmoth}~\ref{th:fedmasslr}
-\textbf{Linear speedup.} To verify the linear speedup results of FedAvg in Theorem~\ref{thm:SGD_scvx}~\ref{thm:SGD_cvx}, we evaluate the number of iterations needed to reach
-$\epsilon$-accuracy in three objectives. We initialize all runs with $\vw_0 = \textbf{0}_d$ and measure. For each configuration, we run three random seeds and report
-the best seed results for each configuration $(E, K)$. In each configure, we
-extensive search the stepsize from $\min(\eta_0, \frac{nc}{1 + t})$, where
+\textbf{Linear speedup of FedAvg and Nesterov accelerated FedAvg.} To verify the linear speedup convergence as shown in Theorem~\ref{thm:SGD_scvx}~\ref{thm:SGD_cvx}~\ref{thm:nesterov_scvx}~\ref{thm:Nesterov_cvx}, we evaluate the number of iterations needed to reach
+$\epsilon$-accuracy in three objectives.
+We initialize all runs with $\vw_0 = \textbf{0}_d$ and measure the number of iterations to reach the target accuracy $\epsilon$.
+For each configuration $(E, K)$, we extensively search the learning rate from $\min(\eta_0, \frac{nc}{1 + t})$, where
 $\eta_0 \in \{0.1, 0.12, 1, 32 \}$ according to different problems and $c$ can
-take the values $c = 2^i \ \forall i \in \ZZ$. As the results shown in Figure~\ref{fig:speedup} (1st row),
-the number of iterations to reach the target accuracy $\epsilon$ decreasing 
-as the number of workers increase.
+take the values $c = 2^i \ \forall i \in \ZZ$. As the results shown in Figure~\ref{fig:speedup},
+the number of iterations decreases as the number of (active) workers increasing, which is consistent for FedAvg and Nesterov accelerated FedAvg across all scenarios.
+For the additional experiments on the impact of $E$, detailed experimental setup, and hyperparameter setting, please refer to the Appendix, section~\ref{sec:expsupp}.
 
-\textbf{Partial participations.} We verify the linear speedup in the
-partial participation settings, where we set $50\%$ of devices
-are active. As the results are shown in Figure~\ref{fig:speedup} (2nd row), the 
-FedAve enjoys linear speedup in various settings even with partial
-device participation.
+% \textbf{Partial participations.} We verify the linear speedup in the
+% partial participation settings, where we set $50\%$ of devices
+% are active. As the results are shown in Figure~\ref{fig:speedup} (2nd row), the 
+% FedAve enjoys linear speedup in various settings even with partial
+% device participation.
 
-\textbf{Nesterov accelerated FedAvg.} In the third row of Figure~\ref{fig:speedup}, 
-we report the last iteration to converge to $\epsilon$-accuracy of Nesterov accelerated FedAvg.
+% \textbf{Nesterov accelerated FedAvg.} In the third row of Figure~\ref{fig:speedup}, 
+% we report the last iteration to converge to $\epsilon$-accuracy of Nesterov accelerated FedAvg.
 The empirical observations align with Theorem~\ref{thm:nesterov_scvx}~\ref{thm:Nesterov_cvx} that the accelerated version of FedAvg can also achieve the linear speedup
-w.r.t the number of workers.
+% w.r.t the number of workers.
 
-\textbf{The impact of $E$.} We further examine how does the number of local steps ($E$) 
-affect convergence. As shown in Figure~\ref{fig:e} in appendix section~\ref{sec:expsupp}, the number of iterations increases as $E$ increase, which slow down the convergence in
-terms of gradient computation. However, it can save communication costs as
-the number of rounds decreased when the $E$ increases. This showcase that
-we need a proper choice of $E$ to trade-off the communication cost and
-convergence speed.
+% \textbf{The impact of $E$.} We further examine how does the number of local steps ($E$) 
+% affect convergence. As shown in Figure~\ref{fig:e} in appendix section~\ref{sec:expsupp}, the number of iterations increases as $E$ increase, which slow down the convergence in
+% terms of gradient computation. However, it can save communication costs as
+% the number of rounds decreased when the $E$ increases. This showcase that
+% we need a proper choice of $E$ to trade-off the communication cost and
+% convergence speed.
 
@@
 \centering
+{\small
 \begin{tabular}{ccc}
@@
 (a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
-	\end{tabular}
-\caption{\textbf{Linear speedup} The first row reports the linear speedup convergence of FedAvg w.r.t the number of workers.
+	\end{tabular}}
+	\vspace{-1em}
+\caption{The linear speedup of FedAvg in full participation, partial participation, and the linear speedup of Nesterov accelerated FedAvg, respectively.
+% \textbf{Linear speedup } The first row reports the linear speedup convergence of FedAvg w.r.t the number of workers.
 The second row reports the linear speedup convergence of FedAvg w.r.t the number of active workers in the partial parcipation setting.
@@
 }
+\vspace{-1em}
 \label{fig:speedup}

### 2020-06-03 source/sec_exp_details.tex
exp and conclusions

@@
 We first consider the strongly convex objective function, where we use
-a regularized binary logistic regression with regularization $\lambda=1/n\approx 1e-5$.
+a regularized binary logistic regression with regularization $\lambda=1/n\approx 2e-5$.
 We evenly distributed on $1, 2, 4, 8, 16, 32$ devices and  report the number of iterations/rounds needed to converge to $\epsilon-$accuracy, where $\epsilon=0.005$.

### 2020-06-03 source/sec_problem.tex
Update sec_problem.tex
minor update problem setup

@@
 updates in FedAvg with full participation. For 
-more details on the notations and setup, please also refer to the Section~\ref{sec:app:notations} in the appendix.
+more details on the notations and setup, please refer to Section~\ref{sec:app:notations} in the appendix.
 
@@
 \subsection{Assumptions}
-
+\label{sec:assumptions}
 We make the following standard assumptions on the objective function $F_1,\dots, F_N$.

### 2020-06-03 source/sec_SGD.tex
sec3,4,5 typo
minor updates for SGD section

@@
 
-In this section, we provide convergence analyses of FedAvg with local
-SGD updates in the general setting with heterogeneous data and partial
+In this section, we provide convergence analyses of FedAvg for convex objectives in the general setting with both heterogeneous data and partial
 participation. We show that for strongly convex and smooth objectives,
@@
 objectives, the rate is $\mathcal{O}(1/\sqrt{NT})$. 
-The proof of FedAvg is deferred in Appendix Section~\ref{sec:app:fedavg}.
+Detailed proofs are deferred to Appendix Section~\ref{sec:app:fedavg}.
 
@@
 
-We first show that FedAvg with local SGD updates has $\mathcal{O}(1/NT)$
-convergence rate for $\mu$-strongly convex and $L$-smooth objectives.
-The result improves on the $\mathcal{O}(1/T)$ result of \cite{li2019convergence}
+We first show that FedAvg has $\mathcal{O}(1/NT)$ convergence rate
+for $\mu$-strongly convex and $L$-smooth objectives. The result
+improves on the $\mathcal{O}(1/T)$ rate of \cite{li2019convergence}
 with a linear speedup in the number of devices $N$. Moreover, it
@@
 without degrading the linear speedup in the number of workers. On
-the other hand, with partial participation, $E$ must be $\mathcal{O}(1)$. 
+the other hand, with partial participation, $E$ must be $\mathcal{O}(1)$
+to guarantee $\mathcal{O}(1/NT)$ convergence.
 \begin{theorem}
-	\label{thm:SGD_scvx}Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$
-	be the average of local parameters of FedAvg with learning rates $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$,
-	where $\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$. Then
-	with full device participation, 
+	\label{thm:SGD_scvx}Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$,
+	$\nu_{\max}=\max_{k}Np_{k}$, and set decaying learning rates $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$
+	with $\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$. Then
+	under Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2} with full device participation, 
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right)
+	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
 	\end{align*}
-	and with partial device participation with $K$ sampled devices at
-	each communication round, 
+	and with partial device participation with at most $K$ sampled devices
+	at each communication round, 
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}\leq\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right)
+	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}\leq\mathcal{O}(\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
 	\end{align*}
-	where $\nu_{\max}=\max_{k}Np_{k}$.
 \end{theorem}
@@
 \textbf{Linear speedup. }We first compare our bound with that in \cite{li2019convergence},
-which is $\mathcal{O}(\frac{1}{NT}\nu_{\max}^{2}\sigma^{2}\kappa/\mu+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{E^{2}G^{2}}{T}\kappa/\mu)$.
-Because the third term $\frac{E^{2}G^{2}}{T}\kappa/\mu$ is also $\mathcal{O}(1/T)$
+which is $\mathcal{O}(\frac{1}{NT}+\frac{E^{2}}{KT}+\frac{E^{2}G^{2}}{T})$.
+Because the term $\frac{E^{2}G^{2}}{T}$ is also $\mathcal{O}(1/T)$
 without a dependence on $N$, for any choice of $E$ their bound cannot
 achieve linear speedup. The improvement of our bound comes from the
-term $\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}$, which now is $\mathcal{O}(1/T^{2})$.
+term $\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}$, which now is $\mathcal{O}(E^{2}/T^{2})$.
 As a result, all leading terms scale with $1/N$ in the full device
 particiaption setting, and with $1/K$ in the partial participation
-setting. This implies that in both settings, there is a linear speedup
-in the number of active workers during the communication round. \\
+setting. This implies that in both settings, there is a \emph{linear
+	speedup} in the number of active workers during the communication
+round. \\
 % To our knowledge, this is the first result that explicitly demonstrates
@@
 in the choice of $E$ between the full and partial participation settings.
-With full participation there is linear speedup $\mathcal{O}(1/N)$
-as long as $E=\mathcal{O}(\sqrt{T/N})$. This corresponds to a communication
-complexity of $\mathcal{O}(\sqrt{NT})$. In contrast, the bound in
-\cite{li2019convergence} does not allow $E$ to scale with $\sqrt{T}$,
-even for full participation. On the other hand, with partial participation,
-the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$ is also a leading term,
-and so $E$ must be $\mathcal{O}(1)$. In this setting, our bound
-still yields a linear speedup in $K$, which is not the case for previous
-analyses. \\
-\textbf{Comparison with related works.} To better understand the significance of obtained bound, we compare 
-our rates to the best-known results in the same setting. 
-In~\cite{haddadpour2019convergence}, the authors prove a linear speedup $\mathcal{O}(1/NT)$ result for strongly
-convex and smooth objectives, with $\mathcal{O}(N^{1/3}T^{1/3})$ communication complexity when considering i.i.d. data. 
+With full participation there is linear speedup $\mathcal{O}(1/NT)$
+as long as $E=\mathcal{O}(\sqrt{T/N})$ since then $\mathcal{O}(E^{2}/T^{2})=\mathcal{O}(1/NT)$
+matches the leading term. This corresponds to a communication complexity
+of $T/E=\mathcal{O}(\sqrt{NT})$. In contrast, the bound in \cite{li2019convergence}
+does not allow $E$ to scale with $\sqrt{T}$ to preserve $\mathcal{O}(1/T)$
+rate, even for full participation. On the other hand, with partial
+participation, $\frac{\kappa E^{2}G^{2}/\mu}{KT}$ is also a leading
+term, and so $E$ must be $\mathcal{O}(1)$. In this case, our bound
+still yields a linear speedup in $K$, which is also confirmed by
+experiments. The $E=\mathcal{O}(1)$ requirement in partial participation
+likely cannot be removed for our sampling schemes, as the sampling
+variance is $\Omega(E^{2}/T^{2})$ and the dependence on $E$ is tight.  \\
+\textbf{Comparison with related works.} To better understand the significance of the obtained bound, we compare our rates to the best-known results in the same setting.
+In~\cite{haddadpour2019convergence}, the authors prove a linear speedup $\mathcal{O}(1/NT)$ result for strongly convex and smooth objectives, with $\mathcal{O}(N^{1/3}T^{2/3})$ communication complexity with i.i.d.
+data and partial participation.
 However, their results build upon the bounded gradient diversity assumption, which leads to a more restricted overparameterized setting (see discussions in Section~\ref{sec:assumptions}).
@@
 \begin{theorem}
-	\label{thm:SGD_cvx}With full device participation and constant learning
+	\label{thm:SGD_cvx}Under assumptions~\ref{ass:lsmooth},\ref{ass:boundedvariance},\ref{ass:subgrad2} and constant learning
 	rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$, 
@@
 	\end{align*}
-	and with partial device participation with $K$ sampled devices at
+	with full participation, and with partial device participation with $K$ sampled devices at
 	each communication round and learning rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{K}{T}})$,
@@
 first result on linear speedup in the general federated learning setting
-with both heterogeneous data and partial participation. \\
+with both heterogeneous data and partial participation for convex objectives.
+\begin{comment}
 \textbf{Learning rate. }The learning rate now depends on the final
@@
 problems with $\mathcal{O}(1/\sqrt{NT})$ linear speedup convergence results. 
+\end{comment}

### 2020-06-03 source/sec_Nesterov.tex
minor changes Nesterov Section

@@
 \subsection{Strongly Convex and Smooth Objectives}
-
+\begin{comment}
 We first show that the Nesterov accelerated FedAvg has $\mathcal{O}(1/NT)$
 convergence rate for $\mu$-strongly convex and $L$-smooth objectives.
+\end{comment}
 The Nesterov Accelerated FedAvg algorithm follows the updates
@@
 \end{align*}
-where $g_{t,k}:=\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})$ is
+where $\mathbf{g}_{t,k}:=\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})$ is
 the stochastic gradient sampled on the $k$-th device at time $t$.  
@@
 	\label{thm:nesterov_scvx}Let $\overline{\mathbf{v}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{v}_{T}^{k}$
-	be the average of local parameters of Nesterov accelerated FedAvg
-	with learning rates $\alpha_{t}=\frac{6}{\mu}\frac{1}{t+\gamma}$
-	and $\beta_{t-1}=\frac{3}{14(t+\gamma)(1-\frac{6}{t+\gamma})\max\{\mu,1\}}$,
-	$\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$. Then with
-	full device participation, 
+	and set learning rates $\alpha_{t}=\frac{6}{\mu}\frac{1}{t+\gamma}$,  $\beta_{t-1}=\frac{3}{14(t+\gamma)(1-\frac{6}{t+\gamma})\max\{\mu,1\}}$.
+Then under Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2} with full device participation,
 	\begin{align*}
@@
 \begin{theorem}
-	\label{thm:Nesterov_cvx}Set learning rates $\alpha_{t}=\beta_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$
-	for Accelerated FedAvg with convex and $L$-smooth objectives. With
-	full device participation, 
+	\label{thm:Nesterov_cvx}Set learning rates $\alpha_{t}=\beta_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$.
+Then under Assumptions~\ref{ass:lsmooth},\ref{ass:boundedvariance},\ref{ass:subgrad2} Nesterov accelerated FedAvg with
+	full device participation has rate
 	\begin{align*}

### 2020-06-03 source/sec_overparameterized.tex
minor updates overparameterized

@@
 and the existence of parameter choices with zero training loss is
-assured~\cite{allen2018convergence,zhang2016understanding}.  Due to the
+ensured~\cite{allen2018convergence,zhang2016understanding}.  Due to the
 property of \textit{automatic variance reduction} in overparameterization,  a
@@
 of FedAvg for the overparameterized strongly convex and smooth problems,
-and we show that it preserves the linear speedup at the same time.  
+and show that it preserves linear speedup at the same time.  
 We then sharpen this result in the special case of
@@
 we further propose a novel momentum-based FedAvg algorithm, which enjoys an
-improved convergence rate over FedAvg. The proof is deferred to Appendix Section~\ref{sec:interpolation}.
+improved convergence rate over FedAvg. Detailed proofs are deferred to Appendix Section~\ref{sec:interpolation}.
 \begin{comment}
@@
 In this section, we consider the standard Empirical Risk Minimization (ERM) setting where $\ell$
-is non-negative, $l$-smooth, and convex, and as before, we assume
-that each $F_{k}(\mathbf{w})$ is $L$-smooth and $\mu$-strongly convex. Note that $l\geq L$. This
+is non-negative, $l$-smooth, and convex, and as before, each $F_{k}(\mathbf{w})$ is $L$-smooth and $\mu$-strongly convex. Note that $l\geq L$. This
 setup includes many important problems in practice. In the overparameterized
@@
 such that $\ell(\mathbf{w}^{\ast};\mathbf{x}_{k}^{j})=0$ for all
-$\mathbf{x}_{k}^{j}$. We first show that FedAvg with local SGD updates
-and communication every $E$ iterations achieves geometric convergence
+$\mathbf{x}_{k}^{j}$. We first show that FedAvg achieves geometric convergence
 with linear speedup in the number of workers. 
@@
 	\label{thm:overparameterized_general}In the overparameterized setting,
-	FedAvg with local SGD updates and communication every $E$ iterations
-	with constant step size $\overline{\alpha}=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})})$
-	has geometric convergence,
+	FedAvg with communication every $E$ iterations
+	and constant step size $\overline{\alpha}=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})})$
+	has geometric convergence
 	\begin{align*}
@@
 %
-\textbf{Linear speedup and Communication Complexity} When $l>L$,
-the speedup factor is on the order of $\mathcal{O}(N/E)$ %\frac{N}{E(l\nu_{\max}+L(N-\nu_{\min}))}/\frac{1}{l\nu_{\max}+L(1-\nu_{\min})}\approx\frac{Nl}{E(l+L(N-1))}=
+\textbf{Linear speedup and Communication Complexity} The linear speedup factor is on the order of $\mathcal{O}(N/E)$ %\frac{N}{E(l\nu_{\max}+L(N-\nu_{\min}))}/\frac{1}{l\nu_{\max}+L(1-\nu_{\min})}\approx\frac{Nl}{E(l+L(N-1))}=
 for $N\leq\mathcal{O}(\frac{l}{L})$, i.e. FedAvg with $N$ workers and communication
@@
 result also illustrates that $E$ can be taken $\mathcal{O}(T^{\beta})$
-for any $\beta<1$, achieving better communication efficiency than
+for any $\beta<1$ to achieve geometric convergence, achieving better communication efficiency than
 the standard FL setting.
@@
 and there exists $\mathbf{w}^{\ast}$ such that $F(\mathbf{w}^{\ast})\equiv0$.
-Define the local Hessian matrix as $H^{k}:=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\mathbf{x}_{k}^{j}(\mathbf{x}_{k}^{j})^{T}$,
-the averaged Hessian matrix as $H:=\sum_{k=1}^{N}p_{k}H^{k}$, and the stochastic Hessian matrix as $\tilde{H}_{t}^{k}:=\xi_{t}^{k}(\xi_{t}^{k})^{T}$. 
+Define the local Hessian matrix as $H^{k}:=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\mathbf{x}_{k}^{j}(\mathbf{x}_{k}^{j})^{T}$, and the stochastic Hessian matrix as $\tilde{H}_{t}^{k}:=\xi_{t}^{k}(\xi_{t}^{k})^{T}$.
 \begin{comment}
@@
 	\label{thm:overparameterized_quadratic}For the overparamterized quadratic
-	problem, FedAvg with local SGD updates and communication every $E$
+	problem, FedAvg with communication every $E$
 	iterations with constant step size $\overline{\alpha}=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})})$

### 2020-06-03 source/sec_conclude.tex
minor fixes conclude.tex

@@
 \end{comment}
-This paper provides a comprehensive and unified analysis of the convergence rate of FedAvg
+This paper provides a comprehensive and unified analysis of the convergence of FedAvg
 and its accelerated variants in a general federated learning problem with heterogeneous local data and partial participation. 
@@
 Furthermore, this work also makes algorithmic contributions. We not only prove that FedAvg can achieve geometric convergence for overparameterized strongly convex smooth problems 
-but also propose the MaSS accelerated Fedavg algorithm, which has provable speedup in convergence rate over FedAvg on quadratic problems. Last but not least, extensive empirical 
-evaluations demonstrate the linear speedup of FedAvg and Nesterov accelerated FedAvg under various settings.
+but also propose the MaSS accelerated FedAvg algorithm, which has provable speedup in convergence rate over FedAvg on quadratic problems. Last but not least, extensive empirical 
+evaluations demonstrate the linear speedup of FedAvg and accelerated FedAvg under various settings.

### 2020-06-03 source/sec_intro.tex
intro updates

--- 2020-06-01 source/sec_intro.tex
@@
 Its success has since motivated a growing line of research efforts into understanding its theoretical convergence guarantees in various settings.
+Most existing convergence guarantees of FedAvg were either established in simplified settings~\cite{stich2018local,khaled2019first,wang2018cooperative,yu2019parallel,khaled2020tighter,li2018federated}, have restrictive assumptions~\cite{haddadpour2019convergence} or provide a suboptimal
+convergence rate~\cite{li2019convergence}. Furthermore, the convergence of
+FedAvg's accelerated variants~\cite{yu2019linear,huo2020faster,liu2019accelerating} in the general heterogeneous and convex setting is yet to be analyzed.
+At the heart of the success of FL is the \emph{linear speedup} property with respect to the number of participating clients that can be achieved with infrequent communications.
+Although results of this kind are abundant in distributed optimization~\cite{yu2019linear,yu2019parallel,stich2018local,khaled2020tighter,stich2018local}, generalization to the FL setting with both types of heterogeneity is currently lacking.
 
- Existing convergence guarantees of FedAvg were either largely established on simplified settings~\cite{stich2018local,khaled2019first,wang2018cooperative,yu2019parallel} or provides a suboptimal 
-convergence rate~\cite{li2019convergence}. Furthermore, the convergence of
-FedAvg's accelerated variants~\cite{yu2019linear,huo2020faster,liu2019accelerating} under the realistic setting is yet to be analyzed.
+\subsection{Related Works}
+\textbf{The Convergence of FedAvg and Linear Speedup.}
+There have been a considerable amount of convergence results in distributed stochastic optimization with infrequent communications that establish linear speedup in the number of clients.
+However, most of these prior works provide
+convergence guarantees under simplified settings, where either statistical
+heterogeneity~\cite{stich2018local,zhou2017convergence,wang2018cooperative,woodworth2018graph} or system
+heterogeneity~\cite{yu2019parallel,wang2019adaptive,khaled2019first,jiang2018linear} are not
+considered, and the associated algorithm is often referred to as local SGD.
+~\cite{li2019convergence} are the first to provide convergence of FedAvg for strongly convex problems considering both statistical heterogeneity and system heterogeneity, yet it
+didn't achieve linear speedup \cite{yu2019parallel,haddadpour2019convergence,stich2018local}.
+In~\cite{haddadpour2019convergence,liang2019variance,huo2020faster,jiang2018linear}, the convergence of FedAvg for non-convex problems has been studied, whereas this paper focuses on convex settings.
+\begin{comment}
+However, \cite{huo2020faster} did not discuss the linear speedup and the theorems in \cite{haddadpour2019convergence,liang2019variance} failed to cover the accelerated version of
+FedAvg. 
+\end{comment}
+
+\textbf{Accelerating Stochastic Federated Learning Algorithms.}
+% The accelerated SGD methods has achieved optimal convergence rate in strongly convex
+% optimization~\cite{nesterov1983method}.
+The convergence of accelerated FL algorithms using momentum (Nesterov and Heavy Ball) updates has been studied mostly in non-convex settings~\cite{yu2019linear,huo2020faster}.
+\cite{yu2019linear} does not consider system heterogeneity, while the result in \cite{huo2020faster} does not have linear speedup.
+Furthermore, the rates cannot improve over SGD-based algorithms in general.
+\cite{liu2019accelerating} provides an acceleration result for FedAvg with accelerated GD.
+\cite{liang2019variance} considers variance reduction based acceleration of FedAvg.
+This paper considers accelerated FedAvg with momentum-based stochastic updates in general and overparameterized convex settings.
+
+\begin{comment}
+	In this work, we provide a novel accelerated FedAvg with improved convergence rates under a popular overparameterized setting. 
+\end{comment}
 
@@
 \textbf{Contributions.}
-This work focuses on providing comprehensive convergence analysis
-of FedAvg and its accelerated variants under a realistic setting.
-Our results show
-that the less number of communication rounds is needed to converge to a
-target accuracy while preserving the linear speedup w.r.t. the number of participating devices.
+We provide a comprehensive convergence analysis
+of FedAvg and its accelerated variants in convex stochastic problems with both system and statistical heterogeneity.
+Our results demonstrate the desirable linear speedup property with respect to the number of participating clients.
 More specifically, for strongly convex and smooth problems, we show 
 that FedAvg admits a convergence rate of 
-{\small{$\cO(1/NT)$}}, and local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}.
-For convex and smooth objectives, we show the rate of {\small{$\cO(1/\sqrt{NT})$}} and local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}.
+{\small{$\cO(1/NT)$}}, and local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}. For convex and smooth objectives, the rate is {\small{$\cO(1/\sqrt{NT})$}}. 
 Moreover, we show that FedAvg with Nesterov acceleration also enjoys the same rates and speedup. 
 %The same conclusions are equally applied for Nesterov accelerated FedAvg. 
-To the best of our knowledge, this is the first linear speedup results for FedAvg with Nesterov accelerated SGD. Furthermore, 
-we provide the first improved convergence rate for Nesterov accelerated FedAvg comparing to FedAvg under the popular overparamterized setting. 
+To the best of our knowledge, this is the first linear speedup result for FedAvg and accelerated variants in the full heterogeneous FL setting. Furthermore, 
+we propose a new variant of FedAvg and provide the first improved convergence rate for stochastic momentum-accelerated FedAvg in the popular overparamterized setting.
 Last but not least, we systematically examine the convergence rate of
@@
 that are accessible to the central server.
-Let $T$ be the total number of SGD iterations performed by each local device, $E$ be the local steps between two consecutive server communications.
+Let $T$ be the total number of stochastic updates performed by each local device, $E$ be the local steps between two consecutive server communications.
 Thus, $T/E$ is

### 2020-06-03 source/app_sumarization.tex
add assumption discussion

@@
 
-% \section{Comparison of convergence rates}
+\section{Comparison of convergence rates with related works.}
+\label{sec:app:comparison}
+In this section, we summarize the convergence results in this work from various perspectives
+and compare our results with related works. In Table~\ref{tb:convergenceratev1} and Table~\ref{tb:convergenceratev2}, we summarize 
+the convergence results in terms of participations and accelerations. In Table~\ref{tb:convergenceratev3}, we compare our 
+results with the best known results in the literatures.
+
+In~\cite{haddadpour2019convergence}, they also provide $\cO(1/NT)$ convergence rate of non-convex problems under Polyak-Łojasiewicz (PL) condition, which
+means their results can directly apply to the strongly convex problems. However, their assumption is based on bounded gradient diversity, defined as follows: 
+\begin{align*}
+	\Lambda(\vw) = \frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}} \leq B
+\end{align*} 
+This is a more restrictive comparing to assuming bounded gradient under the case of target accuracy $\epsilon \rightarrow 0$ and PL condition.
+To see this, consider the gradient diversity at the global optimal $\vw^*$,i.e., $\Lambda(\vw^*) = \frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$.
+For $\Lambda(\vw^*)$ to be bounded, it requires $\|\nabla F_{k}(\mathbf{w}^*)\|_{2}^{2} = 0$, $\forall \ k$.
+This indicates
+$\vw^*$ is also the minimizer of each local objective, which contradicts to the practical setting of heterogeneous data. Therefore, their bound 
+is not effective for arbitrary small $\epsilon$-accuracy under general heterogeneous data while our convergence results still hold in this case.
+% If we only consider the domain $\vw \in \{\vw | \|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2} > c\}$, where $c$ is a positive constant. 
+% The bounded gradient can imply gradient diversity (see Corollary 10 in ~\cite{li2018federated}). In this case, the gradient diversity is a relaxed
 
@@
 \end{tabular}
-\caption{A high level summary of the convergence results for full participation in this paper. This table only hightlights the
+\caption{A high level summary of the convergence results for full participation in this paper. This table only highlights the
 dependence on $T$ (number of iterations), $E$ (the largest number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices.}
@@
 \end{tabular}
-\caption{A high level summary of the convergence results in this paper. This table only hightlights the
+\caption{A high level summary of the convergence results in this paper. This table only highlights the
 dependence on $T$ (number of iterations), $E$ (the largest number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices.
@@
 \centering
-\hspace{-2em}
-{\small
-\begin{tabular}{|c|c|c|c|c|c|c|}
-\hline Reference                 & Convergence rate    & E                           			& Participation & Extra Assumptions  		  & Setting  \\ \hline\hline 
-\cite{li2019convergence}         & $\cO(\frac{E^2}{T})$& $\cO(1)$                    			& Partial       & Bounded gradient   		  & Strongly convex  \\ \hline
-\cite{haddadpour2019convergence} & $\cO(\frac{1}{KT})$ & $\cO(K^{1/3}T^{2/3})^{\dagger}$     			& Partial       & Bounded gradient diversity   & Strongly convex$^{\mathsection}$  \\ \hline
-This work                        & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{\dagger\dagger}$ 			& Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
-\cite{zanette2019tighter}  & $\cO(\frac{1}{\sqrt{NT}})$ & $\cO(N^{-3/2}T^{1/2})$     			& Full        & Bounded gradient             & Convex  \\\hline
-This work      & $\cO\left(\frac{1}{\sqrt{NT}}\right)$ &  $\cO(N^{-3/4}T^{1/4})^{\dagger\dagger}$			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
-This work & $\cO\left(\exp(-\frac{NT}{E\kappa_1})\right)$ & $ \cO(T^{\beta})$                   &  Partial     & Bounded gradient    & Overparameterized (LR)$^{\ddagger}$ \\ \hline
-This work & $\cO\left(\exp(-\frac{NT}{E\sqrt{\kappa_1\tilde{\kappa}}})\right)$ & $ \cO(T^{\beta})$  &  Partial     & Bounded gradient    & Overparameterized LR \\ \hline
+{\tiny
+\begin{tabular}{|c|c|c|c|c|c|c|c|}
+\hline Reference                 & Convergence rate    & E                           			& NonIID & Participation & Extra Assumptions  		  & Setting  \\ \hline\hline 
+\cite{li2019convergence}         & $\cO(\frac{E^2}{T})$& $\cO(1)$                    			& \cmark & Partial       & Bounded gradient   		  & Strongly convex  \\ \hline
+\cite{haddadpour2019convergence} & $\cO(\frac{1}{KT})$ & $\cO(K^{1/3}T^{2/3})^{\dagger}$     	& \cmark$^{\ddagger\ddagger}$ & Partial       & Bounded gradient diversity   & Strongly convex$^{\mathsection}$  \\ \hline
+This work                        & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{\dagger\dagger}$& \cmark			& Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
+\cite{zanette2019tighter}  & $\cO(\frac{1}{\sqrt{NT}})$ & $\cO(N^{-3/2}T^{1/2})$     			 & \cmark& Full        & Bounded gradient             & Convex  \\\hline
+This work      & $\cO\left(\frac{1}{\sqrt{NT}}\right)$ &  $\cO(N^{-3/4}T^{1/4})^{\dagger\dagger}$& \cmark			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
+This work & $\cO\left(\exp(-\frac{NT}{E\kappa_1})\right)$ & $ \cO(T^{\beta})$                   & \cmark&  Partial     & Bounded gradient    & Overparameterized (LR)$^{\ddagger}$ \\ \hline
+This work & $\cO\left(\exp(-\frac{NT}{E\sqrt{\kappa_1\tilde{\kappa}}})\right)$ & $ \cO(T^{\beta})$ & \cmark &  Partial     & Bounded gradient    & Overparameterized LR \\ \hline
 \end{tabular}
 }
-\caption{A high level summary of the convergence results in this paper and their comparison to prior state-of-the-art FL algorithms, considering heterogeneous data. This table only hightlights the
+\caption{A high level summary of the convergence results in this paper and their comparison to prior state-of-the-art FL algorithms, considering heterogeneous data. This table only highlights the
 dependence on $T$ (number of iterations), $E$ (the largest number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices. 
 $\kappa$ is the condition number of the system and $\beta \in (0,1)$. 
-We note that all our results for strongly convex and convex smooth problems are unified results for both FedAvg and accelerated FedAvg. The converegence results for prior arts only apply for FedAvg.}
+We note that all our results for strongly convex and convex smooth problems are unified results for both FedAvg and accelerated FedAvg. The convergence results for prior arts only apply for FedAvg.}
 {\raggedright 
@@
          $^{\dagger\dagger}$ This $E$ is obtained under full participation setting. \\ 
-         $^{\mathsection}$ In~\cite{haddadpour2019convergence}, the convergence rate is for non-convex smooth problem with PL condition.
-Thus, it is also applied for strongly convex problem.
-For unified comparison, we denote it as strongly convex setting here.
+         $^{\ddagger}$ The convergence of FedAvg for general overparameterized strongly convex and overparameterized linear regression share the same result.\\
+         $^{\mathsection}$ In~\cite{haddadpour2019convergence}, the convergence rate is for non-convex smooth problem with PL condition, which also applies for strongly convex problems.
+Therefore, we compare it with our strongly convex results here.
 \\
-         $^{\ddagger}$ The convergence of FedAvg for general overparameterized strongly convex and overparameterized linear regression share the same result.  \par}
+         $^{\ddagger\ddagger}$ The bounded gradient diversity assumption is not applicable for general heterogeneous data when converging to arbitrarily small $\epsilon$-accuracy (see discussions in Sec~\ref{sec:app:comparison}).
+           \par}
 \label{tb:convergenceratev3}

### 2020-06-03 source/sec_problem.tex
update problem.tex

@@
 where $N$ is the number of local devices (users/nodes/workers) and $p_k$ is the $k$-th device's weight satisfying $p_k \geq 0$ and $\sum_{k=1}^N p_k = 1$. 
-In $k$-th local device, there are a finite number of $n_k$ data points:
+In $k$-th local device, there are $n_k$ data points:
 $\vx_k^1, \vx_k^2, \dots, \vx_k^{n_k}$.  
@@
 % \end{align}
-where $\ell$ denotes a user-specified loss function.
-In federated learning, each device stores its local data, which gives rise to its own local objective $F_k$.
-Note that we do not make any assumption on the
+where $\ell$ denotes a user-specified loss function. Each device only has access to its local data, which gives rise to its own local objective $F_k$. Note that we do not make any assumptions on the
 data distributions of each local device. The local minimum
-$\vw^{*}_k= \argmin_{\vw \in R^d} F_k(\vw)$ can be drifted far from
-the global minimum of \eq{\ref{eq:problem}}, which quantifies the 
-heterogeneity of the problem.
-In this paper, we systematically investigate the convergence results, 
-when the local objectives are convex smooth, strongly convex and convex but
-non-smooth.
+$\vw^{*}_k= \argmin_{\vw \in R^d} F_k(\vw)$ can be far from
+the global minimum of \eq{\ref{eq:problem}}.
 
@@
 \subsection{The Federated Averaging (FedAvg) Algorithm}
-We first introduce the standard Federated Average (FedAvg) algorithm.
-The FedAvg updates the model in each device by local Stochastic Gradient Descent (SGD) and send the latest model to the central server every $E$
-steps. The central server conducts a weighted average over the models
+We first introduce the standard Federated Average (FedAvg) algorithm~\cite{mcmahan2016communication}.
+FedAvg updates the model in each device by local Stochastic Gradient Descent (SGD) and send the latest model to the central server every $E$
+steps. The central server conducts a weighted average over the model parameters
 received from active devices and broadcasts the latest averaged model to all devices.
@@
 where $\vw_t^k$ is the local model parameter maintained in the $k$-th device at the $t$-th iteration, $\vg_{t,k}:=\nabla F_{k}(\vw_{t}^{k},\xi_{t}^{k})$ is the stochastic gradient based on $\xi_{t}^{k}$, the data sampled from $k$-th device’s local data uniformly at random.
-$\mathcal{I}_{E}=\{E,2E,\dots\}$ is the set of global synchronization steps.
+$\mathcal{I}_{E}=\{E,2E,\dots\}$ is the set of global communication steps.
 We use $\mathcal{S}_{t+1}$ to represent the set of active devices at $t+1$.
@@
 Since federated learning usually involves an enormous amount of 
-local devices, it is more realistic to consider only a subset of 
+local devices, it is often more realistic to assume only a subset of 
 local devices is active at each communication round (system heterogeneity). In this work,
@@
 the case of \textbf{partial participation} where $|\cS_{t+1}| < N$. 
-In the partial participation scenario, $\cS_{t+1}$ is obtained by two types
+With partial participation, $\cS_{t+1}$ is obtained by two types of
 sampling schemes to simulate practical scenarios~\cite{li2019convergence}.
-The sampling scheme I establishes $\cS_{t+1}$ by i.i.d. sampling the devices according to
-the weight $p_k$ with replacement. 
-The sampling scheme II establishes $\cS_{t+1}$ by uniformly sampling all devices without
-replacement. Both of the sampling schemes
+For example, Sampling scheme I establishes $\cS_{t+1}$ by i.i.d.
+sampling the devices with probability $p_k$ with replacement.
+Both schemes
 guarantee that gradient updates in FedAvg are unbiased stochastic versions of
@@
 We make the following standard assumptions on the objective function $F_1,\dots, F_N$.
-Assumption~\ref{ass:lsmooth} and Assumption~\ref{ass:stroncvx} are common assumptions satisfied by a range of popular objective functions, such as $\ell^{2}$ regularized logistic regression and cross-entropy loss functions.
+Assumptions~\ref{ass:lsmooth} and~\ref{ass:stroncvx} are commonly satisfied by a range of popular objective functions, such as $\ell^{2}$ regularized logistic regression and cross-entropy loss functions.
 
@@
 \textbf{Justfication for Assumptions~\ref{ass:boundedvariance} and~\ref{ass:subgrad2}.} These standard assumptions have been made in many previous
-works in federated learning\cite{yu2019parallel,li2019convergence,stich2018local}. Here we provide one more justification
+works in federated learning\cite{yu2019parallel,li2019convergence,stich2018local}. We provide one more justification
 for their generality: even though they require bounds for all $\mathbf{w}$,

### 2020-06-03 source/sec_problem.tex
updated justification

@@
 \end{assumption}
-
-\textbf{Justfication for Assumptions~\ref{ass:boundedvariance} and~\ref{ass:subgrad2}.} These standard assumptions have been made in many previous
-works in federated learning\cite{yu2019parallel,li2019convergence,stich2018local}. We provide one more justification
-for their generality: even though they require bounds for all $\mathbf{w}$,
+Assumptions~\ref{ass:boundedvariance} and~\ref{ass:subgrad2} have been made in many previous
+works in federated learning\cite{yu2019parallel,li2019convergence,stich2018local}. We provide one more justification for their generality: even though they require bounds for all $\mathbf{w}$,
 they are not much stronger than requiring bounds at the global optimal
@@
 Therefore, there is no substantial difference between these assumptions
-and assuming the bounds at $\mathbf{w}^{\ast}$ only. On the other
-hand, some works such as \cite{haddadpour2019convergence,li2018federated}
-assume bounds on the ``gradient diversity'' or ``dissimilarity'',
-defined as $\frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$.
-Such assumptions are more restrictive than our assumptions, because
-they effectively require $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$ for all local
-objectives. For such ``overparameterized'' settings, we will prove
-that the convergence is in fact geometric.
+and assuming the bounds at $\mathbf{w}^{\ast}$ only. Furthermore, instead of  
+assuming \textit{bounded gradient diversity} as in related works~\cite{haddadpour2019convergence,li2018federated}, 
+we argue that Assumption~\ref{ass:subgrad2} is more general. Because when converging to target accuracy $\epsilon$ and $\epsilon \rightarrow 0$,
+bounded gradient diversity restricts local objectives to the same minimizer as global objective, which contradicts to the heterogeneous data setting, 
+while our assumption is still feasible.
+For more detailed explanations of our assumptions, please refer to Appendix Section~\ref{sec:app:comparison}.
+
+% On the other hand, some works such as \cite{haddadpour2019convergence,li2018federated}
+% assume bounds on the ``gradient diversity'' or ``dissimilarity'',
+% defined as $\frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$.
+% Such assumptions are more restrictive than our assumptions, because
+% they effectively require $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$ for all local
+% objectives. For such ``overparameterized'' settings, we will prove
+% that the convergence is in fact geometric.

### 2020-06-03 source/sec_problem.tex
Create sec_problem.tex

--- 2020-06-03 source/sec_problem.tex
@@
 \end{assumption}
-
-\textbf{Justfication for Assumptions~\ref{ass:boundedvariance} and~\ref{ass:subgrad2}.} These standard assumptions have been made in many previous
-works in federated learning\cite{yu2019parallel,li2019convergence,stich2018local}. Here we provide one more justification
+These standard assumptions have been made in many previous works in federated learning\cite{yu2019parallel,li2019convergence,stich2018local}. 
+Here we provide one more justification
 for their generality: even though they require bounds for all $\mathbf{w}$,

### 2020-06-03 source/sec_intro.tex
update

@@
 To address these challenges, Federated Averaging (FedAvg)~\cite{mcmahan2016communication} was proposed as a particularly effective heuristic, which has enjoyed great empirical success~\cite{47586}.
-Its success has since motivated a growing line of research efforts into understanding its theoretical convergence guarantees in various settings.
-Most existing convergence guarantees of FedAvg were either established in simplified settings~\cite{stich2018local,khaled2019first,wang2018cooperative,yu2019parallel,khaled2020tighter,li2018federated}, have restrictive assumptions~\cite{haddadpour2019convergence} or provide a suboptimal
+This success has since motivated a growing line of research efforts into understanding its theoretical convergence guarantees in various settings.
+Most existing convergence guarantees of FedAvg were either established in simplified settings~\cite{stich2018local,khaled2019first,wang2018cooperative,yu2019parallel,khaled2020tighter,li2018federated,koloskova2020unified}, have restrictive assumptions~\cite{haddadpour2019convergence} or provide a suboptimal
 convergence rate~\cite{li2019convergence}. Furthermore, the convergence of
@@
 At the heart of the success of FL is the \emph{linear speedup} property with respect to the number of participating clients that can be achieved with infrequent communications.
-Although results of this kind are abundant in distributed optimization~\cite{yu2019linear,yu2019parallel,stich2018local,khaled2020tighter,stich2018local}, generalization to the FL setting with both types of heterogeneity is currently lacking.
+Although results of this kind are abundant in distributed and decentralized optimization~\cite{yu2019linear,yu2019parallel,stich2018local,khaled2020tighter,stich2018local,koloskova2020unified}, generalization to the FL setting with both types of heterogeneity is currently lacking.
 
@@
 convergence guarantees under simplified settings, where either statistical
-heterogeneity~\cite{stich2018local,zhou2017convergence,wang2018cooperative,woodworth2018graph} or system
+heterogeneity~\cite{stich2018local,zhou2017convergence,wang2018cooperative,woodworth2018graph,haddadpour2019convergence} or system
 heterogeneity~\cite{yu2019parallel,wang2019adaptive,khaled2019first,jiang2018linear} are not

### 2020-06-03 source/sec_problem.tex

--- 2020-06-03 source/sec_problem.tex
@@
 Assumptions~\ref{ass:boundedvariance} and~\ref{ass:subgrad2} have been made in many previous
-works in federated learning\cite{yu2019parallel,li2019convergence,stich2018local}. We provide one more justification for their generality: even though they require bounds for all $\mathbf{w}$,
-they are not much stronger than requiring bounds at the global optimal
-parameter $\mathbf{w}^{\ast}$. As model average parameters become
+works in federated learning, e.g.~\cite{yu2019parallel,li2019convergence,stich2018local}. We provide some more justification for their generality. As model average parameters become
 closer to $\mathbf{w}^{\ast}$, the $L$-smoothness property implies
@@
 and $\mathbb{E}\|\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})-\nabla F_{k}(\mathbf{w}_{t}^{k})\|^{2}$
-will become close to $\mathbb{E}\|\nabla F_{k}(\mathbf{w}^{\ast},\xi_{t}^{k})\|^{2}$
+approach $\mathbb{E}\|\nabla F_{k}(\mathbf{w}^{\ast},\xi_{t}^{k})\|^{2}$
 and $\mathbb{E}\|\nabla F_{k}(\mathbf{w}^{\ast},\xi_{t}^{k})-\nabla F_{k}(\mathbf{w}^{\ast})\|^{2}$.
 Therefore, there is no substantial difference between these assumptions
-and assuming the bounds at $\mathbf{w}^{\ast}$ only. Furthermore, instead of  
-assuming \textit{bounded gradient diversity} as in related works~\cite{haddadpour2019convergence,li2018federated}, 
-we argue that Assumption~\ref{ass:subgrad2} is more general. Because when converging to target accuracy $\epsilon$ and $\epsilon \rightarrow 0$,
-bounded gradient diversity restricts local objectives to the same minimizer as global objective, which contradicts to the heterogeneous data setting, 
-while our assumption is still feasible.
-For more detailed explanations of our assumptions, please refer to Appendix Section~\ref{sec:app:comparison}.
+and assuming the bounds at $\mathbf{w}^{\ast}$ only.
+Furthermore, compared to assuming \textit{bounded gradient diversity} as in related works~\cite{haddadpour2019convergence,li2018federated}, Assumption~\ref{ass:subgrad2} is much less restrictive.
+When target accuracy $\epsilon \rightarrow 0$,
+bounded gradient diversity restricts local objectives to have the same minimizer as the global objective, contradicting the heterogeneous data setting. 
+For detailed explanations of our assumptions, please refer to Appendix Section~\ref{sec:app:comparison}.
 

### 2020-06-03 source/app_sumarization.tex

@@
 \hline Reference                 & Convergence rate    & E                           			& NonIID & Participation & Extra Assumptions  		  & Setting  \\ \hline\hline 
-\cite{li2019convergence}         & $\cO(\frac{E^2}{T})$& $\cO(1)$                    			& \cmark & Partial       & Bounded gradient   		  & Strongly convex  \\ \hline
-\cite{haddadpour2019convergence} & $\cO(\frac{1}{KT})$ & $\cO(K^{1/3}T^{2/3})^{\dagger}$     	& \cmark$^{\ddagger\ddagger}$ & Partial       & Bounded gradient diversity   & Strongly convex$^{\mathsection}$  \\ \hline
-This work                        & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{\dagger\dagger}$& \cmark			& Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
-\cite{zanette2019tighter}  & $\cO(\frac{1}{\sqrt{NT}})$ & $\cO(N^{-3/2}T^{1/2})$     			 & \cmark& Full        & Bounded gradient             & Convex  \\\hline
-This work      & $\cO\left(\frac{1}{\sqrt{NT}}\right)$ &  $\cO(N^{-3/4}T^{1/4})^{\dagger\dagger}$& \cmark			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
-This work & $\cO\left(\exp(-\frac{NT}{E\kappa_1})\right)$ & $ \cO(T^{\beta})$                   & \cmark&  Partial     & Bounded gradient    & Overparameterized (LR)$^{\ddagger}$ \\ \hline
-This work & $\cO\left(\exp(-\frac{NT}{E\sqrt{\kappa_1\tilde{\kappa}}})\right)$ & $ \cO(T^{\beta})$ & \cmark &  Partial     & Bounded gradient    & Overparameterized LR \\ \hline
+FedAvg\cite{li2019convergence}         & $\cO(\frac{E^2}{T})$& $\cO(1)$                    		& \cmark & Partial       & Bounded gradient   		  & Strongly convex  \\ \hline
+FedAvg\cite{haddadpour2019convergence} & $\cO(\frac{1}{KT})$ & $\cO(K^{-1/3}T^{2/3})^{\dagger}$ & \cmark$^{\ddagger\ddagger}$ & Partial       & Bounded gradient diversity   & Strongly convex$^{\mathsection}$  \\ \hline
+FedAvg\cite{koloskova2020unified} & $\cO(\frac{1}{NT})$ & $\cO(N^{-1/2}T^{1/2})$     	& \cmark & Full       & Bounded gradient   & Strongly convex  \\ \hline
+FedAvg/N-FedAvg                 & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{\ddagger}$ & \cmark	 & Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
+FedAvg\cite{khaled2020tighter}  & $\cO(\frac{1}{\sqrt{NT}})$ & $\cO(N^{-3/2}T^{1/2})$     	    & \cmark& Full        & Bounded gradient             & Convex  \\\hline
+FedAvg\cite{koloskova2020unified} & $\cO(\frac{1}{\sqrt{NT}})$ & $\cO(N^{-3/4}T^{1/4})$    & \cmark & Full       & Bounded gradient             &  Convex  \\ \hline
+FedAvg/N-FedAvg      & $\cO\left(\frac{1}{\sqrt{KT}}\right)$ & $\cO(N^{-3/4}T^{1/4})^{\ddagger}$& \cmark			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
+FedAvg & $\cO\left(\exp(-\frac{NT}{E\kappa_1})\right)$ & $ \cO(T^{\beta})$                   & \cmark&  Partial     & Bounded gradient    & Overparameterized LR\\ \hline
+FedMass & $\cO\left(\exp(-\frac{NT}{E\sqrt{\kappa_1\tilde{\kappa}}})\right)$ & $ \cO(T^{\beta})$ & \cmark &  Partial     & Bounded gradient    & Overparameterized LR \\ \hline
 \end{tabular}
@@
 dependence on $T$ (number of iterations), $E$ (the largest number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices. 
-$\kappa$ is the condition number of the system and $\beta \in (0,1)$. 
-We note that all our results for strongly convex and convex smooth problems are unified results for both FedAvg and accelerated FedAvg. The convergence results for prior arts only apply for FedAvg.}
+$\kappa$ is the condition number of the system and $\beta \in (0,1)$. We denote Nesterov accelerated FedAvg as N-FedAvg in this table.}
 {\raggedright 
          $^{\dagger}$ This $E$ is obtained under i.i.d. setting. \\
-         $^{\dagger\dagger}$ This $E$ is obtained under full participation setting. \\ 
-         $^{\ddagger}$ The convergence of FedAvg for general overparameterized strongly convex and overparameterized linear regression share the same result.\\
+         $^{\ddagger}$ This $E$ is obtained under full participation setting. \\ 
+         % $^{\ddagger}$ The convergence of FedAvg for general overparameterized strongly convex and overparameterized linear regression share the same order in terms of $E, T, N$.\\
          $^{\mathsection}$ In~\cite{haddadpour2019convergence}, the convergence rate is for non-convex smooth problem with PL condition, which also applies for strongly convex problems.

### 2020-06-03 source/sec_intro.tex

@@
 At the heart of the success of FL is the \emph{linear speedup} property with respect to the number of participating clients that can be achieved with infrequent communications.
-Although results of this kind are abundant in distributed and decentralized optimization~\cite{yu2019linear,yu2019parallel,stich2018local,khaled2020tighter,stich2018local,koloskova2020unified}, generalization to the FL setting with both types of heterogeneity is currently lacking.
+Although results of this kind are abundant in distributed and decentralized optimization~\cite{yu2019linear,yu2019parallel,stich2018local,khaled2020tighter,koloskova2020unified}, generalization to the FL setting with both types of heterogeneity is currently lacking.
 
@@
 heterogeneity~\cite{stich2018local,zhou2017convergence,wang2018cooperative,woodworth2018graph,haddadpour2019convergence} or system
-heterogeneity~\cite{yu2019parallel,wang2019adaptive,khaled2019first,jiang2018linear} are not
+heterogeneity~\cite{yu2019parallel,wang2019adaptive,khaled2019first,jiang2018linear,koloskova2020unified} are not
 considered, and the associated algorithm is often referred to as local SGD.

### 2020-06-03 source/sec_overparameterized.tex
updated overparameterized

@@
 
-\subsection{Overparameterized Quadratic Problems}
-We now turn to quadratic problems
-and demonstrate how the general bound in Theorem~\ref{thm:overparameterized_general}
-can be sharpened to explicitly depend on the condition number of the
-system, and how this rate can be further improved with a new FedAvg
-algorithm with momentum updates.
+\subsection{Acceleration of FedAvg with Momentum-based Updates}
 
+We now turn to quadratic problems and show that the bound in Theorem \ref{thm:overparameterized_general} can be improved to $\mathcal{O}(\exp(-\frac{N}{E\kappa_{1}}t))$.
+We then propose a variant of FedAvg that has provable acceleration over FedAvg with SGD updates.
 The local device objectives are now given by the sum of squares {\small$F_{k}(\mathbf{w})=\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(\mathbf{w}^{T}\mathbf{x}_{k}^{j}-z_{k}^{j})^{2}$},
 and there exists $\mathbf{w}^{\ast}$ such that $F(\mathbf{w}^{\ast})\equiv0$.
+Two notions of condition number are important in our results:$\kappa_1$ which is based on local Hessians, and $\tilde{\kappa}$, which is termed the statistical condition number~\cite{liu2018accelerating,jain2017accelerating}.
+For their detailed definitions, please refer to Appendix Section~\ref{sec:interpolation}.
+Here we use the fact $\tilde{\kappa} \leq \kappa_1$.
+
+\begin{comment}
 Define the local Hessian matrix as $H^{k}:=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\mathbf{x}_{k}^{j}(\mathbf{x}_{k}^{j})^{T}$, and the stochastic Hessian matrix as $\tilde{H}_{t}^{k}:=\xi_{t}^{k}(\xi_{t}^{k})^{T}$.
+Define $l$ to be the smallest positive number such that $\mathbb{E}\|\xi_{t}^{k}\|^{2}$$\mathbf{\xi}_{t}^{k}$($\mathbf{\xi}_{t}^{k})^{T}\preceq lH^{k}$
+for all $k$. Note that $l\leq\max_{k,j}\|\mathbf{x}_{k}^{j}\|^{2}$.
+Let $L$ and $\mu$ be lower and upper bounds of non-zero eigenvalues
+of $H^{k}$. Define $\kappa_{1}:=l/\mu$ and $\kappa:=L/\mu$. Following
+\cite{liu2018accelerating,jain2017accelerating}, we define the statistical
+condition number $\tilde{\kappa}$ as the smallest positive real number
+such that $\mathbb{E}\left[\langle\xi_{t}^{k}(H^{k})^{-1},\xi_{t}^{k}\rangle\xi_{t}^{k}(\xi_{t}^{k})^{T}\right] \  \preceq\tilde{\kappa}H^{k}$, for all $k$. 
+The condition numbers $\kappa_{1}$ and $\tilde{\kappa}$
+are important in the characterization of convergence rates for FedAvg
+algorithms. Note that $\kappa_{1}>\kappa$ and $\kappa_{1}>\tilde{\kappa}$.
+\end{comment}
+%details to be moved to the appendix.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{comment}
@@
 \end{comment}
-Define $l$ to be the smallest positive number such that $\mathbb{E}\|\xi_{t}^{k}\|^{2}$$\mathbf{\xi}_{t}^{k}$($\mathbf{\xi}_{t}^{k})^{T}\preceq lH^{k}$
-for all $k$. Note that $l\leq\max_{k,j}\|\mathbf{x}_{k}^{j}\|^{2}$.
-Let $L$ and $\mu$ be lower and upper bounds of non-zero eigenvalues
-of $H^{k}$. Define $\kappa_{1}:=l/\mu$ and $\kappa:=L/\mu$. Following
-\cite{liu2018accelerating,jain2017accelerating}, we define the statistical
-condition number $\tilde{\kappa}$ as the smallest positive real number
-such that $\mathbb{E}\left[\langle\xi_{t}^{k}(H^{k})^{-1},\xi_{t}^{k}\rangle\xi_{t}^{k}(\xi_{t}^{k})^{T}\right] \  \preceq\tilde{\kappa}H^{k}$, for all $k$. 
-The condition numbers $\kappa_{1}$ and $\tilde{\kappa}$
-are important in the characterization of convergence rates for FedAvg
-algorithms. Note that $\kappa_{1}>\kappa$ and $\kappa_{1}>\tilde{\kappa}$.
-
-\subsection{Acceleration of FedAvg with Momentum-based Updates}
-
-We show that the bound in Theorem \ref{thm:overparameterized_general} can be improved to $\mathcal{O}(\exp(-\frac{N}{E\kappa_{1}}t))$ for quadratic problems, then propose a variant of FedAvg that has provable acceleration over FedAvg with SGD updates.
 

### 2020-06-03 source/sec_conclude.tex
Update sec_conclude.tex

@@
 \end{comment}
+\vspace{-1em}
 This paper provides a comprehensive and unified analysis of the convergence of FedAvg

### 2020-06-03 source/sec_impact.tex
updated overparameterized
add covid impact

@@
 
+Federated learning balancing efficiency and privacy is not only an effective
+tool to provide enhanced entertainment, but also a vital approach that can
+save millions of lives in the short future.  Under the unprecedented outbreak
+of Covid'19, coronavirus patients have filled the emergency rooms
+thoroughly. It is imperative to have reliable, robust machine learning
+solutions that could help doctors treat patients. However,  HIPAA regulations
+impose strict restrictions on sharing patients' clinical data: privacy is a
+top priority when developing machine learning techniques at those medical institutions. 
+Therefore, propelling the frontiers of federated learning for more efficient 
+algorithms under the privacy-preserving constraint has never been more
+important.
+
 In this work, we establish solid theoretical foundations for federated

### 2020-06-03 source/sec_impact.tex
add hipaa ref

@@
 thoroughly. It is imperative to have reliable, robust machine learning
-solutions that could help doctors treat patients. However,  HIPAA regulations
+solutions that could help doctors treat patients. However,  Health Insurance Portability and Accountability (HIPAA) regulations~\cite{assistance2003summary}
 impose strict restrictions on sharing patients' clinical data: privacy is a

### 2020-06-04 source/sec_intro.tex
Update on Overleaf.

@@
 service provider), while keeping the training data decentralized~\cite{kairouz2019advances}.
-In recent years, FL has swiftly emerged to be an important learning paradigm~\cite{mcmahan2016communication}--one that enjoys widespread success in applications such as personalized recommendations~\cite{chen2018federated}, virtual assistants~\cite{lamautonomy}, keyboard predictions~\cite{47586} to name a few--for at least two reasons: First, the rapid proliferation of smart devices that are equipped with both computing power and data-capturing capabilities provided the infrastructure core for FL.
+In recent years, FL has swiftly emerged to be an important learning paradigm~\cite{mcmahan2016communication} --one that enjoys widespread success in applications such as personalized recommendation~\cite{chen2018federated}, virtual assistant~\cite{lamautonomy}, keyboard prediction~\cite{47586} to name a few--for at least two reasons: First, the rapid proliferation of smart devices that are equipped with both computing power and data-capturing capabilities provided the infrastructure core for FL.
 Second, the rising awareness of privacy and the exponential growth of computational power (blessed by Moore's law) in mobile devices have made it increasingly attractive to push the computation to the edge.
@@
 considered, and the associated algorithm is often referred to as local SGD.
-~\cite{li2019convergence} are the first to provide convergence of FedAvg for strongly convex problems considering both statistical heterogeneity and system heterogeneity, yet it
+~\cite{li2019convergence} is the first to provide convergence of FedAvg for strongly convex problems considering both statistical heterogeneity and system heterogeneity, yet it
 didn't achieve linear speedup \cite{yu2019parallel,haddadpour2019convergence,stich2018local}.
@@
 %The same conclusions are equally applied for Nesterov accelerated FedAvg. 
-To the best of our knowledge, this is the first linear speedup result for FedAvg and accelerated variants in the full heterogeneous FL setting. Furthermore, 
+To the best of our knowledge, this is the first linear speedup result for FedAvg and accelerated variants in the fully heterogeneous FL setting. Furthermore, 
 we propose a new variant of FedAvg and provide the first improved convergence rate for stochastic momentum-accelerated FedAvg in the popular overparamterized setting.

### 2020-06-04 source/sec_impact.tex
updated impact section

--- 2020-06-01 source/sec_impact.tex
@@
 \section{Broader impact}
-With the widespread use of mobile devices, the powerful sensors
-collected enormous amount of data empowered with machine learning
-solutions can transform people's life in many aspects. The proper
-data integration can produce positive social
-impact in terms of better personalization~\cite{fallah2020personalized}, more efficient information solicitation~\cite{chen2018federated}, and enhanced service~\cite{47586}.
+With the growing availability of mobile devices and their increasing ability to capture data, enormous amounts of information can potentially be integrated to help improve and transform people's lives in many aspects such as personalized and enhanced services~\cite{fallah2020personalized,47586} and precision medicine~\cite{uddin2019artificial,frohlich2018hype}, to name a few.
 
-In the meantime, the increasing concerns of security and
-privacy leaking from both large companies and the entire society
-present significant challenges of proper utilization of local
-data. This trade-off between efficiency and privacy/security
-lies in the center of many industrial applications, including 
+Meanwhile, data breaches from central servers of big companies have become increasingly common, raising concerns about privacy and information security when individual clients' information is aggregated and stored centrally.
+
+The challenge we face is therefore to properly integrate and leverage data from different sources to arrive at better solutions, while at the same time minimizing privacy risks.
+This trade-off between efficiency and privacy/security
+lies at the heart of many applications, including 
 but not limited to recommendation systems~\cite{chen2018federated}, virtual assistants~\cite{lamautonomy}, and mobile keyboard prediction~\cite{47586}.
 
+Federated learning has emerged as a promising solution to the challenge.
+Built on insights from decentralized and distributed optimization, it aggregates information by integrating model \emph{parameters} learned locally at each source, eliminating the need to transmit and store data centrally.
+
+This desirable feature makes federated learning an effective tool to balance efficiency and privacy not just in the aforementioned applications with mobile devices, but also in other settings such as healthcare where it has the potential to save millions of lives.
+Under the unprecedented outbreak
+of Covid'19, coronavirus patients have filled emergency rooms nationwide. It is imperative to have reliable, robust machine learning
+solutions that could help doctors diagnose and treat patients. However,  Health Insurance Portability and Accountability (HIPAA) regulations~\cite{assistance2003summary}
+impose strict restrictions on sharing patients' clinical data: privacy is a
+top priority when developing solutions by leveraging information from different medical institutions. 
+Therefore, propelling the frontiers of federated learning to develop more efficient 
+algorithms under the privacy-preserving constraint has never been more
+important.
+
+Two important aspects where federated learning departs from traditional distributed optimization is the heterogeneity of data at different sources and the fact that not all local devices/sources are able to participate in the model averaging at every iteration.
+In this general setting, the question remains as to whether joint learning still enjoys the desirable feature of \emph{linear speedup} with respect to the number of participating devices/institutions.
+
 In this work, we establish solid theoretical foundations for federated
-learning that achieves efficiency and privacy at the same time. 
-On one hand, the data is isolated in each local devices, thus prevent
-the potential leakage of private information. On the other hand, the
-convergence rates can be improved substantially if the number of
-active local devices increases. In conclusion, our work 
-shows a faithful pathway to simultaneously secure people's privacy and improve
-users' experience.
+learning and demonstrate that linear speedup is possible in a variety of settings with the FedAvg algorithm and its variants, under both data and system heterogeneity, significantly improving previous understanding of federated learning.
+
+We therefore show that federated learning is able to achieve efficiency and privacy at the same time. On one hand, data is isolated in each local device/institution, preventing
+the potential leakage of private information. On the other hand, joint learning can speed up substantially if the number of
+participating devices/institutions increases. In conclusion, our work shows a promising machine learning pathway to simultaneously securing people's privacy and improving
+their lives.

### 2020-06-04 source/sec_intro.tex
minor intro fix

@@
 heterogeneity~\cite{stich2018local,zhou2017convergence,wang2018cooperative,woodworth2018graph,haddadpour2019convergence} or system
-heterogeneity~\cite{yu2019parallel,wang2019adaptive,khaled2019first,jiang2018linear,koloskova2020unified} are not
+heterogeneity~\cite{yu2019parallel,wang2019adaptive,khaled2019first,jiang2018linear,koloskova2020unified} is not
 considered, and the associated algorithm is often referred to as local SGD.
-~\cite{li2019convergence} is the first to provide convergence of FedAvg for strongly convex problems considering both statistical heterogeneity and system heterogeneity, yet it
+~\cite{li2019convergence} is the first to provide convergence of FedAvg for general strongly convex problems considering both statistical heterogeneity and system heterogeneity, yet it
 didn't achieve linear speedup \cite{yu2019parallel,haddadpour2019convergence,stich2018local}.

### 2020-06-04 source/sec_SGD.tex
minor fixes

@@
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}\leq\mathcal{O}(\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right)
 	\end{align*}
@@
 variance is $\Omega(E^{2}/T^{2})$ and the dependence on $E$ is tight.  \\
-\textbf{Comparison with related works.} To better understand the significance of the obtained bound, we compare our rates to the best-known results in the same setting.
-In~\cite{haddadpour2019convergence}, the authors prove a linear speedup $\mathcal{O}(1/NT)$ result for strongly convex and smooth objectives, with $\mathcal{O}(N^{1/3}T^{2/3})$ communication complexity with i.i.d.
+\textbf{Comparison with related works.} To better understand the significance of the obtained bound, we compare our rates to the best-known results in related settings.~\cite{haddadpour2019convergence} proves a linear speedup $\mathcal{O}(1/NT)$ result for strongly convex and smooth objectives, with $\mathcal{O}(N^{1/3}T^{2/3})$ communication complexity with i.i.d.
 data and partial participation.
-However, their results build upon the bounded gradient diversity assumption, which leads to a more restricted overparameterized setting (see discussions in Section~\ref{sec:assumptions}).
+However, their results build on the bounded gradient diversity assumption, which implies the existence of $\mathbf{w}^*$ that minimizes all local objectives (see discussions in Section~\ref{sec:assumptions}), effectively removing system heterogeneity.
+The bound in \cite{koloskova2020unified} matches our bound in the full participation case, but their framework excludes partial particiation (See their Proposition 1).
+
+\begin{comment}
 	In this overparameterized setting, we prove a geometric convergence rate (see Section~\ref{sec:overparameterized}), 
@@
 	better communication complexity. 
+\end{comment}
 

### 2020-06-04 source/sec_SGD.tex
minor fix

@@
 However, their results build on the bounded gradient diversity assumption, which implies the existence of $\mathbf{w}^*$ that minimizes all local objectives (see discussions in Section~\ref{sec:assumptions}), effectively removing system heterogeneity.
-The bound in \cite{koloskova2020unified} matches our bound in the full participation case, but their framework excludes partial particiation (See their Proposition 1).
+The bound in \cite{koloskova2020unified} matches our bound in the full participation case, but their framework excludes partial particiation (see their Proposition 1).
 

### 2020-06-04 source/sec_overparameterized.tex
minor updates
minor fixes
minor fix

@@
 we further propose a novel momentum-based FedAvg algorithm, which enjoys an
-improved convergence rate over FedAvg. Detailed proofs are deferred to Appendix Section~\ref{sec:interpolation}.
+improved convergence rate over FedAvg.
+Detailed proofs are deferred to Appendix Section~\ref{sec:interpolation}.
+In particular, we do not need Assumptions~\ref{ass:boundedvariance} and~\ref{ass:subgrad2} and use modified versions of Assumptions~\ref{ass:lsmooth} and~\ref{ass:stroncvx} detailed in this section.
 \begin{comment}
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{w}}_{T}) & \leq\frac{L}{2}(1-\overline{\alpha})^{T}\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}=\mathcal{O}\left(L\exp(-\frac{\mu}{E}\frac{NT}{l\nu_{\max}+L(N-\nu_{\min})})\cdot\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right).
+	\mathbb{E}F(\overline{\mathbf{w}}_{T}) & \leq\frac{L}{2}(1-\overline{\alpha})^{T}\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}=\mathcal{O}\left(L\exp\left(-\frac{\mu}{E}\frac{NT}{l\nu_{\max}+L(N-\nu_{\min})}\right)\cdot\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right).
 	\end{align*}
@@
 Here we use the fact $\tilde{\kappa} \leq \kappa_1$.
+Recall $\nu_{\max}=\max _k p_k N$ and $\nu_{\min}=\min _k p_k N$.
 
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{w}}_{T}) & \leq\mathcal{O}\left(L(1-\frac{N}{E(\nu_{\max}\kappa_{1}+(N-\nu_{\min}))})^{T}\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right).
+	\mathbb{E}F(\overline{\mathbf{w}}_{T}) & \leq\mathcal{O}\left(L\exp(-\frac{NT}{E(\nu_{\max}\kappa_{1}+(N-\nu_{\min}))})\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right).
 	\end{align*}
@@
 \mathbf{u}_{t}^{k}-\eta_{1}^{k}\mathbf{g}_{t,k} & \text{if }t+1\notin\mathcal{I}_{E},\\
-\sum_{k=1}^{N}p_{k}\left[\mathbf{u}_{t}^{k}-\eta_{1}^{k}\mathbf{g}_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E},
+\sum_{k \in \cS_{t+1}}\left[\mathbf{u}_{t}^{k}-\eta_{1}^{k}\mathbf{g}_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E},
 \end{cases}\\
@@
 that FedMaSS improves the convergence to $\mathcal{O}(\exp(-\frac{NT}{E\sqrt{\kappa_{1}\tilde{\kappa}}}))$.
-To our knowledge, this is the first accelerated result of
+To our knowledge, this is the first acceleration result of
 FedAvg with momentum updates over SGD updates.
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{\vw}_{T}) & \leq\mathcal{O}\left((1-\frac{1}{E}\frac{N}{l\nu_{\max}\sqrt{\kappa_{1}\tilde{\kappa}}+(N-\nu_{\min})})^{T}\|w_{0}-w^{\ast}\|^{2}\right).
+	\mathbb{E}F(\overline{\mathbf{w}}_{T}) & \leq\mathcal{O}\left(L\exp(-\frac{NT}{E(\nu_{\max}\sqrt{\kappa_{1}\tilde{\kappa}}+(N-\nu_{\min}))})\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right).
 	\end{align*}

### 2020-06-04 source/sec_intro.tex
minor fix

@@
 At the heart of the success of FL is the \emph{linear speedup} property with respect to the number of participating clients that can be achieved with infrequent communications.
-Although results of this kind are abundant in distributed and decentralized optimization~\cite{yu2019linear,yu2019parallel,stich2018local,khaled2020tighter,koloskova2020unified}, generalization to the FL setting with both types of heterogeneity is currently lacking.
+Although theoretical results of this kind are abundant in distributed and decentralized optimization~\cite{yu2019linear,yu2019parallel,stich2018local,khaled2020tighter,koloskova2020unified}, generalization to the FL setting with both types of heterogeneity is currently lacking.
 

### 2020-06-04 source/sec_Nesterov.tex
minor fixes
minor fix

@@
 and $\mathcal{O}(1/\sqrt{NT})$ convergence results for Nesterov-accelerated
-FedAvg for convex objectives that match the original FedAvg results.
+FedAvg for convex objectives that match the rates from the previous section.
 As we know from stochastic optimization, Nesterov and other momentum
@@
 momentum updates is able to accelerate over the original FedAvg algorithm. 
-The convergence proof of Nesterov-accelerated FedAvg is deferred to Appendix Section~\ref{sec:app:Nesterovfedavg}.
+Detailed proofs of convergence results in this section are deferred to Appendix Section~\ref{sec:app:Nesterovfedavg}.
 
@@
 \mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E},\\
-\sum_{k=1}^{N}p_{k}\left[\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E}.
+\sum_{k \in \cS_{t+1}}\left[\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E}.
 \end{cases}
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}\leq\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).
+	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).
 	\end{align*}

### 2020-06-04 source/sec_Nesterov.tex
minor fix

@@
     FedAvg algorithms with other momentum-based updates. However, in the
-    stochastic optimization setting, none of these methods will be able
-    to achieve a better rate than the original FedAvg with SGD updates in general.
+    stochastic optimization setting, none of these methods can achieve a better rate than the original FedAvg with SGD updates for general problems~\cite{kidambi2018insufficiency}.
     For this reason, we will instead turn to the overparameterized setting

### 2020-06-04 source/app_sgd_scvx_smooth.tex
minor update in proof

@@
-We first summarize some properties of $L$-smooth and $\mu$-strongly
-convex functions. 
+To facilitate reading, theorems from the main paper are restated and numbered identically. We first summarize some properties of $L$-smooth and $\mu$-strongly
+convex functions~\cite{rockafellar1970convex}. 
 \begin{lemma}
@@
 \end{lemma}
-%
-\begin{assumption}
-	The stochastic gradients have bounded variance 
-	\begin{align*}
-	\mathbb{E}\|\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})-\nabla F_{k}(w_{t}^{k})\|^{2} & \leq\sigma_{k}^{2}
-	\end{align*}
-	for all $k$. Moreover, the expected squared norm of stochastic gradients
-	is uniformly bounded, 
-	\begin{align*}
-	\mathbb{E}\|\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})\|^{2} & \leq G^{2}
-	\end{align*}
-\end{assumption}
 
-\begin{theorem}
+\begin{thm}
 	Assume that $F_{k}$ is $L$-smooth and $\mu$-strongly convex for
@@
 	\end{align*}
-\end{theorem}
+\end{thm}
 

### 2020-06-04 source/app_sgd_scvx_smooth.tex
updated proof

@@
 \begin{thm}
-	Assume that $F_{k}$ is $L$-smooth and $\mu$-strongly convex for
-	all $k$, and let $\nu_{\max}=\max_{k}\frac{1}{N}p_{k}$. Let $\kappa=\frac{L}{\mu}$,
-	$\gamma=\max\{32\kappa,E\}$ where $E$ is the communication delay,
-	and diminishing learning rates $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$.
-	Let $\overline{w}_{T}=\sum_{k=1}^{N}p_{k}w_{T}^{k}$ be the average
-	of local parameters at an arbitrary time $T$.
-
-	With full device participation, 
+	Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$,
+	$\nu_{\max}=\max_{k}Np_{k}$, and set decaying learning rates $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$
+	with $\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$. Then
+	under Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2} with full device participation, 
 	\begin{align*}
-	\mathbb{E}F(\overline{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right)
 	\end{align*}
-	and with partial device participation with $K$ sampled devices at
-	each communication round, 
+	and with partial device participation with at most $K$ sampled devices
+	at each communication round, 
 	\begin{align*}
-	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right)
 	\end{align*}
@@
 \begin{proof}
-		The idea of the proof is to observe that the $L$-smoothness of $F$
-		provides the upper bound
+	The first step is to observe that the $L$-smoothness of $F$ provides
+	the upper bound
 	\begin{align*}
@@
 	\begin{align*}
-		-2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle & =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
+	& -2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle\\
+	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
 	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
@@
 	\begin{align*}
-		2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right] & \leq2\alpha_{t}\left[F(\overline{w}_{t+1})-F(\overline{w}_{t})\right]\\
+	& 2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]\\
+	& \leq2\alpha_{t}\left[F(\overline{w}_{t+1})-F(\overline{w}_{t})\right]\\
 	& \leq2\alpha_{t}\mathbb{E}\langle\nabla F(\overline{w}_{t}),\overline{w}_{t+1}-\overline{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\overline{w}_{t+1}-\overline{w}_{t}\|^{2}\\
@@
 	\begin{align*}
-		\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2} & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}-(\overline{w}_{t}-\overline{w}_{t_{0}})\|^{2}\\
+	& \mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}\\
+	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}-(\overline{w}_{t}-\overline{w}_{t_{0}})\|^{2}\\
 	& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}\|^{2}\\
@@
 	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\
-		& \leq\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}\|g_{i,k}\|^{2}\\
-		& \leq\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{t_{0}}^{2}G^{2}\\
+	& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}E\alpha_{i}^{2}\|g_{i,k}\|^{2}\\
+	& \leq2\sum_{k=1}^{N}p_{k}E^{2}\alpha_{t_{0}}^{2}G^{2}\\
 	& \leq4E^{2}\alpha_{t}^{2}G^{2}
 	\end{align*}
+	\begin{comment}
+	Alternatively, we can also write
+	\begin{align*}
+	\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2} & =\sum_{k=1}^{N}p_{k}\|\overline{w}_{t-1}-\alpha_{t-1}g_{t-1}-w_{t-1}^{k}+\alpha_{t-1}g_{t-1,k}\|^{2}\\
+	& \leq2\sum_{k=1}^{N}p_{k}\left(\|\overline{w}_{t-1}-w_{t-1}^{k}\|^{2}+\|\alpha_{t-1}g_{t-1}-\alpha_{t-1}g_{t-1,k}\|^{2}\right)
+	\end{align*}
+	and write 
+	\begin{align*}
+	\sum_{k}p_{k}\|g_{t-1,k}-g_{t-1}\|^{2}\leq\sum_{k}p_{k}\|g_{t-1,k}\|^{2} & =\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(w_{t-1}^{k})+\nabla F_{k}(w_{t-1}^{k})\|^{2}\\
+	& \leq2\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(w_{t-1}^{k})\|^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(w_{t-1}^{k})\|^{2}\\
+	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(w_{t-1}^{k})-\nabla F_{k}(\overline{w}_{t-1})+\nabla F_{k}(\overline{w}_{t-1})\|^{2}\\
+	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+4\sum_{k}p_{k}\left(\|\nabla F_{k}(w_{t-1}^{k})-\nabla F_{k}(\overline{w}_{t-1})\|^{2}+\|\nabla F_{k}(\overline{w}_{t-1})\|^{2}\right)\\
+	& \leq2\sigma^{2}+4\sum_{k}p_{k}\left(L^{2}\|\overline{w}_{t-1}-w_{t-1}^{k}\|^{2}+\|\nabla F_{k}(\overline{w}_{t-1})\|^{2}\right)
+	\end{align*}
+	so that 
+	\begin{align*}
+	\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\sum_{k=1}^{N}p_{k}\|\overline{w}_{t-1}-w_{t-1}^{k}\|^{2}+4\sigma^{2}\alpha_{t-1}^{2}+8\alpha_{t-1}^{2}\sum_{k}p_{k}\|\nabla F_{k}(\overline{w}_{t-1})\|^{2}
+	\end{align*}
+	and taking expectation, we have 
+	\begin{align*}
+	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t-1}-w_{t-1}^{k}\|^{2}+4\alpha_{t-1}^{2}(\sigma^{2}+2G^{2})
+	\end{align*}
+	Using this recusive relation at most $E$ times to arrive at a communication
+	round and that $\alpha_{t}L\leq1/8$ for all $t$, we have 
+	\begin{align*}
+	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}\leq E\frac{1}{L^{2}}(\sigma^{2}+2G^{2})
+	\end{align*}
+	which gives 
+	\end{comment}
+	{}
 
@@
 	\begin{align*}
-		\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	& \mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2}\\
+	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
 	& =\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\sum_{k=1}^{N}(p_{k}N)^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
@@
 	where $\sigma^{2}=\max_{k}\sigma_{k}^{2}$, and if $\alpha_{t}$
-		satisfies $L\alpha_{t}\leq\frac{1}{8}$. We show next that $\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}}{t^{2}})$.
+	satisfies $L\alpha_{t}\leq\frac{1}{8}$. We show next that $\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}LG^{2}}{t^{2}})$.
 
@@
 	\end{align*}
-		and so it remains to choose $\alpha_{t}$ and $b$ such that $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}\leq b\alpha_{t+1}$.
+	and so it remains to choose $\alpha_{t}$ and $b$ such that $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}\leq b\alpha_{t+1}$
+	and $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}\leq b\alpha_{t+1}^{2}$.
 	Recall that we require $\alpha_{t_{0}}\leq\alpha_{t}$ for any $t-t_{0}\leq E-1$,
@@
 	& =\left(b\frac{t+\gamma-4}{t+\gamma}+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}\\
-		& =b(\frac{t+\gamma-2}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
+	& =b(\frac{t+\gamma-3}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
 	& \leq b(\frac{t+\gamma-1}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
@@
 	\end{align*}
-		and where we have used the facts that 
+	where we have used the facts that 
 	\begin{align*}
@@
 	\begin{align*}
-		\mathbb{E}(F(\overline{w}_{T}))-F^{\ast} & =\mathbb{E}(F(\overline{w}_{T})-F(w^{\ast}))\\
+	& \mathbb{E}(F(\overline{w}_{T}))-F^{\ast}\\
+	& =\mathbb{E}(F(\overline{w}_{T})-F(w^{\ast}))\\
 	& \leq\frac{L}{2}\mathbb{E}\|\overline{w}_{T}-w^{\ast}\|^{2}\leq\frac{L}{2}c\|w_{0}-w^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
 	& =2c\|w_{0}-w^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
-		& \leq2c\|w_{0}-w^{\ast}\|^{2}\kappa\left[\frac{4}{\mu T}\cdot\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}LG^{2}\cdot(\frac{4}{\mu T})^{2}\right]\\
-		& =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa}{\mu^{2}}E^{2}LG^{2}\cdot\frac{1}{T^{2}})
+	& \leq2c\|w_{0}-w^{\ast}\|^{2}\kappa\left[\frac{4}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}LG^{2}\cdot(\frac{4}{\mu(T+\gamma)})^{2}\right]\\
+	& =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
 	\end{align*}
@@
 	With partial participation, the update at each communication round
-		is now given by averages over a subset of sampled devices. The key
-		is to bound 
+	is now given by averages over a subset of sampled devices. When $t+1\notin\mathcal{I}_{E}$,
+	$\overline{v}_{t+1}=\overline{w}_{t+1}$, while when $t+1\notin\mathcal{I}_{E}$,
+	we have $\mathbb{E}\overline{w}_{t+1}=\overline{v}_{t+1}$ by design
+	of the sampling schemes, so that 
 	\begin{align*}
-		\mathbb{E}\|\overline{w}_{t}-\overline{y}_{t}\|^{2} & =\frac{1}{K}\sum_{k}p^{k}\|w_{t}^{k}-\sum_{k}p_{k}w_{t}^{k}\|^{2}
+	\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & =\mathbb{E}\|\overline{w}_{t+1}-\overline{v}_{t+1}+\overline{v}_{t+1}-w^{\ast}\|^{2}\\
+	& =\mathbb{E}\|\overline{w}_{t+1}-\overline{v}_{t+1}\|^{2}+\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2}
 	\end{align*}
-		This can be bounded by the same argument as before and yield 
+	As before, $\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2}\leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
+
+	The key is to bound $\mathbb{E}\|\overline{w}_{t+1}-\overline{v}_{t+1}\|^{2}$.
+	For sampling scheme I we have 
 	\begin{align*}
-		\mathbb{E}\|\overline{w}_{t}-\overline{y}_{t}\|^{2} & \leq\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
+	\mathbb{E}\|\overline{w}_{t+1}-\overline{v}_{t+1}\|^{2} & =\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|w_{t+1}^{k}-\overline{w}_{t+1}\|^{2}\\
+	& \leq\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
 	\end{align*}
-		which results in the bound 
+	while for sampling scheme II 
+	\begin{align*}
+	\mathbb{E}\|\overline{w}_{t+1}-\overline{v}_{t+1}\|^{2} & =\frac{N-K}{N-1}\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|w_{t+1}^{k}-\overline{w}_{t+1}\|^{2}\\
+	& \leq\frac{N-K}{N-1}\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
+	\end{align*}
+	The same argument as the full participation case implies 
 	\begin{align*}
@@
 	\end{align*}
+
+	\begin{comment}
+	\begin{proof}
+	With partial participation, $2E\sum_{\tau=1}^{E-1}\alpha_{t-\tau}^{2}\sum_{k=1}^{N}p_{k}\left(\|\nabla F_{k}(\overline{w}_{t-\tau},\xi_{t-\tau}^{k})\|^{2}+l^{2}\|w_{t-\tau}^{k}-\overline{w}_{t-\tau}\|^{2}\right)$
+	\begin{align*}
+	\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{1}{K}\sum_{k}p^{k}\|w_{t}^{k}-\overline{w}_{t}\|^{2}\\
+	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}\\
+	& \leq\mathbb{E}(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})\cdots(1-\mu\alpha_{t-E})\|\overline{w}_{t-E}-w^{\ast}\|^{2}=O(\frac{1}{t-E}\sigma^{2}+E^{2}LG^{2}\frac{1}{(t-E)^{2}}+\frac{E^{2}G^{2}}{K}\frac{1}{(t-E)^{3/2}})\\
+	& +(\alpha_{t}^{3}+(1-\mu\alpha_{t-1})\alpha_{t-1}^{3}+(1-\mu\alpha_{t-1})(1-\mu\alpha_{t-2})\alpha_{t-2}^{3}+\cdots+(1-\mu\alpha_{t-1})\cdots(1-\mu\alpha_{t-E})\alpha_{t-E}^{3}E^{2}LG^{2}\\
+	& +(\alpha_{t}^{2}+(1-\mu\alpha_{t-1})\alpha_{t-1}^{2}+(1-\mu\alpha_{t-1})(1-\mu\alpha_{t-2})\alpha_{t-2}^{2}+\cdots+(1-\mu\alpha_{t-1})\cdots(1-\mu\alpha_{t-E})\alpha_{t-E}^{2}\sigma^{2}\\
+	& +\frac{4}{K}EG^{2}(\alpha_{t}^{2}+\alpha_{t-1}^{2}+\cdots+\alpha_{t-E}^{2})
+	\end{align*}
+
+	so need ot check 
+	\begin{align*}
+	(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})\cdots(1-\mu\alpha_{t-E+1})b\alpha_{t-E+1}^{\beta}+\alpha_{t}^{2} & \leq\alpha_{t+1}^{\beta}\\
+	\end{align*}
+
+	Alternatively, show that 
+	\begin{align*}
+	\frac{1}{K}\sum_{k}p^{k}\|w_{t}^{k}-\overline{w}_{t}\|^{2} & =O(\alpha_{t}^{2}G^{2}/K)\\
+	\sum_{k}p^{k}\|w_{t}^{k}-\overline{w}_{t}\|^{2}\le\sum_{k}p^{k}\|w_{t}^{k}-\overline{w}_{t_{0}}\|^{2} & \leq\sum_{k}p^{k}\|-\alpha_{t_{0}}g_{t_{0},k}-\alpha_{t_{0}+1}g_{t_{0}+1,k}-\cdots-\alpha_{t-1}g_{t-1,k}\|^{2}\\
+	& =\sum_{k}p^{k}\|\alpha_{t_{0}}g_{t_{0},k}+\alpha_{t_{0}+1}g_{t_{0}+1,k}+\cdots+\alpha_{t-1}g_{t-1,k}\|^{2}\\
+	& =\sum_{k}p^{k}\|\alpha_{t_{0}}\nabla F_{k}(\overline{w}_{t_{0}},\xi_{t_{0}}^{k})+\alpha_{t_{0}}\nabla F_{k}(\overline{w}_{t_{0}}-\alpha_{t_{0}+1}\nabla F_{k}(\overline{w}_{t_{0}},\xi_{t_{0}}^{k}),\xi_{t_{0}+1}^{k})+\cdots+\alpha_{t-1}g_{t-1,k}\|^{2}
+	\end{align*}
+	The intuition is to approximate with the sequence starting at $w^{\ast}$,
+	i.e. 
+	\begin{align*}
+	\sum_{k}p^{k}\|\alpha_{t_{0}}\nabla F_{k}(w^{\ast},\xi_{t_{0}}^{k})+\alpha_{t_{0}+1}\nabla F_{k}(w^{\ast}-\alpha_{t_{0}+1}\nabla F_{k}(w^{\ast},\xi_{t_{0}}^{k}),\xi_{t_{0}+1}^{k})+\cdots+\alpha_{t-1}g_{t-1,k}\|^{2} & \approx\\
+	\end{align*}
+	since $\overline{w}_{t_{0}}\approx w^{\ast}+\frac{1}{t_{0}}$. We
+	can take 
+	\begin{align*}
+	g_{t_{0},k} & =\nabla F_{k}(\overline{w}_{t_{0}},\xi)\approx\nabla F_{k}(\overline{w}_{t_{0}},\xi)+\\
+	\end{align*}
+
+	\begin{align*}
+	\alpha_{t}^{2}+(1-\mu\alpha_{t-1})\alpha_{t-1}^{2}+(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})b\alpha_{t-1} & \le\alpha_{t+1}\\
+	\alpha_{t}^{2}+(1-\mu\alpha_{t-1})\alpha_{t-1}^{2}+(1-\mu\alpha_{t-1})(1-\mu\alpha_{t-2})\alpha_{t-2}^{2}+(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})(1-\mu\alpha_{t-2})b\alpha_{t-2} & \le\alpha_{t+1}\\
+	\dots
+	\end{align*}
+	recall 
+	\begin{align*}
+	(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t} & \leq b\alpha_{t+1}
+	\end{align*}
+
+	\begin{align*}
+	(1-\mu\alpha_{t-1})\alpha_{t-1}^{2}+(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})b\alpha_{t-1} & \leq(1-\mu\alpha_{t})\alpha_{t-1}^{2}+(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})b\alpha_{t-1}\\
+	& \le(1-\mu\alpha_{t})\alpha_{t-1}(\alpha_{t-1}+(1-\mu\alpha_{t-1})b)\\
+	& \leq(1-\mu\alpha_{t})b\alpha_{t}
+	\end{align*}
+	so that 
+	\begin{align*}
+	(1-\mu\alpha_{t})\beta\alpha_{t}+\alpha_{t}^{2} & \leq b\alpha_{t+1}\\
+	\end{align*}
+	\begin{align*}
+	(1-\mu\alpha_{t-1})\alpha_{t-1}^{2}+(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})b\alpha_{t-1} & \leq\alpha_{t}^{2}+(1-\mu\alpha_{t-1})\alpha_{t-1}(\alpha_{t-1}+b(1-\mu\alpha_{t}))\le\\
+	(\alpha_{t-1}+b(1-\mu\alpha_{t}))\alpha_{t}\leq
+	\end{align*}
 	\end{proof}
+	\end{comment}
+
+	One may ask whether the dependence on $E$ in the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$
+	can be removed, or equivalently whether $\sum_{k}p_{k}\|\mathbf{w}_{t+1}^{k}-\overline{\mathbf{w}}_{t+1}\|^{2}$
+	can be independent of $E$. We provide a heuristic argument for why
+	this is not possible. Suppose that each $F_{k}(\mathbf{w})=\frac{1}{n}\sum(\mathbf{x}_{k}^{j}-\mathbf{w})^{2}$
+	contains data points $\mathbf{x}_{k}^{j}$ concentrated around the
+	optimizer $\mathbf{w}^{\ast,k}$, and that the $\mathbf{w}{}^{\ast,k}$'s
+	come in pairs around the origin. As a result, the global objective
+	$F$ is minimized at $\mathbf{w}^{\ast}=0$. Then with $t=O(T)$ large,
+	after a communication round $t$ each device will start with $\overline{\mathbf{w}}_{t}$
+	close to the origin, while after $E=T^{\beta}$ iterations with step
+	size $O(1/T)$, the local parameters are updated towards $\mathbf{w}^{\ast,k}$
+	with $\|\mathbf{w}_{t+E}^{k}\|^{2}\approx\frac{1}{T^{2-2\beta}}$,
+	while $\overline{\mathbf{w}}_{t+E}$ remains close to $\mathbf{w}^{\ast}$,
+	so that $\sum_{k}p_{k}\|\mathbf{w}_{t+E}^{k}-\overline{\mathbf{w}}_{t+E}\|^{2}=O(1/T^{2-2\beta})$,
+	which is at a slower rate than $1/T$ for any $\beta>0$. It follows
+	that $E=O(1)$. 
+\end{proof}

### 2020-06-04 source/app_sgd_cvx_smooth.tex
updated proof

@@
-\begin{theorem}
-	Suppose $F_{k}$ is $L$-smooth and convex for all $k$, and let $\nu_{\max}=\max_{k}\frac{1}{N}p_{k}$.
-	Set learning rate $\alpha_{t}=O(\sqrt{\frac{N}{T}})$.
-
-	With full device participation, 
+\begin{thm}
+	Under assumptions~\ref{ass:lsmooth},\ref{ass:boundedvariance},\ref{ass:subgrad2} and constant learning
+	rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$, 
 	\begin{align*}
-	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T}\right)
 	\end{align*}
-	and with partial device participation with $K$ sampled devices at
-	each communication round, 
+	with full participation, and with partial device participation with $K$ sampled devices at
+	each communication round and learning rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{K}{T}})$,
 	\begin{align*}
-	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}G^{2}}{KT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right)
 	\end{align*}
-\end{theorem}
+\end{thm}
 
@@
 	\begin{align*}
-	-2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle & =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
+	& -2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle\\
+	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
 	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
@@
 	\begin{align*}
-	\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]\\
+	& \|\overline{w}_{t+1}-w^{\ast}\|^{2}\\
+	& \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]\\
 	& +2\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|w_{t}^{k}-\overline{w}_{t}\|^{2}+4\alpha_{t}^{2}L(F(\overline{w}_{t})-F(w^{\ast}))+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
@@
 	\begin{align*}
-	\|\overline{w}_{t+1}-w^{\ast}\|^{2}+\alpha_{t}(F(\overline{w}_{t})-F(w^{\ast})) & \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+	& \|\overline{w}_{t+1}-w^{\ast}\|^{2}+\alpha_{t}(F(\overline{w}_{t})-F(w^{\ast}))\\
+	& \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
 	& \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+\frac{3}{2}\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
@@
 
-	By setting the constant learning rate $\alpha_{t}\equiv\sqrt{\frac{\|w_{0}-w^{\ast}\|^{2}}{\frac{1}{N}\nu_{\max}^{2}\sigma^{2}T+6E^{2}LG^{2}T^{2/3}}}=O(\sqrt{\frac{N}{T}})$,
+	By setting the constant learning rate $\alpha_{t}\equiv\sqrt{\frac{N}{T}}$,
 	we have
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & \leq\frac{1}{T\sqrt{\frac{\|w_{0}-w^{\ast}\|^{2}}{\frac{1}{N}\nu_{\max}^{2}\sigma^{2}T+6E^{2}LG^{2}T^{2/3}}}}\cdot\left(\|w_{0}-w^{\ast}\|^{2}+T\frac{\|w_{0}-w^{\ast}\|^{2}}{\frac{1}{N}\nu_{\max}^{2}\sigma^{2}T}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6E^{2}LG^{2}T\frac{\|w_{0}-w^{\ast}\|^{2}}{6E^{2}LG^{2}T}\right)\\
-	& =\frac{3\|w_{0}-w^{\ast}\|^{2}}{T\sqrt{\frac{\|w_{0}-w^{\ast}\|^{2}}{\frac{1}{N}\nu_{\max}^{2}\sigma^{2}T+6E^{2}LG^{2}T^{2/3}}}}=3\|w_{0}-w^{\ast}\|\sqrt{(\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{6E^{2}LG^{2}}{T^{4/3}})}
+	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & \leq\frac{1}{\sqrt{NT}}\cdot\|w_{0}-w^{\ast}\|^{2}+\frac{1}{\sqrt{NT}}T\cdot\frac{N}{T}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\frac{1}{\sqrt{NT}}T(\sqrt{\frac{N}{T}})^{3}6E^{2}LG^{2}\\
+	& \leq\frac{1}{\sqrt{NT}}\cdot\|w_{0}-w^{\ast}\|^{2}+\frac{1}{\sqrt{NT}}T\cdot\frac{N}{T}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\frac{N}{T}6E^{2}LG^{2}\\
+	& =(\|w_{0}-w^{\ast}\|^{2}+\nu_{\max}^{2}\sigma^{2})\frac{1}{\sqrt{NT}}+\frac{N}{T}6E^{2}LG^{2}
 	\end{align*}
 
-	Similarly, for partial participation, we just need to replace $\frac{\nu_{\max}^{2}\sigma^{2}}{NT}$
-	with $\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+C\frac{E^{2}G^{2}}{KT}$,
-	so that we have 
+	Similarly, for partial participation, we have 
 	\begin{align*}
-	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =O(\sqrt{\frac{\nu_{\max}^{2}\sigma^{2}}{NT}+\frac{E^{2}G^{2}}{KT}+\frac{E^{2}LG^{2}}{T^{4/3}}})
+	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|w_{0}-w^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot(\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+C)+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right)
+	\end{align*}
+	where $C=\frac{4}{K}E^{2}G^{2}$ or $\frac{N-K}{N-1}\frac{4}{K}E^{2}G^{2}$,
+	so that in the former case, with $\alpha_{t}=\sqrt{\frac{K}{T}}$,
+	we have 
+	\begin{align*}
+	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =\mathcal{O}(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T})
 	\end{align*}

### 2020-06-04 source/sec_problem.tex
minor fix
Update sec_problem.tex

@@
 where $N$ is the number of local devices (users/nodes/workers) and $p_k$ is the $k$-th device's weight satisfying $p_k \geq 0$ and $\sum_{k=1}^N p_k = 1$. 
-In $k$-th local device, there are $n_k$ data points:
+In the $k$-th local device, there are $n_k$ data points:
 $\vx_k^1, \vx_k^2, \dots, \vx_k^{n_k}$.  
@@
 data distributions of each local device. The local minimum
-$\vw^{*}_k= \argmin_{\vw \in R^d} F_k(\vw)$ can be far from
+$F^{*}_k= \min_{\vw \in R^d} F_k(\vw)$ can be far from
 the global minimum of \eq{\ref{eq:problem}}.
@@
 \subsection{The Federated Averaging (FedAvg) Algorithm}
-We first introduce the standard Federated Average (FedAvg) algorithm~\cite{mcmahan2016communication}.
+We first introduce the standard Federated Averaging (FedAvg) algorithm~\cite{mcmahan2016communication}.
 FedAvg updates the model in each device by local Stochastic Gradient Descent (SGD) and send the latest model to the central server every $E$
@@
 With partial participation, $\cS_{t+1}$ is obtained by two types of
-sampling schemes to simulate practical scenarios~\cite{li2019convergence}.
-For example, Sampling scheme I establishes $\cS_{t+1}$ by i.i.d.
-sampling the devices with probability $p_k$ with replacement.
+sampling schemes to simulate practical scenarios~\cite{li2019convergence}. For example, one scheme establishes $\cS_{t+1}$ by i.i.d. sampling the devices with probability $p_k$ with replacement. 
 Both schemes
@@
 Assumptions~\ref{ass:boundedvariance} and~\ref{ass:subgrad2} have been made in many previous
-works in federated learning, e.g.~\cite{yu2019parallel,li2019convergence,stich2018local}. We provide some more justification for their generality. As model average parameters become
+works in federated learning, e.g.~\cite{yu2019parallel,li2019convergence,stich2018local}. We provide further justification for their generality. As model average parameters become
 closer to $\mathbf{w}^{\ast}$, the $L$-smoothness property implies

### 2020-06-04 source/sec_SGD.tex
Update sec_SGD.tex

@@
 	with $\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$. Then
-	under Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2} with full device participation, 
+	under Assumptions~\ref{ass:lsmooth} to \ref{ass:subgrad2} with full device participation, 
+	% under Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2} with full device participation, 
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right)
+	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right),
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right)
+	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).
 	\end{align*}
@@
 However, their results build on the bounded gradient diversity assumption, which implies the existence of $\mathbf{w}^*$ that minimizes all local objectives (see discussions in Section~\ref{sec:assumptions}), effectively removing system heterogeneity.
-The bound in \cite{koloskova2020unified} matches our bound in the full participation case, but their framework excludes partial particiation (see their Proposition 1).
+The bound in \cite{koloskova2020unified} matches our bound in the full participation case, but their framework excludes partial particiation \cite[Proposition 1]{koloskova2020unified}.
 
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right)
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right).
 	\end{align*}

### 2020-06-04 source/sec_intro.tex
Update on Overleaf.

@@
 \subsection{Related Works}
-\textbf{The Convergence of FedAvg and Linear Speedup.}
+\textbf{Convergence of FedAvg}
 There have been a considerable amount of convergence results in distributed stochastic optimization with infrequent communications that establish linear speedup in the number of clients.
@@
 % \subsection{Contributions}
-\textbf{Contributions.}
+\subsection{Our Contributions.}
 We provide a comprehensive convergence analysis

### 2020-06-04 source/sec_intro.tex
add table delete conclusion

--- 2020-06-04 source/sec_intro.tex
@@
 
+\begin{table}[h!]
+\centering
+\begin{tabular}{|c|c|c|}\hline 
+	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex \\ \hline \hline
+	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
+	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$      \\ \hline
+\end{tabular}
+\caption{A high level summary of the convergence results for FedAvg and Nesterov accelerated FedAvg in this paper.}
+\label{tb:convergencerateintro}
+\end{table}
+
 % Our contributions
@@
 We provide a comprehensive convergence analysis
-of FedAvg and its accelerated variants in convex stochastic problems with both system and statistical heterogeneity.
+of FedAvg and its accelerated variants in convex stochastic problems with both system and statistical heterogeneity (see Table~\ref{tb:convergencerateintro}).
 Our results demonstrate the desirable linear speedup property with respect to the number of participating clients.

### 2020-06-04 source/main.tex
Update on Overleaf.
Sec 4, 8 minor
updated overparameterized
minor fix
Update main.tex
add table delete conclusion

@@
 \newcommand{\lkxcom}[1]{{\color{red}{#1}}}
+\usepackage{pifont}% http://ctan.org/pkg/pifont
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
+\def\lowcomma{_{\textstyle{,}}}
+\def\lowperiod{_{\textstyle.}}
 
@@
 
-\title{The Blessing of FedAvg's Linear Speedup in Federated Learning}
+\title{Federated Learning's Blessing:\\
+FedAvg has Linear Speedup}
 
@@
 In this work, we provide a
-comprehensive convergence analysis on Federated Averaging (FedAvg),
-a widely used framework in which SGD or Nesterov SGD updates are used for general convex problems with arbitrarily
-heterogeneous data. 
+comprehensive convergence analysis on Federated Averaging (FedAvg), a widely used framework in which SGD or Nesterov SGD updates are used with arbitrarily heterogeneous data. 
 We first show that the convergence rates using both updates enjoy a linear speedup w.r.t. the number of participating workers for all convex smooth problems. 
@@
 \input{sec_intro}
-\input{sec_relatedwork}
+%\input{sec_relatedwork}
 \input{sec_problem}
@@
 
-\input{sec_conclude}
+% \input{sec_conclude}
 
+\newpage
 \input{sec_impact}
-\newpage
-\bibliographystyle{plain}
+\bibliographystyle{unsrt}
 \bibliography{ref.bib}
 
+\newpage
 \appendix

### 2020-06-04 source/sec_intro.tex
Update on Overleaf.

@@
 convergence rate~\cite{li2019convergence}. Furthermore, the convergence of
-FedAvg's accelerated variants~\cite{yu2019linear,huo2020faster,liu2019accelerating} in the general heterogeneous and convex setting is yet to be analyzed.
+FedAvg's accelerated variants~\cite{yu2019linear,huo2020faster,liu2019accelerating} in the general heterogeneous and convex setting is unknown.
 At the heart of the success of FL is the \emph{linear speedup} property with respect to the number of participating clients that can be achieved with infrequent communications.
@@
 
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|}\hline 
-	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex \\ \hline \hline
-	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
-	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$      \\ \hline
-\end{tabular}
-\caption{A high level summary of the convergence results for FedAvg and Nesterov accelerated FedAvg in this paper.}
-\label{tb:convergencerateintro}
-\end{table}
-
 % Our contributions
@@
 FedAvg and its accelerated variants on both a synthetic dataset and  a real-world dataset. The empirical observations are well-aligned with the obtained theoretical findings.
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|c|c|c|}\hline 
+	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex \\ \hline \hline
+	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
+	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$      \\ \hline
+\end{tabular}
+\caption{Our convergence results for FedAvg and Nesterov accelerated FedAvg in this paper.}
+\label{tb:convergencerateintro}
+\end{table}
 

### 2020-06-04 source/sec_intro.tex
Update sec_intro.tex

--- 2020-06-04 source/sec_intro.tex
@@
 \centering
-\begin{tabular}{|c|c|c|}\hline 
-	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex \\ \hline \hline
-	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
-	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$      \\ \hline
+\begin{tabular}{|c|c|c|c|}\hline 
+	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex  & Overparameterized \\ \hline \hline
+	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$   & $\cO(\exp(-\frac{NT}{E}))$    \\ \hline
+	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$ & $\cO(\exp(-\frac{KT}{E}))$     \\ \hline
 \end{tabular}

### 2020-06-04 source/sec_intro.tex

--- 2020-06-04 source/sec_intro.tex
@@
 
+\begin{table}[h!]
+\centering
+\begin{tabular}{|c|c|c|c|}\hline 
+	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex  & Overparameterized \\ \hline \hline
+	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$   & $\cO(\exp(-\frac{NT}{E}))$    \\ \hline
+	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$ & $\cO(\exp(-\frac{KT}{E}))$     \\ \hline
+\end{tabular}
+\caption{A high level summary of the convergence results for FedAvg and Nesterov accelerated FedAvg in this paper.}
+\label{tb:convergencerateintro}
+\end{table}
+
 % Our theoretical results not only cover all general convex objectives, including strongly convex cases, convex smooth cases and convex non-smooth cases, but also provide tighter convergence guarantee for all convex smooth objectives, which shows the convergence rate enjoys a linear speedup w.r.t.

### 2020-06-04 source/sec_intro.tex
Update sec_intro.tex

@@
 \centering
-\begin{tabular}{|c|c|c|}\hline 
-	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex \\ \hline \hline
-	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
-	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$      \\ \hline
-\end{tabular}
-\caption{Our convergence results for FedAvg and Nesterov accelerated FedAvg in this paper.}
-\label{tb:convergencerateintro}
-\end{table}
-
-\begin{table}[h!]
-\centering
+{\small
 \begin{tabular}{|c|c|c|c|}\hline 
@@
 \end{tabular}
-\caption{A high level summary of the convergence results for FedAvg and Nesterov accelerated FedAvg in this paper.}
+}
+\caption{Our convergence results for FedAvg and Nesterov accelerated FedAvg in this paper.}
 \label{tb:convergencerateintro}

### 2020-06-05 source/sec_intro.tex
Update on Overleaf.

@@
 We provide a comprehensive convergence analysis
-of FedAvg and its accelerated variants in convex stochastic problems with both system and statistical heterogeneity (see Table~\ref{tb:convergencerateintro}).
+of FedAvg and its accelerated variants in the presence of both system and statistical heterogeneity (see Table~\ref{tb:convergencerateintro}).
 Our results demonstrate the desirable linear speedup property with respect to the number of participating clients.
@@
 }
-\caption{Our convergence results for FedAvg and Nesterov accelerated FedAvg in this paper.}
+\caption{Our convergence results for FedAvg and accelerated FedAvg in this paper.}
 \label{tb:convergencerateintro}

### 2020-06-05 source/sec_intro.tex
Update sec_intro.tex

--- 2020-06-04 source/sec_intro.tex
@@
 	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex  & Overparameterized \\ \hline \hline
-	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$   & $\cO(\exp(-\frac{NT}{E}))$    \\ \hline
-	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$ & $\cO(\exp(-\frac{KT}{E}))$     \\ \hline
+	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$   & $\cO(\exp(-\frac{NT}{E\kappa }))^{\dagger}$    \\ \hline
+	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$ & $\cO(\exp(-\frac{KT}{E\kappa }))^{\dagger}$     \\ \hline
 \end{tabular}
 }
-\caption{Our convergence results for FedAvg and Nesterov accelerated FedAvg in this paper.}
+\caption{Our convergence results for FedAvg and accelerated FedAvg in this paper.  }
+{\raggedright 
+         $^{\dagger}$ In the overparaterized setting, $\kappa=\kappa_1$ for FedAvg and $\kappa=\sqrt{\kappa_1\tilde{\kappa}}$ for accelerated FedAvg. Since
+$\kappa_{1}\geq\tilde{\kappa}$, this implies a speedup factor of
+$\sqrt{\frac{\kappa_{1}}{\tilde{\kappa}}}$ for accelerated FedAvg.\\
+           \par}
 \label{tb:convergencerateintro}

### 2020-06-05 source/main.tex
Update on Overleaf.

@@
 In this work, we provide a
-comprehensive convergence analysis on Federated Averaging (FedAvg), a widely used framework in which SGD or Nesterov SGD updates are used with arbitrarily heterogeneous data. 
+comprehensive study of Federated Averaging (FedAvg), a widely used framework in which SGD or Nesterov SGD updates are used in the presence of both heterogeneous data and heterogeneous systems. 
 We first show that the convergence rates using both updates enjoy a linear speedup w.r.t. the number of participating workers for all convex smooth problems. 

### 2020-06-05 source/sec_intro.tex
Update sec_intro.tex

@@
 {\raggedright 
-         $^{\dagger}$ In the overparaterized setting, $\kappa=\kappa_1$ for FedAvg and $\kappa=\sqrt{\kappa_1\tilde{\kappa}}$ for accelerated FedAvg. Since
+         $^{\dagger}$ In the overparaterized setting, we have $\kappa=\kappa_1$ for FedAvg and $\kappa=\sqrt{\kappa_1\tilde{\kappa}}$ for accelerated FedAvg. Since
 $\kappa_{1}\geq\tilde{\kappa}$, this implies a speedup factor of

### 2020-06-05 source/sec_intro.tex
update table

@@
 
+%\multirow{2}{*}{Participation}
+% \multirow{2}{*}{\backslashbox{Participation}{Objective function}} 
 \begin{table}[h!]
+\hspace{-1em}
+{\small
 \centering
-{\small
-\begin{tabular}{|c|c|c|c|}\hline 
-	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex  & Overparameterized \\ \hline \hline
-	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$   & $\cO(\exp(-\frac{NT}{E\kappa }))^{\dagger}$    \\ \hline
-	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$ & $\cO(\exp(-\frac{KT}{E\kappa }))^{\dagger}$     \\ \hline
+\begin{tabular}{|c|c|c|c|c|}\hline 
+	\multirow{2}{*}{\backslashbox{{\tiny Participation} }{{\tiny Objective function}}} & \multirow{2}{*}{Strongly Convex}     &\multirow{2}{*}{Convex}    & Overparameterized & Overparameterized \\ 
+	                                &                        &         &     general case                 & linear regression   \\ \hline 
+	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$   & $\cO(\exp(-\frac{NT}{E}))$ & $\cO(\exp(-\frac{NT}{E\kappa }))^{\dagger}$    \\ \hline
+	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$ &  $\cO(\exp(-\frac{NT}{EL}))$ & $\cO(\exp(-\frac{KT}{EL\kappa }))^{\dagger}$     \\ \hline
 \end{tabular}
@@
 $\sqrt{\frac{\kappa_{1}}{\tilde{\kappa}}}$ for accelerated FedAvg.\\
+% $^{*}$ The fourth column presents the results for the general overparaterized setting. The fifth column presents the results for overparaterized linear regression.
            \par}

### 2020-06-05 source/main.tex
update table
Update on Overleaf.

--- 2020-06-04 source/main.tex
@@
      \usepackage[nonatbib]{neurips_2020}
-
+\usepackage{slashbox,multirow}
+% \usepackage{tikz}
 \usepackage[utf8]{inputenc} % allow utf-8 input
@@
 \begin{abstract}
-Federated learning learns
-a model jointly from a set of distributed workers without sharing their private
-data. The characteristics of non-\emph{i.i.d.} data
-across the network and low device participation bring significant challenges
-in understanding the convergence of federated learning algorithms. Prior work
-analyzed the convergence with restrained settings,
-%either \emph{i.i.d.} data, or with full device participation, 
-and often with suboptimal rates. 
-In this work, we provide a
-comprehensive convergence analysis on Federated Averaging (FedAvg), a widely used framework in which SGD or Nesterov SGD updates are used with arbitrarily heterogeneous data. 
-We first show that the convergence rates using both updates enjoy a linear speedup w.r.t. the number of participating workers for all convex smooth problems. 
-In addition, in the overparamterized setting, we show that the convergence of FedAvg is in fact exponential, also with a speedup factor linear in the number of workers. 
-Furthermore, we propose a new accelerated federated learning algorithm that further improves this convergence rate with provable guarantees in the linear regression setting. 
+Federated learning (FL) learns a model jointly from a set of participating devices without sharing each other's privately held data.
+The characteristics of non-\textit{iid} data across the network, low device participation and the mandate that data remain private bring challenges in understanding the convergence of FL algorithms, particularly in regards to how convergence scales with the number of participating devices.
+In this paper, we focus on Federated Averaging (FedAvg)--the most widely used and effective FL algorithm in use today--and provide a comprehensive study of its convergence rate.
+Although FedAvg has recently been studied by an emerging line of literature, it remains open as to how FedAvg's convergence scales with the number of participating devices in the FL setting--a crucial question whose answer would shed light on the performance of FedAvg in large FL systems.
+We fill this gap by establishing convergence guarantees for FedAvg under two classes of problems: strongly convex smooth, convex smooth, and overparametrized strongly convex smooth problems.
+We show that FedAvg enjoys linear speedup in each case, although with different convergence rates.
+For general convex and strongly convex problems, we also characterize the corresponding convergence rates for the Nesterov accelerated FedAvg algorithm in the FL setting: to the best of our knowledge, they are the first linear speedup guarantees for FedAvg when Nesterov acceleration is used.
+We also design a new  accelerated FL algorithm that further improves the convergence rate in overparametrized linear regression problems.
 Empirical studies of the algorithms in various settings have supported our theoretical results.
@@
 
+% (and its related variant local SGD also analyzed in the simpler homogeneous distributed optimization setting)
 \input{sec_intro}

### 2020-06-05 source/sec_problem.tex
Update sec_problem.tex
Update on Overleaf.

@@
 We first introduce the standard Federated Averaging (FedAvg) algorithm~\cite{mcmahan2016communication}.
-FedAvg updates the model in each device by local Stochastic Gradient Descent (SGD) and send the latest model to the central server every $E$
+FedAvg updates the model in each device by local Stochastic Gradient Descent (SGD) and sends the latest model to the central server every $E$
 steps. The central server conducts a weighted average over the model parameters
@@
 Formally, the updates of FedAvg at round $t$ is described as follows:
-\begin{align}
-\vv_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}\vg_{t,k}\\
-\mathbf{w}_{t+1}^{k} &=\left\{
+\begin{align*} 
+\vv_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}\vg_{t,k}, \hspace{1em}
+\mathbf{w}_{t+1}^{k} =\left\{
 \begin{array}{ll}
 \mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\ 
-\sum_{k \in \cS_{t+1}} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}
+\sum_{k \in \cS_{t+1}} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E},
 \end{array}\right.
-\end{align}
+\end{align*}
 where $\vw_t^k$ is the local model parameter maintained in the $k$-th device at the $t$-th iteration, $\vg_{t,k}:=\nabla F_{k}(\vw_{t}^{k},\xi_{t}^{k})$ is the stochastic gradient based on $\xi_{t}^{k}$, the data sampled from $k$-th device’s local data uniformly at random.
@@
 With partial participation, $\cS_{t+1}$ is obtained by two types of
-sampling schemes to simulate practical scenarios~\cite{li2019convergence}. For example, one scheme establishes $\cS_{t+1}$ by i.i.d. sampling the devices with probability $p_k$ with replacement. 
+sampling schemes to simulate practical scenarios~\cite{li2019convergence}.
+For example, one scheme establishes $\cS_{t+1}$ by \emph{i.i.d.} sampling the devices with probability $p_k$ with replacement.
 Both schemes
@@
 We make the following standard assumptions on the objective function $F_1,\dots, F_N$.
-Assumptions~\ref{ass:lsmooth} and~\ref{ass:stroncvx} are commonly satisfied by a range of popular objective functions, such as $\ell^{2}$ regularized logistic regression and cross-entropy loss functions.
+Assumptions~\ref{ass:lsmooth} and~\ref{ass:stroncvx} are commonly satisfied by a range of popular objective functions, such as $\ell^{2}$-regularized logistic regression and cross-entropy loss functions.
 
@@
 and assuming the bounds at $\mathbf{w}^{\ast}$ only.
-Furthermore, compared to assuming \textit{bounded gradient diversity} as in related works~\cite{haddadpour2019convergence,li2018federated}, Assumption~\ref{ass:subgrad2} is much less restrictive.
-When target accuracy $\epsilon \rightarrow 0$,
+Furthermore, compared to assuming \textit{bounded gradient diversity} as in related work~\cite{haddadpour2019convergence,li2018federated}, Assumption~\ref{ass:subgrad2} is much less restrictive.
+When the optimality gap converges to zero,
 bounded gradient diversity restricts local objectives to have the same minimizer as the global objective, contradicting the heterogeneous data setting. 

### 2020-06-05 source/main.tex

@@
 We show that FedAvg enjoys linear speedup in each case, although with different convergence rates.
-For general convex and strongly convex problems, we also characterize the corresponding convergence rates for the Nesterov accelerated FedAvg algorithm in the FL setting: to the best of our knowledge, they are the first linear speedup guarantees for FedAvg when Nesterov acceleration is used.
+For general convex and strongly convex problems, we also characterize the corresponding convergence rates for the Nesterov accelerated FedAvg algorithm in the FL setting: to the best of our knowledge, these are the first linear speedup guarantees for FedAvg when Nesterov acceleration is used.
 We also design a new accelerated FL algorithm that further improves the convergence rate in overparametrized linear regression problems.

### 2020-06-05 source/sec_intro.tex

--- 2020-06-05 source/sec_intro.tex
@@
 Federated learning (FL) is a machine learning setting where many clients (e.g.
-mobile devices or whole organizations) collaboratively train a model under the orchestration of a central server (e.g.
-service provider), while keeping the training data decentralized~\cite{kairouz2019advances}.
-In recent years, FL has swiftly emerged to be an important learning paradigm~\cite{mcmahan2016communication} --one that enjoys widespread success in applications such as personalized recommendation~\cite{chen2018federated}, virtual assistant~\cite{lamautonomy}, keyboard prediction~\cite{47586} to name a few--for at least two reasons: First, the rapid proliferation of smart devices that are equipped with both computing power and data-capturing capabilities provided the infrastructure core for FL.
+mobile devices or organizations) collaboratively train a model under the orchestration of a central server (e.g.
+service provider), while keeping the training data decentralized~\cite{smith2017federated, kairouz2019advances}.
+In recent years, FL has swiftly emerged as an important learning paradigm~\cite{mcmahan2016communication,li2018federated}--one that enjoys widespread success in applications such as personalized recommendation~\cite{chen2018federated}, virtual assistant~\cite{lamautonomy}, and keyboard prediction~\cite{47586}, to name a few--for at least two reasons: First, the rapid proliferation of smart devices that are equipped with both computing power and data-capturing capabilities provided the infrastructure core for FL.
 Second, the rising awareness of privacy and the exponential growth of computational power (blessed by Moore's law) in mobile devices have made it increasingly attractive to push the computation to the edge.
 
-Despite its promise and broad applicability in our current era, the potential value FL delivers is coupled with the unique challenges it brings forth. In particular,
-when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated (and hence private)~\cite{kairouz2019advances}, it faces two challenges that are absent in centralized optimization and traditional distributed optimization~\cite{li2019convergence,smith2017federated}:
-\begin{enumerate}
-\item\textbf{Statistical heterogeneity:} data distributions in different devices are different.
-\item\textbf{System heterogeneity:} only a subset of devices may access the central server at each time. 
-\end{enumerate}
+Despite its promise and broad applicability in our current era, the potential value FL delivers is coupled with the unique challenges it brings forth.
+In particular, when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated (and hence protects privacy)~\cite{kairouz2019advances}, it faces two challenges that are absent in centralized optimization and distributed (stochastic) optimization~\cite{zhou2017convergence, stich2018local,khaled2019first,liang2019variance,wang2018cooperative,woodworth2018graph,wang2019adaptive,jiang2018linear,yu2019parallel,yu2019linear, khaled2020tighter,koloskova2020unified}:
+
+1) \textbf{Data heterogeneity:} data distributions in devices are different (and data can't be shared);
+
+2) \textbf{System heterogeneity:} only a subset of devices may access the central server at each time both because the communications bandwidth profiles vary across devices and because there is no central server that has control over when a device is active.
 
@@
 This success has since motivated a growing line of research efforts into understanding its theoretical convergence guarantees in various settings.
-Most existing convergence guarantees of FedAvg were either established in simplified settings~\cite{stich2018local,khaled2019first,wang2018cooperative,yu2019parallel,khaled2020tighter,li2018federated,koloskova2020unified}, have restrictive assumptions~\cite{haddadpour2019convergence} or provide a suboptimal
-convergence rate~\cite{li2019convergence}. Furthermore, the convergence of
-FedAvg's accelerated variants~\cite{yu2019linear,huo2020faster,liu2019accelerating} in the general heterogeneous and convex setting is unknown.
-At the heart of the success of FL is the \emph{linear speedup} property with respect to the number of participating clients that can be achieved with infrequent communications.
-Although theoretical results of this kind are abundant in distributed and decentralized optimization~\cite{yu2019linear,yu2019parallel,stich2018local,khaled2020tighter,koloskova2020unified}, generalization to the FL setting with both types of heterogeneity is currently lacking.
+For instance, \cite{haddadpour2019convergence} analyzed FedAvg (for non-convex smooth problems satisfying PL conditions) under the assumption that each local device's minimizer is the same as the minimizer of the joint problem (if all devices' data is aggregated together), an overly restrictive assumption.
+Very recently, \cite{li2019convergence} furthered the progress and established an $\cO(\frac{1}{T})$ convergence rate for FedAvg for strongly convex smooth problems. 
+At the same time, \cite{huo2020faster} studied the Nesterov accelerated FedAvg for non-convex smooth problems and established 
+an $\cO(\frac{1}{\sqrt{T}})$ convergence rate to stationary points.
+% (Prior to that, \cite{liu2019accelerating} studied Nesterov accelerated FedAvg using full gradient--rather than stochastic gradient as in both the above-mentioned works and our setting--for strongly convex smooth problems and established an $\cO(\frac{1}{T})$ convergence rate).
 
-\subsection{Related Works}
-\textbf{Convergence of FedAvg}
-There have been a considerable amount of convergence results in distributed stochastic optimization with infrequent communications that establish linear speedup in the number of clients.
-However, most of these prior works provide
-convergence guarantees under simplified settings, where either statistical
-heterogeneity~\cite{stich2018local,zhou2017convergence,wang2018cooperative,woodworth2018graph,haddadpour2019convergence} or system
-heterogeneity~\cite{yu2019parallel,wang2019adaptive,khaled2019first,jiang2018linear,koloskova2020unified} is not
-considered, and the associated algorithm is often referred to as local SGD.
-~\cite{li2019convergence} is the first to provide convergence of FedAvg for general strongly convex problems considering both statistical heterogeneity and system heterogeneity, yet it
-didn't achieve linear speedup \cite{yu2019parallel,haddadpour2019convergence,stich2018local}.
-In~\cite{haddadpour2019convergence,liang2019variance,huo2020faster,jiang2018linear}, the convergence of FedAvg for non-convex problems has been studied, whereas this paper focuses on convex settings.
-\begin{comment}
-However, \cite{huo2020faster} did not discuss the linear speedup and the theorems in \cite{haddadpour2019convergence,liang2019variance} failed to cover the accelerated version of
-FedAvg. 
-\end{comment}
+However, despite these very recent fruitful pioneering efforts into understanding the theoretical convergence properties of FedAvg, it remains open as to how the number of devices--particularly the number of devices that participate in the computation--affect the convergence speed.
+In particular, do we get linear speedup of FedAvg? What about when FedAvg is accelerated? These aspects are currently unexplored in FL. We fill in the gaps here by providing affirmative answers.
 
-\textbf{Accelerating Stochastic Federated Learning Algorithms.}
-% The accelerated SGD methods has achieved optimal convergence rate in strongly convex
-% optimization~\cite{nesterov1983method}.
-The convergence of accelerated FL algorithms using momentum (Nesterov and Heavy Ball) updates has been studied mostly in non-convex settings~\cite{yu2019linear,huo2020faster}.
-\cite{yu2019linear} does not consider system heterogeneity, while the result in \cite{huo2020faster} does not have linear speedup.
-Furthermore, the rates cannot improve over SGD-based algorithms in general.
-\cite{liu2019accelerating} provides an acceleration result for FedAvg with accelerated GD.
-\cite{liang2019variance} considers variance reduction based acceleration of FedAvg.
-This paper considers accelerated FedAvg with momentum-based stochastic updates in general and overparameterized convex settings.
-
-\begin{comment}
-	In this work, we provide a novel accelerated FedAvg with improved convergence rates under a popular overparameterized setting. 
-\end{comment}
-
-% Our contributions
-% \subsection{Contributions}
-\subsection{Our Contributions.}
-We provide a comprehensive convergence analysis
-of FedAvg and its accelerated variants in the presence of both system and statistical heterogeneity (see Table~\ref{tb:convergencerateintro}).
-Our results demonstrate the desirable linear speedup property with respect to the number of participating clients.
-More specifically, for strongly convex and smooth problems, we show 
-that FedAvg admits a convergence rate of 
-{\small{$\cO(1/NT)$}}, and local steps can be as large as {\small{$\cO(\sqrt{T/N})$}}. For convex and smooth objectives, the rate is {\small{$\cO(1/\sqrt{NT})$}}. 
-Moreover, we show that FedAvg with Nesterov acceleration also enjoys the same rates and speedup. 
-%The same conclusions are equally applied for Nesterov accelerated FedAvg. 
-To the best of our knowledge, this is the first linear speedup result for FedAvg and accelerated variants in the fully heterogeneous FL setting. Furthermore, 
-we propose a new variant of FedAvg and provide the first improved convergence rate for stochastic momentum-accelerated FedAvg in the popular overparamterized setting.
-Last but not least, we systematically examine the convergence rate of
-FedAvg and its accelerated variants on both a synthetic dataset and  a real-world dataset. The empirical observations are well-aligned with the obtained theoretical findings.
-
-\begin{table}[h!]
+\begin{table}[t!]
+\hspace{-1em}
+{\small
 \centering
-{\small
-\begin{tabular}{|c|c|c|c|}\hline 
-	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex  & Overparameterized \\ \hline \hline
-	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$   & $\cO(\exp(-\frac{NT}{E\kappa }))^{\dagger}$    \\ \hline
-	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$ & $\cO(\exp(-\frac{KT}{E\kappa }))^{\dagger}$     \\ \hline
+\begin{tabular}{|c|c|c|c|c|}\hline 
+	\multirow{2}{*}{\backslashbox{{\tiny Participation} }{{\tiny Objective function}}} & \multirow{2}{*}{Strongly Convex}     &\multirow{2}{*}{Convex}    & Overparameterized & Overparameterized \\ 
+	                                &                        &         &     general case                 & linear regression   \\ \hline 
+	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$   & $\cO(\exp(-\frac{NT}{E\kappa_1}))$ & $\cO(\exp(-\frac{NT}{E\kappa }))^{\dagger}$    \\ \hline
+	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$ &  $\cO(\exp(-\frac{KT}{E\kappa_1}))$ & $\cO(\exp(-\frac{KT}{E\kappa }))^{\dagger}$     \\ \hline
 \end{tabular}
 }
-\caption{Our convergence results for FedAvg and accelerated FedAvg in this paper.  }
-{\raggedright 
-         $^{\dagger}$ In the overparaterized setting, $\kappa=\kappa_1$ for FedAvg and $\kappa=\sqrt{\kappa_1\tilde{\kappa}}$ for accelerated FedAvg. Since
+\caption{Our convergence results for FedAvg and accelerated FedAvg in this paper. Throughout the paper, $N$ is the total
+number of local devices, and $K \leq N$ is the maximal number of devices that are accessible to the central server.
+$T$ is the total number of stochastic updates performed by each local device, $E$ is the local steps between two consecutive server communications (and hence $T/E$ is the number of communications).
+$^{\dagger}$ In the linear regression setting, we have $\kappa=\kappa_1$ for FedAvg and $\kappa=\sqrt{\kappa_1\tilde{\kappa}}$ for accelerated FedAvg, where $\kappa_1$ and $\sqrt{\kappa_1\tilde{\kappa}}$ are condition numbers defined in Section \ref{sec:overparameterized}.
+Since
 $\kappa_{1}\geq\tilde{\kappa}$, this implies a speedup factor of
-$\sqrt{\frac{\kappa_{1}}{\tilde{\kappa}}}$ for accelerated FedAvg.\\
-           \par}
+$\sqrt{\frac{\kappa_{1}}{\tilde{\kappa}}}$ for accelerated FedAvg.}
+% {\raggedright 
+% $^{*}$ The fourth column presents the results for the general overparaterized setting. The fifth column presents the results for overparaterized linear regression.
+        %   \par}
 \label{tb:convergencerateintro}
+\vspace{-2.5em}
 \end{table}
+
+\textbf{Our Contributions}
+We provide a comprehensive convergence analysis
+of FedAvg and its accelerated variants in the presence of both data and system heterogeneity. 
+Our contributions are threefold.
+
+First, we establish an {\small{$\cO(1/KT)$}} convergence rate  under FedAvg for strongly convex and smooth problems and  an
+{\small{$\cO(1/\sqrt{KT})$}} convergence rate for convex
+and smooth problems (where $K$ is the number of participating devices), thereby establishing that FedAvg enjoys the desirable linear speedup property in the FL setup.
+Prior to our work here, the best and the most related convergence analysis is given by \cite{li2019convergence}, which established an $\cO(\frac{1}{T})$ convergence rate for strongly convex smooth problems under FedAvg.
+Our rate matches the same (and optimal) dependence on $T$, but also completes the picture by establishing the linear dependence on $K$.
+
+Second, we establish the same convergence rates--{\small{$\cO(1/KT)$}} for strongly convex and smooth problems and {\small{$\cO(1/\sqrt{KT})$}} for convex and smooth problems--for Nesterov accelerated FedAvg.
+We analyze the accelerated version of FedAvg here because empirically it tends to perform better; yet, its theoretical convergence guarantee is unknown.
+To the best of our knowledge, these are the first results that provide a linear speedup characterization of Nesterov accelerated FedAvg in those two problem classes (that FedAvg and Nesterov accelerated FedAvg share the same convergence rate is to be expected: this is the case even for centralized stochastic optimization).
+
+Third, we study a subclass of strongly convex smooth problems where the objective is over-parameterized and establish 
+a faster $\cO(\exp(-\frac{KT}{\kappa}))$ convergence rate for FedAvg.
+Within this class, we further consider the linear regression problem and establish an even sharper rate under FedAvg.
+In addition, we propose a new variant of accelerated FedAvg--MaSS accelerated FedAvg--and establish a faster convergence rate (compared to if no acceleration is used).
+This stands in contrast to generic (strongly) convex stochastic problems where theoretically no rate improvement is obtained when one accelerates FedAvg.
+The detailed convergence results are summarized in Table~\ref{tb:convergencerateintro}.
+
+% The detailed convergence results are summarized in Table~\ref{tb:convergencerateintro}, which also include aspects we did not have space to dicuss (i.e.
+how the rate scales with communications frequency etc.).
+Note that the full participation case corresponds to $K=N$.
 
@@
 % evaluation on both synthetic and real-world dataset, which demonstrates that our theoretical indications is well-aligned with the empirical observations.
-
-\textbf{Notations.}
-Throughout the paper, we adopt the following notations. Let $N$ be the total
-number of local devices, and $K \leq N$ is the maximal number of devices
-that are accessible to the central server.
-Let $T$ be the total number of stochastic updates performed by each local device, $E$ be the local steps between two consecutive server communications.
-Thus, $T/E$ is
-the number of communications.
 

### 2020-06-05 source/main.tex

--- 2020-06-05 source/main.tex
@@
 We show that FedAvg enjoys linear speedup in each case, although with different convergence rates.
-For general convex and strongly convex problems, we also characterize the corresponding convergence rates for the Nesterov accelerated FedAvg algorithm in the FL setting: to the best of our knowledge, they are the first linear speedup guarantees for FedAvg when Nesterov acceleration is used.
-We also design a new  accelerated FL algorithm that further improves the convergence rate in overparametrized linear regression problems.
+For each class, we also characterize the corresponding convergence rates for the Nesterov accelerated FedAvg algorithm in the FL setting: to the best of our knowledge, these are the first linear speedup guarantees for FedAvg when Nesterov acceleration is used.
+To accelerate FedAvg, we also design a new momentum-based FL algorithm that further improves the convergence rate in overparametrized linear regression problems.
 Empirical studies of the algorithms in various settings have supported our theoretical results.

### 2020-06-05 source/main.tex

@@
 Federated learning (FL) learns a model jointly from a set of participating devices without sharing each other's privately held data.
-The characteristics of non-\textit{iid} data across the network, low device participation and the mandate that data remain private bring challenges in understanding the convergence of FL algorithms, particularly in regards to how convergence scales with the number of participating devices.
+The characteristics of non-\textit{iid} data across the network, low device participation, and the mandate that data remain private bring challenges in understanding the convergence of FL algorithms, particularly in regards to how convergence scales with the number of participating devices.
 In this paper, we focus on Federated Averaging (FedAvg)--the most widely used and effective FL algorithm in use today--and provide a comprehensive study of its convergence rate.

### 2020-06-05 source/main.tex
Update on Overleaf.

@@
 Although FedAvg has recently been studied by an emerging line of literature, it remains open as to how FedAvg's convergence scales with the number of participating devices in the FL setting--a crucial question whose answer would shed light on the performance of FedAvg in large FL systems.
-We fill this gap by establishing convergence guarantees for FedAvg under two classes of problems: strongly convex smooth, convex smooth, and overparametrized strongly convex smooth problems.
+We fill this gap by establishing convergence guarantees for FedAvg under three classes of problems: strongly convex smooth, convex smooth, and overparameterized strongly convex smooth problems.
 We show that FedAvg enjoys linear speedup in each case, although with different convergence rates.
 For each class, we also characterize the corresponding convergence rates for the Nesterov accelerated FedAvg algorithm in the FL setting: to the best of our knowledge, these are the first linear speedup guarantees for FedAvg when Nesterov acceleration is used.
-To accelerate FedAvg, we also design a new momentum-based FL algorithm that further improves the convergence rate in overparametrized linear regression problems.
+To accelerate FedAvg, we also design a new momentum-based FL algorithm that further improves the convergence rate in overparameterized linear regression problems.
 Empirical studies of the algorithms in various settings have supported our theoretical results.

### 2020-06-09 source/app_nasgd_scvx_smooth.tex
updated nesterov proofs

@@
-We show that FedAvg with Accelerated SGD has $O(1/T)$ rate under
-$\mu$-strong convexity and $L$-smoothness. The proof follows the
-framework of the ICLR paper. The FedAv algorithm with Nesterov Accelerated
-SGD (NASGD) follows the updates
+\begin{thm}
+	Let $\overline{\mathbf{v}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{v}_{T}^{k}$
+	and set learning rates $\alpha_{t}=\frac{6}{\mu}\frac{1}{t+\gamma}$,  $\beta_{t-1}=\frac{3}{14(t+\gamma)(1-\frac{6}{t+\gamma})\max\{\mu,1\}}$.
+Then under Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2} with full device participation,
 	\begin{align*}
-y_{t+1}^{k} & =w_{t}^{k}-\alpha_{t}g_{t,k}\\
-w_{t+1}^{k} & =\begin{cases}
-y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E}\\
-\sum_{k=1}^{N}p_{k}\left[y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E}
-\end{cases}
-\end{align*}
-where 
-\begin{align*}
-g_{t,k} & :=\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})
-\end{align*}
-is the stochastic gradient.
-
-Define the virtual sequences $\overline{y}_{t}=\sum_{k=1}^{N}p_{k}y_{t}^{k}$,
-$\overline{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$, and $\overline{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$.
-We have $\mathbb{E}g_{t}=\overline{g}_{t}$ and $\overline{y}_{t+1}=\overline{w}_{t}-\alpha_{t}g_{t}$,
-and $\overline{w}_{t+1}=\overline{y}_{t+1}+\beta_{t}(\overline{y}_{t+1}-\overline{y}_{t})$.
-
-\begin{theorem}
-	Assume that $F_{k}$ is $L$-smooth and $\mu$-strongly convex for
-	all $k$, and let $\nu_{\max}=\max_{k}\frac{1}{N}p_{k}$. Let $\kappa=\frac{L}{\mu}$,
-	$\gamma=\max\{32\kappa,E\}$ where $E$ is the communication delay,
-	and diminishing learning rates $\alpha_{t}=\frac{9}{\mu}\frac{1}{t+\gamma}$
-	and $\beta_{t-1}=\frac{9}{14\max\{\mu,1\}}\frac{1}{t+\gamma}$. Let
-	$\overline{y}_{T}=\sum_{k=1}^{N}p_{k}y_{T}^{k}$ be the average of
-	local parameters at an arbitrary time $T$.
-
-	With full device participation, 
-	\begin{align*}
-	\mathbb{E}F(\overline{y}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right),
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{y}_{T})-F^{\ast}\leq O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).
 	\end{align*}
-\end{theorem}
+\end{thm}
 %
@@
 \begin{proof}
-	The proof again uses the $L$-smoothness of $F$ to bound 
+Define the virtual sequences $\overline{v}_{t}=\sum_{k=1}^{N}p_{k}v_{t}^{k}$,
+$\overline{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$, and $\overline{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$.
+We have $\mathbb{E}g_{t}=\overline{g}_{t}$ and $\overline{v}_{t+1}=\overline{w}_{t}-\alpha_{t}g_{t}$,
+and $\overline{w}_{t+1}=\overline{v}_{t+1}$ for all $t$. The proof
+again uses the $L$-smoothness of $F$ to bound 
 \begin{align*}
-	\mathbb{E}(F(\overline{y}_{t}))-F^{\ast} & =\mathbb{E}(F(\overline{y}_{t})-F(w^{\ast}))\\
-	& \leq\frac{L}{2}\mathbb{E}\|\overline{y}_{t}-w^{\ast}\|^{2}
+\mathbb{E}(F(\overline{v}_{t}))-F^{\ast} & =\mathbb{E}(F(\overline{v}_{t})-F(w^{\ast}))\\
+& \leq\frac{L}{2}\mathbb{E}\|\overline{v}_{t}-w^{\ast}\|^{2}
 \end{align*}
-
 Our main step is to prove the bound
 \begin{align*}
-	\mathbb{E}\|\overline{y}_{t+1}-w^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\overline{y}_{t}-w^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}L\alpha_{t}^{3}G^{2}
+\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\overline{v}_{t}-w^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+20E^{2}L\alpha_{t}^{3}G^{2}
 \end{align*}
+ for appropriate step sizes $\alpha_{t},\beta_{t}$.
 
@@
 \begin{align*}
-	\|\overline{y}_{t+1}-w^{\ast}\|^{2} & =\|(\overline{w}_{t}-\alpha_{t}g_{t})-w^{\ast}\|^{2}\\
+\|\overline{v}_{t+1}-w^{\ast}\|^{2} & =\|(\overline{w}_{t}-\alpha_{t}g_{t})-w^{\ast}\|^{2}\\
 & =\|(\overline{w}_{t}-\alpha_{t}\overline{g}_{t}-w^{\ast})-\alpha_{t}(g_{t}-\overline{g}_{t})\|^{2}\\
@@
 \begin{align*}
-	\|\overline{w}_{t}-w^{\ast}\|^{2} & =\|\overline{y}_{t}+\beta_{t-1}(\overline{y}_{t}-\overline{y}_{t-1})-w^{\ast}\|^{2}\\
-	& =\|(1+\beta_{t-1})(\overline{y}_{t}-w^{\ast})-\beta_{t-1}(\overline{y}_{t-1}-w^{\ast})\|^{2}\\
-	& =(1+\beta_{t-1})^{2}\|\overline{y}_{t}-w^{\ast}\|^{2}-2\beta_{t-1}(1+\beta_{t-1})\langle\overline{y}_{t}-w^{\ast},\overline{y}_{t-1}-w^{\ast}\rangle+\beta_{t-1}^{2}\|(\overline{y}_{t-1}-w^{\ast})\|^{2}\\
-	& \leq(1+\beta_{t-1})^{2}\|\overline{y}_{t}-w^{\ast}\|^{2}+2\beta_{t-1}(1+\beta_{t-1})\|\overline{y}_{t}-w^{\ast}\|\cdot\|\overline{y}_{t-1}-w^{\ast}\|+\beta_{t-1}^{2}\|(\overline{y}_{t-1}-w^{\ast})\|^{2}
+\|\overline{w}_{t}-w^{\ast}\|^{2} & =\|\overline{v}_{t}+\beta_{t-1}(\overline{v}_{t}-\overline{v}_{t-1})-w^{\ast}\|^{2}\\
+& =\|(1+\beta_{t-1})(\overline{v}_{t}-w^{\ast})-\beta_{t-1}(\overline{v}_{t-1}-w^{\ast})\|^{2}\\
+& =(1+\beta_{t-1})^{2}\|\overline{v}_{t}-w^{\ast}\|^{2}-2\beta_{t-1}(1+\beta_{t-1})\langle\overline{v}_{t}-w^{\ast},\overline{v}_{t-1}-w^{\ast}\rangle+\beta_{t-1}^{2}\|(\overline{v}_{t-1}-w^{\ast})\|^{2}\\
+& \leq(1+\beta_{t-1})^{2}\|\overline{v}_{t}-w^{\ast}\|^{2}+2\beta_{t-1}(1+\beta_{t-1})\|\overline{v}_{t}-w^{\ast}\|\cdot\|\overline{v}_{t-1}-w^{\ast}\|+\beta_{t-1}^{2}\|(\overline{v}_{t-1}-w^{\ast})\|^{2}
 \end{align*}
@@
 \begin{align*}
-	\|\overline{y}_{t+1}-w^{\ast}\|^{2} & \leq(1-\alpha_{t}\mu)(1+\beta_{t-1})^{2}\|\overline{y}_{t}-w^{\ast}\|^{2}+2(1-\alpha_{t}\mu)\beta_{t-1}(1+\beta_{t-1})\|\overline{y}_{t}-w^{\ast}\|\cdot\|\overline{y}_{t-1}-w^{\ast}\|+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-	& +\beta_{t-1}^{2}(1-\alpha_{t}\mu)\|(\overline{y}_{t-1}-w^{\ast})\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}LG^{2}
+\|\overline{v}_{t+1}-w^{\ast}\|^{2} & \leq(1-\alpha_{t}\mu)(1+\beta_{t-1})^{2}\|\overline{v}_{t}-w^{\ast}\|^{2}+2(1-\alpha_{t}\mu)\beta_{t-1}(1+\beta_{t-1})\|\overline{v}_{t}-w^{\ast}\|\cdot\|\overline{v}_{t-1}-w^{\ast}\|+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+& +\beta_{t-1}^{2}(1-\alpha_{t}\mu)\|(\overline{v}_{t-1}-w^{\ast})\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}LG^{2}
 \end{align*}
@@
 can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $w_{t_{0}}^{k}=\overline{w}_{t_{0}}$for
-	all $k$. Moreover, using $\alpha_{t}$ is non-increasing and $\alpha_{t_{0}}\leq2\alpha{}_{t}$
-	for any $t-t_{0}\leq E-1$, we have 
+all $k$. Moreover, using $\alpha_{t}$ is non-increasing, $\alpha_{t_{0}}\leq2\alpha{}_{t}$,
+and $\beta_{t}\leq\alpha_{t}$ for any $t-t_{0}\leq E-1$, we have
 \begin{align*}
@@
 & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-w_{t_{0}}^{k}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\sum_{i=t_{0}}^{t-1}\beta_{i}(y_{i+1}^{k}-y_{i}^{k})-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}\|g_{i,k}\|^{2}+2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\beta_{i}^{2}\|(y_{i+1}^{k}-y_{i}^{k})\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}(\|g_{i,k}\|^{2}+\|(y_{i+1}^{k}-y_{i}^{k})\|^{2})\\
+& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\sum_{i=t_{0}}^{t-1}\beta_{i}(v_{i+1}^{k}-v_{i}^{k})-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\
+& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}\|g_{i,k}\|^{2}+2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\beta_{i}^{2}\|(v_{i+1}^{k}-v_{i}^{k})\|^{2}\\
+& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}(\|g_{i,k}\|^{2}+\|(v_{i+1}^{k}-v_{i}^{k})\|^{2})\\
 & \leq4\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}G^{2}\\
@@
 \end{align*}
-	where we have used $\mathbb{E}\|\overline{y}_{t+1}-\overline{y}_{t}\|^{2}\leq G^{2}$.
+where we have used $\mathbb{E}\|v_{t}^{k}-v_{t-1}^{k}\|^{2}\leq G^{2}$.
 To see this identity for appropriate $\alpha_{t},\beta_{t}$, note
@@
 \begin{align*}
-	y_{t+1}^{k}-y_{t}^{k} & =w_{t}^{k}-w_{t-1}^{k}-(\alpha_{t}g_{t,k}-\alpha_{t-1}g_{t-1,k})\\
-	w_{t+1}^{k}-w_{t}^{k} & =-\alpha_{t}g_{t,k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k})
+v_{t+1}^{k}-v_{t}^{k} & =w_{t}^{k}-w_{t-1}^{k}-(\alpha_{t}g_{t,k}-\alpha_{t-1}g_{t-1,k})\\
+w_{t+1}^{k}-w_{t}^{k} & =-\alpha_{t}g_{t,k}+\beta_{t}(v_{t+1}^{k}-v_{t}^{k})
 \end{align*}
@@
 \begin{align*}
-	y_{t+1}^{k}-y_{t}^{k} & =-\alpha_{t-1}g_{t-1,k}+\beta_{t-1}(y_{t}^{k}-y_{t-1}^{k})-(\alpha_{t}g_{t,k}-\alpha_{t-1}g_{t-1,k})\\
-	& =\beta_{t-1}(y_{t}^{k}-y_{t-1}^{k})-\alpha_{t}g_{t,k}
+v_{t+1}^{k}-v_{t}^{k} & =-\alpha_{t-1}g_{t-1,k}+\beta_{t-1}(v_{t}^{k}-v_{t-1}^{k})-(\alpha_{t}g_{t,k}-\alpha_{t-1}g_{t-1,k})\\
+& =\beta_{t-1}(v_{t}^{k}-v_{t-1}^{k})-\alpha_{t}g_{t,k}
 \end{align*}
-	Since the identity $y_{t+1}^{k}-y_{t}^{k}=\beta_{t-1}(y_{t}^{k}-y_{t-1}^{k})-\alpha_{t}g_{t,k}$
+Since the identity $v_{t+1}^{k}-v_{t}^{k}=\beta_{t-1}(v_{t}^{k}-v_{t-1}^{k})-\alpha_{t}g_{t,k}$
 implies 
 \begin{align*}
-	\mathbb{E}\|y_{t+1}^{k}-y_{t}^{k}\|^{2} & \leq2\beta_{t-1}^{2}\mathbb{E}\|y_{t}^{k}-y_{t-1}^{k}\|^{2}+2\alpha_{t}^{2}G^{2}
+\mathbb{E}\|v_{t+1}^{k}-v_{t}^{k}\|^{2} & \leq2\beta_{t-1}^{2}\mathbb{E}\|v_{t}^{k}-v_{t-1}^{k}\|^{2}+2\alpha_{t}^{2}G^{2}
 \end{align*}
-	as long as $\alpha_{t},\beta_{t}$ satisfy $2\beta_{t-1}^{2}+2\alpha_{t}^{2}\leq1$,
-	we can guarantee that $\mathbb{E}\|y_{t}^{k}-y_{t-1}^{k}\|^{2}\leq G^{2}$
-	for all $k$. This together with Jensen's inequality gives $\mathbb{E}\|\overline{y}_{t}-\overline{y}_{t-1}\|^{2}\leq G^{2}$
+as long as $\alpha_{t},\beta_{t-1}$ satisfy $2\beta_{t-1}^{2}+2\alpha_{t}^{2}\leq1/2$,
+we can guarantee that $\mathbb{E}\|v_{t}^{k}-v_{t-1}^{k}\|^{2}\leq G^{2}$
+for all $k$ by induction. This together with Jensen's inequality
+also gives $\mathbb{E}\|\overline{v}_{t}-\overline{v}_{t-1}\|^{2}\leq G^{2}$
 for all $t$.
@@
 Using the bound on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}$,
-	we can conclude that, with $\nu_{max}:=N\cdot\max_{k}p_{k}$ and $\nu_{min}:=N\cdot\min_{k}p_{k}$,
+we can conclude that, with $\nu_{\max}:=N\cdot\max_{k}p_{k}$,
 
 \begin{align*}
-	\mathbb{E}\|\overline{y}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\overline{y}_{t}-w^{\ast}\|^{2}+16E^{2}L\alpha_{t}^{3}G^{2}+16E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{3}LG^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{y}_{t-1}-w^{\ast})\|^{2}\\
-	& +\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{y}_{t}-w^{\ast}\|\cdot\|\overline{y}_{t-1}-w^{\ast}\|\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\overline{y}_{t}-w^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{y}_{t-1}-w^{\ast})\|^{2}\\
-	& +\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{y}_{t}-w^{\ast}\|\cdot\|\overline{y}_{t-1}-w^{\ast}\|
+\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\overline{v}_{t}-w^{\ast}\|^{2}+16E^{2}L\alpha_{t}^{3}G^{2}+16E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{3}LG^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{v}_{t-1}-w^{\ast})\|^{2}\\
+& +\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{v}_{t}-w^{\ast}\|\cdot\|\overline{v}_{t-1}-w^{\ast}\|\\
+& \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\overline{v}_{t}-w^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{v}_{t-1}-w^{\ast})\|^{2}\\
+& +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{v}_{t}-w^{\ast}\|\cdot\|\overline{v}_{t-1}-w^{\ast}\|
 \end{align*}
-	where $\sigma^{2}=\max_{k}\sigma_{k}^{2}$, and $\alpha_{t}$ satisfies
-	$L\alpha_{t}\leq\frac{1}{8}$. We show next that $\mathbb{E}\|\overline{y}_{t}-w^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{1}{t^{2}})$
+where $\sigma^{2}=\sum_{k}p_{k}\sigma_{k}^{2}$, and $\alpha_{t}$
+satisfies $L\alpha_{t}\leq\frac{1}{5}$. We show next that $\mathbb{E}\|\overline{v}_{t}-w^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}}{t^{2}})$
 by induction.
@@
 \end{align*}
-	where $C=20E^{2}LG^{2}$, $D=\frac{1}{N}\nu_{max}^{2}\sigma^{2}$,
-	and $b$ is to be chosen later. For step sizes we choose $\alpha_{t}=\frac{9}{\mu}\frac{1}{t+\gamma}$
-	and $\beta_{t-1}=\frac{9}{14\mu}\frac{1}{t+\gamma}$ if $\mu>1$ and
-	$\frac{9}{14}\frac{1}{t+\gamma}$ if $\mu<1$, so that $\beta_{t-1}\leq\alpha_{t}$
-	and $(1-\mu\alpha_{t})(1+14\beta_{t-1})^{2}\leq(1-\mu\alpha_{t})(1+\mu\alpha_{t})$.
-	Moreover, $\|\overline{y}_{t-1}-w^{\ast}\|\leq2\|\overline{y}_{t}-w^{\ast}\|$.
-	Recall that we require $\alpha_{t_{0}}\leq\alpha_{t}$ for any $t-t_{0}\leq E-1$,
-	and $L\alpha_{t}\leq\frac{1}{8}$.
+for all iterations until $t$, where $C=20E^{2}LG^{2}$, $D=\frac{1}{N}\nu_{max}^{2}\sigma^{2}$,
+and $b$ is to be chosen later. For step sizes we choose $\alpha_{t}=\frac{6}{\mu}\frac{1}{t+\gamma}$
+and $\beta_{t-1}=\frac{3}{14(t+\gamma)(1-\frac{6}{t+\gamma})\max\{\mu,1\}}$
+where $\gamma=\max\{32\kappa,E\}$, so that $\beta_{t-1}\leq\alpha_{t}$
+and 
+\begin{align*}
+(1-\mu\alpha_{t})(1+14\beta_{t-1}) & \leq(1-\frac{6}{t+\gamma})(1+\frac{3}{(t+\gamma)(1-\frac{6}{t+\gamma})})\\
+& =1-\frac{6}{t+\gamma}+\frac{3}{t+\gamma}=1-\frac{3}{t+\gamma}=1-\frac{\mu\alpha_{t}}{2}
+\end{align*}
+Recall that we also require $\alpha_{t_{0}}\leq2\alpha_{t}$ for any
+$t-t_{0}\leq E-1$, $L\alpha_{t}\leq\frac{1}{5}$, and $2\beta_{t-1}^{2}+2\alpha_{t}^{2}\leq1/2$,
+which we can also check to hold by definition of $\alpha_{t}$ and
+$\beta_{t}$.
 
-	Then the bound for $\mathbb{E}\|\overline{y}_{t+1}-w^{\ast}\|^{2}$
+Moreover, $\mathbb{E}\|\overline{y}_{t}-w^{\ast}\|^{2}\leq b(C\alpha_{t}^{2}+D\alpha_{t})$
+with the chosen step sizes also implies $\|\overline{v}_{t-1}-w^{\ast}\|\leq2\|\overline{v}_{t}-w^{\ast}\|$.
+Therefore the bound for $\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2}$
 can be further simplified with 
 \begin{align*}
-	2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{y}_{t}-w^{\ast}\|\cdot\|\overline{y}_{t-1}-w^{\ast}\| & \leq4\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{y}_{t}-w^{\ast}\|^{2}
+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{v}_{t}-w^{\ast}\|\cdot\|\overline{v}_{t-1}-w^{\ast}\| & \leq4\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{v}_{t}-w^{\ast}\|^{2}
 \end{align*}
@@
 \begin{align*}
-	(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{y}_{t-1}-w^{\ast})\|^{2} & \leq4(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{y}_{t}-w^{\ast})\|^{2}
+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{v}_{t-1}-w^{\ast})\|^{2} & \leq4(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{v}_{t}-w^{\ast})\|^{2}
 \end{align*}
-	Thus 
+so that
 \begin{align*}
-	\mathbb{E}\|\overline{y}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})((1+\beta_{t-1})^{2}+4\beta_{t-1}(1+\beta_{t-1})+4\beta_{t-1}^{2})\|(\overline{y}_{t}-w^{\ast})\|^{2}\\
-	& +20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})(1+14\beta_{t-1})\|(\overline{y}_{t}-w^{\ast})\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\\
-	& \leq b(1-\mu\alpha_{t})^{2}(C\alpha_{t}^{2}+D\alpha_{t})+C\alpha_{t}^{3}+D\alpha_{t}^{2}\\
-	& =(b(1-\mu\alpha_{t})^{2}+\alpha_{t})\alpha_{t}^{2}C+(b(1-\mu\alpha_{t})^{2}+\alpha_{t})\alpha_{t}D
+\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})((1+\beta_{t-1})^{2}+4\beta_{t-1}(1+\beta_{t-1})+4\beta_{t-1}^{2})\mathbb{E}\|(\overline{v}_{t}-w^{\ast})\|^{2}\\
+& +20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
+& \leq\mathbb{E}(1-\mu\alpha_{t})(1+14\beta_{t-1})\|(\overline{v}_{t}-w^{\ast})\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
+& \leq b(1-\frac{\mu\alpha_{t}}{2})(C\alpha_{t}^{2}+D\alpha_{t})+C\alpha_{t}^{3}+D\alpha_{t}^{2}\\
+& =(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}^{2}C+(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}D
 \end{align*}
@@
 \begin{align*}
-	(b(1-\mu\alpha_{t})^{2}+\alpha_{t})\alpha_{t} & \leq b\alpha_{t+1}\\
-	(b(1-\mu\alpha_{t})^{2}+\alpha_{t})\alpha_{t}^{2} & \leq b\alpha_{t+1}^{2}
+(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t} & \leq b\alpha_{t+1}\\
+(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}^{2} & \leq b\alpha_{t+1}^{2}
+\end{align*}
+from which we can conclude $\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2}\leq\alpha_{t+1}^{2}C+\alpha_{t+1}D$.
+
+With $b=\frac{6}{\mu}$, we have
+\begin{align*}
+(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t} & =(b(1-(\frac{3}{t+\gamma})+\frac{6}{\mu(t+\gamma)})\frac{6}{\mu(t+\gamma)}\\
+& =(b\frac{t+\gamma-3}{t+\gamma}+\frac{6}{\mu(t+\gamma)})\frac{6}{\mu(t+\gamma)}\\
+& \leq b(\frac{t+\gamma-1}{t+\gamma})\frac{6}{\mu(t+\gamma)}\\
+& \leq b\frac{6}{\mu(t+\gamma+1)}=b\alpha_{t+1}
+\end{align*}
+where we have used $\frac{t+\gamma-1}{(t+\gamma)^{2}}\leq\frac{1}{t+\gamma+1}$.
+
+Similarly 
+\begin{align*}
+(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}^{2} & =(b(1-(\frac{3}{t+\gamma})+\frac{6}{\mu(t+\gamma)})(\frac{6}{\mu(t+\gamma)})^{2}\\
+& =(b\frac{t+\gamma-3}{t+\gamma}+\frac{6}{\mu(t+\gamma)})(\frac{6}{\mu(t+\gamma)})^{2}\\
+& =b(\frac{t+\gamma-2}{t+\gamma})(\frac{6}{\mu(t+\gamma)})^{2}\\
+& \leq b\frac{36}{\mu^{2}(t+\gamma+1)^{2}}=b\alpha_{t+1}^{2}
+\end{align*}
+where we have used $\frac{t+\gamma-2}{(t+\gamma)^{3}}\leq\frac{1}{(t+\gamma+1)^{2}}$.
+
+Finally, to ensure $\|v_{0}-w^{\ast}\|^{2}\leq b(C\alpha_{0}^{2}+D\alpha_{0})$,
+we can rescale $b$ by $c\|v_{0}-w^{\ast}\|^{2}$ for some $c.$ It
+follows that $\mathbb{E}\|\overline{v}_{t}-w^{\ast}\|^{2}\leq b(C\alpha_{t}^{2}+D\alpha_{t})$
+for all $t$. Thus
+
+\begin{align*}
+\mathbb{E}(F(\overline{w}_{T}))-F^{\ast} & =\mathbb{E}(F(\overline{w}_{T})-F(w^{\ast}))\\
+& \leq\frac{L}{2}\mathbb{E}\|\overline{w}_{T}-w^{\ast}\|^{2}\leq\frac{L}{2}c\|w_{0}-w^{\ast}\|^{2}\frac{6}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
+& =3c\|w_{0}-w^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
+& \leq3c\|w_{0}-w^{\ast}\|^{2}\kappa\left[\frac{6}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{\max}\sigma^{2}+20E^{2}LG^{2}\cdot(\frac{6}{\mu(T+\gamma)})^{2}\right]\\
+& =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{\max}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
 \end{align*}
 
-	We have 
+With partial participation, the same argument in the SGD case yields
 \begin{align*}
-	(b(1-\mu\alpha_{t})^{2}+\alpha_{t})\alpha_{t} & =(b(1-\frac{4}{t+\gamma})^{2}+\frac{4}{\mu(t+\gamma)})\frac{4}{\mu(t+\gamma)}\\
-	& =(b\frac{(t+\gamma-4)^{2}}{(t+\gamma)^{2}}+\frac{4}{\mu(t+\gamma)})\frac{4}{\mu(t+\gamma)}\\
-	& =b(\frac{(t+\gamma-4)^{2}+t+\gamma}{(t+\gamma)^{2}})\frac{4}{\mu(t+\gamma)}\\
-	& =b\frac{4}{\mu(t+\gamma+1)}
-	\end{align*}
-	where we have used 
-	\begin{align*}
-	\frac{(t+\gamma-4)^{2}+t+\gamma}{(t+\gamma)^{2}}\frac{1}{t+\gamma} & \leq\frac{1}{t+\gamma+1}
-	\end{align*}
-	Similarly $(b(1-\mu\alpha_{t})^{2}+\alpha_{t})\alpha_{t}^{2}\leq b\alpha_{t+1}^{2}$.
-	It follows that $\mathbb{E}\|\overline{y}_{t}-w^{\ast}\|^{2}\leq b(C\alpha_{t}^{2}+D\alpha_{t})$
-	for all $t$. Thus 
-	\begin{align*}
-	\mathbb{E}(F(\overline{y}_{T}))-F^{\ast} & =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa}{\mu^{2}}E^{2}LG^{2}\cdot\frac{1}{T^{2}})
-	\end{align*}
-
-	With partial participation, the update at each communication round
-	is now given by averages over a subset of sampled devices. The key
-	is to bound 
-	\begin{align*}
-	\mathbb{E}\|\overline{w}_{t}-\overline{y}_{t}\|^{2} & =\frac{1}{K}\sum_{k}p^{k}\|w_{t}^{k}-\sum_{k}p_{k}w_{t}^{k}\|^{2}
-	\end{align*}
-	This can be bounded by the same argument as before and yield 
-	\begin{align*}
-	\mathbb{E}\|\overline{w}_{t}-\overline{y}_{t}\|^{2} & \leq\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
-	\end{align*}
-	which results in the bound 
-	\begin{align*}
-	\mathbb{E}F(\overline{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+\mathbb{E}F(\overline{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
 \end{align*}

### 2020-06-09 source/app_sumarization.tex
Update app_sumarization.tex

@@
 In this section, we summarize the convergence results in this work from various perspectives
-and compare our results with related works. In Table~\ref{tb:convergenceratev1} and Table~\ref{tb:convergenceratev2}, we summarize 
-the convergence results in terms of participations and accelerations. In Table~\ref{tb:convergenceratev3}, we compare our 
+and compare our results with related works. 
+% In Table~\ref{tb:convergenceratev1} and Table~\ref{tb:convergenceratev2}, we summarize 
+% the convergence results in terms of participations and accelerations. 
+In Table~\ref{tb:convergenceratev3}, we compare our 
 results with the best known results in the literatures.
@@
 
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|}\hline 
-	Algorithm \textbackslash\  Objective function        & Strongly Convex        & Convex \\ \hline \hline
-	FedAvg                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
-	Nesterov accelerated FedAvg    & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    & $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
-\end{tabular}
-\caption{A high level summary of the convergence results for full participation in this paper. This table only highlights the
-dependence on $T$ (number of iterations), $E$ (the largest number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices.}
-\label{tb:convergenceratev1}
-\end{table}
+% \begin{table}[h!]
+% \centering
+% \begin{tabular}{|c|c|c|}\hline 
+% 	Algorithm \textbackslash\  Objective function        & Strongly Convex        & Convex \\ \hline \hline
+% 	FedAvg                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
+% 	Nesterov accelerated FedAvg    & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    & $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
+% \end{tabular}
+% \caption{A high level summary of the convergence results for full participation in this paper. This table only highlights the
+% dependence on $T$ (number of iterations), $E$ (the largest number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices.}
+% \label{tb:convergenceratev1}
+% \end{table}
 
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|}\hline 
-	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex \\ \hline \hline
-	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
-	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$      \\ \hline
-\end{tabular}
-\caption{A high level summary of the convergence results in this paper. This table only highlights the
-dependence on $T$ (number of iterations), $E$ (the largest number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices.
-These convergence rates applied to both FedAvg and Nesterov accelerated FedAvg.}
-\label{tb:convergenceratev2}
-\end{table}
+% \begin{table}[h!]
+% \centering
+% \begin{tabular}{|c|c|c|}\hline 
+% 	Participation \textbackslash\ Objective function            & Strongly Convex        & Convex \\ \hline \hline
+% 	Full                         & $\cO(\frac{1}{NT}+\frac{E^{2}}{T^{2}})$    &  $\mathcal{O}\left(\frac{1}{\sqrt{NT}}+\frac{NE^{2}}{T}\right)$       \\ \hline
+% 	Partial                      &  $\cO\left(\frac{E^{2}}{KT}+\frac{E^{2}}{T^{2}}\right)$   &  $\cO\left(\frac{E^2}{\sqrt{KT}}+\frac{KE^2}{T} \right)$      \\ \hline
+% \end{tabular}
+% \caption{A high level summary of the convergence results in this paper. This table only highlights the
+% dependence on $T$ (number of iterations), $E$ (the largest number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices.
+% These convergence rates applied to both FedAvg and Nesterov accelerated FedAvg.}
+% \label{tb:convergenceratev2}
+% \end{table}
 

### 2020-06-09 source/app_sgd_cvx_smooth.tex

@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T}\right)
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T}\right)
 	\end{align*}
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right)
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right)
 	\end{align*}
@@
 & \leq\frac{1}{\sqrt{NT}}\cdot\|w_{0}-w^{\ast}\|^{2}+\frac{1}{\sqrt{NT}}T\cdot\frac{N}{T}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\frac{N}{T}6E^{2}LG^{2}\\
-	& =(\|w_{0}-w^{\ast}\|^{2}+\nu_{\max}^{2}\sigma^{2})\frac{1}{\sqrt{NT}}+\frac{N}{T}6E^{2}LG^{2}
+& =(\|w_{0}-w^{\ast}\|^{2}+\nu_{\max}^{2}\sigma^{2})\frac{1}{\sqrt{NT}}+\frac{N}{T}6E^{2}LG^{2}\\
+& =O(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T})
 \end{align*}
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|w_{0}-w^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot(\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+C)+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right)
+	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|w_{0}-w^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot(\frac{1}{N}\nu_{\max}\sigma^{2}+C)+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right)
 	\end{align*}
 	where $C=\frac{4}{K}E^{2}G^{2}$ or $\frac{N-K}{N-1}\frac{4}{K}E^{2}G^{2}$,
-	so that in the former case, with $\alpha_{t}=\sqrt{\frac{K}{T}}$,
-	we have 
+	so that with $\alpha_{t}=\sqrt{\frac{K}{T}}$, we have 
 	\begin{align*}
-	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =\mathcal{O}(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T})
+	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =\mathcal{O}(\frac{\nu_{\max}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T})
 	\end{align*}

### 2020-06-10 source/app_nasgd_scvx_smooth.tex
\ov{v}, \ov{w}, \vw

@@
 \begin{proof}
-Define the virtual sequences $\overline{v}_{t}=\sum_{k=1}^{N}p_{k}v_{t}^{k}$,
-$\overline{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$, and $\overline{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$.
-We have $\mathbb{E}g_{t}=\overline{g}_{t}$ and $\overline{v}_{t+1}=\overline{w}_{t}-\alpha_{t}g_{t}$,
-and $\overline{w}_{t+1}=\overline{v}_{t+1}$ for all $t$. The proof
+Define the virtual sequences $\ov{v}_{t}=\sum_{k=1}^{N}p_{k}v_{t}^{k}$,
+$\ov{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$, and $\overline{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$.
+We have $\mathbb{E}g_{t}=\overline{g}_{t}$ and $\ov{v}_{t+1}=\ov{w}_{t}-\alpha_{t}g_{t}$,
+and $\ov{w}_{t+1}=\ov{v}_{t+1}$ for all $t$. The proof
 again uses the $L$-smoothness of $F$ to bound 
 \begin{align*}
-\mathbb{E}(F(\overline{v}_{t}))-F^{\ast} & =\mathbb{E}(F(\overline{v}_{t})-F(w^{\ast}))\\
-& \leq\frac{L}{2}\mathbb{E}\|\overline{v}_{t}-w^{\ast}\|^{2}
+\mathbb{E}(F(\ov{v}_{t}))-F^{\ast} & =\mathbb{E}(F(\ov{v}_{t})-F(\vw^{\ast}))\\
+& \leq\frac{L}{2}\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}
 \end{align*}
@@
 \begin{align*}
-\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\overline{v}_{t}-w^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+20E^{2}L\alpha_{t}^{3}G^{2}
+\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+20E^{2}L\alpha_{t}^{3}G^{2}
 \end{align*}
@@
 \begin{align*}
-\|\overline{v}_{t+1}-w^{\ast}\|^{2} & =\|(\overline{w}_{t}-\alpha_{t}g_{t})-w^{\ast}\|^{2}\\
-& =\|(\overline{w}_{t}-\alpha_{t}\overline{g}_{t}-w^{\ast})-\alpha_{t}(g_{t}-\overline{g}_{t})\|^{2}\\
+\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}g_{t})-\vw^{\ast}\|^{2}\\
+& =\|(\ov{w}_{t}-\alpha_{t}\overline{g}_{t}-\vw^{\ast})-\alpha_{t}(g_{t}-\overline{g}_{t})\|^{2}\\
 & =A_{1}+A_{2}+A_{3}
@@
 \begin{align*}
-A_{1} & =\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2}\\
-A_{2} & =2\alpha_{t}\langle\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t},\overline{g}_{t}-g_{t}\rangle\\
+A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2}\\
+A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t},\overline{g}_{t}-g_{t}\rangle\\
 A_{3} & =\alpha_{t}^{2}\|g_{t}-\overline{g}_{t}\|^{2}
@@
 \begin{align*}
-\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & =\|\overline{w}_{t}-w^{\ast}\|^{2}+2\langle\overline{w}_{t}-w^{\ast},-\alpha_{t}\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2}
+\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2}
 \end{align*}
@@
 \begin{align*}
--2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2} & \leq\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}\mu\|\overline{w}_{t}-w^{\ast}\|^{2}
+-2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2} & \leq\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
 \end{align*}
@@
 \begin{align*}
-\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & \leq(1-\alpha_{t}\mu)\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}
+\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & \leq(1-\alpha_{t}\mu)\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}
 \end{align*}
@@
 \begin{align*}
-\|\overline{w}_{t}-w^{\ast}\|^{2} & =\|\overline{v}_{t}+\beta_{t-1}(\overline{v}_{t}-\overline{v}_{t-1})-w^{\ast}\|^{2}\\
-& =\|(1+\beta_{t-1})(\overline{v}_{t}-w^{\ast})-\beta_{t-1}(\overline{v}_{t-1}-w^{\ast})\|^{2}\\
-& =(1+\beta_{t-1})^{2}\|\overline{v}_{t}-w^{\ast}\|^{2}-2\beta_{t-1}(1+\beta_{t-1})\langle\overline{v}_{t}-w^{\ast},\overline{v}_{t-1}-w^{\ast}\rangle+\beta_{t-1}^{2}\|(\overline{v}_{t-1}-w^{\ast})\|^{2}\\
-& \leq(1+\beta_{t-1})^{2}\|\overline{v}_{t}-w^{\ast}\|^{2}+2\beta_{t-1}(1+\beta_{t-1})\|\overline{v}_{t}-w^{\ast}\|\cdot\|\overline{v}_{t-1}-w^{\ast}\|+\beta_{t-1}^{2}\|(\overline{v}_{t-1}-w^{\ast})\|^{2}
+\|\ov{w}_{t}-\vw^{\ast}\|^{2} & =\|\ov{v}_{t}+\beta_{t-1}(\ov{v}_{t}-\ov{v}_{t-1})-\vw^{\ast}\|^{2}\\
+& =\|(1+\beta_{t-1})(\ov{v}_{t}-\vw^{\ast})-\beta_{t-1}(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
+& =(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}-2\beta_{t-1}(1+\beta_{t-1})\langle\ov{v}_{t}-\vw^{\ast},\ov{v}_{t-1}-\vw^{\ast}\rangle+\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
+& \leq(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+2\beta_{t-1}(1+\beta_{t-1})\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|+\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}
 \end{align*}
@@
 \begin{align*}
-\|\overline{v}_{t+1}-w^{\ast}\|^{2} & \leq(1-\alpha_{t}\mu)(1+\beta_{t-1})^{2}\|\overline{v}_{t}-w^{\ast}\|^{2}+2(1-\alpha_{t}\mu)\beta_{t-1}(1+\beta_{t-1})\|\overline{v}_{t}-w^{\ast}\|\cdot\|\overline{v}_{t-1}-w^{\ast}\|+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-& +\beta_{t-1}^{2}(1-\alpha_{t}\mu)\|(\overline{v}_{t-1}-w^{\ast})\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}LG^{2}
+\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\alpha_{t}\mu)(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+2(1-\alpha_{t}\mu)\beta_{t-1}(1+\beta_{t-1})\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+& +\beta_{t-1}^{2}(1-\alpha_{t}\mu)\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}LG^{2}
 \end{align*}
@@
 
-First we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}$.
+First we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$.
 Since communication is done every $E$ steps, for any $t\geq0$, we
-can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $w_{t_{0}}^{k}=\overline{w}_{t_{0}}$for
+can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $w_{t_{0}}^{k}=\ov{w}_{t_{0}}$for
 all $k$. Moreover, using $\alpha_{t}$ is non-increasing, $\alpha_{t_{0}}\leq2\alpha{}_{t}$,
@@
 \begin{align*}
-\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2} & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}-(\overline{w}_{t}-\overline{w}_{t_{0}})\|^{2}\\
-& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}\|^{2}\\
-& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-w_{t_{0}}^{k}\|^{2}\\
+\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
+& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
+& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
 & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\sum_{i=t_{0}}^{t-1}\beta_{i}(v_{i+1}^{k}-v_{i}^{k})-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\
@@
 \begin{align*}
-v_{t+1}^{k}-v_{t}^{k} & =w_{t}^{k}-w_{t-1}^{k}-(\alpha_{t}g_{t,k}-\alpha_{t-1}g_{t-1,k})\\
-w_{t+1}^{k}-w_{t}^{k} & =-\alpha_{t}g_{t,k}+\beta_{t}(v_{t+1}^{k}-v_{t}^{k})
+v_{t+1}^{k}-v_{t}^{k} & =w_{t}^{k}-\vw_{t-1}^{k}-(\alpha_{t}g_{t,k}-\alpha_{t-1}g_{t-1,k})\\
+w_{t+1}^{k}-\vw_{t}^{k} & =-\alpha_{t}g_{t,k}+\beta_{t}(v_{t+1}^{k}-v_{t}^{k})
 \end{align*}
@@
 for all $k$ by induction. This together with Jensen's inequality
-also gives $\mathbb{E}\|\overline{v}_{t}-\overline{v}_{t-1}\|^{2}\leq G^{2}$
+also gives $\mathbb{E}\|\ov{v}_{t}-\ov{v}_{t-1}\|^{2}\leq G^{2}$
 for all $t$.
 
-Using the bound on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}$,
+Using the bound on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$,
 we can conclude that, with $\nu_{\max}:=N\cdot\max_{k}p_{k}$,
@@
 \begin{align*}
-\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\overline{v}_{t}-w^{\ast}\|^{2}+16E^{2}L\alpha_{t}^{3}G^{2}+16E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{3}LG^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{v}_{t-1}-w^{\ast})\|^{2}\\
-& +\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{v}_{t}-w^{\ast}\|\cdot\|\overline{v}_{t-1}-w^{\ast}\|\\
-& \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\overline{v}_{t}-w^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{v}_{t-1}-w^{\ast})\|^{2}\\
-& +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{v}_{t}-w^{\ast}\|\cdot\|\overline{v}_{t-1}-w^{\ast}\|
+\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+16E^{2}L\alpha_{t}^{3}G^{2}+16E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{3}LG^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
+& +\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|\\
+& \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
+& +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|
 \end{align*}
 where $\sigma^{2}=\sum_{k}p_{k}\sigma_{k}^{2}$, and $\alpha_{t}$
-satisfies $L\alpha_{t}\leq\frac{1}{5}$. We show next that $\mathbb{E}\|\overline{v}_{t}-w^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}}{t^{2}})$
+satisfies $L\alpha_{t}\leq\frac{1}{5}$. We show next that $\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}}{t^{2}})$
 by induction.
@@
 \begin{align*}
-\mathbb{E}\|\overline{y}_{t}-w^{\ast}\|^{2} & \leq b(C\alpha_{t}^{2}+D\alpha_{t})
+\mathbb{E}\|\overline{y}_{t}-\vw^{\ast}\|^{2} & \leq b(C\alpha_{t}^{2}+D\alpha_{t})
 \end{align*}
@@
 
-Moreover, $\mathbb{E}\|\overline{y}_{t}-w^{\ast}\|^{2}\leq b(C\alpha_{t}^{2}+D\alpha_{t})$
-with the chosen step sizes also implies $\|\overline{v}_{t-1}-w^{\ast}\|\leq2\|\overline{v}_{t}-w^{\ast}\|$.
-Therefore the bound for $\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2}$
+Moreover, $\mathbb{E}\|\overline{y}_{t}-\vw^{\ast}\|^{2}\leq b(C\alpha_{t}^{2}+D\alpha_{t})$
+with the chosen step sizes also implies $\|\ov{v}_{t-1}-\vw^{\ast}\|\leq2\|\ov{v}_{t}-\vw^{\ast}\|$.
+Therefore the bound for $\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}$
 can be further simplified with 
 \begin{align*}
-2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{v}_{t}-w^{\ast}\|\cdot\|\overline{v}_{t-1}-w^{\ast}\| & \leq4\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\overline{v}_{t}-w^{\ast}\|^{2}
+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\| & \leq4\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|^{2}
 \end{align*}
@@
 \begin{align*}
-(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{v}_{t-1}-w^{\ast})\|^{2} & \leq4(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\overline{v}_{t}-w^{\ast})\|^{2}
+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2} & \leq4(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t}-\vw^{\ast})\|^{2}
 \end{align*}
@@
 \begin{align*}
-\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})((1+\beta_{t-1})^{2}+4\beta_{t-1}(1+\beta_{t-1})+4\beta_{t-1}^{2})\mathbb{E}\|(\overline{v}_{t}-w^{\ast})\|^{2}\\
+\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})((1+\beta_{t-1})^{2}+4\beta_{t-1}(1+\beta_{t-1})+4\beta_{t-1}^{2})\mathbb{E}\|(\ov{v}_{t}-\vw^{\ast})\|^{2}\\
 & +20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
-& \leq\mathbb{E}(1-\mu\alpha_{t})(1+14\beta_{t-1})\|(\overline{v}_{t}-w^{\ast})\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
+& \leq\mathbb{E}(1-\mu\alpha_{t})(1+14\beta_{t-1})\|(\ov{v}_{t}-\vw^{\ast})\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
 & \leq b(1-\frac{\mu\alpha_{t}}{2})(C\alpha_{t}^{2}+D\alpha_{t})+C\alpha_{t}^{3}+D\alpha_{t}^{2}\\
@@
 \end{align*}
-from which we can conclude $\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2}\leq\alpha_{t+1}^{2}C+\alpha_{t+1}D$.
+from which we can conclude $\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}\leq\alpha_{t+1}^{2}C+\alpha_{t+1}D$.
 
@@
 
-Finally, to ensure $\|v_{0}-w^{\ast}\|^{2}\leq b(C\alpha_{0}^{2}+D\alpha_{0})$,
-we can rescale $b$ by $c\|v_{0}-w^{\ast}\|^{2}$ for some $c.$ It
-follows that $\mathbb{E}\|\overline{v}_{t}-w^{\ast}\|^{2}\leq b(C\alpha_{t}^{2}+D\alpha_{t})$
+Finally, to ensure $\|v_{0}-\vw^{\ast}\|^{2}\leq b(C\alpha_{0}^{2}+D\alpha_{0})$,
+we can rescale $b$ by $c\|v_{0}-\vw^{\ast}\|^{2}$ for some $c.$ It
+follows that $\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}\leq b(C\alpha_{t}^{2}+D\alpha_{t})$
 for all $t$. Thus
@@
 \begin{align*}
-\mathbb{E}(F(\overline{w}_{T}))-F^{\ast} & =\mathbb{E}(F(\overline{w}_{T})-F(w^{\ast}))\\
-& \leq\frac{L}{2}\mathbb{E}\|\overline{w}_{T}-w^{\ast}\|^{2}\leq\frac{L}{2}c\|w_{0}-w^{\ast}\|^{2}\frac{6}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
-& =3c\|w_{0}-w^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
-& \leq3c\|w_{0}-w^{\ast}\|^{2}\kappa\left[\frac{6}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{\max}\sigma^{2}+20E^{2}LG^{2}\cdot(\frac{6}{\mu(T+\gamma)})^{2}\right]\\
+\mathbb{E}(F(\ov{w}_{T}))-F^{\ast} & =\mathbb{E}(F(\ov{w}_{T})-F(\vw^{\ast}))\\
+& \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}\leq\frac{L}{2}c\|w_{0}-\vw^{\ast}\|^{2}\frac{6}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
+& =3c\|w_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
+& \leq3c\|w_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{6}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{\max}\sigma^{2}+20E^{2}LG^{2}\cdot(\frac{6}{\mu(T+\gamma)})^{2}\right]\\
 & =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{\max}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
@@
 \begin{align*}
-\mathbb{E}F(\overline{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+\mathbb{E}F(\ov{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
 \end{align*}

### 2020-06-10 source/app_sgd_cvx_smooth.tex
\ov{v}, \ov{w}, \vw

@@
 	\begin{align*}
-	\|\overline{w}_{t+1}-w^{\ast}\|^{2} & =\|(\overline{w}_{t}-\alpha_{t}g_{t})-w^{\ast}\|^{2}\\
-	& =\|(\overline{w}_{t}-\alpha_{t}\overline{g}_{t}-w^{\ast})-\alpha_{t}(g_{t}-\overline{g}_{t})\|^{2}\\
+	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}g_{t})-\vw^{\ast}\|^{2}\\
+	& =\|(\ov{w}_{t}-\alpha_{t}\overline{g}_{t}-\vw^{\ast})-\alpha_{t}(g_{t}-\overline{g}_{t})\|^{2}\\
 	& =A_{1}+A_{2}+A_{3}
@@
 	\begin{align*}
-	A_{1} & =\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2}\\
-	A_{2} & =2\alpha_{t}\langle\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t},\overline{g}_{t}-g_{t}\rangle\\
+	A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2}\\
+	A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t},\overline{g}_{t}-g_{t}\rangle\\
 	A_{3} & =\alpha_{t}^{2}\|g_{t}-\overline{g}_{t}\|^{2}
@@
 	\begin{align*}
-	\|\overline{w}_{t}-w^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & =\|\overline{w}_{t}-w^{\ast}\|^{2}+2\langle\overline{w}_{t}-w^{\ast},-\alpha_{t}\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2}
+	\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2}
 	\end{align*}
@@
 	\begin{align*}
-	& -2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
-	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(w^{\ast})-F_{k}(w_{t}^{k}))\\
-	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w_{t}^{k})-F_{k}(\overline{w}_{t})+\frac{L}{2}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+F_{k}(w^{\ast})-F_{k}(w_{t}^{k})\right]\\
-	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]
+	& -2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\overline{g}_{t}\rangle\\
+	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))\\
+	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]\\
+	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]
 	\end{align*}
@@
 	\begin{align*}
-	\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]+\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
 	\end{align*}
@@
 	The difference of this bound with that in the strongly convex case
-	is that we no longer have a contraction factor in front of $\|\overline{w}_{t}-w^{\ast}\|^{2}$.
+	is that we no longer have a contraction factor in front of $\|\ov{w}_{t}-\vw^{\ast}\|^{2}$.
 	In the strongly convex case, we were able to cancel $\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}$
-	with $2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]$
+	with $2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]$
 	and obtain only lower order terms. In the convex case, we use a different
-	strategy and preserve $\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]$
+	strategy and preserve $\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]$
 	in order to obtain a telescoping sum.
@@
 	\begin{align*}
-	\|\overline{g}_{t}\|^{2} & =\|\sum_{k}p_{k}\nabla F_{k}(w_{t}^{k})\|^{2}\\
-	& =\|\sum_{k}p_{k}\nabla F_{k}(w_{t}^{k})-\sum_{k}p_{k}\nabla F_{k}(\overline{w}_{t})+\sum_{k}p_{k}\nabla F_{k}(\overline{w}_{t})\|^{2}\\
-	& \leq2\|\sum_{k}p_{k}\nabla F_{k}(w_{t}^{k})-\sum_{k}p_{k}\nabla F_{k}(\overline{w}_{t})\|^{2}+2\|\sum_{k}p_{k}\nabla F_{k}(\overline{w}_{t})\|^{2}\\
-	& \leq2L^{2}\sum_{k}p_{k}\|w_{t}^{k}-\overline{w}_{t}\|^{2}+2\|\sum_{k}p_{k}\nabla F_{k}(\overline{w}_{t})\|^{2}\\
-	& =2L^{2}\sum_{k}p_{k}\|w_{t}^{k}-\overline{w}_{t}\|^{2}+2\|\nabla F(\overline{w}_{t})\|^{2}
+	\|\overline{g}_{t}\|^{2} & =\|\sum_{k}p_{k}\nabla F_{k}(\vw_{t}^{k})\|^{2}\\
+	& =\|\sum_{k}p_{k}\nabla F_{k}(\vw_{t}^{k})-\sum_{k}p_{k}\nabla F_{k}(\ov{w}_{t})+\sum_{k}p_{k}\nabla F_{k}(\ov{w}_{t})\|^{2}\\
+	& \leq2\|\sum_{k}p_{k}\nabla F_{k}(\vw_{t}^{k})-\sum_{k}p_{k}\nabla F_{k}(\ov{w}_{t})\|^{2}+2\|\sum_{k}p_{k}\nabla F_{k}(\ov{w}_{t})\|^{2}\\
+	& \leq2L^{2}\sum_{k}p_{k}\|w_{t}^{k}-\ov{w}_{t}\|^{2}+2\|\sum_{k}p_{k}\nabla F_{k}(\ov{w}_{t})\|^{2}\\
+	& =2L^{2}\sum_{k}p_{k}\|w_{t}^{k}-\ov{w}_{t}\|^{2}+2\|\nabla F(\ov{w}_{t})\|^{2}
 	\end{align*}
-	using $\nabla F(w^{\ast})=0$. Now using the $L$ smoothness of $F$,
-	we have $\|\nabla F(\overline{w}_{t})\|^{2}\leq2L(F(\overline{w}_{t})-F(w^{\ast}))$,
+	using $\nabla F(\vw^{\ast})=0$. Now using the $L$ smoothness of $F$,
+	we have $\|\nabla F(\ov{w}_{t})\|^{2}\leq2L(F(\ov{w}_{t})-F(\vw^{\ast}))$,
 	so that 
 	\begin{align*}
-	& \|\overline{w}_{t+1}-w^{\ast}\|^{2}\\
-	& \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]\\
-	& +2\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|w_{t}^{k}-\overline{w}_{t}\|^{2}+4\alpha_{t}^{2}L(F(\overline{w}_{t})-F(w^{\ast}))+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-	& =\|\overline{w}_{t}-w^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-	& +\alpha_{t}(1-4\alpha_{t}L)(F(w^{\ast})-F(\overline{w}_{t}))
+	& \|\ov{w}_{t+1}-\vw^{\ast}\|^{2}\\
+	& \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]\\
+	& +2\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|w_{t}^{k}-\ov{w}_{t}\|^{2}+4\alpha_{t}^{2}L(F(\ov{w}_{t})-F(\vw^{\ast}))+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+	& =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+	& +\alpha_{t}(1-4\alpha_{t}L)(F(\vw^{\ast})-F(\ov{w}_{t}))
 	\end{align*}
-	Since $F(w^{\ast})\leq F(\overline{w}_{t})$, as long as $4\alpha_{t}L\leq1$,
+	Since $F(\vw^{\ast})\leq F(\ov{w}_{t})$, as long as $4\alpha_{t}L\leq1$,
 	we can ignore the last term, and rearrange the inequality to obtain
 	\begin{align*}
-	& \|\overline{w}_{t+1}-w^{\ast}\|^{2}+\alpha_{t}(F(\overline{w}_{t})-F(w^{\ast}))\\
-	& \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-	& \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+\frac{3}{2}\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+	& \|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast}))\\
+	& \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+	& \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\frac{3}{2}\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
 	\end{align*}
 
-	The same argument as before yields $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}\leq4E^{2}\alpha_{t}^{2}G^{2}$
+	The same argument as before yields $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\leq4E^{2}\alpha_{t}^{2}G^{2}$
 	which gives 
 	\begin{align*}
-	\|\overline{w}_{t+1}-w^{\ast}\|^{2}+\alpha_{t}(F(\overline{w}_{t})-F(w^{\ast})) & \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+6\alpha_{t}^{3}E^{2}LG^{2}\\
-	& \leq\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
+	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+6\alpha_{t}^{3}E^{2}LG^{2}\\
+	& \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\sum_{t=0}^{T}\alpha_{t}(F(\overline{w}_{t})-F(w^{\ast})) & \leq\|w_{0}-w^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}
+	\sum_{t=0}^{T}\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq\|w_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|w_{0}-w^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right)
+	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|w_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right)
 	\end{align*}
@@
 \begin{align*}
-\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & \leq\frac{1}{\sqrt{NT}}\cdot\|w_{0}-w^{\ast}\|^{2}+\frac{1}{\sqrt{NT}}T\cdot\frac{N}{T}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\frac{1}{\sqrt{NT}}T(\sqrt{\frac{N}{T}})^{3}6E^{2}LG^{2}\\
-& \leq\frac{1}{\sqrt{NT}}\cdot\|w_{0}-w^{\ast}\|^{2}+\frac{1}{\sqrt{NT}}T\cdot\frac{N}{T}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\frac{N}{T}6E^{2}LG^{2}\\
-& =(\|w_{0}-w^{\ast}\|^{2}+\nu_{\max}^{2}\sigma^{2})\frac{1}{\sqrt{NT}}+\frac{N}{T}6E^{2}LG^{2}\\
+\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sqrt{NT}}\cdot\|w_{0}-\vw^{\ast}\|^{2}+\frac{1}{\sqrt{NT}}T\cdot\frac{N}{T}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\frac{1}{\sqrt{NT}}T(\sqrt{\frac{N}{T}})^{3}6E^{2}LG^{2}\\
+& \leq\frac{1}{\sqrt{NT}}\cdot\|w_{0}-\vw^{\ast}\|^{2}+\frac{1}{\sqrt{NT}}T\cdot\frac{N}{T}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\frac{N}{T}6E^{2}LG^{2}\\
+& =(\|w_{0}-\vw^{\ast}\|^{2}+\nu_{\max}^{2}\sigma^{2})\frac{1}{\sqrt{NT}}+\frac{N}{T}6E^{2}LG^{2}\\
 & =O(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T})
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|w_{0}-w^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot(\frac{1}{N}\nu_{\max}\sigma^{2}+C)+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right)
+	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|w_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot(\frac{1}{N}\nu_{\max}\sigma^{2}+C)+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right)
 	\end{align*}
@@
 	\begin{align*}
-	\min_{t\leq T}F(\overline{w}_{t})-F(w^{\ast}) & =\mathcal{O}(\frac{\nu_{\max}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T})
+	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & =\mathcal{O}(\frac{\nu_{\max}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T})
 	\end{align*}

### 2020-06-10 source/app_sgd_scvx_smooth.tex
\ov{v}, \ov{w}, \vw

@@
 
-	1. Quadratic upper bound: $0\leq F(w)-F(w')-\langle\nabla F(w'),w-w'\rangle\leq\frac{L}{2}\|w-w'\|^{2}$.
+	1. Quadratic upper bound: $0\leq F(\vw)-F(\vw')-\langle\nabla F(\vw'),\vw-\vw'\rangle\leq\frac{L}{2}\|\vw-\vw'\|^{2}$.
 
-	2. Coercivity: $\frac{1}{L}\|\nabla F(w)-\nabla F(w')\|^{2}\leq\langle\nabla F(w)-\nabla F(w'),w-w'\rangle$.
+	2. Coercivity: $\frac{1}{L}\|\nabla F(\vw)-\nabla F(\vw')\|^{2}\leq\langle\nabla F(\vw)-\nabla F(\vw'),\vw-\vw'\rangle$.
 
-	3. Lower bound: $F(w)\geq F(w')+\langle\nabla F(w'),w-w'\rangle+\frac{1}{2L}\|\nabla F(w)-\nabla F(w')\|^{2}$.
-	In particular, $\|\nabla F(w)\|^{2}\leq2L(F(w)-F(w^{\ast}))$.
+	3. Lower bound: $F(\vw)\geq F(\vw')+\langle\nabla F(\vw'),\vw-\vw'\rangle+\frac{1}{2L}\|\nabla F(\vw)-\nabla F(\vw')\|^{2}$.
+	In particular, $\|\nabla F(\vw)\|^{2}\leq2L(F(\vw)-F(\vw^{\ast}))$.
 
-	4. Optimality gap: $F(w)-F(w^{\ast})\leq$$\langle\nabla F(w),w-w^{\ast}\rangle$.
+	4. Optimality gap: $F(\vw)-F(\vw^{\ast})\leq$$\langle\nabla F(\vw),\vw-\vw^{\ast}\rangle$.
 \end{lemma}
@@
 	\begin{align*}
-	F(w) & \leq F(w')+\langle\nabla F(w'),w-w'\rangle+\frac{1}{2\mu}\|\nabla F(w)-\nabla F(w')\|^{2}\\
-	F(w)-F(w^{\ast}) & \leq\frac{1}{2\mu}\|\nabla F(w)\|^{2}
+	F(\vw) & \leq F(\vw')+\langle\nabla F(\vw'),\vw-\vw'\rangle+\frac{1}{2\mu}\|\nabla F(\vw)-\nabla F(\vw')\|^{2}\\
+	F(\vw)-F(\vw^{\ast}) & \leq\frac{1}{2\mu}\|\nabla F(\vw)\|^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}(F(\overline{w}_{t}))-F^{\ast} & =\mathbb{E}(F(\overline{w}_{t})-F(w^{\ast}))\\
-	& \leq\frac{L}{2}\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}
+	\mathbb{E}(F(\ov{w}_{t}))-F^{\ast} & =\mathbb{E}(F(\ov{w}_{t})-F(\vw^{\ast}))\\
+	& \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}
 	\end{align*}
-	and bound $\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}$.
+	and bound $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$.
 
@@
 	\begin{align*}
-	\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}L\alpha_{t}^{3}G^{2}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}L\alpha_{t}^{3}G^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\alpha_{t}^{2}\|\overline{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(w_{t}^{k})\|^{2}\leq2L\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}(F_{k}(w_{t}^{k})-F_{k}^{\ast})
+	\alpha_{t}^{2}\|\overline{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq2L\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}(F_{k}(\vw_{t}^{k})-F_{k}^{\ast})
 	\end{align*}
@@
 	\begin{align*}
-	\alpha_{t}^{2}\|\overline{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(w_{t}^{k})\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\mathbb{E}\|\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})\|^{2}\leq\alpha_{t}^{2}G^{2}
+	\alpha_{t}^{2}\|\overline{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\mathbb{E}\|\nabla F_{k}(\vw_{t}^{k},\xi_{t}^{k})\|^{2}\leq\alpha_{t}^{2}G^{2}
 	\end{align*}
@@
 	\begin{align*}
-	& -2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
-	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(w^{\ast})-F_{k}(w_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-w^{\ast}\|^{2}\\
-	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w_{t}^{k})-F_{k}(\overline{w}_{t})+\frac{L}{2}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+F_{k}(w^{\ast})-F_{k}(w_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}w_{t}^{k}-w^{\ast}\|^{2}\\
-	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]-\alpha_{t}\mu\|\overline{w}_{t}-w^{\ast}\|^{2}
+	& -2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\overline{g}_{t}\rangle\\
+	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\vw^{\ast}\|^{2}\\
+	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}w_{t}^{k}-\vw^{\ast}\|^{2}\\
+	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
 	\end{align*}
@@
 	\begin{align*}
-	& 2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]\\
-	& \leq2\alpha_{t}\left[F(\overline{w}_{t+1})-F(\overline{w}_{t})\right]\\
-	& \leq2\alpha_{t}\mathbb{E}\langle\nabla F(\overline{w}_{t}),\overline{w}_{t+1}-\overline{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\overline{w}_{t+1}-\overline{w}_{t}\|^{2}\\
-	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\overline{w}_{t}),g_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\overline{w}_{t}),\overline{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& =-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\|\nabla F(\overline{w}_{t})-\overline{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& =-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\|\nabla F(\overline{w}_{t})-\sum_{k}p_{k}\nabla F(w_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\overline{w}_{t})-\nabla F(w_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\overline{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\overline{w}_{t})\|^{2}
+	& 2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]\\
+	& \leq2\alpha_{t}\left[F(\ov{w}_{t+1})-F(\ov{w}_{t})\right]\\
+	& \leq2\alpha_{t}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{w}_{t+1}-\ov{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\ov{w}_{t+1}-\ov{w}_{t}\|^{2}\\
+	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),g_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
+	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\overline{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
+	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\overline{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
+	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\sum_{k}p_{k}\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
+	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\ov{w}_{t})-\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
+	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
+	& \leq-\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
 	\end{align*}
@@
 	cancels the $\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}$ in the bound
-	for $A_{1}$, so that plugging in the bound for $-2\alpha_{t}\langle\overline{w}_{t}-w^{\ast},\overline{g}_{t}\rangle$,
+	for $A_{1}$, so that plugging in the bound for $-2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\overline{g}_{t}\rangle$,
 	we have so far proved 
 	\begin{align*}
-	\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\overline{w}_{t})\|^{2}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
 	\end{align*}
@@
 
-	Now we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}$.
+	Now we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$.
 	Since communication is done every $E$ steps, for any $t\geq0$, we
-	can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $w_{t_{0}}^{k}=\overline{w}_{t_{0}}$for
+	can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $w_{t_{0}}^{k}=\ov{w}_{t_{0}}$for
 	all $k$. Moreover, using $\alpha_{t}$ is non-increasing and $\alpha_{t_{0}}\leq2\alpha{}_{t}$
@@
 	\begin{align*}
-	& \mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}-(\overline{w}_{t}-\overline{w}_{t_{0}})\|^{2}\\
-	& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\overline{w}_{t_{0}}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-w_{t_{0}}^{k}\|^{2}\\
+	& \mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\\
+	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
+	& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
+	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
 	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\
@@
 	\begin{align*}
-	\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2} & =\sum_{k=1}^{N}p_{k}\|\overline{w}_{t-1}-\alpha_{t-1}g_{t-1}-w_{t-1}^{k}+\alpha_{t-1}g_{t-1,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}\left(\|\overline{w}_{t-1}-w_{t-1}^{k}\|^{2}+\|\alpha_{t-1}g_{t-1}-\alpha_{t-1}g_{t-1,k}\|^{2}\right)
+	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & =\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\alpha_{t-1}g_{t-1}-\vw_{t-1}^{k}+\alpha_{t-1}g_{t-1,k}\|^{2}\\
+	& \leq2\sum_{k=1}^{N}p_{k}\left(\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\alpha_{t-1}g_{t-1}-\alpha_{t-1}g_{t-1,k}\|^{2}\right)
 	\end{align*}
@@
 	\begin{align*}
-	\sum_{k}p_{k}\|g_{t-1,k}-g_{t-1}\|^{2}\leq\sum_{k}p_{k}\|g_{t-1,k}\|^{2} & =\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(w_{t-1}^{k})+\nabla F_{k}(w_{t-1}^{k})\|^{2}\\
-	& \leq2\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(w_{t-1}^{k})\|^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(w_{t-1}^{k})\|^{2}\\
-	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(w_{t-1}^{k})-\nabla F_{k}(\overline{w}_{t-1})+\nabla F_{k}(\overline{w}_{t-1})\|^{2}\\
-	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+4\sum_{k}p_{k}\left(\|\nabla F_{k}(w_{t-1}^{k})-\nabla F_{k}(\overline{w}_{t-1})\|^{2}+\|\nabla F_{k}(\overline{w}_{t-1})\|^{2}\right)\\
-	& \leq2\sigma^{2}+4\sum_{k}p_{k}\left(L^{2}\|\overline{w}_{t-1}-w_{t-1}^{k}\|^{2}+\|\nabla F_{k}(\overline{w}_{t-1})\|^{2}\right)
+	\sum_{k}p_{k}\|g_{t-1,k}-g_{t-1}\|^{2}\leq\sum_{k}p_{k}\|g_{t-1,k}\|^{2} & =\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})+\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
+	& \leq2\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})\|^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
+	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})-\nabla F_{k}(\ov{w}_{t-1})+\nabla F_{k}(\ov{w}_{t-1})\|^{2}\\
+	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+4\sum_{k}p_{k}\left(\|\nabla F_{k}(\vw_{t-1}^{k})-\nabla F_{k}(\ov{w}_{t-1})\|^{2}+\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}\right)\\
+	& \leq2\sigma^{2}+4\sum_{k}p_{k}\left(L^{2}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}\right)
 	\end{align*}
@@
 	\begin{align*}
-	\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\sum_{k=1}^{N}p_{k}\|\overline{w}_{t-1}-w_{t-1}^{k}\|^{2}+4\sigma^{2}\alpha_{t-1}^{2}+8\alpha_{t-1}^{2}\sum_{k}p_{k}\|\nabla F_{k}(\overline{w}_{t-1})\|^{2}
+	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+4\sigma^{2}\alpha_{t-1}^{2}+8\alpha_{t-1}^{2}\sum_{k}p_{k}\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t-1}-w_{t-1}^{k}\|^{2}+4\alpha_{t-1}^{2}(\sigma^{2}+2G^{2})
+	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+4\alpha_{t-1}^{2}(\sigma^{2}+2G^{2})
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}\leq E\frac{1}{L^{2}}(\sigma^{2}+2G^{2})
+	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\leq E\frac{1}{L^{2}}(\sigma^{2}+2G^{2})
 	\end{align*}
@@
 
-	Using the bound on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}$,
+	Using the bound on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$,
 	we can conclude that, with $\nu_{max}:=N\cdot\max_{k}p_{k}$ and $\nu_{min}:=N\cdot\min_{k}p_{k}$,
@@
 	\begin{align*}
-	& \mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& =\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\sum_{k=1}^{N}(p_{k}N)^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\nu_{max}^{2}\sum_{k=1}^{N}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
+	& \mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}\\
+	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	& =\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\sum_{k=1}^{N}(p_{k}N)^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\nu_{max}^{2}\sum_{k=1}^{N}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
 	\end{align*}
 	where $\sigma^{2}=\max_{k}\sigma_{k}^{2}$, and if $\alpha_{t}$
-	satisfies $L\alpha_{t}\leq\frac{1}{8}$. We show next that $\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}LG^{2}}{t^{2}})$.
+	satisfies $L\alpha_{t}\leq\frac{1}{8}$. We show next that $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}LG^{2}}{t^{2}})$.
 
 	Let $C\equiv5E^{2}LG^{2}$ and $D\equiv\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
-	Suppose that we have shown $\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}\leq b\cdot(\alpha_{t}D+\alpha_{t}^{2}C)$
+	Suppose that we have shown $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}\leq b\cdot(\alpha_{t}D+\alpha_{t}^{2}C)$
 	for some constant $b$ and $\alpha_{t}$. Then 
 	\begin{align*}
-	\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq b(1-\mu\alpha_{t})(\alpha_{t}D+\alpha_{t}^{2}C)+\alpha_{t}^{2}D+\alpha_{t}^{3}C\\
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq b(1-\mu\alpha_{t})(\alpha_{t}D+\alpha_{t}^{2}C)+\alpha_{t}^{2}D+\alpha_{t}^{3}C\\
 	& =(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}D+(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}C
@@
 	\begin{align*}
-	\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq b\cdot(\alpha_{t+1}D+\alpha_{t+1}^{2}C)
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq b\cdot(\alpha_{t+1}D+\alpha_{t+1}^{2}C)
 	\end{align*}
@@
 	\begin{align*}
-	\|w_{0}-w^{\ast}\|^{2} & \leq b\cdot(\alpha_{0}D+\alpha_{0}^{2}C)\\
+	\|w_{0}-\vw^{\ast}\|^{2} & \leq b\cdot(\alpha_{0}D+\alpha_{0}^{2}C)\\
 	& =b\cdot(\frac{4}{\mu\gamma}D+\frac{16}{\mu^{2}\gamma^{2}}C)
 	\end{align*}
-	we can simply scale $b$ by $c\|w_{0}-w^{\ast}\|^{2}$ for a constant
+	we can simply scale $b$ by $c\|w_{0}-\vw^{\ast}\|^{2}$ for a constant
 	$c$ large enough and the induction step still holds.
@@
 	\begin{align*}
-	\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2} & \leq c\|w_{0}-w^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{t}+C\alpha_{t}^{2})
+	\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2} & \leq c\|w_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{t}+C\alpha_{t}^{2})
 	\end{align*}
@@
 	\begin{align*}
-	& \mathbb{E}(F(\overline{w}_{T}))-F^{\ast}\\
-	& =\mathbb{E}(F(\overline{w}_{T})-F(w^{\ast}))\\
-	& \leq\frac{L}{2}\mathbb{E}\|\overline{w}_{T}-w^{\ast}\|^{2}\leq\frac{L}{2}c\|w_{0}-w^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
-	& =2c\|w_{0}-w^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
-	& \leq2c\|w_{0}-w^{\ast}\|^{2}\kappa\left[\frac{4}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}LG^{2}\cdot(\frac{4}{\mu(T+\gamma)})^{2}\right]\\
+	& \mathbb{E}(F(\ov{w}_{T}))-F^{\ast}\\
+	& =\mathbb{E}(F(\ov{w}_{T})-F(\vw^{\ast}))\\
+	& \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}\leq\frac{L}{2}c\|w_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
+	& =2c\|w_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
+	& \leq2c\|w_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{4}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}LG^{2}\cdot(\frac{4}{\mu(T+\gamma)})^{2}\right]\\
 	& =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
@@
 	is now given by averages over a subset of sampled devices. When $t+1\notin\mathcal{I}_{E}$,
-	$\overline{v}_{t+1}=\overline{w}_{t+1}$, while when $t+1\notin\mathcal{I}_{E}$,
-	we have $\mathbb{E}\overline{w}_{t+1}=\overline{v}_{t+1}$ by design
+	$\ov{v}_{t+1}=\ov{w}_{t+1}$, while when $t+1\notin\mathcal{I}_{E}$,
+	we have $\mathbb{E}\ov{w}_{t+1}=\ov{v}_{t+1}$ by design
 	of the sampling schemes, so that 
 	\begin{align*}
-	\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & =\mathbb{E}\|\overline{w}_{t+1}-\overline{v}_{t+1}+\overline{v}_{t+1}-w^{\ast}\|^{2}\\
-	& =\mathbb{E}\|\overline{w}_{t+1}-\overline{v}_{t+1}\|^{2}+\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}+\ov{v}_{t+1}-\vw^{\ast}\|^{2}\\
+	& =\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2}+\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}
 	\end{align*}
-	As before, $\mathbb{E}\|\overline{v}_{t+1}-w^{\ast}\|^{2}\leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
+	As before, $\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}\leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
 
-	The key is to bound $\mathbb{E}\|\overline{w}_{t+1}-\overline{v}_{t+1}\|^{2}$.
+	The key is to bound $\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2}$.
 	For sampling scheme I we have 
 	\begin{align*}
-	\mathbb{E}\|\overline{w}_{t+1}-\overline{v}_{t+1}\|^{2} & =\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|w_{t+1}^{k}-\overline{w}_{t+1}\|^{2}\\
+	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & =\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|w_{t+1}^{k}-\ov{w}_{t+1}\|^{2}\\
 	& \leq\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
@@
 	\begin{align*}
-	\mathbb{E}\|\overline{w}_{t+1}-\overline{v}_{t+1}\|^{2} & =\frac{N-K}{N-1}\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|w_{t+1}^{k}-\overline{w}_{t+1}\|^{2}\\
+	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & =\frac{N-K}{N-1}\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|w_{t+1}^{k}-\ov{w}_{t+1}\|^{2}\\
 	& \leq\frac{N-K}{N-1}\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
@@
 	\begin{align*}
-	\mathbb{E}F(\overline{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\ov{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
 	\end{align*}
@@
 	\begin{proof}
-	With partial participation, $2E\sum_{\tau=1}^{E-1}\alpha_{t-\tau}^{2}\sum_{k=1}^{N}p_{k}\left(\|\nabla F_{k}(\overline{w}_{t-\tau},\xi_{t-\tau}^{k})\|^{2}+l^{2}\|w_{t-\tau}^{k}-\overline{w}_{t-\tau}\|^{2}\right)$
+	With partial participation, $2E\sum_{\tau=1}^{E-1}\alpha_{t-\tau}^{2}\sum_{k=1}^{N}p_{k}\left(\|\nabla F_{k}(\ov{w}_{t-\tau},\xi_{t-\tau}^{k})\|^{2}+l^{2}\|w_{t-\tau}^{k}-\ov{w}_{t-\tau}\|^{2}\right)$
 	\begin{align*}
-	\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{1}{K}\sum_{k}p^{k}\|w_{t}^{k}-\overline{w}_{t}\|^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})\cdots(1-\mu\alpha_{t-E})\|\overline{w}_{t-E}-w^{\ast}\|^{2}=O(\frac{1}{t-E}\sigma^{2}+E^{2}LG^{2}\frac{1}{(t-E)^{2}}+\frac{E^{2}G^{2}}{K}\frac{1}{(t-E)^{3/2}})\\
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{1}{K}\sum_{k}p^{k}\|w_{t}^{k}-\ov{w}_{t}\|^{2}\\
+	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}\\
+	& \leq\mathbb{E}(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})\cdots(1-\mu\alpha_{t-E})\|\ov{w}_{t-E}-\vw^{\ast}\|^{2}=O(\frac{1}{t-E}\sigma^{2}+E^{2}LG^{2}\frac{1}{(t-E)^{2}}+\frac{E^{2}G^{2}}{K}\frac{1}{(t-E)^{3/2}})\\
 	& +(\alpha_{t}^{3}+(1-\mu\alpha_{t-1})\alpha_{t-1}^{3}+(1-\mu\alpha_{t-1})(1-\mu\alpha_{t-2})\alpha_{t-2}^{3}+\cdots+(1-\mu\alpha_{t-1})\cdots(1-\mu\alpha_{t-E})\alpha_{t-E}^{3}E^{2}LG^{2}\\
@@
 	\begin{align*}
-	\frac{1}{K}\sum_{k}p^{k}\|w_{t}^{k}-\overline{w}_{t}\|^{2} & =O(\alpha_{t}^{2}G^{2}/K)\\
-	\sum_{k}p^{k}\|w_{t}^{k}-\overline{w}_{t}\|^{2}\le\sum_{k}p^{k}\|w_{t}^{k}-\overline{w}_{t_{0}}\|^{2} & \leq\sum_{k}p^{k}\|-\alpha_{t_{0}}g_{t_{0},k}-\alpha_{t_{0}+1}g_{t_{0}+1,k}-\cdots-\alpha_{t-1}g_{t-1,k}\|^{2}\\
+	\frac{1}{K}\sum_{k}p^{k}\|w_{t}^{k}-\ov{w}_{t}\|^{2} & =O(\alpha_{t}^{2}G^{2}/K)\\
+	\sum_{k}p^{k}\|w_{t}^{k}-\ov{w}_{t}\|^{2}\le\sum_{k}p^{k}\|w_{t}^{k}-\ov{w}_{t_{0}}\|^{2} & \leq\sum_{k}p^{k}\|-\alpha_{t_{0}}g_{t_{0},k}-\alpha_{t_{0}+1}g_{t_{0}+1,k}-\cdots-\alpha_{t-1}g_{t-1,k}\|^{2}\\
 	& =\sum_{k}p^{k}\|\alpha_{t_{0}}g_{t_{0},k}+\alpha_{t_{0}+1}g_{t_{0}+1,k}+\cdots+\alpha_{t-1}g_{t-1,k}\|^{2}\\
-	& =\sum_{k}p^{k}\|\alpha_{t_{0}}\nabla F_{k}(\overline{w}_{t_{0}},\xi_{t_{0}}^{k})+\alpha_{t_{0}}\nabla F_{k}(\overline{w}_{t_{0}}-\alpha_{t_{0}+1}\nabla F_{k}(\overline{w}_{t_{0}},\xi_{t_{0}}^{k}),\xi_{t_{0}+1}^{k})+\cdots+\alpha_{t-1}g_{t-1,k}\|^{2}
+	& =\sum_{k}p^{k}\|\alpha_{t_{0}}\nabla F_{k}(\ov{w}_{t_{0}},\xi_{t_{0}}^{k})+\alpha_{t_{0}}\nabla F_{k}(\ov{w}_{t_{0}}-\alpha_{t_{0}+1}\nabla F_{k}(\ov{w}_{t_{0}},\xi_{t_{0}}^{k}),\xi_{t_{0}+1}^{k})+\cdots+\alpha_{t-1}g_{t-1,k}\|^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\sum_{k}p^{k}\|\alpha_{t_{0}}\nabla F_{k}(w^{\ast},\xi_{t_{0}}^{k})+\alpha_{t_{0}+1}\nabla F_{k}(w^{\ast}-\alpha_{t_{0}+1}\nabla F_{k}(w^{\ast},\xi_{t_{0}}^{k}),\xi_{t_{0}+1}^{k})+\cdots+\alpha_{t-1}g_{t-1,k}\|^{2} & \approx\\
+	\sum_{k}p^{k}\|\alpha_{t_{0}}\nabla F_{k}(\vw^{\ast},\xi_{t_{0}}^{k})+\alpha_{t_{0}+1}\nabla F_{k}(\vw^{\ast}-\alpha_{t_{0}+1}\nabla F_{k}(\vw^{\ast},\xi_{t_{0}}^{k}),\xi_{t_{0}+1}^{k})+\cdots+\alpha_{t-1}g_{t-1,k}\|^{2} & \approx\\
 	\end{align*}
-	since $\overline{w}_{t_{0}}\approx w^{\ast}+\frac{1}{t_{0}}$. We
+	since $\ov{w}_{t_{0}}\approx w^{\ast}+\frac{1}{t_{0}}$. We
 	can take 
 	\begin{align*}
-	g_{t_{0},k} & =\nabla F_{k}(\overline{w}_{t_{0}},\xi)\approx\nabla F_{k}(\overline{w}_{t_{0}},\xi)+\\
+	g_{t_{0},k} & =\nabla F_{k}(\ov{w}_{t_{0}},\xi)\approx\nabla F_{k}(\ov{w}_{t_{0}},\xi)+\\
 	\end{align*}

### 2020-06-10 source/app_nasgd_scvx_smooth.tex
\overline{z} ->\ov{z}\overline{g} ->\ov{g} |w_{t}^{k} -> |\vw_{t}^{k}

@@
 Define the virtual sequences $\ov{v}_{t}=\sum_{k=1}^{N}p_{k}v_{t}^{k}$,
-$\ov{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$, and $\overline{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$.
-We have $\mathbb{E}g_{t}=\overline{g}_{t}$ and $\ov{v}_{t+1}=\ov{w}_{t}-\alpha_{t}g_{t}$,
+$\ov{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$, and $\ov{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$.
+We have $\mathbb{E}g_{t}=\ov{g}_{t}$ and $\ov{v}_{t+1}=\ov{w}_{t}-\alpha_{t}g_{t}$,
 and $\ov{w}_{t+1}=\ov{v}_{t+1}$ for all $t$. The proof
@@
 \|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}g_{t})-\vw^{\ast}\|^{2}\\
-& =\|(\ov{w}_{t}-\alpha_{t}\overline{g}_{t}-\vw^{\ast})-\alpha_{t}(g_{t}-\overline{g}_{t})\|^{2}\\
+& =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(g_{t}-\ov{g}_{t})\|^{2}\\
 & =A_{1}+A_{2}+A_{3}
@@
 \begin{align*}
-A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2}\\
-A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t},\overline{g}_{t}-g_{t}\rangle\\
-A_{3} & =\alpha_{t}^{2}\|g_{t}-\overline{g}_{t}\|^{2}
+A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}\\
+A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-g_{t}\rangle\\
+A_{3} & =\alpha_{t}^{2}\|g_{t}-\ov{g}_{t}\|^{2}
 \end{align*}
-$\mathbb{E}A_{2}=0$ by definition of $g_{t}$ and $\overline{g}_{t}$,
+$\mathbb{E}A_{2}=0$ by definition of $g_{t}$ and $\ov{g}_{t}$,
 while for $A_{3}$ we have
 \begin{align*}
-\alpha_{t}^{2}\mathbb{E}\|g_{t}-\overline{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|g_{t}-\mathbb{E}g_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+\alpha_{t}^{2}\mathbb{E}\|g_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|g_{t}-\mathbb{E}g_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
 \end{align*}
@@
 \begin{align*}
-\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2}
+\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2}
 \end{align*}
@@
 \begin{align*}
--2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2} & \leq\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
+-2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2} & \leq\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
 \end{align*}
@@
 \begin{align*}
-\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & \leq(1-\alpha_{t}\mu)\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}
+\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} & \leq(1-\alpha_{t}\mu)\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}
 \end{align*}
@@
 \begin{align*}
-\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
-& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
-& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
+\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
+& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
+& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
 & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\sum_{i=t_{0}}^{t-1}\beta_{i}(v_{i+1}^{k}-v_{i}^{k})-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\

### 2020-06-10 source/app_sgd_cvx_smooth.tex
\overline{z} ->\ov{z}\overline{g} ->\ov{g} |w_{t}^{k} -> |\vw_{t}^{k}

@@
 	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}g_{t})-\vw^{\ast}\|^{2}\\
-	& =\|(\ov{w}_{t}-\alpha_{t}\overline{g}_{t}-\vw^{\ast})-\alpha_{t}(g_{t}-\overline{g}_{t})\|^{2}\\
+	& =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(g_{t}-\ov{g}_{t})\|^{2}\\
 	& =A_{1}+A_{2}+A_{3}
@@
 	\begin{align*}
-	A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2}\\
-	A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t},\overline{g}_{t}-g_{t}\rangle\\
-	A_{3} & =\alpha_{t}^{2}\|g_{t}-\overline{g}_{t}\|^{2}
+	A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}\\
+	A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-g_{t}\rangle\\
+	A_{3} & =\alpha_{t}^{2}\|g_{t}-\ov{g}_{t}\|^{2}
 	\end{align*}
-	$\mathbb{E}A_{2}=0$ by definition of $g_{t}$ and $\overline{g}_{t}$,
+	$\mathbb{E}A_{2}=0$ by definition of $g_{t}$ and $\ov{g}_{t}$,
 	while for $A_{3}$ we have
 	\begin{align*}
-	\alpha_{t}^{2}\mathbb{E}\|g_{t}-\overline{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|g_{t}-\mathbb{E}g_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+	\alpha_{t}^{2}\mathbb{E}\|g_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|g_{t}-\mathbb{E}g_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2}
+	\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2}
 	\end{align*}
@@
 	\begin{align*}
-	& -2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\overline{g}_{t}\rangle\\
+	& -2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle\\
 	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
@@
 	\begin{align*}
-	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
 	\end{align*}
@@
 	is that we no longer have a contraction factor in front of $\|\ov{w}_{t}-\vw^{\ast}\|^{2}$.
-	In the strongly convex case, we were able to cancel $\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}$
+	In the strongly convex case, we were able to cancel $\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$
 	with $2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]$
@@
 	\begin{align*}
-	\|\overline{g}_{t}\|^{2} & =\|\sum_{k}p_{k}\nabla F_{k}(\vw_{t}^{k})\|^{2}\\
+	\|\ov{g}_{t}\|^{2} & =\|\sum_{k}p_{k}\nabla F_{k}(\vw_{t}^{k})\|^{2}\\
 	& =\|\sum_{k}p_{k}\nabla F_{k}(\vw_{t}^{k})-\sum_{k}p_{k}\nabla F_{k}(\ov{w}_{t})+\sum_{k}p_{k}\nabla F_{k}(\ov{w}_{t})\|^{2}\\
 	& \leq2\|\sum_{k}p_{k}\nabla F_{k}(\vw_{t}^{k})-\sum_{k}p_{k}\nabla F_{k}(\ov{w}_{t})\|^{2}+2\|\sum_{k}p_{k}\nabla F_{k}(\ov{w}_{t})\|^{2}\\
-	& \leq2L^{2}\sum_{k}p_{k}\|w_{t}^{k}-\ov{w}_{t}\|^{2}+2\|\sum_{k}p_{k}\nabla F_{k}(\ov{w}_{t})\|^{2}\\
-	& =2L^{2}\sum_{k}p_{k}\|w_{t}^{k}-\ov{w}_{t}\|^{2}+2\|\nabla F(\ov{w}_{t})\|^{2}
+	& \leq2L^{2}\sum_{k}p_{k}\|\vw_{t}^{k}-\ov{w}_{t}\|^{2}+2\|\sum_{k}p_{k}\nabla F_{k}(\ov{w}_{t})\|^{2}\\
+	& =2L^{2}\sum_{k}p_{k}\|\vw_{t}^{k}-\ov{w}_{t}\|^{2}+2\|\nabla F(\ov{w}_{t})\|^{2}
 	\end{align*}
@@
 	& \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]\\
-	& +2\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|w_{t}^{k}-\ov{w}_{t}\|^{2}+4\alpha_{t}^{2}L(F(\ov{w}_{t})-F(\vw^{\ast}))+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+	& +2\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\vw_{t}^{k}-\ov{w}_{t}\|^{2}+4\alpha_{t}^{2}L(F(\ov{w}_{t})-F(\vw^{\ast}))+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
 	& =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\

### 2020-06-10 source/app_sgd_scvx_smooth.tex
\overline{z} ->\ov{z}\overline{g} ->\ov{g} |w_{t}^{k} -> |\vw_{t}^{k}

@@
 	\begin{align*}
-	\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\overline{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\overline{g}_{t}\rangle+\|\alpha_{t}\overline{g}_{t}\|^{2}
+	\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2}
 	\end{align*}
-	and we will show that the third term $\|\alpha_{t}\overline{g}_{t}\|^{2}$
+	and we will show that the third term $\|\alpha_{t}\ov{g}_{t}\|^{2}$
 	can be canceled by an upper bound of the second term. %
@@
 	\begin{align*}
-	\alpha_{t}^{2}\|\overline{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq2L\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}(F_{k}(\vw_{t}^{k})-F_{k}^{\ast})
+	\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq2L\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}(F_{k}(\vw_{t}^{k})-F_{k}^{\ast})
 	\end{align*}
@@
 	\begin{align*}
-	\alpha_{t}^{2}\|\overline{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\mathbb{E}\|\nabla F_{k}(\vw_{t}^{k},\xi_{t}^{k})\|^{2}\leq\alpha_{t}^{2}G^{2}
+	\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\mathbb{E}\|\nabla F_{k}(\vw_{t}^{k},\xi_{t}^{k})\|^{2}\leq\alpha_{t}^{2}G^{2}
 	\end{align*}
@@
 	\begin{align*}
-	& -2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\overline{g}_{t}\rangle\\
+	& -2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle\\
 	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
 	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\vw^{\ast}\|^{2}\\
+	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
 	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}w_{t}^{k}-\vw^{\ast}\|^{2}\\
@@
 	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),g_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\overline{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\overline{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\sum_{k}p_{k}\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\ov{w}_{t})-\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\overline{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
+	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
+	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\ov{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
+	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\sum_{k}p_{k}\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
+	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\ov{w}_{t})-\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
+	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
+	& \leq-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
 	\end{align*}
@@
 
-	Note that the term $-\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}$ exactly
-	cancels the $\alpha_{t}^{2}\|\overline{g}_{t}\|^{2}$ in the bound
-	for $A_{1}$, so that plugging in the bound for $-2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\overline{g}_{t}\rangle$,
+	Note that the term $-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$ exactly
+	cancels the $\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$ in the bound
+	for $A_{1}$, so that plugging in the bound for $-2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle$,
 	we have so far proved 
@@
 	& \mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
-	& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
+	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
+	& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
+	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
 	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\
@@
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{1}{K}\sum_{k}p^{k}\|w_{t}^{k}-\ov{w}_{t}\|^{2}\\
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{1}{K}\sum_{k}p^{k}\|\vw_{t}^{k}-\ov{w}_{t}\|^{2}\\
 	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}\\
@@
 	\begin{align*}
-	\frac{1}{K}\sum_{k}p^{k}\|w_{t}^{k}-\ov{w}_{t}\|^{2} & =O(\alpha_{t}^{2}G^{2}/K)\\
-	\sum_{k}p^{k}\|w_{t}^{k}-\ov{w}_{t}\|^{2}\le\sum_{k}p^{k}\|w_{t}^{k}-\ov{w}_{t_{0}}\|^{2} & \leq\sum_{k}p^{k}\|-\alpha_{t_{0}}g_{t_{0},k}-\alpha_{t_{0}+1}g_{t_{0}+1,k}-\cdots-\alpha_{t-1}g_{t-1,k}\|^{2}\\
+	\frac{1}{K}\sum_{k}p^{k}\|\vw_{t}^{k}-\ov{w}_{t}\|^{2} & =O(\alpha_{t}^{2}G^{2}/K)\\
+	\sum_{k}p^{k}\|\vw_{t}^{k}-\ov{w}_{t}\|^{2}\le\sum_{k}p^{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2} & \leq\sum_{k}p^{k}\|-\alpha_{t_{0}}g_{t_{0},k}-\alpha_{t_{0}+1}g_{t_{0}+1,k}-\cdots-\alpha_{t-1}g_{t-1,k}\|^{2}\\
 	& =\sum_{k}p^{k}\|\alpha_{t_{0}}g_{t_{0},k}+\alpha_{t_{0}+1}g_{t_{0}+1,k}+\cdots+\alpha_{t-1}g_{t-1,k}\|^{2}\\

### 2020-06-10 source/app_nasgd_scvx_smooth.tex
w_{0} -> \vw_{0} g_{t-1} -\vg_{t-1} \overline{p}->\ov{p}}

@@
 $\ov{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$, and $\ov{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$.
-We have $\mathbb{E}g_{t}=\ov{g}_{t}$ and $\ov{v}_{t+1}=\ov{w}_{t}-\alpha_{t}g_{t}$,
+We have $\mathbb{E}\vg_{t}=\ov{g}_{t}$ and $\ov{v}_{t+1}=\ov{w}_{t}-\alpha_{t}\vg_{t}$,
 and $\ov{w}_{t+1}=\ov{v}_{t+1}$ for all $t$. The proof
@@
 \begin{align*}
-\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}g_{t})-\vw^{\ast}\|^{2}\\
-& =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(g_{t}-\ov{g}_{t})\|^{2}\\
+\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}\vg_{t})-\vw^{\ast}\|^{2}\\
+& =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(\vg_{t}-\ov{g}_{t})\|^{2}\\
 & =A_{1}+A_{2}+A_{3}
@@
 A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}\\
-A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-g_{t}\rangle\\
-A_{3} & =\alpha_{t}^{2}\|g_{t}-\ov{g}_{t}\|^{2}
+A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-\vg_{t}\rangle\\
+A_{3} & =\alpha_{t}^{2}\|\vg_{t}-\ov{g}_{t}\|^{2}
 \end{align*}
-$\mathbb{E}A_{2}=0$ by definition of $g_{t}$ and $\ov{g}_{t}$,
+$\mathbb{E}A_{2}=0$ by definition of $\vg_{t}$ and $\ov{g}_{t}$,
 while for $A_{3}$ we have
 \begin{align*}
-\alpha_{t}^{2}\mathbb{E}\|g_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|g_{t}-\mathbb{E}g_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
 \end{align*}
@@
 \begin{align*}
--2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2} & \leq\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
+-2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2} & \leq\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
 \end{align*}
@@
 \begin{align*}
-\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} & \leq(1-\alpha_{t}\mu)\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}
+\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} & \leq(1-\alpha_{t}\mu)\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}
 \end{align*}
@@
 \mathbb{E}(F(\ov{w}_{T}))-F^{\ast} & =\mathbb{E}(F(\ov{w}_{T})-F(\vw^{\ast}))\\
-& \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}\leq\frac{L}{2}c\|w_{0}-\vw^{\ast}\|^{2}\frac{6}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
-& =3c\|w_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
-& \leq3c\|w_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{6}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{\max}\sigma^{2}+20E^{2}LG^{2}\cdot(\frac{6}{\mu(T+\gamma)})^{2}\right]\\
+& \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}\leq\frac{L}{2}c\|\vw_{0}-\vw^{\ast}\|^{2}\frac{6}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
+& =3c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
+& \leq3c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{6}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{\max}\sigma^{2}+20E^{2}LG^{2}\cdot(\frac{6}{\mu(T+\gamma)})^{2}\right]\\
 & =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{\max}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})

### 2020-06-10 source/app_sgd_cvx_smooth.tex
w_{0} -> \vw_{0} g_{t-1} -\vg_{t-1} \overline{p}->\ov{p}}

@@
 	\begin{align*}
-	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}g_{t})-\vw^{\ast}\|^{2}\\
-	& =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(g_{t}-\ov{g}_{t})\|^{2}\\
+	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}\vg_{t})-\vw^{\ast}\|^{2}\\
+	& =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(\vg_{t}-\ov{g}_{t})\|^{2}\\
 	& =A_{1}+A_{2}+A_{3}
@@
 	A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}\\
-	A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-g_{t}\rangle\\
-	A_{3} & =\alpha_{t}^{2}\|g_{t}-\ov{g}_{t}\|^{2}
+	A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-\vg_{t}\rangle\\
+	A_{3} & =\alpha_{t}^{2}\|\vg_{t}-\ov{g}_{t}\|^{2}
 	\end{align*}
-	$\mathbb{E}A_{2}=0$ by definition of $g_{t}$ and $\ov{g}_{t}$,
+	$\mathbb{E}A_{2}=0$ by definition of $\vg_{t}$ and $\ov{g}_{t}$,
 	while for $A_{3}$ we have
 	\begin{align*}
-	\alpha_{t}^{2}\mathbb{E}\|g_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|g_{t}-\mathbb{E}g_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+	\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\sum_{t=0}^{T}\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq\|w_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}
+	\sum_{t=0}^{T}\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq\|\vw_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|w_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right)
+	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|\vw_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right)
 	\end{align*}
@@
 \begin{align*}
-\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sqrt{NT}}\cdot\|w_{0}-\vw^{\ast}\|^{2}+\frac{1}{\sqrt{NT}}T\cdot\frac{N}{T}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\frac{1}{\sqrt{NT}}T(\sqrt{\frac{N}{T}})^{3}6E^{2}LG^{2}\\
-& \leq\frac{1}{\sqrt{NT}}\cdot\|w_{0}-\vw^{\ast}\|^{2}+\frac{1}{\sqrt{NT}}T\cdot\frac{N}{T}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\frac{N}{T}6E^{2}LG^{2}\\
-& =(\|w_{0}-\vw^{\ast}\|^{2}+\nu_{\max}^{2}\sigma^{2})\frac{1}{\sqrt{NT}}+\frac{N}{T}6E^{2}LG^{2}\\
+\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sqrt{NT}}\cdot\|\vw_{0}-\vw^{\ast}\|^{2}+\frac{1}{\sqrt{NT}}T\cdot\frac{N}{T}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\frac{1}{\sqrt{NT}}T(\sqrt{\frac{N}{T}})^{3}6E^{2}LG^{2}\\
+& \leq\frac{1}{\sqrt{NT}}\cdot\|\vw_{0}-\vw^{\ast}\|^{2}+\frac{1}{\sqrt{NT}}T\cdot\frac{N}{T}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\frac{N}{T}6E^{2}LG^{2}\\
+& =(\|\vw_{0}-\vw^{\ast}\|^{2}+\nu_{\max}^{2}\sigma^{2})\frac{1}{\sqrt{NT}}+\frac{N}{T}6E^{2}LG^{2}\\
 & =O(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T})
@@
 	\begin{align*}
-	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|w_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot(\frac{1}{N}\nu_{\max}\sigma^{2}+C)+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right)
+	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|\vw_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot(\frac{1}{N}\nu_{\max}\sigma^{2}+C)+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right)
 	\end{align*}

### 2020-06-10 source/app_sgd_scvx_smooth.tex
w_{0} -> \vw_{0} g_{t-1} -\vg_{t-1} \overline{p}->\ov{p}}

@@
 	& \leq2\alpha_{t}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{w}_{t+1}-\ov{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\ov{w}_{t+1}-\ov{w}_{t}\|^{2}\\
-	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),g_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\ov{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\sum_{k}p_{k}\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\ov{w}_{t})-\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
+	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\vg_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\ov{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\sum_{k}p_{k}\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\ov{w}_{t})-\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	& \leq-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
 	\end{align*}
@@
 	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|g_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
+	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
 	\end{align*}
-	The term $\mathbb{E}\|g_{t}\|^{2}\leq G^{2}$ by assumption.
+	The term $\mathbb{E}\|\vg_{t}\|^{2}\leq G^{2}$ by assumption.
 
@@
 	\begin{align*}
-	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & =\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\alpha_{t-1}g_{t-1}-\vw_{t-1}^{k}+\alpha_{t-1}g_{t-1,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}\left(\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\alpha_{t-1}g_{t-1}-\alpha_{t-1}g_{t-1,k}\|^{2}\right)
+	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & =\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\alpha_{t-1}\vg_{t-1}-\vw_{t-1}^{k}+\alpha_{t-1}g_{t-1,k}\|^{2}\\
+	& \leq2\sum_{k=1}^{N}p_{k}\left(\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\alpha_{t-1}\vg_{t-1}-\alpha_{t-1}g_{t-1,k}\|^{2}\right)
 	\end{align*}
@@
 	\begin{align*}
-	\sum_{k}p_{k}\|g_{t-1,k}-g_{t-1}\|^{2}\leq\sum_{k}p_{k}\|g_{t-1,k}\|^{2} & =\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})+\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
+	\sum_{k}p_{k}\|g_{t-1,k}-\vg_{t-1}\|^{2}\leq\sum_{k}p_{k}\|g_{t-1,k}\|^{2} & =\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})+\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
 	& \leq2\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})\|^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
@@
 	\begin{align*}
-	\|w_{0}-\vw^{\ast}\|^{2} & \leq b\cdot(\alpha_{0}D+\alpha_{0}^{2}C)\\
+	\|\vw_{0}-\vw^{\ast}\|^{2} & \leq b\cdot(\alpha_{0}D+\alpha_{0}^{2}C)\\
 	& =b\cdot(\frac{4}{\mu\gamma}D+\frac{16}{\mu^{2}\gamma^{2}}C)
 	\end{align*}
-	we can simply scale $b$ by $c\|w_{0}-\vw^{\ast}\|^{2}$ for a constant
+	we can simply scale $b$ by $c\|\vw_{0}-\vw^{\ast}\|^{2}$ for a constant
 	$c$ large enough and the induction step still holds.
@@
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2} & \leq c\|w_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{t}+C\alpha_{t}^{2})
+	\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2} & \leq c\|\vw_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{t}+C\alpha_{t}^{2})
 	\end{align*}
@@
 	& =\mathbb{E}(F(\ov{w}_{T})-F(\vw^{\ast}))\\
-	& \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}\leq\frac{L}{2}c\|w_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
-	& =2c\|w_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
-	& \leq2c\|w_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{4}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}LG^{2}\cdot(\frac{4}{\mu(T+\gamma)})^{2}\right]\\
+	& \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}\leq\frac{L}{2}c\|\vw_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
+	& =2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
+	& \leq2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{4}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}LG^{2}\cdot(\frac{4}{\mu(T+\gamma)})^{2}\right]\\
 	& =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})

### 2020-06-10 source/app_nasgd_scvx_smooth.tex
update to vector notations
v_{t+1}^{k} -> \vv_{t+1}^{k}, v_{0}->\vv_{0}, |w_{t+1}^{k} -> |\vw_{t+1}^{k}

g_{i,k}-> \vg_{i,k}

v_{i+1}^{k}  \vv_{i+1}^{k}

@@
 \begin{proof}
-Define the virtual sequences $\ov{v}_{t}=\sum_{k=1}^{N}p_{k}v_{t}^{k}$,
-$\ov{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$, and $\ov{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$.
+Define the virtual sequences $\ov{v}_{t}=\sum_{k=1}^{N}p_{k}\vv_{t}^{k}$,
+$\ov{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$, and $\ov{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}\vg_{t,k}$.
 We have $\mathbb{E}\vg_{t}=\ov{g}_{t}$ and $\ov{v}_{t+1}=\ov{w}_{t}-\alpha_{t}\vg_{t}$,
@@
 \begin{align*}
-\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|\vg_{t,k}-\mathbb{E}\vg_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
 \end{align*}
-again by Jensen's inequality and using the independence of $g_{t,k},g_{t,k'}$.
+again by Jensen's inequality and using the independence of $\vg_{t,k},g_{t,k'}$.
 
@@
 & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
-& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\sum_{i=t_{0}}^{t-1}\beta_{i}(v_{i+1}^{k}-v_{i}^{k})-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\
-& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}\|g_{i,k}\|^{2}+2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\beta_{i}^{2}\|(v_{i+1}^{k}-v_{i}^{k})\|^{2}\\
-& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}(\|g_{i,k}\|^{2}+\|(v_{i+1}^{k}-v_{i}^{k})\|^{2})\\
+& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\sum_{i=t_{0}}^{t-1}\beta_{i}(\vv_{i+1}^{k}-\vv_{i}^{k})-\sum_{i=t_{0}}^{t-1}\alpha_{i}\vg_{i,k}\|^{2}\\
+& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}\|\vg_{i,k}\|^{2}+2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\beta_{i}^{2}\|(\vv_{i+1}^{k}-\vv_{i}^{k})\|^{2}\\
+& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}(\|\vg_{i,k}\|^{2}+\|(\vv_{i+1}^{k}-\vv_{i}^{k})\|^{2})\\
 & \leq4\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}G^{2}\\
@@
 \end{align*}
-where we have used $\mathbb{E}\|v_{t}^{k}-v_{t-1}^{k}\|^{2}\leq G^{2}$.
+where we have used $\mathbb{E}\|\vv_{t}^{k}-\vv_{t-1}^{k}\|^{2}\leq G^{2}$.
 To see this identity for appropriate $\alpha_{t},\beta_{t}$, note
@@
 \begin{align*}
-v_{t+1}^{k}-v_{t}^{k} & =w_{t}^{k}-\vw_{t-1}^{k}-(\alpha_{t}g_{t,k}-\alpha_{t-1}g_{t-1,k})\\
-w_{t+1}^{k}-\vw_{t}^{k} & =-\alpha_{t}g_{t,k}+\beta_{t}(v_{t+1}^{k}-v_{t}^{k})
+\vv_{t+1}^{k}-\vv_{t}^{k} & =\vw_{t}^{k}-\vw_{t-1}^{k}-(\alpha_{t}\vg_{t,k}-\alpha_{t-1}\vg_{t-1,k})\\
+\vw_{t+1}^{k}-\vw_{t}^{k} & =-\alpha_{t}\vg_{t,k}+\beta_{t}(\vv_{t+1}^{k}-\vv_{t}^{k})
 \end{align*}
@@
 \begin{align*}
-v_{t+1}^{k}-v_{t}^{k} & =-\alpha_{t-1}g_{t-1,k}+\beta_{t-1}(v_{t}^{k}-v_{t-1}^{k})-(\alpha_{t}g_{t,k}-\alpha_{t-1}g_{t-1,k})\\
-& =\beta_{t-1}(v_{t}^{k}-v_{t-1}^{k})-\alpha_{t}g_{t,k}
+\vv_{t+1}^{k}-\vv_{t}^{k} & =-\alpha_{t-1}\vg_{t-1,k}+\beta_{t-1}(\vv_{t}^{k}-\vv_{t-1}^{k})-(\alpha_{t}\vg_{t,k}-\alpha_{t-1}\vg_{t-1,k})\\
+& =\beta_{t-1}(\vv_{t}^{k}-\vv_{t-1}^{k})-\alpha_{t}\vg_{t,k}
 \end{align*}
-Since the identity $v_{t+1}^{k}-v_{t}^{k}=\beta_{t-1}(v_{t}^{k}-v_{t-1}^{k})-\alpha_{t}g_{t,k}$
+Since the identity $\vv_{t+1}^{k}-\vv_{t}^{k}=\beta_{t-1}(\vv_{t}^{k}-\vv_{t-1}^{k})-\alpha_{t}\vg_{t,k}$
 implies 
 \begin{align*}
-\mathbb{E}\|v_{t+1}^{k}-v_{t}^{k}\|^{2} & \leq2\beta_{t-1}^{2}\mathbb{E}\|v_{t}^{k}-v_{t-1}^{k}\|^{2}+2\alpha_{t}^{2}G^{2}
+\mathbb{E}\|\vv_{t+1}^{k}-\vv_{t}^{k}\|^{2} & \leq2\beta_{t-1}^{2}\mathbb{E}\|\vv_{t}^{k}-\vv_{t-1}^{k}\|^{2}+2\alpha_{t}^{2}G^{2}
 \end{align*}
 as long as $\alpha_{t},\beta_{t-1}$ satisfy $2\beta_{t-1}^{2}+2\alpha_{t}^{2}\leq1/2$,
-we can guarantee that $\mathbb{E}\|v_{t}^{k}-v_{t-1}^{k}\|^{2}\leq G^{2}$
+we can guarantee that $\mathbb{E}\|\vv_{t}^{k}-\vv_{t-1}^{k}\|^{2}\leq G^{2}$
 for all $k$ by induction. This together with Jensen's inequality
@@
 
-Finally, to ensure $\|v_{0}-\vw^{\ast}\|^{2}\leq b(C\alpha_{0}^{2}+D\alpha_{0})$,
-we can rescale $b$ by $c\|v_{0}-\vw^{\ast}\|^{2}$ for some $c.$ It
+Finally, to ensure $\|\vv_{0}-\vw^{\ast}\|^{2}\leq b(C\alpha_{0}^{2}+D\alpha_{0})$,
+we can rescale $b$ by $c\|\vv_{0}-\vw^{\ast}\|^{2}$ for some $c.$ It
 follows that $\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}\leq b(C\alpha_{t}^{2}+D\alpha_{t})$

### 2020-06-10 source/app_sumarization.tex
Update app_sumarization.tex

@@
 \end{align*} 
-This is a more restrictive comparing to assuming bounded gradient under the case of target accuracy $\epsilon \rightarrow 0$ and PL condition.
-To see this, consider the gradient diversity at the global optimal $\vw^*$,i.e., $\Lambda(\vw^*) = \frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$.
+This is a more restrictive assumption comparing to assuming bounded gradient under the case of target accuracy $\epsilon \rightarrow 0$ and PL condition.
+To see this, consider the gradient diversity at the global optimal $\vw^*$, i.e., $\Lambda(\vw^*) = \frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$.
 For $\Lambda(\vw^*)$ to be bounded, it requires $\|\nabla F_{k}(\mathbf{w}^*)\|_{2}^{2} = 0$, $\forall \ k$.

### 2020-06-10 source/app_notations.tex
sec A, B, E

@@
 \begin{align}
-\overline{\mathbf{v}}_{t+1}=\overline{\mathbf{w}}_{t}-\eta_{t} \mathbf{g}_{t}	\label{eq:vbar}
+\overline{\mathbf{v}}_{t+1}=\overline{\mathbf{w}}_{t}-\eta_{t} \mathbf{g}_{t},	\label{eq:vbar}
 \end{align}
@@
 where $\vg_{t} = \sum_{k=1}^{N} p_{k} \vg_{t,k} $ is one-step stochastic gradient, averaged over all devices. 
-\begin{align}
-\vg_{t,k} \left\{\begin{array}{ll} 
- = \nabla F_{k}\left(\mathbf{w}_{t}^{k},\xi_{t}^{k} \right)  &  \text{smooth}\\
- \in \partial F_{k}\left(w_{t}^{k}, \xi_{t}^{k}\right)  & \text{non-smooth}
- \end{array}\right.
-\end{align}
+\begin{align*}
+\vg_{t,k} = \nabla F_{k}\left(\mathbf{w}_{t}^{k},\xi_{t}^{k} \right),   
+\end{align*}
+% \begin{align}
+% \vg_{t,k} \left\{\begin{array}{ll} 
+%  = \nabla F_{k}\left(\mathbf{w}_{t}^{k},\xi_{t}^{k} \right)  &  \text{smooth}\\
+%  \in \partial F_{k}\left(\vw_{t}^{k}, \xi_{t}^{k}\right)  & \text{non-smooth} \lkxcom{remove non-smooth}
+%  \end{array}\right.
+% \end{align}
 Similarly, we denote the expected one-step gradient $\ov{g}_{t}= \EE_{\xi_t}[\vg_t] = \sum_{k=1}^{N} p_{k} \EE_{\xi_{t}^{k}} \vg_{t,k}$, where
-\begin{align}
-\EE_{\xi_{t}^{k}} \vg_{t,k}  \left\{\begin{array}{ll} 
- = \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)  &  \text{smooth}\\
- \in \partial F_{k}\left(w_{t}^{k}\right)  & \text{non-smooth}
- \end{array}\right.
-\end{align}
+\begin{align*}
+\EE_{\xi_{t}^{k}} \vg_{t,k}  = \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right), 
+\end{align*}
+% \begin{align}
+% \EE_{\xi_{t}^{k}} \vg_{t,k}  \left\{\begin{array}{ll} 
+%  = \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)  &  \text{smooth}\\
+%  \in \partial F_{k}\left(\vw_{t}^{k}\right)  & \text{non-smooth}
+%  \end{array}\right.
+% \end{align}
 and $\xi_t = \{\xi_t^k\}_{k=1}^N$ denotes random samples at all devices at time step $t$. 
@@
 The sampling scheme I establishes $\cS_{t+1}$ by i.i.d. sampling the devices with replacement,
-in this case the upper bound of expected square norm of $\ov{w}_{t+1} - \ov{v}_{t+1}$ is given by:
+in this case the upper bound of expected square norm of $\ov{w}_{t+1} - \ov{v}_{t+1}$ is given by~\cite[Lemma 5]{li2019convergence}:
 \begin{align}
-\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4}{K} \eta_t^2 E^2G^2
+\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4}{K} \eta_t^2 E^2G^2.
 \end{align}
@@
 \begin{align}
-\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4(N - K)}{K(N-1)} \eta_t^2 E^2G^2
+\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4(N - K)}{K(N-1)} \eta_t^2 E^2G^2.
 \end{align}
@@
 \begin{align}
-	\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2 \leq  \eta_t^2 C
+	\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2 \leq  \eta_t^2 C.
 	\label{eq:partialsample}

### 2020-06-10 source/app_notations.tex
notations

@@
 \begin{align*}
-\vg_{t,k} = \nabla F_{k}\left(\mathbf{w}_{t}^{k},\xi_{t}^{k} \right),   
+\vg_{t,k} = \nabla F_{k}\left(\mathbf{w}_{t}^{k},\mathbf{\xi}_{t}^{k} \right),   
 \end{align*}
@@
 % \vg_{t,k} \left\{\begin{array}{ll} 
-%  = \nabla F_{k}\left(\mathbf{w}_{t}^{k},\xi_{t}^{k} \right)  &  \text{smooth}\\
-%  \in \partial F_{k}\left(\vw_{t}^{k}, \xi_{t}^{k}\right)  & \text{non-smooth} \lkxcom{remove non-smooth}
+%  = \nabla F_{k}\left(\mathbf{w}_{t}^{k},\mathbf{\xi}_{t}^{k} \right)  &  \text{smooth}\\
+%  \in \partial F_{k}\left(\vw_{t}^{k}, \mathbf{\xi}_{t}^{k}\right)  & \text{non-smooth} \lkxcom{remove non-smooth}
 %  \end{array}\right.
 % \end{align}
-Similarly, we denote the expected one-step gradient $\ov{g}_{t}= \EE_{\xi_t}[\vg_t] = \sum_{k=1}^{N} p_{k} \EE_{\xi_{t}^{k}} \vg_{t,k}$, where
+Similarly, we denote the expected one-step gradient $\ov{g}_{t}= \EE_{\xi_t}[\vg_t] = \sum_{k=1}^{N} p_{k} \EE_{\mathbf{\xi}_{t}^{k}} \vg_{t,k}$, where
 \begin{align*}
-\EE_{\xi_{t}^{k}} \vg_{t,k}  = \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right), 
+\EE_{\mathbf{\xi}_{t}^{k}} \vg_{t,k}  = \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right), 
 \end{align*}
 % \begin{align}
-% \EE_{\xi_{t}^{k}} \vg_{t,k}  \left\{\begin{array}{ll} 
+% \EE_{\mathbf{\xi}_{t}^{k}} \vg_{t,k}  \left\{\begin{array}{ll} 
 %  = \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)  &  \text{smooth}\\
@@
 % \end{align}
-and $\xi_t = \{\xi_t^k\}_{k=1}^N$ denotes random samples at all devices at time step $t$. 
+and $\mathbf{\xi}_t = \{\mathbf{\xi}_t^k\}_{k=1}^N$ denotes random samples at all devices at time step $t$. 
 Since in this work, we also consider the case of partial participation. The sampling

### 2020-06-10 source/app_sumarization.tex

@@
 \label{sec:app:comparison}
-In this section, we summarize the convergence results in this work from various perspectives
-and compare our results with related works. 
+In this section, we compare our convergence rate with the best-known results in the literature (see Table~\ref{tb:convergenceratev3}). 
 % In Table~\ref{tb:convergenceratev1} and Table~\ref{tb:convergenceratev2}, we summarize 
 % the convergence results in terms of participations and accelerations. 
-In Table~\ref{tb:convergenceratev3}, we compare our 
-results with the best known results in the literatures.
-
-In~\cite{haddadpour2019convergence}, they also provide $\cO(1/NT)$ convergence rate of non-convex problems under Polyak-Łojasiewicz (PL) condition, which
+% In Table~\ref{tb:convergenceratev3}, we compare our 
+% results with the best known results in the literatures. 
+In~\cite{haddadpour2019convergence}, the authors provide $\cO(1/NT)$ convergence rate of non-convex problems under Polyak-Łojasiewicz (PL) condition, which
 means their results can directly apply to the strongly convex problems. However, their assumption is based on bounded gradient diversity, defined as follows: 
@@
 }
-\caption{A high level summary of the convergence results in this paper and their comparison to prior state-of-the-art FL algorithms, considering heterogeneous data. This table only highlights the
-dependence on $T$ (number of iterations), $E$ (the largest number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices. 
+\caption{A high-level summary of the convergence results in this paper compared to prior state-of-the-art FL algorithms. This table only highlights the
+dependence on $T$ (number of iterations), $E$ (the maximal number of local steps), $N$ (the total number of devices), and $K\leq N$ the number of participated devices. 
 $\kappa$ is the condition number of the system and $\beta \in (0,1)$. We denote Nesterov accelerated FedAvg as N-FedAvg in this table.}
@@
          % $^{\ddagger}$ The convergence of FedAvg for general overparameterized strongly convex and overparameterized linear regression share the same order in terms of $E, T, N$.\\
-         $^{\mathsection}$ In~\cite{haddadpour2019convergence}, the convergence rate is for non-convex smooth problem with PL condition, which also applies for strongly convex problems.
-Therefore, we compare it with our strongly convex results here.
-\\
+         $^{\mathsection}$ In~\cite{haddadpour2019convergence}, the convergence rate is for non-convex smooth problems with PL condition, which also applies to strongly convex problems.
+Therefore, we compare it with our strongly convex results here.\\
          $^{\ddagger\ddagger}$ The bounded gradient diversity assumption is not applicable for general heterogeneous data when converging to arbitrarily small $\epsilon$-accuracy (see discussions in Sec~\ref{sec:app:comparison}).

### 2020-06-10 source/app_sgd_scvx_smooth.tex
update to vector notations
v_{t+1}^{k} -> \vv_{t+1}^{k}, v_{0}->\vv_{0}, |w_{t+1}^{k} -> |\vw_{t+1}^{k}

g_{i,k}-> \vg_{i,k}

v_{i+1}^{k}  \vv_{i+1}^{k}
notations
Update app_sgd_scvx_smooth.tex
1. update constant in E^2LG^2

@@
+% !TEX ROOT=./supp-main.tex
 To facilitate reading, theorems from the main paper are restated and numbered identically. We first summarize some properties of $L$-smooth and $\mu$-strongly
@@
 	\end{align*}
-	$\mathbb{E}A_{2}=0$ by definition of $\vg_{t}$ and $\ov{g}_{t}$,
-	while for $A_{3}$ we have
+	By definition of $\vg_{t}$ and $\ov{g}_{t}$ (see \eq{\ref{eq:gradient}}), we have $\mathbb{E}A_{2}=0$.
+	For $A_{3}$, we have the follow upper bound:
 	\begin{align*}
-	\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+	\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|\vg_{t,k}-\mathbb{E}\vg_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
 	\end{align*}
-	again by Jensen's inequality and using the independence of $g_{t,k},g_{t,k'}$.
+	again by Jensen's inequality and using the independence of $\vg_{t,k},\vg_{t,k'}$ \cite[Lemma 2]{li2019convergence}.
 
@@
 	\begin{align*}
-	\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\mathbb{E}\|\nabla F_{k}(\vw_{t}^{k},\xi_{t}^{k})\|^{2}\leq\alpha_{t}^{2}G^{2}
+	\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\mathbb{E}\|\nabla F_{k}(\vw_{t}^{k},\mathbf{\xi}_{t}^{k})\|^{2}\leq\alpha_{t}^{2}G^{2}
 	\end{align*}
@@
 	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle \vw_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
 	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
-	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}w_{t}^{k}-\vw^{\ast}\|^{2}\\
+	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
 	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
@@
 	Since communication is done every $E$ steps, for any $t\geq0$, we
-	can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $w_{t_{0}}^{k}=\ov{w}_{t_{0}}$for
+	can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $\vw_{t_{0}}^{k}=\ov{w}_{t_{0}}$for
 	all $k$. Moreover, using $\alpha_{t}$ is non-increasing and $\alpha_{t_{0}}\leq2\alpha{}_{t}$
@@
 	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}E\alpha_{i}^{2}\|g_{i,k}\|^{2}\\
+	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}\vg_{i,k}\|^{2}\\
+	& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}E\alpha_{i}^{2}\|\vg_{i,k}\|^{2}\\
 	& \leq2\sum_{k=1}^{N}p_{k}E^{2}\alpha_{t_{0}}^{2}G^{2}\\
@@
 	\begin{align*}
-	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & =\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\alpha_{t-1}\vg_{t-1}-\vw_{t-1}^{k}+\alpha_{t-1}g_{t-1,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}\left(\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\alpha_{t-1}\vg_{t-1}-\alpha_{t-1}g_{t-1,k}\|^{2}\right)
+	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & =\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\alpha_{t-1}\vg_{t-1}-\vw_{t-1}^{k}+\alpha_{t-1}\vg_{t-1,k}\|^{2}\\
+	& \leq2\sum_{k=1}^{N}p_{k}\left(\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\alpha_{t-1}\vg_{t-1}-\alpha_{t-1}\vg_{t-1,k}\|^{2}\right)
 	\end{align*}
@@
 	\begin{align*}
-	\sum_{k}p_{k}\|g_{t-1,k}-\vg_{t-1}\|^{2}\leq\sum_{k}p_{k}\|g_{t-1,k}\|^{2} & =\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})+\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
-	& \leq2\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})\|^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
+	\sum_{k}p_{k}\|\vg_{t-1,k}-\vg_{t-1}\|^{2}\leq\sum_{k}p_{k}\|\vg_{t-1,k}\|^{2} & =\sum_{k}p_{k}\|\vg_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})+\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
+	& \leq2\sum_{k}p_{k}\|\vg_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})\|^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
 	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})-\nabla F_{k}(\ov{w}_{t-1})+\nabla F_{k}(\ov{w}_{t-1})\|^{2}\\
@@
 	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\nu_{max}^{2}\sum_{k=1}^{N}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
+	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
 	\end{align*}
-	where $\sigma^{2}=\max_{k}\sigma_{k}^{2}$, and if $\alpha_{t}$
+	where in the last inequality we use $\sigma^{2}=\max_{k}\sigma_{k}^{2}$, and assume $\alpha_{t}$
 	satisfies $L\alpha_{t}\leq\frac{1}{8}$. We show next that $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}LG^{2}}{t^{2}})$.
 
-	Let $C\equiv5E^{2}LG^{2}$ and $D\equiv\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
+	Let $C\equiv6E^{2}LG^{2}$ and $D\equiv\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
 	Suppose that we have shown $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}\leq b\cdot(\alpha_{t}D+\alpha_{t}^{2}C)$
@@
 	and $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}\leq b\alpha_{t+1}^{2}$.
-	Recall that we require $\alpha_{t_{0}}\leq\alpha_{t}$ for any $t-t_{0}\leq E-1$,
+	Recall that we require $\alpha_{t_{0}}\leq2\alpha_{t}$ for any $t-t_{0}\leq E-1$,
 	and $L\alpha_{t}\leq\frac{1}{8}$. If we let $\alpha_{t}=\frac{4}{\mu(t+\gamma)}$
@@
 	& =2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
-	& \leq2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{4}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}LG^{2}\cdot(\frac{4}{\mu(T+\gamma)})^{2}\right]\\
+	& \leq2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{4}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}LG^{2}\cdot(\frac{4}{\mu(T+\gamma)})^{2}\right]\\
 	& =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
@@
 	\end{align*}
-	As before, $\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}\leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
+	As before, $\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}\leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
 
@@
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & =\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|w_{t+1}^{k}-\ov{w}_{t+1}\|^{2}\\
+	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & =\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|\vw_{t+1}^{k}-\ov{w}_{t+1}\|^{2}\\
 	& \leq\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
@@
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & =\frac{N-K}{N-1}\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|w_{t+1}^{k}-\ov{w}_{t+1}\|^{2}\\
+	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & =\frac{N-K}{N-1}\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|\vw_{t+1}^{k}-\ov{w}_{t+1}\|^{2}\\
 	& \leq\frac{N-K}{N-1}\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
@@
 	\begin{proof}
-	With partial participation, $2E\sum_{\tau=1}^{E-1}\alpha_{t-\tau}^{2}\sum_{k=1}^{N}p_{k}\left(\|\nabla F_{k}(\ov{w}_{t-\tau},\xi_{t-\tau}^{k})\|^{2}+l^{2}\|w_{t-\tau}^{k}-\ov{w}_{t-\tau}\|^{2}\right)$
+	With partial participation, $2E\sum_{\tau=1}^{E-1}\alpha_{t-\tau}^{2}\sum_{k=1}^{N}p_{k}\left(\|\nabla F_{k}(\ov{w}_{t-\tau},\mathbf{\xi}_{t-\tau}^{k})\|^{2}+l^{2}\|w_{t-\tau}^{k}-\ov{w}_{t-\tau}\|^{2}\right)$
 	\begin{align*}
@@
 	\frac{1}{K}\sum_{k}p^{k}\|\vw_{t}^{k}-\ov{w}_{t}\|^{2} & =O(\alpha_{t}^{2}G^{2}/K)\\
-	\sum_{k}p^{k}\|\vw_{t}^{k}-\ov{w}_{t}\|^{2}\le\sum_{k}p^{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2} & \leq\sum_{k}p^{k}\|-\alpha_{t_{0}}g_{t_{0},k}-\alpha_{t_{0}+1}g_{t_{0}+1,k}-\cdots-\alpha_{t-1}g_{t-1,k}\|^{2}\\
-	& =\sum_{k}p^{k}\|\alpha_{t_{0}}g_{t_{0},k}+\alpha_{t_{0}+1}g_{t_{0}+1,k}+\cdots+\alpha_{t-1}g_{t-1,k}\|^{2}\\
-	& =\sum_{k}p^{k}\|\alpha_{t_{0}}\nabla F_{k}(\ov{w}_{t_{0}},\xi_{t_{0}}^{k})+\alpha_{t_{0}}\nabla F_{k}(\ov{w}_{t_{0}}-\alpha_{t_{0}+1}\nabla F_{k}(\ov{w}_{t_{0}},\xi_{t_{0}}^{k}),\xi_{t_{0}+1}^{k})+\cdots+\alpha_{t-1}g_{t-1,k}\|^{2}
+	\sum_{k}p^{k}\|\vw_{t}^{k}-\ov{w}_{t}\|^{2}\le\sum_{k}p^{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2} & \leq\sum_{k}p^{k}\|-\alpha_{t_{0}}g_{t_{0},k}-\alpha_{t_{0}+1}g_{t_{0}+1,k}-\cdots-\alpha_{t-1}\vg_{t-1,k}\|^{2}\\
+	& =\sum_{k}p^{k}\|\alpha_{t_{0}}g_{t_{0},k}+\alpha_{t_{0}+1}g_{t_{0}+1,k}+\cdots+\alpha_{t-1}\vg_{t-1,k}\|^{2}\\
+	& =\sum_{k}p^{k}\|\alpha_{t_{0}}\nabla F_{k}(\ov{w}_{t_{0}},\mathbf{\xi}_{t_{0}}^{k})+\alpha_{t_{0}}\nabla F_{k}(\ov{w}_{t_{0}}-\alpha_{t_{0}+1}\nabla F_{k}(\ov{w}_{t_{0}},\mathbf{\xi}_{t_{0}}^{k}),\mathbf{\xi}_{t_{0}+1}^{k})+\cdots+\alpha_{t-1}\vg_{t-1,k}\|^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\sum_{k}p^{k}\|\alpha_{t_{0}}\nabla F_{k}(\vw^{\ast},\xi_{t_{0}}^{k})+\alpha_{t_{0}+1}\nabla F_{k}(\vw^{\ast}-\alpha_{t_{0}+1}\nabla F_{k}(\vw^{\ast},\xi_{t_{0}}^{k}),\xi_{t_{0}+1}^{k})+\cdots+\alpha_{t-1}g_{t-1,k}\|^{2} & \approx\\
+	\sum_{k}p^{k}\|\alpha_{t_{0}}\nabla F_{k}(\vw^{\ast},\mathbf{\xi}_{t_{0}}^{k})+\alpha_{t_{0}+1}\nabla F_{k}(\vw^{\ast}-\alpha_{t_{0}+1}\nabla F_{k}(\vw^{\ast},\mathbf{\xi}_{t_{0}}^{k}),\mathbf{\xi}_{t_{0}+1}^{k})+\cdots+\alpha_{t-1}\vg_{t-1,k}\|^{2} & \approx\\
 	\end{align*}

### 2020-06-11 source/app_sgd_scvx_smooth.tex
counterexample lemma

@@
 	\end{comment}
+\end{proof}
 
 One may ask whether the dependence on $E$ in the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$
-	can be removed, or equivalently whether $\sum_{k}p_{k}\|\mathbf{w}_{t+1}^{k}-\overline{\mathbf{w}}_{t+1}\|^{2}$
-	can be independent of $E$. We provide a heuristic argument for why
-	this is not possible. Suppose that each $F_{k}(\mathbf{w})=\frac{1}{n}\sum(\mathbf{x}_{k}^{j}-\mathbf{w})^{2}$
-	contains data points $\mathbf{x}_{k}^{j}$ concentrated around the
-	optimizer $\mathbf{w}^{\ast,k}$, and that the $\mathbf{w}{}^{\ast,k}$'s
-	come in pairs around the origin. As a result, the global objective
-	$F$ is minimized at $\mathbf{w}^{\ast}=0$. Then with $t=O(T)$ large,
-	after a communication round $t$ each device will start with $\overline{\mathbf{w}}_{t}$
-	close to the origin, while after $E=T^{\beta}$ iterations with step
-	size $O(1/T)$, the local parameters are updated towards $\mathbf{w}^{\ast,k}$
-	with $\|\mathbf{w}_{t+E}^{k}\|^{2}\approx\frac{1}{T^{2-2\beta}}$,
-	while $\overline{\mathbf{w}}_{t+E}$ remains close to $\mathbf{w}^{\ast}$,
-	so that $\sum_{k}p_{k}\|\mathbf{w}_{t+E}^{k}-\overline{\mathbf{w}}_{t+E}\|^{2}=O(1/T^{2-2\beta})$,
-	which is at a slower rate than $1/T$ for any $\beta>0$. It follows
-	that $E=O(1)$. 
+can be removed, or equivalently whether $\sum_{k}p_{k}\|\mathbf{w}_{t}^{k}-\overline{\mathbf{w}}_{t}\|^{2}=O(1/T^{2})$
+can be independent of $E$. We provide a simple counterexample that
+shows that this is not possible in general. 
+\begin{lemma}
+	There exists a dataset such that if $E=\mathcal{O}(T^{\beta})$ for
+	any $\beta>0$ then $\sum_{k}p_{k}\|\mathbf{w}_{t}^{k}-\overline{\mathbf{w}}_{t}\|^{2}=\Omega(\frac{1}{T^{2-2\beta}})$
+	.
+\end{lemma}
+\begin{proof}
+	Suppose that we have an even number of devices and each $F_{k}(\mathbf{w})=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}(\mathbf{x}_{k}^{j}-\mathbf{w})^{2}$
+	contains data points $\mathbf{x}_{k}^{j}=\mathbf{w}^{\ast,k}$, with
+	$n_{k}\equiv n$. Moreover, the $\mathbf{w}{}^{\ast,k}$'s come in
+	pairs around the origin. As a result, the global objective $F$ is
+	minimized at $\mathbf{w}^{\ast}=0$. Moreover, if we start from $\overline{\mathbf{w}}_{0}=0$,
+	then by design of the dataset the updates in local steps exactly cancel
+	each other at each iteration, resulting in $\overline{\mathbf{w}}_{t}=0$
+	for all $t$. On the other hand, if $E=T^{\beta}$, then starting
+	from any $t=\mathcal{O}(T)$ with constant step size $\mathcal{O}(\frac{1}{T})$,
+	after $E$ iterations of local steps, the local parameters are updated
+	towards $\mathbf{w}^{\ast,k}$ with $\|\mathbf{w}_{t+E}^{k}\|^{2}=\Omega((T^{\beta}\cdot\frac{1}{T})^{2})=\Omega(\frac{1}{T^{2-2\beta}})$.
+	This implies that 
+	\begin{align*}
+	\sum_{k}p_{k}\|\mathbf{w}_{t+E}^{k}-\overline{\mathbf{w}}_{t+E}\|^{2} & =\sum_{k}p_{k}\|\mathbf{w}_{t+E}^{k}\|^{2}\\
+	& =\Omega(\frac{1}{T^{2-2\beta}})
+	\end{align*}
+	which is at a slower rate than $\frac{1}{T^{2}}$ for any $\beta>0$.
+	Thus the sampling variance $\mathbb{E}\|\overline{\mathbf{w}}_{t+1}-\overline{\mathbf{v}}_{t+1}\|^{2}=\Omega(\sum_{k}p_{k}\mathbb{E}\|\mathbf{w}_{t+1}^{k}-\overline{\mathbf{w}}_{t+1}\|^{2})$
+	decays at a slower rate than $\frac{1}{T^{2}}$, resulting in a convergence
+	rate slower than $O(\frac{1}{T})$ with partial participation. 
 \end{proof}

### 2020-08-07 rebuttal/main.tex
rebutal template

--- 2020-06-05 source/main.tex
@@
 
-% if you need to pass options to natbib, use, e.g.:
-%     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2020
+\usepackage{neurips_2019_author_response}
 
-% ready for submission
-% \usepackage{neurips_2020}
-
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
-%     \usepackage[preprint]{neurips_2020}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-%     \usepackage[final]{neurips_2020}
-
-% to avoid loading the natbib package, add option nonatbib:
-     \usepackage[nonatbib]{neurips_2020}
-\usepackage{slashbox,multirow}
-% \usepackage{tikz}
 \usepackage[utf8]{inputenc} % allow utf-8 input
@@
 \usepackage{microtype}      % microtypography
-\usepackage{amsthm}     % for theorems
-\usepackage{algorithm,algorithmic}
-\usepackage{mathtools}
-\usepackage{amsmath}
-\allowdisplaybreaks
-\usepackage{amssymb}
-\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
-\usepackage{balance}
-\usepackage{comment}
-\newcommand{\eq}[1]{{Eq~(#1)}}
-\newcommand{\figref}[1]{Figure~\ref{#1}}
-\newtheorem{theorem}{Theorem}
-\newtheorem{thm}{Theorem}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{remark}{Remark}
-\newcommand{\lkxcom}[1]{{\color{red}{#1}}}
-\usepackage{pifont}% http://ctan.org/pkg/pifont
-\newcommand{\cmark}{\ding{51}}%
-\newcommand{\xmark}{\ding{55}}%
-\def\lowcomma{_{\textstyle{,}}}
-\def\lowperiod{_{\textstyle.}}
+\usepackage{color}
+\usepackage{lipsum}
+\usepackage{wrapfig}
+\usepackage{graphicx}
+\usepackage{caption}
 
+\input{bmacros}
 \newcommand{\ov}[1]{{\overline{\mathbf{#1}}}}
-\input{bmacros}
-\newtheorem*{assumption*}{Assumption}
+\begin{document}
+The authors would like to thank three reviewers for their valuable feedback. 
+% Based on their comments, upon acceptance of this paper, we shall improve in 
+% the final version (a) Refine the presentation, fix typo, figures, and properly 
+% located assumptions etc. (b) More experiments in terms of 
+Our point-to-point response are provided below.
 
-\title{Federated Learning's Blessing:\\
-FedAvg has Linear Speedup}
+\input{r1}
+\input{r2}
+\input{r3}
 
-\author{}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-Federated learning (FL) learns a model jointly from a set of participating devices without sharing each other's privately held data.
-The characteristics of non-\textit{iid} data across the network, low device participation and the mandate that data remain private bring challenges in understanding the convergence of FL algorithms, particularly in regards to how convergence scales with the number of participating devices.
-In this paper, we focus on Federated Averaging (FedAvg)--the most widely used and effective FL algorithm in use today--and provide a comprehensive study of its convergence rate.
-Although FedAvg has recently been studied by an emerging line of literature, it remains open as to how FedAvg's convergence scales with the number of participating devices in the FL setting--a crucial question whose answer would shed light on the performance of FedAvg in large FL systems.
-We fill this gap by establishing convergence guarantees for FedAvg under two classes of problems: strongly convex smooth, convex smooth, and overparametrized strongly convex smooth problems.
-We show that FedAvg enjoys linear speedup in each case, although with different convergence rates.
-For each class, we also characterize the corresponding convergence rates for the Nesterov accelerated FedAvg algorithm in the FL setting: to the best of our knowledge, these are the first linear speedup guarantees for FedAvg when Nesterov acceleration is used.
-To accelerate FedAvg, we also design a new momentum-based FL algorithm that further improves the convergence rate in overparametrized linear regression problems.
-Empirical studies of the algorithms in various settings have supported our theoretical results.
-\end{abstract}
-
-% (and its related variant local SGD also analyzed in the simpler homogeneous distributed optimization setting)
-\input{sec_intro}
-%\input{sec_relatedwork}
-\input{sec_problem}
-\input{sec_SGD}
-\input{sec_Nesterov}
-\input{sec_overparameterized}
-\input{sec_exp}
-
-% \input{sec_conclude}
-
-\newpage
-\input{sec_impact}
-\bibliographystyle{unsrt}
-\bibliography{ref.bib}
-
-\newpage
-\appendix
-\input{sec_appendix}
 \end{document}
-
-\iffalse
-
-\endif

### 2020-08-07 rebuttal/r1.tex
rebutal template

{\color{blue}\textbf{Reviewer 1:}} We thank the reviewer for the valuable feedback. Your concerns are addressed as follows:

\textbf{\textit{1.Difference with [21].}} Compared to [21], the improvement 
of convergence comes from the improved bound of $2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]$ (as we mentioned in L489, Appendix C.1). Because of this improvement, we are able to cancel square of gradient norm ($\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$) in the upper bound of A1 (see L483) while in [21] this term remains as leading term ($6\eta_t^2L \Gamma$).

\textbf{\textit{2. The overparameterized case seem to be out of context here with no experimental results to back them up.}} The overparameterized 
setting is an important setting, especially in the era of deep learning
where the neural network can achieve zero training loss. We provide the 
geometric rate with linear speed up in this setting including a general overparameterized setting and overparameterized linear regression. 
Experimentally, we verify the linear speedup results of overparameterized setting in 3rd column, Figure 1 and Figure 2.

\textbf{\textit{3. The novelty and main results.}}

### 2020-08-07 rebuttal/r3.tex
rebutal template

{\color{blue}\textbf{Reviewer 3:}} We thank the reviewer for the valuable feedback. Your concerns are addressed as follows: \\
\textbf{\textit{1.Comparison with SCAFFOLD.}}: We thank the reviewer point out 
relavent reference, we will cite the paper in the new version.

\textbf{\textit{2. The convergence of Nestrov Accelerated FedAvg.}}

\textbf{\textit{3. The geometrical rates of overparametrized linear regression.}}

### 2020-08-08 rebuttal/r1.tex
Update r1.tex

@@
 
-\textbf{\textit{1.Difference with [21].}} Compared to [21], the improvement 
+\textbf{\textit{1. Improvement compared to [20] and [21].}} Compared to [21], the improvement 
 of convergence comes from the improved bound of $2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]$ (as we mentioned in L489, Appendix C.1).
-Because of this improvement, we are able to cancel square of gradient norm ($\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$) in the upper bound of A1 (see L483) while in [21] this term remains as leading term ($6\eta_t^2L \Gamma$).
+Because of this improvement, we are able to cancel square of gradient norm ($\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$) in the upper bound of A1 (defined in L483, also see explanations in L492) while in [21] this term remains as leading term ($6\eta_t^2L \Gamma$).
+
+Our work is not an extension of [20], we have discussed that our assumption is
+a more relaxed assumption comparing to the bounded gradient diversity in [20] (see Appendix B).
 
 \textbf{\textit{2. The overparameterized case seem to be out of context here with no experimental results to back them up.}} The overparameterized 
-setting is an important setting, especially in the era of deep learning
+setting is an important setting [29, 30], especially in the era of deep learning
 where the neural network can achieve zero training loss. We provide the 
@@
 \textbf{\textit{3. The novelty and main results.}}
+We would like to emphasize that the techniques used for different settings
+to achieve linear speedup are quite different (strongly convex see L489, L492, convex smooth see L553). The main contributions include
+the linear speedup results in strongly convex, convex smooth, and overparameterized
+setting, for both FedAvg and its accelerated variants (this is the first linear speedup results of accelerated FedAvg). We further propose FedMass,
+which is the first algorithm that enjoys an improved convergence rate comparing to
+FedAvg. We respectfully disagree that our main contribution is the improvement over [21] as the linear speedup results under other settings are not a trivial extension of the strongly convex case.

### 2020-08-08 rebuttal/r3.tex
r3 zhaonan

@@
 {\color{blue}\textbf{Reviewer 3:}} We thank the reviewer for the valuable feedback. Your concerns are addressed as follows: \\
-\textbf{\textit{1.Comparison with SCAFFOLD.}}: We thank the reviewer point out 
-relavent reference, we will cite the paper in the new version.
+\textbf{\textit{1.Comparison with SCAFFOLD.}}: 
+1.
+The analyses in SCAFFOLD do not imply $E=\cO(\sqrt{T})$ but in fact also require $E=O(1)$ under partial participation (sampling).
+To see this, it is easier to examine Theorem V in their paper, which is stated in terms of optimality gaps, but otherwise equivalent to their Theorem I.
+Notice that when $S<N$, the second term in $M^2$ gives rise to a term $O(1/R)$ in the bound for the strongly convex case (ignoring constants), and since $R=T/E$ in our notation, this term is $O(E/T)$.
+Therefore, in order to achieve a $O(1/T)$ convergence rate for the optimality gap, it must be the case that $E=O(1)$ as well.
+Similarly for the general convex case.
+Thus in terms of communication complexity, our results imply the same requirements for both full and partial participation as that in SCAFFOLD.
+
+2.
+Furthermore, the sampling procedure analyzed in our paper is different from that in SCAFFOLD, as we allow the sampling probability to scale with device specific weights and sample with replacement, whereas in SCAFFOLD the sampling is uniform without replacement.
+
+3.
+Under our sampling schemes, we explicitly analyze the contribution of sampling variance to the optimality gap, and in Lemma 10 of our paper in the Appendix, we provided a problem instance that lower bounds the sampling variance, showing that $E=O(1)$ cannot be improved in general when there is partial participation with sampling.
+
+4.
+As there has not been any linear speedup analysis of Nesterov accelerated FedAvg for convex problems, our result is a first in this regard and completes the picture.
+Our analyses of FedAvg and Nesterov FedAvg are also unified, highlighting the common elements and distinctions for the two algorithms, which has not been done by previous studies.
+
+5.
+Finally, our geometric rates in Theorem 5 are for general overparameterized strongly convex objectives rather than just linear regression, and as SCAFFOLD is a variance reduction based algorithm, our result, which is on the FedAvg algorithm, is not directly comparable to the geometric convergence result in SCAFFOLD.
 

### 2020-08-08 rebuttal/main.tex
update template to 2020
update

@@
 
-\usepackage{neurips_2019_author_response}
+\usepackage{neurips_2020_author_response}
 
@@
 \usepackage{caption}
-
+\usepackage{enumitem}
+\setlist{leftmargin=2.5mm}
 \input{bmacros}
@@
 
-\input{r1}
-\input{r2}
-\input{r3}
+% \input{r1}
+
+{\color{blue}\textbf{Reviewer 1:}} We thank the reviewer for the valuable feedback. Your concerns are addressed as follows:
+
+\textbf{\textit{1. Improvement compared to [20] and [21].}} Compared to [21], the improvement 
+of convergence comes from the improved bound of $2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]$ (as we mentioned in L489, Appendix C.1).
+Because of this improvement, we are able to cancel square of gradient norm ($\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$) in the upper bound of A1 (defined in L483, also see explanations in L492) while in [21] this term remains as leading term ($6\eta_t^2L \Gamma$).
+
+Our work is not an extension of [20], we have discussed that our assumption is
+a more relaxed assumption compared to the bounded gradient diversity in [20] (see L44 and Appendix B).
+
+\textbf{\textit{2. The overparameterized case seem to be out of context here with no experimental results to back them up.}} The overparameterized 
+setting is an important setting [29, 30], especially in the era of deep learning
+where the neural network can achieve zero training loss. We provide the 
+geometric rate with linear speed up in this setting including a general overparameterized setting and overparameterized linear regression. 
+Experimentally, we verify the linear speedup results of overparameterized setting in 3rd column, Figure 1 and Figure 2.
+
+\textbf{\textit{3. The novelty and main results.}}
+We would like to emphasize that the techniques used for different settings
+to achieve linear speedup are quite different (strongly convex see L489, L492, convex smooth see L553). The main contributions include
+the linear speedup results in strongly convex, convex smooth, and overparameterized
+setting, for both FedAvg and its accelerated variants (this is the first linear speedup results of accelerated FedAvg). We further propose FedMass,
+which is the first algorithm that enjoys an improved convergence rate comparing to
+FedAvg. We respectfully disagree that our main contribution is the improvement over [21] as the linear speedup results under other settings are not a trivial extension of the strongly convex case.
+
+% \input{r2}
+{\color{blue}\textbf{Reviewer 2:}} We thank the reviewer for the valuable feedback. Your concerns are addressed as follows: \\
+\textbf{\textit{1.The range of N/K.}}
+
+% \input{r3}
+
+{\color{blue}\textbf{Reviewer 3:}} We thank the reviewer for the valuable feedback. Your concerns are addressed as follows: \\
+\textbf{\textit{1.Comparison with SCAFFOLD.}}: 
+\begin{itemize}
+	\item The analyses in SCAFFOLD do not imply $E=\cO(\sqrt{T})$ but in fact also require $E=O(1)$ under partial participation (sampling).
+To see this, it is easier to examine Theorem V in their paper, which is stated in terms of optimality gaps, but otherwise equivalent to their Theorem I.
+Notice that when $S<N$, the second term in $M^2$ gives rise to a term $O(1/R)$ in the bound for the strongly convex case (ignoring constants), and since $R=T/E$ in our notation, this term is $O(E/T)$.
+Therefore, in order to achieve a $O(1/T)$ convergence rate for the optimality gap, it must be the case that $E=O(1)$ as well.
+Similarly for the general convex case.
+Thus in terms of communication complexity, our results imply the same requirements for both full and partial participation as that in SCAFFOLD.
+	\item Furthermore, the sampling procedure analyzed in our paper is different from that in SCAFFOLD, as we allow the sampling probability to scale with device specific weights and sample with replacement, whereas in SCAFFOLD the sampling is uniform without replacement.
+	\item  Under our sampling schemes, we explicitly analyze the contribution of sampling variance to the optimality gap, and in Lemma 10 of our paper in the Appendix, we provided a problem instance that lower bounds the sampling variance, showing that $E=O(1)$ cannot be improved in general when there is partial participation with sampling.
+\end{itemize}
+
+\textbf{\textit{2. The convergence of Nestrov Accelerated FedAvg.}}
+As there has not been any linear speedup analysis of Nesterov accelerated FedAvg for convex problems, our result is a first in this regard and completes the picture.
+Our analyses of FedAvg and Nesterov FedAvg are also unified, highlighting the common elements and distinctions for the two algorithms, which has not been done by previous studies.
+
+\textbf{\textit{3. The geometrical rates of overparametrized linear regression.}}
+The geometric rates in Theorem 5 are for general overparameterized strongly convex objectives rather than just linear regression, and as SCAFFOLD is a variance reduction based algorithm, our result, which is on the FedAvg algorithm, is not directly comparable to the geometric convergence result in SCAFFOLD.
 

### 2020-08-10 rebuttal/acresponse.tex
Update on Overleaf.

\documentclass{article}

\usepackage{neurips_2020_author_response}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{color}
\usepackage{lipsum}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{enumitem}
\setlist{leftmargin=2mm}
\input{bmacros}
\newcommand{\ov}[1]{{\overline{\mathbf{#1}}}}
\begin{document}
Dear Area Chair,

We would like to emphasize that we respectfully disagree with the evaluation from R1 and R3. 
R1 mainly challenge the writing, contributions, and comparison with prior work. Regarding the writing, the other reviewers think the paper is well-written and is easy to follow. Regarding the contributions and comparison with prior work, we explicitly summarized those in the table and did thorough comparison in the appendix.  We also explicitly mention what is the improvement leads to the linear speedup comparing to prior prove in the appendix. 
R1 might think our work is an extension of [21]. This is correct for the strongly convex case. (and we did specify the non-trivial step compared to [21] in the appendix (L483, L492). However,  the general convex and overparameterized setting as well as the Nesterov results use very different techniques and it is not fair to require a simple highlight of non-trivial step compared to prior work [20], as discussed in Appendix B, we use a more general assumption than [20].

R3 mentioned a missing reference (SCAFFOLD) while his/her conclusions of that paper are not correct. The missing reference’s analysis for FedAvg (c.f. Theorem V) didn’t achieve linear speedup in general, as we discuss in detail in our response to R3. On the other hand, our results achieve linear speedup under more flexible sampling strategies for any number of participating devices. Also,  SCAFFOLD didn’t achieve better communication complexity as R3 claimed. We have the same communication complexity and we guarantee the linear speedup.

In light of the above discussions, we kindly request the area chair to re-consider the evaluation from R1 and R3.

\end{document}

### 2020-09-09 rebuttal/main.tex

--- 2020-08-07 rebuttal/main.tex
@@
 \usepackage{caption}
-
+\usepackage{enumitem}
+\setlist{leftmargin=2mm}
 \input{bmacros}
@@
 \begin{document}
-The authors would like to thank three reviewers for their valuable feedback. 
+The authors would like to thank reviewers for their valuable feedback. 
 % Based on their comments, upon acceptance of this paper, we shall improve in 
@@
 
-\input{r1}
-\input{r2}
-\input{r3}
+% \input{r1}
+\vspace{-1.5mm}
+{\color{blue}\textbf{Reviewer 1:}} 
+\vspace{-0.4mm}
+\textbf{\textit{1. Improvement compared to [21].}} Compared to [21], the improvement 
+of convergence comes from the improved bound of {\small $2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]$} (discussed in L489, Appendix C.1).
+Because of this improvement, we are able to cancel square of gradient norm ($\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$) in the upper bound of A1 (defined in L483, also see explanations in L492) while in [21] this term remains as leading term ($6\eta_t^2L \Gamma$), hence a significant improvement.
+We agree with you this part of the appendix should be brought to the main text and carefully explained to highlight our contributions.
+\vspace{-1.5mm}
+
+\textbf{\textit{2. Improvement compared to [20].}}
+Our work goes much beyond [20], which studies the highly special case where each local device’s minimizer is the same as the minimizer of the joint problem with all devices' data aggregated (this is referred to as bounded gradient diversity (BGD) in [20], see L44 and Appendix B), which never occurs in practice .
+\vspace{-1.5mm}
+
+\textbf{\textit{3. The overparameterized case seem to be out of context here with no experimental results to back them up.}} The overparameterized 
+setting is an important setting [29, 30], especially in the era of deep learning
+where the neural network can achieve zero training loss. We provide the 
+geometric rate with linear speedup in this setting including a general overparameterized setting and overparameterized linear regression. 
+Experimentally, we verify the linear speedup results of overparameterized setting in 3rd column, Figure 1 and Figure 2. 
+\vspace{-1.5mm}
+
+\textbf{\textit{4. The novelty and main results.}}
+We would like to emphasize that the techniques used for different settings
+to achieve linear speedup are quite different (strongly convex see L489, L492, convex smooth see L553). The main contributions include
+the linear speedup results in strongly convex, convex smooth, and overparameterized
+setting, for both FedAvg and its accelerated variants (this is the first linear speedup results of Nesterov accelerated FedAvg for convex problems). We further propose FedMass,
+which is the first momentum-based acceleration algorithm that enjoys an improved convergence rate compared to
+FedAvg. We respectfully disagree that our main contribution is the improvement over [21] as the linear speedup results under other settings are not trivial extensions of the strongly convex case.
+\vspace{-1.5mm}
+
+\textbf{\textit{5. Writing.}} 
+We will improve the organization of the paper upon acceptance, including bringing back certain proofs in appendix to highlight contributions.
+% As you pointed out, the organization of the paper can be improved.
+Certain proofs currently in the appendix should be brought back and clearly exposed to highlight the contributions.
+We will do so in the camera-ready.
+\vspace{-1.5mm}
+
+% \input{r2}
+{\color{blue}\textbf{Reviewer 2:}} 
+\textbf{\textit{1.The valid range of the number of (participating) devices--(K) N for linear speedup.}} We agree that the valid range of $N$ and $K$ that allows linear speedup is an important issue, although it is straightforward to make it explicit: we just need to make sure all terms in the bound are of the same order (and have linear speedup).
+For example, in the case of general convex problems with partial participation (Theorem 2), with local step $E=\cO(1)$, as long as $K=O(T^{1/3})$ all three terms in the bound are $\cO(1/\sqrt{KT})$.
+Meanwhile, previous literature discussing linear speedup e.g.
+[15, 17] also did not provide explicit discussion of range of $N$ and $K$.
+% Similar to  the literature discussing linear speedup [15, 17], we didn't provide explicit result and we admit it worth further exploration.
+To give a more concrete answer, in general convex case with partial participation, we have local step $E=\cO(1)$, we can determine the optimal $K$ is {\small $\left( \sqrt{T}(\gamma_{\max}\delta^2 + G^2)/G^2L\right)^{2/3}$} as long as  $K \leq N$.
+This provides
+% an upper bound of $K$ that we have no benefit to further increase $K$, which might serve as a valid range of $K$.
+
+% While we could have stated the dependence of $N$ on $T$ and $E$, in order to better illustrate the communication complexity, we decided to equivalently state the dependence of $E$ on $N$ and $T$ instead in discussions following our results.
+We agree that for the optimal choice of $E$, we could be more explicit about the range of $N$.
+%%e require $N\leq T$, and with larger $N$, we achieve faster convergence but communicate more as $E$ decreases.
+With partial participation, we have local step $E=\cO(1)$, and the optimal $K$ is $\left( \frac{\sqrt{T}(\gamma_{\max}\delta^2 + G^2)}{2G^2L}\right)^{2/3}$ as long as  $K \leq N$.
+
+% \textbf{\textit{2.Explicit constant dependence.}}
+% We agree that including the constants makes the results more precise.
+% \input{r3}
+\vspace{-1.5mm}
+
+{\color{blue}\textbf{Reviewer 3:}} Thank you for the valuable feedback and pointing out the missing reference, we will cite the paper in the new version.
+Also, there are substantial differences between our work and SCAFFOLD, which we hope to clarify here.
+\\
+\textbf{\textit{1.Comparison with SCAFFOLD.}}
+% \vspace{-1.2em}
+% \begin{itemize}
+	\textbf{We prove linear speedup for any number of devices for FedAvg while SCAFFOLD didn't.} SCAFFOLD's results for FedAvg do not enjoy linear speedup for any number of devices while our results apply to any valid $K\leq N$.
+Under our notation, i.e.
+$N$: number of devices; $K$: number of participating devices; $T$: iteration number; $E$: number of local steps, Theorem V of the SCAFFOLD paper provides an optimality gap of {\small $\cO\left((1-\frac{K}{N})E/T\right)$} for the strongly convex case.
+% 	To achieve linear speedup, it requires $1-\frac{K}{N} = 1/K$. Therefore, the activated devices number needs to satisfy {\small $K =
+% \cO\left((1\pm\sqrt{1-4/N})N/2\right)$} for linear speedup, which is much more restricted than our results. The same conclusion applies for general convex case.
+With partial participation, and when $K=O(1)$, their convergence rate is $\cO(E/T)$ which does not have linear speedup, while our optimality gap bound has linear speedup for any valid number of participating devices.
+(See Theorems 1-4 in our paper.)
+	\textbf{We achieve the same communication complexity.}
+	Under partial participation, the FedAvg analyses in SCAFFOLD do not imply $E=\cO(\sqrt{T})$, but requires $E=\cO(1)$.
+For example, the strongly convex result {\small $\cO((1-\frac{K}{N})E/T)$} in Theorem V is $\cO(E/T)$ when $K=O(1)$ and is $\cO(E/KT)$ when $K=O(N)$.
+	In either case, to achieve a $\cO(1/T)$ convergence rate, it requires $E=\cO(1)$ as well. Similarly for the general convex case. 
+% \textbf{We achieve the same communication complexity.} Under partial participation (sampling), the analyses in SCAFFOLD do not imply $E=\cO(\sqrt{T})$, but in fact also require $E=\cO(1)$.
+To see this, it is easier to examine Theorem V in their paper, which is stated in terms of optimality gaps (same as our paper), but otherwise equivalent to their Theorem I.
+Notice that when $S<N$, the second term in $M^2$ gives rise to a term $\cO((1-\frac{S}{N})/R)$ in the bound for the strongly convex case, and since $R=T/E$ in our notation, this term is $\cO(E/T)$ when $S=O(1)$ and is $\cO(E/NT)$ when $S=O(N)$.
+Therefore, in order to achieve a $\cO(1/T)$ convergence rate for the optimality gap, it must be the case that $E=\cO(1)$ as well.
+Similarly for the general convex case.
+Thus in terms of communication complexity, our results imply the same requirements for both full and partial participation as that in SCAFFOLD.
+In addition, their result does not have linear speedup when $S$ is small, while ours does.
+% 	\vspace{-0.5em}
+	\textbf{We allow more flexible sampling schemes.} Furthermore, the sampling procedure analyzed in our paper is more general compared to SCAFFOLD, as we allow the sampling probability to scale with device specific weights and sample with replacement, whereas in SCAFFOLD the sampling is uniform without replacement.
+The sampling procedure is important and different schemes could imply different bounds with different proof techniques.
+% 	\vspace{-0.5em}
+	\textbf{Our bound is tight.} Under our sampling schemes, we explicitly analyze the contribution of sampling variance to the optimality gap.
+We provided a problem instance that lower bounds the sampling variance, showing that $E=\cO(1)$ cannot be improved in general when there is partial participation with our sampling procedure (see Lemma 10).
+% \end{itemize}
+% \vspace{-0.5em}
+\vspace{-1.5mm}
+
+\textbf{\textit{2. The first and unified linear speedup results of Nestrov Accelerated FedAvg.}}
+Nesterov Accelerated FedAvg has been studied by prior arts [16,22,35].
+However, there has not been any linear speedup analysis of Nesterov accelerated FedAvg for convex problems.
+Our result is the first to complete the picture.
+Our paper is also the first to provide a unified analysis for FedAvg and Nesterov accelerated FedAvg for convex problems, highlighting the common elements and distinctions for the two algorithms, whereas previous analyses of the two algorithms only exist in the non-convex case (only with full participation) [17] and are scattered and use different notation and proof techniques [16, 17].
+\\
+% \vspace{-0.5em}
+\textbf{\textit{3. The first geometrical rates of general overparametrized strongly convex problem for FedAvg.}}
+The geometric rates in Theorem 5 are for general overparameterized strongly convex objectives rather than just linear regression, and as SCAFFOLD is a variance reduction-based algorithm, our result, which is on the FedAvg algorithm, is not directly comparable to the geometric convergence result for SCAFFOLD.
 

### 2020-09-09 source/r1.tex

--- 2020-08-08 rebuttal/r1.tex
@@
 Our work is not an extension of [20], we have discussed that our assumption is
-a more relaxed assumption comparing to the bounded gradient diversity in [20] (see Appendix B).
+a more relaxed assumption compared to the bounded gradient diversity in [20] (see L44 and Appendix B).
 

### 2020-09-09 source/r3.tex

--- 2020-08-07 rebuttal/r3.tex
@@
 {\color{blue}\textbf{Reviewer 3:}} We thank the reviewer for the valuable feedback. Your concerns are addressed as follows: \\
-\textbf{\textit{1.Comparison with SCAFFOLD.}}: We thank the reviewer point out 
-relavent reference, we will cite the paper in the new version.
+\textbf{\textit{1.Comparison with SCAFFOLD.}}: 
+\begin{itemize}
+	\item The analyses in SCAFFOLD do not imply $E=\cO(\sqrt{T})$ but in fact also require $E=O(1)$ under partial participation (sampling).
+To see this, it is easier to examine Theorem V in their paper, which is stated in terms of optimality gaps, but otherwise equivalent to their Theorem I.
+Notice that when $S<N$, the second term in $M^2$ gives rise to a term $O(1/R)$ in the bound for the strongly convex case (ignoring constants), and since $R=T/E$ in our notation, this term is $O(E/T)$.
+Therefore, in order to achieve a $O(1/T)$ convergence rate for the optimality gap, it must be the case that $E=O(1)$ as well.
+Similarly for the general convex case.
+Thus in terms of communication complexity, our results imply the same requirements for both full and partial participation as that in SCAFFOLD.
+	\item Furthermore, the sampling procedure analyzed in our paper is different from that in SCAFFOLD, as we allow the sampling probability to scale with device specific weights and sample with replacement, whereas in SCAFFOLD the sampling is uniform without replacement.
+	\item  Under our sampling schemes, we explicitly analyze the contribution of sampling variance to the optimality gap, and in Lemma 10 of our paper in the Appendix, we provided a problem instance that lower bounds the sampling variance, showing that $E=O(1)$ cannot be improved in general when there is partial participation with sampling.
+\end{itemize}
 
 \textbf{\textit{2. The convergence of Nestrov Accelerated FedAvg.}}
+As there has not been any linear speedup analysis of Nesterov accelerated FedAvg for convex problems, our result is a first in this regard and completes the picture.
+Our analyses of FedAvg and Nesterov FedAvg are also unified, highlighting the common elements and distinctions for the two algorithms, which has not been done by previous studies.
 
 \textbf{\textit{3. The geometrical rates of overparametrized linear regression.}}
+The geometric rates in Theorem 5 are for general overparameterized strongly convex objectives rather than just linear regression, and as SCAFFOLD is a variance reduction based algorithm, our result, which is on the FedAvg algorithm, is not directly comparable to the geometric convergence result in SCAFFOLD.

### 2020-09-09 source-iclr/app_sgd_scvx_smooth.tex
iclr from nips version 1.

--- 2020-06-11 source/app_sgd_scvx_smooth.tex
@@
 \begin{proof}
-	The first step is to observe that the $L$-smoothness of $F$ provides
+	The proof builds on ideas from~\cite{li2019convergence}. The first step is to observe that the $L$-smoothness of $F$ provides
 	the upper bound
@@
 
-	Now we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$.
+	Now we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$ following \cite{li2019convergence}.
 	Since communication is done every $E$ steps, for any $t\geq0$, we

### 2020-09-09 source-iclr/sec_Nesterov.tex
iclr from nips version 1.

--- 2020-06-04 source/sec_Nesterov.tex
@@
 \end{comment}
-The Nesterov Accelerated FedAvg algorithm follows the updates
+The Nesterov Accelerated FedAvg algorithm follows the updates:
 \begin{align*}
@@
 \mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E},\\
-\sum_{k \in \cS_{t+1}}\left[\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E}.
+\sum_{k \in \cS_{t+1}}\left[\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E},
 \end{cases}
@@
 accelerated variant. In particular, to achieve $\mathcal{O}(1/NT)$
-linear speedup, $T$ iterations of the algorithm requires only $\mathcal{O}(\sqrt{NT})$
+linear speedup, $T$ iterations of the algorithm require only $\mathcal{O}(\sqrt{NT})$
 communication rounds with full participation.

### 2020-09-09 source-iclr/sec_SGD.tex
iclr from nips version 1.

--- 2020-06-04 source/sec_SGD.tex
@@
 
-We first show that FedAvg has $\mathcal{O}(1/NT)$ convergence rate
+We first show that FedAvg has an $\mathcal{O}(1/NT)$ convergence rate
 for $\mu$-strongly convex and $L$-smooth objectives. The result
@@
 implies a distinction in communication efficiency that guarantees
-this linear speedup for FedAvg with full and partial device participations.
+this linear speedup for FedAvg with full and partial device participation.
 With full participation, $E$ can be chosen as large as $\mathcal{O}(\sqrt{T/N})$
@@
 As a result, all leading terms scale with $1/N$ in the full device
-particiaption setting, and with $1/K$ in the partial participation
+participation setting, and with $1/K$ in the partial participation
 setting. This implies that in both settings, there is a \emph{linear
-	speedup} in the number of active workers during the communication
-round. \\
+	speedup} in the number of active workers during a communication
+round.
+We also emphasize that the reason one cannot recover the full participation bound by setting $K=N$ in the partial participation bound is due to the variance generated by sampling which depends on $E$.
+\\
 % To our knowledge, this is the first result that explicitly demonstrates
@@
 still yields a linear speedup in $K$, which is also confirmed by
-experiments. The $E=\mathcal{O}(1)$ requirement in partial participation
+experiments. The requirement $E=\mathcal{O}(1)$ in partial participation
 likely cannot be removed for our sampling schemes, as the sampling
 variance is $\Omega(E^{2}/T^{2})$ and the dependence on $E$ is tight.  \\
-\textbf{Comparison with related works.} To better understand the significance of the obtained bound, we compare our rates to the best-known results in related settings.~\cite{haddadpour2019convergence} proves a linear speedup $\mathcal{O}(1/NT)$ result for strongly convex and smooth objectives, with $\mathcal{O}(N^{1/3}T^{2/3})$ communication complexity with i.i.d.
-data and partial participation.
+\textbf{Comparison with related work.} To better understand the significance of the obtained bound, we compare our rates to the best-known results in related settings.
+\cite{haddadpour2019convergence} proves a linear speedup $\mathcal{O}(1/NT)$ result for strongly convex and smooth objectives, with $\mathcal{O}(N^{1/3}T^{2/3})$ communication complexity with \emph{i.i.d.} data and partial participation.
 However, their results build on the bounded gradient diversity assumption, which implies the existence of $\mathbf{w}^*$ that minimizes all local objectives (see discussions in Section~\ref{sec:assumptions}), effectively removing system heterogeneity.
-The bound in \cite{koloskova2020unified} matches our bound in the full participation case, but their framework excludes partial particiation \cite[Proposition 1]{koloskova2020unified}.
+The bound in \cite{koloskova2020unified} matches our bound in the full participation case, but their framework excludes partial participation \cite[Proposition 1]{koloskova2020unified}.
 
@@
 convex case the learning rate decays as $\mathcal{O}(1/t)$. Such
-a requirement $\alpha_{t}=\mathcal{O}(\sqrt{N/T})$ is also present
-in the works \cite{haddadpour2019convergence,yu2019parallel} on non-convex
+a requirement $\alpha_{t}=\mathcal{O}(\sqrt{N/T})$ also presents
+in \cite{haddadpour2019convergence,yu2019parallel} on non-convex
 problems with $\mathcal{O}(1/\sqrt{NT})$ linear speedup convergence results. 

### 2020-09-09 source-iclr/sec_exp.tex
iclr from nips version 1.

--- 2020-06-03 source/sec_exp.tex
@@
 
-In this section, we empirically examine the linear speedup convergence of FedAvg and Nesterov accelerated FedAvg on various scenarios, including strongly convex function, convex smooth function, and overparameterized setting, as proved in previous theorems.
+In this section, we empirically examine the linear speedup convergence of FedAvg and Nesterov accelerated FedAvg in various settings, including strongly convex function, convex smooth function, and overparameterized objectives, as analyzed in previous sections.
 
-\textbf{Setup.} Follow the experimental setting in~\cite{stich2018local}, we
+\textbf{Setup.} Following the experimental setting in~\cite{stich2018local}, we
 conduct experiments on both synthetic datasets and real-world dataset
 w8a~\cite{platt1998fast} $(d=300, n=49749)$. We consider the distributed
-objectives $F(\vw) = \sum_{k=1}^N p_kF_k(\vw)$, and the objective function on
+objectives $F(\vw) = \sum_{k=1}^N p_kF_k(\vw)$, and the objective function on the
 $k$-th local device includes three cases: 1) \textbf{Strongly convex
@@
 % ~\ref{th:strongcvxsmth}~\ref{th:cvxsmoth}~\ref{th:fedmasslr}
-\textbf{Linear speedup of FedAvg and Nesterov accelerated FedAvg.} To verify the linear speedup convergence as shown in Theorem~\ref{thm:SGD_scvx}~\ref{thm:SGD_cvx}~\ref{thm:nesterov_scvx}~\ref{thm:Nesterov_cvx}, we evaluate the number of iterations needed to reach
+\textbf{Linear speedup of FedAvg and Nesterov accelerated FedAvg.} To verify the linear speedup convergence as shown in Theorems~\ref{thm:SGD_scvx}~\ref{thm:SGD_cvx}~\ref{thm:nesterov_scvx}~\ref{thm:Nesterov_cvx}, we evaluate the number of iterations needed to reach
 $\epsilon$-accuracy in three objectives.
@@
 the number of iterations decreases as the number of (active) workers increasing, which is consistent for FedAvg and Nesterov accelerated FedAvg across all scenarios.
-For the additional experiments on the impact of $E$, detailed experimental setup, and hyperparameter setting, please refer to the Appendix, section~\ref{sec:expsupp}.
+For additional experiments on the impact of $E$, detailed experimental setup, and hyperparameter setting, please refer to the Appendix Section~\ref{sec:expsupp}.
 

### 2020-09-09 source-iclr/sec_intro.tex
iclr from nips version 1.

--- 2020-06-05 source/sec_intro.tex
@@
 \section{Introduction}
-Federated learning (FL) is a machine learning setting where many clients (e.g.
-mobile devices or organizations) collaboratively train a model under the orchestration of a central server (e.g.
-service provider), while keeping the training data decentralized~\cite{smith2017federated, kairouz2019advances}.
+Federated learning (FL) is a machine learning setting where many clients (e.g., mobile devices or organizations) collaboratively train a model under the orchestration of a central server (e.g., service provider), while keeping the training data decentralized~\cite{smith2017federated, kairouz2019advances}.
 In recent years, FL has swiftly emerged as an important learning paradigm~\cite{mcmahan2016communication,li2018federated}--one that enjoys widespread success in applications such as personalized recommendation~\cite{chen2018federated}, virtual assistant~\cite{lamautonomy}, and keyboard prediction~\cite{47586}, to name a few--for at least two reasons: First, the rapid proliferation of smart devices that are equipped with both computing power and data-capturing capabilities provided the infrastructure core for FL.
@@
 
-However, despite these very recent fruitful pioneering efforts into understanding the theoretical convergence properties of FedAvg, it remains open as to how the number of devices--particularly the number of devices that participate in the computation--affect the convergence speed.
+However, despite these very recent fruitful pioneering efforts into understanding the theoretical convergence properties of FedAvg, it remains open as to how the number of devices--particularly the number of devices that participate in the computation--affects the convergence speed.
 In particular, do we get linear speedup of FedAvg? What about when FedAvg is accelerated? These aspects are currently unexplored in FL. We fill in the gaps here by providing affirmative answers.

### 2020-09-09 source-iclr/sec_intro_old.tex
iclr from nips version 1.

% !TEX ROOT=./main.tex

\section{Introduction and Related Works}

\begin{table}[h!]
	\centering
	\small
	\begin{tabular}{|c|c|c|c|c|}\hline
		paper         &  Cvx Non-smth & Cvx L-smth & Strongly Cvx Non-smth& Strongly Cvx L-smth \\ \hline
		Accelerated	SGD   &   $\cO(\frac{1}{\sqrt{T}})$     & $\cO(\frac{1}{\sqrt{KT}})$   &    $\cO(\frac{1}{T})$   &  $\cO(\frac{1}{KT}+\frac{E^2}{T^2})$    \\\hline
		SGD    &    $\cO(\frac{1}{\sqrt{T}})$  &   $\cO(\frac{1}{\sqrt{KT}})$  &  $\cO(\frac{1}{T})$  & $\cO(\frac{1}{T})$\cite{li2019convergence,haddadpour2019convergence}, $\cO(\frac{1}{KT}+\frac{E^2}{T^2})$ ours    \\\hline
	\end{tabular}
	\caption{Summary of related work on federated learning, heterogeneous data.}
\end{table}

+ exponential acceleration of MaSS for linear regression: for $E=1$, 
\begin{align*}
\|\overline{w}_{t+1}-w^{\ast}\|^{2} & \leq C\cdot\exp(-\frac{t}{\sqrt{(\kappa_{N})(\tilde{\kappa}_{N})}})\leq C\cdot\exp(-\frac{Nt}{\sqrt{(\kappa_{1})(\tilde{\kappa})}})
\end{align*}

Most recent result by Koloskova et. al gives same bounds for convex and strongly convex.

### 2020-09-09 source-iclr/sec_sgd_scvx_smooth.tex
iclr from nips version 1.

--- 2020-04-19 source/sec_sgd_scvx_smooth.tex
@@
-We show that FedAvg with SGD has $O(1/NT)$ rate under $\mu$-strong
-convexity and $L$-smoothness. The result improves on the ICLR paper's
-$O(1/T)$ with a linear speedup in the number of devices $N$. {[}TODO:
-Add more comment here.{]} The FedAv algorithm with SGD follows the
-updates
-\begin{align*}
-y_{t+1}^{k} & =w_{t}^{k}-\alpha_{t}g_{t,k}\\
-w_{t+1}^{k} & =\begin{cases}
-y_{t+1}^{k} & \text{if }t+1\notin\mathcal{I}_{E}\\
-\sum_{k=1}^{N}p_{k}y_{t+1}^{k} & \text{if }t+1\in\mathcal{I}_{E}
-\end{cases}
-\end{align*}
-where 
-\begin{align*}
-g_{t,k} & :=\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})
-\end{align*}
-is the stochastic gradient.
-
-Define the virtual sequences $\overline{y}_{t}=\sum_{k=1}^{N}p_{k}y_{t}^{k}$,
-$\overline{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$, and $\overline{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$.
-We have $\mathbb{E}g_{t}=\overline{g}_{t}$ and $\overline{y}_{t+1}=\overline{w}_{t}-\alpha_{t}g_{t}$,
-and $\overline{w}_{t+1}=\overline{y}_{t+1}$ for all $t$.
-
 We first summarize some properties of $L$-smooth and $\mu$-strongly
 convex functions. 
+\begin{lemma}
+	Let $F$ be a convex $L$-smooth function. Then we have the following
+	inequalities:
 
+	1. Quadratic upper bound: $0\leq F(w)-F(w')-\langle\nabla F(w'),w-w'\rangle\leq\frac{L}{2}\|w-w'\|^{2}$.
+
+	2. Coercivity: $\frac{1}{L}\|\nabla F(w)-\nabla F(w')\|^{2}\leq\langle\nabla F(w)-\nabla F(w'),w-w'\rangle$.
+
+	3. Lower bound: $F(w)\geq F(w')+\langle\nabla F(w'),w-w'\rangle+\frac{1}{2L}\|\nabla F(w)-\nabla F(w')\|^{2}$.
+	In particular, $\|\nabla F(w)\|^{2}\leq2L(F(w)-F(w^{\ast}))$.
+
+	4. Optimality gap: $F(w)-F(w^{\ast})\leq$$\langle\nabla F(w),w-w^{\ast}\rangle$.
+\end{lemma}
+%
+\begin{lemma}
+	Let $F$ be a $\mu$-strongly convex function. Then 
+	\begin{align*}
+	F(w) & \leq F(w')+\langle\nabla F(w'),w-w'\rangle+\frac{1}{2\mu}\|\nabla F(w)-\nabla F(w')\|^{2}\\
+	F(w)-F(w^{\ast}) & \leq\frac{1}{2\mu}\|\nabla F(w)\|^{2}
+	\end{align*}
+\end{lemma}
+%
 \begin{assumption}
@@
 	so that $\alpha_{t}\leq2\alpha_{t+E}$, and where $0\leq c\leq1$
-	is small enough such that the following hold: 
+	is small enough such that
+	$\alpha_{t} \leq\frac{1}{4L}$ and 
+	$\alpha_{t} \leq\frac{1}{N}$ for all $t\geq0$.
+
+	Then with full device participation, if $\nu_{max}=\max_{k}\frac{1}{N}p_{k}$,
+	we have 
 	\begin{align*}
-	\alpha_{t} & \leq\frac{1}{4L}\\
-	\alpha_{t} & \leq\frac{1}{N}
+	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq C\frac{\kappa}{\gamma+T}\cdot(8E^{2}LG^{2}/T+\frac{1}{N}\nu_{max}^{2}\sigma^{2})
 	\end{align*}
-	for all $t\geq0$.
-
-	Then with full device participation, as long as $E\leq O(\sqrt{\frac{N}{T}})$,
+	so that as long as $E=O(\sqrt{\frac{N}{T}})$, 
 	\begin{align*}
-	\mathbb{E}F(w_{T})-F^{\ast} & \leq C\frac{\kappa}{N(\gamma+t)}\\
-	C & =5LG^{2}+\nu_{max}^{2}\sigma^{2}
+	\mathbb{E}F(w_{T})-F^{\ast} & \leq C(8LG^{2}+\nu_{max}^{2}\sigma^{2})\frac{\kappa}{N(\gamma+T)}
 	\end{align*}
-	and with partial participation, with $E=O(1)$ and $K$ sampled devices
-	at each communication round, 
+	and with partial participation with $K$ sampled devices at each
+	communication round, $\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}$
 	\begin{align*}
-	\mathbb{E}F(w_{T})-F^{\ast} & \leq(4E^{2}G^{2}+\nu_{max}^{2}\sigma^{2})\frac{\kappa}{K(\gamma+t)}+\frac{4E^{2}LG^{2}}{t^{2}}
+	\mathbb{E}F(\overline{w}_{T})-F^{\ast}\leq C\frac{\kappa}{\gamma+T}\cdot(8E^{2}LG^{2}/T+\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{4}{K}E^{2}G^{2})
+	\end{align*}
+	so that if $E=O(1)$, 
+	\begin{align*}
+	\mathbb{E}F(\overline{w}_{T})-F^{\ast} & \leq C(4E^{2}G^{2}+\nu_{max}^{2}\sigma^{2})\frac{\kappa}{K(\gamma+T)}+\frac{8E^{2}LG^{2}}{T^{2}}
 	\end{align*}
 \end{theorem}
+
 \begin{proof}
@@
 	can be canceled by an upper bound of the second term. %
+	\begin{comment}
+	The last term is straightforward to bound by the convexity of $\|\cdot\|^{2}$
+	and $L$-smoothness of $F_{k}$,
+	\begin{align*}
+	\alpha_{t}^{2}\|\overline{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(w_{t}^{k})\|^{2}\leq2L\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}(F_{k}(w_{t}^{k})-F_{k}^{\ast})
+	\end{align*}
+	or 
+	\begin{align*}
+	\alpha_{t}^{2}\|\overline{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(w_{t}^{k})\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\mathbb{E}\|\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})\|^{2}\leq\alpha_{t}^{2}G^{2}
+	\end{align*}
+	\end{comment}
 
@@
 	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-w^{\ast},\nabla F_{k}(w_{t}^{k})\rangle\\
-	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(w^{\ast})-F_{k}(w_{t}^{k}))\\
-	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w_{t}^{k})-F_{k}(\overline{w}_{t})+\frac{L}{2}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+F_{k}(w^{\ast})-F_{k}(w_{t}^{k})\right]\\
-	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]
+	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\overline{w}_{t}-w_{t}^{k},\nabla F_{k}(w_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(w^{\ast})-F_{k}(w_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|w_{t}^{k}-w^{\ast}\|^{2}\\
+	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w_{t}^{k})-F_{k}(\overline{w}_{t})+\frac{L}{2}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+F_{k}(w^{\ast})-F_{k}(w_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}w_{t}^{k}-w^{\ast}\|^{2}\\
+	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\overline{w}_{t}-w_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(w^{\ast})-F_{k}(\overline{w}_{t})\right]-\alpha_{t}\mu\|\overline{w}_{t}-w^{\ast}\|^{2}
 	\end{align*}
@@
 	\end{align*}
-	where $\sigma^{2}=\max_{k}\sigma_{k}^{2}$. The leading term in the
-	bound is $\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}$, which
-	one summed using the recursive relations gives $O(\alpha_{t})$, but
-	the constant now is scaled down by $1/N$. To achieve this linear
-	speedup, we need $\sum_{t=0}^{T}(1-\mu\alpha_{t})\|w_{0}-w^{\ast}\|^{2}$
-	to be small enough. \textbf{Note also that the above bound implies
-		that $E$ can be chosen to be as large as $O(\sqrt{\frac{T}{N}})$
-		without degrading the performance, since $E^{2}\alpha_{t}^{3}=O(\alpha_{t}^{2}/N)$
+	where $\sigma^{2}=\max_{k}\sigma_{k}^{2}$. We show next that $\|\overline{w}_{t}-w^{\ast}\|^{2}=O(\frac{\alpha_{t}}{N})=O(\frac{1}{Nt})$.
+	Before we do, we remark briefly on the choice of $E$. The leading
+	term among terms other than $\mathbb{E}(1-\mu\alpha_{t})\|\overline{w}_{t}-w^{\ast}\|^{2}$
+	is $\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}$. This means
+	we can take $E=O(\sqrt{\frac{1}{N\alpha_{t}}})=O(\sqrt{\frac{t}{N}})$
+	which means $4E^{2}L\alpha_{t}^{3}G^{2}=O(\frac{\alpha_{t}}{N})$
+	will be on the same order as the leading term. This is in contrast
+	with previous results where $E$ cannot scale with $t$. Thus our
+	bound yields a more efficient communication complexity than current
+	analyses. We note however that this is only possible when there is
+	full participation, and when $\nu_{max}$ is close to 1. When there
+	is partial participation, however, $E$ has to stay $O(1)$, as we
+	will soon see. Thus our result hightlights the difference between
+	the communication efficiency of full and partial participation when
+	an $O(1/T)$ bound on the optimality gap is desired.
+
+	\textbf{}%
+	\begin{comment}
+	The leading term on the right hand side the bound is $\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}$,
+	which when summed using the recursive relations gives $O(\alpha_{t})=O(1/t)$,
+	but the constant now is scaled down by $1/N$. \textbf{To achieve
+	this linear speedup, we need $\sum_{t=0}^{T}(1-\mu\alpha_{t})\|w_{0}-w^{\ast}\|^{2}$
+	to be small enough. Note also that the above bound implies that $E$
+	can be chosen to be as large as $O(\sqrt{\frac{T}{N}})$ without degrading
+	the performance, since $E^{2}\alpha_{t}^{3}=O(\alpha_{t}^{2}/N)$
 	which is the order of the leading term. As we will see, in the partial
@@
 	$O(T^{\beta})$ for any $\beta>0$. }
+	\end{comment}
 
-	To complete the proof of convergence, for $\alpha_{t}=\frac{\delta}{t+\gamma}$
-	for $\delta>\frac{1}{\mu}$ and $\gamma>0$ such that $\alpha_{1}\leq\min\{\frac{1}{\mu},\frac{1}{4L}\}$and
+	To complete the proof of convergence, let $B=$ $B_{t}=4E^{2}L\alpha_{t}G^{2}+4E^{2}L^{2}\alpha_{t}^{2}G^{2}+\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\alpha_{t}LG^{2}$.
+	Set the learning rate $\alpha_{t}=\frac{\delta}{t+\gamma}$ for $\delta>\frac{1}{\mu}$
+	and $\gamma=\max\{8\kappa-1,E\}$ such that $\alpha_{1}\leq\min\{\frac{1}{\mu},\frac{1}{4L}\}$and
 	$\alpha_{t}\leq2\alpha_{t+E}$, we can show that 
@@
 	\end{align*}
-	where $v=\max\{\frac{\delta^{2}B'}{\delta\mu-c+1},(\gamma+1)\|w_{0}-w^{\ast}\|^{2}\}$
+	where $v=\max\{\frac{\delta^{2}B}{\delta\mu-1},(\gamma+1)\|w_{0}-w^{\ast}\|^{2}\}$,
 	by induction. The definition of $v$ ensures that 
@@
 	\mathbb{E}\|\overline{w}_{t+1}-w^{\ast}\|^{2}\leq & (1-\alpha_{t}\mu)\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2}+\alpha_{t}^{2}B'\\
-	& \leq(1-\frac{\delta\mu}{t+\gamma})\frac{v}{t+\gamma}+\frac{\delta^{2}B'}{(t+\gamma)^{2}}\\
-	& =\frac{t+\gamma-1}{(t+\gamma)^{2}}v+\left[\frac{\delta^{2}B'}{(t+\gamma)^{2}}-\frac{\delta\mu-1}{(t+\gamma)^{2}}v\right]\\
-	& \leq\frac{v}{t+\gamma+1}
+	& \leq(1-\frac{\delta\mu}{t+\gamma})\frac{v}{N(t+\gamma)}+\frac{\delta^{2}B'}{N^{2}(t+\gamma)^{2}}\\
+	& =\frac{t+\gamma-1}{N(t+\gamma)^{2}}v+\left[\frac{\delta^{2}B'}{N^{2}(t+\gamma)^{2}}-\frac{\delta\mu-1}{N(t+\gamma)^{2}}v\right]\\
+	& \leq\frac{t+\gamma-1}{N(t+\gamma)^{2}}v+\left[\frac{\delta^{2}B'}{N(t+\gamma)^{2}}-\frac{\delta\mu-1}{N(t+\gamma)^{2}}v\right]\\
+	& \leq\frac{1}{N}\frac{v}{t+\gamma+1}
 	\end{align*}
@@
 	\end{align*}
+	where $B'=8E^{2}L^{2}G^{2}/t+\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
 

### 2020-09-10 source-iclr/sec_exp.tex
Update on Overleaf.

@@
 take the values $c = 2^i \ \forall i \in \ZZ$. As the results shown in Figure~\ref{fig:speedup},
-the number of iterations decreases as the number of (active) workers increasing, which is consistent for FedAvg and Nesterov accelerated FedAvg across all scenarios.
+sthe number of iterations decreases as the number of (active) workers increasing, which is consistent for FedAvg and Nesterov accelerated FedAvg across all scenarios.
 For additional experiments on the impact of $E$, detailed experimental setup, and hyperparameter setting, please refer to the Appendix Section~\ref{sec:expsupp}.

### 2020-09-14 source-iclr/app_sumarization.tex
add scaffold and analysis

--- 2020-06-10 source/app_sumarization.tex
@@
 
-\section{Comparison of convergence rates with related works.}
+\section{Comparison of Convergence Rates with Related Works}
 \label{sec:app:comparison}
@@
 % The bounded gradient can imply gradient diversity (see Corollary 10 in ~\cite{li2018federated}). In this case, the gradient diversity is a relaxed
+
+In~\cite{karimireddy2019scaffold}, the linear speedup convergence rate of FedAvg are provided for strongly convex, general convex, and non-convex problems 
+under full participation setting.
+However, their rate does not enjoy linear speedup for any number of devices while our results apply to any valid $K\leq N$.
+For example,  they provides an optimality gap of $\cO\left((1-\frac{K}{N})E/T\right)$ for the strongly convex case \cite[Theorem V]{karimireddy2019scaffold}.
+With partial participation, and when $K=O(1)$, their convergence rate is $\cO(E/T)$ which does not have linear speedup. 
+Under partial participation, the FedAvg analyses in~\cite{karimireddy2019scaffold} requires $E=\cO(1)$.
+For example, the strongly convex result $\cO((1-\frac{K}{N})E/T)$ in Theorem V is $\cO(E/T)$ when $K=O(1)$ and is $\cO(E/NT)$ when $K=\cO(N)$.
+In either case, to achieve a $\cO(1/T)$ convergence rate, it requires $E=\cO(1)$ as well.
+Similar conclusion also holds for the general convex problem.
 
@@
 \begin{table}[h!]
+\hspace{-4em}
+{
 \centering
-{\tiny
+\tiny
 \begin{tabular}{|c|c|c|c|c|c|c|c|}
@@
 FedAvg\cite{koloskova2020unified} & $\cO(\frac{1}{NT})$ & $\cO(N^{-1/2}T^{1/2})$     	& \cmark & Full       & Bounded gradient   & Strongly convex  \\ \hline
+FedAvg\cite{karimireddy2019scaffold} & $\cO(\frac{1}{NT})^{\dagger\dagger}$ & $\cO(1)^{\dagger\dagger}$   	& \cmark & Partial       & Bounded gradient dissimilarity   & Strongly convex  \\ \hline
 FedAvg/N-FedAvg                 & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{\ddagger}$ & \cmark	 & Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
@@
 FedAvg\cite{koloskova2020unified} & $\cO(\frac{1}{\sqrt{NT}})$ & $\cO(N^{-3/4}T^{1/4})$    & \cmark & Full       & Bounded gradient             &  Convex  \\ \hline
+FedAvg\cite{karimireddy2019scaffold} & $\cO(\frac{1}{\sqrt{NT}})^{\dagger\dagger}$ & $\cO(1)^{\dagger\dagger}$  & \cmark & Partial       & Bounded gradient dissimilarity   &  Convex  \\ \hline
 FedAvg/N-FedAvg      & $\cO\left(\frac{1}{\sqrt{KT}}\right)$ & $\cO(N^{-3/4}T^{1/4})^{\ddagger}$& \cmark			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
@@
 Therefore, we compare it with our strongly convex results here.\\
-         $^{\ddagger\ddagger}$ The bounded gradient diversity assumption is not applicable for general heterogeneous data when converging to arbitrarily small $\epsilon$-accuracy (see discussions in Sec~\ref{sec:app:comparison}).
+         $^{\ddagger\ddagger}$ The bounded gradient diversity assumption is not applicable for general heterogeneous data when converging to arbitrarily small $\epsilon$-accuracy (see discussions in Sec~\ref{sec:app:comparison}).\\
+         $^{\dagger\dagger}$ Although the results in~\cite{karimireddy2019scaffold} is applicable for partial participation setting, their results only achieve linear speedup under full participation setting $K=N$ while we show linear speedup convergence for $K\leq N$ (see discussions in Sec~\ref{sec:app:comparison}).
+Under partial participation, the communication complexity is $\cO(1)$.)
            \par}

### 2020-09-14 source-iclr/sec_intro.tex
add scaffold and analysis

@@
 Despite its promise and broad applicability in our current era, the potential value FL delivers is coupled with the unique challenges it brings forth.
-In particular, when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated (and hence protects privacy)~\cite{kairouz2019advances}, it faces two challenges that are absent in centralized optimization and distributed (stochastic) optimization~\cite{zhou2017convergence, stich2018local,khaled2019first,liang2019variance,wang2018cooperative,woodworth2018graph,wang2019adaptive,jiang2018linear,yu2019parallel,yu2019linear, khaled2020tighter,koloskova2020unified}:
+In particular, when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated~\cite{kairouz2019advances}, it faces two challenges that are absent in centralized optimization and distributed (stochastic) optimization~\cite{zhou2017convergence, stich2018local,khaled2019first,liang2019variance,wang2018cooperative,woodworth2018graph,wang2019adaptive,jiang2018linear,yu2019parallel,yu2019linear, khaled2020tighter,koloskova2020unified}:
 

### 2020-09-16 source-iclr/app_sgd_scvx_smooth.tex
proof th1 minor

@@
 % !TEX ROOT=./supp-main.tex
-To facilitate reading, theorems from the main paper are restated and numbered identically. We first summarize some properties of $L$-smooth and $\mu$-strongly
-convex functions~\cite{rockafellar1970convex}. 
-\begin{lemma}
-	Let $F$ be a convex $L$-smooth function. Then we have the following
-	inequalities:
-
-	1. Quadratic upper bound: $0\leq F(\vw)-F(\vw')-\langle\nabla F(\vw'),\vw-\vw'\rangle\leq\frac{L}{2}\|\vw-\vw'\|^{2}$.
-
-	2. Coercivity: $\frac{1}{L}\|\nabla F(\vw)-\nabla F(\vw')\|^{2}\leq\langle\nabla F(\vw)-\nabla F(\vw'),\vw-\vw'\rangle$.
-
-	3. Lower bound: $F(\vw)\geq F(\vw')+\langle\nabla F(\vw'),\vw-\vw'\rangle+\frac{1}{2L}\|\nabla F(\vw)-\nabla F(\vw')\|^{2}$.
-	In particular, $\|\nabla F(\vw)\|^{2}\leq2L(F(\vw)-F(\vw^{\ast}))$.
-
-	4. Optimality gap: $F(\vw)-F(\vw^{\ast})\leq$$\langle\nabla F(\vw),\vw-\vw^{\ast}\rangle$.
-\end{lemma}
-%
-\begin{lemma}
-	Let $F$ be a $\mu$-strongly convex function. Then 
-	\begin{align*}
-	F(\vw) & \leq F(\vw')+\langle\nabla F(\vw'),\vw-\vw'\rangle+\frac{1}{2\mu}\|\nabla F(\vw)-\nabla F(\vw')\|^{2}\\
-	F(\vw)-F(\vw^{\ast}) & \leq\frac{1}{2\mu}\|\nabla F(\vw)\|^{2}
-	\end{align*}
-\end{lemma}
 

### 2020-09-16 source-iclr/app_techlemma.tex
proof th1 minor

--- 2020-06-10 source/app_sgd_scvx_smooth.tex
@@
-To facilitate reading, theorems from the main paper are restated and numbered identically. We first summarize some properties of $L$-smooth and $\mu$-strongly
-convex functions~\cite{rockafellar1970convex}. 
+To facilitate reading, we first summarize some technical lemmas regarding the properties of $L$-smooth and $\mu$-strongly
+convex functions~\cite{rockafellar1970convex} which are useful for the proof later on. 
 \begin{lemma}
@@
 	\begin{align*}
-	F(\vw) & \leq F(\vw')+\langle\nabla F(\vw'),\vw-\vw'\rangle+\frac{1}{2\mu}\|\nabla F(\vw)-\nabla F(\vw')\|^{2}\\
-	F(\vw)-F(\vw^{\ast}) & \leq\frac{1}{2\mu}\|\nabla F(\vw)\|^{2}
+	& F(\vw)  \leq F(\vw')+\langle\nabla F(\vw'),\vw-\vw'\rangle+\frac{1}{2\mu}\|\nabla F(\vw)-\nabla F(\vw')\|^{2}\\
+	& F(\vw)-F(\vw^{\ast})  \leq\frac{1}{2\mu}\|\nabla F(\vw)\|^{2}
 	\end{align*}
 \end{lemma}
-
-\begin{thm}
-	Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$,
-	$\nu_{\max}=\max_{k}Np_{k}$, and set decaying learning rates $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$
-	with $\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$. Then
-	under Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2} with full device participation, 
-	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right)
-	\end{align*}
-	and with partial device participation with at most $K$ sampled devices
-	at each communication round, 
-	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right)
-	\end{align*}
-\end{thm}
-
-\begin{proof}
-	The first step is to observe that the $L$-smoothness of $F$ provides
-	the upper bound
-	\begin{align*}
-	\mathbb{E}(F(\ov{w}_{t}))-F^{\ast} & =\mathbb{E}(F(\ov{w}_{t})-F(\vw^{\ast}))\\
-	& \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}
-	\end{align*}
-	and bound $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$.
-
-	Our main step is to prove the bound 
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}L\alpha_{t}^{3}G^{2}
-	\end{align*}
-
-	We have 
-	\begin{align*}
-	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}\vg_{t})-\vw^{\ast}\|^{2}\\
-	& =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(\vg_{t}-\ov{g}_{t})\|^{2}\\
-	& =A_{1}+A_{2}+A_{3}
-	\end{align*}
-	where 
-	\begin{align*}
-	A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}\\
-	A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-\vg_{t}\rangle\\
-	A_{3} & =\alpha_{t}^{2}\|\vg_{t}-\ov{g}_{t}\|^{2}
-	\end{align*}
-	$\mathbb{E}A_{2}=0$ by definition of $\vg_{t}$ and $\ov{g}_{t}$,
-	while for $A_{3}$ we have
-	\begin{align*}
-	\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
-	\end{align*}
-	again by Jensen's inequality and using the independence of $g_{t,k},g_{t,k'}$.
-
-	Next we bound $A_{1}$: 
-	\begin{align*}
-	\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2}
-	\end{align*}
-	and we will show that the third term $\|\alpha_{t}\ov{g}_{t}\|^{2}$
-	can be canceled by an upper bound of the second term. %
-	\begin{comment}
-	The last term is straightforward to bound by the convexity of $\|\cdot\|^{2}$
-	and $L$-smoothness of $F_{k}$,
-	\begin{align*}
-	\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq2L\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}(F_{k}(\vw_{t}^{k})-F_{k}^{\ast})
-	\end{align*}
-	or 
-	\begin{align*}
-	\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\mathbb{E}\|\nabla F_{k}(\vw_{t}^{k},\xi_{t}^{k})\|^{2}\leq\alpha_{t}^{2}G^{2}
-	\end{align*}
-	\end{comment}
-
-	Now 
-	\begin{align*}
-	& -2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
-	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}w_{t}^{k}-\vw^{\ast}\|^{2}\\
-	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
-	\end{align*}
-	For the second term, which is negative, we can ignore it, but this
-	yields a suboptimal bound that fails to provide the desired linear
-	speedup. Instead, we upper bound it using the following derivation:
-	\begin{align*}
-	& 2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]\\
-	& \leq2\alpha_{t}\left[F(\ov{w}_{t+1})-F(\ov{w}_{t})\right]\\
-	& \leq2\alpha_{t}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{w}_{t+1}-\ov{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\ov{w}_{t+1}-\ov{w}_{t}\|^{2}\\
-	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\vg_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\ov{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\sum_{k}p_{k}\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\ov{w}_{t})-\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
-	\end{align*}
-	where we have used the smoothness of $F$ twice.
-
-	Note that the term $-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$ exactly
-	cancels the $\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$ in the bound
-	for $A_{1}$, so that plugging in the bound for $-2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle$,
-	we have so far proved 
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
-	\end{align*}
-	The term $\mathbb{E}\|\vg_{t}\|^{2}\leq G^{2}$ by assumption.
-
-	Now we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$.
-	Since communication is done every $E$ steps, for any $t\geq0$, we
-	can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $w_{t_{0}}^{k}=\ov{w}_{t_{0}}$for
-	all $k$. Moreover, using $\alpha_{t}$ is non-increasing and $\alpha_{t_{0}}\leq2\alpha{}_{t}$
-	for any $t-t_{0}\leq E-1$, we have 
-	\begin{align*}
-	& \mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
-	& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}g_{i,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}E\alpha_{i}^{2}\|g_{i,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}E^{2}\alpha_{t_{0}}^{2}G^{2}\\
-	& \leq4E^{2}\alpha_{t}^{2}G^{2}
-	\end{align*}
-	\begin{comment}
-	Alternatively, we can also write
-	\begin{align*}
-	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & =\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\alpha_{t-1}\vg_{t-1}-\vw_{t-1}^{k}+\alpha_{t-1}g_{t-1,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}\left(\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\alpha_{t-1}\vg_{t-1}-\alpha_{t-1}g_{t-1,k}\|^{2}\right)
-	\end{align*}
-	and write 
-	\begin{align*}
-	\sum_{k}p_{k}\|g_{t-1,k}-\vg_{t-1}\|^{2}\leq\sum_{k}p_{k}\|g_{t-1,k}\|^{2} & =\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})+\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
-	& \leq2\sum_{k}p_{k}\|g_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})\|^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
-	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})-\nabla F_{k}(\ov{w}_{t-1})+\nabla F_{k}(\ov{w}_{t-1})\|^{2}\\
-	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+4\sum_{k}p_{k}\left(\|\nabla F_{k}(\vw_{t-1}^{k})-\nabla F_{k}(\ov{w}_{t-1})\|^{2}+\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}\right)\\
-	& \leq2\sigma^{2}+4\sum_{k}p_{k}\left(L^{2}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}\right)
-	\end{align*}
-	so that 
-	\begin{align*}
-	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+4\sigma^{2}\alpha_{t-1}^{2}+8\alpha_{t-1}^{2}\sum_{k}p_{k}\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}
-	\end{align*}
-	and taking expectation, we have 
-	\begin{align*}
-	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+4\alpha_{t-1}^{2}(\sigma^{2}+2G^{2})
-	\end{align*}
-	Using this recusive relation at most $E$ times to arrive at a communication
-	round and that $\alpha_{t}L\leq1/8$ for all $t$, we have 
-	\begin{align*}
-	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\leq E\frac{1}{L^{2}}(\sigma^{2}+2G^{2})
-	\end{align*}
-	which gives 
-	\end{comment}
-	{}
-
-	Using the bound on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$,
-	we can conclude that, with $\nu_{max}:=N\cdot\max_{k}p_{k}$ and $\nu_{min}:=N\cdot\min_{k}p_{k}$,
-
-	\begin{align*}
-	& \mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& =\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\sum_{k=1}^{N}(p_{k}N)^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\nu_{max}^{2}\sum_{k=1}^{N}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
-	\end{align*}
-	where $\sigma^{2}=\max_{k}\sigma_{k}^{2}$, and if $\alpha_{t}$
-	satisfies $L\alpha_{t}\leq\frac{1}{8}$. We show next that $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}LG^{2}}{t^{2}})$.
-
-	Let $C\equiv5E^{2}LG^{2}$ and $D\equiv\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
-	Suppose that we have shown $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}\leq b\cdot(\alpha_{t}D+\alpha_{t}^{2}C)$
-	for some constant $b$ and $\alpha_{t}$. Then 
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq b(1-\mu\alpha_{t})(\alpha_{t}D+\alpha_{t}^{2}C)+\alpha_{t}^{2}D+\alpha_{t}^{3}C\\
-	& =(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}D+(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}C
-	\end{align*}
-	and so it remains to choose $\alpha_{t}$ and $b$ such that $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}\leq b\alpha_{t+1}$
-	and $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}\leq b\alpha_{t+1}^{2}$.
-	Recall that we require $\alpha_{t_{0}}\leq\alpha_{t}$ for any $t-t_{0}\leq E-1$,
-	and $L\alpha_{t}\leq\frac{1}{8}$. If we let $\alpha_{t}=\frac{4}{\mu(t+\gamma)}$
-	where $\gamma=\max\{E,32\kappa\}$, then we may check that $\alpha_{t}$
-	satisfies both requirements.
-
-	Setting $b=\frac{4}{\mu}$, we have 
-	\begin{align*}
-	(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t} & =\left(b(1-\frac{4}{t+\gamma})+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}\\
-	& =\left(b\frac{t+\gamma-4}{t+\gamma}+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}\\
-	& =b(\frac{t+\gamma-3}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
-	& \leq b(\frac{t+\gamma-1}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
-	& \leq b\frac{4}{\mu(t+\gamma+1)}=b\alpha_{t+1}
-	\end{align*}
-	and 
-	\begin{align*}
-	(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2} & =\left(b(1-\frac{4}{t+\gamma})+\frac{4}{\mu(t+\gamma)}\right)\frac{16}{\mu^{2}(t+\gamma)^{2}}\\
-	& =\left(b\frac{t+\gamma-4}{t+\gamma}+\frac{4}{\mu(t+\gamma)}\right)\frac{16}{\mu^{2}(t+\gamma)^{2}}\\
-	& =b(\frac{t+\gamma-2}{t+\gamma})\frac{16}{\mu^{2}(t+\gamma)^{2}}\\
-	& \leq b\frac{16}{\mu^{2}(t+\gamma+1)^{2}}=b\alpha_{t+1}^{2}
-	\end{align*}
-	where we have used the facts that 
-	\begin{align*}
-	\frac{t+\gamma-1}{(t+\gamma)^{2}} & \leq\frac{1}{(t+\gamma+1)}\\
-	\frac{t+\gamma-2}{(t+\gamma)^{3}} & \leq\frac{1}{(t+\gamma+1)^{2}}
-	\end{align*}
-	for $\gamma\geq1$.
-
-	Thus we have shown 
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq b\cdot(\alpha_{t+1}D+\alpha_{t+1}^{2}C)
-	\end{align*}
-	for our choice of $\alpha_{t}$ and $b$. Now to ensure 
-	\begin{align*}
-	\|\vw_{0}-\vw^{\ast}\|^{2} & \leq b\cdot(\alpha_{0}D+\alpha_{0}^{2}C)\\
-	& =b\cdot(\frac{4}{\mu\gamma}D+\frac{16}{\mu^{2}\gamma^{2}}C)
-	\end{align*}
-	we can simply scale $b$ by $c\|\vw_{0}-\vw^{\ast}\|^{2}$ for a constant
-	$c$ large enough and the induction step still holds.
-
-	It follows that 
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2} & \leq c\|\vw_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{t}+C\alpha_{t}^{2})
-	\end{align*}
-	for all $t\geq0$.
-
-	Finally, the $L$-smoothness of $F$ implies 
-	\begin{align*}
-	& \mathbb{E}(F(\ov{w}_{T}))-F^{\ast}\\
-	& =\mathbb{E}(F(\ov{w}_{T})-F(\vw^{\ast}))\\
-	& \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}\leq\frac{L}{2}c\|\vw_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
-	& =2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
-	& \leq2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{4}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}LG^{2}\cdot(\frac{4}{\mu(T+\gamma)})^{2}\right]\\
-	& =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
-	\end{align*}
-
-	With partial participation, the update at each communication round
-	is now given by averages over a subset of sampled devices. When $t+1\notin\mathcal{I}_{E}$,
-	$\ov{v}_{t+1}=\ov{w}_{t+1}$, while when $t+1\notin\mathcal{I}_{E}$,
-	we have $\mathbb{E}\ov{w}_{t+1}=\ov{v}_{t+1}$ by design
-	of the sampling schemes, so that 
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}+\ov{v}_{t+1}-\vw^{\ast}\|^{2}\\
-	& =\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2}+\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}
-	\end{align*}
-	As before, $\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}\leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
-
-	The key is to bound $\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2}$.
-	For sampling scheme I we have 
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & =\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|w_{t+1}^{k}-\ov{w}_{t+1}\|^{2}\\
-	& \leq\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
-	\end{align*}
-	while for sampling scheme II 
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & =\frac{N-K}{N-1}\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|w_{t+1}^{k}-\ov{w}_{t+1}\|^{2}\\
-	& \leq\frac{N-K}{N-1}\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
-	\end{align*}
-	The same argument as the full participation case implies 
-	\begin{align*}
-	\mathbb{E}F(\ov{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
-	\end{align*}
-
-	\begin{comment}
-	\begin{proof}
-	With partial participation, $2E\sum_{\tau=1}^{E-1}\alpha_{t-\tau}^{2}\sum_{k=1}^{N}p_{k}\left(\|\nabla F_{k}(\ov{w}_{t-\tau},\xi_{t-\tau}^{k})\|^{2}+l^{2}\|w_{t-\tau}^{k}-\ov{w}_{t-\tau}\|^{2}\right)$
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{1}{K}\sum_{k}p^{k}\|\vw_{t}^{k}-\ov{w}_{t}\|^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+5E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})\cdots(1-\mu\alpha_{t-E})\|\ov{w}_{t-E}-\vw^{\ast}\|^{2}=O(\frac{1}{t-E}\sigma^{2}+E^{2}LG^{2}\frac{1}{(t-E)^{2}}+\frac{E^{2}G^{2}}{K}\frac{1}{(t-E)^{3/2}})\\
-	& +(\alpha_{t}^{3}+(1-\mu\alpha_{t-1})\alpha_{t-1}^{3}+(1-\mu\alpha_{t-1})(1-\mu\alpha_{t-2})\alpha_{t-2}^{3}+\cdots+(1-\mu\alpha_{t-1})\cdots(1-\mu\alpha_{t-E})\alpha_{t-E}^{3}E^{2}LG^{2}\\
-	& +(\alpha_{t}^{2}+(1-\mu\alpha_{t-1})\alpha_{t-1}^{2}+(1-\mu\alpha_{t-1})(1-\mu\alpha_{t-2})\alpha_{t-2}^{2}+\cdots+(1-\mu\alpha_{t-1})\cdots(1-\mu\alpha_{t-E})\alpha_{t-E}^{2}\sigma^{2}\\
-	& +\frac{4}{K}EG^{2}(\alpha_{t}^{2}+\alpha_{t-1}^{2}+\cdots+\alpha_{t-E}^{2})
-	\end{align*}
-
-	so need ot check 
-	\begin{align*}
-	(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})\cdots(1-\mu\alpha_{t-E+1})b\alpha_{t-E+1}^{\beta}+\alpha_{t}^{2} & \leq\alpha_{t+1}^{\beta}\\
-	\end{align*}
-
-	Alternatively, show that 
-	\begin{align*}
-	\frac{1}{K}\sum_{k}p^{k}\|\vw_{t}^{k}-\ov{w}_{t}\|^{2} & =O(\alpha_{t}^{2}G^{2}/K)\\
-	\sum_{k}p^{k}\|\vw_{t}^{k}-\ov{w}_{t}\|^{2}\le\sum_{k}p^{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2} & \leq\sum_{k}p^{k}\|-\alpha_{t_{0}}g_{t_{0},k}-\alpha_{t_{0}+1}g_{t_{0}+1,k}-\cdots-\alpha_{t-1}g_{t-1,k}\|^{2}\\
-	& =\sum_{k}p^{k}\|\alpha_{t_{0}}g_{t_{0},k}+\alpha_{t_{0}+1}g_{t_{0}+1,k}+\cdots+\alpha_{t-1}g_{t-1,k}\|^{2}\\
-	& =\sum_{k}p^{k}\|\alpha_{t_{0}}\nabla F_{k}(\ov{w}_{t_{0}},\xi_{t_{0}}^{k})+\alpha_{t_{0}}\nabla F_{k}(\ov{w}_{t_{0}}-\alpha_{t_{0}+1}\nabla F_{k}(\ov{w}_{t_{0}},\xi_{t_{0}}^{k}),\xi_{t_{0}+1}^{k})+\cdots+\alpha_{t-1}g_{t-1,k}\|^{2}
-	\end{align*}
-	The intuition is to approximate with the sequence starting at $w^{\ast}$,
-	i.e. 
-	\begin{align*}
-	\sum_{k}p^{k}\|\alpha_{t_{0}}\nabla F_{k}(\vw^{\ast},\xi_{t_{0}}^{k})+\alpha_{t_{0}+1}\nabla F_{k}(\vw^{\ast}-\alpha_{t_{0}+1}\nabla F_{k}(\vw^{\ast},\xi_{t_{0}}^{k}),\xi_{t_{0}+1}^{k})+\cdots+\alpha_{t-1}g_{t-1,k}\|^{2} & \approx\\
-	\end{align*}
-	since $\ov{w}_{t_{0}}\approx w^{\ast}+\frac{1}{t_{0}}$. We
-	can take 
-	\begin{align*}
-	g_{t_{0},k} & =\nabla F_{k}(\ov{w}_{t_{0}},\xi)\approx\nabla F_{k}(\ov{w}_{t_{0}},\xi)+\\
-	\end{align*}
-
-	\begin{align*}
-	\alpha_{t}^{2}+(1-\mu\alpha_{t-1})\alpha_{t-1}^{2}+(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})b\alpha_{t-1} & \le\alpha_{t+1}\\
-	\alpha_{t}^{2}+(1-\mu\alpha_{t-1})\alpha_{t-1}^{2}+(1-\mu\alpha_{t-1})(1-\mu\alpha_{t-2})\alpha_{t-2}^{2}+(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})(1-\mu\alpha_{t-2})b\alpha_{t-2} & \le\alpha_{t+1}\\
-	\dots
-	\end{align*}
-	recall 
-	\begin{align*}
-	(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t} & \leq b\alpha_{t+1}
-	\end{align*}
-
-	\begin{align*}
-	(1-\mu\alpha_{t-1})\alpha_{t-1}^{2}+(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})b\alpha_{t-1} & \leq(1-\mu\alpha_{t})\alpha_{t-1}^{2}+(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})b\alpha_{t-1}\\
-	& \le(1-\mu\alpha_{t})\alpha_{t-1}(\alpha_{t-1}+(1-\mu\alpha_{t-1})b)\\
-	& \leq(1-\mu\alpha_{t})b\alpha_{t}
-	\end{align*}
-	so that 
-	\begin{align*}
-	(1-\mu\alpha_{t})\beta\alpha_{t}+\alpha_{t}^{2} & \leq b\alpha_{t+1}\\
-	\end{align*}
-	\begin{align*}
-	(1-\mu\alpha_{t-1})\alpha_{t-1}^{2}+(1-\mu\alpha_{t})(1-\mu\alpha_{t-1})b\alpha_{t-1} & \leq\alpha_{t}^{2}+(1-\mu\alpha_{t-1})\alpha_{t-1}(\alpha_{t-1}+b(1-\mu\alpha_{t}))\le\\
-	(\alpha_{t-1}+b(1-\mu\alpha_{t}))\alpha_{t}\leq
-	\end{align*}
-	\end{proof}
-	\end{comment}
-
-	One may ask whether the dependence on $E$ in the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$
-	can be removed, or equivalently whether $\sum_{k}p_{k}\|\mathbf{w}_{t+1}^{k}-\overline{\mathbf{w}}_{t+1}\|^{2}$
-	can be independent of $E$. We provide a heuristic argument for why
-	this is not possible. Suppose that each $F_{k}(\mathbf{w})=\frac{1}{n}\sum(\mathbf{x}_{k}^{j}-\mathbf{w})^{2}$
-	contains data points $\mathbf{x}_{k}^{j}$ concentrated around the
-	optimizer $\mathbf{w}^{\ast,k}$, and that the $\mathbf{w}{}^{\ast,k}$'s
-	come in pairs around the origin. As a result, the global objective
-	$F$ is minimized at $\mathbf{w}^{\ast}=0$. Then with $t=O(T)$ large,
-	after a communication round $t$ each device will start with $\overline{\mathbf{w}}_{t}$
-	close to the origin, while after $E=T^{\beta}$ iterations with step
-	size $O(1/T)$, the local parameters are updated towards $\mathbf{w}^{\ast,k}$
-	with $\|\mathbf{w}_{t+E}^{k}\|^{2}\approx\frac{1}{T^{2-2\beta}}$,
-	while $\overline{\mathbf{w}}_{t+E}$ remains close to $\mathbf{w}^{\ast}$,
-	so that $\sum_{k}p_{k}\|\mathbf{w}_{t+E}^{k}-\overline{\mathbf{w}}_{t+E}\|^{2}=O(1/T^{2-2\beta})$,
-	which is at a slower rate than $1/T$ for any $\beta>0$. It follows
-	that $E=O(1)$. 
-\end{proof}

### 2020-09-17 source-iclr/app_sgd_scvx_smooth.tex
svx appendix

@@
 
+To convey our proof clearly and highlight the significance of our results comparing to prior work, 
+we first state the follow lemmas and defer the proofs of these lemmas latter. 
+\begin{lemma}[\textbf{One round progress}] Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$,
+suppose our functions satisfies Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2}, and set step size $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$
+	with $\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$, the updates of FedAvg satisfy
+	\begin{align*}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}L\alpha_{t}^{3}G^{2}.
+	\end{align*}
+\label{lem:scvxoner}
+\end{lemma}
+
+\begin{lemma}[\textbf{Bounding gradient variance (Lemma 2 \cite{li2019convergence})} ]
+Given Assumption~\ref{ass:boundedvariance}, the upper bound of gradient variance is given as follows,
+\begin{align*}
+	\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} \leq \sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}.
+	\end{align*}
+\label{lem:bgv}
+\end{lemma}
+
+\begin{lemma}[\textbf{Bounding the divergence of $\vw_t^k$ (Lemma 3 \cite{li2019convergence})} ]
+Given Assumption~\ref{ass:subgrad2}, $\alpha_t$ is non-increasing and $\alpha_t \leq 2\alpha_{t+E}$ for all $t\geq 0$, we have
+	\begin{align*}
+	\mathbb{E}\left[\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} \right]\leq 4E^{2}\alpha_{t}^{2}G^{2}.
+	\end{align*}
+\label{lem:bdw}
+\end{lemma}
+
+We first restate Theorem~\ref{th:scvx_sgd} to facilitate reading and then see its proof based on previous lemmas.
 \begin{thm}
@@
 	\begin{align*}
-	\mathbb{E}(F(\ov{w}_{t}))-F^{\ast} & =\mathbb{E}(F(\ov{w}_{t})-F(\vw^{\ast}))\\
-	& \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}
+	\mathbb{E}(F(\ov{w}_{t}))-F^{\ast} & \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}
 	\end{align*}
@@
 
-	Our main step is to prove the bound 
+	Use lemma~\ref{lem:scvxoner}, we have the follow upper bound for the one round progress: 
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}L\alpha_{t}^{3}G^{2}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}L\alpha_{t}^{3}G^{2}.
 	\end{align*}
 
-	We have 
-	\begin{align*}
-	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}\vg_{t})-\vw^{\ast}\|^{2}\\
-	& =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(\vg_{t}-\ov{g}_{t})\|^{2}\\
-	& =A_{1}+A_{2}+A_{3}
-	\end{align*}
-	where 
-	\begin{align*}
-	A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}\\
-	A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-\vg_{t}\rangle\\
-	A_{3} & =\alpha_{t}^{2}\|\vg_{t}-\ov{g}_{t}\|^{2}
-	\end{align*}
-	By definition of $\vg_{t}$ and $\ov{g}_{t}$ (see \eq{\ref{eq:gradient}}), we have $\mathbb{E}A_{2}=0$.
-	For $A_{3}$, we have the follow upper bound:
-	\begin{align*}
-	\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|\vg_{t,k}-\mathbb{E}\vg_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
-	\end{align*}
-	again by Jensen's inequality and using the independence of $\vg_{t,k},\vg_{t,k'}$ \cite[Lemma 2]{li2019convergence}.
-
-	Next we bound $A_{1}$: 
-	\begin{align*}
-	\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2}
-	\end{align*}
-	and we will show that the third term $\|\alpha_{t}\ov{g}_{t}\|^{2}$
-	can be canceled by an upper bound of the second term. %
-	\begin{comment}
-	The last term is straightforward to bound by the convexity of $\|\cdot\|^{2}$
-	and $L$-smoothness of $F_{k}$,
-	\begin{align*}
-	\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq2L\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}(F_{k}(\vw_{t}^{k})-F_{k}^{\ast})
-	\end{align*}
-	or 
-	\begin{align*}
-	\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\mathbb{E}\|\nabla F_{k}(\vw_{t}^{k},\mathbf{\xi}_{t}^{k})\|^{2}\leq\alpha_{t}^{2}G^{2}
-	\end{align*}
-	\end{comment}
-
-	Now 
-	\begin{align*}
-	& -2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle \vw_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
-	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
-	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
-	\end{align*}
-	For the second term, which is negative, we can ignore it, but this
-	yields a suboptimal bound that fails to provide the desired linear
-	speedup. Instead, we upper bound it using the following derivation:
-	\begin{align*}
-	& 2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]\\
-	& \leq2\alpha_{t}\left[F(\ov{w}_{t+1})-F(\ov{w}_{t})\right]\\
-	& \leq2\alpha_{t}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{w}_{t+1}-\ov{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\ov{w}_{t+1}-\ov{w}_{t}\|^{2}\\
-	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\vg_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\ov{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\sum_{k}p_{k}\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\ov{w}_{t})-\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
-	\end{align*}
-	where we have used the smoothness of $F$ twice.
-
-	Note that the term $-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$ exactly
-	cancels the $\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$ in the bound
-	for $A_{1}$, so that plugging in the bound for $-2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle$,
-	we have so far proved 
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
-	\end{align*}
-	The term $\mathbb{E}\|\vg_{t}\|^{2}\leq G^{2}$ by assumption.
-
-	Now we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$ following \cite{li2019convergence}.
-	Since communication is done every $E$ steps, for any $t\geq0$, we
-	can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $\vw_{t_{0}}^{k}=\ov{w}_{t_{0}}$for
-	all $k$. Moreover, using $\alpha_{t}$ is non-increasing and $\alpha_{t_{0}}\leq2\alpha{}_{t}$
-	for any $t-t_{0}\leq E-1$, we have 
-	\begin{align*}
-	& \mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
-	& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}\vg_{i,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}E\alpha_{i}^{2}\|\vg_{i,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}E^{2}\alpha_{t_{0}}^{2}G^{2}\\
-	& \leq4E^{2}\alpha_{t}^{2}G^{2}
-	\end{align*}
-	\begin{comment}
-	Alternatively, we can also write
-	\begin{align*}
-	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & =\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\alpha_{t-1}\vg_{t-1}-\vw_{t-1}^{k}+\alpha_{t-1}\vg_{t-1,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}\left(\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\alpha_{t-1}\vg_{t-1}-\alpha_{t-1}\vg_{t-1,k}\|^{2}\right)
-	\end{align*}
-	and write 
-	\begin{align*}
-	\sum_{k}p_{k}\|\vg_{t-1,k}-\vg_{t-1}\|^{2}\leq\sum_{k}p_{k}\|\vg_{t-1,k}\|^{2} & =\sum_{k}p_{k}\|\vg_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})+\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
-	& \leq2\sum_{k}p_{k}\|\vg_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})\|^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
-	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})-\nabla F_{k}(\ov{w}_{t-1})+\nabla F_{k}(\ov{w}_{t-1})\|^{2}\\
-	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+4\sum_{k}p_{k}\left(\|\nabla F_{k}(\vw_{t-1}^{k})-\nabla F_{k}(\ov{w}_{t-1})\|^{2}+\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}\right)\\
-	& \leq2\sigma^{2}+4\sum_{k}p_{k}\left(L^{2}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}\right)
-	\end{align*}
-	so that 
-	\begin{align*}
-	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+4\sigma^{2}\alpha_{t-1}^{2}+8\alpha_{t-1}^{2}\sum_{k}p_{k}\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}
-	\end{align*}
-	and taking expectation, we have 
-	\begin{align*}
-	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+4\alpha_{t-1}^{2}(\sigma^{2}+2G^{2})
-	\end{align*}
-	Using this recusive relation at most $E$ times to arrive at a communication
-	round and that $\alpha_{t}L\leq1/8$ for all $t$, we have 
-	\begin{align*}
-	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\leq E\frac{1}{L^{2}}(\sigma^{2}+2G^{2})
-	\end{align*}
-	which gives 
-	\end{comment}
-	{}
-
-	Using the bound on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$,
-	we can conclude that, with $\nu_{max}:=N\cdot\max_{k}p_{k}$ and $\nu_{min}:=N\cdot\min_{k}p_{k}$,
-
-	\begin{align*}
-	& \mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& =\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\sum_{k=1}^{N}(p_{k}N)^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\nu_{max}^{2}\sum_{k=1}^{N}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
-	\end{align*}
-	where in the last inequality we use $\sigma^{2}=\max_{k}\sigma_{k}^{2}$, and assume $\alpha_{t}$
-	satisfies $L\alpha_{t}\leq\frac{1}{8}$. We show next that $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}LG^{2}}{t^{2}})$.
+	We show next that $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}LG^{2}}{t^{2}})$.
 
@@
 
+\subsubsection{Deferred proofs of key lemmas}
+Here we first rewrite the proof of lemma \ref{lem:bgv} \ref{lem:bdw} from~\cite{li2019convergence} with slight modification for the consistence and completeness of this work. 
+\begin{proof}[Proof of lemma~\ref{lem:bgv}]
+	\begin{align*}
+	\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\sum_{k=1}^{N}p_{k}^{2}\|\vg_{t,k}-\mathbb{E}\vg_{t,k}\|^{2}\leq \sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+	\end{align*}
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:bdw}]
+	Now we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$ following \cite{li2019convergence}.
+	Since communication is done every $E$ steps, for any $t\geq0$, we
+	can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $\vw_{t_{0}}^{k}=\ov{w}_{t_{0}}$for
+	all $k$. Moreover, using $\alpha_{t}$ is non-increasing and $\alpha_{t_{0}}\leq2\alpha{}_{t}$
+	for any $t-t_{0}\leq E-1$, we have 
+		\begin{align*}
+	& \mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\\
+	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
+	& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
+	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
+	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}\vg_{i,k}\|^{2}\\
+	& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}E\alpha_{i}^{2}\|\vg_{i,k}\|^{2}\\
+	& \leq2\sum_{k=1}^{N}p_{k}E^{2}\alpha_{t_{0}}^{2}G^{2}\\
+	& \leq4E^{2}\alpha_{t}^{2}G^{2}
+	\end{align*}
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:scvxoner}]
+	We have 
+	\begin{align*}
+	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}\vg_{t})-\vw^{\ast}\|^{2}\\
+	& =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(\vg_{t}-\ov{g}_{t})\|^{2}\\
+	& =A_{1}+A_{2}+A_{3}
+	\end{align*}
+	where 
+	\begin{align*}
+	A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}\\
+	A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-\vg_{t}\rangle\\
+	A_{3} & =\alpha_{t}^{2}\|\vg_{t}-\ov{g}_{t}\|^{2}
+	\end{align*}
+	By definition of $\vg_{t}$ and $\ov{g}_{t}$ (see \eq{\ref{eq:gradient}}), we have $\mathbb{E}A_{2}=0$.
+	For $A_{3}$, we have the follow upper bound:
+	\begin{align*}
+	\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|\vg_{t,k}-\mathbb{E}\vg_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+	\end{align*}
+	again by Jensen's inequality and using the independence of $\vg_{t,k},\vg_{t,k'}$ \cite[Lemma 2]{li2019convergence}.
+
+	Next we bound $A_{1}$: 
+	\begin{align*}
+	\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2}
+	\end{align*}
+	and we will show that the third term $\|\alpha_{t}\ov{g}_{t}\|^{2}$
+	can be canceled by an upper bound of the second term. %
+	\begin{comment}
+	The last term is straightforward to bound by the convexity of $\|\cdot\|^{2}$
+	and $L$-smoothness of $F_{k}$,
+	\begin{align*}
+	\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq2L\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}(F_{k}(\vw_{t}^{k})-F_{k}^{\ast})
+	\end{align*}
+	or 
+	\begin{align*}
+	\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\mathbb{E}\|\nabla F_{k}(\vw_{t}^{k},\mathbf{\xi}_{t}^{k})\|^{2}\leq\alpha_{t}^{2}G^{2}
+	\end{align*}
+	\end{comment}
+
+	Now 
+	\begin{align*}
+	& -2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle\\
+	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle \vw_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
+	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
+	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
+	\end{align*}
+	For the second term, which is negative, we can ignore it, but this
+	yields a suboptimal bound that fails to provide the desired linear
+	speedup. Instead, we upper bound it using the following derivation:
+	\begin{align*}
+	& 2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]\\
+	& \leq2\alpha_{t}\left[F(\ov{w}_{t+1})-F(\ov{w}_{t})\right]\\
+	& \leq2\alpha_{t}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{w}_{t+1}-\ov{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\ov{w}_{t+1}-\ov{w}_{t}\|^{2}\\
+	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\vg_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\ov{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\sum_{k}p_{k}\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\ov{w}_{t})-\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	& \leq-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
+	\end{align*}
+	where we have used the smoothness of $F$ twice.
+
+	Note that the term $-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$ exactly
+	cancels the $\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$ in the bound
+	for $A_{1}$, so that plugging in the bound for $-2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle$,
+	we have so far proved 
+	\begin{align*}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
+	\end{align*}
+	The term $\mathbb{E}\|\vg_{t}\|^{2}\leq G^{2}$ by assumption. 
+	Using lemma~\ref{lem:bdw}, we have
+	\begin{align*}
+     \mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} \leq 4E^{2}\alpha_{t}^{2}G^{2}
+	\end{align*}
+	\begin{comment}
+	Alternatively, we can also write
+	\begin{align*}
+	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & =\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\alpha_{t-1}\vg_{t-1}-\vw_{t-1}^{k}+\alpha_{t-1}\vg_{t-1,k}\|^{2}\\
+	& \leq2\sum_{k=1}^{N}p_{k}\left(\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\alpha_{t-1}\vg_{t-1}-\alpha_{t-1}\vg_{t-1,k}\|^{2}\right)
+	\end{align*}
+	and write 
+	\begin{align*}
+	\sum_{k}p_{k}\|\vg_{t-1,k}-\vg_{t-1}\|^{2}\leq\sum_{k}p_{k}\|\vg_{t-1,k}\|^{2} & =\sum_{k}p_{k}\|\vg_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})+\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
+	& \leq2\sum_{k}p_{k}\|\vg_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})\|^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
+	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})-\nabla F_{k}(\ov{w}_{t-1})+\nabla F_{k}(\ov{w}_{t-1})\|^{2}\\
+	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+4\sum_{k}p_{k}\left(\|\nabla F_{k}(\vw_{t-1}^{k})-\nabla F_{k}(\ov{w}_{t-1})\|^{2}+\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}\right)\\
+	& \leq2\sigma^{2}+4\sum_{k}p_{k}\left(L^{2}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}\right)
+	\end{align*}
+	so that 
+	\begin{align*}
+	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+4\sigma^{2}\alpha_{t-1}^{2}+8\alpha_{t-1}^{2}\sum_{k}p_{k}\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}
+	\end{align*}
+	and taking expectation, we have 
+	\begin{align*}
+	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+4\alpha_{t-1}^{2}(\sigma^{2}+2G^{2})
+	\end{align*}
+	Using this recusive relation at most $E$ times to arrive at a communication
+	round and that $\alpha_{t}L\leq1/8$ for all $t$, we have 
+	\begin{align*}
+	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\leq E\frac{1}{L^{2}}(\sigma^{2}+2G^{2})
+	\end{align*}
+	which gives 
+	\end{comment}
+
+	Using the bound on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$,
+	we can conclude that, with $\nu_{max}:=N\cdot\max_{k}p_{k}$ and $\nu_{min}:=N\cdot\min_{k}p_{k}$,
+
+	\begin{align*}
+	& \mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}\\
+	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	& =\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\sum_{k=1}^{N}(p_{k}N)^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\nu_{max}^{2}\sum_{k=1}^{N}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
+	\end{align*}
+	where in the last inequality we use $\sigma^{2}=\max_{k}\sigma_{k}^{2}$, and assume $\alpha_{t}$
+	satisfies $L\alpha_{t}\leq\frac{1}{8}$.
+
+\end{proof}
+
 One may ask whether the dependence on $E$ in the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$

### 2020-09-17 source-iclr/app_sumarization.tex
proof th1 minor
svx appendix

@@
 FedAvg\cite{koloskova2020unified} & $\cO(\frac{1}{NT})$ & $\cO(N^{-1/2}T^{1/2})$     	& \cmark & Full       & Bounded gradient   & Strongly convex  \\ \hline
-FedAvg\cite{karimireddy2019scaffold} & $\cO(\frac{1}{NT})^{\dagger\dagger}$ & $\cO(1)^{\dagger\dagger}$   	& \cmark & Partial       & Bounded gradient dissimilarity   & Strongly convex  \\ \hline
+FedAvg\cite{karimireddy2019scaffold} & $\cO(\frac{1}{NT})^{\dagger\dagger}$ & $\cO(N^{-1/2}T^{1/2})^{\dagger\dagger}$   	& \cmark & Partial       & Bounded gradient dissimilarity   & Strongly convex  \\ \hline
 FedAvg/N-FedAvg                 & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{\ddagger}$ & \cmark	 & Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
@@
 FedAvg\cite{koloskova2020unified} & $\cO(\frac{1}{\sqrt{NT}})$ & $\cO(N^{-3/4}T^{1/4})$    & \cmark & Full       & Bounded gradient             &  Convex  \\ \hline
-FedAvg\cite{karimireddy2019scaffold} & $\cO(\frac{1}{\sqrt{NT}})^{\dagger\dagger}$ & $\cO(1)^{\dagger\dagger}$  & \cmark & Partial       & Bounded gradient dissimilarity   &  Convex  \\ \hline
+FedAvg\cite{karimireddy2019scaffold} & $\cO(\frac{1}{\sqrt{NT}})^{\dagger\dagger}$ & $\cO(N^{-3/4}T^{1/4})^{\dagger\dagger}$  & \cmark & Partial       & Bounded gradient dissimilarity   &  Convex  \\ \hline
 FedAvg/N-FedAvg      & $\cO\left(\frac{1}{\sqrt{KT}}\right)$ & $\cO(N^{-3/4}T^{1/4})^{\ddagger}$& \cmark			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
@@
          $^{\dagger\dagger}$ Although the results in~\cite{karimireddy2019scaffold} is applicable for partial participation setting, their results only achieve linear speedup under full participation setting $K=N$ while we show linear speedup convergence for $K\leq N$ (see discussions in Sec~\ref{sec:app:comparison}).
-Under partial participation, the communication complexity is $\cO(1)$.)
+The $E$ in the table is obtained under full participation.
+Under partial participation, the communication complexity is $E=\cO(1)$.
            \par}

### 2020-09-21 source-iclr/app_sgd_scvx_smooth.tex
strongly convex proof

@@
 we first state the follow lemmas and defer the proofs of these lemmas latter. 
-\begin{lemma}[\textbf{One round progress}] Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$,
+\begin{lemma}[\textbf{One step progress}] Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$,
 suppose our functions satisfies Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2}, and set step size $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$
@@
 \begin{proof}
-	The proof builds on ideas from~\cite{li2019convergence}. The first step is to observe that the $L$-smoothness of $F$ provides
-	the upper bound
+	% The proof builds on ideas from~\cite{li2019convergence}. The first step is to observe that the $L$-smoothness of $F$ provides
+	% the upper bound
+	% \begin{align*}
+	% \mathbb{E}(F(\ov{w}_{t}))-F^{\ast} & \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}
+	% \end{align*}
+	% and bound $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$. 
+	The roadmap of this proof contains three steps. Firstly, we establish the recursive relationship between $\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}$
+	and $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$. In another word, we show the upper bound of progress of FedAvg from step $t$ to step $t+1$. 
+	Secondly, we upper bound $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$ by induction proof using the recursive relationship in the previous step. 
+	Thirdly, we use the property of L-smoothness to bound the optimality gap.
+
+	With lemma~\ref{lem:scvxoner}, we have the following upper bound for the one step progress: 
 	\begin{align*}
-	\mathbb{E}(F(\ov{w}_{t}))-F^{\ast} & \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}L\alpha_{t}^{3}G^{2}.
 	\end{align*}
-	and bound $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$.
-
-	Use lemma~\ref{lem:scvxoner}, we have the follow upper bound for the one round progress: 
+	We show next that $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}LG^{2}}{t^{2}})$ by induction proof. 
+	To simplify the presentation, we denote $C\equiv6E^{2}LG^{2}$ and $D\equiv\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
+	Suppose that the result is established at step $t$, i.e., $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}\leq b\cdot(\alpha_{t}D+\alpha_{t}^{2}C)$
+	for some constant $b$ and $\alpha_{t}$. Then we have the following inequality using lemma~\ref{lem:scvxoner}:
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}L\alpha_{t}^{3}G^{2}.
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq (b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}D+(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}C
 	\end{align*}
-
-	We show next that $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}LG^{2}}{t^{2}})$.
-
-	Let $C\equiv6E^{2}LG^{2}$ and $D\equiv\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
-	Suppose that we have shown $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}\leq b\cdot(\alpha_{t}D+\alpha_{t}^{2}C)$
-	for some constant $b$ and $\alpha_{t}$. Then 
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq b(1-\mu\alpha_{t})(\alpha_{t}D+\alpha_{t}^{2}C)+\alpha_{t}^{2}D+\alpha_{t}^{3}C\\
-	& =(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}D+(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}C
-	\end{align*}
-	and so it remains to choose $\alpha_{t}$ and $b$ such that $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}\leq b\alpha_{t+1}$
+	To establish the results at step $t+1$, it remains to choose $\alpha_{t}$ and $b$ such that $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}\leq b\alpha_{t+1}$
 	and $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}\leq b\alpha_{t+1}^{2}$.
@@
 	and $L\alpha_{t}\leq\frac{1}{8}$. If we let $\alpha_{t}=\frac{4}{\mu(t+\gamma)}$
-	where $\gamma=\max\{E,32\kappa\}$, then we may check that $\alpha_{t}$
-	satisfies both requirements.
-
-	Setting $b=\frac{4}{\mu}$, we have 
+	where $\gamma=\max\{E,32\kappa\}$ and set $b=\frac{4}{\mu}$, we have:
 	\begin{align*}
-	(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t} & =\left(b(1-\frac{4}{t+\gamma})+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}\\
-	& =\left(b\frac{t+\gamma-4}{t+\gamma}+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}\\
-	& =b(\frac{t+\gamma-3}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
-	& \leq b(\frac{t+\gamma-1}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
-	& \leq b\frac{4}{\mu(t+\gamma+1)}=b\alpha_{t+1}
+	& (b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}  =\left(b(1-\frac{4}{t+\gamma})+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}
+	% & =\left(b\frac{t+\gamma-4}{t+\gamma}+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}\\
+	% & =b(\frac{t+\gamma-3}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
+	% & \leq b(\frac{t+\gamma-1}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
+	 \leq b\frac{4}{\mu(t+\gamma+1)}=b\alpha_{t+1}\\
+	& (b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2} 
+	 =b(\frac{t+\gamma-2}{t+\gamma})\frac{16}{\mu^{2}(t+\gamma)^{2}} \leq b\frac{16}{\mu^{2}(t+\gamma+1)^{2}}=b\alpha_{t+1}^{2}
 	\end{align*}
-	and 
+	where we have used the following inequalities:
 	\begin{align*}
-	(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2} & =\left(b(1-\frac{4}{t+\gamma})+\frac{4}{\mu(t+\gamma)}\right)\frac{16}{\mu^{2}(t+\gamma)^{2}}\\
-	& =\left(b\frac{t+\gamma-4}{t+\gamma}+\frac{4}{\mu(t+\gamma)}\right)\frac{16}{\mu^{2}(t+\gamma)^{2}}\\
-	& =b(\frac{t+\gamma-2}{t+\gamma})\frac{16}{\mu^{2}(t+\gamma)^{2}}\\
-	& \leq b\frac{16}{\mu^{2}(t+\gamma+1)^{2}}=b\alpha_{t+1}^{2}
+	\frac{t+\gamma-1}{(t+\gamma)^{2}} \leq\frac{1}{(t+\gamma+1)}   \hspace{2em}
+	\frac{t+\gamma-2}{(t+\gamma)^{3}} \leq\frac{1}{(t+\gamma+1)^{2}}  \hspace{2em}  \forall\ \gamma\geq1
 	\end{align*}
-	where we have used the facts that 
-	\begin{align*}
-	\frac{t+\gamma-1}{(t+\gamma)^{2}} & \leq\frac{1}{(t+\gamma+1)}\\
-	\frac{t+\gamma-2}{(t+\gamma)^{3}} & \leq\frac{1}{(t+\gamma+1)^{2}}
-	\end{align*}
-	for $\gamma\geq1$.
-
-	Thus we have shown 
+	Thus we have established the result at step $t+1$ assuming the result is correct at step $t$:
 	\begin{align*}
@@
 	\end{align*}
-	for our choice of $\alpha_{t}$ and $b$. Now to ensure 
+	% for our choice of $\alpha_{t}$ and $b$. Now to ensure 
+	At step $t=0$, we can ensure the following inequality by scaling $b$ with $c\|\vw_{0}-\vw^{\ast}\|^{2}$ for a sufficiently large constant
+	$c$.
 	\begin{align*}
-	\|\vw_{0}-\vw^{\ast}\|^{2} & \leq b\cdot(\alpha_{0}D+\alpha_{0}^{2}C)\\
-	& =b\cdot(\frac{4}{\mu\gamma}D+\frac{16}{\mu^{2}\gamma^{2}}C)
+	\|\vw_{0}-\vw^{\ast}\|^{2} \leq b\cdot(\alpha_{0}D+\alpha_{0}^{2}C) =b\cdot(\frac{4}{\mu\gamma}D+\frac{16}{\mu^{2}\gamma^{2}}C)
 	\end{align*}
-	we can simply scale $b$ by $c\|\vw_{0}-\vw^{\ast}\|^{2}$ for a constant
-	$c$ large enough and the induction step still holds.
-
 	It follows that 
-	\begin{align*}
+	\begin{align}
 	\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2} & \leq c\|\vw_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{t}+C\alpha_{t}^{2})
-	\end{align*}
+	\label{eq:th1:1}
+	\end{align}
 	for all $t\geq0$.
@@
 	\begin{align*}
-	& \mathbb{E}(F(\ov{w}_{T}))-F^{\ast}\\
-	& =\mathbb{E}(F(\ov{w}_{T})-F(\vw^{\ast}))\\
-	& \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}\leq\frac{L}{2}c\|\vw_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
+	\mathbb{E}(F(\ov{w}_{T}))-F^{\ast} & \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}\\ 
+	& \leq\frac{L}{2}c\|\vw_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
 	& =2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
@@
 	\end{align*}
+	where in the first line, we use the property of L-smooth function (see Lemma~\ref{lem:lsmooth}), and in the second line, we use the conclusion in \eq{\ref{eq:th1:1}}.
 
@@
 	\end{align*}
-	As before, $\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}\leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
+	Using Lemma~\ref{lem:scvxoner}, we have $\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}\leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
 
@@
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & =\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|\vw_{t+1}^{k}-\ov{w}_{t+1}\|^{2}\\
-	& \leq\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
+	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & \leq\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & =\frac{N-K}{N-1}\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|\vw_{t+1}^{k}-\ov{w}_{t+1}\|^{2}\\
-	& \leq\frac{N-K}{N-1}\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
+	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & \leq\frac{N-K}{N-1}\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}\vg_{t})-\vw^{\ast}\|^{2}\\
-	& =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(\vg_{t}-\ov{g}_{t})\|^{2}\\
-	& =A_{1}+A_{2}+A_{3}
+	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}\vg_{t})-\vw^{\ast}\|^{2} =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(\vg_{t}-\ov{g}_{t})\|^{2}\\
+	& = \underbrace{\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}}_{A_1} + \underbrace{2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-\vg_{t}\rangle}_{A_2} + \underbrace{\alpha_{t}^{2}\|\vg_{t}-\ov{g}_{t}\|^{2}}_{A_3}
 	\end{align*}
-	where 
+	where we denote: 
 	\begin{align*}
@@
 	By definition of $\vg_{t}$ and $\ov{g}_{t}$ (see \eq{\ref{eq:gradient}}), we have $\mathbb{E}A_{2}=0$.
-	For $A_{3}$, we have the follow upper bound:
+	For $A_{3}$, we have the follow upper bound (see Lemma~\ref{lem:bgv}):
 	\begin{align*}
-	\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|\vg_{t,k}-\mathbb{E}\vg_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+	\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
 	\end{align*}
-	again by Jensen's inequality and using the independence of $\vg_{t,k},\vg_{t,k'}$ \cite[Lemma 2]{li2019convergence}.
 
@@
 	and we will show that the third term $\|\alpha_{t}\ov{g}_{t}\|^{2}$
-	can be canceled by an upper bound of the second term. %
+	can be canceled by an upper bound of the second term, which is one of major improvement comparing to prior art~\cite{li2019convergence}.
 	\begin{comment}
@@
 	\end{comment}
-
-	Now 
+	The upper bound of second term can be derived as follows:
 	\begin{align*}
 	& -2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle \vw_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
-	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
-	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
+	=& -2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	=& -2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle \vw_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	\leq&-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
+	\leq& 2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
+	=& \alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
 	\end{align*}
@@
 	& 2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]\\
-	& \leq2\alpha_{t}\left[F(\ov{w}_{t+1})-F(\ov{w}_{t})\right]\\
-	& \leq2\alpha_{t}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{w}_{t+1}-\ov{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\ov{w}_{t+1}-\ov{w}_{t}\|^{2}\\
-	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\vg_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\ov{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\sum_{k}p_{k}\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\ov{w}_{t})-\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
+	\leq& 2\alpha_{t}\left[F(\ov{w}_{t+1})-F(\ov{w}_{t})\right]\\
+	\leq& 2\alpha_{t}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{w}_{t+1}-\ov{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\ov{w}_{t+1}-\ov{w}_{t}\|^{2}\\
+	 =& -2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\vg_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	=&-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	=&-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\ov{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	=&-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\sum_{k}p_{k}\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+   \leq&-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\ov{w}_{t})-\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	\leq&-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	\leq&-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
 	\end{align*}
@@
 	\end{align*}
-	The term $\mathbb{E}\|\vg_{t}\|^{2}\leq G^{2}$ by assumption. 
-	Using lemma~\ref{lem:bdw}, we have
-	\begin{align*}
-     \mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} \leq 4E^{2}\alpha_{t}^{2}G^{2}
-	\end{align*}
+	Under assumption~\ref{ass:subgrad2}, we have $\mathbb{E}\|\vg_{t}\|^{2}\leq G^{2}$.
+Furthermore, we can plug in $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} \leq 4E^{2}\alpha_{t}^{2}G^{2}$ to the above inequality (see Lemma~\ref{lem:bdw}).
+
 	\begin{comment}
@@
 	\end{comment}
-
-	Using the bound on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$,
-	we can conclude that, with $\nu_{max}:=N\cdot\max_{k}p_{k}$ and $\nu_{min}:=N\cdot\min_{k}p_{k}$,
-
+	Therefore, we can conclude that, with $\nu_{max}:=N\cdot\max_{k}p_{k}$ and $\nu_{min}:=N\cdot\min_{k}p_{k}$, 
 	\begin{align*}
 	& \mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& =\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\sum_{k=1}^{N}(p_{k}N)^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\nu_{max}^{2}\sum_{k=1}^{N}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
+	\leq& \mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	= &\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\sum_{k=1}^{N}(p_{k}N)^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	\leq &\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\nu_{max}^{2}\sum_{k=1}^{N}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	\leq &\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
 	\end{align*}

### 2020-09-24 source-iclr/app_sgd_cvx_smooth.tex
Update on Overleaf.

--- 2020-06-10 source/app_sgd_cvx_smooth.tex
@@
-\begin{thm}
-	Under assumptions~\ref{ass:lsmooth},\ref{ass:boundedvariance},\ref{ass:subgrad2} and constant learning
-	rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$, 
+% !TEX ROOT=./supp-main.tex
+In this section we provide the proof of the convergence result for FedAvg with convex and smooth objectives.
+The key step is a one step progress result analogous to that in the strongly convex case, and their proofs share identical components as well.
+\begin{lemma} [\textbf{One step progress, convex case}]
+Let $\overline{\mathbf{w}}_{t}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{t}^{k}$ in FedAvg.
+Under assumptions~\ref{ass:lsmooth},\ref{ass:boundedvariance},\ref{ass:subgrad2}, the following bound holds for all $t$:
 \begin{align*}
-	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T}\right)
+	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
 	\end{align*}
-	with full participation, and with partial device participation with $K$ sampled devices at
-	each communication round and learning rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{K}{T}})$,
-	\begin{align*}
-	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right)
-	\end{align*}
-\end{thm}
-
+	\label{lem:cvxoner}
+\end{lemma}
 \begin{proof}
-	We again start by bounding the term
-	\begin{align*}
-	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}\vg_{t})-\vw^{\ast}\|^{2}\\
-	& =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(\vg_{t}-\ov{g}_{t})\|^{2}\\
-	& =A_{1}+A_{2}+A_{3}
-	\end{align*}
-	where 
-	\begin{align*}
-	A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}\\
-	A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-\vg_{t}\rangle\\
-	A_{3} & =\alpha_{t}^{2}\|\vg_{t}-\ov{g}_{t}\|^{2}
-	\end{align*}
-	$\mathbb{E}A_{2}=0$ by definition of $\vg_{t}$ and $\ov{g}_{t}$,
-	while for $A_{3}$ we have
-	\begin{align*}
-	\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
-	\end{align*}
-	again by Jensen's inequality and using the independence of $g_{t,k},g_{t,k'}$.
-
-	Next we bound $A_{1}$: 
-	\begin{align*}
-	\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2}
-	\end{align*}
-	Using the convexity and $L$-smoothness of $F_{k}$, 
-	\begin{align*}
-	& -2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))\\
-	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]\\
-	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]
-	\end{align*}
-	which results in 
+    The first part of the proof follows directly from \eq{\ref{eq:common one step}} in the proof of Lemma \ref{lem:scvxoner}.
+Setting $\mu=0$ in \eq{\ref{eq:common one step}} (since we are in the convex setting instead of strongly convex), we obtain
     \begin{align*}
@@
 	\end{align*}
-
 	The difference of this bound with that in the strongly convex case
-	is that we no longer have a contraction factor in front of $\|\ov{w}_{t}-\vw^{\ast}\|^{2}$.
+	is that we no longer have a contraction factor of $1-\mu\alpha_t$ in front of $\|\ov{w}_{t}-\vw^{\ast}\|^{2}$.
 	In the strongly convex case, we were able to cancel $\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$
@@
 	\end{align*}
+\end{proof}
+With the one step progress result, we can now prove the convergence result in the convex setting, which we restate below.
+\begin{thm}
+	Under assumptions~\ref{ass:lsmooth},\ref{ass:boundedvariance},\ref{ass:subgrad2} and constant learning
+	rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$, FedAvg satisfies
+	\begin{align*}
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T}\right)
+	\end{align*}
+	with full participation, and with partial device participation with $K$ sampled devices at
+	each communication round and learning rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{K}{T}})$,
+	\begin{align*}
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right)
+	\end{align*}
+\end{thm}
+
+\begin{proof}
+	We first prove the bound for full participation. Applying Lemma~\ref{lem:cvxoner}, we have
+	\begin{align*}
+	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
+	\end{align*}
 	Summing the inequalities from $t=0$ to $t=T$, we obtain 
@@
 	\end{align*}
-	where $C=\frac{4}{K}E^{2}G^{2}$ or $\frac{N-K}{N-1}\frac{4}{K}E^{2}G^{2}$,
+	where $C=\frac{4}{K}E^{2}G^{2}$ or $\frac{N-K}{N-1}\frac{4}{K}E^{2}G^{2}$ depending on the sampling scheme,
 	so that with $\alpha_{t}=\sqrt{\frac{K}{T}}$, we have 

### 2020-09-24 source-iclr/app_sgd_scvx_smooth.tex
Update app_sgd_scvx_smooth.tex
Update on Overleaf.

@@
 
-To convey our proof clearly and highlight the significance of our results comparing to prior work, 
-we first state the follow lemmas and defer the proofs of these lemmas latter. 
-\begin{lemma}[\textbf{One step progress}] Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$,
-suppose our functions satisfies Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2}, and set step size $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$
-	with $\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$, the updates of FedAvg satisfy
+To organize our proofs more effectively and highlight the significance of our results compared to prior works, 
+we first state the following key lemmas used in proofs of main results and defer their proofs to later. 
+\begin{lemma}[\textbf{One step progress, strongly convex}] Let $\overline{\mathbf{w}}_{t}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{t}^{k}$, and
+suppose our functions satisfy Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2}, and set step size $\alpha_{t}=\frac{4}{\mu(\gamma+t)}$
+	with $\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$, then the updates of FedAvg with full participation satisfy
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}L\alpha_{t}^{3}G^{2}.
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}L\alpha_{t}^{3}G^{2}.
 	\end{align*}
@@
 \end{lemma}
-
+We emphasize that the above lemma is the key step that allows us to obtain a bound that improves on the convergence result of~\cite{li2019convergence} with linear speedup.
+Its proof will make use of the following two results.
 \begin{lemma}[\textbf{Bounding gradient variance (Lemma 2 \cite{li2019convergence})} ]
@@
 \begin{lemma}[\textbf{Bounding the divergence of $\vw_t^k$ (Lemma 3 \cite{li2019convergence})} ]
-Given Assumption~\ref{ass:subgrad2}, $\alpha_t$ is non-increasing and $\alpha_t \leq 2\alpha_{t+E}$ for all $t\geq 0$, we have
+Given Assumption~\ref{ass:subgrad2}, and assume that $\alpha_t$ is non-increasing and $\alpha_t \leq 2\alpha_{t+E}$ for all $t\geq 0$, we have
 	\begin{align*}
@@
 
-We first restate Theorem~\ref{th:scvx_sgd} to facilitate reading and then see its proof based on previous lemmas.
+We now restate Theorem~\ref{th:scvx_sgd} from the main text and then prove it using Lemma~\ref{lem:scvxoner}.
 \begin{thm}
-	Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$,
-	$\nu_{\max}=\max_{k}Np_{k}$, and set decaying learning rates $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$
+	Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$ in FedAvg,
+	$\nu_{\max}=\max_{k}Np_{k}$, and set decaying learning rates $\alpha_{t}=\frac{4}{\mu(\gamma+t)}$
 	with $\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$. Then
@@
 	% and bound $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$. 
-	The roadmap of this proof contains three steps. Firstly, we establish the recursive relationship between $\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}$
-	and $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$. In another word, we show the upper bound of progress of FedAvg from step $t$ to step $t+1$. 
-	Secondly, we upper bound $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$ by induction proof using the recursive relationship in the previous step. 
-	Thirdly, we use the property of L-smoothness to bound the optimality gap.
+	The road map of the proof for full device participation contains three steps. First, we establish a recursive relationship between $\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}$
+	and $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$, upper bounding the progress of FedAvg from step $t$ to step $t+1$. 
+	Second, we show that $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}=、\mathcal{O}(\frac{\nu_{max}^{2}\sigma^{2}/\mu}{tN}+\frac{E^{2}LG^{2}/\mu^2}{t^{2}})$ by induction using the recursive relationship from the previous step.
+	Third, we use the property of $L$-smoothness to bound the optimality gap by $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$.
 
-	With lemma~\ref{lem:scvxoner}, we have the following upper bound for the one step progress: 
+	By Lemma~\ref{lem:scvxoner}, we have the following upper bound for the one step progress: 
 	\begin{align*}
@@
 	\end{align*}
-	We show next that $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}LG^{2}}{t^{2}})$ by induction proof. 
+	We show next that $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}=O(\frac{\nu_{max}^{2}\sigma^{2}/\mu}{tN}+\frac{E^{2}LG^{2}/\mu^2}{t^{2}})$ using induction. 
 	To simplify the presentation, we denote $C\equiv6E^{2}LG^{2}$ and $D\equiv\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
-	Suppose that the result is established at step $t$, i.e., $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}\leq b\cdot(\alpha_{t}D+\alpha_{t}^{2}C)$
-	for some constant $b$ and $\alpha_{t}$. Then we have the following inequality using lemma~\ref{lem:scvxoner}:
+	Suppose that we have the bound $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}\leq b\cdot(\alpha_{t}D+\alpha_{t}^{2}C)$
+	for some constant $b$ and learning rates $\alpha_{t}$. Then the one step progress from Lemma~\ref{lem:scvxoner} becomes:
 	\begin{align*}
@@
 	\end{align*}
-	To establish the results at step $t+1$, it remains to choose $\alpha_{t}$ and $b$ such that $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}\leq b\alpha_{t+1}$
+	To establish the result at step $t+1$, it remains to choose $\alpha_{t}$ and $b$ such that $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}\leq b\alpha_{t+1}$
 	and $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}\leq b\alpha_{t+1}^{2}$.
-	Recall that we require $\alpha_{t_{0}}\leq2\alpha_{t}$ for any $t-t_{0}\leq E-1$,
-	and $L\alpha_{t}\leq\frac{1}{8}$. If we let $\alpha_{t}=\frac{4}{\mu(t+\gamma)}$
-	where $\gamma=\max\{E,32\kappa\}$ and set $b=\frac{4}{\mu}$, we have:
+	%Recall that we require $\alpha_{t_{0}}\leq2\alpha_{t}$ for any $t-t_{0}\leq E-1$, and $L\alpha_{t}\leq\frac{1}{8}$. 
+	If we let $\alpha_{t}=\frac{4}{\mu(t+\gamma)}$
+	where $\gamma=\max\{E,32\kappa\}$ (choice of $\gamma$ required to guarantee the one step progress) and set $b=\frac{4}{\mu}$, we have:
 	\begin{align*}
@@
 	At step $t=0$, we can ensure the following inequality by scaling $b$ with $c\|\vw_{0}-\vw^{\ast}\|^{2}$ for a sufficiently large constant
-	$c$.
+	$c$:
 	\begin{align*}
@@
 	& \leq2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{4}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}LG^{2}\cdot(\frac{4}{\mu(T+\gamma)})^{2}\right]\\
-	& =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
+	& =\mathcal{O}(\frac{\kappa}{\mu}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
 	\end{align*}
-	where in the first line, we use the property of L-smooth function (see Lemma~\ref{lem:lsmooth}), and in the second line, we use the conclusion in \eq{\ref{eq:th1:1}}.
+	where in the first line, we use the property of $L$-smooth function (see Lemma~\ref{lem:lsmooth}), and in the second line, we use the conclusion in \eq{\ref{eq:th1:1}}.
 
 	With partial participation, the update at each communication round
-	is now given by averages over a subset of sampled devices. When $t+1\notin\mathcal{I}_{E}$,
-	$\ov{v}_{t+1}=\ov{w}_{t+1}$, while when $t+1\notin\mathcal{I}_{E}$,
+	is now given by weighted averages over a subset of sampled devices. When $t+1\notin\mathcal{I}_{E}$,
+	$\ov{v}_{t+1}=\ov{w}_{t+1}$, while when $t+1\in\mathcal{I}_{E}$,
 	we have $\mathbb{E}\ov{w}_{t+1}=\ov{v}_{t+1}$ by design
-	of the sampling schemes, so that 
+	of the sampling schemes~(\cite{li2019convergence}, Lemma 4), so that 
 	\begin{align*}
@@
 	\end{align*}
-	Using Lemma~\ref{lem:scvxoner}, we have $\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}\leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
+	This in particular implies $\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}\leq \mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$ for all $t$.
+	Since $\ov{v}_t =\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$ always averages over all devices, the full participation one step progress result Lemma~\ref{lem:scvxoner} applied to $\ov{v}_t$  implies
+	\begin{align*}
+	   \mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{v}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\\ &
+	   \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
+	\end{align*}
 
-	The key is to bound $\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2}$.
-	For sampling scheme I we have 
+	The bound for $\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2}$ for the two sampling schemes we consider is provided in \eq{\ref{eq:partialsample}}, and applying it we can write the one step progress for partial participation as
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & \leq\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}+6E^{2}L\alpha_{t}^{3}G^{2},
 	\end{align*}
-	while for sampling scheme II 
+	and the same arguments using induction and $L$-smoothness as the full participation case implies 
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & \leq\frac{N-K}{N-1}\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
-	\end{align*}
-	The same argument as the full participation case implies 
-	\begin{align*}
-	\mathbb{E}F(\ov{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\ov{w}_{T})-F^{\ast}=\mathcal{O}(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
 	\end{align*}
@@
 \subsubsection{Deferred proofs of key lemmas}
-Here we first rewrite the proof of lemma \ref{lem:bgv} \ref{lem:bdw} from~\cite{li2019convergence} with slight modification for the consistence and completeness of this work. 
+Here we first rewrite the proofs of lemmas \ref{lem:bgv} and  \ref{lem:bdw} from~\cite{li2019convergence} with slight modifications for the consistency and completeness of this work, since later we will use modified versions of these results in the convergence proof for Nesterov accelerated FedAvg.
 \begin{proof}[Proof of lemma~\ref{lem:bgv}]
@@
 	& \mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
-	& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}\vg_{i,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}E\alpha_{i}^{2}\|\vg_{i,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}E^{2}\alpha_{t_{0}}^{2}G^{2}\\
-	& \leq4E^{2}\alpha_{t}^{2}G^{2}
+= & \mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
+\leq &\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
+	= &\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
+	= &\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}\vg_{i,k}\|^{2}\\
+	\leq & 2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}E\alpha_{i}^{2}\|\vg_{i,k}\|^{2}\\
+	\leq & 2\sum_{k=1}^{N}p_{k}E^{2}\alpha_{t_{0}}^{2}G^{2}\\
+	\leq & 4E^{2}\alpha_{t}^{2}G^{2}
 	\end{align*}
 \end{proof}
-
+Based on the results of Lemma~\ref{lem:bgv},~\ref{lem:bdw}, we now prove the upper bound of one step SGD progress. This 
+proof improves on the previous work~\cite{li2019convergence} and is the first to reveal the linear speedup of convergence of FedAvg. 
 \begin{proof}[Proof of lemma~\ref{lem:scvxoner}]
@@
 	By definition of $\vg_{t}$ and $\ov{g}_{t}$ (see \eq{\ref{eq:gradient}}), we have $\mathbb{E}A_{2}=0$.
-	For $A_{3}$, we have the follow upper bound (see Lemma~\ref{lem:bgv}):
+	For $A_{3}$, we have the following upper bound (see Lemma~\ref{lem:bgv}):
 	\begin{align*}
@@
 	\end{comment}
-	The upper bound of second term can be derived as follows:
+	The upper bound of second term can be derived as follows, 	using the strong convexity and $L$-smoothness of $F_{k}$:
 	\begin{align*}
@@
 	\end{align*}
-	For the second term, which is negative, we can ignore it, but this
+	We record the bound we have obtained so far, as it will also be used in the proof for convex case: 
+	\begin{align*}
+	& \mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}\\
+	\leq& \mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} \numberthis \label{eq:common one step}
+	\end{align*}
+	For the term $2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]$, which is negative, we can ignore it, but this
 	yields a suboptimal bound that fails to provide the desired linear
@@
 	Note that the term $-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$ exactly
-	cancels the $\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$ in the bound
-	for $A_{1}$, so that plugging in the bound for $-2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle$,
+	cancels the $\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$ in the bound in \eq{\ref{eq:common one step}}, so that plugging in the bound for $-2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle$,
 	we have so far proved 
@@
 	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
+	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2} \numberthis \label{eq:common recursion}
 	\end{align*}
-	Under assumption~\ref{ass:subgrad2}, we have $\mathbb{E}\|\vg_{t}\|^{2}\leq G^{2}$.
-Furthermore, we can plug in $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} \leq 4E^{2}\alpha_{t}^{2}G^{2}$ to the above inequality (see Lemma~\ref{lem:bdw}).
+	Under Assumption~\ref{ass:subgrad2}, we have $\mathbb{E}\|\vg_{t}\|^{2}\leq G^{2}$.
+Furthermore, we can check that our choice of $\alpha_t$ satisfies $\alpha_t$ is non-increasing and $\alpha_t \leq 2\alpha_{t+E}$, so we may plug in the bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} \leq 4E^{2}\alpha_{t}^{2}G^{2}$ to the above inequality (see Lemma~\ref{lem:bdw}).
 
@@
 	\end{align*}
-	where in the last inequality we use $\sigma^{2}=\max_{k}\sigma_{k}^{2}$, and assume $\alpha_{t}$
+	where in the last inequality we use $\sigma^2=\sum_{k=1}^{N}p_k\sigma_{k}^{2}$, and that by construction $\alpha_{t}$
 	satisfies $L\alpha_{t}\leq\frac{1}{8}$.
@@
 One may ask whether the dependence on $E$ in the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$
-can be removed, or equivalently whether $\sum_{k}p_{k}\|\mathbf{w}_{t}^{k}-\overline{\mathbf{w}}_{t}\|^{2}=O(1/T^{2})$
+can be removed, or equivalently whether $\sum_{k}p_{k}\|\mathbf{w}_{t}^{k}-\overline{\mathbf{w}}_{t}\|^{2}=\mathcal{O}(1/T^{2})$
 can be independent of $E$. We provide a simple counterexample that
 shows that this is not possible in general. 
-\begin{lemma}
+\begin{proposition}
 	There exists a dataset such that if $E=\mathcal{O}(T^{\beta})$ for
@@
 	.
-\end{lemma}
+\end{proposition}
 \begin{proof}
@@
 	decays at a slower rate than $\frac{1}{T^{2}}$, resulting in a convergence
-	rate slower than $O(\frac{1}{T})$ with partial participation. 
+	rate slower than $\mathcal{O}(\frac{1}{T})$ with partial participation. 
 \end{proof}

### 2020-09-24 source-iclr/sec_Nesterov.tex
Update on Overleaf.

@@
 \begin{align*}
-\mathbf{v}_{t+1}^{k} & =\mathbf{w}_{t}^{k}-\alpha_{t}\mathbf{g}_{t,k},\\
-\mathbf{w}_{t+1}^{k} & =\begin{cases}
+\mathbf{v}_{t+1}^{k} & =\mathbf{w}_{t}^{k}-\alpha_{t}\mathbf{g}_{t,k}, \hspace{1em}
+\mathbf{w}_{t+1}^{k} =\begin{cases}
 \mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E},\\
-\sum_{k \in \cS_{t+1}}\left[\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E},
+\sum_{k \in \cS_{t+1}}q_k\left[\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E},
 \end{cases}
@@
 where $\mathbf{g}_{t,k}:=\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})$ is
-the stochastic gradient sampled on the $k$-th device at time $t$.  
+the stochastic gradient sampled on the $k$-th device at time $t$, and $q_k$ again depends on participation and sampling schemes.  
 \begin{theorem}
-	\label{thm:nesterov_scvx}Let $\overline{\mathbf{v}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{v}_{T}^{k}$
+	\label{thm:nesterov_scvx}Let $\overline{\mathbf{v}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{v}_{T}^{k}$ in Nesterov accelerated FedAvg,
 	and set learning rates $\alpha_{t}=\frac{6}{\mu}\frac{1}{t+\gamma}$,  $\beta_{t-1}=\frac{3}{14(t+\gamma)(1-\frac{6}{t+\gamma})\max\{\mu,1\}}$.

### 2020-09-24 source-iclr/sec_SGD.tex
Update sec_SGD.tex
Update on Overleaf.

@@
 \begin{theorem}
-	\label{thm:SGD_scvx}Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$,
-	$\nu_{\max}=\max_{k}Np_{k}$, and set decaying learning rates $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$
+	\label{thm:SGD_scvx}Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$ in FedAvg,
+	$\nu_{\max}=\max_{k}Np_{k}$, and set decaying learning rates $\alpha_{t}=\frac{4}{\mu(\gamma+t)}$
 	with $\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$. Then
@@
 	\end{align*}
+	\label{th:scvx_sgd}
 \end{theorem}
@@
 \cite{haddadpour2019convergence} proves a linear speedup $\mathcal{O}(1/NT)$ result for strongly convex and smooth objectives, with $\mathcal{O}(N^{1/3}T^{2/3})$ communication complexity with \emph{i.i.d.} data and partial participation.
-However, their results build on the bounded gradient diversity assumption, which implies the existence of $\mathbf{w}^*$ that minimizes all local objectives (see discussions in Section~\ref{sec:assumptions}), effectively removing system heterogeneity.
+However, their results build on the bounded gradient diversity assumption, which implies the existence of $\mathbf{w}^*$ that minimizes all local objectives (see discussions in Section~\ref{sec:assumptions} and Appendix~\ref{sec:app:comparison}), effectively removing system heterogeneity.
 The bound in \cite{koloskova2020unified} matches our bound in the full participation case, but their framework excludes partial participation \cite[Proposition 1]{koloskova2020unified}.
@@
 \begin{theorem}
-	\label{thm:SGD_cvx}Under assumptions~\ref{ass:lsmooth},\ref{ass:boundedvariance},\ref{ass:subgrad2} and constant learning
-	rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$, 
+	\label{thm:SGD_cvx}Under Assumptions~\ref{ass:lsmooth},\ref{ass:boundedvariance},\ref{ass:subgrad2} and constant learning
+	rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$, FedAvg satisfies
 	\begin{align*}

### 2020-09-24 source-iclr/sec_intro.tex
Update on Overleaf.

@@
 In recent years, FL has swiftly emerged as an important learning paradigm~\cite{mcmahan2016communication,li2018federated}--one that enjoys widespread success in applications such as personalized recommendation~\cite{chen2018federated}, virtual assistant~\cite{lamautonomy}, and keyboard prediction~\cite{47586}, to name a few--for at least two reasons: First, the rapid proliferation of smart devices that are equipped with both computing power and data-capturing capabilities provided the infrastructure core for FL.
-Second, the rising awareness of privacy and the exponential growth of computational power (blessed by Moore's law) in mobile devices have made it increasingly attractive to push the computation to the edge.
+Second, the rising awareness of privacy and the exponential growth of computational power in mobile devices have made it increasingly attractive to push the computation to the edge.
 
@@
 
-1) \textbf{Data heterogeneity:} data distributions in devices are different (and data can't be shared);
+1) \textbf{Data heterogeneity:} data distributions in devices are different (and data cannot be shared);
 
-2) \textbf{System heterogeneity:} only a subset of devices may access the central server at each time both because the communications bandwidth profiles vary across devices and because there is no central server that has control over when a device is active.
+2) \textbf{System heterogeneity:} only a subset of devices may access the central server at each time both because the communications bandwidth profiles vary across devices and because there is no central server that has control over when a device is active (the presence of ``stragglers'').
+% Gaps in the theoretical understanding.
 
-% Gaps in the theoretical understanding. 
 To address these challenges, Federated Averaging (FedAvg)~\cite{mcmahan2016communication} was proposed as a particularly effective heuristic, which has enjoyed great empirical success~\cite{47586}.
 This success has since motivated a growing line of research efforts into understanding its theoretical convergence guarantees in various settings.
-For instance, \cite{haddadpour2019convergence} analyzed FedAvg (for non-convex smooth problems satisfying PL conditions) under the assumption that each local device's minimizer is the same as the minimizer of the joint problem (if all devices' data is aggregated together), an overly restrictive assumption.
-Very recently, \cite{li2019convergence} furthered the progress and established an $\cO(\frac{1}{T})$ convergence rate for FedAvg for strongly convex smooth problems. 
+For instance, \cite{haddadpour2019convergence} analyzed FedAvg (for non-convex smooth problems satisfying PL conditions) under the assumption that each local device's minimizer is the same as the minimizer of the joint problem (if all devices' data is aggregated together), an overly restrictive assumption that restricts the extent of data heterogeneity.
+Very recently, \cite{li2019convergence} furthered the progress and established an $\cO(\frac{1}{T})$ convergence rate for FedAvg for strongly convex smooth Federated learning problems with both data and system heterogeneity.
+A concurrent result in the same setting~\cite{karimireddy2019scaffold} established a similar $\cO(\frac{1}{T})$ result that allows for a linear speedup when the number of participating devices is large.
 At the same time, \cite{huo2020faster} studied the Nesterov accelerated FedAvg for non-convex smooth problems and established 
@@
 However, despite these very recent fruitful pioneering efforts into understanding the theoretical convergence properties of FedAvg, it remains open as to how the number of devices--particularly the number of devices that participate in the computation--affects the convergence speed.
-In particular, do we get linear speedup of FedAvg? What about when FedAvg is accelerated? These aspects are currently unexplored in FL. We fill in the gaps here by providing affirmative answers.
+In particular, is linear speedup of FedAvg a universal phenomenon across different settings and algorithms?
+What about when FedAvg is accelerated with momentum updates?
+Does the presence of both data and system heterogeneity in Federated learning require major technical novelties over results from distributed and decentralized optimization?
+These aspects are currently unexplored in FL.
+We fill in the gaps here by providing affirmative answers.
 
@@
 }
-\caption{Our convergence results for FedAvg and accelerated FedAvg in this paper. Throughout the paper, $N$ is the total
+\caption{\small Our convergence results for FedAvg and accelerated FedAvg in this paper. Throughout the paper, $N$ is the total
 number of local devices, and $K \leq N$ is the maximal number of devices that are accessible to the central server.

### 2020-09-24 source-iclr/sec_problem.tex
Update on Overleaf.

--- 2020-06-05 source/sec_problem.tex
@@
 \subsection{The Federated Averaging (FedAvg) Algorithm}
-We first introduce the standard Federated Averaging (FedAvg) algorithm~\cite{mcmahan2016communication}.
+We first introduce the standard Federated Averaging (FedAvg)  algorithm which was first proposed by~\cite{mcmahan2016communication}.
 FedAvg updates the model in each device by local Stochastic Gradient Descent (SGD) and sends the latest model to the central server every $E$
@@
 Formally, the updates of FedAvg at round $t$ is described as follows:
-\begin{align*} 
+\begin{align}
+\label{eq:fedavg updates}
 \vv_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}\vg_{t,k}, \hspace{1em}
@@
 \mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\ 
-\sum_{k \in \cS_{t+1}} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E},
+\sum_{k \in \cS_{t+1}} q_k \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E},
 \end{array}\right.
-\end{align*}
-where $\vw_t^k$ is the local model parameter maintained in the $k$-th device at the $t$-th iteration, $\vg_{t,k}:=\nabla F_{k}(\vw_{t}^{k},\xi_{t}^{k})$ is the stochastic gradient based on $\xi_{t}^{k}$, the data sampled from $k$-th device’s local data uniformly at random.
-$\mathcal{I}_{E}=\{E,2E,\dots\}$ is the set of global communication steps.
-We use $\mathcal{S}_{t+1}$ to represent the set of active devices at $t+1$.
+\end{align}
+where $\vw_t^k$ is the local model parameter maintained in the $k$-th device at the $t$-th iteration and $\vg_{t,k}:=\nabla F_{k}(\vw_{t}^{k},\xi_{t}^{k})$ is the stochastic gradient based on $\xi_{t}^{k}$, the data point sampled from $k$-th device’s local data uniformly at random.
+$\mathcal{I}_{E}=\{E,2E,\dots\}$ is the set of global communication steps, when local parameters from a set of active devices are averaged and broadcast to all devices.
+We use $\mathcal{S}_{t+1}$ to represent the (random) set of active devices at $t+1$.
+$q_k$ is a set of averaging weights that are specific to the sampling procedure used to obtain the set of active devices $\mathcal{S}_{t+1}$.
 
@@
 we consider both the case of \textbf{full participation} where the model is
-averaging over all devices at the communication round, i.e.,
-$\mathbf{w}_{t+1}^{k} = \sum_{k=1}^N p_k \mathbf{v}_{t+1}^{k}$, and
+averaged over all devices at each communication round, in which case $q_k=p_k$ for all $k$ and 
+$\mathbf{w}_{t+1}^{k} = \sum_{k=1}^N p_k \mathbf{v}_{t+1}^{k}$ if $t+1 \in \mathcal{I}_{E}$, and
 the case of \textbf{partial participation} where $|\cS_{t+1}| < N$. 
-With partial participation, $\cS_{t+1}$ is obtained by two types of
-sampling schemes to simulate practical scenarios~\cite{li2019convergence}.
-For example, one scheme establishes $\cS_{t+1}$ by \emph{i.i.d.} sampling the devices with probability $p_k$ with replacement.
+With partial participation, we follow~\cite{li2019convergence} and~\cite{li2018federated} and assume that $\cS_{t+1}$ is obtained by one of two types of
+sampling schemes to simulate practical scenarios.
+One scheme establishes $\cS_{t+1}$ by \emph{i.i.d.} sampling the devices with probability $p_k$ with replacement, and uses $q_k=\frac{1}{K}$, while the other scheme samples $\cS_{t+1}$ uniformly \emph{i.i.d.} from all devices without replacement, and uses $q_k=p_k\frac{N}{K}$.
 Both schemes
 guarantee that gradient updates in FedAvg are unbiased stochastic versions of
-updates in FedAvg with full participation. For 
-more details on the notations and setup, please refer to Section~\ref{sec:app:notations} in the appendix.
+updates in FedAvg with full participation, which is important in the theoretical analysis of convergence.
+Because the original sampling scheme and weights proposed by \cite{mcmahan2016communication} lacks this nice property, it is not considered in this paper.
+For
+more details on the notations and setup as well as properties of the two sampling schemes, please refer to Section~\ref{sec:app:notations} in the appendix.
 
@@
 bounded gradient diversity restricts local objectives to have the same minimizer as the global objective, contradicting the heterogeneous data setting. 
-For detailed explanations of our assumptions, please refer to Appendix Section~\ref{sec:app:comparison}.
+For detailed discussions of our assumptions, please refer to Appendix Section~\ref{sec:app:comparison}.
 

### 2020-09-24 source-iclr/sec_SGD.tex
Update on Overleaf.

@@
 
-In this section, we provide convergence analyses of FedAvg for convex objectives in the general setting with both heterogeneous data and partial
-participation. We show that for strongly convex and smooth objectives,
+In this section, we provide convergence analyses of FedAvg for convex objectives in the general setting with both heterogeneous data (data heterogeneity) and partial
+participation (system heterogeneity). We show that for strongly convex and smooth objectives,
 the convergence of the optimality gap of averaged parameters across
 devices is $\mathcal{O}(1/NT)$, while for convex and smooth
-objectives, the rate is $\mathcal{O}(1/\sqrt{NT})$. 
+objectives, the rate is $\mathcal{O}(1/\sqrt{NT})$. [\textcolor{red}{ADD RELATION TO PREVIOUS WORKS}]
 Detailed proofs are deferred to Appendix Section~\ref{sec:app:fedavg}.
@@
 The bound in \cite{koloskova2020unified} matches our bound in the full participation case, but their framework excludes partial participation \cite[Proposition 1]{koloskova2020unified}.
+\textcolor{red}{ADD COMPARISON WITH SCAFFOLD PAPER AND OTHER LOCAL SGD RESULTS}
 

### 2020-09-24 source-iclr/sec_intro.tex
Update on Overleaf.

@@
 However, despite these very recent fruitful pioneering efforts into understanding the theoretical convergence properties of FedAvg, it remains open as to how the number of devices--particularly the number of devices that participate in the computation--affects the convergence speed.
-In particular, is linear speedup of FedAvg a universal phenomenon across different settings and algorithms?
+In particular, is linear speedup of FedAvg a universal phenomenon across different settings and for any number of devices?
 What about when FedAvg is accelerated with momentum updates?
-Does the presence of both data and system heterogeneity in Federated learning require major technical novelties over results from distributed and decentralized optimization?
-These aspects are currently unexplored in FL.
+Does the presence of both data and system heterogeneity in FL require major technical novelties over results from distributed and decentralized optimization?
+These aspects are currently unexplored or underexplored in FL.
 We fill in the gaps here by providing affirmative answers.
@@
 and smooth problems (where $K$ is the number of participating devices), thereby establishing that FedAvg enjoys the desirable linear speedup property in the FL setup.
-Prior to our work here, the best and the most related convergence analysis is given by \cite{li2019convergence}, which established an $\cO(\frac{1}{T})$ convergence rate for strongly convex smooth problems under FedAvg.
-Our rate matches the same (and optimal) dependence on $T$, but also completes the picture by establishing the linear dependence on $K$.
+Prior to our work here, the best and the most related convergence analysis is given by \cite{li2019convergence} and~\cite{karimireddy2019scaffold}, which established an $\cO(\frac{1}{T})$ convergence rate for strongly convex smooth problems under FedAvg.
+Our rate matches the same (and optimal) dependence on $T$, but also completes the picture by establishing the linear dependence on $K$, for any $K\leq N$, where $N$ is the total number of devices, whereas~\cite{li2019convergence} does not have linear speedup analysis, and~\cite{karimireddy2019scaffold} only allows linear speedup close to full participation ($K=\mathcal{O}(N)$).
+As for convex and smooth problems, there was no prior work that established the {\small{$\cO(\frac{1}{\sqrt{T}})$}} rate under both system and data heterogeneity.
 
@@
 To the best of our knowledge, these are the first results that provide a linear speedup characterization of Nesterov accelerated FedAvg in those two problem classes (that FedAvg and Nesterov accelerated FedAvg share the same convergence rate is to be expected: this is the case even for centralized stochastic optimization).
+Prior to our results here, the most relevant results~\cite{yu2019linear} only concern the non-convex setting, where convergence is measured with respect to stationary points (vanishing of gradient norms, rather than optimality gaps).
 
 Third, we study a subclass of strongly convex smooth problems where the objective is over-parameterized and establish 
-a faster $\cO(\exp(-\frac{KT}{\kappa}))$ convergence rate for FedAvg.
+a faster $\cO(\exp(-\frac{KT}{\kappa}))$ convergence rate for FedAvg, in contrast to the $\cO(\exp(-\frac{T}{\kappa}))$ rate for individual solvers~\cite{ma2017power}.
 Within this class, we further consider the linear regression problem and establish an even sharper rate under FedAvg.
-In addition, we propose a new variant of accelerated FedAvg--MaSS accelerated FedAvg--and establish a faster convergence rate (compared to if no acceleration is used).
+In addition, we propose a new variant of accelerated FedAvg based on a momentum update of~\cite{liu2018accelerating}--MaSS accelerated FedAvg--and establish a faster convergence rate (compared to if no acceleration is used).
 This stands in contrast to generic (strongly) convex stochastic problems where theoretically no rate improvement is obtained when one accelerates FedAvg.
 The detailed convergence results are summarized in Table~\ref{tb:convergencerateintro}.
+
+\textcolor{red}{ADD COMPARISON WITH LOCAL SGD AND OTHER DISTRIBUTED OPTIMIZATION RESULTS}
 

### 2020-09-24 source-iclr/sec_problem.tex
Update on Overleaf.

@@
 data distributions of each local device. The local minimum
-$F^{*}_k= \min_{\vw \in R^d} F_k(\vw)$ can be far from
-the global minimum of \eq{\ref{eq:problem}}.
+$F^{*}_k= \min_{\vw} F_k(\vw)$ can be far from
+the global minimum of \eq{\ref{eq:problem}} (data heterogeneity).
 
@@
 the case of \textbf{partial participation} where $|\cS_{t+1}| < N$.
+
 With partial participation, we follow~\cite{li2019convergence} and~\cite{li2018federated} and assume that $\cS_{t+1}$ is obtained by one of two types of
 sampling schemes to simulate practical scenarios.
-One scheme establishes $\cS_{t+1}$ by \emph{i.i.d.} sampling the devices with probability $p_k$ with replacement, and uses $q_k=\frac{1}{K}$, while the other scheme samples $\cS_{t+1}$ uniformly \emph{i.i.d.} from all devices without replacement, and uses $q_k=p_k\frac{N}{K}$.
+One scheme establishes $\cS_{t+1}$ by \emph{i.i.d.} sampling the devices with probability $p_k$ with replacement, and uses $q_k=\frac{1}{K}$, where $K=|\mathcal{S}_{t+1}|$, while the other scheme samples $\cS_{t+1}$ uniformly \emph{i.i.d.} from all devices without replacement, and uses $q_k=p_k\frac{N}{K}$.
 Both schemes
@@
 
-\begin{assumption}[L-smooth]
+\begin{assumption}[$L$-smooth]
 	$F_{1}, \cdots, F_{N}$ are all $L$-smooth: for all  $\mathbf{v}$  and $\mathbf{w}$, $F_{k}(\mathbf{v}) \leq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{L}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$.
@@
 The variance of stochastic gradients in each device is bounded: $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq \sigma_{k}^{2}$,
-	for $k=1, \cdots, N$ and any $\mathbf{w}_{t}^{k}$. Let $\sigma^2=\sum_{k=1}^{N}p_k\sigma_{k}^{2}$.
+	for $k=1, \cdots, N$ and any $\mathbf{w}_{t}^{k}$. Let $\sigma^2:=\sum_{k=1}^{N}p_k\sigma_{k}^{2}$.
 	\label{ass:boundedvariance}
@@
 Therefore, there is no substantial difference between these assumptions
-and assuming the bounds at $\mathbf{w}^{\ast}$ only.
+and assuming the bounds at $\mathbf{w}^{\ast}$ only~\cite{koloskova2020unified}.
 Furthermore, compared to assuming \textit{bounded gradient diversity} as in related work~\cite{haddadpour2019convergence,li2018federated}, Assumption~\ref{ass:subgrad2} is much less restrictive.
@@
 bounded gradient diversity restricts local objectives to have the same minimizer as the global objective, contradicting the heterogeneous data setting. 
-For detailed discussions of our assumptions, please refer to Appendix Section~\ref{sec:app:comparison}.
+For detailed discussions of our assumptions and additional comparisons with related works, please refer to Appendix Section~\ref{sec:app:comparison}.
 

### 2020-09-24 source-iclr/app_sgd_scvx_smooth.tex
D2 first round

@@
 	and $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$, upper bounding the progress of FedAvg from step $t$ to step $t+1$. 
-	Second, we show that $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}=、\mathcal{O}(\frac{\nu_{max}^{2}\sigma^{2}/\mu}{tN}+\frac{E^{2}LG^{2}/\mu^2}{t^{2}})$ by induction using the recursive relationship from the previous step.
+	Second, we show that $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}=\mathcal{O}(\frac{\nu_{max}^{2}\sigma^{2}/\mu}{tN}+\frac{E^{2}LG^{2}/\mu^2}{t^{2}})$ by induction using the recursive relationship from the previous step.
 	Third, we use the property of $L$-smoothness to bound the optimality gap by $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$.
@@
 	\begin{align*}
-	& \mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}\\
-	\leq& \mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} \numberthis \label{eq:common one step}
+	 \mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}
+	\leq & \mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\\
+	 & +2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} \numberthis \label{eq:common one step}
 	\end{align*}

### 2020-09-25 source-iclr/app_sgd_scvx_smooth.tex
Update on Overleaf.

@@
 shows that this is not possible in general. 
-\begin{proposition}
+\begin{proposition} \label{prop:tight}
 	There exists a dataset such that if $E=\mathcal{O}(T^{\beta})$ for

### 2020-09-25 source-iclr/main.tex
iclr from nips version 1.
update exp
add scaffold and analysis
proof th1 minor
Update main.tex
Update on Overleaf.

--- 2020-06-05 source/main.tex
@@
 % ready for submission
-% \usepackage{neurips_2020}
+\usepackage{iclr2021_conference,times}
 
@@
 % to avoid loading the natbib package, add option nonatbib:
-     \usepackage[nonatbib]{neurips_2020}
+     % \usepackage[nonatbib]{neurips_2020}
 \usepackage{slashbox,multirow}
@@
 \newtheorem{thm}{Theorem}
-\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{lemma}{Lemma}
 \newtheorem{remark}{Remark}
+\newtheorem{proposition}{Proposition}
 \newcommand{\lkxcom}[1]{{\color{red}{#1}}}
@@
 \input{bmacros}
+
+\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
+
 \newtheorem*{assumption*}{Assumption}
+\newtheorem*{claim*}{Claim}
 
@@
 FedAvg has Linear Speedup}
-
-\author{}
+\author{%
+  Zhaonan Qu\thanks{Equal contribution} \\
+ Stanford University\\
+  \texttt{zhaonanq@stanford.edu} \\
+  \And
+  Kaixiang Lin$^*$ \\
+  Michigan State University \\
+  \texttt{linkaixi@msu.edu} \\
+  \AND
+  Jayant Kalagnanam \\
+  IBM Research \\
+  \texttt{jayant@us.ibm.com} \\
+  \And
+  Zhaojian Li \\
+  Michigan State University \\
+  \texttt{lizhaoj1@egr.msu.edu} \\
+  \And
+  Jiayu Zhou\\
+  Michigan State University \\
+  \texttt{jiayuz@msu.edu} \\
+  \And
+  Zhengyuan Zhou\\
+  New York University \\
+  \texttt{zzhou@stern.nyu.edu} \\
+}
 
@@
 \begin{abstract}
-Federated learning (FL) learns a model jointly from a set of participating devices without sharing each other's privately held data.
-The characteristics of non-\textit{iid} data across the network, low device participation and the mandate that data remain private bring challenges in understanding the convergence of FL algorithms, particularly in regards to how convergence scales with the number of participating devices.
-In this paper, we focus on Federated Averaging (FedAvg)--the most widely used and effective FL algorithm in use today--and provide a comprehensive study of its convergence rate.
-Although FedAvg has recently been studied by an emerging line of literature, it remains open as to how FedAvg's convergence scales with the number of participating devices in the FL setting--a crucial question whose answer would shed light on the performance of FedAvg in large FL systems.
-We fill this gap by establishing convergence guarantees for FedAvg under two classes of problems: strongly convex smooth, convex smooth, and overparametrized strongly convex smooth problems.
-We show that FedAvg enjoys linear speedup in each case, although with different convergence rates.
-For each class, we also characterize the corresponding convergence rates for the Nesterov accelerated FedAvg algorithm in the FL setting: to the best of our knowledge, these are the first linear speedup guarantees for FedAvg when Nesterov acceleration is used.
-To accelerate FedAvg, we also design a new momentum-based FL algorithm that further improves the convergence rate in overparametrized linear regression problems.
+\small Federated learning (FL) learns a model jointly from a set of participating devices without sharing each other's privately held data.
+The characteristics of non-\textit{i.i.d.} data across the network, low device participation, high communication costs, and the mandate that data remain private bring challenges in understanding the convergence of FL algorithms, particularly in regards to how convergence scales with the number of participating devices.
+In this paper, we focus on Federated Averaging (FedAvg)--arguably the most popular and effective FL algorithm class in use today--and provide a unified and comprehensive study of its convergence rate.
+Although FedAvg has recently been studied by an emerging line of literature, it remains open as to how FedAvg's convergence scales with the number of participating devices in the fully heterogeneous FL setting--a crucial question whose answer would shed light on the performance of FedAvg in large FL systems.
+We fill this gap by providing a unified analysis that establishes convergence guarantees for FedAvg under three classes of problems: strongly convex smooth, convex smooth, and overparameterized strongly convex smooth problems.
+We show that FedAvg enjoys linear speedup in each case, although with different convergence rates and communciation efficiencies.
+While there have been linear speedup results from distributed optimization that assumes full participation, ours are the first to establish linear speedup for FedAvg under both statistical and system heterogeneity.
+For strongly convex and convex problems, we also characterize the corresponding convergence rates for the Nesterov accelerated FedAvg algorithm, which are the first linear speedup guarantees for momentum variants of FedAvg in the convex setting.
+To provably accelerate FedAvg, we design a new momentum-based FL algorithm that further improves the convergence rate in overparameterized linear regression problems.
 Empirical studies of the algorithms in various settings have supported our theoretical results.
@@
 \input{sec_Nesterov}
-\input{sec_overparameterized}
+%\input{sec_overparameterized}
 \input{sec_exp}
@@
 
+% \newpage
+% \input{sec_impact}
 \newpage
-\input{sec_impact}
-\bibliographystyle{unsrt}
+\bibliographystyle{iclr2021_conference}
 \bibliography{ref.bib}

### 2020-09-25 source-iclr/sec_Nesterov.tex
Update on Overleaf.

@@
 updates focus on the non-convex smooth case \cite{huo2020faster,yu2019linear,li2018federated}.
-In this section, we complete the picture with $\mathcal{O}(1/NT)$
-and $\mathcal{O}(1/\sqrt{NT})$ convergence results for Nesterov-accelerated
-FedAvg for convex objectives that match the rates from the previous section.
-As we know from stochastic optimization, Nesterov and other momentum
-updates may fail to accelerate over SGD \cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2016influence}.
-Therefore in Section~\ref{sec:overparameterized} we will specialize to overparameterized
-problems where we demonstrate that a particular FedAvg variant with
-momentum updates is able to accelerate over the original FedAvg algorithm. 
-Detailed proofs of convergence results in this section are deferred to Appendix Section~\ref{sec:app:Nesterovfedavg}.
+In this section, we complete the picture by providing the first $\mathcal{O}(1/KT)$
+and $\mathcal{O}(1/\sqrt{KT})$ convergence results for Nesterov-accelerated
+FedAvg for convex objectives that match the rates from the previous section. Detailed proofs of convergence results in this section are deferred to Appendix Section~\ref{sec:app:Nesterovfedavg}.
 
@@
 \end{theorem}
+The key step in the proof of this result is a one step contraction that involves three time steps, due to the update format of Nesterov SGD, but shares similar elements as that for FedAvg:
+\begin{align*}
+\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
+& +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|
+\end{align*}
 %
@@
 
-We now show that the optimality gap of Nesterov Accelerated FedAvg has $\mathcal{O}(1/\sqrt{NT})$ rate. This result complements the strongly convex case in the previous
+We now show that the optimality gap of Nesterov Accelerated FedAvg has $\mathcal{O}(1/\sqrt{KT})$ rate. This result complements the strongly convex case in the previous
 part, as well as the non-convex smooth setting in \cite{huo2020faster,yu2019linear,li2018federated},
-where a similar $\mathcal{O}(1/\sqrt{NT})$ rate is given in terms
+where a similar $\mathcal{O}(1/\sqrt{KT})$ rate is given in terms
 of averaged gradient norm. 
@@
 	It is possible to extend the results in this section to accelerated
-    FedAvg algorithms with other momentum-based updates. However, in the
-    stochastic optimization setting, none of these methods can achieve a better rate than the original FedAvg with SGD updates for general problems~\cite{kidambi2018insufficiency}.
+    FedAvg algorithms with other momentum-based updates. However, as know from stochastic optimization, Nesterov and other momentum
+updates may fail to accelerate over SGD (\cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2016influence}).
     For this reason, we will instead turn to the overparameterized setting
-    \cite{ma2017power,liu2018accelerating,canziani2016analysis} in the
-    next section where we show that FedAvg enjoys geometric convergence
-    and it is possible to improve its convergence rate with momentum-based updates.
+    \cite{ma2017power,liu2018accelerating,canziani2016analysis} in Section~\ref{sec:app:overparameterized} of the appendix where we show that FedAvg enjoys geometric convergence
+    and it is possible to improve its convergence rate with a new momentum-based FedAvg variant, which we term ``FedMaSS''.

### 2020-09-25 source-iclr/sec_SGD.tex
Update on Overleaf.

--- 2020-09-24 source-iclr/sec_SGD.tex
@@
 
-In this section, we provide convergence analyses of FedAvg for convex objectives in the general setting with both heterogeneous data and partial
-participation. We show that for strongly convex and smooth objectives,
+In this section, we provide convergence analyses of FedAvg for convex objectives in the general setting with both heterogeneous data (statistical heterogeneity) and partial
+participation (system heterogeneity). We show that for strongly convex and smooth objectives,
 the convergence of the optimality gap of averaged parameters across
-devices is $\mathcal{O}(1/NT)$, while for convex and smooth
-objectives, the rate is $\mathcal{O}(1/\sqrt{NT})$. 
+devices is $\mathcal{O}(1/KT)$, while for convex and smooth
+objectives, the rate is $\mathcal{O}(1/\sqrt{KT})$.
+Our results improve upon~\cite{li2019convergence,karimireddy2019scaffold} by showing linear speedup for any number of participating devices, and upon~\cite{khaled2020tighter,koloskova2020unified} by allowing system heterogeneity.
+The proofs also highlight similarities and distinctions between the strongly convex and convex settings.
 Detailed proofs are deferred to Appendix Section~\ref{sec:app:fedavg}.
@@
 
-We first show that FedAvg has an $\mathcal{O}(1/NT)$ convergence rate
-for $\mu$-strongly convex and $L$-smooth objectives. The result
-improves on the $\mathcal{O}(1/T)$ rate of \cite{li2019convergence}
-with a linear speedup in the number of devices $N$. Moreover, it
+We first show that FedAvg has an $\mathcal{O}(1/KT)$ convergence rate
+for $\mu$-strongly convex and $L$-smooth objectives. The result relies on a technical improvement over the analysis in~\cite{li2019convergence}. Moreover, it
 implies a distinction in communication efficiency that guarantees
@@
 the other hand, with partial participation, $E$ must be $\mathcal{O}(1)$
-to guarantee $\mathcal{O}(1/NT)$ convergence.
+to guarantee $\mathcal{O}(1/KT)$ convergence.
 \begin{theorem}
@@
 \end{theorem}
-%
+A crucial ingredient in the proof of Theorem~\ref{th:scvx_sgd} is a one step contraction bound 
+	\begin{align*}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}L\alpha_{t}^{3}G^{2} 
+	\end{align*}
+that improves upon the analysis in~\cite{li2019convergence} to deliver the linear speedup.
+Moreover, analogues of this bound appear in the analyses of the convex setting as well as the Nesterov accelerated FedAvg.\\
 \textbf{Linear speedup. }We first compare our bound with that in \cite{li2019convergence},
@@
 	speedup} in the number of active workers during a communication
-round.
-We also emphasize that the reason one cannot recover the full participation bound by setting $K=N$ in the partial participation bound is due to the variance generated by sampling which depends on $E$.
-\\
+round. We also emphasize that the reason one cannot recover the full participation bound by setting $K=N$ in the partial participation bound is due to the variance generated by sampling.  \\
 % To our knowledge, this is the first result that explicitly demonstrates
@@
 still yields a linear speedup in $K$, which is also confirmed by
-experiments. The requirement $E=\mathcal{O}(1)$ in partial participation
-likely cannot be removed for our sampling schemes, as the sampling
-variance is $\Omega(E^{2}/T^{2})$ and the dependence on $E$ is tight.  \\
-\textbf{Comparison with related work.} To better understand the significance of the obtained bound, we compare our rates to the best-known results in related settings.
-\cite{haddadpour2019convergence} proves a linear speedup $\mathcal{O}(1/NT)$ result for strongly convex and smooth objectives, with $\mathcal{O}(N^{1/3}T^{2/3})$ communication complexity with \emph{i.i.d.} data and partial participation.
-However, their results build on the bounded gradient diversity assumption, which implies the existence of $\mathbf{w}^*$ that minimizes all local objectives (see discussions in Section~\ref{sec:assumptions} and Appendix~\ref{sec:app:comparison}), effectively removing system heterogeneity.
+experiments. The requirement $E=\mathcal{O}(1)$ in partial participation cannot be removed for our sampling schemes, as the sampling
+variance is $\Omega(E^{2}/T^{2})$ and the dependence on $E$ is tight (see Proposition~\ref{prop:tight} in Section~\ref{sec:app:fedavg} of the appendix).  \\
+\textbf{Comparison with related works.} To better understand the significance of the obtained bound, we compare our rates to the best-known results in related settings.
+\cite{haddadpour2019convergence} proves a linear speedup $\mathcal{O}(1/KT)$ result for strongly convex and smooth objectives\footnote{Their result applies to a larger class of non-convex objectives that satisfy the Polyak-Lojasiewicz condition.}, with $\mathcal{O}(K^{1/3}T^{2/3})$ communication complexity with non-\emph{i.i.d.} data and partial participation.
+However, their results build on the bounded gradient diversity assumption, which implies the existence of $\mathbf{w}^*$ that minimizes all local objectives (see discussions in Section~\ref{sec:assumptions} and Appendix~\ref{sec:app:comparison}), effectively removing statistical heterogeneity.
 The bound in \cite{koloskova2020unified} matches our bound in the full participation case, but their framework excludes partial participation \cite[Proposition 1]{koloskova2020unified}.
+The result of~\cite{karimireddy2019scaffold} applies to the full FL setting, but only has linear speedup when $K=\mathcal{O}(N)$, i.e.
+close to full participation, whereas our result has linear speedup for any number of participating devices.
 
@@
 \subsection{Convex Smooth Objectives}
-Next we provide linear speedup analyses of FedAvg with convex and
-smooth objectives and show that the optimality gap is $\mathcal{O}(1/\sqrt{NT})$. 
+Next we provide linear speedup analysis of FedAvg with convex and
+smooth objectives and show that the optimality gap is $\mathcal{O}(1/\sqrt{KT})$. 
 This result complements the strongly convex case in the previous part, as well as the non-convex
 smooth setting in \cite{jiang2018linear,yu2019parallel,haddadpour2019convergence},
-where $\mathcal{O}(1/\sqrt{NT})$ results are given in terms of averaged
-gradient norm. 
+where $\mathcal{O}(1/\sqrt{KT})$ results are given in terms of averaged
+gradient norm, and it also extends the result in~\cite{khaled2020tighter}, which has linear speedup in the convex setting, but only for full participation.
 \begin{theorem}
@@
 \end{theorem}
+The proof again relies on a one step bound, without contraction, which is then summed over time steps: 
+\begin{align*}
+	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
+	\end{align*}
 %

### 2020-09-25 source-iclr/sec_exp_details.tex
Update on Overleaf.

--- 2020-06-03 source/sec_exp_details.tex
@@
 
-\section{Details on Experiments and additional results}
+\section{Details on Experiments and Additional Results}
 % \subsection{Experiments}
@@
 
-We described the precise procedure to reproduce the results in this paper.
+We describe the precise procedure to reproduce the results in this paper.
 As we mentioned in Section~\ref{sec:exp}, we empirically verified the
@@
 In the next following subsections,
-we introduce parametering searching in each objective separately.
+we introduce parameter searching in each objective separately.
 
@@
 \subsection{Partial Participation}
-To examine the linear speedup of FedAvg in partial participation experiments,
+To examine the linear speedup of FedAvg in partial participation setting,
 we evenly distributed data on $4, 8, 16, 32, 64, 128$ devices and 
-uniformly sample $50\%$ devices without replacement. All other hyperparameters
-are the same as previous sections.
+uniformly sample $50\%$ devices without replacement. 
+All other hyperparameters are the same as previous sections.
 
@@
 \begin{align*}
-y_{t+1}^{k} & =w_{t}^{k}-\alpha_{t}g_{t,k}\\
-w_{t+1}^{k} & =\begin{cases}
-y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E}\\
-\sum_{k\in\mathcal{S}_{t+1}}\left(y_{t+1}^{k}+\beta_{t}(y_{t+1}^{k}-y_{t}^{k})\right) & \text{if }t+1\in\mathcal{I}_{E}
+\vy_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}\vg_{t,k}\\
+\vw_{t+1}^{k} & =\begin{cases}
+\vy_{t+1}^{k}+\beta_{t}(\vy_{t+1}^{k}-\vy_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E}\\
+\sum_{k\in\mathcal{S}_{t+1}}\left(\vy_{t+1}^{k}+\beta_{t}(\vy_{t+1}^{k}-\vy_{t}^{k})\right) & \text{if }t+1\in\mathcal{I}_{E}
 \end{cases}

### 2020-09-25 source-iclr/sec_intro.tex
Update on Overleaf.

@@
 \section{Introduction}
-Federated learning (FL) is a machine learning setting where many clients (e.g., mobile devices or organizations) collaboratively train a model under the orchestration of a central server (e.g., service provider), while keeping the training data decentralized~\cite{smith2017federated, kairouz2019advances}.
-In recent years, FL has swiftly emerged as an important learning paradigm~\cite{mcmahan2016communication,li2018federated}--one that enjoys widespread success in applications such as personalized recommendation~\cite{chen2018federated}, virtual assistant~\cite{lamautonomy}, and keyboard prediction~\cite{47586}, to name a few--for at least two reasons: First, the rapid proliferation of smart devices that are equipped with both computing power and data-capturing capabilities provided the infrastructure core for FL.
-Second, the rising awareness of privacy and the exponential growth of computational power in mobile devices have made it increasingly attractive to push the computation to the edge.
+Federated learning (FL) is a machine learning paradigm where many clients (e.g., mobile devices or organizations) collaboratively train a model under the orchestration of a central server (e.g., service provider), while keeping the training data decentralized (\cite{smith2017federated, kairouz2019advances}).
+In recent years, FL has swiftly emerged as an important learning paradigm (\cite{mcmahan2016communication,li2018federated})--one that enjoys widespread success in applications such as personalized recommendation (\cite{chen2018federated}), virtual assistant (\cite{lam2019protecting}), and keyboard prediction (\cite{47586}), to name a few--for at least three reasons: First, the rapid proliferation of smart devices that are equipped with both computing power and data-capturing capabilities provided the infrastructure core for FL.
+Second, the rising awareness of privacy and the explosive growth of computational power in mobile devices have made it increasingly attractive to push the computation to the edge.
+Third, the empirical success of communication-efficient FL algorithms has enabled increasingly larger-scale parallel computing and learning with less communication overhead.
 
 Despite its promise and broad applicability in our current era, the potential value FL delivers is coupled with the unique challenges it brings forth.
-In particular, when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated~\cite{kairouz2019advances}, it faces two challenges that are absent in centralized optimization and distributed (stochastic) optimization~\cite{zhou2017convergence, stich2018local,khaled2019first,liang2019variance,wang2018cooperative,woodworth2018graph,wang2019adaptive,jiang2018linear,yu2019parallel,yu2019linear, khaled2020tighter,koloskova2020unified}:
+In particular, when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated (\cite{kairouz2019advances}), it faces two challenges that are absent in centralized optimization and distributed (stochastic) optimization (\cite{zhou2017convergence, stich2018local,khaled2019first,liang2019variance,wang2018cooperative,woodworth2018graph,wang2019adaptive,jiang2018linear,yu2019parallel,yu2019linear, khaled2020tighter,koloskova2020unified}):
 
-1) \textbf{Data heterogeneity:} data distributions in devices are different (and data cannot be shared);
+1) \textbf{Data (statistical) heterogeneity:} data distributions in devices are different (and data cannot be shared);
 
@@
 
-To address these challenges, Federated Averaging (FedAvg)~\cite{mcmahan2016communication} was proposed as a particularly effective heuristic, which has enjoyed great empirical success~\cite{47586}.
+To address these challenges, Federated Averaging (FedAvg)~\cite{mcmahan2016communication} was proposed as a particularly effective heuristic, which has enjoyed great empirical success.
 This success has since motivated a growing line of research efforts into understanding its theoretical convergence guarantees in various settings.
@@
 Very recently, \cite{li2019convergence} furthered the progress and established an $\cO(\frac{1}{T})$ convergence rate for FedAvg for strongly convex smooth Federated learning problems with both data and system heterogeneity.
-A concurrent result in the same setting~\cite{karimireddy2019scaffold} established a similar $\cO(\frac{1}{T})$ result that allows for a linear speedup when the number of participating devices is large.
+A similar result in the same setting~\cite{karimireddy2019scaffold} also established an $\cO(\frac{1}{T})$ result that allows for a linear speedup when the number of participating devices is large.
 At the same time, \cite{huo2020faster} studied the Nesterov accelerated FedAvg for non-convex smooth problems and established 
@@
 What about when FedAvg is accelerated with momentum updates?
-Does the presence of both data and system heterogeneity in FL require major technical novelties over results from distributed and decentralized optimization?
+Does the presence of both data and system heterogeneity in FL imply different communication complexities and require technical novelties over results in distributed and decentralized optimization?
 These aspects are currently unexplored or underexplored in FL.
@@
 $T$ is the total number of stochastic updates performed by each local device, $E$ is the local steps between two consecutive server communications (and hence $T/E$ is the number of communications).
-$^{\dagger}$ In the linear regression setting, we have $\kappa=\kappa_1$ for FedAvg and $\kappa=\sqrt{\kappa_1\tilde{\kappa}}$ for accelerated FedAvg, where $\kappa_1$ and $\sqrt{\kappa_1\tilde{\kappa}}$ are condition numbers defined in Section \ref{sec:overparameterized}.
+$^{\dagger}$ In the linear regression setting, we have $\kappa=\kappa_1$ for FedAvg and $\kappa=\sqrt{\kappa_1\tilde{\kappa}}$ for momentum accelerated FedAvg (FedMaSS), where $\kappa_1$ and $\sqrt{\kappa_1\tilde{\kappa}}$ are condition numbers defined in Section \ref{sec:app:overparameterized}.
 Since
@@
 \textbf{Our Contributions}
-We provide a comprehensive convergence analysis
+We provide a comprehensive and unified convergence analysis
 of FedAvg and its accelerated variants in the presence of both data and system heterogeneity. 
@@
 As for convex and smooth problems, there was no prior work that established the {\small{$\cO(\frac{1}{\sqrt{T}})$}} rate under both system and data heterogeneity.
+Our unified analysis highlights the common elements and distinctions between the strongly and convex settings.
 
@@
 To the best of our knowledge, these are the first results that provide a linear speedup characterization of Nesterov accelerated FedAvg in those two problem classes (that FedAvg and Nesterov accelerated FedAvg share the same convergence rate is to be expected: this is the case even for centralized stochastic optimization).
-Prior to our results here, the most relevant results~\cite{yu2019linear} only concern the non-convex setting, where convergence is measured with respect to stationary points (vanishing of gradient norms, rather than optimality gaps).
+Prior to our results here, the most relevant results~\cite{yu2019linear,li2018federated,huo2020faster} only concern the non-convex setting, where convergence is measured with respect to stationary points (vanishing of gradient norms, rather than optimality gaps).
+Our unified analysis of Nesterov FedAvg also illustrates the technical similarities and distinctions compared to the original FedAvg algorithm, whereas prior works (in the non-convex setting) were scattered and used different notations.
 
@@
 
-\textcolor{red}{ADD COMPARISON WITH LOCAL SGD AND OTHER DISTRIBUTED OPTIMIZATION RESULTS}
+\textbf{Connections with Distributed and Decentralized Optimization}
+Federated learning is closely related to distributed and decentralized optimization, and as such it is important to discuss connections and distinctions between our work and related results from that literature.
+First, when there is neither system heterogeneity, i.e.
+all devices participate in parameter averaging during a communication round, nor statistical heterogeneity, i.e.
+all devices have access to a common set of stochastic gradients, FedAvg coincides with the ``Local SGD'' of~\cite{stich2018local}, which showed the linear speedup rate $\mathcal{O}(1/NT)$ for strongly convex and smooth functions.
+When there is only data heterogeneity, some works have continued to use the term Local SGD to refer to FedAvg, while others subsume it in more general frameworks that include decentralized model averaging based on a network topology or a mixing matrix.
+They have provided linear speedup analyses for strongly convex and convex problems, e.g.~\cite{khaled2020tighter,koloskova2020unified} as well as non-convex problems, e.g.~\cite{jiang2018linear,yu2019parallel,wang2018cooperative}.
+However, these results do not consider system heterogeneity, i.e.
+the presence of stragglers in the device network.
+Even with decentralized model averaging, the assumptions usually imply that model averages over all devices is the same as decentralized model averages based on network topology (e.g.~\cite{koloskova2020unified} Proposition 1), which precludes system heterogeneity as defined in this paper and prevalent in FL problems.
+For momentum accelerated FedAvg,~\cite{yu2019linear} provided linear speedup analysis for non-convex problems, while results for strongly convex and convex settings are entirely lacking, even without system heterogeneity.~\cite{karimireddy2019scaffold} considers both types of heterogeneities, but their rate implies a linear speedup only when the number of stragglers is negligible.
+In contrast, our linear speedup analyses consider both types of heterogeneity present in the full federated learning setting, and are valid for any number of participating devices.
+We also highlight a distinction in communication efficiency when system heterogeneity is present.
+Moreover, our results for Nesterov accelerated FedAvg completes the picture for strongly convex and convex problems.
+For a detailed comparison with related works, please refer to Table~\ref{tb:convergenceratev3} in Appendix Section~\ref{sec:app:comparison}.
 

### 2020-09-25 source-iclr/sec_problem.tex
Update on Overleaf.

@@
 
-With partial participation, we follow~\cite{li2019convergence} and~\cite{li2018federated} and assume that $\cS_{t+1}$ is obtained by one of two types of
+With partial participation, we follow~\cite{li2018federated,karimireddy2019scaffold,li2019convergence} and assume that $\cS_{t+1}$ is obtained by one of two types of
 sampling schemes to simulate practical scenarios.
@@
 bounded gradient diversity restricts local objectives to have the same minimizer as the global objective, contradicting the heterogeneous data setting. 
-For detailed discussions of our assumptions and additional comparisons with related works, please refer to Appendix Section~\ref{sec:app:comparison}.
+For detailed discussions of our assumptions, please refer to Appendix Section~\ref{sec:app:comparison}.
 

### 2020-09-25 source-iclr/app_unifiedframework.tex
Update on Overleaf.

In this section, we provide a unified proof framework for FedAvg and its accelerated variants.

### 2020-09-27 source-iclr/app_sgd_cvx_smooth.tex
D2 first round
Update on Overleaf.

@@
     \begin{align*}
-	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} \\ 
+	& +2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
 	\end{align*}
@@
 	strategy and preserve $\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]$
-	in order to obtain a telescoping sum.
+	in order to obtain the desired optimality gap.
 
@@
 	& \|\ov{w}_{t+1}-\vw^{\ast}\|^{2}\\
-	& \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]\\
+	\leq & \|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]\\
 	& +2\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\vw_{t}^{k}-\ov{w}_{t}\|^{2}+4\alpha_{t}^{2}L(F(\ov{w}_{t})-F(\vw^{\ast}))+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-	& =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-	& +\alpha_{t}(1-4\alpha_{t}L)(F(\vw^{\ast})-F(\ov{w}_{t}))
+	= & \|\ov{w}_{t}-\vw^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right] \\ 
+	 & +\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+	 +\alpha_{t}(1-4\alpha_{t}L)(F(\vw^{\ast})-F(\ov{w}_{t}))
 	\end{align*}
@@
 	& \|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast}))\\
-	& \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-	& \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\frac{3}{2}\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+ \leq & \|\ov{w}_{t}-\vw^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+	\leq & \|\ov{w}_{t}-\vw^{\ast}\|^{2}+\frac{3}{2}\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
 	\end{align*}
@@
 & =(\|\vw_{0}-\vw^{\ast}\|^{2}+\nu_{\max}^{2}\sigma^{2})\frac{1}{\sqrt{NT}}+\frac{N}{T}6E^{2}LG^{2}\\
-& =O(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T})
+& =\mathcal{O}(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T})
 \end{align*}

### 2020-09-27 source-iclr/sec_Nesterov.tex
Update on Overleaf.

@@
 \end{theorem}
-The key step in the proof of this result is a one step contraction that involves three time steps, due to the update format of Nesterov SGD, but shares similar elements as that for FedAvg:
+Similar to FedAvg, the key step in the proof of this result is a recursive contraction bound, but different in that it involves three time steps, due to the update format of Nesterov SGD:
 \begin{align*}
@@
 \end{align*}
+and from this one can again use induction and $L$-smoothness to obtain the desired bound.
 %
@@
 
-We now show that the optimality gap of Nesterov Accelerated FedAvg has $\mathcal{O}(1/\sqrt{KT})$ rate. This result complements the strongly convex case in the previous
+We now show that the optimality gap of Nesterov Accelerated FedAvg has $\mathcal{O}(1/\sqrt{KT})$ rate for convex and smooth objectives.
+This result complements the strongly convex case in the previous
 part, as well as the non-convex smooth setting in \cite{huo2020faster,yu2019linear,li2018federated},

### 2020-09-27 source-iclr/sec_SGD.tex
Update on Overleaf.

@@
 \end{theorem}
+\textbf{Proof sketch.} Because our unified analyses of results in the main text follow the same framework with variations in technical details, we first give an outline of proof  for Theorem~\ref{th:scvx_sgd} to illustrate the main ideas.
+For full participation, the main ingredient is a recursive contraction bound
+	\begin{align*}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6\alpha_{t}^{3}LE^{2}G^{2}
+	\end{align*}
+ where the $\mathcal{O}(\alpha_{t}^{3}E^{2}G^{2})$ term is the key improvement over the bound in~\cite{li2019convergence}, which has $\mathcal{O}(\alpha_{t}^{2}E^2 G^2)$ instead.
+We then use induction to obtain a non-recursive bound on $\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}$, which is converted to a bound on $\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}$ using $L$-smoothness.
+For partial participation, an additional sampling variance term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order is added to the contraction bound.
+
+\begin{comment}
 A crucial ingredient in the proof of Theorem~\ref{th:scvx_sgd} is a one step contraction bound 
@@
 Moreover, analogues of this bound appear in the analyses of the convex setting as well as the Nesterov accelerated FedAvg.\\
-\textbf{Linear speedup. }We first compare our bound with that in \cite{li2019convergence},
+\end{comment}
+
+\textbf{Linear speedup. }We compare our bound with that in \cite{li2019convergence},
 which is $\mathcal{O}(\frac{1}{NT}+\frac{E^{2}}{KT}+\frac{E^{2}G^{2}}{T})$.
@@
 \end{theorem}
-The proof again relies on a one step bound, without contraction, which is then summed over time steps: 
+The analysis again relies on a recursive bound, but without contraction: 
 \begin{align*}
@@
 	\end{align*}
+which is then summed over time steps to give the desired bound, with $\alpha_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$. \\
 %

### 2020-09-28 source-iclr/app_unifiedframework.tex
appendix C: UNIFIED FRAMEWORK v1

@@
-In this section, we provide a unified proof framework for FedAvg and its accelerated variants.
+% !TEX ROOT=./main.tex
+This section provides a unified proof framework for the convergence analysis of FedAvg and its accelerated variants. To facilitate our presentation of the unified framework,
+we first outline the FedAvg in Alg~\ref{alg:fedavg} and we will take 
+the strongly convex problem as an example to illustrate our analysis. 
+The corresponding proof for general convex functions follows the 
+same framework.
+
+% The main roadmap of the convergence proof in this work contain the
+% following steps. 
+% First, the goal of our analysis is to provide the convergence rate 
+% of FedAvg and its accelerated variants, which means the we would like
+% to show how does the optimality gap decrease as we conduct more 
+% iterations 
+First, recall that our goal is to analyze the convergence rate of FedAvg algorithm, i.e., $F(\vw_t) - F^* \leq \cO(1/T)$ in strongly convex 
+case. 
+To prove this bound, we need to connect the optimality gap with the Frobenius norm
+of current model and the optimal model ($\| \vw_t - \vw^*\|^2$). Then, we 
+analyze the one step progress of FedAvg. This step rigorously shows how 
+much the distance between $\vw_{t+1}$ and $\vw^*$ becomes closer when 
+we take one step stochastic gradient descent in FedAvg. More specifically,
+we provide an upper bound of $\|\vw_{t+1} - \vw^*\|^2$ using the previous
+step $\| \vw_t - \vw^*\|^2$. This will justify a solid theoretical foundation for the step in line 9, Alg~\ref{alg:fedavg}. More importantly, our 
+prove also motivate that we should include more clients to accelerate the
+convergence as we show $\cO(1/NT)$.  Notice that this is the central step of our analysis, which is the main source of improvement of our results comparing to 
+prior arts. 
+Using one step progress bound, we can connect $\|\vw_t - \vw^*\|^2$ to 
+$\|\vw_0 - \vw^*\|^2$ by induction. Therefore, we have an upper bound
+of $\|\vw_t - \vw^*\|^2$ and lastly, we can use the upper bound
+of $\|\vw_t - \vw^*\|^2$ to bound the optimality gap. 
+For strongly convex problems, we can easily connect the optimality gap
+and using L-smooth properties.
+
+In conclusion, our proof contains two main steps: 1) \textbf{one step progress bound} prove how much progress we have made by taking one step
+SGD in FedAvg. 2) \textbf{Iterative proof} bounds the up to date progress using one step progress bound iteratively.
+
+\begin{algorithm}[h!]\small
+\begin{algorithmic}[1]
+\STATE \textbf{Server input:} initial model $\vw_0$, initial step size $\alpha_0$, local steps $E$. 
+\STATE \textbf{Client input:} 
+\FOR {each round $r = 0, 1, ..., R$, where $r = t*E$} 
+\STATE  Sample clients $\cS_t \subseteq \{1,...,N\}$
+\STATE Broadcast $\vw$ to all clients $k \in \cS_t$
+\FOR {each client $k \subseteq \cS_t$}
+\STATE initialize local model $\vw_t^k = \vw$
+\FOR {$t = r * E + 1, \dots, (r+1)*E$}
+\STATE $\vw_{t+1}^{k}  =\vw_{t}^{k}-\alpha_{t}\vg_{t,k}$
+\ENDFOR
+\ENDFOR
+\STATE Average the local models at server end: $\ov{w}_t = \sum_{k\in \cS_t} \vw_t^k$.
+\ENDFOR 
+\end{algorithmic}
+\caption{\textsc{FedAvg}: Federated Averaging}% $S$ = cA3C($\theta, \Theta_v$
+\label{alg:fedavg}
+\end{algorithm}

### 2020-09-30 source-iclr/sec_Nesterov.tex
Update on Overleaf.
iclr r1

@@
 Then under Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2} with full device participation,
-	\begin{align*}
+	% \vspace{-1em}
+	\small{\begin{align*}
 	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right),
-	\end{align*}
+	\end{align*}}
+	% \vspace{-1em}
 	and with partial device participation with $K$ sampled devices at
 	each communication round, 
+	\small{
 	\begin{align*}
 	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).
-	\end{align*}
+	\end{align*}}
 \end{theorem}
-Similar to FedAvg, the key step in the proof of this result is a recursive contraction bound, but different in that it involves three time steps, due to the update format of Nesterov SGD:
-\begin{align*}
-\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
-& +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|
-\end{align*}
-and from this one can again use induction and $L$-smoothness to obtain the desired bound.
+Similar to FedAvg, the key step in the proof of this result is a recursive contraction bound, but different in that it involves three time steps, due to the update format of Nesterov SGD (see Lemma~\ref{lem:nest-scvxoner} in Appendix~\ref{sec:convexsmoothsgd}).
+% \begin{align*}
+% \mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
+% & +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|
+% \end{align*}
+Then we can again use induction and $L$-smoothness to obtain the desired bound.
 %
@@
 	full device participation has rate
-	\begin{align*}
+	% \vspace{-0em}
+	\small{\begin{align*}
 		\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T}\right),
-	\end{align*}
+		\end{align*}}
+		% \vspace{-1em}
 	and with partial device participation with $K$ sampled devices at
 	each communication round, 
-	\begin{align*}
+	\small{\begin{align*}
 		\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right).
-	\end{align*}
+		\end{align*}}
 \end{theorem}

### 2020-11-22 source-iclr/sec_Nesterov.tex
Update on Overleaf.

@@
 
-A natural extension of the FedAvg algorithm is to use momentum-based
-local updates instead of local SGD updates. To our knowledge, the
+{\color{red}A natural extension of the FedAvg algorithm is to use momentum-based 
+local updates instead of local SGD updates in order to accelerate FedAvg. As we know from stochastic optimization, Nesterov and other momentum
+updates fail to provably accelerate over SGD (\cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2020federated}).
+This is in contrast to the classical acceleration result of Nesterov-accelerated gradient descent over GD.
+Thus in the FL setting, the best provable convergence rate for FedAvg with Nesterov updates is the same as FedAvg with SGD updates.
+Nevertheless, Nesterov and other momentum updates are frequently used in practice, in both non-FL and FL settings, and are observed to perform better empirically.
+In fact, previous works such as \cite{stich2018local} on FedAvg with vanilla SGD uses FedAvg with Nesterov or other momentum updates in their experiments to achieve target accuracy.
+Because of the popularity of Nesterov and other momentum-based methods, understanding the linear speedup behavior of FedAvg with such local updates is important.} To our knowledge, the
 only convergence analyses of FedAvg with momentum-based stochastic
-updates focus on the non-convex smooth case \cite{huo2020faster,yu2019linear,li2018federated}.
+updates focus on the non-convex smooth case \cite{huo2020faster,yu2019linear,li2018federated}, and no results existed in the convex smooth settings.
 In this section, we complete the picture by providing the first $\mathcal{O}(1/KT)$
 and $\mathcal{O}(1/\sqrt{KT})$ convergence results for Nesterov-accelerated
-FedAvg for convex objectives that match the rates from the previous section. Detailed proofs of convergence results in this section are deferred to Appendix Section~\ref{sec:app:Nesterovfedavg}.
+FedAvg for convex objectives that match the rates for FedAvg with SGD updates. Detailed proofs of convergence results in this section are deferred to Appendix Section~\ref{sec:app:Nesterovfedavg}.
 
@@
 
-We now show that the optimality gap of Nesterov Accelerated FedAvg has $\mathcal{O}(1/\sqrt{KT})$ rate for convex and smooth objectives.
+We now show that the optimality gap of Nesterov-accelerated FedAvg has $\mathcal{O}(1/\sqrt{KT})$ rate for convex and smooth objectives.
 This result complements the strongly convex case in the previous
@@
 \end{theorem}
-%
+
+We emphasize again that in the stochastic optimization setting, the optimal convergence rate that FedAvg wtih Nesterov udpates can achieve is the same as FedAvg with SGD updates.
+However, due to the popularity and superior performance of momentum methods in practice, it is still important to understand the linear speedup behavior of such FedAvg variants.
+Our results in this section fill exactly this gap.
+
+	\begin{comment}
 	It is possible to extend the results in this section to accelerated
@@
 updates may fail to accelerate over SGD (\cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2016influence}).
-    For this reason, we will instead turn to the overparameterized setting
+    For this reason, we instead turn to the overparameterized setting
     \cite{ma2017power,liu2018accelerating,canziani2016analysis} in Section~\ref{sec:app:overparameterized} of the appendix where we show that FedAvg enjoys geometric convergence
     and it is possible to improve its convergence rate with a new momentum-based FedAvg variant, which we term ``FedMaSS''.
+	\end{comment}

### 2020-11-22 source-iclr/sec_intro.tex
Update on Overleaf.

@@
 Despite its promise and broad applicability in our current era, the potential value FL delivers is coupled with the unique challenges it brings forth.
-In particular, when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated (\cite{kairouz2019advances}), it faces two challenges that are absent in centralized optimization and distributed (stochastic) optimization (\cite{zhou2017convergence, stich2018local,khaled2019first,liang2019variance,wang2018cooperative,woodworth2018graph,wang2019adaptive,jiang2018linear,yu2019parallel,yu2019linear, khaled2020tighter,koloskova2020unified}):
+In particular, when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated (\cite{kairouz2019advances}), it faces two challenges that are absent in centralized optimization and distributed (stochastic) optimization (\cite{zhou2017convergence, stich2018local,khaled2019first,liang2019variance,wang2018cooperative,woodworth2018graph,wang2019adaptive,jiang2018linear,yu2019parallel,yu2019linear, khaled2020tighter,koloskova2020unified,woodworth2020local,woodworth2020minibatch}):
 
@@
 all devices have access to a common set of stochastic gradients, FedAvg coincides with the ``Local SGD'' of~\cite{stich2018local}, which showed the linear speedup rate $\mathcal{O}(1/NT)$ for strongly convex and smooth functions.
+\cite{woodworth2020local} and \cite{woodworth2020minibatch} further improved the communication complexity that guarantees the linear speedup rate.
 When there is only data heterogeneity, some works have continued to use the term Local SGD to refer to FedAvg, while others subsume it in more general frameworks that include decentralized model averaging based on a network topology or a mixing matrix.
@@
 Even with decentralized model averaging, the assumptions usually imply that model averages over all devices is the same as decentralized model averages based on network topology (e.g.~\cite{koloskova2020unified} Proposition 1), which precludes system heterogeneity as defined in this paper and prevalent in FL problems.
-For momentum accelerated FedAvg,~\cite{yu2019linear} provided linear speedup analysis for non-convex problems, while results for strongly convex and convex settings are entirely lacking, even without system heterogeneity.~\cite{karimireddy2019scaffold} considers both types of heterogeneities, but their rate implies a linear speedup only when the number of stragglers is negligible.
+For momentum accelerated FedAvg,~\cite{yu2019linear} provided linear speedup analysis for non-convex problems, while results for strongly convex and convex settings are entirely lacking, even without system heterogeneity.~\cite{karimireddy2019scaffold} considers both types of heterogeneities for FedAvg, but their rate implies a linear speedup only when the number of stragglers is negligible.
 In contrast, our linear speedup analyses consider both types of heterogeneity present in the full federated learning setting, and are valid for any number of participating devices.

### 2021-02-06 source-iclrw/ICLR Rebuttal.tex
Add iclr workshop

\documentclass{article}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{color}
\usepackage{lipsum}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{enumitem}
\usepackage{comment}
\setlist{leftmargin=2mm}
\input{bmacros}
\newcommand{\ov}[1]{{\overline{\mathbf{#1}}}}
\begin{document}
% Based on their comments, upon acceptance of this paper, we shall improve in 
% the final version (a) Refine the presentation, fix typo, figures, and properly 
% located assumptions etc. (b) More experiments in terms of

% \input{r1}
{\color{blue}\textbf{Reviewer 1:}}

Thank you very much for your valuable and constructive feedback! Our point-to-point responses are provided below.

\textbf{\textit{1. Dependence on $E$ and comparison to distributed SGD.}} 
\begin{comment}
To clarify, $E$ is the number of local updates between consecutive communication rounds, and $T/E$ is the total number of communication rounds. Convergence speed is faster when the latter increases, but decreases when $E$ increases.
This is because due to the heterogeneity in data at local sources, when the number of local SGD updates $E$ increases, the parameter at each local client drifts further towards the optimizer of the local optimization problem rather than the global problem, and so the convergence of FedAvg is expected to degrade with increased $E$. However, this does not mean that we should always do model averaging at every iteration, as distributed SGD does, because communication can be costly, and this was one of the original motivations for Federated Learning. So increasing $E$ does not improve the convergence speed in terms of number of total updates, but decreases the number of communications required to achieve a rate (which can translate to faster wall clock time when communication is costly). For example, our results imply that the number of communication rounds $T/E$ can be as small as $\mathcal{O}(\sqrt{NT})$ in the full participation case without degrading the $\mathcal{O}(1/NT)$ linear speedup rate, whereas distributed SGD needs $T$ communications (since it averages at every iteration) to achieve the same $\mathcal{O}(1/NT)$ convergence rate. See the discussion under communication complexity on page 6. We apologize if this point was not made clear in the text, and can clarify it further in the main text if the reviewer thinks that it is necessary.
\end{comment}

In our notation, $E$ is the number of local updates between consecutive communication rounds, $T$ is the total number of iterations or equivalently the total number of SGD updates, and
$T/E$ is the total number of communication rounds.

The fact that an increased $E$ can lead to worse \emph{iteration complexity} holds true universally in both distributed optimization and federated learning settings, as the decrease in communication frequency among local clients results in slower convergence to the global optimizer. The benefit of increasing $E$ is to reduce the communication cost, as it decreases the number of communication rounds $T/E$ required to achieve an iteration complexity, which can then translate to faster \emph{wall clock time} when communication is costly. For example, our results imply that the number of communication rounds $T/E$ can be as small as $\mathcal{O}(\sqrt{NT})$ in the full participation case without degrading the $\mathcal{O}(1/NT)$ linear speedup rate, whereas distributed SGD needs $T$ communications (since it averages at every iteration) to achieve the same $\mathcal{O}(1/NT)$ convergence rate. So it is in this sense that our algorithm is faster compared to distributed SGD.

We apologize that this point was not made clear in the text, and have clarified it further in the main text of the updated version, e.g. in the discussions under communication complexity on page 6, as well as Table 2 in the appendix, where we provide the largest possible number of local updates $E$ that guarantees the linear speedup convergence.

\begin{comment}
Due to the heterogeneity in data at local sources, when the number of local SGD updates $E$ increases, the parameter at each local client drifts further towards the optimizer of the local optimization problem rather than the global problem, and so the convergence of FedAvg is expected to degrade with increased $E$.
\end{comment}

{\color{blue}\textbf{Reviewer 2:}}

Thank you very much for your valuable and constructive feedback! Our point-to-point responses are provided below.

\textbf{\textit{1. Choice of $E$ in partial participation.}}
The choice of $E$ in the partial participation case is tight due to the sampling variance of our sampling schemes. We have incorporated this observation on page 6 of the main text in the discussion under communication complexity, where we remark that the requirement $E=\mathcal{O}(1)$ cannot be removed for our particular sampling schemes, as the dependence on $E$ of the sampling variance $\mathcal{O}(E^2/T^2)$ is tight, and also refer to Proposition 1 in the Appendix, where we provide a problem instance that demonstrates the tightness. We can definitely place further emphasis on this point in the next round of updates if the reviewer thinks that is necessary.

\textbf{\textit{2. Significance of results in full FL setting.}} 
We agree that there has been a line of works in distributed optimization that demonstrate the linear speedup phenomenon in simpler settings without heterogeneous data or partial participation, and we have discussed and compared with these works in the introduction and after our main theorems. However, what distinguishes the full FL setting from these distributed optimization settings is the presence of both types of heterogeneities, and prior works were not able to demonstrate the linear speedup phenomenon with arbitrary number of participating devices in the full FL setting, so our work is the first to do so. Moreover, we provide a unified analysis of the results for convex and strongly convex problems, for both SGD-based and momentum-based FedAvg, and highlight the common elements and differences in their analysis, which also contributes to the understanding of linear speedup behavior of these algorithms under both data heterogeneity and system heterogeneity.

\textbf{\textit{3. Response to follow-up question.}}

Thanks for your response! The $E^2 G^2/\sqrt{KT}$ term in the bound in Theorem 2 indeed also comes from the sampling variance, in a way that is similar to the term $\frac{\kappa E^2 G^2 \mu}{KT}$ in Theorem 1. We apologize that we did not directly address the concern about $E^2 G^2/\sqrt{KT}$, and have incorporated it in the discussion following Theorem 2, as well as the updated proof for Theorem 2, where we spell out how the sampling variance $\mathbb{E}\|\ov{w}_t-\ov{v}_t\|^2$ becomes $E^2 G^2/\sqrt{KT}$ in the final bound. Essentially, Proposition 1 demonstrates that the sampling variance $\mathbb{E}\|\ov{w}_t-\ov{v}_t\|^2=\mathcal{O}(\alpha_t^2 E^2)$ cannot be made independent of $E$, which is true independent of whether we are in the strongly convex or convex case, since the sampling scheme does not depend on the convexity in any way. Due to this dependence on $E$, the extra term $E^2 G^2/\sqrt{KT}$ in the bound in Theorem 2 in the partial participation case cannot be made independent of $E$. Since it is the leading term, this restricts $E=\mathcal{O}(1)$ for partial participation.

{\color{blue}\textbf{Reviewer 3:}}

Thank you very much for your valuable and constructive feedback! Our point-to-point responses are provided below.

\textbf{\textit{1. Difference between Lemma 3 and Lemma 6.}}
While Lemma 3 and Lemma 6 have a similar form, the crucial difference between the two is that the bound in Lemma 3, under strong convexity, contains a contraction factor of $(1-\mu \alpha_t)$ in front of $\mathbb{E}\|\mathbf{\overline{\mathbf{w}}_{t}-\mathbf{w}^{*}}\|^2$, which is not present in the bound in Lemma 6 for convex problems. This contraction factor is what allows one to obtain an $\mathcal{O}(1/KT)$ rate for strongly convex problems, compared to the $\mathcal{O}(1/\sqrt{KT})$ rate for convex problems. The term $\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast}))$ actually also appears in the bound in Lemma 3, but since we do not need it, thanks to the contraction factor, we simplified the bound by dropping this non-negative term. Thus even though they look similar, their proofs are different and we have highlighted the differences with more emphasis in the paper, e.g. on page 19.

\textbf{\textit{2. Comparison with [2] and [3].}}
Thank you for pointing out our omission of these excellent references, and we have incorporated them in the updated version. We verified that even though the algorithm they study is also referred to as FedAvg, the setting in the two works do not allow either heterogeneous data or partial participation, in contrast to the FL setting considered in this paper. Thus even in the full participation case, our results are not directly comparable as we assume that the data that is available to each client is different, which departs from the classical distributed optimization setting. We have included comparison with these works in the updated version, but since Table 2 only includes papers that study data heterogeneous settings, we do not put these two works there, although can definitely change that if the reviewer thinks it is necessary.

\textbf{\textit{3,4. Non-acceleration of stochastic Nesterov FedAvg.}}
Indeed as the reviewer points out, full Nesterov-accelerated gradient descent with constant step size provably accelerates over vanilla GD. However, this breaks down with stochastic methods, and Nesterov SGD does not accelerate over SGD even in the non-FL setting. See e.g. \cite{kidambi2018insufficiency,liu2018accelerating,yuan2020federated}. Thus the best we can hope for in the FL setting for FedAvg with stochastic Nesterov updates is the same convergence rate with linear speedup as FedAvg with SGD, and \cite{yu2019linear} proves that this is the case for non-convex problems in terms of convergence to stationary points. No results existed in convex and strongly convex settings, so our results fill this gap.

While optimal theoretical results for Nesterov FedAvg are limited to the same rates as FedAvg with vanilla SGD, in practice FedAvg with Nesterov is observed to perform better empirically. In fact, many previous works such as \cite{stich2018local} on FedAvg with vanilla SGD uses Nesterov or other momentum versions in their experiments to achieve target accuracy. We have incorporated this observation in the section on Nesterov FedAvg to better motivate Theorems 3 and 4.

We agree that the algorithm in \cite{yuan2020federated} achieves the same linear speedup rates with a better communication complexity for general convex and strongly convex problems, but their setting does not allow either data heterogeneity or partial participation. Although their proposed accelerated algorithm is interesting from both theoretical and practical perspectives, Nesterov and other momentum-based algorithms are more commonly used in practice in both non-FL and FL settings. Therefore, we believe that our results on the convergence rates of Neseterov FedAvg provide theoretical understanding of the linear speedup behavior in the full FL setting and complement the result in \cite{yu2019linear} for non-convex problems to complete the picture.

Given this discussion, we believe that our optimal linear speedup results on Nesterov FedAvg are relevant due to the popularity of momentum methods in practice and the lack of theoretical understanding in convex settings. The overparameterized results are to illustrate the geometric convergence and to provide a provable acceleration of the geometric rate. Because of the relevance of the Nesterov results and the fact that their analyses can be unified with those of Theorems 1 and 2 to provide insights on the linear speedup, we decided to emphasize these and delegated the overparameterized results to the appendix.

In the updated version, we have emphasized that the bounds are the same for Nesterov and vanilla SGD, but can simplify it further. In particular, by ``reciting'' the bound do you mean e.g. referring to Theorem 1 in Theorem 3 directly?

\begin{comment}
{If the reviewer believes that changes in emphasis are still necessary given our responses and Reviewer 4's comments on the overparameterized setting, we are definitely willing to make the necessary changes.}
\end{comment}

{\color{blue}\textbf{Reviewer 4:}}

Thank you very much for your valuable and constructive feedback! Our point-to-point responses are provided below.

\textbf{\textit{1. Condition numbers in the bounds in the overparameterized setting.}}
The condition numbers $\kappa_1$ and $\tilde{\kappa}$ are defined in Section H.2. These quantities are defined in terms of the Hessian and stochastic Hessian matrices at each device and the global Hessian matrix, and are fixed and finite quantities given a particular problem instance. We use the fact that $\tilde{\kappa}\leq \kappa_1$. As the number of participating devices goes to infinity, the condition numbers converge to $\kappa:=L/\mu$, where $L$ and $\mu$ are the Lipschitz and smoothness constants of the global optimization objective. This is observed in e.g. \cite{liu2018accelerating}.

\textbf{\textit{2. Relevance of the overparameterized setting.}}
Our results in the overparameterized setting indeed assume strong convexity. While it is true that many overparameterized deep learning systems are non-convex, here by overparameterization we only mean that a global optimizer exists that yields 0 global loss, which does not restrict the problem to be non-convex. This type of overparameterization setting with strong convexity was studied in the works \cite{jain2017accelerating,liu2018accelerating,ma2017power} in the non-FL setting in order to develop provable accelerations over SGD, which motivated us to study extensions to the FL setting. Our main results highlight the geometric convergence with linear speedup in the overparameterized setting, which improves on the general strongly convex and convex settings, and provide a provable acceleration of momentum-based stochastic FedAvg algorithm over the vanilla FedAvg. While overparameterization is an important setting in practice, we agree that the applicability of our results is limited by the strong convexity assumption, and that our presentation of this topic in the main text is limited by constraints on space. We can remove the results in this section and include them in an extended version or a separate work since they are not as closely connected to the non-overparameterized setting, and we believe that they warrant a more thorough treatment and discussion to highlight their importance, which is limited by the current draft.

\begin{comment}
since it serves as "bonus" results on top of non-overparameterized settings.
\end{comment}

\textbf{Summary of updates}
We thank all reviewers for their positive feedback and appreciation of our work! 
We provided point-to-point responses to address the concerns of all reviewers separately and
incorporated the reviewers' suggestions in the updated version. We are glad to see R1 raise 
the score after reading our response. In the meantime, we are looking forward to 
responses from the other reviewers to clarify if there are further concerns.

We are glad that the reviewers have found our theory fill a gap in the literature (R2), 
appreciate its novelty (R3), and consider it qualified (R4) and new for the popular FedAvg algorithm under various settings (R1). We are also pleased that reviewers found our paper is well motivated (R4), well written (R3, R4) and appreciated our discussions of existing literature (R3, R4). We have further revised the paper to clarify all conceptual misunderstandings and provided individual responses as well. In particular, we have made the following updates and response in the revisions to address the reviewers' concerns:
1. Dependence on $E$ and comparison to distributed SGD (2nd paragraph, page 6). (R1)
2. Why $E$ has to be $\mathcal{O}(1)$ in theorem 2. (2nd paragraph, page 7, also see last sentence in 2nd paragraph, page 6 as Theorem 1 and Theorem 2 share the same reason that requires $E=\mathcal{O}(1)$) (R2)
3. Missing related works Woodworth et al. (2020b;a) (third paragraph, page 6) (R3)
4. The explanations on condition number and over-parameterized case. (see Appendix G2 first paragraph and our rebuttals) (R4).

Overall, we thank the reviewers for their positive and constructive comments that help improving this work. We would like to emphasize that this work provides the first linear speedup results under the more realistic non-IID data and partial participation Federated Learning scenarios for FedAvg and its accelerated variants, under various convex settings.

If there's any remaining concerns, please feel free to let us know, and we would be happy to provide a followup discussion.

\bibliographystyle{plain}
\bibliography{ref.bib}

\end{document}

### 2021-02-06 source-iclrw/app_notations.tex
Add iclr workshop

% !TEX ROOT=./main.tex

\section{Additional Notations and Bounds for Sampling Schemes}
\label{sec:app:notations}
In this section, we introduce additional notations that are used throughout
the proofs. Following common practice, e.g.~\cite{stich2018local,li2019convergence}, we define two virtual sequences $\overline{\mathbf{v}}_{t}=\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$ and $\overline{\mathbf{w}}_{t}=\sum_{k=1}^{N} p_{k} \mathbf{w}_{t}^{k}$, where we recall the FedAvg updates from \eqref{eq:fedavg updates}:
\begin{align*}
\vv_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}\vg_{t,k}, \hspace{1em}
\mathbf{w}_{t+1}^{k} =\left\{
\begin{array}{ll}
\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\ 
\sum_{k \in \cS_{t+1}} q_k \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}.
\end{array}\right.
\end{align*}
The following observations apply to FedAvg updates, while Nesterov accelerated FedAvg requires modifications. For full device participation or partial participation with $t \notin \cI_E$, note that
$\ov{v}_t = \ov{w}_t =\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$. For partial participation with $t \in \cI_E$, $\ov{w}_t \neq \ov{v}_t$ since $\ov{v}_t=\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$ while $\ov{w}_t=\sum_{k\in \cS_t}q_k\mathbf{w}_{t}^{k}$. However, we can
use unbiased sampling strategies such that $ \EE_{\cS_t} \ov{w}_t = \ov{v}_t$.
Note that $\overline{\mathbf{v}}_{t+1}$ is one-step SGD from $\overline{\mathbf{w}}_{t}$. 
\begin{align}
\overline{\mathbf{v}}_{t+1}=\overline{\mathbf{w}}_{t}-\alpha_{t} \mathbf{g}_{t},	\label{eq:vbar}
\end{align}
% $t+1 \in \cI_E$, we can fetch $\overline{\vw}_{t+1}$, we can communicate $\overline{\vw}_{t+1}$ to all devices.
where $\vg_{t} = \sum_{k=1}^{N} p_{k} \vg_{t,k} $ is the one-step stochastic gradient averaged over all devices. 
\begin{align*}
\vg_{t,k} = \nabla F_{k}\left(\mathbf{w}_{t}^{k},\mathbf{\xi}_{t}^{k} \right),   
\end{align*}
% \begin{align}
% \vg_{t,k} \left\{\begin{array}{ll} 
%  = \nabla F_{k}\left(\mathbf{w}_{t}^{k},\mathbf{\xi}_{t}^{k} \right)  &  \text{smooth}\\
%  \in \partial F_{k}\left(\vw_{t}^{k}, \mathbf{\xi}_{t}^{k}\right)  & \text{non-smooth} \lkxcom{remove non-smooth}
%  \end{array}\right.
% \end{align}
Similarly, we denote the expected one-step gradient $\ov{g}_{t}= \EE_{\xi_t}[\vg_t] = \sum_{k=1}^{N} p_{k} \EE_{\mathbf{\xi}_{t}^{k}} \vg_{t,k}$, where
\begin{align}
\EE_{\mathbf{\xi}_{t}^{k}} \vg_{t,k}  = \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right), 
\label{eq:gradient}
\end{align}
% \begin{align}
% \EE_{\mathbf{\xi}_{t}^{k}} \vg_{t,k}  \left\{\begin{array}{ll} 
%  = \nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)  &  \text{smooth}\\
%  \in \partial F_{k}\left(\vw_{t}^{k}\right)  & \text{non-smooth}
%  \end{array}\right.
% \end{align}
and $\mathbf{\xi}_t = \{\mathbf{\xi}_t^k\}_{k=1}^N$ denotes random samples at all devices at time step $t$. \\
Since in this work we also consider the case of partial participation, the sampling strategy to approximate the system heterogeneity can also affect the convergence. Here we
follow the prior works~\cite{li2019convergence} and~\cite{li2018federated} and consider two types of sampling
schemes that guarantee $ \EE_{\cS_t} \ov{w}_t = \ov{v}_t$. 
The sampling scheme I establishes $\cS_{t+1}$ by \emph{i.i.d.} sampling the devices according to probabilities $p_k$ with replacement, and setting $q_k=\frac{1}{K}$.
In this case the upper bound of expected square norm of $\ov{w}_{t+1} - \ov{v}_{t+1}$ is given by~\cite[Lemma 5]{li2019convergence}:
\begin{align}
\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4}{K} \alpha_t^2 E^2G^2.
\end{align}
The sampling scheme II establishes $\cS_{t+1}$ by uniformly sampling all devices without
replacement and setting $q_k=p_k\frac{N}{K}$, in which case we have
\begin{align}
\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2	\leq \frac{4(N - K)}{K(N-1)} \alpha_t^2 E^2G^2.
\end{align}
We summarize these upper bounds as follows: 
\begin{align}
	\EE_{\cS_{t+1}}\left\|\ov{w}_{t+1} - \ov{v}_{t+1}\right\|^2 \leq  \frac{4}{K} \alpha_t^2 E^2G^2.
	\label{eq:partialsample}
\end{align}
and this bound will be used in the convergence proof of the partial participation result.

### 2021-02-06 source-iclrw/app_sumarization.tex
Add iclr workshop

--- 2020-09-17 source-iclr/app_sumarization.tex
@@
 FedAvg\cite{karimireddy2019scaffold} & $\cO(\frac{1}{NT})^{\dagger\dagger}$ & $\cO(N^{-1/2}T^{1/2})^{\dagger\dagger}$   	& \cmark & Partial       & Bounded gradient dissimilarity   & Strongly convex  \\ \hline
-FedAvg/N-FedAvg                 & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{\ddagger}$ & \cmark	 & Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
+FedAvg/N-FedAvg (our work)                 & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{\ddagger}$ & \cmark	 & Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
 FedAvg\cite{khaled2020tighter}  & $\cO(\frac{1}{\sqrt{NT}})$ & $\cO(N^{-3/2}T^{1/2})$     	    & \cmark& Full        & Bounded gradient             & Convex  \\\hline
@@
 FedAvg\cite{karimireddy2019scaffold} & $\cO(\frac{1}{\sqrt{NT}})^{\dagger\dagger}$ & $\cO(N^{-3/4}T^{1/4})^{\dagger\dagger}$  & \cmark & Partial       & Bounded gradient dissimilarity   &  Convex  \\ \hline
-FedAvg/N-FedAvg      & $\cO\left(\frac{1}{\sqrt{KT}}\right)$ & $\cO(N^{-3/4}T^{1/4})^{\ddagger}$& \cmark			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
+FedAvg/N-FedAvg (our work)      & $\cO\left(\frac{1}{\sqrt{KT}}\right)$ & $\cO(N^{-3/4}T^{1/4})^{\ddagger}$& \cmark			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
 FedAvg & $\cO\left(\exp(-\frac{NT}{E\kappa_1})\right)$ & $ \cO(T^{\beta})$                   & \cmark&  Partial     & Bounded gradient    & Overparameterized LR\\ \hline

### 2021-02-06 source-iclrw/sec_SGD.tex
Add iclr workshop

--- 2020-09-27 source-iclr/sec_SGD.tex
@@
 We then use induction to obtain a non-recursive bound on $\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}$, which is converted to a bound on $\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}$ using $L$-smoothness.
-For partial participation, an additional sampling variance term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order is added to the contraction bound.
+For partial participation, an additional term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order resulting from sampling variance is added to the contraction bound.
+To facilitate the understanding of our analysis, please refer to a high-level summary in Appendix~\ref{sec:app:sum}.
 
@@
 achieve linear speedup. The improvement of our bound comes from the
-term $\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}$, which now is $\mathcal{O}(E^{2}/T^{2})$.
-As a result, all leading terms scale with $1/N$ in the full device
+term $\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}$, which now is $\mathcal{O}(E^{2}/T^{2})$ and so is not of leading order. As a result, all leading terms scale with $1/N$ in the full device
 participation setting, and with $1/K$ in the partial participation
@@
 in the choice of $E$ between the full and partial participation settings.
-With full participation there is linear speedup $\mathcal{O}(1/NT)$
-as long as $E=\mathcal{O}(\sqrt{T/N})$ since then $\mathcal{O}(E^{2}/T^{2})=\mathcal{O}(1/NT)$
+{\color{red}With full participation, the term involving $E$, $\mathcal{O}(E^{2}/T^{2})$, is not of leading order $\mathcal{O}(1/T)$, so we can increase $E$ and reduce the number of communication rounds without degrading the linear speedup in iteration complexity $\mathcal{O}(1/NT)$, as long as $E=\mathcal{O}(\sqrt{T/N})$, since then $\mathcal{O}(E^{2}/T^{2})=\mathcal{O}(1/NT)$
 matches the leading term. This corresponds to a communication complexity
-of $T/E=\mathcal{O}(\sqrt{NT})$. In contrast, the bound in \cite{li2019convergence}
+of $T/E=\mathcal{O}(\sqrt{NT})$.} In contrast, the bound in \cite{li2019convergence}
 does not allow $E$ to scale with $\sqrt{T}$ to preserve $\mathcal{O}(1/T)$
@@
 still yields a linear speedup in $K$, which is also confirmed by
-experiments. The requirement $E=\mathcal{O}(1)$ in partial participation cannot be removed for our sampling schemes, as the sampling
-variance is $\Omega(E^{2}/T^{2})$ and the dependence on $E$ is tight (see Proposition~\ref{prop:tight} in Section~\ref{sec:app:fedavg} of the appendix).  \\
+experiments.
+{\color{red}The requirement that $E=\mathcal{O}(1)$ in order to achieve linear speedup in partial participation cannot be removed for our sampling schemes, as the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$ comes from variance in the sampling process, which is $\mathcal{O}(E^{2}/T^{2})$.
+In Proposition~\ref{prop:tight} in Section~\ref{sec:app:fedavg} of the appendix, we provide a problem instance where the dependence of the sampling variance on $E$ is tight.}  \\
 \textbf{Comparison with related works.} To better understand the significance of the obtained bound, we compare our rates to the best-known results in related settings.
@@
 close to full participation, whereas our result has linear speedup for any number of participating devices.
+{\color{red}When there is no data heterogeneity, i.e.
+in the classical distributed optimization paradigm, communication complexity can be further improved, e.g.
+\cite{woodworth2020local,woodworth2020minibatch}, but such results are not directly comparable to ours since we consider the setting where individual devices have access to different datasets.}
 
@@
 \begin{align*}
-	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq \mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
 	\end{align*}
@@
 be $O(1)$ in order to achieve linear speedup of $\mathcal{O}(1/\sqrt{KT})$.
-Our result again demonstrates the difference in communication complexities
+{\color{red}This is again due to the fact that the sampling variance $\mathbb{E}\|\ov{w}_t-\ov{v}_t\|^2=\mathcal{O}(\alpha_t^2 E^2G^2)$ cannot be made independent of $E$, as illustrated by Proposition \ref{prop:tight}.
+See also the proof in Section \ref{sec:app:fedavg} for how the sampling variance and the term $E^2G^2/\sqrt{KT}$ are related.}  Our result again demonstrates the difference in communication complexities
 between full and partial participation, and is to our knowledge the

### 2021-02-06 source-iclrw/app_assumptions.tex
Shrink Sec 2

--- 2020-09-25 source-iclr/sec_problem.tex
@@
 
-\section{Setup}
-
-% \subsection{The Federated Learning Problem}
-
-In this paper, we study the following federated learning problem:
-\begin{align}
-	\min _{\mathbf{w}}\left\{F(\mathbf{w}) \triangleq \sum\nolimits_{k=1}^{N} p_{k} F_{k}(\mathbf{w})\right\},
-	\label{eq:problem}
-\end{align}
-where $N$ is the number of local devices (users/nodes/workers) and $p_k$ is the $k$-th device's weight satisfying $p_k \geq 0$ and $\sum_{k=1}^N p_k = 1$. 
-In the $k$-th local device, there are $n_k$ data points:
-$\vx_k^1, \vx_k^2, \dots, \vx_k^{n_k}$.  
-The local objective $F_k(\cdot)$ is defined as:
-% \begin{align}
-$F_{k}(\mathbf{w}) \triangleq \frac{1}{n_{k}} \sum_{j=1}^{n_{k}} \ell\left(\mathbf{w} ; \vx_k^j\right)$,
-% \label{eq:localloss}
-% \end{align}
-where $\ell$ denotes a user-specified loss function. Each device only has access to its local data, which gives rise to its own local objective $F_k$. Note that we do not make any assumptions on the
-data distributions of each local device. The local minimum
-$F^{*}_k= \min_{\vw} F_k(\vw)$ can be far from
-the global minimum of \eq{\ref{eq:problem}} (data heterogeneity).
-
-% A key difference with many problems in the distributed optimization setting is that data on local devices is heterogeneous, i.e.
-$F_k(w)$ may have different minimizers $w^{\ast}_k$.
-Moreover, each local device only has access to stochastic gradients based on data stored on the local device.
-In this paper, we investigate local objectives that are convex or strongly convex and smooth, although convergence results for non-smooth objectives are also given in the appendix.
-
-\subsection{The Federated Averaging (FedAvg) Algorithm}
-We first introduce the standard Federated Averaging (FedAvg)  algorithm which was first proposed by~\cite{mcmahan2016communication}.
-FedAvg updates the model in each device by local Stochastic Gradient Descent (SGD) and sends the latest model to the central server every $E$
-steps. The central server conducts a weighted average over the model parameters
-received from active devices and broadcasts the latest averaged model to all devices.
-Formally, the updates of FedAvg at round $t$ is described as follows:
-\begin{align}
-\label{eq:fedavg updates}
-\vv_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}\vg_{t,k}, \hspace{1em}
-\mathbf{w}_{t+1}^{k} =\left\{
-\begin{array}{ll}
-\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\ 
-\sum_{k \in \cS_{t+1}} q_k \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E},
-\end{array}\right.
-\end{align}
-where $\vw_t^k$ is the local model parameter maintained in the $k$-th device at the $t$-th iteration and $\vg_{t,k}:=\nabla F_{k}(\vw_{t}^{k},\xi_{t}^{k})$ is the stochastic gradient based on $\xi_{t}^{k}$, the data point sampled from $k$-th device’s local data uniformly at random.
-$\mathcal{I}_{E}=\{E,2E,\dots\}$ is the set of global communication steps, when local parameters from a set of active devices are averaged and broadcast to all devices.
-We use $\mathcal{S}_{t+1}$ to represent the (random) set of active devices at $t+1$.
-$q_k$ is a set of averaging weights that are specific to the sampling procedure used to obtain the set of active devices $\mathcal{S}_{t+1}$.
-
-Since federated learning usually involves an enormous amount of 
-local devices, it is often more realistic to assume only a subset of 
-local devices is active at each communication round (system heterogeneity). In this work,
-we consider both the case of \textbf{full participation} where the model is
-averaged over all devices at each communication round, in which case $q_k=p_k$ for all $k$ and 
-$\mathbf{w}_{t+1}^{k} = \sum_{k=1}^N p_k \mathbf{v}_{t+1}^{k}$ if $t+1 \in \mathcal{I}_{E}$, and
-the case of \textbf{partial participation} where $|\cS_{t+1}| < N$.
-
-With partial participation, we follow~\cite{li2018federated,karimireddy2019scaffold,li2019convergence} and assume that $\cS_{t+1}$ is obtained by one of two types of
-sampling schemes to simulate practical scenarios.
-One scheme establishes $\cS_{t+1}$ by \emph{i.i.d.} sampling the devices with probability $p_k$ with replacement, and uses $q_k=\frac{1}{K}$, where $K=|\mathcal{S}_{t+1}|$, while the other scheme samples $\cS_{t+1}$ uniformly \emph{i.i.d.} from all devices without replacement, and uses $q_k=p_k\frac{N}{K}$.
-Both schemes
-guarantee that gradient updates in FedAvg are unbiased stochastic versions of
-updates in FedAvg with full participation, which is important in the theoretical analysis of convergence.
-Because the original sampling scheme and weights proposed by \cite{mcmahan2016communication} lacks this nice property, it is not considered in this paper.
-For
-more details on the notations and setup as well as properties of the two sampling schemes, please refer to Section~\ref{sec:app:notations} in the appendix.
-
-% Thus when $t+1 \notin \mathcal{I}_{E}$, each device updates its local parameter based on local
-% stochastic gradient descent, and every $E$ iterations, parameters are aggregated
-% across devices based on weights $p_{k}$ and then broadcast back to
-% each local device.
-
-% Communication and broadcasting of local parameters
-% is costly, so the efficiency of the FedAvg algorithm requires that
-% $E$ is relatively large, while at the same time ensuring convergence
-% to a global minimizer of the global objective $\sum_{k}p_{k}F_{k}(w)$.
-
-% In practical applications, it may not be possible to collect local
-% parameters from \emph{all }devices at a specified communication time.
-% Thus the above \emph{full participation }FedAvg algorithm may not
-% be feasible. Instead, a more practical algorithm is to allow only
-% a random subset $\mathcal{S}_{t+1}$ of all devices to participate in parameter aggregation, while
-% broadcasting the aggregated parameter to all devices. This gives rise
-% to the \emph{partial participation }FedAvg algorithm.
-
-% The updates of FedAvg with partial device activation is given by: 
-% \begin{align} 
-% \mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\
-% \sum_{k \in \cS_{t+1}} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}\end{array}\right.
-% \end{align}
-
-% where $\mathcal{S}_{t+1}$ is obtained using some sampling scheme.
-We consider two types of common unbiased sampling strategies.
-The first one samples each device with probability $p_k$, with replacement, while the second one samples each device uniformly without replacement.
-Both of these sampling schemes guarantee that updates in FedAvg with partial participation are unbiased stochastic versions of updates in FedAvg with full participation, during the communication round.
-
-\subsection{Assumptions}
+\section{Assumptions}
 \label{sec:assumptions}
@@
 bounded gradient diversity restricts local objectives to have the same minimizer as the global objective, contradicting the heterogeneous data setting. 
-For detailed discussions of our assumptions, please refer to Appendix Section~\ref{sec:app:comparison}.
-
-% On the other hand, some works such as \cite{haddadpour2019convergence,li2018federated}
-% assume bounds on the ``gradient diversity'' or ``dissimilarity'',
-% defined as $\frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$.
-% Such assumptions are more restrictive than our assumptions, because
-% they effectively require $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$ for all local
-% objectives. For such ``overparameterized'' settings, we will prove
-% that the convergence is in fact geometric.
+For detailed discussions of our assumptions, please also refer to Appendix Section~\ref{sec:app:comparison}.

### 2021-02-06 source-iclrw/app_sumarization.tex
Shrink Sec 2

@@
 
-\section{Comparison of Convergence Rates with Related Works}
+\section{Connections with Distributed and Decentralized Optimization and Comparison of Convergence Rates with Related Works}
 \label{sec:app:comparison}
-In this section, we compare our convergence rate with the best-known results in the literature (see Table~\ref{tb:convergenceratev3}). 
+Federated learning is closely related to distributed and decentralized optimization, and as such it is important to discuss connections and distinctions between our work and related results from that literature.
+First, when there is neither system heterogeneity, i.e.
+all devices participate in parameter averaging during a communication round, nor statistical heterogeneity, i.e.
+all devices have access to a common set of stochastic gradients, FedAvg coincides with the ``Local SGD'' of~\cite{stich2018local}, which showed the linear speedup rate $\mathcal{O}(1/NT)$ for strongly convex and smooth functions.
+\cite{woodworth2020local} and \cite{woodworth2020minibatch} further improved the communication complexity that guarantees the linear speedup rate.
+When there is only data heterogeneity, some works have continued to use the term Local SGD to refer to FedAvg, while others subsume it in more general frameworks that include decentralized model averaging based on a network topology or a mixing matrix.
+They have provided linear speedup analyses for strongly convex and convex problems, e.g.~\cite{khaled2020tighter,koloskova2020unified} as well as non-convex problems, e.g.~\cite{jiang2018linear,yu2019parallel,wang2018cooperative}.
+However, these results do not consider system heterogeneity, i.e.
+the presence of stragglers in the device network.
+Even with decentralized model averaging, the assumptions usually imply that model averages over all devices is the same as decentralized model averages based on network topology (e.g.~\cite{koloskova2020unified} Proposition 1), which precludes system heterogeneity as defined in this paper and prevalent in FL problems.
+For momentum accelerated FedAvg,~\cite{yu2019linear} provided linear speedup analysis for non-convex problems, while results for strongly convex and convex settings are entirely lacking, even without system heterogeneity.~\cite{karimireddy2019scaffold} considers both types of heterogeneities for FedAvg, but their rate implies a linear speedup only when the number of stragglers is negligible.
+In contrast, our linear speedup analyses consider both types of heterogeneity present in the full federated learning setting, and are valid for any number of participating devices.
+We also highlight a distinction in communication efficiency when system heterogeneity is present.
+Moreover, our results for Nesterov accelerated FedAvg completes the picture for strongly convex and convex problems.
+For a detailed comparison with best-known results in the literature, please refer to Table~\ref{tb:convergenceratev3}.
+
+% In this section, we compare our convergence rate with the best-known results in the literature (see Table~\ref{tb:convergenceratev3}). 
 % In Table~\ref{tb:convergenceratev1} and Table~\ref{tb:convergenceratev2}, we summarize 

### 2021-02-06 source-iclrw/sec_SGD.tex
Shrink Sec 2

@@
 be $O(1)$ in order to achieve linear speedup of $\mathcal{O}(1/\sqrt{KT})$.
-{\color{red}This is again due to the fact that the sampling variance $\mathbb{E}\|\ov{w}_t-\ov{v}_t\|^2=\mathcal{O}(\alpha_t^2 E^2G^2)$ cannot be made independent of $E$, as illustrated by Proposition \ref{prop:tight}.
-See also the proof in Section \ref{sec:app:fedavg} for how the sampling variance and the term $E^2G^2/\sqrt{KT}$ are related.}  Our result again demonstrates the difference in communication complexities
+This is again due to the fact that the sampling variance $\mathbb{E}\|\ov{w}_t-\ov{v}_t\|^2=\mathcal{O}(\alpha_t^2 E^2G^2)$ cannot be made independent of $E$, as illustrated by Proposition \ref{prop:tight}.
+See also the proof in Section \ref{sec:app:fedavg} for how the sampling variance and the term $E^2G^2/\sqrt{KT}$ are related.
+Our result again demonstrates the difference in communication complexities
 between full and partial participation, and is to our knowledge the

### 2021-02-06 source-iclrw/app_fedavgalg.tex
4 pages v1 todo: sec 4, 5 transition

--- 2020-09-25 source-iclr/sec_problem.tex
@@
 
-\section{Setup}
-
-% \subsection{The Federated Learning Problem}
-
-In this paper, we study the following federated learning problem:
-\begin{align}
-	\min _{\mathbf{w}}\left\{F(\mathbf{w}) \triangleq \sum\nolimits_{k=1}^{N} p_{k} F_{k}(\mathbf{w})\right\},
-	\label{eq:problem}
-\end{align}
-where $N$ is the number of local devices (users/nodes/workers) and $p_k$ is the $k$-th device's weight satisfying $p_k \geq 0$ and $\sum_{k=1}^N p_k = 1$. 
-In the $k$-th local device, there are $n_k$ data points:
-$\vx_k^1, \vx_k^2, \dots, \vx_k^{n_k}$.  
-The local objective $F_k(\cdot)$ is defined as:
-% \begin{align}
-$F_{k}(\mathbf{w}) \triangleq \frac{1}{n_{k}} \sum_{j=1}^{n_{k}} \ell\left(\mathbf{w} ; \vx_k^j\right)$,
-% \label{eq:localloss}
-% \end{align}
-where $\ell$ denotes a user-specified loss function. Each device only has access to its local data, which gives rise to its own local objective $F_k$. Note that we do not make any assumptions on the
-data distributions of each local device. The local minimum
-$F^{*}_k= \min_{\vw} F_k(\vw)$ can be far from
-the global minimum of \eq{\ref{eq:problem}} (data heterogeneity).
-
-% A key difference with many problems in the distributed optimization setting is that data on local devices is heterogeneous, i.e.
-$F_k(w)$ may have different minimizers $w^{\ast}_k$.
-Moreover, each local device only has access to stochastic gradients based on data stored on the local device.
-In this paper, we investigate local objectives that are convex or strongly convex and smooth, although convergence results for non-smooth objectives are also given in the appendix.
-
-\subsection{The Federated Averaging (FedAvg) Algorithm}
-We first introduce the standard Federated Averaging (FedAvg)  algorithm which was first proposed by~\cite{mcmahan2016communication}.
+\section{The Federated Averaging (FedAvg) Algorithm}
+We introduce the standard Federated Averaging (FedAvg) algorithm which was first proposed by~\cite{mcmahan2016communication}.
 FedAvg updates the model in each device by local Stochastic Gradient Descent (SGD) and sends the latest model to the central server every $E$
@@
 more details on the notations and setup as well as properties of the two sampling schemes, please refer to Section~\ref{sec:app:notations} in the appendix.
-
-% Thus when $t+1 \notin \mathcal{I}_{E}$, each device updates its local parameter based on local
-% stochastic gradient descent, and every $E$ iterations, parameters are aggregated
-% across devices based on weights $p_{k}$ and then broadcast back to
-% each local device.
-
-% Communication and broadcasting of local parameters
-% is costly, so the efficiency of the FedAvg algorithm requires that
-% $E$ is relatively large, while at the same time ensuring convergence
-% to a global minimizer of the global objective $\sum_{k}p_{k}F_{k}(w)$.
-
-% In practical applications, it may not be possible to collect local
-% parameters from \emph{all }devices at a specified communication time.
-% Thus the above \emph{full participation }FedAvg algorithm may not
-% be feasible. Instead, a more practical algorithm is to allow only
-% a random subset $\mathcal{S}_{t+1}$ of all devices to participate in parameter aggregation, while
-% broadcasting the aggregated parameter to all devices. This gives rise
-% to the \emph{partial participation }FedAvg algorithm.
-
-% The updates of FedAvg with partial device activation is given by: 
-% \begin{align} 
-% \mathbf{v}_{t+1}^{k} &=\mathbf{w}_{t}^{k}-\eta_{t} \nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right) \\ \mathbf{w}_{t+1}^{k} &=\left\{\begin{array}{ll}\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\
-% \sum_{k \in \cS_{t+1}} \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E}\end{array}\right.
-% \end{align}
-
-% where $\mathcal{S}_{t+1}$ is obtained using some sampling scheme.
-We consider two types of common unbiased sampling strategies.
-The first one samples each device with probability $p_k$, with replacement, while the second one samples each device uniformly without replacement.
-Both of these sampling schemes guarantee that updates in FedAvg with partial participation are unbiased stochastic versions of updates in FedAvg with full participation, during the communication round.
-
-\subsection{Assumptions}
-\label{sec:assumptions}
-We make the following standard assumptions on the objective function $F_1,\dots, F_N$.
-Assumptions~\ref{ass:lsmooth} and~\ref{ass:stroncvx} are commonly satisfied by a range of popular objective functions, such as $\ell^{2}$-regularized logistic regression and cross-entropy loss functions.
-
- %Assumption~\ref{ass:lsmooth} is made to achieve linear speedup of convergence rate with respect to the number of workers, though it is not necessary for FedAvg and its variants to convergence.
-
-\begin{assumption}[$L$-smooth]
-	$F_{1}, \cdots, F_{N}$ are all $L$-smooth: for all  $\mathbf{v}$  and $\mathbf{w}$, $F_{k}(\mathbf{v}) \leq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{L}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$.
-	\label{ass:lsmooth}
-\end{assumption}
-\begin{assumption}[Strongly-convex]
-	$	F_{1}, \cdots, F_{N} \text { are all } \mu \text { -strongly convex: for all v and } \mathbf{w}, F_{k}(\mathbf{v}) \geq F_{k}(\mathbf{w})+(\mathbf{v}-\mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{\mu}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$
-	\label{ass:stroncvx}
-\end{assumption}
-\begin{assumption}[Bounded local variance]
-	Let $\xi_{t}^{k}$ be sampled from the $k$-th device's local data uniformly at random.
-The variance of stochastic gradients in each device is bounded: $\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)-\nabla F_{k}\left(\mathbf{w}_{t}^{k}\right)\right\|^{2} \leq \sigma_{k}^{2}$,
-	for $k=1, \cdots, N$ and any $\mathbf{w}_{t}^{k}$. Let $\sigma^2:=\sum_{k=1}^{N}p_k\sigma_{k}^{2}$.
-	\label{ass:boundedvariance}
-\end{assumption}
-\begin{assumption}[Bounded local gradient]
-	The expected squared norm of stochastic gradients is uniformly bounded. i.e.,
-	$\mathbb{E}\left\|\nabla F_{k}\left(\mathbf{w}_{t}^{k}, \xi_{t}^{k}\right)\right\|^{2} \leq G^{2}$, for all $k = 1,..., N$ and $t=0, \dots, T-1$.
-	\label{ass:subgrad2}
-\end{assumption}
-Assumptions~\ref{ass:boundedvariance} and~\ref{ass:subgrad2} have been made in many previous
-works in federated learning, e.g.~\cite{yu2019parallel,li2019convergence,stich2018local}. We provide further justification for their generality. As model average parameters become
-closer to $\mathbf{w}^{\ast}$, the $L$-smoothness property implies
-that $\mathbb{E}\|\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})\|^{2}$
-and $\mathbb{E}\|\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})-\nabla F_{k}(\mathbf{w}_{t}^{k})\|^{2}$
-approach $\mathbb{E}\|\nabla F_{k}(\mathbf{w}^{\ast},\xi_{t}^{k})\|^{2}$
-and $\mathbb{E}\|\nabla F_{k}(\mathbf{w}^{\ast},\xi_{t}^{k})-\nabla F_{k}(\mathbf{w}^{\ast})\|^{2}$.
-Therefore, there is no substantial difference between these assumptions
-and assuming the bounds at $\mathbf{w}^{\ast}$ only~\cite{koloskova2020unified}.
-Furthermore, compared to assuming \textit{bounded gradient diversity} as in related work~\cite{haddadpour2019convergence,li2018federated}, Assumption~\ref{ass:subgrad2} is much less restrictive.
-When the optimality gap converges to zero,
-bounded gradient diversity restricts local objectives to have the same minimizer as the global objective, contradicting the heterogeneous data setting. 
-For detailed discussions of our assumptions, please refer to Appendix Section~\ref{sec:app:comparison}.
-
-% On the other hand, some works such as \cite{haddadpour2019convergence,li2018federated}
-% assume bounds on the ``gradient diversity'' or ``dissimilarity'',
-% defined as $\frac{\sum_{k}p_{k}\|\nabla F_{k}(\mathbf{w})\|_{2}^{2}}{\|\sum_{k}p_{k}\nabla F_{k}(\mathbf{w})\|_{2}^{2}}$.
-% Such assumptions are more restrictive than our assumptions, because
-% they effectively require $\nabla F_{k}(\mathbf{w}^{\ast})\equiv0$ for all local
-% objectives. For such ``overparameterized'' settings, we will prove
-% that the convergence is in fact geometric.

### 2021-02-06 source-iclrw/sec_SGD.tex
4 pages v1 todo: sec 4, 5 transition

@@
 In this section, we provide convergence analyses of FedAvg for convex objectives in the general setting with both heterogeneous data (statistical heterogeneity) and partial
-participation (system heterogeneity). We show that for strongly convex and smooth objectives,
-the convergence of the optimality gap of averaged parameters across
-devices is $\mathcal{O}(1/KT)$, while for convex and smooth
-objectives, the rate is $\mathcal{O}(1/\sqrt{KT})$.
+participation (system heterogeneity). 
+% We show that for strongly convex and smooth objectives,
+% the convergence of the optimality gap of averaged parameters across
+% devices is $\mathcal{O}(1/KT)$, while for convex and smooth
+% objectives, the rate is $\mathcal{O}(1/\sqrt{KT})$.
 Our results improve upon~\cite{li2019convergence,karimireddy2019scaffold} by showing linear speedup for any number of participating devices, and upon~\cite{khaled2020tighter,koloskova2020unified} by allowing system heterogeneity.
@@
 \end{theorem}
+\begin{comment}
 \textbf{Proof sketch.} Because our unified analyses of results in the main text follow the same framework with variations in technical details, we first give an outline of proof  for Theorem~\ref{th:scvx_sgd} to illustrate the main ideas.
@@
 To facilitate the understanding of our analysis, please refer to a high-level summary in Appendix~\ref{sec:app:sum}.
-
+\end{comment}
 \begin{comment}
@@
 in the choice of $E$ between the full and partial participation settings.
-{\color{red}With full participation, the term involving $E$, $\mathcal{O}(E^{2}/T^{2})$, is not of leading order $\mathcal{O}(1/T)$, so we can increase $E$ and reduce the number of communication rounds without degrading the linear speedup in iteration complexity $\mathcal{O}(1/NT)$, as long as $E=\mathcal{O}(\sqrt{T/N})$, since then $\mathcal{O}(E^{2}/T^{2})=\mathcal{O}(1/NT)$
+With full participation, the term involving $E$, $\mathcal{O}(E^{2}/T^{2})$, is not of leading order $\mathcal{O}(1/T)$, so we can increase $E$ and reduce the number of communication rounds without degrading the linear speedup in iteration complexity $\mathcal{O}(1/NT)$, as long as $E=\mathcal{O}(\sqrt{T/N})$, since then $\mathcal{O}(E^{2}/T^{2})=\mathcal{O}(1/NT)$
 matches the leading term. This corresponds to a communication complexity
-of $T/E=\mathcal{O}(\sqrt{NT})$.} In contrast, the bound in \cite{li2019convergence}
+of $T/E=\mathcal{O}(\sqrt{NT})$. In contrast, the bound in \cite{li2019convergence}
 does not allow $E$ to scale with $\sqrt{T}$ to preserve $\mathcal{O}(1/T)$
@@
 experiments.
-{\color{red}The requirement that $E=\mathcal{O}(1)$ in order to achieve linear speedup in partial participation cannot be removed for our sampling schemes, as the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$ comes from variance in the sampling process, which is $\mathcal{O}(E^{2}/T^{2})$.
-In Proposition~\ref{prop:tight} in Section~\ref{sec:app:fedavg} of the appendix, we provide a problem instance where the dependence of the sampling variance on $E$ is tight.}  \\
+The requirement that $E=\mathcal{O}(1)$ in order to achieve linear speedup in partial participation cannot be removed for our sampling schemes, as the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$ comes from variance in the sampling process, which is $\mathcal{O}(E^{2}/T^{2})$.
+In Proposition~\ref{prop:tight} in Section~\ref{sec:app:fedavg} of the appendix, we provide a problem instance where the dependence of the sampling variance on $E$ is tight.
+\\
 \textbf{Comparison with related works.} To better understand the significance of the obtained bound, we compare our rates to the best-known results in related settings.
@@
 close to full participation, whereas our result has linear speedup for any number of participating devices.
-{\color{red}When there is no data heterogeneity, i.e.
+When there is no data heterogeneity, i.e.
 in the classical distributed optimization paradigm, communication complexity can be further improved, e.g.
-\cite{woodworth2020local,woodworth2020minibatch}, but such results are not directly comparable to ours since we consider the setting where individual devices have access to different datasets.}
+\cite{woodworth2020local,woodworth2020minibatch}, but such results are not directly comparable to ours since we consider the setting where individual devices have access to different datasets.
 

### 2021-02-06 source-iclrw/sec_intro.tex
Shrink Sec 2
4 pages v1 todo: sec 4, 5 transition

--- 2020-11-22 source-iclr/sec_intro.tex
@@
 To address these challenges, Federated Averaging (FedAvg)~\cite{mcmahan2016communication} was proposed as a particularly effective heuristic, which has enjoyed great empirical success.
-This success has since motivated a growing line of research efforts into understanding its theoretical convergence guarantees in various settings.
-For instance, \cite{haddadpour2019convergence} analyzed FedAvg (for non-convex smooth problems satisfying PL conditions) under the assumption that each local device's minimizer is the same as the minimizer of the joint problem (if all devices' data is aggregated together), an overly restrictive assumption that restricts the extent of data heterogeneity.
-Very recently, \cite{li2019convergence} furthered the progress and established an $\cO(\frac{1}{T})$ convergence rate for FedAvg for strongly convex smooth Federated learning problems with both data and system heterogeneity.
+This success has since motivated a growing line of research efforts \cite{haddadpour2019convergence,li2019convergence,karimireddy2019scaffold,huo2020faster} into understanding its theoretical convergence guarantees in various settings.
+% For instance, \cite{haddadpour2019convergence} analyzed FedAvg (for non-convex smooth problems satisfying PL conditions) under the assumption that each local device's minimizer is the same as the minimizer of the joint problem (if all devices' data is aggregated together), an overly restrictive assumption that restricts the extent of data heterogeneity.
+% Very recently, \cite{li2019convergence} furthered the progress and established an $\cO(\frac{1}{T})$ convergence rate for FedAvg for strongly convex smooth Federated learning problems with both data and system heterogeneity.
 A similar result in the same setting~\cite{karimireddy2019scaffold} also established an $\cO(\frac{1}{T})$ result that allows for a linear speedup when the number of participating devices is large.
-At the same time, \cite{huo2020faster} studied the Nesterov accelerated FedAvg for non-convex smooth problems and established 
-an $\cO(\frac{1}{\sqrt{T}})$ convergence rate to stationary points. 
+% At the same time, \cite{huo2020faster} studied the Nesterov accelerated FedAvg for non-convex smooth problems and established 
+% an $\cO(\frac{1}{\sqrt{T}})$ convergence rate to stationary points. 
 % (Prior to that, \cite{liu2019accelerating} studied Nesterov accelerated FedAvg using full gradient--rather than stochastic gradient as in both the above-mentioned works and our setting--for strongly convex smooth problems and established an $\cO(\frac{1}{T})$ convergence rate).
-
 However, despite these very recent fruitful pioneering efforts into understanding the theoretical convergence properties of FedAvg, it remains open as to how the number of devices--particularly the number of devices that participate in the computation--affects the convergence speed.
@@
 We provide a comprehensive and unified convergence analysis
-of FedAvg and its accelerated variants in the presence of both data and system heterogeneity. 
+of FedAvg and its accelerated variants considering both data and system heterogeneity. 
 Our contributions are threefold.
-
 First, we establish an {\small{$\cO(1/KT)$}} convergence rate  under FedAvg for strongly convex and smooth problems and  an
 {\small{$\cO(1/\sqrt{KT})$}} convergence rate for convex
-and smooth problems (where $K$ is the number of participating devices), thereby establishing that FedAvg enjoys the desirable linear speedup property in the FL setup.
+and smooth problems (where $K$ is the number of participating devices), thereby establishing that FedAvg enjoys the desirable linear speedup convergence, which improves the best known results\cite{li2019convergence,karimireddy2019scaffold} .
+\begin{comment}
 Prior to our work here, the best and the most related convergence analysis is given by \cite{li2019convergence} and~\cite{karimireddy2019scaffold}, which established an $\cO(\frac{1}{T})$ convergence rate for strongly convex smooth problems under FedAvg.
@@
 Our unified analysis highlights the common elements and distinctions between the strongly and convex settings.
-
+\end{comment}
 Second, we establish the same convergence rates--{\small{$\cO(1/KT)$}} for strongly convex and smooth problems and {\small{$\cO(1/\sqrt{KT})$}} for convex and smooth problems--for Nesterov accelerated FedAvg.
+\begin{comment}
 We analyze the accelerated version of FedAvg here because empirically it tends to perform better; yet, its theoretical convergence guarantee is unknown.
@@
 Our unified analysis of Nesterov FedAvg also illustrates the technical similarities and distinctions compared to the original FedAvg algorithm, whereas prior works (in the non-convex setting) were scattered and used different notations.
-
+\end{comment}
 Third, we study a subclass of strongly convex smooth problems where the objective is over-parameterized and establish 
 a faster $\cO(\exp(-\frac{KT}{\kappa}))$ convergence rate for FedAvg, in contrast to the $\cO(\exp(-\frac{T}{\kappa}))$ rate for individual solvers~\cite{ma2017power}. 
+\begin{comment}
 Within this class, we further consider the linear regression problem and establish an even sharper rate under FedAvg.
@@
 The detailed convergence results are summarized in Table~\ref{tb:convergencerateintro}.
+\end{comment}
+
+\begin{comment}
 
@@
 For a detailed comparison with related works, please refer to Table~\ref{tb:convergenceratev3} in Appendix Section~\ref{sec:app:comparison}.
+
+\end{comment}
 

### 2021-02-06 source-iclrw/sec_problem.tex
Shrink Sec 2
4 pages v1 todo: sec 4, 5 transition

--- 2020-09-25 source-iclr/sec_problem.tex
@@
 
-\subsection{The Federated Averaging (FedAvg) Algorithm}
-We first introduce the standard Federated Averaging (FedAvg)  algorithm which was first proposed by~\cite{mcmahan2016communication}.
-FedAvg updates the model in each device by local Stochastic Gradient Descent (SGD) and sends the latest model to the central server every $E$
-steps. The central server conducts a weighted average over the model parameters
-received from active devices and broadcasts the latest averaged model to all devices.
-Formally, the updates of FedAvg at round $t$ is described as follows:
-\begin{align}
-\label{eq:fedavg updates}
-\vv_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}\vg_{t,k}, \hspace{1em}
-\mathbf{w}_{t+1}^{k} =\left\{
-\begin{array}{ll}
-\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\ 
-\sum_{k \in \cS_{t+1}} q_k \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E},
-\end{array}\right.
-\end{align}
-where $\vw_t^k$ is the local model parameter maintained in the $k$-th device at the $t$-th iteration and $\vg_{t,k}:=\nabla F_{k}(\vw_{t}^{k},\xi_{t}^{k})$ is the stochastic gradient based on $\xi_{t}^{k}$, the data point sampled from $k$-th device’s local data uniformly at random.
-$\mathcal{I}_{E}=\{E,2E,\dots\}$ is the set of global communication steps, when local parameters from a set of active devices are averaged and broadcast to all devices.
-We use $\mathcal{S}_{t+1}$ to represent the (random) set of active devices at $t+1$.
-$q_k$ is a set of averaging weights that are specific to the sampling procedure used to obtain the set of active devices $\mathcal{S}_{t+1}$.
-
-Since federated learning usually involves an enormous amount of 
-local devices, it is often more realistic to assume only a subset of 
-local devices is active at each communication round (system heterogeneity). In this work,
-we consider both the case of \textbf{full participation} where the model is
-averaged over all devices at each communication round, in which case $q_k=p_k$ for all $k$ and 
-$\mathbf{w}_{t+1}^{k} = \sum_{k=1}^N p_k \mathbf{v}_{t+1}^{k}$ if $t+1 \in \mathcal{I}_{E}$, and
-the case of \textbf{partial participation} where $|\cS_{t+1}| < N$.
-
-With partial participation, we follow~\cite{li2018federated,karimireddy2019scaffold,li2019convergence} and assume that $\cS_{t+1}$ is obtained by one of two types of
-sampling schemes to simulate practical scenarios.
-One scheme establishes $\cS_{t+1}$ by \emph{i.i.d.} sampling the devices with probability $p_k$ with replacement, and uses $q_k=\frac{1}{K}$, where $K=|\mathcal{S}_{t+1}|$, while the other scheme samples $\cS_{t+1}$ uniformly \emph{i.i.d.} from all devices without replacement, and uses $q_k=p_k\frac{N}{K}$.
-Both schemes
-guarantee that gradient updates in FedAvg are unbiased stochastic versions of
-updates in FedAvg with full participation, which is important in the theoretical analysis of convergence.
-Because the original sampling scheme and weights proposed by \cite{mcmahan2016communication} lacks this nice property, it is not considered in this paper.
-For
-more details on the notations and setup as well as properties of the two sampling schemes, please refer to Section~\ref{sec:app:notations} in the appendix.
-
 % Thus when $t+1 \notin \mathcal{I}_{E}$, each device updates its local parameter based on local
@@
 Both of these sampling schemes guarantee that updates in FedAvg with partial participation are unbiased stochastic versions of updates in FedAvg with full participation, during the communication round.
-
-\subsection{Assumptions}
+\begin{comment}
+{\subsection{Assumptions}
 \label{sec:assumptions}
@@
 bounded gradient diversity restricts local objectives to have the same minimizer as the global objective, contradicting the heterogeneous data setting. 
-For detailed discussions of our assumptions, please refer to Appendix Section~\ref{sec:app:comparison}.
+For detailed discussions of our assumptions, please refer to Appendix Section~\ref{sec:app:comparison}.}
+\end{comment}
 

### 2021-02-06 source-iclrw/app_nasgd_scvx_smooth.tex
4 pages v2

--- 2020-06-10 source/app_nasgd_scvx_smooth.tex
@@
+Recall that the Nesterov accelerated FedAvg follows the updates 
+\begin{align*}
+\mathbf{v}_{t+1}^{k} & =\mathbf{w}_{t}^{k}-\alpha_{t}\mathbf{g}_{t,k}, \hspace{1em}
+\mathbf{w}_{t+1}^{k} =\begin{cases}
+\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E},\\
+\sum_{k \in \cS_{t+1}}q_k\left[\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E}.
+\end{cases}
+\end{align*}
+
+The proofs of convergence results for Nesterov Accelerated FedAvg consists of components that are direct analogues of the FedAvg case.
+We first state these analogue results before proving the main theorem.
+Like before, the proofs of the lemmas are deferred to after the main proof.
+
+\begin{lemma}[\textbf{One step progress, Nesterov}] Let $\overline{\mathbf{v}}_{t}=\sum_{k=1}^{N}p_{k}\mathbf{v}_{t}^{k}$ in Nesterov accelerated FedAvg,
+and suppose our functions satisfy Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2}, and set step sizes $\alpha_{t}=\frac{6}{\mu}\frac{1}{t+\gamma}$,  $\beta_{t-1}=\frac{3}{14(t+\gamma)(1-\frac{6}{t+\gamma})\max\{\mu,1\}}$
+	with $\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$, the updates of Nesterov accelerated FedAvg satisfy
+\begin{align*}
+\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
+& +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|
+\end{align*}
+\label{lem:nest-scvxoner}
+\end{lemma}
+
+The one step progress result makes use of the same bound on the gradient variance in~Lemma~\ref{lem:bgv}, as well as a divergence bound analogous to Lemma~\ref{lem:bdw}, which we state below.
+\begin{lemma}[\textbf{Bounding the divergence of $\vw_t^k$, Nesterov}]
+Given Assumption~\ref{ass:subgrad2}, and assume that $\alpha_t$ is non-increasing, $\alpha_t \leq 2\alpha_{t+E}$, and $2\beta_{t-1}^{2}+2\alpha_{t}^{2}\leq1/2$ for all $t\geq 0$, $\overline{\mathbf{w}}_{t}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{t}^{k}$ in Nesterov accelerated FedAvg satisfies
+	\begin{align*}
+	\mathbb{E}\left[\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} \right]\leq16(E-1)^{2}\alpha_{t}^{2}G^{2}.
+	\end{align*}
+\label{lem:nest-bdw}
+\end{lemma}
+
 \begin{thm}
-	Let $\overline{\mathbf{v}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{v}_{T}^{k}$
+	Let $\overline{\mathbf{v}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{v}_{T}^{k}$ in Nesterov accelerated FedAvg
 	and set learning rates $\alpha_{t}=\frac{6}{\mu}\frac{1}{t+\gamma}$,  $\beta_{t-1}=\frac{3}{14(t+\gamma)(1-\frac{6}{t+\gamma})\max\{\mu,1\}}$.
@@
 %
-\textbf{}%
+
 \begin{comment}
@@
 \begin{proof}
-Define the virtual sequences $\ov{v}_{t}=\sum_{k=1}^{N}p_{k}\vv_{t}^{k}$,
-$\ov{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$, and $\ov{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}\vg_{t,k}$.
-We have $\mathbb{E}\vg_{t}=\ov{g}_{t}$ and $\ov{v}_{t+1}=\ov{w}_{t}-\alpha_{t}\vg_{t}$,
-and $\ov{w}_{t+1}=\ov{v}_{t+1}$ for all $t$. The proof
-again uses the $L$-smoothness of $F$ to bound 
+We first prove the result for full participation. Applying the one step progress bound in Lemma~\ref{lem:nest-scvxoner}, we have
 \begin{align*}
-\mathbb{E}(F(\ov{v}_{t}))-F^{\ast} & =\mathbb{E}(F(\ov{v}_{t})-F(\vw^{\ast}))\\
-& \leq\frac{L}{2}\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}
+\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
+& +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|
 \end{align*}
-Our main step is to prove the bound
+Recall that we require $\alpha_{t_{0}}\leq2\alpha_{t}$ for any
+$t-t_{0}\leq E-1$, $L\alpha_{t}\leq\frac{1}{5}$, and $2\beta_{t-1}^{2}+2\alpha_{t}^{2}\leq1/2$ in order for Lemmas~\ref{lem:nest-bdw} and~\ref{lem:nest-scvxoner} to hold,
+which we can check by definition of $\alpha_{t}$ and
+$\beta_{t}$.\\
+ We show next that $\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}=\mathcal{O}(\frac{\nu_{max}^{2}\sigma^{2}/\mu}{tN}+\frac{E^{2}LG^{2}/\mu^2}{t^{2}})$ 
+by induction. Assume that we have shown 
 \begin{align*}
-\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+20E^{2}L\alpha_{t}^{3}G^{2}
+\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2} & \leq b(C\alpha_{t}^{2}+D\alpha_{t})
 \end{align*}
- for appropriate step sizes $\alpha_{t},\beta_{t}$.
+for all iterations until $t$, where $C=20E^{2}LG^{2}$, $D=\frac{1}{N}\nu_{max}^{2}\sigma^{2}$,
+and $b$ is some constant to be chosen later. For step sizes recall that we choose $\alpha_{t}=\frac{6}{\mu}\frac{1}{t+\gamma}$
+and $\beta_{t-1}=\frac{3}{14(t+\gamma)(1-\frac{6}{t+\gamma})\max\{\mu,1\}}$
+where $\gamma=\max\{32\kappa,E\}$, so that $\beta_{t-1}\leq\alpha_{t}$
+and 
+\begin{align*}
+(1-\mu\alpha_{t})(1+14\beta_{t-1}) & \leq(1-\frac{6}{t+\gamma})(1+\frac{3}{(t+\gamma)(1-\frac{6}{t+\gamma})})\\
+& =1-\frac{6}{t+\gamma}+\frac{3}{t+\gamma}=1-\frac{3}{t+\gamma}=1-\frac{\mu\alpha_{t}}{2}
+\end{align*}
 
-We have 
+Moreover, $\mathbb{E}\|\ov{v}_{t-1}-\vw^{\ast}\|^{2}\leq b(C\alpha_{t-1}^{2}+D\alpha_{t-1})\leq 4b(C\alpha_{t}^{2}+D\alpha_{t})$
+with the chosen step sizes.
+Therefore the bound for $\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}$
+can be further simplified with 
 \begin{align*}
-\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}\vg_{t})-\vw^{\ast}\|^{2}\\
-& =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(\vg_{t}-\ov{g}_{t})\|^{2}\\
-& =A_{1}+A_{2}+A_{3}
+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\| & \leq4\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\cdot b(C\alpha_{t}^{2}+D\alpha_{t})
 \end{align*}
-where 
+and 
 \begin{align*}
-A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}\\
-A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-\vg_{t}\rangle\\
-A_{3} & =\alpha_{t}^{2}\|\vg_{t}-\ov{g}_{t}\|^{2}
-\end{align*}
-$\mathbb{E}A_{2}=0$ by definition of $\vg_{t}$ and $\ov{g}_{t}$,
-while for $A_{3}$ we have
-\begin{align*}
-\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|\vg_{t,k}-\mathbb{E}\vg_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
-\end{align*}
-again by Jensen's inequality and using the independence of $\vg_{t,k},g_{t,k'}$.
-
-Next we bound $A_{1}$: 
-\begin{align*}
-\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2}
-\end{align*}
-Same as the SGD case, 
-\begin{align*}
--2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2} & \leq\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\mathbb{E}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2} & \leq4(1-\alpha_{t}\mu)\beta_{t-1}^{2}\cdot b(C\alpha_{t}^{2}+D\alpha_{t})
 \end{align*}
@@
 \begin{align*}
-\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} & \leq(1-\alpha_{t}\mu)\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}
+\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})((1+\beta_{t-1})^{2}+4\beta_{t-1}(1+\beta_{t-1})+4\beta_{t-1}^{2})\cdot b(C\alpha_{t}^{2}+D\alpha_{t})\\
+& +20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
+& \leq\mathbb{E}(1-\mu\alpha_{t})(1+14\beta_{t-1})\cdot b(C\alpha_{t}^{2}+D\alpha_{t})+20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
+& \leq b(1-\frac{\mu\alpha_{t}}{2})(C\alpha_{t}^{2}+D\alpha_{t})+C\alpha_{t}^{3}+D\alpha_{t}^{2}\\
+& =(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}^{2}C+(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}D
+\end{align*}
+and so it remains to choose $b$ such that 
+\begin{align*}
+(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t} & \leq b\alpha_{t+1}\\
+(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}^{2} & \leq b\alpha_{t+1}^{2}
+\end{align*}
+from which we can conclude $\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}\leq\alpha_{t+1}^{2}C+\alpha_{t+1}D$.
+
+With $b=\frac{6}{\mu}$, we have
+\begin{align*}
+(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t} & =(b(1-(\frac{3}{t+\gamma})+\frac{6}{\mu(t+\gamma)})\frac{6}{\mu(t+\gamma)}\\
+& =(b\frac{t+\gamma-3}{t+\gamma}+\frac{6}{\mu(t+\gamma)})\frac{6}{\mu(t+\gamma)}\\
+& \leq b(\frac{t+\gamma-1}{t+\gamma})\frac{6}{\mu(t+\gamma)}\\
+& \leq b\frac{6}{\mu(t+\gamma+1)}=b\alpha_{t+1}
+\end{align*}
+where we have used $\frac{t+\gamma-1}{(t+\gamma)^{2}}\leq\frac{1}{t+\gamma+1}$.
+
+Similarly 
+\begin{align*}
+(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}^{2} & =(b(1-(\frac{3}{t+\gamma})+\frac{6}{\mu(t+\gamma)})(\frac{6}{\mu(t+\gamma)})^{2}\\
+& =(b\frac{t+\gamma-3}{t+\gamma}+\frac{6}{\mu(t+\gamma)})(\frac{6}{\mu(t+\gamma)})^{2}\\
+& =b(\frac{t+\gamma-2}{t+\gamma})(\frac{6}{\mu(t+\gamma)})^{2}\\
+& \leq b\frac{36}{\mu^{2}(t+\gamma+1)^{2}}=b\alpha_{t+1}^{2}
+\end{align*}
+where we have used $\frac{t+\gamma-2}{(t+\gamma)^{3}}\leq\frac{1}{(t+\gamma+1)^{2}}$.
+
+Finally, to ensure $\|\vv_{0}-\vw^{\ast}\|^{2}\leq b(C\alpha_{0}^{2}+D\alpha_{0})$,
+we can rescale $b$ by $c\|\vv_{0}-\vw^{\ast}\|^{2}$ for some $c.$ It
+follows that $\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}\leq b(C\alpha_{t}^{2}+D\alpha_{t})$
+for all $t\geq0$. Using the $L$-smooothness of $F$,
+\begin{align*}
+\mathbb{E}(F(\ov{v}_{T}))-F^{\ast} & =\mathbb{E}(F(\ov{v}_{T})-F(\vw^{\ast}))\\
+& \leq\frac{L}{2}\mathbb{E}\|\ov{v}_{T}-\vw^{\ast}\|^{2}\leq\frac{L}{2}c\|\vv_{0}-\vw^{\ast}\|^{2}\frac{6}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
+& =3c\|\vv_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
+& \leq3c\|\vv_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{6}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{\max}\sigma^{2}+20E^{2}LG^{2}\cdot(\frac{6}{\mu(T+\gamma)})^{2}\right]\\
+& =\mathcal{O}(\frac{\kappa}{\mu}\frac{1}{N}\nu_{\max}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
 \end{align*}
 
-Different from the SGD case, we have 
+With partial participation, the same argument with an added term for sampling error yields
 \begin{align*}
-\|\ov{w}_{t}-\vw^{\ast}\|^{2} & =\|\ov{v}_{t}+\beta_{t-1}(\ov{v}_{t}-\ov{v}_{t-1})-\vw^{\ast}\|^{2}\\
-& =\|(1+\beta_{t-1})(\ov{v}_{t}-\vw^{\ast})-\beta_{t-1}(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
-& =(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}-2\beta_{t-1}(1+\beta_{t-1})\langle\ov{v}_{t}-\vw^{\ast},\ov{v}_{t-1}-\vw^{\ast}\rangle+\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
-& \leq(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+2\beta_{t-1}(1+\beta_{t-1})\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|+\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}
+\mathbb{E}F(\ov{w}_{T})-F^{\ast}=\mathcal{O}(\frac{\kappa\nu_{\max}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
 \end{align*}
-which gives 
-\begin{align*}
-\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\alpha_{t}\mu)(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+2(1-\alpha_{t}\mu)\beta_{t-1}(1+\beta_{t-1})\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-& +\beta_{t-1}^{2}(1-\alpha_{t}\mu)\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}LG^{2}
-\end{align*}
-and we will using this recursive relation to obtain the desired bound.
+\end{proof}
 
-First we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$.
+\subsubsection{Deferred proofs of key lemmas}
+
+\begin{proof}[Proof of lemma~\ref{lem:nest-bdw}]
+	The proof of bound for $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$ in the Nesterov accelerated FedAvg follows a similar logic as in Lemma~\ref{lem:bdw}, but requires extra reasoning.
 Since communication is done every $E$ steps, for any $t\geq0$, we
@@
 for all $t$. 
+\end{proof}
 
-Using the bound on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$,
-we can conclude that, with $\nu_{\max}:=N\cdot\max_{k}p_{k}$,
+Now we are ready to prove the one step progress result for Nesterov accelerated FedAvg.
+The first part of the proof is identical to that of the FedAvg case, while the main recursion takes a different form.
+\begin{proof}[Proof of lemma~\ref{lem:nest-scvxoner}]
+We again have 
+\begin{align*}
+\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}\vg_{t})-\vw^{\ast}\|^{2}
+\end{align*}
+and using exactly the same derivation as the FedAvg case, we can obtain the following bound (same as \eq{\ref{eq:common recursion}} in the proof of Lemma~\ref{lem:scvxoner}):
+\begin{align*}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
+	\end{align*}
 
+Different from the FedAvg case, we no longer have $\ov{w}_{t}=\ov{v}_{t}$. Instead,
 \begin{align*}
-\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+16E^{2}L\alpha_{t}^{3}G^{2}+16E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{3}LG^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
-& +\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|\\
+\|\ov{w}_{t}-\vw^{\ast}\|^{2} & =\|\ov{v}_{t}+\beta_{t-1}(\ov{v}_{t}-\ov{v}_{t-1})-\vw^{\ast}\|^{2}\\
+& =\|(1+\beta_{t-1})(\ov{v}_{t}-\vw^{\ast})-\beta_{t-1}(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
+& =(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}-2\beta_{t-1}(1+\beta_{t-1})\langle\ov{v}_{t}-\vw^{\ast},\ov{v}_{t-1}-\vw^{\ast}\rangle+\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
+& \leq(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+2\beta_{t-1}(1+\beta_{t-1})\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|+\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}
+\end{align*}
+which gives a recursion involving both $\ov{v}_{t}$ and $\ov{v}_{t-1}$:
+\begin{align*}
+\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\alpha_{t}\mu)(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+2(1-\alpha_{t}\mu)\beta_{t-1}(1+\beta_{t-1})\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+& +\beta_{t-1}^{2}(1-\alpha_{t}\mu)\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}LG^{2}
+\end{align*}
+and we will using this recursive relation to obtain the desired bound.
+
+We can check that our choice of $\alpha_t$ and $\beta_t$ satisfy $\alpha_t$ is non-increasing, $\alpha_t \leq 2\alpha_{t+E}$, and $2\beta_{t-1}^{2}+2\alpha_{t}^{2}\leq1/2$ for all $t\geq 0$, so that we can apply the bound from Lemma~\ref{lem:nest-bdw} on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$ to conclude that, with $\nu_{\max}:=N\cdot\max_{k}p_{k}$,
+\begin{align*}
+\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+16E^{2}L\alpha_{t}^{3}G^{2}+16E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{3}LG^{2}\\
+& +(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2} +\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|\\
 & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
@@
 \end{align*}
-where $\sigma^{2}=\sum_{k}p_{k}\sigma_{k}^{2}$, and $\alpha_{t}$
-satisfies $L\alpha_{t}\leq\frac{1}{5}$. We show next that $\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}}{t^{2}})$
-by induction.
-
-Assume that we have shown 
-\begin{align*}
-\mathbb{E}\|\overline{y}_{t}-\vw^{\ast}\|^{2} & \leq b(C\alpha_{t}^{2}+D\alpha_{t})
-\end{align*}
-for all iterations until $t$, where $C=20E^{2}LG^{2}$, $D=\frac{1}{N}\nu_{max}^{2}\sigma^{2}$,
-and $b$ is to be chosen later. For step sizes we choose $\alpha_{t}=\frac{6}{\mu}\frac{1}{t+\gamma}$
-and $\beta_{t-1}=\frac{3}{14(t+\gamma)(1-\frac{6}{t+\gamma})\max\{\mu,1\}}$
-where $\gamma=\max\{32\kappa,E\}$, so that $\beta_{t-1}\leq\alpha_{t}$
-and 
-\begin{align*}
-(1-\mu\alpha_{t})(1+14\beta_{t-1}) & \leq(1-\frac{6}{t+\gamma})(1+\frac{3}{(t+\gamma)(1-\frac{6}{t+\gamma})})\\
-& =1-\frac{6}{t+\gamma}+\frac{3}{t+\gamma}=1-\frac{3}{t+\gamma}=1-\frac{\mu\alpha_{t}}{2}
-\end{align*}
-Recall that we also require $\alpha_{t_{0}}\leq2\alpha_{t}$ for any
-$t-t_{0}\leq E-1$, $L\alpha_{t}\leq\frac{1}{5}$, and $2\beta_{t-1}^{2}+2\alpha_{t}^{2}\leq1/2$,
-which we can also check to hold by definition of $\alpha_{t}$ and
-$\beta_{t}$.
-
-Moreover, $\mathbb{E}\|\overline{y}_{t}-\vw^{\ast}\|^{2}\leq b(C\alpha_{t}^{2}+D\alpha_{t})$
-with the chosen step sizes also implies $\|\ov{v}_{t-1}-\vw^{\ast}\|\leq2\|\ov{v}_{t}-\vw^{\ast}\|$.
-Therefore the bound for $\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}$
-can be further simplified with 
-\begin{align*}
-2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\| & \leq4\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|^{2}
-\end{align*}
-and 
-\begin{align*}
-(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2} & \leq4(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t}-\vw^{\ast})\|^{2}
-\end{align*}
-so that
-\begin{align*}
-\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})((1+\beta_{t-1})^{2}+4\beta_{t-1}(1+\beta_{t-1})+4\beta_{t-1}^{2})\mathbb{E}\|(\ov{v}_{t}-\vw^{\ast})\|^{2}\\
-& +20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
-& \leq\mathbb{E}(1-\mu\alpha_{t})(1+14\beta_{t-1})\|(\ov{v}_{t}-\vw^{\ast})\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
-& \leq b(1-\frac{\mu\alpha_{t}}{2})(C\alpha_{t}^{2}+D\alpha_{t})+C\alpha_{t}^{3}+D\alpha_{t}^{2}\\
-& =(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}^{2}C+(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}D
-\end{align*}
-and so it remains to choose $b$ such that 
-\begin{align*}
-(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t} & \leq b\alpha_{t+1}\\
-(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}^{2} & \leq b\alpha_{t+1}^{2}
-\end{align*}
-from which we can conclude $\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}\leq\alpha_{t+1}^{2}C+\alpha_{t+1}D$.
-
-With $b=\frac{6}{\mu}$, we have
-\begin{align*}
-(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t} & =(b(1-(\frac{3}{t+\gamma})+\frac{6}{\mu(t+\gamma)})\frac{6}{\mu(t+\gamma)}\\
-& =(b\frac{t+\gamma-3}{t+\gamma}+\frac{6}{\mu(t+\gamma)})\frac{6}{\mu(t+\gamma)}\\
-& \leq b(\frac{t+\gamma-1}{t+\gamma})\frac{6}{\mu(t+\gamma)}\\
-& \leq b\frac{6}{\mu(t+\gamma+1)}=b\alpha_{t+1}
-\end{align*}
-where we have used $\frac{t+\gamma-1}{(t+\gamma)^{2}}\leq\frac{1}{t+\gamma+1}$.
-
-Similarly 
-\begin{align*}
-(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}^{2} & =(b(1-(\frac{3}{t+\gamma})+\frac{6}{\mu(t+\gamma)})(\frac{6}{\mu(t+\gamma)})^{2}\\
-& =(b\frac{t+\gamma-3}{t+\gamma}+\frac{6}{\mu(t+\gamma)})(\frac{6}{\mu(t+\gamma)})^{2}\\
-& =b(\frac{t+\gamma-2}{t+\gamma})(\frac{6}{\mu(t+\gamma)})^{2}\\
-& \leq b\frac{36}{\mu^{2}(t+\gamma+1)^{2}}=b\alpha_{t+1}^{2}
-\end{align*}
-where we have used $\frac{t+\gamma-2}{(t+\gamma)^{3}}\leq\frac{1}{(t+\gamma+1)^{2}}$.
-
-Finally, to ensure $\|\vv_{0}-\vw^{\ast}\|^{2}\leq b(C\alpha_{0}^{2}+D\alpha_{0})$,
-we can rescale $b$ by $c\|\vv_{0}-\vw^{\ast}\|^{2}$ for some $c.$ It
-follows that $\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}\leq b(C\alpha_{t}^{2}+D\alpha_{t})$
-for all $t$. Thus
-
-\begin{align*}
-\mathbb{E}(F(\ov{w}_{T}))-F^{\ast} & =\mathbb{E}(F(\ov{w}_{T})-F(\vw^{\ast}))\\
-& \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}\leq\frac{L}{2}c\|\vw_{0}-\vw^{\ast}\|^{2}\frac{6}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
-& =3c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
-& \leq3c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{6}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{\max}\sigma^{2}+20E^{2}LG^{2}\cdot(\frac{6}{\mu(T+\gamma)})^{2}\right]\\
-& =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{\max}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
-\end{align*}
-
-With partial participation, the same argument in the SGD case yields
-\begin{align*}
-\mathbb{E}F(\ov{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
-\end{align*}
+where we have used $\sigma^{2}=\sum_{k}p_{k}\sigma_{k}^{2}$, and by construction our $\alpha_{t}$
+satisfies $L\alpha_{t}\leq\frac{1}{5}$.
 \end{proof}

### 2021-02-06 source-iclrw/app_sgd_cvx_smooth.tex
4 pages v2

--- 2020-09-27 source-iclr/app_sgd_cvx_smooth.tex
@@
 
-	Similarly, for partial participation, we have 
+	For partial participation, the one step progress bound in Lemma \ref{lem:cvxoner} is updated in a similar manner as the strongly convex case in \eqref{eqn:scvx-partial-oner} to incorporate the sampling variance.
+More precisely, with partial participation,
 	\begin{align*}
-	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|\vw_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot(\frac{1}{N}\nu_{\max}\sigma^{2}+C)+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right)
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}+\ov{v}_{t+1}-\vw^{\ast}\|^{2}\\
+	& =\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2}+\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2},
 	\end{align*}
-	where $C=\frac{4}{K}E^{2}G^{2}$ or $\frac{N-K}{N-1}\frac{4}{K}E^{2}G^{2}$ depending on the sampling scheme,
+	where $\mathbb{E}\ov{w}_{t+1}=\ov{v}_{t+1}$ for all $t$, by the unbiasedness of our sampling schemes.
+Since $\ov{v}_t =\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$ always averages over all devices, the full participation one step progress bound in Lemma~\ref{lem:cvxoner} applied to $\ov{v}_t$ implies
+	\begin{align*}
+	\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{v}_{t})-F(\vw^{\ast})) & \leq\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}\\ &
+	   \leq\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
+	\end{align*}
+
+	The bound for $\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2}$ for the two sampling schemes we consider is provided in \eq{\ref{eq:partialsample}}, and applying it to the above bound we can write the one step progress for partial participation as
+	\begin{align*}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} + \alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}(\frac{1}{N}\nu_{max}^{2}\sigma^{2}+C)+6E^{2}L\alpha_{t}^{3}G^{2},
+	\end{align*}
+	where $C=\frac{4}{K}E^{2}G^{2}$ or $\frac{N-K}{N-1}\frac{4}{K}E^{2}G^{2}$ depending on the sampling scheme.
+
+	Summing up the one-step progress over $t$, 
+	\begin{align*}
+	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|\vw_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot(\frac{1}{N}\nu_{\max}\sigma^{2}+C)+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right),
+	\end{align*}
 	so that with $\alpha_{t}=\sqrt{\frac{K}{T}}$, we have 
 	\begin{align*}
-	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & =\mathcal{O}(\frac{\nu_{\max}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T})
+	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & =\mathcal{O}(\frac{\nu_{\max}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}).
 	\end{align*}

### 2021-02-06 source-iclrw/sec_conclude.tex
4 pages v2

--- 2020-06-03 source/sec_conclude.tex
@@
 We show that both Nesterov accelerated FedAvg and FedAvg can achieve linear speedup convergence for convex smooth problems and strongly 
-convex smooth problems.  
-Furthermore, this work also makes algorithmic contributions. We not only prove that FedAvg can achieve geometric convergence for overparameterized strongly convex smooth problems 
-but also propose the MaSS accelerated FedAvg algorithm, which has provable speedup in convergence rate over FedAvg on quadratic problems. Last but not least, extensive empirical 
+convex smooth problems. We further prove that FedAvg can achieve geometric convergence for overparameterized strongly convex smooth problems. Last but not least, extensive empirical 
 evaluations demonstrate the linear speedup of FedAvg and accelerated FedAvg under various settings.

### 2021-02-06 source-iclrw/sec_intro.tex
4 pages v2

@@
 \end{table}
-
 \textbf{Our Contributions}
@@
 {\small{$\cO(1/\sqrt{KT})$}} convergence rate for convex
-and smooth problems (where $K$ is the number of participating devices), thereby establishing that FedAvg enjoys the desirable linear speedup convergence, which improves the best known results\cite{li2019convergence,karimireddy2019scaffold} .
+and smooth problems (where $K$ is the number of participating devices), thereby establishing that FedAvg enjoys the desirable linear speedup convergence, which improves the best known results~\cite{li2019convergence,karimireddy2019scaffold} (see Section~\ref{sec:sgd}).
 \begin{comment}
@@
 \end{comment}
-Second, we establish the same convergence rates--{\small{$\cO(1/KT)$}} for strongly convex and smooth problems and {\small{$\cO(1/\sqrt{KT})$}} for convex and smooth problems--for Nesterov accelerated FedAvg.
+Second, we establish the same convergence rates--{\small{$\cO(1/KT)$}} for strongly convex and smooth problems and {\small{$\cO(1/\sqrt{KT})$}} for convex and smooth problems--for Nesterov accelerated FedAvg (see Appendix Section~\ref{sec:Nesterov}).
 \begin{comment}
@@
 Third, we study a subclass of strongly convex smooth problems where the objective is over-parameterized and establish 
-a faster $\cO(\exp(-\frac{KT}{\kappa}))$ convergence rate for FedAvg, in contrast to the $\cO(\exp(-\frac{T}{\kappa}))$ rate for individual solvers~\cite{ma2017power}. 
+a faster $\cO(\exp(-\frac{KT}{\kappa}))$ convergence rate for FedAvg, in contrast to the $\cO(\exp(-\frac{T}{\kappa}))$ rate for individual solvers~\cite{ma2017power} (see Appendix Section~\ref{sec:app:overparameterized}).
+In addition to the theoretical contributions, we also conduct extensive empirical evaluations demonstrating the linear speedup of FedAvg and accelerated FedAvg under various settings (see Appendix Section~\ref{sec:exp}).
 \begin{comment}

### 2021-06-26 source-iclrw/main.tex
4 pages v1 todo: sec 4, 5 transition
4 pages v2

--- 2020-09-25 source-iclr/main.tex
@@
 
+\iclrfinalcopy
+
 \title{Federated Learning's Blessing:\\
@@
   \texttt{linkaixi@msu.edu} \\
-  \AND
-  Jayant Kalagnanam \\
-  IBM Research \\
-  \texttt{jayant@us.ibm.com} \\
   \And
@@
   \texttt{jiayuz@msu.edu} \\
-  \And
-  Zhengyuan Zhou\\
-  New York University \\
-  \texttt{zzhou@stern.nyu.edu} \\
+  %\And
+  %Zhengyuan Zhou\\
+  %New York University \\
+  %\texttt{zzhou@stern.nyu.edu} \\
 }
@@
 We fill this gap by providing a unified analysis that establishes convergence guarantees for FedAvg under three classes of problems: strongly convex smooth, convex smooth, and overparameterized strongly convex smooth problems.
-We show that FedAvg enjoys linear speedup in each case, although with different convergence rates and communciation efficiencies.
+We show that FedAvg enjoys linear speedup in each case, although with different convergence rates and communication efficiencies.
 While there have been linear speedup results from distributed optimization that assumes full participation, ours are the first to establish linear speedup for FedAvg under both statistical and system heterogeneity.
@@
 \input{sec_SGD}
-\input{sec_Nesterov}
-%\input{sec_overparameterized}
 \input{sec_exp}
 
-% \input{sec_conclude}
+\input{sec_conclude}
+
+\subsubsection*{Acknowledgements}
+This research is supported in part by Stanford Interdisciplinary Graduate Fellowship (ZQ),  NSF CMMI-2030411 (ZL), NSF IIS-1749940 (JZ), ONR N00014-20-1-2382 (JZ).
 

### 2021-06-27 source-TNNLS/main.tex
TNNLS version

--- 2020-09-25 source-iclr/main.tex
@@
-\documentclass{article}
+\documentclass[journal]{IEEEtran}
+\ifCLASSINFOpdf
+  % \usepackage[pdftex]{graphicx}
+  % declare the path(s) where your graphic files are
+  % \graphicspath{{../pdf/}{../jpeg/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+\else
+  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
+  % will default to the driver specified in the system graphics.cfg if no
+  % driver is specified.
+  % \usepackage[dvips]{graphicx}
+  % declare the path(s) where your graphic files are
+  % \graphicspath{{../eps/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  % \DeclareGraphicsExtensions{.eps}
+\fi
 
-% if you need to pass options to natbib, use, e.g.:
-%     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2020
-
-% ready for submission
-\usepackage{iclr2021_conference,times}
-
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
-    % \usepackage[preprint]{neurips_2020}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-%     \usepackage[final]{neurips_2020}
-
-% to avoid loading the natbib package, add option nonatbib:
-     % \usepackage[nonatbib]{neurips_2020}
 \usepackage{slashbox,multirow}
@@
 \usepackage{amsmath}
+\usepackage{time}
 \allowdisplaybreaks
@@
 
-\title{Federated Learning's Blessing:\\
-FedAvg has Linear Speedup}
-\author{%
-  Zhaonan Qu\thanks{Equal contribution} \\
- Stanford University\\
-  \texttt{zhaonanq@stanford.edu} \\
-  \And
-  Kaixiang Lin$^*$ \\
-  Michigan State University \\
-  \texttt{linkaixi@msu.edu} \\
-  \AND
-  Jayant Kalagnanam \\
-  IBM Research \\
-  \texttt{jayant@us.ibm.com} \\
-  \And
-  Zhaojian Li \\
-  Michigan State University \\
-  \texttt{lizhaoj1@egr.msu.edu} \\
-  \And
-  Jiayu Zhou\\
-  Michigan State University \\
-  \texttt{jiayuz@msu.edu} \\
-  \And
-  Zhengyuan Zhou\\
-  New York University \\
-  \texttt{zzhou@stern.nyu.edu} \\
+% correct bad hyphenation here
+\hyphenation{}
+
+\begin{document}
+
+\title{Federated Learning's Blessing: FedAvg has Linear Speedup}
+
+\author{Zhaonan Qu$^*$, Kaixiang Lin$^*$, Zhaojian Li, Jiayu Zhou, and Zhengyuan Zhou% <-this % stops a space
+\thanks{$^*$ The first two authors contribute equally.}
+\thanks{Zhaonan Qu is with the Department of Economics, Stanford University, CA 94305, USA. Email: {zhaonanq}@stanford.edu}% <-this % stops a space
+\thanks{Kaixiang Lin was with the Department of Computer Science and Engineering, Michigan State University, Lansing, MI, 48824,
+USA. Email: {linkaixi}@msu.edu.
+}% <-this % stops a space
+\thanks{Zhaojian Li is with the Department of
+Mechanical Engineering, Michigan State University, Lansing, MI, 48824,
+USA. Email: {lizhaoj1}@msu.edu.}
+\thanks{Jiayu Zhou is with the Department of Computer Science and Engineering, Michigan State University, Lansing, MI, 48824,
+USA. Email: {jiayuz}@msu.edu.}
+\thanks{Zhengyuan Zhou is with the Department of Technology, Operations, and Statistics at Stern School of Business, New York University, New York, NY 10003,
+USA. Email: {zzhou}@stern.nyu.edu.}
 }
 
-\begin{document}
+% The paper headers
+\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
+{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}
 
@@
 \begin{abstract}
-\small Federated learning (FL) learns a model jointly from a set of participating devices without sharing each other's privately held data.
+Federated learning (FL) learns a model jointly from a set of participating devices without sharing each other's privately held data.
 The characteristics of non-\textit{i.i.d.} data across the network, low device participation, high communication costs, and the mandate that data remain private bring challenges in understanding the convergence of FL algorithms, particularly in regards to how convergence scales with the number of participating devices.
@@
 
-% (and its related variant local SGD also analyzed in the simpler homogeneous distributed optimization setting)
+% Note that keywords are not normally used for peerreview papers.
+\begin{IEEEkeywords}
+IEEE, IEEEtran, journal, \LaTeX, paper, template.
+\end{IEEEkeywords}
+
+\IEEEpeerreviewmaketitle
+
 \input{sec_intro}
@@
 
-% \input{sec_conclude}
+% \bibliographystyle{IEEE}
+\newpage
+\bibliographystyle{abbrv}
+%\bibliographystyle{unsrt}
+\bibliography{ref}
 
-% \newpage
-% \input{sec_impact}
+\appendices
+
+\input{sec_appendix}
+
+% use section* for acknowledgment
+\section*{Acknowledgment}
+
+% Can use something like this to put references on a page
+% by themselves when using endfloat and the captionsoff option.
+\ifCLASSOPTIONcaptionsoff
   \newpage
-\bibliographystyle{iclr2021_conference}
-\bibliography{ref.bib}
+\fi
 
-\newpage
-\appendix
-\input{sec_appendix}
+% 
+% \begin{thebibliography}{1}
+% \bibitem{IEEEhowto:kopka}
+
+% \end{thebibliography}
+
+\begin{IEEEbiography}{Zhaonan Qu}
+Biography text here.
+\end{IEEEbiography}
+
+% if you will not have a photo at all:
+\begin{IEEEbiographynophoto}{Kaixiang Lin}
+Biography text here.
+\end{IEEEbiographynophoto}
+
+% insert where needed to balance the two columns on the last page with
+% biographies
+%\newpage
+
+\begin{IEEEbiographynophoto}{Jane Doe}
+Biography text here.
+\end{IEEEbiographynophoto}
+
+% You can push biographies down or up by placing
+% a \vfill before or after them. The appropriate
+% use of \vfill depends on what kind of text is
+% on the last page and whether or not the columns
+% are being equalized.
+
+%\vfill
+
+% Can be used to pull up biographies so that the bottom of the last one
+% is flush with the other column.
+%\enlargethispage{-5in}
+
+% that's all folks
 \end{document}
-
-\iffalse
-
-\endif

### 2021-06-27 source-TNNLS/sec_Nesterov.tex
TNNLS version

--- 2020-11-22 source-iclr/sec_Nesterov.tex
@@
 
-{\color{red}A natural extension of the FedAvg algorithm is to use momentum-based 
+A natural extension of the FedAvg algorithm is to use momentum-based 
 local updates instead of local SGD updates in order to accelerate FedAvg. As we know from stochastic optimization, Nesterov and other momentum
@@
 In fact, previous works such as \cite{stich2018local} on FedAvg with vanilla SGD uses FedAvg with Nesterov or other momentum updates in their experiments to achieve target accuracy.
-Because of the popularity of Nesterov and other momentum-based methods, understanding the linear speedup behavior of FedAvg with such local updates is important.} To our knowledge, the
+Because of the popularity of Nesterov and other momentum-based methods, understanding the linear speedup behavior of FedAvg with such local updates is important.
+To our knowledge, the
 only convergence analyses of FedAvg with momentum-based stochastic

### 2021-06-27 source-TNNLS/sec_intro.tex
TNNLS version
Update on Overleaf.

--- 2020-11-22 source-iclr/sec_intro.tex
@@
 
-\begin{table}[t!]
-\hspace{-1em}
-{\small
+\begin{table*}[th!]
 \centering
@@
 \end{tabular}
-}
-\caption{\small Our convergence results for FedAvg and accelerated FedAvg in this paper. Throughout the paper, $N$ is the total
+\caption{Our convergence results for FedAvg and accelerated FedAvg in this paper. Throughout the paper, $N$ is the total
 number of local devices, and $K \leq N$ is the maximal number of devices that are accessible to the central server.
@@
 \label{tb:convergencerateintro}
-\vspace{-2.5em}
-\end{table}
+\end{table*}
 
@@
 For a detailed comparison with related works, please refer to Table~\ref{tb:convergenceratev3} in Appendix Section~\ref{sec:app:comparison}.
+
+\textbf{New related works since last time}: \cite{zhu2019multi,xu2020ternary,yang2020federated,chen2019communication,sattler2020clustered,konevcny2016federated,yuan2020stochastic,dinh2020federated,li2020acceleration,nedic2009distributed,duchi2011dual}
 

### 2021-06-27 source-TNNLS/main.tex
Update on Overleaf.
Add bio and reformat appendix

@@
 \begin{IEEEkeywords}
-IEEE, IEEEtran, journal, \LaTeX, paper, template.
+Federated learning, Optimization. 
 \end{IEEEkeywords}
@@
 \input{sec_exp}
+\input{sec_conclude}
+
+\input{sec_ack}
 
@@
 
+\begin{comment}
+
+\begin{IEEEbiography}{Zhaonan Qu}
+Biography text here.
+\end{IEEEbiography}
+
+% if you will not have a photo at all:
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fig/bio_kaixiang.png}}]{Kaixiang Lin}
+Dr. Kaixiang Lin is an applied scientist at Amazon Lab 126. Before joining Amazon, 
+Kaixiang received his Ph.D. degree in Computer Science at Michigan State University in 2020. 
+\end{IEEEbiography}
+
+% insert where needed to balance the two columns on the last page with
+% biographies
+%\newpage
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fig/bio_jiayu.png}}]{Jiayu Zhou}
+Dr. Jiayu Zhou is an assistant professor at
+ Department of Computer Science and Engineering, Michigan State University.
+ Before joining MSU, Jiayu was a staff research scientist at Samsung Research
+ America. Jiayu received his Ph.D. degree in computer science at Arizona
+ State University in 2014. 
+\end{IEEEbiography}
+
+\end{comment}
+
 \appendices
@@
 \input{sec_appendix}
-
-% use section* for acknowledgment
-\section*{Acknowledgment}
 
@@
 % \end{thebibliography}
-
-\begin{IEEEbiography}{Zhaonan Qu}
-Biography text here.
-\end{IEEEbiography}
-
-% if you will not have a photo at all:
-\begin{IEEEbiographynophoto}{Kaixiang Lin}
-Biography text here.
-\end{IEEEbiographynophoto}
-
-% insert where needed to balance the two columns on the last page with
-% biographies
-%\newpage
-
-\begin{IEEEbiographynophoto}{Jane Doe}
-Biography text here.
-\end{IEEEbiographynophoto}
 

### 2021-07-04 source-TNNLS/app_sgd_scvx_smooth.tex
Add bio and reformat appendix
Update on Overleaf.

--- 2020-09-25 source-iclr/app_sgd_scvx_smooth.tex
@@
-% !TEX ROOT=./supp-main.tex
+% !TEX ROOT=./main.tex
 
@@
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}L\alpha_{t}^{3}G^{2}.
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}\\
+	+& \alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}L\alpha_{t}^{3}G^{2}.
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}L\alpha_{t}^{3}G^{2}.
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}\\
+	& +\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}L\alpha_{t}^{3}G^{2}.
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq (b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}D+(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}C
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq (b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}D\\
+	& +(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}C
 	\end{align*}
@@
 	\begin{align*}
-	& (b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}  =\left(b(1-\frac{4}{t+\gamma})+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}
+	(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}  &=\left(b(1-\frac{4}{t+\gamma})+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}\\
 	% & =\left(b\frac{t+\gamma-4}{t+\gamma}+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}\\
@@
 	% & \leq b(\frac{t+\gamma-1}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
-	 \leq b\frac{4}{\mu(t+\gamma+1)}=b\alpha_{t+1}\\
-	& (b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2} 
-	 =b(\frac{t+\gamma-2}{t+\gamma})\frac{16}{\mu^{2}(t+\gamma)^{2}} \leq b\frac{16}{\mu^{2}(t+\gamma+1)^{2}}=b\alpha_{t+1}^{2}
+	 &\leq b\frac{4}{\mu(t+\gamma+1)}=b\alpha_{t+1}\\
+	 (b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2} &=b(\frac{t+\gamma-2}{t+\gamma})\frac{16}{\mu^{2}(t+\gamma)^{2}} \\
+	 & \leq b\frac{16}{\mu^{2}(t+\gamma+1)^{2}}=b\alpha_{t+1}^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\frac{t+\gamma-1}{(t+\gamma)^{2}} \leq\frac{1}{(t+\gamma+1)}   \hspace{2em}
-	\frac{t+\gamma-2}{(t+\gamma)^{3}} \leq\frac{1}{(t+\gamma+1)^{2}}  \hspace{2em}  \forall\ \gamma\geq1
+	\frac{t+\gamma-1}{(t+\gamma)^{2}} \leq\frac{1}{(t+\gamma+1)},\hspace{0.2em} 
+	\frac{t+\gamma-2}{(t+\gamma)^{3}} \leq\frac{1}{(t+\gamma+1)^{2}}, \hspace{0.2em}  \forall\ \gamma\geq1
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}(F(\ov{w}_{T}))-F^{\ast} & \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}\\ 
-	& \leq\frac{L}{2}c\|\vw_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
-	& =2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
-	& \leq2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{4}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}LG^{2}\cdot(\frac{4}{\mu(T+\gamma)})^{2}\right]\\
-	& =\mathcal{O}(\frac{\kappa}{\mu}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
+	& \mathbb{E}(F(\ov{w}_{T}))-F^{\ast} \leq \frac{L}{2}\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}\\ 
+	\leq &\frac{L}{2}c\|\vw_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
+	=& 2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
+	\leq &2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{4}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}LG^{2}\cdot(\frac{4}{\mu(T+\gamma)})^{2}\right]\\
+	=& \mathcal{O}(\frac{\kappa}{\mu}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
 	\end{align*}
@@
 	\begin{align*}
-	   \mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{v}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\\ &
-	   \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
+	   \mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{v}_{t}-\vw^{\ast}\|^{2}\\
+	   & +6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\\ &
+	   \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}\\
+	   & +6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
 	\end{align*}
@@
 	The bound for $\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2}$ for the two sampling schemes we consider is provided in \eq{\ref{eq:partialsample}}, and applying it we can write the one step progress for partial participation as
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}+6E^{2}L\alpha_{t}^{3}G^{2},
-	\end{align*}
+	\begin{align}\label{eqn:scvx-partial-oner}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\nonumber\\
+	&+\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}+6E^{2}L\alpha_{t}^{3}G^{2},
+	\end{align}
 	and the same arguments using induction and $L$-smoothness as the full participation case implies 
 	\begin{align*}
-	\mathbb{E}F(\ov{w}_{T})-F^{\ast}=\mathcal{O}(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\ov{w}_{T})-F^{\ast}=\mathcal{O}(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}
+	+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
 	\end{align*}
@@
 		\begin{align*}
-	& \mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\\
-= & \mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
-\leq &\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
-	= &\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
-	= &\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}\vg_{i,k}\|^{2}\\
-	\leq & 2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}E\alpha_{i}^{2}\|\vg_{i,k}\|^{2}\\
-	\leq & 2\sum_{k=1}^{N}p_{k}E^{2}\alpha_{t_{0}}^{2}G^{2}\\
-	\leq & 4E^{2}\alpha_{t}^{2}G^{2}
+	& \mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}
+= \mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
+&\leq \mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2}
+= \mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
+	&= \mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}\vg_{i,k}\|^{2}
+	\leq  2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}E\alpha_{i}^{2}\|\vg_{i,k}\|^{2}\\
+	&\leq  2\sum_{k=1}^{N}p_{k}E^{2}\alpha_{t_{0}}^{2}G^{2}
+	\leq  4E^{2}\alpha_{t}^{2}G^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}\vg_{t})-\vw^{\ast}\|^{2} =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(\vg_{t}-\ov{g}_{t})\|^{2}\\
-	& = \underbrace{\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}}_{A_1} + \underbrace{2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-\vg_{t}\rangle}_{A_2} + \underbrace{\alpha_{t}^{2}\|\vg_{t}-\ov{g}_{t}\|^{2}}_{A_3}
+	&\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}  =\|(\ov{w}_{t}-\alpha_{t}\vg_{t})-\vw^{\ast}\|^{2}\\ &=\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(\vg_{t}-\ov{g}_{t})\|^{2} = \underbrace{\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}}_{A_1} \\ &+\underbrace{2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-\vg_{t}\rangle}_{A_2} + \underbrace{\alpha_{t}^{2}\|\vg_{t}-\ov{g}_{t}\|^{2}}_{A_3}
 	\end{align*}
@@
 	\begin{align*}
-	\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2}
+	&\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} \\ &=\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2}
 	\end{align*}
 	and we will show that the third term $\|\alpha_{t}\ov{g}_{t}\|^{2}$
-	can be canceled by an upper bound of the second term, which is one of major improvement comparing to prior art~\cite{li2019convergence}.
+	can be canceled by an upper bound of the second term, which is one of our major improvements compared to prior works~\cite{li2019convergence}.
 	\begin{comment}
@@
 	\begin{align*}
-	& -2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle\\
-	=& -2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	=& -2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle \vw_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	\leq&-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
-	\leq& 2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
-	=& \alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
+	 -&2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle= -2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	= &-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	-&2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle \vw_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	\leq&-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	+&2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
+	\leq & 2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
+	= &\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]\\
+	-&\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
 	\end{align*}
@@
 	\begin{align*}
-	 \mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}
-	\leq & \mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\\
-	 & +2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} \numberthis \label{eq:common one step}
+	 &\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}
+	\leq  \mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\\
+	 +&2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} \numberthis \label{eq:common one step}
 	\end{align*}
@@
 	\leq& 2\alpha_{t}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{w}_{t+1}-\ov{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\ov{w}_{t+1}-\ov{w}_{t}\|^{2}\\
-	 =& -2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\vg_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
 	=&-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	=&-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\ov{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	=&-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\sum_{k}p_{k}\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-   \leq&-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\ov{w}_{t})-\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	=&-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\ov{g}_t\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
 	\leq&-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\

### 2021-07-04 source-TNNLS/sec_problem.tex
TNNLS version
Update on Overleaf.

--- 2020-09-25 source-iclr/sec_problem.tex
@@
 \label{eq:fedavg updates}
-\vv_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}\vg_{t,k}, \hspace{1em}
-\mathbf{w}_{t+1}^{k} =\left\{
+\vv_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}\vg_{t,k}, \nonumber\\
+\mathbf{w}_{t+1}^{k} & =\left\{
 \begin{array}{ll}
-\mathbf{v}_{t+1}^{k} & \text { if } t+1 \notin \mathcal{I}_{E}, \\ 
+\mathbf{v}_{t+1}^{k}  \text { if } t+1 \notin \mathcal{I}_{E}, \\ 
 \sum_{k \in \cS_{t+1}} q_k \mathbf{v}_{t+1}^{k} & \text { if } t+1 \in \mathcal{I}_{E},
@@
 
-\begin{assumption}[$L$-smooth]
-	$F_{1}, \cdots, F_{N}$ are all $L$-smooth: for all  $\mathbf{v}$  and $\mathbf{w}$, $F_{k}(\mathbf{v}) \leq F_{k}(\mathbf{w})+(\mathbf{v}- \\ \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{L}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$.
+\begin{assumption}[$L$-smoothness]
+	$F_{1}, \cdots, F_{N}$ are all $L$-smooth: for all  $\mathbf{v}$  and $\mathbf{w}$, 
+	\begin{equation}
+	  F_{k}(\mathbf{v}) \leq F_{k}(\mathbf{w})+(\mathbf{v}-  \mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{L}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}.
+	\end{equation}
 	\label{ass:lsmooth}
 \end{assumption}
-\begin{assumption}[Strongly-convex]
-	$	F_{1}, \cdots, F_{N} \text { are all } \mu \text { -strongly convex: for all v and } \mathbf{w}, F_{k}(\mathbf{v}) \geq F_{k}(\mathbf{w})+(\mathbf{v}-\mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{\mu}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$
+\begin{assumption}[Strong convexity]
+	$F_{1}, \cdots, F_{N}$ are all $\mu$-strongly convex: for all $v$ and  $\mathbf{w}$,
+	\begin{equation}
+	   F_{k}(\mathbf{v}) \geq F_{k}(\mathbf{w})+(\mathbf{v}-\mathbf{w})^{T} \nabla F_{k}(\mathbf{w})+\frac{\mu}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}.
+	\end{equation}
 	\label{ass:stroncvx}

### 2021-07-04 source-nips/MaSS_FedAvg_1.tex
Update on Overleaf.

%% LyX 2.3.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{algorithm,algpseudocode}

\renewcommand\[{\begin{equation}}
\renewcommand\]{\end{equation}}

\makeatother

\usepackage{babel}
\providecommand{\theoremname}{Theorem}

\begin{document}

\section{Exponential Convergence of FedAvg with MaSS for Linear Regression
Problem with 0 Loss}

The MaSS algorithm is a variant of the Nesterov acceleration, where
in the update a non-negative multiple of the gradient is added to
the Nesterov parameter update, to correct for the ``over-descent''
of the Nesterov update. In {[}Liu\&Belkin{]}, the authors show that
for stochastic gradient methods in the interpolation setting, Nesterov
update is not able to achieve accleration over vanilla SGD, whereas
the MaSS algorithm is able to achieve acceleration both theoretically
and empirically. We adapt the MaSS algorithm to our federated learning
setting, and prove that for certain class of convex objectives with
0 global objective value, the FedAvg with MaSS update achieves exponential
convergence and acceleration over SGD.

We first show that FedAvg with MaSS has exponential convergence for
linear regression when the global objective has 0 as its global minimum.
Since the global loss is given by 
\begin{align*}
F(w) & =\sum_{k=1}^{N}p_{k}F_{k}(w)\\
F_{k}(w) & =\frac{1}{2n_{k}}\sum_{j=1}^{n_{k}}(w^{T}x_{k,j}-z_{k,j})^{2}
\end{align*}
 and there exists $w^{\ast}$ such that $F(w^{\ast})=0$, in particular
this implies that $F_{k}^{\ast}=F_{k}(w^{\ast})=F^{\ast}=0$ for all
$k$.

\subsection{Notation and Definitions}

Following Liu\&Belkin and Jain et al., we define some condition number
related quantities. Let $H^{k}=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}x_{k,j}x_{k,j}$
be the Hessian matrix of $F_{k}$. Let $L^{k}$ and $\mu^{k}$ be
the largest and smallest non-zero eigenvalues of the Hessians $H^{k}$.
For a mini-batch $\{\tilde{x}_{j}\}_{j=1}^{m}$ of $m$ samples from
device $k$, let $\tilde{H}_{m}^{k}=\frac{1}{m}\sum_{j=1}^{m}\tilde{x}_{j}\tilde{x}_{j}^{T}$
be the unbiased mini-batch estimate of $H^{k}$. Let $L_{1}^{k}$
be the smallest positive number such that 
\begin{align*}
\mathbb{E}\left[\|\tilde{x}\|^{2}\tilde{x}\tilde{x}^{T}\right] & \preceq L_{1}^{k}H^{k}
\end{align*}
 and define 
\begin{align*}
L_{m}^{k} & =L_{1}^{k}/m+(m-1)L/m
\end{align*}

Define the $m$-stochastic condition number as $\kappa_{m}^{k}:=L_{m}^{k}/\mu$.
Define the statistical condition number $\tilde{\kappa}^{k}$ as the
smallest positive real number such that 
\begin{align*}
\mathbb{E}\left[\langle\tilde{x}(H^{k})^{-1},\tilde{x}\rangle\tilde{x}\tilde{x}^{T}\right] & \preceq\tilde{\kappa}^{k}H^{k}
\end{align*}
 Note that $L_{m}^{k}\geq L$, and $\kappa_{m}^{k}\geq\kappa^{k}$,
while $\tilde{\kappa}^{k}\leq\kappa^{1}$.

\subsection{FedAvg with MaSS}

The FedAvg algorithm with MaSS follows the updates
\begin{align*}
w_{t+1}^{k} & =\begin{cases}
u_{t}^{k}-\eta_{1}^{k}g_{t,k} & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}\left[u_{t}^{k}-\eta_{1}^{k}g_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}\\
u_{t+1}^{k} & =w_{t+1}^{k}+\gamma^{k}(w_{t+1}^{k}-w_{t}^{k})+\eta_{2}^{k}g_{t,k}
\end{align*}
 where we note that the natural parameter is $w_{t}$, while $u_{t}$
is an auxiliary parameter, which we initialize to be $u_{0}^{k}\equiv u_{0}$,
and 
\begin{align*}
g_{t,k} & :=\nabla F_{k}(w_{t}^{k},\xi_{t}^{k})
\end{align*}
 is the stochastic gradient and 
\begin{align*}
g_{t} & =\sum_{k=1}^{N}p_{k}g_{t,k}
\end{align*}
is the averaged stochastic gradient. When $\eta_{2}^{k}\equiv0$,
this reduces to the FedAvg algorithm with Nesterov updates.

We note that the update can equivalently be written as 
\begin{align*}
v_{t+1}^{k} & =(1-\alpha^{k})v_{t}^{k}+\alpha^{k}w_{t}^{k}-\delta^{k}g_{t,k}\\
w_{t+1}^{k} & =\begin{cases}
u_{t}^{k}-\eta^{k}g_{t,k} & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}\left[u_{t}^{k}-\eta^{k}g_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}\\
u_{t+1}^{k} & =\frac{\alpha^{k}}{1+\alpha^{k}}v_{t+1}^{k}+\frac{1}{1+\alpha^{k}}w_{t+1}^{k}
\end{align*}
 where there is a bijection between the parameters 
\begin{align*}
\frac{1-\alpha^{k}}{1+\alpha^{k}} & =\gamma^{k}\\
\eta^{k} & =\eta_{1}^{k}\\
\frac{\eta^{k}-\alpha^{k}\delta^{k}}{1+\alpha^{k}} & =\eta_{2}^{k}
\end{align*}
 and we further introduce an auxiliary parameter $v_{t}^{k}$, which
is initialized at $v_{0}^{k}\equiv v_{0}$. We also note that when
$\delta_{t}=\frac{\eta_{t}}{\alpha_{t}}$, the update reduces to the
Nesterov accelerated SGD. This version of the FedAvg with MaSS algorithm
is used for analyzing the exponential convergence.

As before, define the virtual sequences $\overline{w}_{t}=\sum_{k=1}^{N}p_{k}w_{t}^{k}$,
$\overline{v}_{t}=\sum_{k=1}^{N}p_{k}v_{t}^{k}$, $\overline{u}_{t}=\sum_{k=1}^{N}p_{k}u_{t}^{k}$,
and $\overline{g}_{t}=\sum_{k=1}^{N}p_{k}\mathbb{E}g_{t,k}$. We have
$\mathbb{E}g_{t}=\overline{g}_{t}$ and $\overline{w}_{t+1}=\overline{u}_{t}-\eta_{t}g_{t}$,
$\overline{v}_{t+1}=(1-\alpha^{k})\overline{v}_{t}+\alpha^{k}\overline{w}_{t}-\delta^{k}g_{t}$,
and $\overline{w}_{t+1}=\frac{\alpha^{k}}{1+\alpha^{k}}\overline{v}_{t+1}+\frac{1}{1+\alpha^{k}}\overline{y}_{t+1}$.

For the linear regression problem in the interpolation setting, we
can write 
\begin{align*}
F(w) & =\frac{1}{2}(w-w^{\ast})^{T}H(w-w^{\ast})\\
 & =\frac{1}{2}\|w-w^{\ast}\|_{H}^{2}
\end{align*}
 and similarly $F^{k}(w)=\frac{1}{2}\|w-w^{\ast}\|_{H^{k}}^{2}$,
so that 
\begin{align*}
g_{t,k} & =\tilde{H}_{t}^{k}(w_{t}^{k}-w^{\ast})\\
g_{t} & =\sum_{k=1}^{N}p_{k}\tilde{H}_{t}^{k}(w_{t}^{k}-w^{\ast})
\end{align*}

\subsection{Exponential Convergence when Global Loss is 0}

We now present the exponential convergence result using FedAvg with
MaSS updates. On each device, local data is stored and mini-batch
gradient descent with batch size $m^{k}$ is performed. We allow for
varying batch sizes across devices. 
\begin{thm}
(FedAvg with MaSS, Full Participation) Let the hyperparameters satisfy
the requirements of {[}Liu\&Belkin{]}. More precisely, let $\tilde{\kappa}_{m}:=\tilde{\kappa}/m^{k}+(m^{k}-1)/m^{k}$,
and let the hyper parameters satisfy 
\begin{align*}
\eta^{k}(m)=\frac{1}{L_{m}}, & \alpha^{k}(m)=\frac{1}{\sqrt{\kappa_{m}^{k}\tilde{\kappa}_{m}^{k}}},\delta^{k}(m)=\frac{\eta^{k}}{\alpha^{k}\tilde{\kappa}_{m}^{k}}
\end{align*}
Let $\mathcal{I}_{E}=\ell E$ where $\ell\in\mathbb{N}$ be the set
of communication rounds, then the updates 
\begin{align*}
v_{t+1}^{k} & =(1-\alpha^{k})v_{t}^{k}+\alpha^{k}w_{t}^{k}-\delta^{k}g_{t,k}\\
w_{t+1}^{k} & =\begin{cases}
u_{t}^{k}-\eta^{k}g_{t,k} & \text{if }t+1\notin\mathcal{I}_{E}\\
\sum_{k=1}^{N}p_{k}\left[u_{t}^{k}-\eta^{k}g_{t,k}\right] & \text{if }t+1\in\mathcal{I}_{E}
\end{cases}\\
u_{t+1}^{k} & =\frac{\alpha^{k}}{1+\alpha^{k}}v_{t+1}^{k}+\frac{1}{1+\alpha^{k}}w_{t+1}^{k}
\end{align*}
 where $g_{t,k}$ are the mini-batch stochastic gradients achieves
the convergence
\begin{align*}
\|\overline{w}_{t}-w^{\ast}\|^{2} & \leq C\cdot(1-1/\sqrt{\kappa_{m}}\\
\end{align*}
\end{thm}
\begin{proof}
Note that at each communication round we update the $w_{t+1}^{k}$
parameters to be the average across devices while fixing $v_{t+1}^{k}$.
This automatically adjusts the $u_{t+1}^{k}$ parameter at each device
by the relation 
\begin{align*}
u_{t+1}^{k} & =\frac{\alpha^{k}}{1+\alpha^{k}}v_{t+1}^{k}+\frac{1}{1+\alpha^{k}}w_{t+1}^{k}
\end{align*}
 valid for all $t\geq0$.

\textbf{For now assume $\delta^{k}=\delta$ and $\alpha^{k}=\alpha$
is equal across devices. Need to relax later. }

Theorems 2 and 3 of the Liu\&Belkin paper gives the bound 
\begin{align*}
\mathbb{E}\left[\|v_{E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|u_{E}^{k}-\eta^{k}g_{E,k}-w^{\ast}\|^{2}\right] & \leq(1-\alpha^{k})^{E}(\|v_{0}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|w_{0}^{k}-w^{\ast}\|^{2})
\end{align*}
for all $k$, where $E$ is the first communication round. Note that
$w_{E}^{k}=\overline{w}_{E}^{k}\neq u_{E}^{k}-\eta^{k}g_{t,k}$.

It follows from convexity that 
\begin{align*}
\mathbb{E}\left[\sum_{k=1}^{N}p_{k}\|v_{E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|\overline{w}_{E}-w^{\ast}\|^{2}\right] & \leq\sum_{k=1}^{N}p_{k}\mathbb{E}\left[\|v_{E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|u_{E}^{k}-\eta^{k}g_{E,k}-w^{\ast}\|^{2}\right]\\
 & \leq\sum_{k=1}^{N}p_{k}(1-\alpha^{k})^{E}(\|v_{0}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|w_{0}^{k}-w^{\ast}\|^{2})
\end{align*}
Since $w_{E}^{k}=\overline{w}_{E}$ for all devices, applying the
per-device result again starting at $t=E$ instead of $t=0$, for
each device we have the bound
\begin{align*}
\mathbb{E}\left[\|v_{2E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|u_{2E}^{k}-\eta^{k}g_{2E,k}-w^{\ast}\|^{2}\right] & \leq(1-\alpha^{k})^{E}\mathbb{E}(\|v_{E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|w_{E}^{k}-w^{\ast}\|^{2})\\
 & =(1-\alpha^{k})^{E}\mathbb{E}(\|v_{E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|\overline{w}_{E}-w^{\ast}\|^{2})
\end{align*}

Here we emphasize that $w_{E}^{k}$ results from broadcasting and
so is the same across all devices, while $v_{E}^{k}$ remains distinct
on each device (and is only auxiliary). Then by convexity and summing
the above inequalities across devices we have 
\begin{align*}
\mathbb{E}\left[\sum_{k=1}^{N}p_{k}\|v_{2E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|\overline{w}_{2E}-w^{\ast}\|^{2}\right] & \leq\sum_{k=1}^{N}p_{k}\mathbb{E}\left[\|v_{2E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|u_{2E}^{k}-\eta^{k}g_{2E,k}-w^{\ast}\|^{2}\right]\\
 & \leq\sum_{k=1}^{N}p_{k}(1-\alpha^{k})^{E}\mathbb{E}(\|v_{E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|\overline{w}_{E}-w^{\ast}\|^{2})\\
 & \leq\sum_{k=1}^{N}p_{k}(1-\alpha^{k})^{2E}(\|v_{0}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|w_{0}^{k}-w^{\ast}\|^{2})
\end{align*}
 and by induction we can show that 
\begin{align*}
\mathbb{E}\left[\sum_{k=1}^{N}p_{k}\|v_{\ell E}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|\overline{w}_{\ell E}-w^{\ast}\|^{2}\right] & \leq\sum_{k=1}^{N}p_{k}(1-\alpha^{k})^{\ell E}(\|v_{0}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|w_{0}^{k}-w^{\ast}\|^{2})
\end{align*}
 and more generally
\begin{align*}
\mathbb{E}\left[\sum_{k=1}^{N}p_{k}\|v_{t}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|\overline{w}_{t}-w^{\ast}\|^{2}\right] & \leq\sum_{k=1}^{N}p_{k}(1-\alpha^{k})^{t}(\|v_{0}^{k}-w^{\ast}\|_{H_{k}^{-1}}^{2}+\frac{\delta^{k}}{\alpha^{k}}\|w_{0}^{k}-w^{\ast}\|^{2})
\end{align*}
 In particular, this implies 
\begin{align*}
\mathbb{E}\|\overline{w}_{t}-w^{\ast}\|^{2} & \leq C\cdot\sum_{k=1}^{N}p_{k}(1-\alpha^{k})^{t}
\end{align*}
 which is exponential convergence. 
\end{proof}

\end{document}

### 2021-07-04 source-nips/sec_Nesterov.tex
Update on Overleaf.

--- 2020-06-04 source/sec_Nesterov.tex
@@
 FedAvg for convex objectives that match the rates from the previous section.
+Our work the first to provide a unified analysis for FedAvg and Nesterov accelerated FedAvg for convex problems, highlighting the common elements
+and distinctions for the two algorithms, whereas previous analysis of momentum accelerated FedAvg algorithms only exist in the non-convex
+case (only with full participation)~\cite{yu2019linear} and are scattered and use different notation and proof techniques [16, 17].
+
 As we know from stochastic optimization, Nesterov and other momentum

### 2021-07-04 source-nips/sec_noncvx.tex
Update on Overleaf.

% !TEX ROOT=./main.tex

\section{Non-convex convergence rate}

{\centering
\begin{tabular}{l} 
Algorithm 1 MKL-SGD \\
\hline 1: Initialize $\boldsymbol{w}_{\mathbf{0}}$ \\
2: Given samples $D=\left(\boldsymbol{x}_{\boldsymbol{t}}, y_{t}\right)_{t=1}^{\infty}$ \\
3: for $t=1, \ldots$ do \\
4: Choose a set $S_{t}$ of $k$ samples \\
5: Select $i_{t}=\arg \min _{i \in S_{t}} f_{i}\left(\boldsymbol{w}_{t}\right)$ \\
6: Update $\boldsymbol{w}_{t+1}=\boldsymbol{w}_{t}-\eta_t \nabla f_{i_{t}}\left(\boldsymbol{w}_{t}\right)$ \\
7: end for \\
8: Return $\boldsymbol{w}_{t}$ \\
\hline
\end{tabular}}
In the algorithm, we require the step size satisfies
\begin{equation}
\sum_{t=1}^{\infty} \eta_t = \infty \text{ and } \sum_{t=1}^{\infty}\eta_t^2 < \infty
\label{eq:stepsize}
\end{equation}

\subsection{Problem Setup and Notation}
In this subsection, we analyze the convergence rate of MKL-SGD for Non-convex objective. Formally, 
we study the problem of 
 $$\min _{\boldsymbol{w}} F(\boldsymbol{w}),$$
we denote the objective function on clean dataset as $F(\vw) = \frac{1}{\|G\|} \sum_{i \in G} f_i(\vw)$ and
the objective function on noisy dataset $F_a(\vw) =\frac{1}{N} \sum_{i=1}^N f_i(\vw)$, where the
$N$ data point includes both clean data and abnormal data. 
We use $p_i(\vw_t)$ denote the probability that sample $i$ will be used for computing
gradient using MKL-SGD.

\begin{assumption}[Gradient purity]
The expected gradient of MKL-SGD in noisy dataset move in the same direction as the 
expected gradient in clean dataset.
$\langle \grad F(\vw_t),\sum_{i=1}^N p_i(\vw_t)\grad f_i(\vw_t) - \grad F(\vw_t) \rangle \geq 0$
\label{ass:gp}
\end{assumption}

\begin{remark}
	The Assumption~\ref{ass:gp} is relatively weak in terms of gradient norm as it tolerates the
	where noisy gradient is persistent, while it has strict requirement on the noisy gradient moves
	the object toward the correct direction with sufficient distance.
\end{remark}
In the noiseless case, we have $ \sum_{i=1}^N p_i(\vw_t)\grad f_i(\vw_t) = \grad F(\vw_t)$, the 
assumption is satisfied and the results reduce to regular non-convex objective convergence 
results.

\begin{assumption}[L-smooth]
	$F(\mathbf{v}) \leq F(\mathbf{w})+\langle \mathbf{v}-\mathbf{w}, \nabla F(\mathbf{w}) \rangle +\frac{L}{2}\|\mathbf{v}-\mathbf{w}\|_{2}^{2}$.
	\label{ass:lsmooth}
\end{assumption}

\begin{assumption}[Bounded gradient norm]
	The expected square of gradient norm is bounded $\EE \|\grad F_a(\vw_t)\|_2^2  \leq G$: 
	\label{ass:bgn}
\end{assumption}

\begin{theorem}
	With assumption~\ref{ass:gp}, \ref{ass:lsmooth}, \ref{ass:bgn}, the MKL-SGD is running with
	step size $\eta_t$ satisfying \eq{\ref{eq:stepsize}}, then, with $A_T =\sum_{t=0}^T\eta_t $, 
	\begin{equation}
		\EE \frac{1}{A_T}\sum_{t=0}^T \eta_t\|\grad F(\vw_t)\|_2^2 \rightarrow 0
	\end{equation}
\end{theorem}
\begin{proof}
$\vw_{t+1} = \vw_t - \eta_t \grad f_i(\vw_t)$
\begin{align*}
	F(\vw_{t+1}) - F(\vw_{t}) & \leq \langle \grad F(\vw_t), \vw_{t+1} - \vw_t \rangle + \frac{L}{2}\|\vw_{t+1} - \vw_t\|_2^2  \\
                              &  = - \eta_t \langle \grad F(\vw_t), \grad f_i(\vw_t)\rangle  + \frac{L}{2}\|\vw_{t+1} - \vw_t\|_2^2  \\
                              & \leq - \eta_t \langle \grad F(\vw_t), \grad f_i(\vw_t)\rangle  + \frac{\eta_t^2 L}{2}\|\grad f_i(\vw_t)\|_2^2 
\end{align*}
Take expectation w.r.t $i$, condition on $\vw_t$, 
\begin{align*}
	\EE[F(\vw_{t+1})] - F(\vw_{t}) & \leq - \eta_t \langle \grad F(\vw_t),\sum_{i=1}^N p_i(\vw_t)\grad f_i(\vw_t) \rangle + \sum_{i=1}^N p_i(\vw_t) \frac{\eta_t^2 L}{2}\|\grad f_i(\vw_t)\|_2^2 \\
	&  \leq - \eta_t \langle \grad F(\vw_t),\sum_{i=1}^N p_i(\vw_t)\grad f_i(\vw_t) - \grad F(\vw_t) \rangle - \eta_t \|\grad F(\vw_t)\|_2^2 +  \frac{\eta_t^2 LG}{2}\\
\end{align*}
Take total expectation: 
\begin{align*}
	\eta_t\EE \|\grad F(\vw_t)\|_2^2 &\leq \EE[F(\vw_{t})] - \EE[F(\vw_{t+1})] + \EE[- \eta_t \langle \grad F(\vw_t),\sum_{i=1}^N p_i(\vw_t)\grad f_i(\vw_t) - \grad F(\vw_t) \rangle] + \frac{\eta_t^2 LG}{2}
\end{align*}
Summing both sides from $t = 0$ to $T$
\begin{align*}
	\sum_{t=0}^T \eta_t\EE \|\grad F(\vw_t)\|_2^2 &\leq \EE[F(\vw_{1})] - F_{\inf} -\sum_{t=0}^T \eta_t \EE[ \langle \grad F(\vw_t),\sum_{i=1}^N p_i(\vw_t)\grad f_i(\vw_t) - \grad F(\vw_t) \rangle] + \sum_{t=0}^T \frac{\eta_t^2 LG}{2} \\
	& \leq \EE[F(\vw_{1})] - F_{\inf} + \sum_{t=0}^T \frac{\eta_t^2 LG}{2} 
\end{align*}
The learning rate satisfies\eq{\ref{eq:stepsize}}, therefore, $\EE \frac{1}{A_T}\sum_{t=0}^T \eta_t \|\grad F(\vw_t)\|_2^2 \rightarrow 0$

\end{proof}

### 2021-07-04 source-nips/sec_overparameterized.tex
Update on Overleaf.

--- 2020-06-04 source/sec_overparameterized.tex
@@
 We then sharpen this result in the special case of
-quadratic problems.
+linear regression.
 Inspired by recent advances in accelerating SGD~\cite{liu2019accelerating,jain2017accelerating}, 
@@
 new accelerated FedAvg algorithm by adapting the MaSS algorithm of
-\cite{liu2018accelerating} to the federated learning setting. We
+\cite{liu2018accelerating} to the FL setting. We
 show that it achieves geometric convergence for overparameterized
-quadratic problems with rate $\mathcal{O}(\exp(-\frac{NT}{E\sqrt{\kappa_{1}\tilde{\kappa}}}))$,
+linear regression with an $\mathcal{O}(\exp(-\frac{NT}{E\sqrt{\kappa_{1}\tilde{\kappa}}}))$ rate,
 where $\tilde{\kappa}$ is a ``statistical condition number''\cite{liu2018accelerating,jain2017accelerating}
@@
 \subsection{Geometric Convergence of FedAvg in the Overparameterized Setting}
-Recall the federated learning problem $\min_{w}\sum_{k=1}^{N}p_{k}F_{k}(\mathbf{w})$
+Recall the FL problem $\min_{w}\sum_{k=1}^{N}p_{k}F_{k}(\mathbf{w})$
 with $F_{k}(\mathbf{w})=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}\ell(\mathbf{w};\mathbf{x}_{k}^{j})$.
@@
 	and constant step size $\overline{\alpha}=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})})$
-	has geometric convergence
+	has geometric convergence:
 	\begin{align*}
@@
 
-\subsection{Acceleration of FedAvg with Momentum-based Updates}
+\subsection{Overparameterized Linear Regression Problems}
 
-We now turn to quadratic problems and show that the bound in Theorem \ref{thm:overparameterized_general} can be improved to $\mathcal{O}(\exp(-\frac{N}{E\kappa_{1}}t))$.
+We now turn to quadratic problems and show that the bound in Theorem~\ref{thm:overparameterized_general} can be improved to $\mathcal{O}(\exp(-\frac{N}{E\kappa_{1}}t))$ for a larger range of $N$.
 We then propose a variant of FedAvg that has provable acceleration over FedAvg with SGD updates.
@@
 and there exists $\mathbf{w}^{\ast}$ such that $F(\mathbf{w}^{\ast})\equiv0$.
-Two notions of condition number are important in our results:$\kappa_1$ which is based on local Hessians, and $\tilde{\kappa}$, which is termed the statistical condition number~\cite{liu2018accelerating,jain2017accelerating}.
+Two notions of condition number are important in our results: $\kappa_1$ which is based on local Hessians, and $\tilde{\kappa}$, which is termed the statistical condition number~\cite{liu2018accelerating,jain2017accelerating}.
 For their detailed definitions, please refer to Appendix Section~\ref{sec:interpolation}.
@@
 \begin{comment}
-In general $H$ has zero eigevalues. However, because the null space
+In general $H$ has zero eigenvalues. However, because the null space
 of $H$ and range of $H$ are orthogonal, in our subsequence analysis
@@
 \begin{theorem}
-	\label{thm:overparameterized_quadratic}For the overparamterized quadratic
-	problem, FedAvg with communication every $E$
+	\label{thm:overparameterized_quadratic}For the overparamterized linear regression problem, FedAvg with communication every $E$
 	iterations with constant step size $\overline{\alpha}=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})})$
@@
 \begin{theorem}
-	\label{thm:overparameterized_MaSS}For the overparamterized quadratic
-	problem, FedMaSS with communication every $E$ iterations and constant
+	\label{thm:overparameterized_MaSS}For the overparamterized linear regression problem, FedMaSS with communication every $E$ iterations and constant
 	step sizes $\overline{\eta}_1=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+\mu(N-\nu_{\min})}),\overline{\eta}_{2}=\frac{\overline{\eta}_{1}(1-\frac{1}{\tilde{\kappa}})}{1+\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}}},\overline{\gamma}=\frac{1-\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}}}{1+\frac{1}{\sqrt{\kappa_{1}\tilde{\kappa}}}}$

### 2021-07-04 source/app_sgd_cvx_smooth.tex
update to vector notations
v_{t+1}^{k} -> \vv_{t+1}^{k}, v_{0}->\vv_{0}, |w_{t+1}^{k} -> |\vw_{t+1}^{k}

g_{i,k}-> \vg_{i,k}

v_{i+1}^{k}  \vv_{i+1}^{k}
Update app_sgd_cvx_smooth.tex
minor
Update on Overleaf.

@@
-\begin{thm}
-	Under assumptions~\ref{ass:lsmooth},\ref{ass:boundedvariance},\ref{ass:subgrad2} and constant learning
-	rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$, 
+% !TEX ROOT=./supp-main.tex
+In this section we provide the proof of the convergence result for FedAvg with convex and smooth objectives.
+The key step is a one step progress result analogous to that in the strongly convex case, and their proofs share identical components as well.
+\begin{lemma} [\textbf{One step progress, convex case}]
+Let $\overline{\mathbf{w}}_{t}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{t}^{k}$ in FedAvg.
+Under assumptions~\ref{ass:lsmooth},\ref{ass:boundedvariance},\ref{ass:subgrad2}, the following bound holds for all $t$:
 \begin{align*}
-	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T}\right)
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq \mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
 	\end{align*}
-	with full participation, and with partial device participation with $K$ sampled devices at
-	each communication round and learning rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{K}{T}})$,
+	\label{lem:cvxoner}
+\end{lemma}
+\begin{proof}
+    The first part of the proof follows directly from \eq{\ref{eq:common one step}} in the proof of Lemma \ref{lem:scvxoner}.
+Setting $\mu=0$ in \eq{\ref{eq:common one step}} (since we are in the convex setting instead of strongly convex), we obtain
     \begin{align*}
-	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right)
+	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} \\ 
+	& +2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
 	\end{align*}
-\end{thm}
-
-\begin{proof}
-	We again start by bounding the term
-	\begin{align*}
-	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}\vg_{t})-\vw^{\ast}\|^{2}\\
-	& =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(\vg_{t}-\ov{g}_{t})\|^{2}\\
-	& =A_{1}+A_{2}+A_{3}
-	\end{align*}
-	where 
-	\begin{align*}
-	A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}\\
-	A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-\vg_{t}\rangle\\
-	A_{3} & =\alpha_{t}^{2}\|\vg_{t}-\ov{g}_{t}\|^{2}
-	\end{align*}
-	$\mathbb{E}A_{2}=0$ by definition of $\vg_{t}$ and $\ov{g}_{t}$,
-	while for $A_{3}$ we have
-	\begin{align*}
-	\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|g_{t,k}-\mathbb{E}g_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
-	\end{align*}
-	again by Jensen's inequality and using the independence of $g_{t,k},g_{t,k'}$.
-
-	Next we bound $A_{1}$: 
-	\begin{align*}
-	\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2}
-	\end{align*}
-	Using the convexity and $L$-smoothness of $F_{k}$, 
-	\begin{align*}
-	& -2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle w_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))\\
-	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]\\
-	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]
-	\end{align*}
-	which results in 
-	\begin{align*}
-	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
-	\end{align*}
-
 	The difference of this bound with that in the strongly convex case
-	is that we no longer have a contraction factor in front of $\|\ov{w}_{t}-\vw^{\ast}\|^{2}$.
+	is that we no longer have a contraction factor of $1-\mu\alpha_t$ in front of $\|\ov{w}_{t}-\vw^{\ast}\|^{2}$.
 	In the strongly convex case, we were able to cancel $\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$
@@
 	& \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
+	\end{align*}
+\end{proof}
+With the one step progress result, we can now prove the convergence result in the convex setting, which we restate below.
+\begin{thm}
+	Under assumptions~\ref{ass:lsmooth},\ref{ass:boundedvariance},\ref{ass:subgrad2} and constant learning
+	rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$, FedAvg satisfies
+	\begin{align*}
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T}\right)
+	\end{align*}
+	with full participation, and with partial device participation with $K$ sampled devices at
+	each communication round and learning rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{K}{T}})$,
+	\begin{align*}
+	\min_{t\leq T}F(\overline{\mathbf{w}}_{t})-F(\mathbf{w}^{\ast}) & =\mathcal{O}\left(\frac{\nu_{\max}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right)
+	\end{align*}
+\end{thm}
+
+\begin{proof}
+	We first prove the bound for full participation. Applying Lemma~\ref{lem:cvxoner}, we have
+	\begin{align*}
+	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
 	\end{align*}
@@
 & =(\|\vw_{0}-\vw^{\ast}\|^{2}+\nu_{\max}^{2}\sigma^{2})\frac{1}{\sqrt{NT}}+\frac{N}{T}6E^{2}LG^{2}\\
-& =O(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T})
+& =\mathcal{O}(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T})
 \end{align*}
 
-	Similarly, for partial participation, we have 
+	{\color{red}For partial participation, the one step progress bound in Lemma \ref{lem:cvxoner} is updated in a similar manner as the strongly convex case in \eqref{eqn:scvx-partial-oner} to incorporate the sampling variance.
+More precisely, with partial participation,
 	\begin{align*}
-	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|\vw_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot(\frac{1}{N}\nu_{\max}\sigma^{2}+C)+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right)
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}+\ov{v}_{t+1}-\vw^{\ast}\|^{2}\\
+	& =\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2}+\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2},
 	\end{align*}
-	where $C=\frac{4}{K}E^{2}G^{2}$ or $\frac{N-K}{N-1}\frac{4}{K}E^{2}G^{2}$,
+	where $\mathbb{E}\ov{w}_{t+1}=\ov{v}_{t+1}$ for all $t$, by the unbiasedness of our sampling schemes.
+Since $\ov{v}_t =\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$ always averages over all devices, the full participation one step progress bound in Lemma~\ref{lem:cvxoner} applied to $\ov{v}_t$ implies
+	\begin{align*}
+	\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{v}_{t})-F(\vw^{\ast})) & \leq\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}\\ &
+	   \leq\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
+	\end{align*}
+
+	The bound for $\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2}$ for the two sampling schemes we consider is provided in \eq{\ref{eq:partialsample}}, and applying it to the above bound we can write the one step progress for partial participation as
+	\begin{align*}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} + \alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}(\frac{1}{N}\nu_{max}^{2}\sigma^{2}+C)+6E^{2}L\alpha_{t}^{3}G^{2},
+	\end{align*}
+	where $C=\frac{4}{K}E^{2}G^{2}$ or $\frac{N-K}{N-1}\frac{4}{K}E^{2}G^{2}$ depending on the sampling scheme.
+	}
+
+	Summing up the one-step progress over $t$, 
+	\begin{align*}
+	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|\vw_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot(\frac{1}{N}\nu_{\max}\sigma^{2}+C)+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right),
+	\end{align*}
 	so that with $\alpha_{t}=\sqrt{\frac{K}{T}}$, we have 
 	\begin{align*}
-	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & =\mathcal{O}(\frac{\nu_{\max}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T})
+	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & =\mathcal{O}(\frac{\nu_{\max}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}).
 	\end{align*}

### 2021-07-04 source/app_sgd_scvx_smooth.tex
Update on Overleaf.

@@
-% !TEX ROOT=./supp-main.tex
-To facilitate reading, theorems from the main paper are restated and numbered identically. We first summarize some properties of $L$-smooth and $\mu$-strongly
-convex functions~\cite{rockafellar1970convex}. 
-\begin{lemma}
-	Let $F$ be a convex $L$-smooth function. Then we have the following
-	inequalities:
+% !TEX ROOT=./main.tex
 
-	1. Quadratic upper bound: $0\leq F(\vw)-F(\vw')-\langle\nabla F(\vw'),\vw-\vw'\rangle\leq\frac{L}{2}\|\vw-\vw'\|^{2}$.
-
-	2. Coercivity: $\frac{1}{L}\|\nabla F(\vw)-\nabla F(\vw')\|^{2}\leq\langle\nabla F(\vw)-\nabla F(\vw'),\vw-\vw'\rangle$.
-
-	3. Lower bound: $F(\vw)\geq F(\vw')+\langle\nabla F(\vw'),\vw-\vw'\rangle+\frac{1}{2L}\|\nabla F(\vw)-\nabla F(\vw')\|^{2}$.
-	In particular, $\|\nabla F(\vw)\|^{2}\leq2L(F(\vw)-F(\vw^{\ast}))$.
-
-	4. Optimality gap: $F(\vw)-F(\vw^{\ast})\leq$$\langle\nabla F(\vw),\vw-\vw^{\ast}\rangle$.
+To organize our proofs more effectively and highlight the significance of our results compared to prior works, 
+we first state the following key lemmas used in proofs of main results and defer their proofs to later. 
+\begin{lemma}[\textbf{One step progress, strongly convex}] Let $\overline{\mathbf{w}}_{t}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{t}^{k}$, and
+suppose our functions satisfy Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2}, and set step size $\alpha_{t}=\frac{4}{\mu(\gamma+t)}$
+	with $\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$, then the updates of FedAvg with full participation satisfy
+	\begin{align*}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}\\
+	+& \alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}L\alpha_{t}^{3}G^{2}.
+	\end{align*}
+\label{lem:scvxoner}
 \end{lemma}
-%
-\begin{lemma}
-	Let $F$ be a $\mu$-strongly convex function. Then 
+We emphasize that the above lemma is the key step that allows us to obtain a bound that improves on the convergence result of~\cite{li2019convergence} with linear speedup.
+Its proof will make use of the following two results.
+\begin{lemma}[\textbf{Bounding gradient variance (Lemma 2 \cite{li2019convergence})} ]
+Given Assumption~\ref{ass:boundedvariance}, the upper bound of gradient variance is given as follows,
 \begin{align*}
-	F(\vw) & \leq F(\vw')+\langle\nabla F(\vw'),\vw-\vw'\rangle+\frac{1}{2\mu}\|\nabla F(\vw)-\nabla F(\vw')\|^{2}\\
-	F(\vw)-F(\vw^{\ast}) & \leq\frac{1}{2\mu}\|\nabla F(\vw)\|^{2}
+	\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} \leq \sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}.
 	\end{align*}
+\label{lem:bgv}
 \end{lemma}
 
+\begin{lemma}[\textbf{Bounding the divergence of $\vw_t^k$ (Lemma 3 \cite{li2019convergence})} ]
+Given Assumption~\ref{ass:subgrad2}, and assume that $\alpha_t$ is non-increasing and $\alpha_t \leq 2\alpha_{t+E}$ for all $t\geq 0$, we have
+	\begin{align*}
+	\mathbb{E}\left[\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} \right]\leq 4E^{2}\alpha_{t}^{2}G^{2}.
+	\end{align*}
+\label{lem:bdw}
+\end{lemma}
+
+We now restate Theorem~\ref{th:scvx_sgd} from the main text and then prove it using Lemma~\ref{lem:scvxoner}.
 \begin{thm}
-	Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$,
-	$\nu_{\max}=\max_{k}Np_{k}$, and set decaying learning rates $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$
+	Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$ in FedAvg,
+	$\nu_{\max}=\max_{k}Np_{k}$, and set decaying learning rates $\alpha_{t}=\frac{4}{\mu(\gamma+t)}$
 	with $\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$. Then
@@
 \begin{proof}
-	The first step is to observe that the $L$-smoothness of $F$ provides
-	the upper bound
+	% The proof builds on ideas from~\cite{li2019convergence}. The first step is to observe that the $L$-smoothness of $F$ provides
+	% the upper bound
+	% \begin{align*}
+	% \mathbb{E}(F(\ov{w}_{t}))-F^{\ast} & \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}
+	% \end{align*}
+	% and bound $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$. 
+	The road map of the proof for full device participation contains three steps. First, we establish a recursive relationship between $\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}$
+	and $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$, upper bounding the progress of FedAvg from step $t$ to step $t+1$. 
+	Second, we show that $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}=\mathcal{O}(\frac{\nu_{max}^{2}\sigma^{2}/\mu}{tN}+\frac{E^{2}LG^{2}/\mu^2}{t^{2}})$ by induction using the recursive relationship from the previous step.
+	Third, we use the property of $L$-smoothness to bound the optimality gap by $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$.
+
+	By Lemma~\ref{lem:scvxoner}, we have the following upper bound for the one step progress: 
 	\begin{align*}
-	\mathbb{E}(F(\ov{w}_{t}))-F^{\ast} & =\mathbb{E}(F(\ov{w}_{t})-F(\vw^{\ast}))\\
-	& \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}\\
+	& +\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}L\alpha_{t}^{3}G^{2}.
 	\end{align*}
-	and bound $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$.
-
-	Our main step is to prove the bound 
+	We show next that $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}=O(\frac{\nu_{max}^{2}\sigma^{2}/\mu}{tN}+\frac{E^{2}LG^{2}/\mu^2}{t^{2}})$ using induction. 
+	To simplify the presentation, we denote $C\equiv6E^{2}LG^{2}$ and $D\equiv\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
+	Suppose that we have the bound $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}\leq b\cdot(\alpha_{t}D+\alpha_{t}^{2}C)$
+	for some constant $b$ and learning rates $\alpha_{t}$. Then the one step progress from Lemma~\ref{lem:scvxoner} becomes:
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+5E^{2}L\alpha_{t}^{3}G^{2}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq (b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}D\\
+	& +(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}C
 	\end{align*}
-
-	We have 
+	To establish the result at step $t+1$, it remains to choose $\alpha_{t}$ and $b$ such that $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}\leq b\alpha_{t+1}$
+	and $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}\leq b\alpha_{t+1}^{2}$.
+	%Recall that we require $\alpha_{t_{0}}\leq2\alpha_{t}$ for any $t-t_{0}\leq E-1$, and $L\alpha_{t}\leq\frac{1}{8}$. 
+	If we let $\alpha_{t}=\frac{4}{\mu(t+\gamma)}$
+	where $\gamma=\max\{E,32\kappa\}$ (choice of $\gamma$ required to guarantee the one step progress) and set $b=\frac{4}{\mu}$, we have:
 	\begin{align*}
-	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & =\|(\ov{w}_{t}-\alpha_{t}\vg_{t})-\vw^{\ast}\|^{2}\\
-	& =\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(\vg_{t}-\ov{g}_{t})\|^{2}\\
-	& =A_{1}+A_{2}+A_{3}
-	\end{align*}
-	where 
-	\begin{align*}
-	A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}\\
-	A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-\vg_{t}\rangle\\
-	A_{3} & =\alpha_{t}^{2}\|\vg_{t}-\ov{g}_{t}\|^{2}
-	\end{align*}
-	By definition of $\vg_{t}$ and $\ov{g}_{t}$ (see \eq{\ref{eq:gradient}}), we have $\mathbb{E}A_{2}=0$.
-	For $A_{3}$, we have the follow upper bound:
-	\begin{align*}
-	\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\|\vg_{t,k}-\mathbb{E}\vg_{t,k}\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
-	\end{align*}
-	again by Jensen's inequality and using the independence of $\vg_{t,k},\vg_{t,k'}$ \cite[Lemma 2]{li2019convergence}.
-
-	Next we bound $A_{1}$: 
-	\begin{align*}
-	\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} & =\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2}
-	\end{align*}
-	and we will show that the third term $\|\alpha_{t}\ov{g}_{t}\|^{2}$
-	can be canceled by an upper bound of the second term. %
-	\begin{comment}
-	The last term is straightforward to bound by the convexity of $\|\cdot\|^{2}$
-	and $L$-smoothness of $F_{k}$,
-	\begin{align*}
-	\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq2L\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}(F_{k}(\vw_{t}^{k})-F_{k}^{\ast})
-	\end{align*}
-	or 
-	\begin{align*}
-	\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\mathbb{E}\|\nabla F_{k}(\vw_{t}^{k},\mathbf{\xi}_{t}^{k})\|^{2}\leq\alpha_{t}^{2}G^{2}
-	\end{align*}
-	\end{comment}
-
-	Now 
-	\begin{align*}
-	& -2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	& =-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle \vw_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
-	& \leq-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle+2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
-	& \leq2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
-	& =\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]-\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
-	\end{align*}
-	For the second term, which is negative, we can ignore it, but this
-	yields a suboptimal bound that fails to provide the desired linear
-	speedup. Instead, we upper bound it using the following derivation:
-	\begin{align*}
-	& 2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]\\
-	& \leq2\alpha_{t}\left[F(\ov{w}_{t+1})-F(\ov{w}_{t})\right]\\
-	& \leq2\alpha_{t}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{w}_{t+1}-\ov{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\ov{w}_{t+1}-\ov{w}_{t}\|^{2}\\
-	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\vg_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& =-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\ov{g}_{t}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& =-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\sum_{k}p_{k}\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\ov{w}_{t})-\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	& \leq-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
-	\end{align*}
-	where we have used the smoothness of $F$ twice.
-
-	Note that the term $-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$ exactly
-	cancels the $\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$ in the bound
-	for $A_{1}$, so that plugging in the bound for $-2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle$,
-	we have so far proved 
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
-	\end{align*}
-	The term $\mathbb{E}\|\vg_{t}\|^{2}\leq G^{2}$ by assumption.
-
-	Now we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$.
-	Since communication is done every $E$ steps, for any $t\geq0$, we
-	can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $\vw_{t_{0}}^{k}=\ov{w}_{t_{0}}$for
-	all $k$. Moreover, using $\alpha_{t}$ is non-increasing and $\alpha_{t_{0}}\leq2\alpha{}_{t}$
-	for any $t-t_{0}\leq E-1$, we have 
-	\begin{align*}
-	& \mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
-	& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
-	& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}\vg_{i,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}E\alpha_{i}^{2}\|\vg_{i,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}E^{2}\alpha_{t_{0}}^{2}G^{2}\\
-	& \leq4E^{2}\alpha_{t}^{2}G^{2}
-	\end{align*}
-	\begin{comment}
-	Alternatively, we can also write
-	\begin{align*}
-	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & =\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\alpha_{t-1}\vg_{t-1}-\vw_{t-1}^{k}+\alpha_{t-1}\vg_{t-1,k}\|^{2}\\
-	& \leq2\sum_{k=1}^{N}p_{k}\left(\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\alpha_{t-1}\vg_{t-1}-\alpha_{t-1}\vg_{t-1,k}\|^{2}\right)
-	\end{align*}
-	and write 
-	\begin{align*}
-	\sum_{k}p_{k}\|\vg_{t-1,k}-\vg_{t-1}\|^{2}\leq\sum_{k}p_{k}\|\vg_{t-1,k}\|^{2} & =\sum_{k}p_{k}\|\vg_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})+\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
-	& \leq2\sum_{k}p_{k}\|\vg_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})\|^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
-	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})-\nabla F_{k}(\ov{w}_{t-1})+\nabla F_{k}(\ov{w}_{t-1})\|^{2}\\
-	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+4\sum_{k}p_{k}\left(\|\nabla F_{k}(\vw_{t-1}^{k})-\nabla F_{k}(\ov{w}_{t-1})\|^{2}+\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}\right)\\
-	& \leq2\sigma^{2}+4\sum_{k}p_{k}\left(L^{2}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}\right)
-	\end{align*}
-	so that 
-	\begin{align*}
-	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+4\sigma^{2}\alpha_{t-1}^{2}+8\alpha_{t-1}^{2}\sum_{k}p_{k}\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}
-	\end{align*}
-	and taking expectation, we have 
-	\begin{align*}
-	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+4\alpha_{t-1}^{2}(\sigma^{2}+2G^{2})
-	\end{align*}
-	Using this recusive relation at most $E$ times to arrive at a communication
-	round and that $\alpha_{t}L\leq1/8$ for all $t$, we have 
-	\begin{align*}
-	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\leq E\frac{1}{L^{2}}(\sigma^{2}+2G^{2})
-	\end{align*}
-	which gives 
-	\end{comment}
-	{}
-
-	Using the bound on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$,
-	we can conclude that, with $\nu_{max}:=N\cdot\max_{k}p_{k}$ and $\nu_{min}:=N\cdot\min_{k}p_{k}$,
-
-	\begin{align*}
-	& \mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& =\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\sum_{k=1}^{N}(p_{k}N)^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\nu_{max}^{2}\sum_{k=1}^{N}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
-	\end{align*}
-	where in the last inequality we use $\sigma^{2}=\max_{k}\sigma_{k}^{2}$, and assume $\alpha_{t}$
-	satisfies $L\alpha_{t}\leq\frac{1}{8}$. We show next that $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}=O(\frac{1}{tN}+\frac{E^{2}LG^{2}}{t^{2}})$.
-
-	Let $C\equiv6E^{2}LG^{2}$ and $D\equiv\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
-	Suppose that we have shown $\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}\leq b\cdot(\alpha_{t}D+\alpha_{t}^{2}C)$
-	for some constant $b$ and $\alpha_{t}$. Then 
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq b(1-\mu\alpha_{t})(\alpha_{t}D+\alpha_{t}^{2}C)+\alpha_{t}^{2}D+\alpha_{t}^{3}C\\
-	& =(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}D+(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}C
-	\end{align*}
-	and so it remains to choose $\alpha_{t}$ and $b$ such that $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}\leq b\alpha_{t+1}$
-	and $(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2}\leq b\alpha_{t+1}^{2}$.
-	Recall that we require $\alpha_{t_{0}}\leq2\alpha_{t}$ for any $t-t_{0}\leq E-1$,
-	and $L\alpha_{t}\leq\frac{1}{8}$. If we let $\alpha_{t}=\frac{4}{\mu(t+\gamma)}$
-	where $\gamma=\max\{E,32\kappa\}$, then we may check that $\alpha_{t}$
-	satisfies both requirements.
-
-	Setting $b=\frac{4}{\mu}$, we have 
-	\begin{align*}
-	(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t} & =\left(b(1-\frac{4}{t+\gamma})+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}\\
-	& =\left(b\frac{t+\gamma-4}{t+\gamma}+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}\\
-	& =b(\frac{t+\gamma-3}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
-	& \leq b(\frac{t+\gamma-1}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
-	& \leq b\frac{4}{\mu(t+\gamma+1)}=b\alpha_{t+1}
-	\end{align*}
-	and 
-	\begin{align*}
-	(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2} & =\left(b(1-\frac{4}{t+\gamma})+\frac{4}{\mu(t+\gamma)}\right)\frac{16}{\mu^{2}(t+\gamma)^{2}}\\
-	& =\left(b\frac{t+\gamma-4}{t+\gamma}+\frac{4}{\mu(t+\gamma)}\right)\frac{16}{\mu^{2}(t+\gamma)^{2}}\\
-	& =b(\frac{t+\gamma-2}{t+\gamma})\frac{16}{\mu^{2}(t+\gamma)^{2}}\\
+	(b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}  &=\left(b(1-\frac{4}{t+\gamma})+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}\\
+	% & =\left(b\frac{t+\gamma-4}{t+\gamma}+\frac{4}{\mu(t+\gamma)}\right)\frac{4}{\mu(t+\gamma)}\\
+	% & =b(\frac{t+\gamma-3}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
+	% & \leq b(\frac{t+\gamma-1}{t+\gamma})\frac{4}{\mu(t+\gamma)}\\
+	 &\leq b\frac{4}{\mu(t+\gamma+1)}=b\alpha_{t+1}\\
+	 (b(1-\mu\alpha_{t})+\alpha_{t})\alpha_{t}^{2} &=b(\frac{t+\gamma-2}{t+\gamma})\frac{16}{\mu^{2}(t+\gamma)^{2}} \\
 	 & \leq b\frac{16}{\mu^{2}(t+\gamma+1)^{2}}=b\alpha_{t+1}^{2}
 	\end{align*}
-	where we have used the facts that 
+	where we have used the following inequalities:
 	\begin{align*}
-	\frac{t+\gamma-1}{(t+\gamma)^{2}} & \leq\frac{1}{(t+\gamma+1)}\\
-	\frac{t+\gamma-2}{(t+\gamma)^{3}} & \leq\frac{1}{(t+\gamma+1)^{2}}
+	\frac{t+\gamma-1}{(t+\gamma)^{2}} \leq\frac{1}{(t+\gamma+1)},\hspace{0.2em} 
+	\frac{t+\gamma-2}{(t+\gamma)^{3}} \leq\frac{1}{(t+\gamma+1)^{2}}, \hspace{0.2em}  \forall\ \gamma\geq1
 	\end{align*}
-	for $\gamma\geq1$.
-
-	Thus we have shown 
+	Thus we have established the result at step $t+1$ assuming the result is correct at step $t$:
 	\begin{align*}
@@
 	\end{align*}
-	for our choice of $\alpha_{t}$ and $b$. Now to ensure 
+	% for our choice of $\alpha_{t}$ and $b$. Now to ensure 
+	At step $t=0$, we can ensure the following inequality by scaling $b$ with $c\|\vw_{0}-\vw^{\ast}\|^{2}$ for a sufficiently large constant
+	$c$:
 	\begin{align*}
-	\|\vw_{0}-\vw^{\ast}\|^{2} & \leq b\cdot(\alpha_{0}D+\alpha_{0}^{2}C)\\
-	& =b\cdot(\frac{4}{\mu\gamma}D+\frac{16}{\mu^{2}\gamma^{2}}C)
+	\|\vw_{0}-\vw^{\ast}\|^{2} \leq b\cdot(\alpha_{0}D+\alpha_{0}^{2}C) =b\cdot(\frac{4}{\mu\gamma}D+\frac{16}{\mu^{2}\gamma^{2}}C)
 	\end{align*}
-	we can simply scale $b$ by $c\|\vw_{0}-\vw^{\ast}\|^{2}$ for a constant
-	$c$ large enough and the induction step still holds.
-
 	It follows that 
-	\begin{align*}
+	\begin{align}
 	\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2} & \leq c\|\vw_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{t}+C\alpha_{t}^{2})
-	\end{align*}
+	\label{eq:th1:1}
+	\end{align}
 	for all $t\geq0$.
@@
 	\begin{align*}
-	& \mathbb{E}(F(\ov{w}_{T}))-F^{\ast}\\
-	& =\mathbb{E}(F(\ov{w}_{T})-F(\vw^{\ast}))\\
-	& \leq\frac{L}{2}\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}\leq\frac{L}{2}c\|\vw_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
-	& =2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
-	& \leq2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{4}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}LG^{2}\cdot(\frac{4}{\mu(T+\gamma)})^{2}\right]\\
-	& =O(\frac{\kappa}{\mu}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
+	& \mathbb{E}(F(\ov{w}_{T}))-F^{\ast} \leq \frac{L}{2}\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}\\ 
+	\leq &\frac{L}{2}c\|\vw_{0}-\vw^{\ast}\|^{2}\frac{4}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
+	=& 2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
+	\leq &2c\|\vw_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{4}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}LG^{2}\cdot(\frac{4}{\mu(T+\gamma)})^{2}\right]\\
+	=& \mathcal{O}(\frac{\kappa}{\mu}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
 	\end{align*}
+	where in the first line, we use the property of $L$-smooth function (see Lemma~\ref{lem:lsmooth}), and in the second line, we use the conclusion in \eq{\ref{eq:th1:1}}.
 
 	With partial participation, the update at each communication round
-	is now given by averages over a subset of sampled devices. When $t+1\notin\mathcal{I}_{E}$,
-	$\ov{v}_{t+1}=\ov{w}_{t+1}$, while when $t+1\notin\mathcal{I}_{E}$,
+	is now given by weighted averages over a subset of sampled devices. When $t+1\notin\mathcal{I}_{E}$,
+	$\ov{v}_{t+1}=\ov{w}_{t+1}$, while when $t+1\in\mathcal{I}_{E}$,
 	we have $\mathbb{E}\ov{w}_{t+1}=\ov{v}_{t+1}$ by design
-	of the sampling schemes, so that 
+	of the sampling schemes~(\cite{li2019convergence}, Lemma 4), so that 
 	\begin{align*}
@@
 	\end{align*}
-	As before, $\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}\leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}$.
+	This in particular implies $\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}\leq \mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}$ for all $t$.
+	Since $\ov{v}_t =\sum_{k=1}^{N} p_{k} \mathbf{v}_{t}^{k}$ always averages over all devices, the full participation one step progress result Lemma~\ref{lem:scvxoner} applied to $\ov{v}_t$  implies
+	\begin{align*}
+	   \mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{v}_{t}-\vw^{\ast}\|^{2}\\
+	   & +6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\\ &
+	   \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}\\
+	   & +6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
+	\end{align*}
 
-	The key is to bound $\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2}$.
-	For sampling scheme I we have 
+	The bound for $\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2}$ for the two sampling schemes we consider is provided in \eq{\ref{eq:partialsample}}, and applying it we can write the one step progress for partial participation as
+	\begin{align}\label{eqn:scvx-partial-oner}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}\nonumber\\
+	&+\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}+6E^{2}L\alpha_{t}^{3}G^{2},
+	\end{align}
+	and the same arguments using induction and $L$-smoothness as the full participation case implies 
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & =\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|\vw_{t+1}^{k}-\ov{w}_{t+1}\|^{2}\\
-	& \leq\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
-	\end{align*}
-	while for sampling scheme II 
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\ov{v}_{t+1}\|^{2} & =\frac{N-K}{N-1}\frac{1}{K}\sum_{k}p_{k}\mathbb{E}\|\vw_{t+1}^{k}-\ov{w}_{t+1}\|^{2}\\
-	& \leq\frac{N-K}{N-1}\frac{4}{K}\alpha_{t}^{2}E^{2}G^{2}
-	\end{align*}
-	The same argument as the full participation case implies 
-	\begin{align*}
-	\mathbb{E}F(\ov{w}_{T})-F^{\ast}=O(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
+	\mathbb{E}F(\ov{w}_{T})-F^{\ast}=\mathcal{O}(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}
+	+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}})
 	\end{align*}
@@
 
+\subsubsection{Deferred proofs of key lemmas}
+Here we first rewrite the proofs of lemmas \ref{lem:bgv} and  \ref{lem:bdw} from~\cite{li2019convergence} with slight modifications for the consistency and completeness of this work, since later we will use modified versions of these results in the convergence proof for Nesterov accelerated FedAvg.
+\begin{proof}[Proof of lemma~\ref{lem:bgv}]
+	\begin{align*}
+	\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} & =\mathbb{E}\|\vg_{t}-\mathbb{E}\vg_{t}\|^{2}=\sum_{k=1}^{N}p_{k}^{2}\|\vg_{t,k}-\mathbb{E}\vg_{t,k}\|^{2}\leq \sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+	\end{align*}
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:bdw}]
+	Now we bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$ following \cite{li2019convergence}.
+	Since communication is done every $E$ steps, for any $t\geq0$, we
+	can find a $t_{0}\leq t$ such that $t-t_{0}\leq E-1$ and $\vw_{t_{0}}^{k}=\ov{w}_{t_{0}}$for
+	all $k$. Moreover, using $\alpha_{t}$ is non-increasing and $\alpha_{t_{0}}\leq2\alpha{}_{t}$
+	for any $t-t_{0}\leq E-1$, we have 
+		\begin{align*}
+	& \mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}
+= \mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
+&\leq \mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2}
+= \mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
+	&= \mathbb{E}\sum_{k=1}^{N}p_{k}\|-\sum_{i=t_{0}}^{t-1}\alpha_{i}\vg_{i,k}\|^{2}
+	\leq  2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}E\alpha_{i}^{2}\|\vg_{i,k}\|^{2}\\
+	&\leq  2\sum_{k=1}^{N}p_{k}E^{2}\alpha_{t_{0}}^{2}G^{2}
+	\leq  4E^{2}\alpha_{t}^{2}G^{2}
+	\end{align*}
+\end{proof}
+Based on the results of Lemma~\ref{lem:bgv},~\ref{lem:bdw}, we now prove the upper bound of one step SGD progress. This 
+proof improves on the previous work~\cite{li2019convergence} and is the first to reveal the linear speedup of convergence of FedAvg. 
+\begin{proof}[Proof of lemma~\ref{lem:scvxoner}]
+	We have 
+	\begin{align*}
+	&\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}  =\|(\ov{w}_{t}-\alpha_{t}\vg_{t})-\vw^{\ast}\|^{2}\\ &=\|(\ov{w}_{t}-\alpha_{t}\ov{g}_{t}-\vw^{\ast})-\alpha_{t}(\vg_{t}-\ov{g}_{t})\|^{2} = \underbrace{\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}}_{A_1} \\ &+\underbrace{2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-\vg_{t}\rangle}_{A_2} + \underbrace{\alpha_{t}^{2}\|\vg_{t}-\ov{g}_{t}\|^{2}}_{A_3}
+	\end{align*}
+	where we denote: 
+	\begin{align*}
+	A_{1} & =\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2}\\
+	A_{2} & =2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t},\ov{g}_{t}-\vg_{t}\rangle\\
+	A_{3} & =\alpha_{t}^{2}\|\vg_{t}-\ov{g}_{t}\|^{2}
+	\end{align*}
+	By definition of $\vg_{t}$ and $\ov{g}_{t}$ (see \eq{\ref{eq:gradient}}), we have $\mathbb{E}A_{2}=0$.
+	For $A_{3}$, we have the following upper bound (see Lemma~\ref{lem:bgv}):
+	\begin{align*}
+	\alpha_{t}^{2}\mathbb{E}\|\vg_{t}-\ov{g}_{t}\|^{2} \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+	\end{align*}
+
+	Next we bound $A_{1}$: 
+	\begin{align*}
+	&\|\ov{w}_{t}-\vw^{\ast}-\alpha_{t}\ov{g}_{t}\|^{2} \\ &=\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\langle\ov{w}_{t}-\vw^{\ast},-\alpha_{t}\ov{g}_{t}\rangle+\|\alpha_{t}\ov{g}_{t}\|^{2}
+	\end{align*}
+	and we will show that the third term $\|\alpha_{t}\ov{g}_{t}\|^{2}$
+	can be canceled by an upper bound of the second term, which is one of our major improvements compared to prior works~\cite{li2019convergence}.
+	\begin{comment}
+	The last term is straightforward to bound by the convexity of $\|\cdot\|^{2}$
+	and $L$-smoothness of $F_{k}$,
+	\begin{align*}
+	\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq2L\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}(F_{k}(\vw_{t}^{k})-F_{k}^{\ast})
+	\end{align*}
+	or 
+	\begin{align*}
+	\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} & \leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\|\nabla F_{k}(\vw_{t}^{k})\|^{2}\leq\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}\mathbb{E}\|\nabla F_{k}(\vw_{t}^{k},\mathbf{\xi}_{t}^{k})\|^{2}\leq\alpha_{t}^{2}G^{2}
+	\end{align*}
+	\end{comment}
+	The upper bound of second term can be derived as follows, 	using the strong convexity and $L$-smoothness of $F_{k}$:
+	\begin{align*}
+	 -&2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle= -2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	= &-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	-&2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle \vw_{t}^{k}-\vw^{\ast},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	\leq&-2\alpha_{t}\sum_{k=1}^{N}p_{k}\langle\ov{w}_{t}-\vw_{t}^{k},\nabla F_{k}(\vw_{t}^{k})\rangle\\
+	+&2\alpha_{t}\sum_{k=1}^{N}p_{k}(F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k}))-\alpha_{t}\mu\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
+	\leq & 2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw_{t}^{k})-F_{k}(\ov{w}_{t})+\frac{L}{2}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+F_{k}(\vw^{\ast})-F_{k}(\vw_{t}^{k})\right]-\alpha_{t}\mu\|\sum_{k=1}^{N}p_{k}\vw_{t}^{k}-\vw^{\ast}\|^{2}\\
+	= &\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]\\
+	-&\alpha_{t}\mu\|\ov{w}_{t}-\vw^{\ast}\|^{2}
+	\end{align*}
+	We record the bound we have obtained so far, as it will also be used in the proof for convex case: 
+	\begin{align*}
+	 &\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}
+	\leq  \mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\\
+	 +&2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{2}\|\ov{g}_{t}\|^{2} \numberthis \label{eq:common one step}
+	\end{align*}
+	For the term $2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]$, which is negative, we can ignore it, but this
+	yields a suboptimal bound that fails to provide the desired linear
+	speedup. Instead, we upper bound it using the following derivation:
+	{\small
+	\begin{align*}
+	& 2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]\\
+	\leq& 2\alpha_{t}\left[F(\ov{w}_{t+1})-F(\ov{w}_{t})\right]\\
+	\leq& 2\alpha_{t}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{w}_{t+1}-\ov{w}_{t}\rangle+\alpha_{t}L\mathbb{E}\|\ov{w}_{t+1}-\ov{w}_{t}\|^{2}\\
+	=&-2\alpha_{t}^{2}\mathbb{E}\langle\nabla F(\ov{w}_{t}),\ov{g}_{t}\rangle+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	=&-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\ov{g}_t\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	\leq&-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\ov{w}_{t})-\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	\leq&-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
+	\leq&-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
+	\end{align*}}
+	where we have used the smoothness of $F$ twice.
+
+	Note that the term $-\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$ exactly
+	cancels the $\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}$ in the bound in \eq{\ref{eq:common one step}}, so that plugging in the bound for $-2\alpha_{t}\langle\ov{w}_{t}-\vw^{\ast},\ov{g}_{t}\rangle$,
+	we have so far proved 
+	\begin{align*}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2} \numberthis \label{eq:common recursion}
+	\end{align*}
+	Under Assumption~\ref{ass:subgrad2}, we have $\mathbb{E}\|\vg_{t}\|^{2}\leq G^{2}$.
+Furthermore, we can check that our choice of $\alpha_t$ satisfies $\alpha_t$ is non-increasing and $\alpha_t \leq 2\alpha_{t+E}$, so we may plug in the bound $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} \leq 4E^{2}\alpha_{t}^{2}G^{2}$ to the above inequality (see Lemma~\ref{lem:bdw}).
+
+	\begin{comment}
+	Alternatively, we can also write
+	\begin{align*}
+	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & =\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\alpha_{t-1}\vg_{t-1}-\vw_{t-1}^{k}+\alpha_{t-1}\vg_{t-1,k}\|^{2}\\
+	& \leq2\sum_{k=1}^{N}p_{k}\left(\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\alpha_{t-1}\vg_{t-1}-\alpha_{t-1}\vg_{t-1,k}\|^{2}\right)
+	\end{align*}
+	and write 
+	\begin{align*}
+	\sum_{k}p_{k}\|\vg_{t-1,k}-\vg_{t-1}\|^{2}\leq\sum_{k}p_{k}\|\vg_{t-1,k}\|^{2} & =\sum_{k}p_{k}\|\vg_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})+\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
+	& \leq2\sum_{k}p_{k}\|\vg_{t-1,k}-\nabla F_{k}(\vw_{t-1}^{k})\|^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})\|^{2}\\
+	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+2\sum_{k}p_{k}\|\nabla F_{k}(\vw_{t-1}^{k})-\nabla F_{k}(\ov{w}_{t-1})+\nabla F_{k}(\ov{w}_{t-1})\|^{2}\\
+	& \leq2\sum_{k}p_{k}\sigma_{k}^{2}+4\sum_{k}p_{k}\left(\|\nabla F_{k}(\vw_{t-1}^{k})-\nabla F_{k}(\ov{w}_{t-1})\|^{2}+\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}\right)\\
+	& \leq2\sigma^{2}+4\sum_{k}p_{k}\left(L^{2}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}\right)
+	\end{align*}
+	so that 
+	\begin{align*}
+	\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+4\sigma^{2}\alpha_{t-1}^{2}+8\alpha_{t-1}^{2}\sum_{k}p_{k}\|\nabla F_{k}(\ov{w}_{t-1})\|^{2}
+	\end{align*}
+	and taking expectation, we have 
+	\begin{align*}
+	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & \leq2(1+4L^{2}\alpha_{t-1}^{2})\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t-1}-\vw_{t-1}^{k}\|^{2}+4\alpha_{t-1}^{2}(\sigma^{2}+2G^{2})
+	\end{align*}
+	Using this recusive relation at most $E$ times to arrive at a communication
+	round and that $\alpha_{t}L\leq1/8$ for all $t$, we have 
+	\begin{align*}
+	\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\leq E\frac{1}{L^{2}}(\sigma^{2}+2G^{2})
+	\end{align*}
+	which gives 
+	\end{comment}
+	Therefore, we can conclude that, with $\nu_{max}:=N\cdot\max_{k}p_{k}$ and $\nu_{min}:=N\cdot\min_{k}p_{k}$, 
+	\begin{align*}
+	& \mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}\\
+	\leq& \mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	= &\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\sum_{k=1}^{N}(p_{k}N)^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	\leq &\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\nu_{max}^{2}\sum_{k=1}^{N}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	\leq &\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}
+	\end{align*}
+	where in the last inequality we use $\sigma^2=\sum_{k=1}^{N}p_k\sigma_{k}^{2}$, and that by construction $\alpha_{t}$
+	satisfies $L\alpha_{t}\leq\frac{1}{8}$.
+
+\end{proof}
+
 One may ask whether the dependence on $E$ in the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$
-can be removed, or equivalently whether $\sum_{k}p_{k}\|\mathbf{w}_{t}^{k}-\overline{\mathbf{w}}_{t}\|^{2}=O(1/T^{2})$
+can be removed, or equivalently whether $\sum_{k}p_{k}\|\mathbf{w}_{t}^{k}-\overline{\mathbf{w}}_{t}\|^{2}=\mathcal{O}(1/T^{2})$
 can be independent of $E$. We provide a simple counterexample that
 shows that this is not possible in general. 
-\begin{lemma}
+\begin{proposition} \label{prop:tight}
 	There exists a dataset such that if $E=\mathcal{O}(T^{\beta})$ for

### 2021-07-04 source/app_sumarization.tex
Update on Overleaf.

@@
 
-\begin{table}[h!]
+\begin{table*}[h!]
+{
+\scriptsize
 \centering
-{\tiny
 \begin{tabular}{|c|c|c|c|c|c|c|c|}
@@
 FedAvg\cite{koloskova2020unified} & $\cO(\frac{1}{NT})$ & $\cO(N^{-1/2}T^{1/2})$     	& \cmark & Full       & Bounded gradient   & Strongly convex  \\ \hline
-FedAvg/N-FedAvg                 & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{\ddagger}$ & \cmark	 & Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
+FedAvg\cite{karimireddy2019scaffold} & $\cO(\frac{1}{NT})^{\dagger\dagger}$ & $\cO(N^{-1/2}T^{1/2})^{\dagger\dagger}$   	& \cmark & Partial       & Bounded gradient dissimilarity   & Strongly convex  \\ \hline
+FedAvg/N-FedAvg (our work)                 & $\cO(\frac{1}{KT})$ & $\cO(N^{-1/2}T^{1/2})^{\ddagger}$ & \cmark	 & Partial       & Bounded gradient             & Strongly convex  \\\hline\hline
 FedAvg\cite{khaled2020tighter}  & $\cO(\frac{1}{\sqrt{NT}})$ & $\cO(N^{-3/2}T^{1/2})$     	    & \cmark& Full        & Bounded gradient             & Convex  \\\hline
 FedAvg\cite{koloskova2020unified} & $\cO(\frac{1}{\sqrt{NT}})$ & $\cO(N^{-3/4}T^{1/4})$    & \cmark & Full       & Bounded gradient             &  Convex  \\ \hline
-FedAvg/N-FedAvg      & $\cO\left(\frac{1}{\sqrt{KT}}\right)$ & $\cO(N^{-3/4}T^{1/4})^{\ddagger}$& \cmark			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
+FedAvg\cite{karimireddy2019scaffold} & $\cO(\frac{1}{\sqrt{NT}})^{\dagger\dagger}$ & $\cO(N^{-3/4}T^{1/4})^{\dagger\dagger}$  & \cmark & Partial       & Bounded gradient dissimilarity   &  Convex  \\ \hline
+FedAvg/N-FedAvg (our work)      & $\cO\left(\frac{1}{\sqrt{KT}}\right)$ & $\cO(N^{-3/4}T^{1/4})^{\ddagger}$& \cmark			& Partial     & Bounded gradient            &  Convex   \\ \hline\hline
 FedAvg & $\cO\left(\exp(-\frac{NT}{E\kappa_1})\right)$ & $ \cO(T^{\beta})$                   & \cmark&  Partial     & Bounded gradient    & Overparameterized LR\\ \hline
@@
 Therefore, we compare it with our strongly convex results here.\\
-         $^{\ddagger\ddagger}$ The bounded gradient diversity assumption is not applicable for general heterogeneous data when converging to arbitrarily small $\epsilon$-accuracy (see discussions in Sec~\ref{sec:app:comparison}).
+         $^{\ddagger\ddagger}$ The bounded gradient diversity assumption is not applicable for general heterogeneous data when converging to arbitrarily small $\epsilon$-accuracy (see discussions in Sec~\ref{sec:app:comparison}).\\
+         $^{\dagger\dagger}$ Although the results in~\cite{karimireddy2019scaffold} is applicable for partial participation setting, their results only achieve linear speedup under full participation setting $K=N$ while we show linear speedup convergence for $K\leq N$ (see discussions in Sec~\ref{sec:app:comparison}).
+The $E$ in the table is obtained under full participation.
+Under partial participation, the communication complexity is $E=\cO(1)$.
            \par}
 \label{tb:convergenceratev3}
-\end{table}
+\end{table*}

### 2021-07-04 source/app_techlemma.tex
Update on Overleaf.

--- 2020-09-16 source-iclr/app_techlemma.tex
@@
-To facilitate reading, we first summarize some technical lemmas regarding the properties of $L$-smooth and $\mu$-strongly
-convex functions~\cite{rockafellar1970convex} which are useful for the proof later on. 
+To facilitate reading, we first summarize some basic properties of $L$-smooth and $\mu$-strongly
+convex functions, found in e.g.~\cite{rockafellar1970convex}, which are used in various steps of proofs in the appendix. 
 \begin{lemma}
@@
 	4. Optimality gap: $F(\vw)-F(\vw^{\ast})\leq$$\langle\nabla F(\vw),\vw-\vw^{\ast}\rangle$.
+\label{lem:lsmooth}
 \end{lemma}

### 2021-07-04 source/app_unifiedframework.tex
Update on Overleaf.

% !TEX ROOT=./main.tex

% This section provides a unified proof framework for the convergence analysis of FedAvg and its accelerated variants. To facilitate our presentation of the unified framework,
% we first outline the FedAvg in Alg~\ref{alg:fedavg} and we will take 
% the strongly convex problem as an example to illustrate our analysis. 
% The corresponding proof for general convex functions follows the 
% same framework.

% The main roadmap of the convergence proof in this work contain the
% following steps. 
% First, the goal of our analysis is to provide the convergence rate 
% of FedAvg and its accelerated variants, which means the we would like
% to show how does the optimality gap decrease as we conduct more 
% iterations 
% First, recall that our goal is to analyze the convergence rate of FedAvg algorithm, i.e., $F(\vw_t) - F^* \leq \cO(1/T)$ in strongly convex 
% case. 
% To prove this bound, we need to connect the optimality gap with the Frobenius norm
% of current model and the optimal model ($\| \vw_t - \vw^*\|^2$). Then, we 
% analyze the one step progress of FedAvg. This step rigorously shows how 
% much the distance between $\vw_{t+1}$ and $\vw^*$ becomes closer when 
% we take one step stochastic gradient descent in FedAvg. More specifically,
% we provide an upper bound of $\|\vw_{t+1} - \vw^*\|^2$ using the previous
% step $\| \vw_t - \vw^*\|^2$. This will justify a solid theoretical foundation for the step in line 9, Alg~\ref{alg:fedavg}. More importantly, our 
% prove also motivate that we should include more clients to accelerate the
% convergence as we show $\cO(1/NT)$.  Notice that this is the central step of our analysis, which is the main source of improvement of our results comparing to 
% prior arts. 
% Using one step progress bound, we can connect $\|\vw_t - \vw^*\|^2$ to 
% $\|\vw_0 - \vw^*\|^2$ by induction. Therefore, we have an upper bound
% of $\|\vw_t - \vw^*\|^2$ and lastly, we can use the upper bound
% of $\|\vw_t - \vw^*\|^2$ to bound the optimality gap. 
% For strongly convex problems, we can easily connect the optimality gap
% and using L-smooth properties.

% In conclusion, our proof contains two main steps: 1) \textbf{one step progress bound} prove how much progress we have made by taking one step
% SGD in FedAvg. 2) \textbf{Iterative proof} bounds the up to date progress using one step progress bound iteratively. 
To facilitate the understanding of our analysis and highlight the improvement of our work comparing to prior arts, we summarize the general steps 
used in the proofs across the various settings. In this section,
we take the strongly convex case as an example to illustrate our analysis. The corresponding proof for general convex functions follows the 
same framework.

\begin{algorithm}[h!]\small
\begin{algorithmic}[1]
\STATE \textbf{Server input:} initial model $\vw_0$, initial step size $\alpha_0$, local steps $E$. 
\STATE \textbf{Client input:} 
\FOR {each round $r = 0, 1, ..., R$, where $r = t*E$} 
\STATE  Sample clients $\cS_t \subseteq \{1,...,N\}$
\STATE Broadcast $\vw$ to all clients $k \in \cS_t$
\FOR {each client $k \subseteq \cS_t$}
\STATE initialize local model $\vw_t^k = \vw$
\FOR {$t = r * E + 1, \dots, (r+1)*E$}
\STATE $\vw_{t+1}^{k}  =\vw_{t}^{k}-\alpha_{t}\vg_{t,k}$
\ENDFOR
\ENDFOR
\STATE Average the local models at server end: $\ov{w}_t = \sum_{k\in \cS_t} \vw_t^k$.
\ENDFOR 
\end{algorithmic}
\caption{\textsc{FedAvg}: Federated Averaging}% $S$ = cA3C($\theta, \Theta_v$
\label{alg:fedavg}
\end{algorithm}

\textbf{One step progress bound} \\ 
This step establishes the progress of distance ($\|\ov{w}_{t}-\vw^{\ast}\|^{2}$) to optimal solution after one step SGD update (see line 9, Alg~\ref{alg:fedavg}), as the following equation shows:
\begin{align*}
	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq \cO(\eta_t\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2} + \alpha_t^2 \sigma^2/N + \alpha_t^3E^2G^2).
\end{align*}
% \lkxcom{explain variance here and the improvement?}
The above bound consists of three main ingredients, the distance to optima
in previous step (with $\eta_t \in (0, 1)$ to obtained a contraction bound), 
the variance of stochastic gradients in local clients (second term), the variance
across different clients (third term). 
Notice that the third term in this bound is the primary source of improvement in the rate.  Comparing to the bound in~\cite{li2019convergence}, we improve 
the third term from $\cO(\alpha_t^2E^2G^2)$ to $\cO(\alpha_t^2E^2G^2)$, which enables the linear speedup in the convergence rate.

\textbf{Iterative deduction}\\
This step uses the \textit{one step progress bound} iteratively to 
connect the the current distance to optimal solution with the initial distance ($\|\ov{w}_{0}-\vw^{\ast}\|^{2}$), as follows:
\begin{align*}
	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq \cO(\mathbb{E}\|\ov{w}_{0}-\vw^{\ast}\|^{2} \frac{1}{T}).
\end{align*}
% \lkxcom{distinguish the difference between strongly convex and general convex?}
Then we can use the distance to optima to upper bound the optimality gap ($F(\vw_t) - F^* \leq \cO(1/T)$), as follows:
\begin{align*}
	\mathbb{E}(F(\ov{w}_{t}))-F^{\ast} & \leq\cO(\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}). 
\end{align*}
The convergence rate of the optimality gap is equally obtained as the convergence rate of the distance to optima.

\textbf{From full participation to partial participation}\\
There are three sources of variances that affect the convergence rate.
The first two sources come from the variances of within local clients and
across clients (second and third term in one step progress bound). 
The partial participation, which involves a sampling procedure, is the third
source of variance. Therefore, comparing to the rate in full participation, this will add another term of variance into the convergence rate, where we follow a similar derivation as in~\cite{li2019convergence}.

### 2021-07-04 source/sec_Nesterov.tex
Update on Overleaf.

--- 2020-06-04 source/sec_Nesterov.tex
@@
 A natural extension of the FedAvg algorithm is to use momentum-based 
-local updates instead of local SGD updates. To our knowledge, the
+local updates instead of local SGD updates in order to accelerate FedAvg. As we know from stochastic optimization, Nesterov and other momentum
+updates fail to provably accelerate over SGD (\cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2020federated}).
+This is in contrast to the classical acceleration result of Nesterov-accelerated gradient descent over GD.
+Thus in the FL setting, the best provable convergence rate for FedAvg with Nesterov updates is the same as FedAvg with SGD updates.
+Nevertheless, Nesterov and other momentum updates are frequently used in practice, in both non-FL and FL settings, and are observed to perform better empirically.
+In fact, previous works such as \cite{stich2018local} on FedAvg with vanilla SGD uses FedAvg with Nesterov or other momentum updates in their experiments to achieve target accuracy.
+Because of the popularity of Nesterov and other momentum-based methods, understanding the linear speedup behavior of FedAvg with such local updates is important.
+To our knowledge, the
 only convergence analyses of FedAvg with momentum-based stochastic
-updates focus on the non-convex smooth case \cite{huo2020faster,yu2019linear,li2018federated}.
-In this section, we complete the picture with $\mathcal{O}(1/NT)$
-and $\mathcal{O}(1/\sqrt{NT})$ convergence results for Nesterov-accelerated
-FedAvg for convex objectives that match the rates from the previous section.
-As we know from stochastic optimization, Nesterov and other momentum
-updates may fail to accelerate over SGD \cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2016influence}.
-Therefore in Section~\ref{sec:overparameterized} we will specialize to overparameterized
-problems where we demonstrate that a particular FedAvg variant with
-momentum updates is able to accelerate over the original FedAvg algorithm. 
-Detailed proofs of convergence results in this section are deferred to Appendix Section~\ref{sec:app:Nesterovfedavg}.
+updates focus on the non-convex smooth case \cite{huo2020faster,yu2019linear,li2018federated}, and no results existed in the convex smooth settings.
+In this section, we complete the picture by providing the first $\mathcal{O}(1/KT)$
+and $\mathcal{O}(1/\sqrt{KT})$ convergence results for Nesterov-accelerated
+FedAvg for convex objectives that match the rates for FedAvg with SGD updates. Detailed proofs of convergence results in this section are deferred to Appendix Section~\ref{sec:app:Nesterovfedavg}.
 
@@
 \end{comment}
-The Nesterov Accelerated FedAvg algorithm follows the updates
+The Nesterov Accelerated FedAvg algorithm follows the updates:
 \begin{align*}
-\mathbf{v}_{t+1}^{k} & =\mathbf{w}_{t}^{k}-\alpha_{t}\mathbf{g}_{t,k},\\
-\mathbf{w}_{t+1}^{k} & =\begin{cases}
+\mathbf{v}_{t+1}^{k} & =\mathbf{w}_{t}^{k}-\alpha_{t}\mathbf{g}_{t,k}, \\
+\mathbf{w}_{t+1}^{k} &=\begin{cases}
 \mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E},\\
-\sum_{k \in \cS_{t+1}}\left[\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E}.
+\sum_{k \in \cS_{t+1}}q_k\left[\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E},
 \end{cases}
@@
 where $\mathbf{g}_{t,k}:=\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})$ is
-the stochastic gradient sampled on the $k$-th device at time $t$.  
+the stochastic gradient sampled on the $k$-th device at time $t$, and $q_k$ again depends on participation and sampling schemes.  
 \begin{theorem}
-	\label{thm:nesterov_scvx}Let $\overline{\mathbf{v}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{v}_{T}^{k}$
+	\label{thm:nesterov_scvx}Let $\overline{\mathbf{v}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{v}_{T}^{k}$ in Nesterov accelerated FedAvg,
 	and set learning rates $\alpha_{t}=\frac{6}{\mu}\frac{1}{t+\gamma}$,  $\beta_{t-1}=\frac{3}{14(t+\gamma)(1-\frac{6}{t+\gamma})\max\{\mu,1\}}$.
 Then under Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2} with full device participation,
-	\begin{align*}
+	% \vspace{-1em}
+	\small{\begin{align*}
 	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right),
-	\end{align*}
+	\end{align*}}
+	% \vspace{-1em}
 	and with partial device participation with $K$ sampled devices at
 	each communication round, 
+	\small{
 	\begin{align*}
 	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).
-	\end{align*}
+	\end{align*}}
 \end{theorem}
+Similar to FedAvg, the key step in the proof of this result is a recursive contraction bound, but different in that it involves three time steps, due to the update format of Nesterov SGD (see Lemma~\ref{lem:nest-scvxoner} in Appendix~\ref{sec:convexsmoothsgd}).
+% \begin{align*}
+% \mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
+% & +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|
+% \end{align*}
+Then we can again use induction and $L$-smoothness to obtain the desired bound.
 %
@@
 
-We now show that the optimality gap of Nesterov Accelerated FedAvg has $\mathcal{O}(1/\sqrt{NT})$ rate. This result complements the strongly convex case in the previous
+We now show that the optimality gap of Nesterov-accelerated FedAvg has $\mathcal{O}(1/\sqrt{KT})$ rate for convex and smooth objectives.
+This result complements the strongly convex case in the previous
 part, as well as the non-convex smooth setting in \cite{huo2020faster,yu2019linear,li2018federated},
-where a similar $\mathcal{O}(1/\sqrt{NT})$ rate is given in terms
+where a similar $\mathcal{O}(1/\sqrt{KT})$ rate is given in terms
 of averaged gradient norm. 
@@
 	full device participation has rate
-	\begin{align*}
+	% \vspace{-0em}
+	\small{\begin{align*}
 		\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T}\right),
-	\end{align*}
+		\end{align*}}
+		% \vspace{-1em}
 	and with partial device participation with $K$ sampled devices at
 	each communication round, 
-	\begin{align*}
+	\small{\begin{align*}
 		\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right).
-	\end{align*}
+		\end{align*}}
 \end{theorem}
-%
+
+We emphasize again that in the stochastic optimization setting, the optimal convergence rate that FedAvg wtih Nesterov udpates can achieve is the same as FedAvg with SGD updates.
+However, due to the popularity and superior performance of momentum methods in practice, it is still important to understand the linear speedup behavior of such FedAvg variants.
+Our results in this section fill exactly this gap.
+
+	\begin{comment}
 	It is possible to extend the results in this section to accelerated
-    FedAvg algorithms with other momentum-based updates. However, in the
-    stochastic optimization setting, none of these methods will be able
-    to achieve a better rate than the original FedAvg with SGD updates in general.
-    For this reason, we will instead turn to the overparameterized setting
-    \cite{ma2017power,liu2018accelerating,canziani2016analysis} in the
-    next section where we show that FedAvg enjoys geometric convergence
-    and it is possible to improve its convergence rate with momentum-based updates.
+    FedAvg algorithms with other momentum-based updates. However, as know from stochastic optimization, Nesterov and other momentum
+updates may fail to accelerate over SGD (\cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2016influence}).
+    For this reason, we instead turn to the overparameterized setting
+    \cite{ma2017power,liu2018accelerating,canziani2016analysis} in Section~\ref{sec:app:overparameterized} of the appendix where we show that FedAvg enjoys geometric convergence
+    and it is possible to improve its convergence rate with a new momentum-based FedAvg variant, which we term ``FedMaSS''.
+	\end{comment}

### 2021-07-04 source/sec_SGD.tex
Update on Overleaf.

@@
 
-We first show that FedAvg has $\mathcal{O}(1/NT)$ convergence rate
-for $\mu$-strongly convex and $L$-smooth objectives. The result
-improves on the $\mathcal{O}(1/T)$ rate of \cite{li2019convergence}
-with a linear speedup in the number of devices $N$. Moreover, it
+We first show that FedAvg has an $\mathcal{O}(1/KT)$ convergence rate
+for $\mu$-strongly convex and $L$-smooth objectives. The result relies on a technical improvement over the analysis in~\cite{li2019convergence}. Moreover, it
 implies a distinction in communication efficiency that guarantees
-this linear speedup for FedAvg with full and partial device participations.
+this linear speedup for FedAvg with full and partial device participation.
 With full participation, $E$ can be chosen as large as $\mathcal{O}(\sqrt{T/N})$
@@
 the other hand, with partial participation, $E$ must be $\mathcal{O}(1)$
-to guarantee $\mathcal{O}(1/NT)$ convergence.
+to guarantee $\mathcal{O}(1/KT)$ convergence.
 \begin{theorem}
-	\label{thm:SGD_scvx}Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$,
-	$\nu_{\max}=\max_{k}Np_{k}$, and set decaying learning rates $\alpha_{t}=\frac{1}{4\mu(\gamma+t)}$
+	\label{thm:SGD_scvx}Let $\overline{\mathbf{w}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{w}_{T}^{k}$ in FedAvg,
+	$\nu_{\max}=\max_{k}Np_{k}$, and set decaying learning rates $\alpha_{t}=\frac{4}{\mu(\gamma+t)}$
 	with $\gamma=\max\{32\kappa,E\}$ and $\kappa=\frac{L}{\mu}$. Then
@@
 	at each communication round, 
+	\begin{align*}{\small
+	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).}
+	\end{align*}
+	\label{th:scvx_sgd}
+\end{theorem}
+\textbf{Proof sketch.} Because our unified analyses of results in the main text follow the same framework with variations in technical details, we first give an outline of proof  for Theorem~\ref{th:scvx_sgd} to illustrate the main ideas.
+For full participation, the main ingredient is a recursive contraction bound
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} &  \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}\\
+	& +\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6\alpha_{t}^{3}LE^{2}G^{2}
 	\end{align*}
-\end{theorem}
-%
-\textbf{Linear speedup. }We first compare our bound with that in \cite{li2019convergence},
+ where the $\mathcal{O}(\alpha_{t}^{3}E^{2}G^{2})$ term is the key improvement over the bound in~\cite{li2019convergence}, which has $\mathcal{O}(\alpha_{t}^{2}E^2 G^2)$ instead.
+We then use induction to obtain a non-recursive bound on $\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}$, which is converted to a bound on $\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}$ using $L$-smoothness.
+For partial participation, an additional term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order resulting from sampling variance is added to the contraction bound.
+To facilitate the understanding of our analysis, please refer to a high-level summary in Appendix~\ref{sec:app:sum}.
+
+\begin{comment}
+A crucial ingredient in the proof of Theorem~\ref{th:scvx_sgd} is a one step contraction bound 
+	\begin{align*}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}L\alpha_{t}^{3}G^{2} 
+	\end{align*}
+that improves upon the analysis in~\cite{li2019convergence} to deliver the linear speedup.
+Moreover, analogues of this bound appear in the analyses of the convex setting as well as the Nesterov accelerated FedAvg.\\
+\end{comment}
+
+\textbf{Linear speedup. }We compare our bound with that in \cite{li2019convergence},
 which is $\mathcal{O}(\frac{1}{NT}+\frac{E^{2}}{KT}+\frac{E^{2}G^{2}}{T})$.
@@
 achieve linear speedup. The improvement of our bound comes from the
-term $\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}$, which now is $\mathcal{O}(E^{2}/T^{2})$.
-As a result, all leading terms scale with $1/N$ in the full device
-particiaption setting, and with $1/K$ in the partial participation
+term $\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}$, which now is $\mathcal{O}(E^{2}/T^{2})$ and so is not of leading order. As a result, all leading terms scale with $1/N$ in the full device
+participation setting, and with $1/K$ in the partial participation
 setting. This implies that in both settings, there is a \emph{linear
-	speedup} in the number of active workers during the communication
-round. \\
+	speedup} in the number of active workers during a communication
+round. We also emphasize that the reason one cannot recover the full participation bound by setting $K=N$ in the partial participation bound is due to the variance generated by sampling. \\
 % To our knowledge, this is the first result that explicitly demonstrates
@@
 in the choice of $E$ between the full and partial participation settings.
-With full participation there is linear speedup $\mathcal{O}(1/NT)$
-as long as $E=\mathcal{O}(\sqrt{T/N})$ since then $\mathcal{O}(E^{2}/T^{2})=\mathcal{O}(1/NT)$
+With full participation, the term involving $E$, $\mathcal{O}(E^{2}/T^{2})$, is not of leading order $\mathcal{O}(1/T)$, so we can increase $E$ and reduce the number of communication rounds without degrading the linear speedup in iteration complexity $\mathcal{O}(1/NT)$, as long as $E=\mathcal{O}(\sqrt{T/N})$, since then $\mathcal{O}(E^{2}/T^{2})=\mathcal{O}(1/NT)$
 matches the leading term. This corresponds to a communication complexity
@@
 still yields a linear speedup in $K$, which is also confirmed by
-experiments. The $E=\mathcal{O}(1)$ requirement in partial participation
-likely cannot be removed for our sampling schemes, as the sampling
-variance is $\Omega(E^{2}/T^{2})$ and the dependence on $E$ is tight.  \\
-\textbf{Comparison with related works.} To better understand the significance of the obtained bound, we compare our rates to the best-known results in related settings.~\cite{haddadpour2019convergence} proves a linear speedup $\mathcal{O}(1/NT)$ result for strongly convex and smooth objectives, with $\mathcal{O}(N^{1/3}T^{2/3})$ communication complexity with i.i.d.
-data and partial participation.
-However, their results build on the bounded gradient diversity assumption, which implies the existence of $\mathbf{w}^*$ that minimizes all local objectives (see discussions in Section~\ref{sec:assumptions}), effectively removing system heterogeneity.
-The bound in \cite{koloskova2020unified} matches our bound in the full participation case, but their framework excludes partial particiation \cite[Proposition 1]{koloskova2020unified}.
+experiments.
+The requirement that $E=\mathcal{O}(1)$ in order to achieve linear speedup in partial participation cannot be removed for our sampling schemes, as the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$ comes from variance in the sampling process, which is $\mathcal{O}(E^{2}/T^{2})$.
+In Proposition~\ref{prop:tight} in Section~\ref{sec:app:fedavg} of the appendix, we provide a problem instance where the dependence of the sampling variance on $E$ is tight.
+\\
+\textbf{Comparison with related works.} To better understand the significance of the obtained bound, we compare our rates to the best-known results in related settings.
+\cite{haddadpour2019convergence} proves a linear speedup $\mathcal{O}(1/KT)$ result for strongly convex and smooth objectives\footnote{Their result applies to a larger class of non-convex objectives that satisfy the Polyak-Lojasiewicz condition.}, with $\mathcal{O}(K^{1/3}T^{2/3})$ communication complexity with non-\emph{i.i.d.} data and partial participation.
+However, their results build on the bounded gradient diversity assumption, which implies the existence of $\mathbf{w}^*$ that minimizes all local objectives (see discussions in Section~\ref{sec:assumptions} and Appendix~\ref{sec:app:comparison}), effectively removing statistical heterogeneity.
+The bound in \cite{koloskova2020unified} matches our bound in the full participation case, but their framework excludes partial participation \cite[Proposition 1]{koloskova2020unified}.
+The result of~\cite{karimireddy2019scaffold} applies to the full FL setting, but only has linear speedup when $K=\mathcal{O}(N)$, i.e.
+close to full participation, whereas our result has linear speedup for any number of participating devices.
+When there is no data heterogeneity, i.e.
+in the classical distributed optimization paradigm, communication complexity can be further improved, e.g.
+\cite{woodworth2020local,woodworth2020minibatch}, but such results are not directly comparable to ours since we consider the setting where individual devices have access to different datasets.
 
@@
 \subsection{Convex Smooth Objectives}
-Next we provide linear speedup analyses of FedAvg with convex and
-smooth objectives and show that the optimality gap is $\mathcal{O}(1/\sqrt{NT})$. 
+Next we provide linear speedup analysis of FedAvg with convex and
+smooth objectives and show that the optimality gap is $\mathcal{O}(1/\sqrt{KT})$. 
 This result complements the strongly convex case in the previous part, as well as the non-convex
 smooth setting in \cite{jiang2018linear,yu2019parallel,haddadpour2019convergence},
-where $\mathcal{O}(1/\sqrt{NT})$ results are given in terms of averaged
-gradient norm. 
+where $\mathcal{O}(1/\sqrt{KT})$ results are given in terms of averaged
+gradient norm, and it also extends the result in~\cite{khaled2020tighter}, which has linear speedup in the convex setting, but only for full participation.
 \begin{theorem}
-	\label{thm:SGD_cvx}Under assumptions~\ref{ass:lsmooth},\ref{ass:boundedvariance},\ref{ass:subgrad2} and constant learning
-	rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$, 
+	\label{thm:SGD_cvx}Under Assumptions~\ref{ass:lsmooth},\ref{ass:boundedvariance},\ref{ass:subgrad2} and constant learning
+	rate $\alpha_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$, FedAvg satisfies
 	\begin{align*}
@@
 \end{theorem}
+The analysis again relies on a recursive bound, but without contraction: 
+\begin{align*}
+	& \mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) \\
+ \leq &\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
+	\end{align*}
+which is then summed over time steps to give the desired bound, with $\alpha_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$. \\
 %
@@
 be $O(1)$ in order to achieve linear speedup of $\mathcal{O}(1/\sqrt{KT})$.
+This is again due to the fact that the sampling variance $\mathbb{E}\|\ov{w}_t-\ov{v}_t\|^2=\mathcal{O}(\alpha_t^2 E^2G^2)$ cannot be made independent of $E$, as illustrated by Proposition \ref{prop:tight}.
+See also the proof in Section \ref{sec:app:fedavg} for how the sampling variance and the term $E^2G^2/\sqrt{KT}$ are related.
 Our result again demonstrates the difference in communication complexities

### 2021-07-04 source/sec_exp_details.tex
sec A, B, E
Update on Overleaf.

@@
 convergence speed. 
-\begin{figure}
+\begin{figure*}
 \centering
@@
 \label{fig:e}
-\end{figure}
+\end{figure*}

### 2021-07-04 source/sec_intro.tex
Update on Overleaf.

@@
 \section{Introduction}
-Federated learning (FL) is a machine learning setting where many clients (e.g.
-mobile devices or organizations) collaboratively train a model under the orchestration of a central server (e.g.
-service provider), while keeping the training data decentralized~\cite{smith2017federated, kairouz2019advances}.
-In recent years, FL has swiftly emerged as an important learning paradigm~\cite{mcmahan2016communication,li2018federated}--one that enjoys widespread success in applications such as personalized recommendation~\cite{chen2018federated}, virtual assistant~\cite{lamautonomy}, and keyboard prediction~\cite{47586}, to name a few--for at least two reasons: First, the rapid proliferation of smart devices that are equipped with both computing power and data-capturing capabilities provided the infrastructure core for FL.
-Second, the rising awareness of privacy and the exponential growth of computational power (blessed by Moore's law) in mobile devices have made it increasingly attractive to push the computation to the edge.
+Federated learning (FL) is a machine learning paradigm where many clients (e.g., mobile devices or organizations) collaboratively train a model under the orchestration of a central server (e.g., service provider), while keeping the training data decentralized (\cite{smith2017federated, kairouz2019advances}).
+In recent years, FL has swiftly emerged as an important learning paradigm (\cite{mcmahan2016communication,li2018federated})--one that enjoys widespread success in applications such as personalized recommendation (\cite{chen2018federated}), virtual assistant (\cite{lam2019protecting}), and keyboard prediction (\cite{47586}), to name a few--for at least three reasons: First, the rapid proliferation of smart devices that are equipped with both computing power and data-capturing capabilities provided the infrastructure core for FL.
+Second, the rising awareness of privacy and the explosive growth of computational power in mobile devices have made it increasingly attractive to push the computation to the edge.
+Third, the empirical success of communication-efficient FL algorithms has enabled increasingly larger-scale parallel computing and learning with less communication overhead.
 
 Despite its promise and broad applicability in our current era, the potential value FL delivers is coupled with the unique challenges it brings forth.
-In particular, when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated (and hence protects privacy)~\cite{kairouz2019advances}, it faces two challenges that are absent in centralized optimization and distributed (stochastic) optimization~\cite{zhou2017convergence, stich2018local,khaled2019first,liang2019variance,wang2018cooperative,woodworth2018graph,wang2019adaptive,jiang2018linear,yu2019parallel,yu2019linear, khaled2020tighter,koloskova2020unified}:
+In particular, when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated (\cite{kairouz2019advances}), it faces two challenges that are absent in centralized optimization and distributed (stochastic) optimization (\cite{zhou2017convergence, stich2018local,khaled2019first,liang2019variance,wang2018cooperative,woodworth2018graph,wang2019adaptive,jiang2018linear,yu2019parallel,yu2019linear, khaled2020tighter,koloskova2020unified,woodworth2020local,woodworth2020minibatch}):
 
-1) \textbf{Data heterogeneity:} data distributions in devices are different (and data can't be shared);
+1) \textbf{Data (statistical) heterogeneity:} data distributions in devices are different (and data cannot be shared);
 
-2) \textbf{System heterogeneity:} only a subset of devices may access the central server at each time both because the communications bandwidth profiles vary across devices and because there is no central server that has control over when a device is active.
+2) \textbf{System heterogeneity:} only a subset of devices may access the central server at each time both because the communications bandwidth profiles vary across devices and because there is no central server that has control over when a device is active (the presence of ``stragglers'').
+% Gaps in the theoretical understanding.
 
-% Gaps in the theoretical understanding. 
-To address these challenges, Federated Averaging (FedAvg)~\cite{mcmahan2016communication} was proposed as a particularly effective heuristic, which has enjoyed great empirical success~\cite{47586}.
+To address these challenges, Federated Averaging (FedAvg)~\cite{mcmahan2016communication} was proposed as a particularly effective heuristic, which has enjoyed great empirical success.
 This success has since motivated a growing line of research efforts into understanding its theoretical convergence guarantees in various settings.
-For instance, \cite{haddadpour2019convergence} analyzed FedAvg (for non-convex smooth problems satisfying PL conditions) under the assumption that each local device's minimizer is the same as the minimizer of the joint problem (if all devices' data is aggregated together), an overly restrictive assumption.
-Very recently, \cite{li2019convergence} furthered the progress and established an $\cO(\frac{1}{T})$ convergence rate for FedAvg for strongly convex smooth problems. 
+For instance, \cite{haddadpour2019convergence} analyzed FedAvg (for non-convex smooth problems satisfying PL conditions) under the assumption that each local device's minimizer is the same as the minimizer of the joint problem (if all devices' data is aggregated together), an overly restrictive assumption that restricts the extent of data heterogeneity.
+Very recently, \cite{li2019convergence} furthered the progress and established an $\cO(\frac{1}{T})$ convergence rate for FedAvg for strongly convex smooth Federated learning problems with both data and system heterogeneity.
+A similar result in the same setting~\cite{karimireddy2019scaffold} also established an $\cO(\frac{1}{T})$ result that allows for a linear speedup when the number of participating devices is large.
 At the same time, \cite{huo2020faster} studied the Nesterov accelerated FedAvg for non-convex smooth problems and established 
@@
 
-However, despite these very recent fruitful pioneering efforts into understanding the theoretical convergence properties of FedAvg, it remains open as to how the number of devices--particularly the number of devices that participate in the computation--affect the convergence speed.
-In particular, do we get linear speedup of FedAvg? What about when FedAvg is accelerated? These aspects are currently unexplored in FL. We fill in the gaps here by providing affirmative answers.
+However, despite these very recent fruitful pioneering efforts into understanding the theoretical convergence properties of FedAvg, it remains open as to how the number of devices--particularly the number of devices that participate in the computation--affects the convergence speed.
+In particular, is linear speedup of FedAvg a universal phenomenon across different settings and for any number of devices?
+What about when FedAvg is accelerated with momentum updates?
+Does the presence of both data and system heterogeneity in FL imply different communication complexities and require technical novelties over results in distributed and decentralized optimization?
+These aspects are currently unexplored or underexplored in FL.
+We fill in the gaps here by providing affirmative answers.
 
-\begin{table}[t!]
-\hspace{-1em}
-{\small
+\begin{table*}[th!]
 \centering
@@
 \end{tabular}
-}
 \caption{Our convergence results for FedAvg and accelerated FedAvg in this paper. Throughout the paper, $N$ is the total
@@
 \label{tb:convergencerateintro}
-\vspace{-2.5em}
-\end{table}
+\end{table*}
 
 \textbf{Our Contributions}
-We provide a comprehensive convergence analysis
+We provide a comprehensive and unified convergence analysis
 of FedAvg and its accelerated variants in the presence of both data and system heterogeneity. 
@@
 and smooth problems (where $K$ is the number of participating devices), thereby establishing that FedAvg enjoys the desirable linear speedup property in the FL setup.
-Prior to our work here, the best and the most related convergence analysis is given by \cite{li2019convergence}, which established an $\cO(\frac{1}{T})$ convergence rate for strongly convex smooth problems under FedAvg.
-Our rate matches the same (and optimal) dependence on $T$, but also completes the picture by establishing the linear dependence on $K$.
+Prior to our work here, the best and the most related convergence analysis is given by \cite{li2019convergence} and~\cite{karimireddy2019scaffold}, which established an $\cO(\frac{1}{T})$ convergence rate for strongly convex smooth problems under FedAvg.
+Our rate matches the same (and optimal) dependence on $T$, but also completes the picture by establishing the linear dependence on $K$, for any $K\leq N$, where $N$ is the total number of devices, whereas~\cite{li2019convergence} does not have linear speedup analysis, and~\cite{karimireddy2019scaffold} only allows linear speedup close to full participation ($K=\mathcal{O}(N)$).
+As for convex and smooth problems, there was no prior work that established the {\small{$\cO(\frac{1}{\sqrt{T}})$}} rate under both system and data heterogeneity.
+Our unified analysis highlights the common elements and distinctions between the strongly and convex settings.
 
@@
 To the best of our knowledge, these are the first results that provide a linear speedup characterization of Nesterov accelerated FedAvg in those two problem classes (that FedAvg and Nesterov accelerated FedAvg share the same convergence rate is to be expected: this is the case even for centralized stochastic optimization).
+Prior to our results here, the most relevant results~\cite{yu2019linear,li2018federated,huo2020faster} only concern the non-convex setting, where convergence is measured with respect to stationary points (vanishing of gradient norms, rather than optimality gaps).
+Our unified analysis of Nesterov FedAvg also illustrates the technical similarities and distinctions compared to the original FedAvg algorithm, whereas prior works (in the non-convex setting) were scattered and used different notations.
 
 Third, we study a subclass of strongly convex smooth problems where the objective is over-parameterized and establish 
-a faster $\cO(\exp(-\frac{KT}{\kappa}))$ convergence rate for FedAvg.
+a faster $\cO(\exp(-\frac{KT}{\kappa}))$ convergence rate for FedAvg, in contrast to the $\cO(\exp(-\frac{T}{\kappa}))$ rate for individual solvers~\cite{ma2017power}.
 Within this class, we further consider the linear regression problem and establish an even sharper rate under FedAvg.
-In addition, we propose a new variant of accelerated FedAvg--MaSS accelerated FedAvg--and establish a faster convergence rate (compared to if no acceleration is used).
+In addition, we propose a new variant of accelerated FedAvg based on a momentum update of~\cite{liu2018accelerating}--MaSS accelerated FedAvg--and establish a faster convergence rate (compared to if no acceleration is used).
 This stands in contrast to generic (strongly) convex stochastic problems where theoretically no rate improvement is obtained when one accelerates FedAvg.
 The detailed convergence results are summarized in Table~\ref{tb:convergencerateintro}.
+
+\textbf{Connections with Distributed and Decentralized Optimization}
+Federated learning is closely related to distributed and decentralized optimization, and as such it is important to discuss connections and distinctions between our work and related results from that literature.
+First, when there is neither system heterogeneity, i.e.
+all devices participate in parameter averaging during a communication round, nor statistical heterogeneity, i.e.
+all devices have access to a common set of stochastic gradients, FedAvg coincides with the ``Local SGD'' of~\cite{stich2018local}, which showed the linear speedup rate $\mathcal{O}(1/NT)$ for strongly convex and smooth functions.
+\cite{woodworth2020local} and \cite{woodworth2020minibatch} further improved the communication complexity that guarantees the linear speedup rate.
+When there is only data heterogeneity, some works have continued to use the term Local SGD to refer to FedAvg, while others subsume it in more general frameworks that include decentralized model averaging based on a network topology or a mixing matrix.
+They have provided linear speedup analyses for strongly convex and convex problems, e.g.~\cite{khaled2020tighter,koloskova2020unified} as well as non-convex problems, e.g.~\cite{jiang2018linear,yu2019parallel,wang2018cooperative}.
+However, these results do not consider system heterogeneity, i.e.
+the presence of stragglers in the device network.
+Even with decentralized model averaging, the assumptions usually imply that model averages over all devices is the same as decentralized model averages based on network topology (e.g.~\cite{koloskova2020unified} Proposition 1), which precludes system heterogeneity as defined in this paper and prevalent in FL problems.
+For momentum accelerated FedAvg,~\cite{yu2019linear} provided linear speedup analysis for non-convex problems, while results for strongly convex and convex settings are entirely lacking, even without system heterogeneity.~\cite{karimireddy2019scaffold} considers both types of heterogeneities for FedAvg, but their rate implies a linear speedup only when the number of stragglers is negligible.
+In contrast, our linear speedup analyses consider both types of heterogeneity present in the full federated learning setting, and are valid for any number of participating devices.
+We also highlight a distinction in communication efficiency when system heterogeneity is present.
+Moreover, our results for Nesterov accelerated FedAvg completes the picture for strongly convex and convex problems.
+For a detailed comparison with related works, please refer to Table~\ref{tb:convergenceratev3} in Appendix Section~\ref{sec:app:comparison}.
+
+\textbf{New related works since last time}: \cite{zhu2019multi,xu2020ternary,yang2020federated,chen2019communication,sattler2020clustered,konevcny2016federated,yuan2020stochastic,dinh2020federated,li2020acceleration,nedic2009distributed,duchi2011dual,liu2016collective,yuan2016convergence,lu2011gossip,ram2010distributed,yuan2018optimal,horvath2019stochastic,wangni2017gradient,alistarh2017qsgd}
 

### 2021-07-04 source/app_sgd_scvx_smooth.tex
Update on Overleaf.

@@
 	speedup. Instead, we upper bound it using the following derivation:
-	{\small
+	{
 	\begin{align*}
@@
 	=&-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\|\nabla F(\ov{w}_{t})-\ov{g}_t\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
-	\leq&-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-\sum_{k}p_{k}\|\nabla F(\ov{w}_{t})-\nabla F(\vw_{t}^{k})\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
 	\leq&-\alpha_{t}^{2}\left[\|\nabla F(\ov{w}_{t})\|^{2}+\|\ov{g}_{t}\|^{2}-L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\right]+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}\\
@@
 	\end{align*}}
-	where we have used the smoothness of $F$ twice.
+	where we have used that $\ov{g}_t = \sum_k p_k \nabla F(\vw_{t}^{k})$ and the smoothness of $F$ twice.
 
@@
 	we have so far proved 
+	{\small
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2} \numberthis \label{eq:common recursion}
-	\end{align*}
+	&\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}  \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\\
+	&+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+	+\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2} \numberthis \label{eq:common recursion}
+	\end{align*}}
 	Under Assumption~\ref{ass:subgrad2}, we have $\mathbb{E}\|\vg_{t}\|^{2}\leq G^{2}$.
@@
 	\begin{align*}
-	& \mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}\\
-	\leq& \mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	= &\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\sum_{k=1}^{N}(p_{k}N)^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
-	\leq &\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\nu_{max}^{2}\sum_{k=1}^{N}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	& \mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}
+	\leq \mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}\\
+	+&4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	 =& \mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+4E^{2}L\alpha_{t}^{3}G^{2}+4E^{2}L^{2}\alpha_{t}^{4}G^{2}\\
+	+& \alpha_{t}^{2}\frac{1}{N^{2}}\sum_{k=1}^{N}(p_{k}N)^{2}\sigma_{k}^{2}+\alpha_{t}^{3}LG^{2}\\
+	\leq &\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N^{2}}\nu_{max}^{2}\sum_{k=1}^{N}\sigma_{k}^{2}\\
 	\leq &\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+6E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}

### 2021-07-04 source/app_nasgd_scvx_smooth.tex
Reformat Appendix F and G

@@
+% !TEX ROOT=./main.tex
 Recall that the Nesterov accelerated FedAvg follows the updates 
@@
 \begin{align*}
-\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
-& +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|
+&\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} \\
+\leq & \mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}\\
+&+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2} +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
+&+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|
 \end{align*}
@@
 	each communication round, 
+	{\small
 	\begin{align*}
 	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).
-	\end{align*}
+	\end{align*}}
 \end{thm}
@@
 \begin{align*}
-\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
-& +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|
+&\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} \\
+\leq &\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}\\
+&+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2} +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
+&+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|
 \end{align*}
@@
 (1-\mu\alpha_{t})(1+14\beta_{t-1}) & \leq(1-\frac{6}{t+\gamma})(1+\frac{3}{(t+\gamma)(1-\frac{6}{t+\gamma})})\\
-& =1-\frac{6}{t+\gamma}+\frac{3}{t+\gamma}=1-\frac{3}{t+\gamma}=1-\frac{\mu\alpha_{t}}{2}
+& =1-\frac{6}{t+\gamma}+\frac{3}{t+\gamma}=1-\frac{\mu\alpha_{t}}{2}
 \end{align*}
@@
 \begin{align*}
-2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\| & \leq4\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\cdot b(C\alpha_{t}^{2}+D\alpha_{t})
+&2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\| \\
+\leq & 4\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\cdot b(C\alpha_{t}^{2}+D\alpha_{t})
 \end{align*}
@@
 \begin{align*}
-(1-\alpha_{t}\mu)\beta_{t-1}^{2}\mathbb{E}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2} & \leq4(1-\alpha_{t}\mu)\beta_{t-1}^{2}\cdot b(C\alpha_{t}^{2}+D\alpha_{t})
+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\mathbb{E}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2} 
+\leq 4(1-\alpha_{t}\mu)\beta_{t-1}^{2}\cdot b(C\alpha_{t}^{2}+D\alpha_{t})
 \end{align*}
@@
 \begin{align*}
-\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})((1+\beta_{t-1})^{2}+4\beta_{t-1}(1+\beta_{t-1})+4\beta_{t-1}^{2})\cdot b(C\alpha_{t}^{2}+D\alpha_{t})\\
+&\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} \\
+ \leq&(1-\mu\alpha_{t})((1+\beta_{t-1})^{2}+4\beta_{t-1}(1+\beta_{t-1})+4\beta_{t-1}^{2})\cdot b(C\alpha_{t}^{2}+D\alpha_{t})\\
 & +20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
-& \leq\mathbb{E}(1-\mu\alpha_{t})(1+14\beta_{t-1})\cdot b(C\alpha_{t}^{2}+D\alpha_{t})+20E^{2}L\alpha_{t}^{3}G^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
-& \leq b(1-\frac{\mu\alpha_{t}}{2})(C\alpha_{t}^{2}+D\alpha_{t})+C\alpha_{t}^{3}+D\alpha_{t}^{2}\\
-& =(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}^{2}C+(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}D
+\leq &\mathbb{E}(1-\mu\alpha_{t})(1+14\beta_{t-1})\cdot b(C\alpha_{t}^{2}+D\alpha_{t})+20E^{2}L\alpha_{t}^{3}G^{2}\\
+&+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
+\leq & b(1-\frac{\mu\alpha_{t}}{2})(C\alpha_{t}^{2}+D\alpha_{t})+C\alpha_{t}^{3}+D\alpha_{t}^{2}\\
+= &(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}^{2}C+(b(1-\frac{\mu\alpha_{t}}{2})+\alpha_{t})\alpha_{t}D
 \end{align*}
@@
 \begin{align*}
-\mathbb{E}(F(\ov{v}_{T}))-F^{\ast} & =\mathbb{E}(F(\ov{v}_{T})-F(\vw^{\ast}))\\
-& \leq\frac{L}{2}\mathbb{E}\|\ov{v}_{T}-\vw^{\ast}\|^{2}\leq\frac{L}{2}c\|\vv_{0}-\vw^{\ast}\|^{2}\frac{6}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
-& =3c\|\vv_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
-& \leq3c\|\vv_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{6}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{\max}\sigma^{2}+20E^{2}LG^{2}\cdot(\frac{6}{\mu(T+\gamma)})^{2}\right]\\
-& =\mathcal{O}(\frac{\kappa}{\mu}\frac{1}{N}\nu_{\max}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
+&\mathbb{E}(F(\ov{v}_{T}))-F^{\ast} \\
+=&\mathbb{E}(F(\ov{v}_{T})-F(\vw^{\ast}))\\
+\leq & \frac{L}{2}\mathbb{E}\|\ov{v}_{T}-\vw^{\ast}\|^{2}\leq\frac{L}{2}c\|\vv_{0}-\vw^{\ast}\|^{2}\frac{6}{\mu}(D\alpha_{T}+C\alpha_{T}^{2})\\
+= & 3c\|\vv_{0}-\vw^{\ast}\|^{2}\kappa(D\alpha_{T}+C\alpha_{T}^{2})\\
+\leq &3c\|\vv_{0}-\vw^{\ast}\|^{2}\kappa\left[\frac{6}{\mu(T+\gamma)}\cdot\frac{1}{N}\nu_{\max}\sigma^{2}+20E^{2}LG^{2}\cdot(\frac{6}{\mu(T+\gamma)})^{2}\right]\\
+ =&\mathcal{O}(\frac{\kappa}{\mu}\frac{1}{N}\nu_{\max}\sigma^{2}\cdot\frac{1}{T}+\frac{\kappa^{2}}{\mu}E^{2}G^{2}\cdot\frac{1}{T^{2}})
 \end{align*}
@@
 \begin{align*}
-\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} & =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
-& \leq\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
-& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
-& =\mathbb{E}\sum_{k=1}^{N}p_{k}\|\sum_{i=t_{0}}^{t-1}\beta_{i}(\vv_{i+1}^{k}-\vv_{i}^{k})-\sum_{i=t_{0}}^{t-1}\alpha_{i}\vg_{i,k}\|^{2}\\
-& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}\|\vg_{i,k}\|^{2}+2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\beta_{i}^{2}\|(\vv_{i+1}^{k}-\vv_{i}^{k})\|^{2}\\
-& \leq2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}(\|\vg_{i,k}\|^{2}+\|(\vv_{i+1}^{k}-\vv_{i}^{k})\|^{2})\\
-& \leq4\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}G^{2}\\
-& \leq4(E-1)^{2}\alpha_{t_{0}}^{2}G^{2}\leq16(E-1)^{2}\alpha_{t}^{2}G^{2}
+& \mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} \\ 
+=&\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}-(\ov{w}_{t}-\ov{w}_{t_{0}})\|^{2}\\
+\leq &\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\ov{w}_{t_{0}}\|^{2}\\
+= &\mathbb{E}\sum_{k=1}^{N}p_{k}\|\vw_{t}^{k}-\vw_{t_{0}}^{k}\|^{2}\\
+= &\mathbb{E}\sum_{k=1}^{N}p_{k}\|\sum_{i=t_{0}}^{t-1}\beta_{i}(\vv_{i+1}^{k}-\vv_{i}^{k})-\sum_{i=t_{0}}^{t-1}\alpha_{i}\vg_{i,k}\|^{2}\\
+\leq &2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}\|\vg_{i,k}\|^{2}\\
+&+2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\beta_{i}^{2}\|(\vv_{i+1}^{k}-\vv_{i}^{k})\|^{2}\\
+\leq &2\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}(\|\vg_{i,k}\|^{2}+\|(\vv_{i+1}^{k}-\vv_{i}^{k})\|^{2})\\
+\leq &4\sum_{k=1}^{N}p_{k}\mathbb{E}\sum_{i=t_{0}}^{t-1}(E-1)\alpha_{i}^{2}G^{2}\\
+\leq &4(E-1)^{2}\alpha_{t_{0}}^{2}G^{2}\leq16(E-1)^{2}\alpha_{t}^{2}G^{2}
 \end{align*}
@@
 \begin{align*}
-\vv_{t+1}^{k}-\vv_{t}^{k} & =-\alpha_{t-1}\vg_{t-1,k}+\beta_{t-1}(\vv_{t}^{k}-\vv_{t-1}^{k})-(\alpha_{t}\vg_{t,k}-\alpha_{t-1}\vg_{t-1,k})\\
-& =\beta_{t-1}(\vv_{t}^{k}-\vv_{t-1}^{k})-\alpha_{t}\vg_{t,k}
+ & \vv_{t+1}^{k}-\vv_{t}^{k} \\
+=&-\alpha_{t-1}\vg_{t-1,k}+\beta_{t-1}(\vv_{t}^{k}-\vv_{t-1}^{k})-(\alpha_{t}\vg_{t,k}-\alpha_{t-1}\vg_{t-1,k})\\
+=&\beta_{t-1}(\vv_{t}^{k}-\vv_{t-1}^{k})-\alpha_{t}\vg_{t,k}
 \end{align*}
@@
 \begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+	 &\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} \\
+	& \leq\mathbb{E}(1-\mu\alpha_{t})\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
 	& +\alpha_{t}^{2}L^{2}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}L\mathbb{E}\|\vg_{t}\|^{2}-\alpha_{t}^{2}\|\nabla F(\ov{w}_{t})\|^{2}
@@
 \begin{align*}
-\|\ov{w}_{t}-\vw^{\ast}\|^{2} & =\|\ov{v}_{t}+\beta_{t-1}(\ov{v}_{t}-\ov{v}_{t-1})-\vw^{\ast}\|^{2}\\
-& =\|(1+\beta_{t-1})(\ov{v}_{t}-\vw^{\ast})-\beta_{t-1}(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
-& =(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}-2\beta_{t-1}(1+\beta_{t-1})\langle\ov{v}_{t}-\vw^{\ast},\ov{v}_{t-1}-\vw^{\ast}\rangle+\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
-& \leq(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+2\beta_{t-1}(1+\beta_{t-1})\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|+\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}
+&\|\ov{w}_{t}-\vw^{\ast}\|^{2} \\
+ =&\|\ov{v}_{t}+\beta_{t-1}(\ov{v}_{t}-\ov{v}_{t-1})-\vw^{\ast}\|^{2}\\
+=&\|(1+\beta_{t-1})(\ov{v}_{t}-\vw^{\ast})-\beta_{t-1}(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
+=&(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}-2\beta_{t-1}(1+\beta_{t-1})\langle\ov{v}_{t}-\vw^{\ast},\ov{v}_{t-1}-\vw^{\ast}\rangle\\
+&+\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
+\leq&(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+2\beta_{t-1}(1+\beta_{t-1})\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|\\
+&+\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}
 \end{align*}
@@
 \begin{align*}
-\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\alpha_{t}\mu)(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+2(1-\alpha_{t}\mu)\beta_{t-1}(1+\beta_{t-1})\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-& +\beta_{t-1}^{2}(1-\alpha_{t}\mu)\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}LG^{2}
+&\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} \\
+\leq&(1-\alpha_{t}\mu)(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}\\
+& +2(1-\alpha_{t}\mu)\beta_{t-1}(1+\beta_{t-1})\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|\\
+&+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+\beta_{t-1}^{2}(1-\alpha_{t}\mu)\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
+&+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{3}LG^{2}
 \end{align*}
@@
 We can check that our choice of $\alpha_t$ and $\beta_t$ satisfy $\alpha_t$ is non-increasing, $\alpha_t \leq 2\alpha_{t+E}$, and $2\beta_{t-1}^{2}+2\alpha_{t}^{2}\leq1/2$ for all $t\geq 0$, so that we can apply the bound from Lemma~\ref{lem:nest-bdw} on $\mathbb{E}\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}$ to conclude that, with $\nu_{\max}:=N\cdot\max_{k}p_{k}$,
+{\small
 \begin{align*}
-\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+16E^{2}L\alpha_{t}^{3}G^{2}+16E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{3}LG^{2}\\
-& +(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2} +\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|\\
-& \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
-& +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|
+& \mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} \\
+\leq &\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+16E^{2}L\alpha_{t}^{3}G^{2}\\
+&+16E^{2}L^{2}\alpha_{t}^{4}G^{2}+\alpha_{t}^{3}LG^{2} +(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2} \\
+&+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|\\
+\leq& \mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}\\
+&+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}\\
+&+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|
 \end{align*}
+}
 where we have used $\sigma^{2}=\sum_{k}p_{k}\sigma_{k}^{2}$, and by construction our $\alpha_{t}$

### 2021-07-04 source/app_sgd_cvx_smooth.tex

@@
 \begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq \mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
+	&\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast}))\\
+	& \leq \mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
 	\end{align*}
@@
     \begin{align*}
-	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} \\ 
+	&\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}  \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2} \\ 
 	& +2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+\alpha_{t}^{2}\|\ov{g}_{t}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
@@
 	\begin{align*}
-	& \|\ov{w}_{t+1}-\vw^{\ast}\|^{2}\\
-	\leq & \|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]\\
-	& +2\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\vw_{t}^{k}-\ov{w}_{t}\|^{2}+4\alpha_{t}^{2}L(F(\ov{w}_{t})-F(\vw^{\ast}))+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
-	= & \|\ov{w}_{t}-\vw^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right] \\ 
-	 & +\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
+	& \|\ov{w}_{t+1}-\vw^{\ast}\|^{2} \leq  \|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}\\
+	+&2\alpha_{t}\sum_{k=1}^{N}p_{k}\left[F_{k}(\vw^{\ast})-F_{k}(\ov{w}_{t})\right]+2\alpha_{t}^{2}L^{2}\sum_{k}p_{k}\|\vw_{t}^{k}-\ov{w}_{t}\|^{2}\\
+	+&4\alpha_{t}^{2}L(F(\ov{w}_{t})-F(\vw^{\ast}))+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+	= & \|\ov{w}_{t}-\vw^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2} \\ 
+	 +& \alpha_{t}(F(\vw^{\ast})-F(\ov{w}_{t}))
 	 +\alpha_{t}(1-4\alpha_{t}L)(F(\vw^{\ast})-F(\ov{w}_{t}))
@@
 	Since $F(\vw^{\ast})\leq F(\ov{w}_{t})$, as long as $4\alpha_{t}L\leq1$,
-	we can ignore the last term, and rearrange the inequality to obtain
+	the last term is negative and can be ignored. Rearranging the inequality, we obtain
 	\begin{align*}
-	& \|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast}))\\
- \leq & \|\ov{w}_{t}-\vw^{\ast}\|^{2}+(2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
+	& \|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast}))
+ \leq \|\ov{w}_{t}-\vw^{\ast}\|^{2}\\
+ +& (2\alpha_{t}^{2}L^{2}+\alpha_{t}L)\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}\\
 	\leq & \|\ov{w}_{t}-\vw^{\ast}\|^{2}+\frac{3}{2}\alpha_{t}L\sum_{k=1}^{N}p_{k}\|\ov{w}_{t}-\vw_{t}^{k}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}
@@
 	\begin{align*}
-	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+6\alpha_{t}^{3}E^{2}LG^{2}\\
-	& \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
+	&\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast}))\\  \leq &\|\ov{w}_{t}-\vw^{\ast}\|^{2}
+	+\alpha_{t}^{2}\sum_{k=1}^{N}p_{k}^{2}\sigma_{k}^{2}+6\alpha_{t}^{3}E^{2}LG^{2}\\ 
+	\leq &\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
+	&\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast}))\\
+	\leq&\|\ov{w}_{t}-\vw^{\ast}\|^{2}
+	+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\sum_{t=0}^{T}\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq\|\vw_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}
+	&\sum_{t=0}^{T}\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) \\ 
+	\leq & \|\vw_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}
 	\end{align*}
@@
  \begin{align*}
-	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|\vw_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right)
+	& \min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) \leq \frac{1}{\sum_{t=0}^{T}\alpha_{t}} \cdot (\|\vw_{0}-\vw^{\ast}\|^{2}\\
+	+&\sum_{t=0}^{T}\alpha_{t}^{2}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2})
 	\end{align*}
@@
 \begin{align*}
-\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sqrt{NT}}\cdot\|\vw_{0}-\vw^{\ast}\|^{2}+\frac{1}{\sqrt{NT}}T\cdot\frac{N}{T}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\frac{1}{\sqrt{NT}}T(\sqrt{\frac{N}{T}})^{3}6E^{2}LG^{2}\\
-& \leq\frac{1}{\sqrt{NT}}\cdot\|\vw_{0}-\vw^{\ast}\|^{2}+\frac{1}{\sqrt{NT}}T\cdot\frac{N}{T}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\frac{N}{T}6E^{2}LG^{2}\\
-& =(\|\vw_{0}-\vw^{\ast}\|^{2}+\nu_{\max}^{2}\sigma^{2})\frac{1}{\sqrt{NT}}+\frac{N}{T}6E^{2}LG^{2}\\
-& =\mathcal{O}(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T})
+&\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast})  \leq\frac{1}{\sqrt{NT}}\cdot\|\vw_{0}-\vw^{\ast}\|^{2}\\
++&\frac{1}{\sqrt{NT}}T\cdot\frac{N}{T}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\frac{1}{\sqrt{NT}}T(\sqrt{\frac{N}{T}})^{3}6E^{2}LG^{2}\\
+ \leq&\frac{1}{\sqrt{NT}}\cdot\|\vw_{0}-\vw^{\ast}\|^{2}+\frac{1}{\sqrt{NT}}T\cdot\frac{N}{T}\cdot\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+\frac{N}{T}6E^{2}LG^{2}\\
+ =&(\|\vw_{0}-\vw^{\ast}\|^{2}+\nu_{\max}^{2}\sigma^{2})\frac{1}{\sqrt{NT}}+\frac{N}{T}6E^{2}LG^{2}\\
+ =&\mathcal{O}(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T})
 \end{align*}
 
-	{\color{red}For partial participation, the one step progress bound in Lemma \ref{lem:cvxoner} is updated in a similar manner as the strongly convex case in \eqref{eqn:scvx-partial-oner} to incorporate the sampling variance.
+For partial participation, the one step progress bound in Lemma \ref{lem:cvxoner} is updated in a similar manner as the strongly convex case in \eqref{eqn:scvx-partial-oner} to incorporate the sampling variance.
 More precisely, with partial participation,
@@
 	\begin{align*}
-	\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{v}_{t})-F(\vw^{\ast})) & \leq\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}\\ &
-	   \leq\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
+	&\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{v}_{t})-F(\vw^{\ast})) \\ \leq &\mathbb{E}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}
+	+6\alpha_{t}^{3}E^{2}LG^{2}\\
+	 \leq&\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
 	\end{align*}
@@
 	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} + \alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) & \leq\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}(\frac{1}{N}\nu_{max}^{2}\sigma^{2}+C)+6E^{2}L\alpha_{t}^{3}G^{2},
+	&\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} + \alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast}))\\ \leq&\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}
+	+\alpha_{t}^{2}(\frac{1}{N}\nu_{max}^{2}\sigma^{2}+C)+6E^{2}L\alpha_{t}^{3}G^{2},
 	\end{align*}
 	where $C=\frac{4}{K}E^{2}G^{2}$ or $\frac{N-K}{N-1}\frac{4}{K}E^{2}G^{2}$ depending on the sampling scheme.
-	}
 
@@
 	\begin{align*}
-	\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) & \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\left(\|\vw_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot(\frac{1}{N}\nu_{\max}\sigma^{2}+C)+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right),
+	&\min_{t\leq T}F(\ov{w}_{t})-F(\vw^{\ast}) \leq\frac{1}{\sum_{t=0}^{T}\alpha_{t}}\cdot\\ &\left(\|\vw_{0}-\vw^{\ast}\|^{2}+\sum_{t=0}^{T}\alpha_{t}^{2}\cdot(\frac{1}{N}\nu_{\max}\sigma^{2}+C)+\sum_{t=0}^{T}\alpha_{t}^{3}\cdot6E^{2}LG^{2}\right),
 	\end{align*}

### 2021-07-04 source/sec_SGD.tex
Update on Overleaf.

@@
 	at each communication round, 
-	\begin{align*}{\small
-	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).}
-	\end{align*}
+	{\small \begin{align*}
+	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).
+	\end{align*}}
 	\label{th:scvx_sgd}
 \end{theorem}
-\textbf{Proof sketch.} Because our unified analyses of results in the main text follow the same framework with variations in technical details, we first give an outline of proof  for Theorem~\ref{th:scvx_sgd} to illustrate the main ideas.
+Because our unified analyses of results in the main text follow the same framework with variations in technical details, we first give an outline of proof  for Theorem~\ref{th:scvx_sgd} to illustrate the main ideas.
+
+To facilitate the understanding of our analysis and highlight the improvement of our work comparing to prior arts, we summarize the general steps 
+used in the proofs across the various settings. In this section,
+we take the strongly convex case as an example to illustrate our analysis. The corresponding proof for general convex functions follows the 
+same framework.
+
+\textbf{One step progress bound} \\ 
+This step establishes the progress of distance ($\|\ov{w}_{t}-\vw^{\ast}\|^{2}$) to optimal solution after one step of the FedAvg algorithm, as the following equation shows:
+\begin{align*}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq \cO(\eta_t\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2} + \alpha_t^2 \sigma^2/N + \alpha_t^3E^2G^2).
+\end{align*}
+% \lkxcom{explain variance here and the improvement?}
+The above bound consists of three main ingredients, the distance to optima
+in previous step (with $\eta_t \in (0, 1)$ to obtained a contraction bound), 
+the variance of stochastic gradients in local clients (second term), the variance
+across different clients (third term). 
+Notice that the third term in this bound is the primary source of improvement in the rate.  Comparing to the bound in~\cite{li2019convergence}, we improve 
+the third term from $\cO(\alpha_t^2E^2G^2)$ to $\cO(\alpha_t^2E^2G^2)$, which enables the linear speedup in the convergence rate.
+
+\textbf{Iterative deduction}\\
+This step uses the \textit{one step progress bound} iteratively to 
+connect the the current distance to optimal solution with the initial distance ($\|\ov{w}_{0}-\vw^{\ast}\|^{2}$), as follows:
+\begin{align*}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq \cO(\mathbb{E}\|\ov{w}_{0}-\vw^{\ast}\|^{2} \frac{1}{T}).
+\end{align*}
+% \lkxcom{distinguish the difference between strongly convex and general convex?}
+Then we can use the distance to optima to upper bound the optimality gap ($F(\vw_t) - F^* \leq \cO(1/T)$), as follows:
+\begin{align*}
+	\mathbb{E}(F(\ov{w}_{t}))-F^{\ast} & \leq\cO(\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}). 
+\end{align*}
+The convergence rate of the optimality gap is equally obtained as the convergence rate of the distance to optima.
+
+\textbf{From full participation to partial participation}\\
+There are three sources of variances that affect the convergence rate.
+The first two sources come from the variances of within local clients and
+across clients (second and third term in one step progress bound). 
+The partial participation, which involves a sampling procedure, is the third
+source of variance.
+Therefore, comparing to the rate in full participation, this will add another term of variance into the convergence rate, where we follow a similar derivation as in~\cite{li2019convergence}.
+
 For full participation, the main ingredient is a recursive contraction bound 

### 2021-07-05 source/sec_exp.tex
Update exp

@@
 
-\section{Numerical Experiments}
-\label{sec:exp}
-
-In this section, we empirically examine the linear speedup convergence of FedAvg and Nesterov accelerated FedAvg on various scenarios, including strongly convex function, convex smooth function, and overparameterized setting, as proved in previous theorems.
-
-\textbf{Setup.} Follow the experimental setting in~\cite{stich2018local}, we
-conduct experiments on both synthetic datasets and real-world dataset
-w8a~\cite{platt1998fast} $(d=300, n=49749)$. We consider the distributed
-objectives $F(\vw) = \sum_{k=1}^N p_kF_k(\vw)$, and the objective function on
-$k$-th local device includes three cases: 1) \textbf{Strongly convex
-objective}: the regularized binary logistic regression problem, $F_k(\vw) =
-\frac{1}{N_k} \sum_{i=1}^{N_k} \log( 1+ \exp(-y_i^k \vw^T\vx_i^k) + \frac{\lambda}{2}
-\|\vw\|^2$. The regularization parameter is set to $\lambda = 1/n \approx
-2e-5$. 2) \textbf{Convex smooth objective}: the binary logistic regression
-problem without regularization. 3) \textbf{Overparameterized setting}:
-the linear regression problem without adding noise to the label, $F_k(\vw) =
-\frac{1}{N_k} \sum_{i=1}^{N_k} (\vw^T\vx_i^k + b  - y_i^k)^2$.  
-% In the first
-% two cases, we use the w8a dataset to conduct empirical evaluation. 
-% In the overparameterized setting, we use features from w8a dataset and 
-% generate a ground truth $w$ to label each instance, which satisfies the
-% overparameterized setting.
-
-% ~\ref{th:strongcvxsmth}~\ref{th:cvxsmoth}~\ref{th:fedmasslr}
-\textbf{Linear speedup of FedAvg and Nesterov accelerated FedAvg.} To verify the linear speedup convergence as shown in Theorem~\ref{thm:SGD_scvx}~\ref{thm:SGD_cvx}~\ref{thm:nesterov_scvx}~\ref{thm:Nesterov_cvx}, we evaluate the number of iterations needed to reach
-$\epsilon$-accuracy in three objectives.
-We initialize all runs with $\vw_0 = \textbf{0}_d$ and measure the number of iterations to reach the target accuracy $\epsilon$.
-For each configuration $(E, K)$, we extensively search the learning rate from $\min(\eta_0, \frac{nc}{1 + t})$, where
-$\eta_0 \in \{0.1, 0.12, 1, 32 \}$ according to different problems and $c$ can
-take the values $c = 2^i \ \forall i \in \ZZ$. As the results shown in Figure~\ref{fig:speedup},
-the number of iterations decreases as the number of (active) workers increasing, which is consistent for FedAvg and Nesterov accelerated FedAvg across all scenarios.
-For the additional experiments on the impact of $E$, detailed experimental setup, and hyperparameter setting, please refer to the Appendix, section~\ref{sec:expsupp}.
-
-% \textbf{Partial participations.} We verify the linear speedup in the
-% partial participation settings, where we set $50\%$ of devices
-% are active. As the results are shown in Figure~\ref{fig:speedup} (2nd row), the 
-% FedAve enjoys linear speedup in various settings even with partial
-% device participation.
-
-% \textbf{Nesterov accelerated FedAvg.} In the third row of Figure~\ref{fig:speedup}, 
-% we report the last iteration to converge to $\epsilon$-accuracy of Nesterov accelerated FedAvg.
-The empirical observations align with Theorem~\ref{thm:nesterov_scvx}~\ref{thm:Nesterov_cvx} that the accelerated version of FedAvg can also achieve the linear speedup
-% w.r.t the number of workers.
-
-% \textbf{The impact of $E$.} We further examine how does the number of local steps ($E$) 
-% affect convergence. As shown in Figure~\ref{fig:e} in appendix section~\ref{sec:expsupp}, the number of iterations increases as $E$ increase, which slow down the convergence in
-% terms of gradient computation. However, it can save communication costs as
-% the number of rounds decreased when the $E$ increases. This showcase that
-% we need a proper choice of $E$ to trade-off the communication cost and
-% convergence speed.
-
-\begin{figure}[t!]
+\begin{figure*}[ht!]
 \centering
@@
 \label{fig:speedup}
-\end{figure}
+\end{figure*}
+
+\section{Numerical Experiments}
+\label{sec:exp}
+
+In this section, we empirically examine the linear speedup convergence of FedAvg and Nesterov accelerated FedAvg in various settings, including strongly convex function, convex smooth function, and overparameterized objectives, as analyzed in previous sections.
+
+\textbf{Setup.} Following the experimental setting in~\cite{stich2018local}, we
+conduct experiments on both synthetic datasets and real-world dataset
+w8a~\cite{platt1998fast}, which consists of $n = 49749 $ samples with
+feature dimension $d = 300$. The label is either positive one or negative one.
+The dataset has sparse binary features in $\{0, 1\}$. Each sample
+has 11.15 non-zero feature values out of $300$ features on average.
+We set the batch size equal to four across all experiments.
+For all the results, we set random seeds as $0, 1, 2$
+and report the best convergence rate across the three folds.
+
+In general, we consider the distributed
+objectives $F(\vw) = \sum_{k=1}^N p_kF_k(\vw)$, and the objective function on the
+$k$-th local device includes three cases: 1) Strongly convex
+objective, 2) Convex smooth objective, and 3) Overparameterized setting.
+To verify the linear speedup convergence as shown in Theorems~\ref{thm:SGD_scvx}~\ref{thm:SGD_cvx}~\ref{thm:nesterov_scvx}~\ref{thm:Nesterov_cvx}, we evaluate the number of iterations needed to reach
+$\epsilon$-accuracy in three objectives.
+We initialize all runs with $\vw_0 = \textbf{0}_d$ and measure the number of iterations to reach the target accuracy $\epsilon$.
+For each configuration $(E, K)$, we extensively search the learning rate from $\min(\eta_0, \frac{nc}{1 + t})$, where
+$\eta_0 \in \{0.1, 0.12, 1, 32 \}$ according to different problems and $c$ can
+take the values $c = 2^i \ \forall i \in \ZZ$. As the results shown in Figure~\ref{fig:speedup},
+the number of iterations decreases as the number of (active) workers increasing, which is consistent for FedAvg across all three objectives under the setting of full participation (first row, Figure~\ref{fig:speedup}) and partial participation (second row, Figure~\ref{fig:speedup}).
+For Nesterov accelerated FedAvg, we also empirically verified the linear speedup is achieved for all three objectives under full participation.
+In the next following subsections,we introduce parameter searching in for different types of objectives separately.
+
+\subsection{Strongly Convex Objectives}
+We first consider the strongly convex objective function, where we use
+a regularized binary logistic regression ($F_k(\vw) =
+\frac{1}{N_k} \sum_{i=1}^{N_k} \log( 1+ \exp(-y_i^k \vw^T\vx_i^k) + \frac{\lambda}{2}
+\|\vw\|^2$) with regularization $\lambda=1/n\approx 2e-5$.
+We evenly distributed on $1, 2, 4, 8, 16, 32$ devices and  report the number of iterations/rounds needed to converge to $\epsilon-$accuracy, where $\epsilon=0.005$.
+The optimal objective function value $f^*$
+is set as $f^* = 0.126433176216545$.
+This is determined numerically and we follow the setting in~\cite{stich2018local}.
+The learning rate is decayed as the $\eta_t = \min(\eta_0, \frac{nc}{1 + t})$, where we extensively search the best learning rate $c \in \{2^{-1}c_0, 2^{-2}c_0, c_0, 2c_0, 2^{2}c_0\}$.
+In this case, we search the initial learning rate $\eta_0\in \{1, 32\}$ and $c_0 = 1/8$.
+
+\subsection{Convex Smooth Objectives}
+We also use binary logistic regression without regularization.
+The setting is almost same as its regularized counter part.
+We also evenly distributed all the samples on $1, 2, 4, 8, 16, 32$ devices.
+The figure shows the number of iterations needed to converge to $\epsilon-$accuracy, where $\epsilon=0.02$.
+The optiaml objective function value is set as $f^*=0.11379089057514849$, determined numerically.
+The learning rate is decayed as the $\eta_t = \min(\eta_0, \frac{nc}{1 + t})$, where we extensively search the best learning rate $c \in \{2^{-1}c_0, 2^{-2}c_0, c_0, 2c_0, 2^{2}c_0\}$.
+In this case, we search the initial learning rate $\eta_0\in \{1, 32\}$ and $c_0 = 1/8$.
+
+\subsection{Linear regression}
+We use the linear regression problem ($F_k(\vw) =
+\frac{1}{N_k} \sum_{i=1}^{N_k} (\vw^T\vx_i^k + b  - y_i^k)^2$)) without adding noise to the label
+for the experiments under overparameterized setting.
+For linear regression, we use the same feature vectors from w8a dataset 
+and generate ground truth $[\vw^*, b^*]$ from a multivariate normal distribution
+with zero mean and standard deviation one. Then we generate label 
+based on $y_i = \vx_i^t\vw^* + b^*$. This procedure will ensure we satisfy
+the over-parameterized setting as required in our theorems. 
+We also evenly distributed all the samples on $1, 2, 4, 8, 16, 32$ devices.
+The figure shows the number of iterations needed to converge to $\epsilon-$accuracy, where $\epsilon=0.02$.
+The optiaml objective function value is $f^*=0$.
+The learning rate is decayed as the $\eta_t = \min(\eta_0, \frac{nc}{1 + t})$, where we extensively search the best learning rate $c \in \{2^{-1}c_0, 2^{-2}c_0, c_0, 2c_0, 2^{2}c_0\}$.
+In this case, we search the initial learning rate $\eta_0\in \{0.1, 0.12\}$ and $c_0 = 1/256$.
+
+\subsection{Partial Participation}
+To examine the linear speedup of FedAvg in partial participation setting,
+we evenly distributed data on $4, 8, 16, 32, 64, 128$ devices and 
+uniformly sample $50\%$ devices without replacement. 
+All other hyperparameters are the same as previous sections.
+
+\subsection{Nesterov accelerated FedAvg}
+The experiments of Nesterov accelerated FedAvg (the update formula is given as follows) uses the same setting as
+previous three sections for vanilla FedAvg.
+\begin{align*}
+\vy_{t+1}^{k} & =\vw_{t}^{k}-\alpha_{t}\vg_{t,k}\\
+\vw_{t+1}^{k} & =\begin{cases}
+\vy_{t+1}^{k}+\beta_{t}(\vy_{t+1}^{k}-\vy_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E}\\
+\sum_{k\in\mathcal{S}_{t+1}}\left(\vy_{t+1}^{k}+\beta_{t}(\vy_{t+1}^{k}-\vy_{t}^{k})\right) & \text{if }t+1\in\mathcal{I}_{E}
+\end{cases}
+\end{align*}
+We set $\beta_t = 0.1$ and search $\alpha_t$ in the same way as $\eta_t$
+in FedAvg.
+
+\subsection{The impact of $E$.}
+In this subsection, we further examine how does the number of local steps ($E$) 
+affect convergence. As shown in Figure~\ref{fig:e}, the number of iterations increases as $E$ increase, which slow down the convergence in
+terms of gradient computation. However, it can save communication costs as
+the number of rounds decreased when the $E$ increases. This showcase that
+we need a proper choice of $E$ to trade-off the communication cost and
+convergence speed. 
+\begin{figure*}
+\centering
+	\begin{tabular}{ccc}
+	\hspace{-2em} \includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsT-min-w8a-epsilon0131-reg1e-05.pdf} &
+	\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupEpochsT-min-w8a-epsilon0134-reg0.pdf} & 
+	\includegraphics[width=0.33\textwidth]{fig/paper-linregression-newspeedupEpochsT-min-linearregressionw8a-epsilon002-reg0.pdf} \\
+	\hspace{-2em} \includegraphics[width=0.33\textwidth]{fig/paper-stronglycvxsmthspeedupEpochsRounds-min-w8a-epsilon0131-reg1e-05.pdf} &
+	\includegraphics[width=0.33\textwidth]{fig/paper-cvxsmoothspeedupEpochsRounds-min-w8a-epsilon0134-reg0.pdf} & 
+	\includegraphics[width=0.33\textwidth]{fig/paper-linregression-newspeedupEpochsRounds-min-linearregressionw8a-epsilon002-reg0.pdf} \\
+(a) Strongly convex objective & (b) Convex smooth objective & (c) Linear regression
+	\end{tabular}
+\caption{The convergence of FedAvg w.r.t the number of local steps $E$. }
+\label{fig:e}
+\end{figure*}
+
+% ~\ref{th:strongcvxsmth}~\ref{th:cvxsmoth}~\ref{th:fedmasslr}
+% \textbf{Linear speedup of FedAvg and Nesterov accelerated FedAvg.}
+
+% \textbf{Partial participations.} We verify the linear speedup in the
+% partial participation settings, where we set $50\%$ of devices
+% are active. As the results are shown in Figure~\ref{fig:speedup} (2nd row), the 
+% FedAve enjoys linear speedup in various settings even with partial
+% device participation.
+
+% \textbf{Nesterov accelerated FedAvg.} In the third row of Figure~\ref{fig:speedup}, 
+% we report the last iteration to converge to $\epsilon$-accuracy of Nesterov accelerated FedAvg.
+The empirical observations align with Theorem~\ref{thm:nesterov_scvx}~\ref{thm:Nesterov_cvx} that the accelerated version of FedAvg can also achieve the linear speedup
+% w.r.t the number of workers.
+
+% \textbf{The impact of $E$.} We further examine how does the number of local steps ($E$) 
+% affect convergence. As shown in Figure~\ref{fig:e} in appendix section~\ref{sec:expsupp}, the number of iterations increases as $E$ increase, which slow down the convergence in
+% terms of gradient computation. However, it can save communication costs as
+% the number of rounds decreased when the $E$ increases. This showcase that
+% we need a proper choice of $E$ to trade-off the communication cost and
+% convergence speed.
 

### 2021-07-05 source/sec_SGD.tex

--- 2021-07-04 source/sec_SGD.tex
@@
 	at each communication round, 
-	\begin{align*}{\small
-	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).}
-	\end{align*}
+	{\small \begin{align*}
+	\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).
+	\end{align*}}
 	\label{th:scvx_sgd}
 \end{theorem}
-\textbf{Proof sketch.} Because our unified analyses of results in the main text follow the same framework with variations in technical details, we first give an outline of proof  for Theorem~\ref{th:scvx_sgd} to illustrate the main ideas.
-For full participation, the main ingredient is a recursive contraction bound
+\textbf{Outline of proof.} Because our unified analyses of results in the main text follow the same framework with variations in technical details, we first give an outline of proof for Theorem~\ref{th:scvx_sgd} to illustrate the main ideas and highlight the improvement of our analyses over prior works.\\
+\textit{One-step progress bound.} The analysis starts with the full participation setting, and establishes a recursive contraction bound for the distance to the optimal solution after one step of the FedAvg algorithm:
 \begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} &  \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}\\
-	& +\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6\alpha_{t}^{3}LE^{2}G^{2}
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq (1-\alpha_t)\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+ \cO(\alpha_t^2 \sigma^2/N + \alpha_t^3E^2G^2).
 \end{align*}
- where the $\mathcal{O}(\alpha_{t}^{3}E^{2}G^{2})$ term is the key improvement over the bound in~\cite{li2019convergence}, which has $\mathcal{O}(\alpha_{t}^{2}E^2 G^2)$ instead.
-We then use induction to obtain a non-recursive bound on $\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}$, which is converted to a bound on $\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}$ using $L$-smoothness.
-For partial participation, an additional term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order resulting from sampling variance is added to the contraction bound.
-To facilitate the understanding of our analysis, please refer to a high-level summary in Appendix~\ref{sec:app:sum}.
+% \lkxcom{explain variance here and the improvement?}
+The above bound consists of three components: the distance to the optimal solution 
+in the previous step, 
+the variance of stochastic gradients in local clients, and the variance
+across different clients. This last term $\mathcal{O}(\alpha_{t}^{3}E^{2}G^{2})$ is the key improvement over ~\cite{li2019convergence}, which has $\mathcal{O}(\alpha_{t}^{2}E^2 G^2)$ instead.\\
+\textit{Iterative deduction.} This step applies induction to the recursive one-step progress bounds to obtain a non-recursive bound on $\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}$ in terms of the initial distance ($\|\ov{w}_{0}-\vw^{\ast}\|^{2}$), which is then converted to a bound on $\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}$ using $L$-smoothness.\\
+\textit{From full to partial participation.} In addition to the variances arising from stochastic updates at local clients and averaging across clients, partial participation, which involves a sampling procedure, adds a third source of variance.
+An additional term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order resulting from the sampling variance is added to the recursive contraction bound.
 

### 2021-07-05 source/sec_SGD.tex
Update on Overleaf.

@@
 \textit{Iterative deduction.} This step applies induction to the recursive one-step progress bounds to obtain a non-recursive bound on $\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}$ in terms of the initial distance ($\|\ov{w}_{0}-\vw^{\ast}\|^{2}$), which is then converted to a bound on $\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}$ using $L$-smoothness.\\
-\textit{From full to partial participation.} In addition to the variances arising from stochastic updates at local clients and averaging across clients, partial participation, which involves a sampling procedure, adds a third source of variance.
-An additional term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order resulting from the sampling variance is added to the recursive contraction bound.
+\textit{From full to partial participation.} In addition to the variances arising from stochastic updates at local clients and averaging across clients, partial participation, which involves a sampling procedure, adds a an additional term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order resulting from the sampling variance is added to the recursive contraction bound.
 

### 2021-07-05 source/sec_SGD.tex
Update on Overleaf.

@@
 \textit{Iterative deduction.} This step applies induction to the recursive one-step progress bounds to obtain a non-recursive bound on $\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}$ in terms of the initial distance ($\|\ov{w}_{0}-\vw^{\ast}\|^{2}$), which is then converted to a bound on $\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}$ using $L$-smoothness.\\
-\textit{From full to partial participation.} In addition to the variances arising from stochastic updates at local clients and averaging across clients, partial participation, which involves a sampling procedure, adds a an additional term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order resulting from the sampling variance is added to the recursive contraction bound.
+\textit{From full to partial participation.} In addition to the variances arising from stochastic updates at local clients and averaging across clients, partial participation, which involves a sampling procedure, gives adds a an additional term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order resulting from the sampling variance is added to the recursive contraction bound.
 

### 2021-07-05 source/sec_SGD.tex
Update on Overleaf.

--- 2021-07-05 source/sec_SGD.tex
@@
 \textit{Iterative deduction.} This step applies induction to the recursive one-step progress bounds to obtain a non-recursive bound on $\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}$ in terms of the initial distance ($\|\ov{w}_{0}-\vw^{\ast}\|^{2}$), which is then converted to a bound on $\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}$ using $L$-smoothness.\\
-\textit{From full to partial participation.} In addition to the variances arising from stochastic updates at local clients and averaging across clients, partial participation, which involves a sampling procedure, adds a an additional term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order resulting from the sampling variance is added to the recursive contraction bound.
+\textit{From full to partial participation.} In addition to the variances arising from stochastic updates at local clients and averaging across clients, partial participation, which involves a sampling procedure, gives rise to a third variance term.
+adds a an additional term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order resulting from the sampling variance is added to the recursive contraction bound.
 

### 2021-07-05 source/sec_SGD.tex

@@
 \textit{From full to partial participation.} In addition to the variances arising from stochastic updates at local clients and averaging across clients, partial participation, which involves a sampling procedure, gives rise to a third variance term.
-adds a an additional term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order resulting from the sampling variance is added to the recursive contraction bound.
+An additional term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order resulting from the sampling variance is added to the recursive contraction bound.
 

### 2021-07-11 source/main.tex
rebutal template
Update on Overleaf.

--- 2020-06-05 source/main.tex
@@
-\documentclass{article}
+\documentclass[journal]{IEEEtran}
+\ifCLASSINFOpdf
+  % \usepackage[pdftex]{graphicx}
+  % declare the path(s) where your graphic files are
+  % \graphicspath{{../pdf/}{../jpeg/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+\else
+  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
+  % will default to the driver specified in the system graphics.cfg if no
+  % driver is specified.
+  % \usepackage[dvips]{graphicx}
+  % declare the path(s) where your graphic files are
+  % \graphicspath{{../eps/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  % \DeclareGraphicsExtensions{.eps}
+\fi
 
-% if you need to pass options to natbib, use, e.g.:
-%     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2020
-
-% ready for submission
-% \usepackage{neurips_2020}
-
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
-%     \usepackage[preprint]{neurips_2020}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-%     \usepackage[final]{neurips_2020}
-
-% to avoid loading the natbib package, add option nonatbib:
-     \usepackage[nonatbib]{neurips_2020}
 \usepackage{slashbox,multirow}
@@
 \usepackage{comment}
+
+%%%%%%
 \newcommand{\eq}[1]{{Eq~(#1)}}
@@
 \input{bmacros}
+
+\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
+
 \newtheorem*{assumption*}{Assumption}
+\newtheorem*{claim*}{Claim}
 
-\title{Federated Learning's Blessing:\\
-FedAvg has Linear Speedup}
-
-\author{}
+% correct bad hyphenation here
+\hyphenation{}
 
 \begin{document}
+
+\title{Unified Linear Speedup Analyses of Communication-Efficient Stochastic FedAvg and Nesterov FedAvg}
+
+\author{Zhaonan Qu$^*$, Kaixiang Lin$^*$, Zhaojian Li, Jiayu Zhou, and Zhengyuan Zhou% <-this % stops a space
+\thanks{$^*$ The first two authors contributed equally to this work.}
+\thanks{Zhaonan Qu is with the Department of Economics, Stanford University, CA 94305, USA. Email: {zhaonanq}@stanford.edu}% <-this % stops a space
+\thanks{Kaixiang Lin was with the Department of Computer Science and Engineering, Michigan State University, Lansing, MI, 48824,
+USA. Email: {linkaixi}@msu.edu.
+}% <-this % stops a space
+\thanks{Zhaojian Li is with the Department of
+Mechanical Engineering, Michigan State University, Lansing, MI, 48824,
+USA. Email: {lizhaoj1}@msu.edu.}
+\thanks{Jiayu Zhou is with the Department of Computer Science and Engineering, Michigan State University, Lansing, MI, 48824,
+USA. Email: {jiayuz}@msu.edu.}
+\thanks{Zhengyuan Zhou is with the Department of Technology, Operations, and Statistics at Stern School of Business, New York University, New York, NY 10003,
+USA. Email: {zzhou}@stern.nyu.edu.}
+}
+
+% The paper headers
+\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
+{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}
 
@@
 Federated learning (FL) learns a model jointly from a set of participating devices without sharing each other's privately held data.
-The characteristics of non-\textit{iid} data across the network, low device participation and the mandate that data remain private bring challenges in understanding the convergence of FL algorithms, particularly in regards to how convergence scales with the number of participating devices.
-In this paper, we focus on Federated Averaging (FedAvg)--the most widely used and effective FL algorithm in use today--and provide a comprehensive study of its convergence rate.
-Although FedAvg has recently been studied by an emerging line of literature, it remains open as to how FedAvg's convergence scales with the number of participating devices in the FL setting--a crucial question whose answer would shed light on the performance of FedAvg in large FL systems.
-We fill this gap by establishing convergence guarantees for FedAvg under two classes of problems: strongly convex smooth, convex smooth, and overparametrized strongly convex smooth problems.
-We show that FedAvg enjoys linear speedup in each case, although with different convergence rates.
-For each class, we also characterize the corresponding convergence rates for the Nesterov accelerated FedAvg algorithm in the FL setting: to the best of our knowledge, these are the first linear speedup guarantees for FedAvg when Nesterov acceleration is used.
-To accelerate FedAvg, we also design a new momentum-based FL algorithm that further improves the convergence rate in overparametrized linear regression problems.
+The characteristics of non-\textit{i.i.d.} data across the network, low device participation, high communication costs, and the mandate that data remain private bring challenges in understanding the convergence of FL algorithms, particularly in regards to how convergence scales with the number of participating devices.
+In this paper, we focus on Federated Averaging (FedAvg)--arguably the most popular and effective FL algorithm class in use today--and provide a unified and comprehensive study of its convergence rate.
+Although FedAvg has recently been studied by an emerging line of literature, it remains open as to how FedAvg's convergence scales with the number of participating devices in the fully heterogeneous FL setting--a crucial question whose answer would shed light on the performance of FedAvg in large FL systems.
+We fill this gap by providing a unified analysis that establishes convergence guarantees for FedAvg under three classes of problems: strongly convex smooth, convex smooth, and overparameterized strongly convex smooth problems.
+We show that FedAvg enjoys linear speedup in each case, although with different convergence rates and communication efficiencies.
+While there have been linear speedup results from distributed optimization that assumes full participation, ours are the first to establish linear speedup for FedAvg under both statistical and system heterogeneity.
+For strongly convex and convex problems, we also characterize the corresponding convergence rates for the Nesterov accelerated FedAvg algorithm, which are the first linear speedup guarantees for momentum variants of FedAvg in the convex setting.
+To provably accelerate FedAvg, we design a new momentum-based FL algorithm that further improves the convergence rate in overparameterized linear regression problems.
 Empirical studies of the algorithms in various settings have supported our theoretical results.
@@
 
-% (and its related variant local SGD also analyzed in the simpler homogeneous distributed optimization setting)
+% Note that keywords are not normally used for peerreview papers.
+\begin{IEEEkeywords}
+Federated learning, convex stochastic optimization, linear speedup, momentum updates.
+\end{IEEEkeywords}
+
+\IEEEpeerreviewmaketitle
+
 \input{sec_intro}
@@
 \input{sec_exp}
+\input{sec_conclude}
 
-% \input{sec_conclude}
+\input{sec_ack}
 
+% \bibliographystyle{IEEE}
 \newpage
-\input{sec_impact}
-\bibliographystyle{unsrt}
-\bibliography{ref.bib}
+\bibliographystyle{abbrv}
+%\bibliographystyle{unsrt}
+\bibliography{ref}
 
+\begin{comment}
+
+\begin{IEEEbiography}{Zhaonan Qu}
+Biography text here.
+\end{IEEEbiography}
+
+% if you will not have a photo at all:
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fig/bio_kaixiang.png}}]{Kaixiang Lin}
+Dr. Kaixiang Lin is an applied scientist at Amazon Lab 126. Before joining Amazon, 
+Kaixiang received his Ph.D. degree in Computer Science at Michigan State University in 2020. 
+\end{IEEEbiography}
+
+% insert where needed to balance the two columns on the last page with
+% biographies
+%\newpage
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fig/bio_jiayu.png}}]{Jiayu Zhou}
+Dr. Jiayu Zhou is an assistant professor at
+ Department of Computer Science and Engineering, Michigan State University.
+ Before joining MSU, Jiayu was a staff research scientist at Samsung Research
+ America. Jiayu received his Ph.D. degree in computer science at Arizona
+ State University in 2014. 
+\end{IEEEbiography}
+
+\end{comment}
+
+\appendices
+
+\input{sec_appendix}
+
+% Can use something like this to put references on a page
+% by themselves when using endfloat and the captionsoff option.
+\ifCLASSOPTIONcaptionsoff
   \newpage
-\appendix
-\input{sec_appendix}
+\fi
+
+% 
+% \begin{thebibliography}{1}
+% \bibitem{IEEEhowto:kopka}
+
+% \end{thebibliography}
+
+% You can push biographies down or up by placing
+% a \vfill before or after them. The appropriate
+% use of \vfill depends on what kind of text is
+% on the last page and whether or not the columns
+% are being equalized.
+
+%\vfill
+
+% Can be used to pull up biographies so that the bottom of the last one
+% is flush with the other column.
+%\enlargethispage{-5in}
+
+% that's all folks
 \end{document}
-
-\iffalse
-
-\endif

### 2021-07-11 source/sec_Nesterov.tex
Update on Overleaf.

@@
 A natural extension of the FedAvg algorithm is to use momentum-based 
-local updates instead of local SGD updates in order to accelerate FedAvg. As we know from stochastic optimization, Nesterov and other momentum
-updates fail to provably accelerate over SGD (\cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2020federated}).
-This is in contrast to the classical acceleration result of Nesterov-accelerated gradient descent over GD.
+local updates instead of local SGD updates. As we know from stochastic optimization, Nesterov and other momentum
+updates may fail to provably accelerate over SGD (\cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2020federated}).
+This is in contrast to the classical acceleration result of Nesterov-accelerated gradient descent over GD, as well as recent acceleration results on FedAvg with momentum-based \emph{gradient} updates \cite{liu2019accelerating,yang2020federated}.
 Thus in the FL setting, the best provable convergence rate for FedAvg with Nesterov updates is the same as FedAvg with SGD updates.
@@
 To our knowledge, the
-only convergence analyses of FedAvg with momentum-based stochastic
-updates focus on the non-convex smooth case \cite{huo2020faster,yu2019linear,li2018federated}, and no results existed in the convex smooth settings.
+only convergence analyses of FedAvg with momentum-based \emph{stochastic} updates focus on the non-convex smooth case \cite{huo2020faster,yu2019linear,li2018federated}, and no results exist in the convex smooth settings.
 In this section, we complete the picture by providing the first $\mathcal{O}(1/KT)$
@@
 where $\mathbf{g}_{t,k}:=\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})$ is
-the stochastic gradient sampled on the $k$-th device at time $t$, and $q_k$ again depends on participation and sampling schemes.  
+the stochastic gradient sampled on the $k$-th device at time $t$, $\beta_t$ is the momentum parameter for local Nesterov updates, and $q_k$ again depends on participation and sampling schemes.  
 \begin{theorem}
@@
 \end{theorem}
-Similar to FedAvg, the key step in the proof of this result is a recursive contraction bound, but different in that it involves three time steps, due to the update format of Nesterov SGD (see Lemma~\ref{lem:nest-scvxoner} in Appendix~\ref{sec:convexsmoothsgd}).
+Similar to FedAvg, the key step in the proof of this result is a recursive contraction bound, but different in that it involves three time steps, due to the update format of Nesterov SGD:
+\begin{align*}
+&\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} \\
+\leq & \mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+\mathcal{O}(E^{2}L\alpha_{t}^{3}G^{2})\\
+&+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2} +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}\\
+&+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|.
+\end{align*} 
 % \begin{align*}
@@
 % \end{align*}
-Then we can again use induction and $L$-smoothness to obtain the desired bound.
+Then we can again use induction and $L$-smoothness to obtain the desired bound in Theorem \ref{thm:nesterov_scvx}.
 %
-To our knowledge, this is the first convergence result for Nesterov
+To the best of our knowledge, this is the first convergence result for Nesterov
 accelerated FedAvg in the strongly convex and smooth setting. The
-same discussion about linear speedup of FedAvg applies to the Nesterov
+same discussion about linear speedup of FedAvg applies to its Nesterov
 accelerated variant. In particular, to achieve $\mathcal{O}(1/NT)$
@@
 	% \vspace{-0em}
-	\small{\begin{align*}
+	{\begin{align*}
 		\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T}\right),
@@
 	each communication round, 
-	\small{\begin{align*}
+	{\begin{align*}
 		\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right).
@@
 \end{theorem}
+
+The analogous one-step recursive bound in the convex smooth setting for Nesterov accelerated FedAvg takes the following form: 
+\begin{align*}
+	&\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\eta_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) \\
+	\leq & \mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+2\eta_{t}\frac{\beta_{t}^{2}}{1-\beta_{t}}G^{2}\\
+	& +\mathcal{O}(LE^{2}\alpha_{t}^{2}\eta_{t}G^{2}+\eta_{t}^{2}\nu_{\max}\frac{1}{N}\sigma^{2})
+	\end{align*}
+where $\eta_{t}=\frac{\alpha_{t}}{1-\beta_{t}}$. This is then summed over time steps to derive the final bound in \ref{thm:Nesterov_cvx}.
 

### 2021-07-11 source/sec_SGD.tex
Update on Overleaf.

--- 2021-07-05 source/sec_SGD.tex
@@
 \begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq (1-\alpha_t)\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+ \cO(\alpha_t^2 \sigma^2/N + \alpha_t^3E^2G^2).
+	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq (1-\alpha_t)\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+ \cO(\alpha_t^2 \sigma^2/N + \alpha_t^3E^2LG^2).
 \end{align*}
@@
 \textit{Iterative deduction.} This step applies induction to the recursive one-step progress bounds to obtain a non-recursive bound on $\mathbb{E}\|\ov{w}_{T}-\vw^{\ast}\|^{2}$ in terms of the initial distance ($\|\ov{w}_{0}-\vw^{\ast}\|^{2}$), which is then converted to a bound on $\mathbb{E}F(\overline{\mathbf{w}}_{T})-F^{\ast}$ using $L$-smoothness.\\
-\textit{From full to partial participation.} In addition to the variances arising from stochastic updates at local clients and averaging across clients, partial participation, which involves a sampling procedure, gives adds a an additional term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order resulting from the sampling variance is added to the recursive contraction bound.
-
-\begin{comment}
-A crucial ingredient in the proof of Theorem~\ref{th:scvx_sgd} is a one step contraction bound 
-	\begin{align*}
-	\mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2} & \leq(1-\mu\alpha_{t})\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{max}^{2}\sigma^{2}+6E^{2}L\alpha_{t}^{3}G^{2} 
-	\end{align*}
-that improves upon the analysis in~\cite{li2019convergence} to deliver the linear speedup.
-Moreover, analogues of this bound appear in the analyses of the convex setting as well as the Nesterov accelerated FedAvg.\\
-\end{comment}
-
+\textit{From full to partial participation.} In addition to the variances arising from stochastic updates at local clients and averaging across clients, partial participation, which involves a sampling procedure, gives rise to a third type of variance: an additional term $\mathcal{O}(\frac{1}{K} \alpha_t^2 E^2G^2)$ of leading order resulting from the sampling variance is added to the recursive contraction bound.
 \textbf{Linear speedup. }We compare our bound with that in \cite{li2019convergence},
@@
 \subsection{Convex Smooth Objectives}
-Next we provide linear speedup analysis of FedAvg with convex and
+Next, we provide linear speedup analysis of FedAvg with convex and
 smooth objectives and show that the optimality gap is $\mathcal{O}(1/\sqrt{KT})$. 
@@
 	& \mathbb{E}\|\ov{w}_{t+1}-\vw^{\ast}\|^{2}+\alpha_{t}(F(\ov{w}_{t})-F(\vw^{\ast})) \\
- \leq &\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\alpha_{t}^{2}\frac{1}{N}\nu_{\max}^{2}\sigma^{2}+6\alpha_{t}^{3}E^{2}LG^{2}
+ \leq &\mathbb{E}\|\ov{w}_{t}-\vw^{\ast}\|^{2}+\cO(\alpha_t^2 \nu_{\max}^{2} \sigma^2/N + \alpha_t^3E^2LG^2).
 	\end{align*}

### 2021-07-11 source/sec_intro.tex
Update on Overleaf.

@@
 \section{Introduction}
-Federated learning (FL) is a machine learning paradigm where many clients (e.g., mobile devices or organizations) collaboratively train a model under the orchestration of a central server (e.g., service provider), while keeping the training data decentralized (\cite{smith2017federated, kairouz2019advances}).
+Federated learning (FL) \cite{konevcny2016federated} is a machine learning paradigm where many clients (e.g., mobile devices or organizations) collaboratively train a model under the orchestration of a central server (e.g., service provider), while keeping the training data decentralized (\cite{smith2017federated, kairouz2019advances}).
 In recent years, FL has swiftly emerged as an important learning paradigm (\cite{mcmahan2016communication,li2018federated})--one that enjoys widespread success in applications such as personalized recommendation (\cite{chen2018federated}), virtual assistant (\cite{lam2019protecting}), and keyboard prediction (\cite{47586}), to name a few--for at least three reasons: First, the rapid proliferation of smart devices that are equipped with both computing power and data-capturing capabilities provided the infrastructure core for FL.
 Second, the rising awareness of privacy and the explosive growth of computational power in mobile devices have made it increasingly attractive to push the computation to the edge.
-Third, the empirical success of communication-efficient FL algorithms has enabled increasingly larger-scale parallel computing and learning with less communication overhead.
+Third, empirical success of communication-efficient FL algorithms has enabled increasingly larger-scale parallel computing and learning with less communication overhead.
 
 Despite its promise and broad applicability in our current era, the potential value FL delivers is coupled with the unique challenges it brings forth.
-In particular, when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated (\cite{kairouz2019advances}), it faces two challenges that are absent in centralized optimization and distributed (stochastic) optimization (\cite{zhou2017convergence, stich2018local,khaled2019first,liang2019variance,wang2018cooperative,woodworth2018graph,wang2019adaptive,jiang2018linear,yu2019parallel,yu2019linear, khaled2020tighter,koloskova2020unified,woodworth2020local,woodworth2020minibatch}):
+In particular, when FL learns a single statistical model using data from across all the devices while keeping each individual device's data isolated (\cite{kairouz2019advances}), it faces two challenges that are absent in centralized optimization and distributed (stochastic) optimization (\cite{nedic2009distributed,ram2010distributed,duchi2011dual,stich2018local,woodworth2018graph,yu2019parallel,yu2019linear, khaled2020tighter,koloskova2020unified}):
 
@@
 $T$ is the total number of stochastic updates performed by each local device, $E$ is the local steps between two consecutive server communications (and hence $T/E$ is the number of communications).
-$^{\dagger}$ In the linear regression setting, we have $\kappa=\kappa_1$ for FedAvg and $\kappa=\sqrt{\kappa_1\tilde{\kappa}}$ for momentum accelerated FedAvg (FedMaSS), where $\kappa_1$ and $\sqrt{\kappa_1\tilde{\kappa}}$ are condition numbers defined in Section \ref{sec:app:overparameterized}.
+$^{\dagger}$ In the linear regression setting, we have $\kappa=\kappa_1$ for FedAvg and $\kappa=\sqrt{\kappa_1\tilde{\kappa}}$ for momentum accelerated FedAvg (FedMaSS), where $\kappa_1$ and $\sqrt{\kappa_1\tilde{\kappa}}$ are condition numbers defined in Section \ref{sec:overparameterized}.
 Since
@@
 
-\textbf{New related works since last time}: \cite{zhu2019multi,xu2020ternary,yang2020federated,chen2019communication,sattler2020clustered,konevcny2016federated,yuan2020stochastic,dinh2020federated,li2020acceleration,nedic2009distributed,duchi2011dual,liu2016collective,yuan2016convergence,lu2011gossip,ram2010distributed,yuan2018optimal,horvath2019stochastic,wangni2017gradient,alistarh2017qsgd}
+\textbf{Other Related Works} Since the proposal of the original FedAvg algorithm, an active area of research in Federated Learning has focused on designing new algorithms that address the various challenges in FL, such as dealing with diverging data distributions \cite{sattler2020clustered}, improving communication efficiency \cite{chen2019communication,xu2020ternary,zhu2019multi}, reducing communication cost via compression or sparsification \cite{alistarh2017qsgd,wangni2017gradient,jiang2019improving,li2020acceleration}, and accelerating stochastic optimization methods with variance reduction \cite{horvath2019stochastic,dinh2020federated} Nevertheless, FedAvg remains a practically popular algorithm whose linear speedup behavior is desirable.
+In the distributed optimization setting, there have been a line of works \cite{liu2016collective,yuan2016convergence,yuan2020stochastic,yuan2018optimal} that deal with constrained optimization problems, often under convexity assumptions.
+Convergence results, especially with linear speedup analyses, in constrained Federated Learning settings are relatively less common \cite{wang2019adaptive}, and remains an open research area that deserves more attention.
 

### 2021-07-11 source/sec_overparameterized.tex
Update on Overleaf.

@@
 learning setting. 
-In this section, we provide the first geometric convergence rate
+In this section, we provide geometric convergence rate analysis
 of FedAvg for the overparameterized strongly convex and smooth problems,
@@
 	and constant step size $\overline{\alpha}=\mathcal{O}(\frac{1}{E}\frac{N}{l\nu_{\max}+L(N-\nu_{\min})})$
-	has geometric convergence
+	has geometric convergence:
 	\begin{align*}
-	\mathbb{E}F(\overline{\mathbf{w}}_{T}) & \leq\frac{L}{2}(1-\overline{\alpha})^{T}\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}=\mathcal{O}\left(L\exp\left(-\frac{\mu}{E}\frac{NT}{l\nu_{\max}+L(N-\nu_{\min})}\right)\cdot\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right).
+	&\mathbb{E}F(\overline{\mathbf{w}}_{T}) \leq\frac{L}{2}(1-\overline{\alpha})^{T}\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\\
+	=&\mathcal{O}\left(L\exp\left(-\frac{\mu}{E}\frac{NT}{l\nu_{\max}+L(N-\nu_{\min})}\right)\cdot\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right).
 	\end{align*}

### 2021-07-11 source/main.tex
Add bio

@@
 
-\begin{comment}
+% \begin{comment}
 
+\newpage
 \begin{IEEEbiography}{Zhaonan Qu}
@@
 
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fig/bio_zhaojian.png}}]{Zhaojian Li}
+Dr.
+Zhaojian Li is an assistant Professor in the Department of Mechanical Engineering at Michigan State University.
+His main research interests include Robotics and Autonomous Vehicles, Intelligent Transportation System, Reinforcement Learning, Vehicle Dynamics, and Optimal Control.
+\end{IEEEbiography}
+
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fig/bio_jiayu.png}}]{Jiayu Zhou}
@@
 
-\end{comment}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fig/bio_zhengyuan.png}}]{Zhengyuan Zhou}
+Dr.
+Zhengyuan Zhou is an assistant professor in the Department of Technology, Operations, and Statistics at Stern School of Business, New York University.
+Before joining NYU Stern, he obtained Ph.D.
+degree in the Information System Laboratory in Department of Electrical Engineering at Stanford University.
+\end{IEEEbiography}
+
+% \end{comment}
 

### 2021-07-11 source/sec_intro.tex

@@
 
-\textbf{Other Related Works} Since the proposal of the original FedAvg algorithm, an active area of research in Federated Learning has focused on designing new algorithms that address the various challenges in FL, such as dealing with diverging data distributions \cite{sattler2020clustered}, improving communication efficiency \cite{chen2019communication,xu2020ternary,zhu2019multi}, reducing communication cost via compression or sparsification \cite{alistarh2017qsgd,wangni2017gradient,jiang2019improving,li2020acceleration}, and accelerating stochastic optimization methods with variance reduction \cite{horvath2019stochastic,dinh2020federated} Nevertheless, FedAvg remains a practically popular algorithm whose linear speedup behavior is desirable.
+\textbf{Other Related Works} Since the proposal of the original FedAvg algorithm, an active area of research in Federated Learning has focused on designing new algorithms that address the various challenges in FL, such as dealing with diverging data distributions \cite{sattler2020clustered}, improving communication efficiency \cite{chen2019communication,xu2020ternary,zhu2019multi,zhou2021communication}, reducing communication cost via compression or sparsification \cite{alistarh2017qsgd,wangni2017gradient,jiang2019improving,li2020acceleration}, and accelerating stochastic optimization methods with variance reduction \cite{horvath2019stochastic,dinh2020federated} Nevertheless, FedAvg remains a practically popular algorithm whose linear speedup behavior is desirable.
 In the distributed optimization setting, there have been a line of works \cite{liu2016collective,yuan2016convergence,yuan2020stochastic,yuan2018optimal} that deal with constrained optimization problems, often under convexity assumptions.

### 2021-07-11 source/sec_overparameterized.tex

@@
 	has geometric convergence:
+	{\small 
 	\begin{align*}
 	\mathbb{E}F(\overline{\mathbf{w}}_{T}) & \leq\mathcal{O}\left(L\exp(-\frac{NT}{E(\nu_{\max}\kappa_{1}+(N-\nu_{\min}))})\|\mathbf{w}_{0}-\mathbf{w}^{\ast}\|^{2}\right).
-	\end{align*}
+	\end{align*}}
 \end{theorem}

### 2021-07-11 source/main.tex
Update on Overleaf.

@@
 \newpage
-\begin{IEEEbiography}{Zhaonan Qu}
-Biography text here.
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fig/bio_zhaonan.jpg}}]{Zhaonan Qu}
+Zhaonan Qu is a PhD student at the Department of Economics at Stanford University.
+His research interest is at the intersection of optimization, machine learning, and causal inference.
+He received a B.S.
+in mathematics from Princeton University and an M.A.
+in mathematics from NYU's Courant Institute.
 \end{IEEEbiography}

### 2021-07-11 source/main.tex
Update bio
Update on Overleaf.

@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fig/bio_zhaonan.jpg}}]{Zhaonan Qu}
-Zhaonan Qu is a PhD student at the Department of Economics at Stanford University.
+Zhaonan Qu is a PhD student in the Department of Economics at Stanford University.
 His research interest is at the intersection of optimization, machine learning, and causal inference.

### 2021-07-12 source/main.tex
Update on Overleaf.

@@
 Zhaonan Qu is a PhD student in the Department of Economics at Stanford University.
-His research interest is at the intersection of optimization, machine learning, and causal inference.
+His research interests include optimization, machine learning, and causal inference.
 He received a B.S.
@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fig/bio_kaixiang.png}}]{Kaixiang Lin}
-Dr. Kaixiang Lin is an applied scientist at Amazon Lab 126. Before joining Amazon, 
-Kaixiang received his Ph.D. degree in Computer Science at Michigan State University in 2020. 
+Dr.
+Kaixiang Lin received his Ph.D.
+degree in Computer Science at Michigan State University in 2020.
+He regularly serves as a program committee member or reviewer for top-tier machine learning/data ming conferences/journals, including Neurips, ICML, ICLR, TKDE, etc.
+% Before joining MSU, he received the B.S. degree from the University of Science and Technology of China, Hefei, China, in 2014. 
 \end{IEEEbiography}
@@
 %\newpage
-
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fig/bio_zhaojian.png}}]{Zhaojian Li}
@@
 His main research interests include Robotics and Autonomous Vehicles, Intelligent Transportation System, Reinforcement Learning, Vehicle Dynamics, and Optimal Control.
+He is a senior member of IEEE and recipient of NSF CAREER Award.
 \end{IEEEbiography}
@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fig/bio_jiayu.png}}]{Jiayu Zhou}
-Dr. Jiayu Zhou is an assistant professor at
- Department of Computer Science and Engineering, Michigan State University.
- Before joining MSU, Jiayu was a staff research scientist at Samsung Research
- America. Jiayu received his Ph.D. degree in computer science at Arizona
- State University in 2014. 
+Dr.
+Jiayu Zhou is an Associate Professor in the Department of Computer Science and Engineering at Michigan State University.
+He received his Ph.D.
+degree in computer science from Arizona State University in 2014.
+Dr.
+Zhou has a broad research interest in large-scale machine learning, data mining, and biomedical informatics, with a focus on the transfer and multi-task learning.
+His research has been funded by the National Science Foundation, National Institutes of Health, and Office of Naval Research, and published more than 100 peer-reviewed journal and conference papers in data mining and machine learning.
+Dr.
+Zhou is a recipient of the National Science Foundation CAREER Award (2018).
+His papers received the Best Student Paper Award in 2014 IEEE International Conference on Data Mining (ICDM), the Best Student Paper Award at the 2016 International Symposium on Biomedical Imaging (ISBI), and Best Paper Award at 2016 IEEE International Conference on Big Data (BigData).
 \end{IEEEbiography}

### 2021-07-18 source/sec_intro.tex
Update on Overleaf.

--- 2021-07-11 source/sec_intro.tex
@@
 
-\textbf{Other Related Works} Since the proposal of the original FedAvg algorithm, an active area of research in Federated Learning has focused on designing new algorithms that address the various challenges in FL, such as dealing with diverging data distributions \cite{sattler2020clustered}, improving communication efficiency \cite{chen2019communication,xu2020ternary,zhu2019multi}, reducing communication cost via compression or sparsification \cite{alistarh2017qsgd,wangni2017gradient,jiang2019improving,li2020acceleration}, and accelerating stochastic optimization methods with variance reduction \cite{horvath2019stochastic,dinh2020federated} Nevertheless, FedAvg remains a practically popular algorithm whose linear speedup behavior is desirable.
+\textbf{Other Related Works} Since the proposal of the original FedAvg algorithm, an active area of research in Federated Learning has focused on designing new algorithms that address the various challenges in FL, such as dealing with diverging data distributions \cite{sattler2020clustered}, improving communication efficiency \cite{sattler2019robust,chen2019communication,xu2020ternary,zhu2019multi,zhou2021communication}, reducing communication cost via compression or sparsification \cite{alistarh2017qsgd,wangni2017gradient,jiang2019improving,li2020acceleration}, and accelerating stochastic optimization methods with variance reduction \cite{horvath2019stochastic,dinh2020federated} Nevertheless, FedAvg remains a practically popular algorithm whose linear speedup behavior is desirable.
 In the distributed optimization setting, there have been a line of works \cite{liu2016collective,yuan2016convergence,yuan2020stochastic,yuan2018optimal} that deal with constrained optimization problems, often under convexity assumptions.

### 2021-07-18 source/sec_intro.tex
change table 1

--- 2021-07-11 source/sec_intro.tex
@@
 \centering
+\caption{The summary of our convergence results for FedAvg and accelerated FedAvg.}
 \begin{tabular}{|c|c|c|c|c|}\hline 
@@
 \end{tabular}
-\caption{Our convergence results for FedAvg and accelerated FedAvg in this paper. Throughout the paper, $N$ is the total
-number of local devices, and $K \leq N$ is the maximal number of devices that are accessible to the central server.
+% \caption{Our convergence results for FedAvg and accelerated FedAvg in this paper. Throughout the paper, $N$ is the total
+% number of local devices, and $K \leq N$ is the maximal number of devices that are accessible to the central server.
 $T$ is the total number of stochastic updates performed by each local device, $E$ is the local steps between two consecutive server communications (and hence $T/E$ is the number of communications).
@@
 Since
-$\kappa_{1}\geq\tilde{\kappa}$, this implies a speedup factor of
-$\sqrt{\frac{\kappa_{1}}{\tilde{\kappa}}}$ for accelerated FedAvg.}
+% $\kappa_{1}\geq\tilde{\kappa}$, this implies a speedup factor of
+% $\sqrt{\frac{\kappa_{1}}{\tilde{\kappa}}}$ for accelerated FedAvg.}
 % {\raggedright 

### 2021-07-18 source/sec_intro.tex

@@
 
-\textbf{Other Related Works} Since the proposal of the original FedAvg algorithm, an active area of research in Federated Learning has focused on designing new algorithms that address the various challenges in FL, such as dealing with diverging data distributions \cite{sattler2020clustered}, improving communication efficiency \cite{chen2019communication,xu2020ternary,zhu2019multi,zhou2021communication}, reducing communication cost via compression or sparsification \cite{alistarh2017qsgd,wangni2017gradient,jiang2019improving,li2020acceleration}, and accelerating stochastic optimization methods with variance reduction \cite{horvath2019stochastic,dinh2020federated} Nevertheless, FedAvg remains a practically popular algorithm whose linear speedup behavior is desirable.
+\textbf{Other Related Works} Since the proposal of the original FedAvg algorithm, an active area of research in Federated Learning has focused on designing new algorithms that address the various challenges in FL, such as dealing with diverging data distributions \cite{sattler2020clustered}, improving communication efficiency \cite{sattler2019robust,chen2019communication,xu2020ternary,zhu2019multi,zhou2021communication}, reducing communication cost via compression or sparsification \cite{alistarh2017qsgd,wangni2017gradient,jiang2019improving,li2020acceleration}, and accelerating stochastic optimization methods with variance reduction \cite{horvath2019stochastic,dinh2020federated} Nevertheless, FedAvg remains a practically popular algorithm whose linear speedup behavior is desirable.
 In the distributed optimization setting, there have been a line of works \cite{liu2016collective,yuan2016convergence,yuan2020stochastic,yuan2018optimal} that deal with constrained optimization problems, often under convexity assumptions.

### 2021-07-18 source/main.tex
change table 1
Update on Overleaf.

@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fig/bio_zhaonan.jpg}}]{Zhaonan Qu}
-Zhaonan Qu is a PhD student in the Department of Economics at Stanford University.
+Zhaonan Qu is a PhD student in the Department of Economics at Stanford University with a PhD Minor in Operations Research.
 His research interests include optimization, machine learning, and causal inference.
+He is supported by a Stanford Interdisciplinary Graduate Fellowship.
 He received a B.S.
@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fig/bio_kaixiang.png}}]{Kaixiang Lin}
-Dr.
-Kaixiang Lin received his Ph.D.
-degree in Computer Science at Michigan State University in 2020.
+Dr. Kaixiang Lin received his Ph.D. degree in Computer Science at 
+Michigan State University in 2020. His research interests include 
+reinforcement learning, data mining, and machine learning. 
 He regularly serves as a program committee member or reviewer for top-tier machine learning/data ming conferences/journals, including Neurips, ICML, ICLR, TKDE, etc.
